### 2021-03-13 docs/index.Rmd
minor

---
title: '**The Illusion Game**'
subtitle: Results
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: no
    df_print: kable
    code_folding: hide
    code_download: yes
  word_document:
    reference_docx: utils/Template_Word.docx
    highlight: pygments
    toc: no
    toc_depth: 3
    df_print: kable
    number_sections: yes
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '2'
    latex_engine: xelatex
editor_options:
  chunk_output_type: console
bibliography: utils/bibliography.bib
csl: utils/apa.csl
---

<!-- 
!!!! IMPORTANT: run `source("utils/render.R")` to publish instead of clicking on 'Knit'
-->

```{r setup, warning=FALSE, message=TRUE, include=FALSE}
# Set up the environment (or use local alternative `source("utils/config.R")`)
source("https://raw.githubusercontent.com/RealityBending/TemplateResults/main/utils/config.R")

fast <- FALSE  # Make this false to skip the chunks
```

# Packages & Data

## Packages

This document was prepared on `r format(Sys.Date())`.

```{r warning=FALSE, message=TRUE, results='asis'}
library(bayestestR)
library(parameters)
library(performance)
library(report)
library(see)
library(ggplot2)

summary(report::report(sessionInfo()))
```

## Data

```{r warning=FALSE, message=TRUE, results='asis'}
df <- read.csv("../analysis/data.csv")

# cat(paste("The data consists of", report::report_participants(df, participants = "Participant")))
```

# Descriptive Stats {.tabset}

```{r child=if (fast == FALSE) '1_descriptive.Rmd'}
```

# Full Code

The full script of executive code contained in this document is reproduced here.

```{r full_code, ref.label=knitr::all_labels(), eval=FALSE}
```

# Package References

```{r warning=FALSE, message=FALSE, results='asis'}
report::cite_packages(sessionInfo())
```

# References

### 2021-03-13 experiment.js

@@ -189,7 +189,7 @@ var fullscreen = {
 // Instructions
 var delboeuf_instructions = {
     type: "html-button-response",
-    choices: ["Start"],
+    choices: ["Got it!"],
     stimulus: function () {
         if (systemInfo().screen_touchscreen == false) {
             return ("<p>In this experiment, two red circles will appear side by side each other " +
@@ -303,7 +303,7 @@ var delboeuf_debrief = {
 // Instructions
 var ebbinghaus_instructions = {
     type: "html-button-response",
-    choices: ["Start"],
+    choices: ["Got it!"],
     stimulus: function () {
         if (systemInfo().screen_touchscreen == false) {
             return ("<p>In this experiment, two red circles will appear side by side each other " +
@@ -417,7 +417,7 @@ var ebbinghaus_debrief = {
 // Instructions
 var mullerlyer_instructions = {
     type: "html-button-response",
-    choices: ["Start"],
+    choices: ["Got it!"],
     stimulus: function () {
         if (systemInfo().screen_touchscreen == false) {
             return ("<p>In this experiment, two red lines will appear " +
@@ -532,7 +532,7 @@ var mullerlyer_debrief = {
 // Instructions
 var ponzo_instructions = {
     type: "html-button-response",
-    choices: ["Start"],
+    choices: ["Got it!"],
     stimulus: function () {
         if (systemInfo().screen_touchscreen == false) {
             return ("<p>In this experiment, two red lines will appear " +

### 2021-03-13 analysis/distribution.R
modify colnames in data, setwd() in R scripts, preload instructions

@@ -6,8 +6,9 @@ library(dplyr)
 # library(jsonlite)
 library(ggplot2)
 
-data <- read.csv("analysis/data.csv") %>% 
-  dplyr::rename(Illusion_Type = Block) %>% 
+setwd("analysis")
+
+data <- read.csv("data.csv") %>% 
   dplyr::mutate(Illusion_Type = as.factor(tools::toTitleCase(Illusion_Type)),
          PlayedBefore = as.factor(PlayedBefore),
          Correct = ifelse(Correct == "TRUE", 1, 0)) 

### 2021-03-13 analysis/make_data.R

@@ -1,6 +1,8 @@
 library(rjson)
 
-data_path <- "data/"
+setwd("analysis")
+
+data_path <- "../data/"
 data <- data.frame()  # Initialize empty dataframe
 
 
@@ -26,7 +28,7 @@ preprocess_session_info <- function(df) {
   names(df) <- tools::toTitleCase(names(df))
   df$Duration_Consent <- df$Rt
   df[c("Button_pressed", "Stimulus", "Screen" ,"Rt", "Internal_node_id", "Time_elapsed", "Trial_index", "Trial_type", "Response")] <- NULL
-  names(df)[names(df) == "Participant_id"] <- "Participant_ID"
+  names(df)[names(df) == "Participant_id"] <- "Participant"
   df
 }
 
@@ -53,6 +55,7 @@ preprocess_trial <- function(df) {
   df <- df[!stringr::str_detect(names(df), pattern="click_")]  # remove click_x and click_y
   names(df) <- tools::toTitleCase(names(df))
   df[c("Screen" ,"Internal_node_id", "Time_elapsed", "Trial_type")] <- NULL
+  names(df)[names(df) == "Block"] = "Illusion_Type"
   names(df)[names(df) == "Illusion_strength"] = "Illusion_Strength"
   names(df)[names(df) == "Illusion_difference"] = "Illusion_Difference"
   names(df)[names(df) == "Rt"] <- "RT"
@@ -63,7 +66,7 @@ preprocess_results <- function(df, per_block = TRUE) {
   scores_cols <- c("accuracy", "rt_mean", "rt_mean_correct", "inverse_efficiency_score")
   if(per_block) {
     df <- df[c(scores_cols, "block")] # keep only these values
-    names(df)[names(df) == "block"] <- "Block"
+    names(df)[names(df) == "block"] <- "Illusion_Type"
   } else {
     df <- df[scores_cols]
   }
@@ -77,7 +80,7 @@ preprocess_results <- function(df, per_block = TRUE) {
     colnames(df) <- paste0('Grand_', colnames(df))
     df
   } else {
-    colnames(df)[colnames(df) != "Block"] <- paste0('Block_', colnames(df[!(colnames(df) %in% "Block")]))
+    colnames(df)[colnames(df) != "Illusion_Type"] <- paste0('Block_', colnames(df[!(colnames(df) %in% "Block")]))
     df
   }
 }
@@ -165,7 +168,7 @@ for(file in list.files(data_path)) {
   }
 
[…]

### 2021-03-13 analysis/make_scores.R

@@ -1,8 +1,9 @@
 library(dplyr)
 library(jsonlite)
 
-data <- read.csv("analysis/data.csv") %>%
-  dplyr::rename(Illusion_Type = Block) %>%
+setwd("analysis")
+
+data <- read.csv("data.csv") %>%
   dplyr::mutate(Illusion_Type = as.factor(tools::toTitleCase(Illusion_Type)),
          PlayedBefore = as.factor(PlayedBefore),
          Correct = ifelse(Correct == "TRUE", 1, 0))

### 2021-03-13 experiment.js

@@ -94,6 +94,16 @@ var session_info = {
 var trial_number = 1 // trial indexing variable starts at 1 for convenience
 var block_number = 0 // block indexing variables (should block 0 be there as practice block?)
 
+// Set demo images and preload them
+var instructions_stimuli = ['utils/Delboeuf_Demo.png', 'utils/Ebbinghaus_Demo.png', 'utils/MullerLyer_Demo.png', 'utils/Ponzo_Demo.png',
+    'utils/answer/answer_leftright_keyboard.PNG', 'utils/answer/answer_leftright_touch.PNG',
+    'utils/answer/answer_updown_keyboard.PNG', 'utils/answer/answer_updown_touch.PNG']
+
+var instructions_preload = {
+    type: 'preload',
+    trials: instructions_stimuli
+}
+
 
 // Welcome + Informed Consent
 var welcome = {
@@ -670,7 +680,7 @@ var end_experiment = {
 
 /* ----------------- Initialize experiment ----------------- */
 jsPsych.init({
-    timeline: [fullscreen, welcome, participant_info_general, participant_info_repetition, if_not_repeated,
+    timeline: [fullscreen, instructions_preload, welcome, participant_info_general, participant_info_repetition, if_not_repeated,
         delboeuf_preload, delboeuf_instructions, test_delboeuf_procedure, delboeuf_debrief,
         ebbinghaus_preload, ebbinghaus_instructions, test_ebbinghaus_procedure, ebbinghaus_debrief,
         mullerlyer_preload, mullerlyer_instructions, test_mullerlyer_procedure, mullerlyer_debrief,

### 2021-03-13 analysis/make_scores.R
trigger again

@@ -32,10 +32,10 @@ for(type in df_scores$Illusion_Type){
 
 
 # Save as js
-jsonlite::write_json(scores, "variables_scores.js")
-text <-  paste("var scores =", readr::read_file("variables_scores.js"))
+jsonlite::write_json(scores, "../variables_scores.js")
+text <-  paste("var scores =", readr::read_file("../variables_scores.js"))
 
 
-file <- file("variables_scores.js")
+file <- file("../variables_scores.js")
 writeLines(text, file)
 close(file)

### 2021-03-14 analysis/make_data.R
minor

@@ -187,4 +187,9 @@ for(file in list.files(data_path)) {
 #   dplyr::select(Participant_ID, Age, Initials, PlayedBefore, Stimulus, Illusion_Strength, Illusion_Difference, everything())
 
 
+data <- data %>%
+  dplyr::mutate(Illusion_Type = as.factor(tools::toTitleCase(Illusion_Type)),
+                PlayedBefore = as.factor(PlayedBefore),
+                Correct = ifelse(Correct == "TRUE", 1, 0))
+
 write.csv(data, "data.csv", row.names = FALSE)

### 2021-03-14 analysis/make_scores.R

@@ -3,11 +3,6 @@ library(jsonlite)
 
 setwd("analysis")
 
-data <- read.csv("data.csv") %>%
-  dplyr::mutate(Illusion_Type = as.factor(tools::toTitleCase(Illusion_Type)),
-         PlayedBefore = as.factor(PlayedBefore),
-         Correct = ifelse(Correct == "TRUE", 1, 0))
-
 
 # Get scores by illusions
 df_scores <- data %>%

### 2021-03-14 analysis/make_data.R
fix

@@ -1,4 +1,5 @@
 library(rjson)
+library(dplyr)
 
 setwd("analysis")
 

### 2021-03-14 analysis/make_scores.R

@@ -5,7 +5,7 @@ setwd("analysis")
 
 
 # Get scores by illusions
-df_scores <- data %>%
+df_scores <- read.csv('data.csv') %>%
   dplyr::group_by(Illusion_Type) %>%
   dplyr::summarize(
     IES_Mean = mean(Block_IES, na.rm=TRUE),

### 2021-03-14 variables_scores.js
Make variables_scores.js

@@ -1,2 +1,2 @@
-var scores = {"Delboeuf":{"IES_Mean":[913.2278],"IES_SD":[192.6265]},"Ebbinghaus":{"IES_Mean":[627.6319],"IES_SD":[321.3488]},"Mullerlyer":{"IES_Mean":[1099.9454],"IES_SD":[291.6726]},"Ponzo":{"IES_Mean":[933.1764],"IES_SD":[509.8769]},"Total":{"IES_Mean":[893.4954],"IES_SD":[328.8812]}}
+var scores = {"delboeuf":{"IES_Mean":[842.4133],"IES_SD":[223.3891]},"ebbinghaus":{"IES_Mean":[644.4352],"IES_SD":[288.9538]},"mullerlyer":{"IES_Mean":[1082.9683],"IES_SD":[262.6984]},"ponzo":{"IES_Mean":[889.5584],"IES_SD":[463.711]},"Total":{"IES_Mean":[864.8438],"IES_SD":[309.6881]}}
 

### 2022-07-31 analysis.Rmd
update data

---
title: "Illusion Game Validation"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3)

cache <- FALSE
fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

## Introduction

Behavioural findings regarding the Illusion Game.

## Methods

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

# This is a local folder containing raw data from unzipped pavlovia
dfraw <- read.csv("data/study1.csv") |> 
  mutate(Screen_Refresh = as.character(Screen_Refresh),
         Illusion_Side = as.factor(Illusion_Side))
```

### Exclusions

For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).

```{r message=FALSE, warning=FALSE}
dfsub <- dfraw |>
  group_by(Participant) |>
  summarize(Error = sum(Error) / n(),
            RT = mean(RT)) |>
  ungroup()

dfraw |>
  group_by(Participant, Illusion_Type, Block) |>
  summarize(ErrorRate_per_block = sum(Error) / n()) |>
  ungroup() |>
  estimate_density(at = c("Illusion_Type", "Block")) |>
  ggplot(aes(x = x, y = y)) +
  geom_line(aes(color = Illusion_Type, linetype= as.factor(Block)))

df <- dfraw |>
  group_by(Participant, Illusion_Type, Block) |>
  mutate(ErrorRate_per_block = sum(Error) / n()) |> 
  ungroup() |> 
  filter(ErrorRate_per_block < 0.5) |> 
  select(-ErrorRate_per_block)

nrow(dfraw) - nrow(df)
```

### Participants

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  group_by(Participant) |>
  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS) |> 
  slice(1) |> 
  ungroup()
```

`r report::report_participants(dfsub, age="Age", sex="Sex")`

```{r }
plot_distribution <- function(dfsub, what="Age", title=what, subtitle="", fill="orange") {
  dfsub |> 
    ggplot(aes_string(x = what)) +
    geom_density(fill = fill) + 
    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype="dashed") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggtitle(title, subtitle=subtitle) +
    theme_modern() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5),
          plot.subtitle = element_text(face = "italic", hjust = 0.5),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank())
}

plot_waffle <- function(data, title="Nationality") {
  data |> 
    # mutate(label = emojifont::fontawesome('fa-twitter')) |> 
    ggplot(aes(x, y, fill = group)) + 
    ggwaffle::geom_waffle() +
    # geom_point() +
    # geom_text(aes(label=label), family='fontawesome-webfont', size=4) +
    coord_equal() + 
    ggtitle(title) +
    labs(fill = "") +
    theme_void() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}
```

```{r }
p1 <- plot_distribution(dfsub, "Age", fill="#FF9800")
p2 <- plot_distribution(dfsub, "Duration", title="Total Duration", subtitle="in minutes", fill="#F44336")
p3 <- plot_distribution(dfsub, "Break_Duration", title="Break Duration", subtitle="in minutes", fill="#3F51B5")

p4 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Sex)) |> 
  plot_waffle("Sex") +
  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63"))

p5 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Education)) |> 
  plot_waffle("Education") +
  scale_fill_viridis_d()

p6 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Nationality)) |> 
  plot_waffle("Nationality") +
  scale_fill_viridis_d()

p7 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Ethnicity)) |> 
  plot_waffle("Ethnicity") +
  scale_fill_viridis_d()

p8 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Screen_Resolution)) |> 
  plot_waffle("Screen Resolution") +
  scale_fill_viridis_d()

p9 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Device_OS)) |> 
  plot_waffle("Device OS") +
  scale_fill_viridis_d()

# p10 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Screen_Refresh)) |> 
#   plot_waffle("Screen Refresh Rate") +
#   scale_fill_viridis_d()

(p1 / p2 / p3) | (p4 / p5 / p6) | (p7 / p8 / p9) 
```

## Results

### Manipulation Checks

We tested whether at illusion strength = 0, no effect of right / left (no bias of illusion at 0).

```{r message=FALSE, warning=FALSE}
model <- glmmTMB::glmmTMB(Error ~ Illusion_Type / Illusion_Side + (1|Participant), 
                          data = filter(df, Illusion_Strength == 0), 
                          family="binomial")

parameters(model, effects="fixed") |> 
  tail(10)
```

### Full Models

#### Outliers

```{r message=FALSE, warning=FALSE}
p1 <- estimate_density(df, select="RT", at="Participant") |> 
  normalize(select = "y") |> 
  ggplot(aes(x=x, y=y)) +
  geom_area(data=normalize(estimate_density(df, select="RT"), select="y"), alpha=0.2) +
  geom_line(aes(color=Participant, group=Participant)) + 
  geom_vline(xintercept = 2500, linetype = "dashed", color = "red") +
  scale_color_material_d("rainbow", guide="none") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(xlim = c(0, 5000)) +
  theme_modern() +
  theme(axis.text.y = element_blank())

df$Outlier <- df$RT > 2500

p2 <- df |> 
  group_by(Participant) |> 
  summarize(Outlier = sum(Outlier) / n()) |> 
  mutate(Participant = fct_reorder(Participant, Outlier)) |> 
  ggplot(aes(x = Participant, y = Outlier)) +
  geom_bar(stat="identity", aes(fill=Participant)) +
  scale_fill_material_d(guide="none") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
  theme_modern("rainbow") +
  theme(axis.text.x = element_blank())

p1 | p2

df <- filter(df, Outlier == FALSE)
```

#### Errors

```{r message=FALSE, warning=FALSE}
model_errors <- function(df, illusion = "Delboeuf") {
  data <- filter(df, Illusion_Type == illusion)

  # formula <- brms::bf(
  #   # Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4) + (1 | Participant),
  #   Error ~ s(Illusion_Strength, by=Illusion_Difference, k = 7) + (1 | Participant),
  #   family = "bernoulli"
  # )
  # model <- brms::brm(formula,
  #   data = data,
  #   algorithm = "meanfield",
  #   refresh = 0
  # )
  # model <- glmmTMB::glmmTMB(
  #   Error ~ Illusion_Difference * Illusion_Strength + (1 | Participant),
  #   data = data,
  #   family = "binomial"
  # )
  model <- mgcv::gamm(
    Error ~ s(Illusion_Difference, by = Illusion_Strength, k=4),
    random = list(Participant = ~1),
    data = data,
    family = "binomial"
  )

  n_lines <- 5
  pred <- estimate_relation(model, at = c("Illusion_Strength", "Illusion_Difference"), length = c(30, n_lines)) |>
    mutate(Illusion_Difference = as.factor(Illusion_Difference),
           Illusion_Type = illusion)

  p <- pred |>
    ggplot(aes(x = Illusion_Strength, y = Predicted, group = Illusion_Difference)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference), alpha = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept=c(0.05, 0.5, 0.95), linetype="dotted", alpha=0.5) +
    geom_line(aes(color = Illusion_Difference)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)) +
    scale_fill_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)) +
    theme_modern() +
    labs(
      title = paste0(illusion, " Illusion"),
      color = "Difficulty", fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  list(p = p, model = model, pred = pred)
}
```

```{r message=FALSE, warning=FALSE}
rez_ponzo <- model_errors(df, illusion = "Ponzo")
rez_mullerlyer <- model_errors(df, illusion = "Müller-Lyer")
rez_delboeuf <- model_errors(df, illusion = "Delboeuf")
rez_ebbinghaus <- model_errors(df, illusion = "Ebbinghaus")
rez_contrast <- model_errors(df, illusion = "Contrast")
rez_white <- model_errors(df, illusion = "White")
rez_poggendorff <- model_errors(df, illusion = "Poggendorff")
rez_zollner <- model_errors(df, illusion = "Zöllner")
rez_rodframe <- model_errors(df, illusion = "Rod-Frame")
rez_verticalhorizontal <- model_errors(df, illusion = "Vertical-Horizontal")
```

```{r message=FALSE, warning=FALSE}
rez_ponzo$p
rez_mullerlyer$p
rez_delboeuf$p
rez_ebbinghaus$p
rez_verticalhorizontal$p
rez_rodframe$p
rez_zollner$p
rez_poggendorff$p
rez_contrast$p
rez_white$p
```

<!-- ```{r message=FALSE, warning=FALSE} -->
<!-- rbind( -->
<!--   rez_delboeuf$pred, -->
<!--   rez_ebbinghaus$pred, -->
<!--   rez_ponzo$pred, -->
<!--   rez_mullerlyer$pred, -->
<!--   rez_contrast$pred, -->
<!--   rez_white$pred, -->
<!--   rez_poggendorff$pred, -->
<!--   rez_zollner$pred, -->
<!--   rez_rodframe$pred, -->
<!--   rez_verticalhorizontal$pred -->
<!--   ) |>  -->
<!--   ggplot(aes(x = Illusion_Strength, y = Predicted, group = Illusion_Difference)) + -->
<!--   geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference), alpha = 0.2) + -->
<!--   # geom_vline(xintercept = 0, linetype = "dotted") + -->
<!--   geom_line(aes(color = Illusion_Difference)) + -->
<!--   scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) + -->
<!--   scale_x_continuous(expand = c(0, 0)) + -->
<!--   scale_color_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(5)) + -->
<!--   scale_fill_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(5)) + -->
<!--   theme_modern() + -->
<!--   labs( -->
<!--     title = paste0(illusion, " Illusion"), -->
<!--     color = "Difficulty", fill = "Difficulty", -->
<!--     y = "Probability of Error", -->
<!--     x = "Illusion Strength" -->
<!--   ) + -->
<!--   theme(plot.title = element_text(face = "bold", hjust = 0.5)) + -->
<!--   facet_wrap(~Illusion_Type, scales="free") -->
<!-- ``` -->

#### RT

<!-- ```{r message=FALSE, warning=FALSE} -->
<!-- run_model_rt <- function(df, illusion = "Delboeuf") { -->
<!--   data <- df |> -->
<!--     filter(Illusion_Type == illusion) |>  -->
<!--     mutate(Error = ifelse(Error == 1, "Error", "Correct")) -->

<!--   model <- lme4::lmer( -->
<!--     RT ~ Error * (Illusion_Difference * Illusion_Strength) + (1 | Participant), -->
<!--     data = data -->
<!--   ) -->

<!--   pred <- estimate_relation(model, at = c("Illusion_Strength", "Illusion_Difference", "Error"), length = c(30, 5)) |> -->
<!--     mutate(Illusion_Difference = as.factor(Illusion_Difference), -->
<!--            Illusion_Type = illusion) -->

<!--   p <- pred |> -->
<!--     mutate(Predicted = ifelse(Error == "Error", -Predicted, Predicted)) |>  -->
<!--     ggplot(aes(x = Illusion_Strength, y = Predicted, group = interaction(Illusion_Difference, Error))) + -->
<!--     # geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference), alpha = 0.2) + -->
<!--     geom_line(aes(color = Illusion_Difference)) + -->
<!--     scale_y_continuous(expand = c(0, 0)) + -->
<!--     scale_x_continuous(expand = c(0, 0)) + -->
<!--     scale_color_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(5)) + -->
<!--     scale_fill_manual(values = colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(5)) + -->
<!--     theme_modern() + -->
<!--     labs( -->
<!--       title = paste0(illusion, " Illusion"), -->
<!--       color = "Difficulty", fill = "Difficulty", -->
<!--       y = "Reaction Time", -->
<!--       x = "Illusion Strength" -->
<!--     ) + -->
<!--     theme(plot.title = element_text(face = "bold", hjust = 0.5)) -->

<!--   list(p = p, model = model, pred = pred) -->

<!-- } -->
<!-- ``` -->
<!-- ```{r message=FALSE, warning=FALSE} -->
<!-- run_model_rt(df, illusion = "Delboeuf")$p -->
<!-- run_model_rt(df, illusion = "Ebbinghaus")$p -->
<!-- run_model_rt(df, illusion = "Ponzo")$p -->
<!-- run_model_rt(df, illusion = "Müller-Lyer")$p -->
<!-- run_model_rt(df, illusion = "Contrast")$p -->
<!-- run_model_rt(df, illusion = "White")$p -->
<!-- run_model_rt(df, illusion = "Poggendorff")$p -->
<!-- run_model_rt(df, illusion = "Zöllner")$p -->
<!-- run_model_rt(df, illusion = "Rod-Frame")$p -->
<!-- run_model_rt(df, illusion = "Vertical-Horizontal")$p -->
<!-- ``` -->

### Inter-Illusion Parameters Standardization

### Individual Score Approximation

correlation between inter-individual indices extracted from full model with that of individual models (with and without priors)

<!-- ### Drift Model -->

## References

### 2022-07-31 preprocessing.R

@@ -1,65 +1,83 @@
 # Preprocessing Function
-preprocess_IllusionGame <- function(file) {
+preprocess_raw <- function(file) {
 
-  json <- jsonlite::fromJSON(file)
+  data <- read.csv(file)
+
+  # Demographics
+  dem <- data[data$screen == "demographics" & !is.na(data$screen), "response"]
 
   # Info
-  info <- json[json$screen == "participant_info" & !is.na(json$screen), ]
+  info <- data[data$screen == "browser_info" & !is.na(data$screen), ]
 
 
-
-  # Trials
-  trials <- json[!is.na(json$block),]
-
-
-  data <- data.frame(Participant = trials$participant_id,
-                     Date = info$date,
-                     Time = info$time,
-                     Duration = json[json$screen == "final_results" & !is.na(json$screen), "time_elapsed"] / 1000 / 60,
-                     Break_Duration = json[json$screen == "break" & !is.na(json$screen), "rt"] / 1000 / 60,
-                     Screen_Resolution = paste0(trials$screen_width, "x", trials$screen_height),
-                     Screen_Refresh = trials$vsync_rate,
-                     Browser = trials$browser,
-                     Browser_Version = trials$browser_version,
-                     Device = ifelse(trials$mobile == TRUE, "Mobile", "Desktop"),
-                     Device_OS = trials$os,
-                     Illusion_Type = trials$type,
-                     Block = trials$block_number,
-                     Trial = trials$trial_number,
-                     Illusion_Strength = trials$illusion_strength,
-                     Illusion_Side = sign(trials$illusion_difference),
-                     Illusion_Difference = abs(trials$illusion_difference),
-                     Answer = unlist(trials$response),
-                     Error = as.integer(!as.logical(trials$correct)),
-                     ISI = json[json$screen == "fixation" & !is.na(json$screen), "trial_duration"],
-                     RT = trials$rt)
+  trials <- data[data$screen == "Trial", ]
+  df <- data.frame(Participant = trials$participant_id,
+                   Age = as.numeric(jsonlite::fromJSON(dem[1])$age),
+                   Sex = jsonlite::fromJSON(dem[2])$sex,
+                   Education = jsonlite::fromJSON(dem[2])$education,
[…]

### 2022-08-01 analysis.Rmd
script

@@
 
-# This is a local folder containing raw data from unzipped pavlovia
-dfraw <- read.csv("data/study1.csv") |> 
-  mutate(Screen_Refresh = as.character(Screen_Refresh),
-         Illusion_Side = as.factor(Illusion_Side))
+df <- read.csv("data/study1.csv") |>
+  mutate(
+    Screen_Refresh = as.character(Screen_Refresh),
+    Illusion_Side = as.factor(Illusion_Side),
+    Block = as.factor(Block),
+    Education = fct_relevel(Education, "Master", "Bachelor", "High School", "Other")
+  )
 ```
 
-### Exclusions
+### Exclusions {.tabset}
+
+```{r message=FALSE, warning=FALSE}
+# plot(estimate_density(dfraw[dfraw$Participant == "60684f29dbfe1bb2059e5e27_rkqoy", "RT"]))
+outliers <- c(
+  # Half of the trials are of very short RT
+  # Prolific Status: REJECTED
+  "60684f29dbfe1bb2059e5e27_rkqoy",
+  # Error rate of 47.9%
+  # Prolific Status: RETURN SUBMITTED
+  "61280140171ec546e87ed8cb_qdlgy",
+  # Error rate of 46.2%
+  # Prolific Status: RETURN SUBMITTED
+  "614f36fd81c78b7a125c4262_6ax4g",
+  # Error rate of 42.1% and very large RT SD
+  # Prolific Status: RETURN SUBMITTED
+  "5d398380b37ab1000111fac3_2nxxh",
+  # Block n2 with very short RTs
+  # Prolific Status: RETURN SUBMITTED
+  "5e860198a846e30497df5189_6e43s",
+  # Error rate of 46.2% and short RTs
+  # Prolific Status: NA
+  "615f319bb341cf2f306d858d_qsaq5"
+)
+
+```
+
+We removed `r length(outliers)` participants upon inspection of the average error rage (when close to 50%, suggesting random answers) and/or when the reaction time distribution was implausibly fast.
 
@@
 
+#### Error Rate
+
 ```{r message=FALSE, warning=FALSE}
-dfsub <- dfraw |>
+dfsub <- df |>
   group_by(Participant) |>
-  summarize(Error = sum(Error) / n(),
-            RT = mean(RT)) |>
-  ungroup()
+  summarize(
+    # n = n(),
+    Error = sum(Error) / n(),
+    RT_Mean = mean(RT),
+    RT_SD = sd(RT),
+  ) |>
+  ungroup() |>
+  arrange(desc(Error))
 
-dfraw |>
+knitr::kable(dfsub) |> 
+  kableExtra::row_spec(which(dfsub$Participant %in% outliers), background  = "#EF9A9A")
+```
+
+#### Reaction Time
+
+```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=20}
+# RT distribution
+p <- estimate_density(df, select = "RT", at = c("Participant", "Block")) |>
+  group_by(Participant) |>
+  normalize(select = "y") |>
+  ungroup() |>
+  mutate(color = ifelse(Participant %in% outliers, "red", "blue")) |>
+  ggplot(aes(x = x, y = y)) +
+  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
+  geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block)) +
+  geom_vline(xintercept = 2500, linetype = "dashed", color = "red") +
+  scale_color_manual(values=c("red"="red", "blue"="blue"), guide = "none") +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  coord_cartesian(xlim = c(0, 3000)) +
+  theme_modern() +
+  theme(axis.text.y = element_blank()) +
+  facet_wrap(~Participant) +
+  labs(y = "", x = "Reaction Time (ms)")
+p
+# ggsave("figures/outliers.png", p, width=20, height=15)
+
+# Filter out
+df <- filter(df, !Participant %in% outliers)
+```
+
+#### Error Rate per Illusion Block
+
+```{r message=FALSE, warning=FALSE}
+temp <- df |>
   group_by(Participant, Illusion_Type, Block) |>
@@
   ungroup() |> 
+  arrange(desc(ErrorRate_per_block))
+
+temp2 <- temp |> 
+  filter(ErrorRate_per_block >= 0.5) |> 
+  group_by(Illusion_Type, Block) |> 
+  summarize(n = n()) |> 
+  arrange(desc(n), Illusion_Type) |> 
+  ungroup() |> 
+  mutate(n_trials = cumsum(n * 56),
+         p_trials = n_trials / nrow(df))
+
+# knitr::kable(temp2)
+
+p1 <- temp |>
   estimate_density(at = c("Illusion_Type", "Block")) |>
   ggplot(aes(x = x, y = y)) +
-  geom_line(aes(color = Illusion_Type, linetype= as.factor(Block)))
+  geom_line(aes(color = Illusion_Type, linetype = Block)) + 
+  geom_vline(xintercept = 0.5, linetype = "dashed") +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(y = "Distribution", x = "Error Rate") +
+  theme_modern()
 
-df <- dfraw |>
+p2 <- temp2 |> 
+  mutate(Block = fct_rev(Block)) |> 
+  ggplot(aes(x = Illusion_Type, y = p_trials)) +
+  geom_bar(stat="identity", aes(fill = Block)) +
+  scale_x_discrete(expand = c(0, 0)) +
+  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
+  labs(y = "Percentage of Trials Removed", x = "Illusion Type") +
+  theme_modern()
+
+p1 | p2
+
+# Drop
+df <- df |>
   group_by(Participant, Illusion_Type, Block) |>
@@
 
-nrow(dfraw) - nrow(df)
+rm(temp, temp2)
 ```
@@
 ```{r }
-plot_distribution <- function(dfsub, what="Age", title=what, subtitle="", fill="orange") {
+plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", fill = "orange") {
   dfsub |>
@@
     geom_density(fill = fill) +
-    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype="dashed") +
+    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype = "dashed") +
     scale_x_continuous(expand = c(0, 0)) +
     scale_y_continuous(expand = c(0, 0)) +
-    ggtitle(title, subtitle=subtitle) +
+    ggtitle(title, subtitle = subtitle) +
     theme_modern() +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5),
+    theme(
+      plot.title = element_text(face = "bold", hjust = 0.5),
       plot.subtitle = element_text(face = "italic", hjust = 0.5),
@@
       axis.title.y = element_blank(),
-          axis.text.y = element_blank())
+      axis.text.y = element_blank()
+    )
 }
 
-plot_waffle <- function(data, title="Nationality") {
-  data |> 
+plot_waffle <- function(dfsub, what = "Nationality") {
+  ggwaffle::waffle_iron(dfsub, what) |>
     # mutate(label = emojifont::fontawesome('fa-twitter')) |>
@@
     coord_equal() +
-    ggtitle(title) +
+    ggtitle(what) +
     labs(fill = "") +
@@
 
-```{r }
-p1 <- plot_distribution(dfsub, "Age", fill="#FF9800")
-p2 <- plot_distribution(dfsub, "Duration", title="Total Duration", subtitle="in minutes", fill="#F44336")
-p3 <- plot_distribution(dfsub, "Break_Duration", title="Break Duration", subtitle="in minutes", fill="#3F51B5")
+```{r fig.width=20, fig.height=15}
+p1 <- plot_distribution(dfsub, "Age", fill = "#FF9800")
+p2 <- plot_distribution(dfsub, "Duration", title = "Total Duration", subtitle = "in minutes", fill = "#F44336")
+p3 <- plot_distribution(dfsub, "Break_Duration", title = "Break Duration", subtitle = "in minutes", fill = "#3F51B5")
 
-p4 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Sex)) |> 
-  plot_waffle("Sex") +
-  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63"))
+p4 <- plot_waffle(dfsub, "Sex") +
+  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))
 
-p5 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Education)) |> 
-  plot_waffle("Education") +
+p5 <- plot_waffle(dfsub, "Education") +
   scale_fill_viridis_d()
 
-p6 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Nationality)) |> 
-  plot_waffle("Nationality") +
-  scale_fill_viridis_d()
+p6 <- plot_waffle(dfsub, "Nationality") +
+  scale_fill_metro_d()
 
-p7 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Ethnicity)) |> 
-  plot_waffle("Ethnicity") +
-  scale_fill_viridis_d()
+p7 <- plot_waffle(dfsub, "Ethnicity") +
+  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0"))
 
-p8 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Screen_Resolution)) |> 
-  plot_waffle("Screen Resolution") +
-  scale_fill_viridis_d()
+p8 <- plot_waffle(dfsub, "Screen_Resolution") +
+  scale_fill_pizza_d()
 
-p9 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Device_OS)) |> 
-  plot_waffle("Device OS") +
-  scale_fill_viridis_d()
+p9 <- plot_waffle(dfsub, "Device_OS") +
+  scale_fill_bluebrown_d()
 
-# p10 <- ggwaffle::waffle_iron(dfsub, ggwaffle::aes_d(group = Screen_Refresh)) |> 
-#   plot_waffle("Screen Refresh Rate") +
+# p10 <- plot_waffle(dfsub, "Screen_Refresh") +
 #   scale_fill_viridis_d()
@@
 
-### Manipulation Checks
-
-We tested whether at illusion strength = 0, no effect of right / left (no bias of illusion at 0).
+### Outliers
 
 ```{r message=FALSE, warning=FALSE}
-model <- glmmTMB::glmmTMB(Error ~ Illusion_Type / Illusion_Side + (1|Participant), 
-                          data = filter(df, Illusion_Strength == 0), 
-                          family="binomial")
-
-parameters(model, effects="fixed") |> 
-  tail(10)
-```
-
-### Full Models
-
-#### Outliers
-
-```{r message=FALSE, warning=FALSE}
-p1 <- estimate_density(df, select="RT", at="Participant") |> 
+p1 <- estimate_density(df, select = "RT", at = "Participant") |>
+  group_by(Participant) |>
   normalize(select = "y") |>
-  ggplot(aes(x=x, y=y)) +
-  geom_area(data=normalize(estimate_density(df, select="RT"), select="y"), alpha=0.2) +
-  geom_line(aes(color=Participant, group=Participant)) + 
-  geom_vline(xintercept = 2500, linetype = "dashed", color = "red") +
-  scale_color_material_d("rainbow", guide="none") +
+  ungroup() |>
+  ggplot(aes(x = x, y = y)) +
+  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
+  geom_line(aes(color = Participant, group = Participant)) +
+  geom_vline(xintercept = 3000, linetype = "dashed", color = "red") +
+  scale_color_material_d("rainbow", guide = "none") +
   scale_x_continuous(expand = c(0, 0)) +
   scale_y_continuous(expand = c(0, 0)) +
-  coord_cartesian(xlim = c(0, 5000)) +
+  coord_cartesian(xlim = c(0, 3500)) +
   theme_modern() +
-  theme(axis.text.y = element_blank())
+  theme(axis.text.y = element_blank()) +
+  # facet_wrap(~Participant) +
+  labs(y = "", x = "Reaction Time (ms)")
 
-df$Outlier <- df$RT > 2500
+df$Outlier <- df$RT > 3000
 
@@
   ggplot(aes(x = Participant, y = Outlier)) +
-  geom_bar(stat="identity", aes(fill=Participant)) +
-  scale_fill_material_d(guide="none") +
+  geom_bar(stat = "identity", aes(fill = Participant)) +
+  scale_fill_material_d("rainbow", guide = "none") +
   scale_x_discrete(expand = c(0, 0)) +
   scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
-  theme_modern("rainbow") +
+  see::theme_modern() +
   theme(axis.text.x = element_blank())
@@
 p1 | p2
+```
 
+We removed `r sum(df$Outlier)` (`r insight::format_value(sum(df$Outlier) / nrow(df), as_percent=TRUE)`) outliers trials (RT > 3000 ms).
+
+```{r message=FALSE, warning=FALSE}
 df <- filter(df, Outlier == FALSE)
 ```
+
+### Manipulation Checks
+
+We tested whether at illusion strength = 0, no effect of right / left (no bias of illusion at 0).
+
+```{r message=FALSE, warning=FALSE}
+model <- glmmTMB::glmmTMB(Error ~ Illusion_Type / Illusion_Side + (1 | Participant),
+  data = filter(df, Illusion_Strength == 0),
+  family = "binomial"
+)
+
+parameters(model, effects = "fixed") |>
+  tail(10) |> 
+  arrange(p)
+
+# select(filter(df, Error == 0, Illusion_Type == "Müller-Lyer"), Illusion_Side, Answer)
+# select(filter(df, Error == 0, Illusion_Type == "Vertical-Horizontal"), Illusion_Side, Answer)
+# select(filter(df, Error == 0, Illusion_Type == "Zöllner"), Illusion_Side, Answer)
+```
+
+Müller-Lyer: when no illusion, it is easier to answer that the upper line is longer.
+
+### Full Models
 
@@
 model_errors <- function(df, illusion = "Delboeuf") {
-  data <- filter(df, Illusion_Type == illusion)
+  data <- filter(df, Illusion_Type == illusion) |>
+    mutate(.dots_side = ifelse(Error == 1, "bottom", "top"))
 
@@
   # )
-  model <- mgcv::gamm(
-    Error ~ s(Illusion_Difference, by = Illusion_Strength, k=4),
-    random = list(Participant = ~1),
+  # model <- mgcv::gamm(
+  #   # Error ~ s(Illusion_Difference, by = Illusion_Strength, k = 4),
+  #   Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4),
+  #   random = list(Participant = ~1),
+  #   data = data,
+  #   family = "binomial"
+  # )
+  model <- glmmTMB::glmmTMB(
+    # Error ~ s(Illusion_Difference, by = Illusion_Strength, k = 4),
+    Error ~ poly(Illusion_Difference, 2) * poly(Illusion_Strength, 2) * Illusion_Side + (1 | Participant),
     data = data,
@@
   n_lines <- 5
-  pred <- estimate_relation(model, at = c("Illusion_Strength", "Illusion_Difference"), length = c(30, n_lines)) |>
-    mutate(Illusion_Difference = as.factor(Illusion_Difference),
-           Illusion_Type = illusion)
+  pred <- estimate_relation(model,
+                            at = c("Illusion_Strength", "Illusion_Difference"),
+                            # at = list(Illusion_Strength = unique(data$Illusion_Strength), Illusion_Difference = unique(data$Illusion_Difference)), 
+                            length = c(5, n_lines)) |>
+    mutate(
+      Illusion_Difference = as.factor(Illusion_Difference),
+      Illusion_Type = illusion
+    )
 
   p <- pred |>
-    ggplot(aes(x = Illusion_Strength, y = Predicted, group = Illusion_Difference)) +
-    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference), alpha = 0.2) +
+    ggplot(aes(x = Illusion_Strength, y = Predicted)) +
+    geom_dots(
+      data = data, dotsize = 2,
+      aes(y = Error, group = Error, side = .dots_side),
+      fill = ifelse(data$Error == 1, "red", "green"), color = NA
+    ) +
+    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference, group = Illusion_Difference), alpha = 0.2) +
     geom_vline(xintercept = 0, linetype = "dashed") +
-    geom_hline(yintercept=c(0.05, 0.5, 0.95), linetype="dotted", alpha=0.5) +
-    geom_line(aes(color = Illusion_Difference)) +
+    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
+    geom_line(aes(color = Illusion_Difference, group = Illusion_Difference)) +
     scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
@@
 rez_ebbinghaus <- model_errors(df, illusion = "Ebbinghaus")
+rez_verticalhorizontal <- model_errors(df, illusion = "Vertical-Horizontal")
+rez_rodframe <- model_errors(df, illusion = "Rod-Frame")
+rez_zollner <- model_errors(df, illusion = "Zöllner")
+rez_poggendorff <- model_errors(df, illusion = "Poggendorff")
 rez_contrast <- model_errors(df, illusion = "Contrast")
 rez_white <- model_errors(df, illusion = "White")
-rez_poggendorff <- model_errors(df, illusion = "Poggendorff")
-rez_zollner <- model_errors(df, illusion = "Zöllner")
-rez_rodframe <- model_errors(df, illusion = "Rod-Frame")
-rez_verticalhorizontal <- model_errors(df, illusion = "Vertical-Horizontal")
 ```
@@
 ```{r message=FALSE, warning=FALSE}
+parameters::parameters(rez_ponzo$model)[c("Parameter", "p")]
 rez_ponzo$p

### 2022-08-01 preprocessing.R

@@ -1,7 +1,11 @@
 # Preprocessing Function
 preprocess_raw <- function(file) {
-
   data <- read.csv(file)
+
+  if (!"final_results" %in% data$screen) {
+    print(paste0("Warning: Incomplete data for ", file))
+    return(data.frame())
+  }
 
   # Demographics
   dem <- data[data$screen == "demographics" & !is.na(data$screen), "response"]
@@ -11,34 +15,37 @@ preprocess_raw <- function(file) {
 
 
   trials <- data[data$screen == "Trial", ]
-  df <- data.frame(Participant = trials$participant_id,
-                   Age = as.numeric(jsonlite::fromJSON(dem[1])$age),
-                   Sex = jsonlite::fromJSON(dem[2])$sex,
-                   Education = jsonlite::fromJSON(dem[2])$education,
-                   Nationality = tools::toTitleCase(jsonlite::fromJSON(dem[1])$nationality),
-                   Ethnicity = tools::toTitleCase(jsonlite::fromJSON(dem[1])$ethnicity),
-                   Date = ifelse(is.null(info$date), NA, info$date),
-                   Time = ifelse(is.null(info$time), NA, info$time),
-                   Duration = as.numeric(data[data$screen == "final_results", "time_elapsed"]) / 1000 / 60,
-                   Break_Duration = as.numeric(data[data$screen == "break" & !is.na(data$screen), "rt"]) / 1000 / 60,
-                   Screen_Resolution = paste0(trials$screen_width, "x", trials$screen_height),
-                   Screen_Refresh = trials$vsync_rate,
-                   Browser = trials$browser,
-                   Browser_Version = trials$browser_version,
-                   Device = ifelse(trials$mobile == TRUE, "Mobile", "Desktop"),
-                   Device_OS = trials$os,
-                   Illusion_Type = trials$type,
-                   Block = ifelse(trials$block_number > 10, 2, 1),
-                   Block_Order = as.numeric(trials$block_number),
-                   Trial = as.numeric(trials$trial_number),
-                   Stimulus = gsub(".png", "", gsub("stimuli/", "", trials$stimulus)),
[…]

### 2022-08-02 study1.Rmd
update script

---
title: "Illusion Game Validation (Study 1)"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3)

cache <- FALSE
fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

# Introduction

Behavioural findings regarding the Illusion Game.

# Methods

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

df <- read.csv("data/study1.csv") |>
  mutate(
    Date = as.Date(Date),
    Participant = fct_reorder(Participant, Date),
    Screen_Refresh = as.character(Screen_Refresh),
    Illusion_Side = as.factor(Illusion_Side),
    Block = as.factor(Block),
    Education = fct_relevel(Education, "Master", "Bachelor", "High School", "Other")
  )
```

## Exclusions {.tabset}

```{r message=FALSE, warning=FALSE}
# plot(estimate_density(dfraw[dfraw$Participant == "60684f29dbfe1bb2059e5e27_rkqoy", "RT"]))

# Dear participant, thank you for participating in our study. Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which make the data unusable for us. We understand that you might have been in a hurry, or had some other issues, and kindly ask you to return your participation. We hope to open-up more slots in the future would you be interested to participate again.

outliers <- c(
  # Half of the trials are of very short RT
  # Prolific Status: REJECTED
  "60684f29dbfe1bb2059e5e27_rkqoy",
  # Error rate of 47.9%
  # Prolific Status: RETURN REQUESTED
  "61280140171ec546e87ed8cb_qdlgy",
  # Error rate of 46.2%
  # Prolific Status: RETURN REQUESTED
  "614f36fd81c78b7a125c4262_6ax4g",
  # Error rate of 42.1% and very large RT SD
  # Prolific Status: RETURN REQUESTED
  "5d398380b37ab1000111fac3_2nxxh",
  # Block n2 with very short RTs
  # Prolific Status: RETURNED
  "5e860198a846e30497df5189_6e43s",
  # Error rate of 46.2% and short RTs
  # Prolific Status: RETURN REQUESTED
  "615f319bb341cf2f306d858d_qsaq5"
)
```

We removed `r length(outliers)` participants upon inspection of the average error rage (when close to 50%, suggesting random answers) and/or when the reaction time distribution was implausibly fast.

For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).

### Error Rate

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  group_by(Participant) |>
  summarize(
    # n = n(),
    Error = sum(Error) / n(),
    RT_Mean = mean(RT),
    RT_SD = sd(RT),
  ) |>
  ungroup() |>
  arrange(desc(Error))

knitr::kable(dfsub) |> 
  kableExtra::row_spec(which(dfsub$Participant %in% outliers), background  = "#EF9A9A")
```

### Error Rate per Illusion Block

```{r message=FALSE, warning=FALSE}
temp <- df |>
  group_by(Participant, Illusion_Type, Block) |>
  summarize(ErrorRate_per_block = sum(Error) / n()) |>
  ungroup() |> 
  arrange(desc(ErrorRate_per_block))

temp2 <- temp |> 
  filter(ErrorRate_per_block >= 0.5) |> 
  group_by(Illusion_Type, Block) |> 
  summarize(n = n()) |> 
  arrange(desc(n), Illusion_Type) |> 
  ungroup() |> 
  mutate(n_trials = cumsum(n * 56),
         p_trials = n_trials / nrow(df))

# knitr::kable(temp2)

p1 <- temp |>
  estimate_density(at = c("Illusion_Type", "Block")) |>
  ggplot(aes(x = x, y = y)) +
  geom_line(aes(color = Illusion_Type, linetype = Block)) + 
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(y = "Distribution", x = "Error Rate") +
  theme_modern()

p2 <- temp2 |> 
  mutate(Block = fct_rev(Block)) |> 
  ggplot(aes(x = Illusion_Type, y = p_trials)) +
  geom_bar(stat="identity", aes(fill = Block)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
  labs(y = "Percentage of Trials Removed", x = "Illusion Type") +
  theme_modern() +
  theme(axis.text.x = element_text(angle=45, hjust = 1))

p1 | p2

# Drop
df <- df |>
  group_by(Participant, Illusion_Type, Block) |>
  mutate(ErrorRate_per_block = sum(Error) / n()) |>
  ungroup() |> 
  filter(ErrorRate_per_block < 0.5) |>
  select(-ErrorRate_per_block)

rm(temp, temp2)
```

### Reaction Time Distribution

```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=20}
# RT distribution
p <- estimate_density(df, select = "RT", at = c("Participant", "Block")) |>
  group_by(Participant) |>
  normalize(select = "y") |>
  ungroup() |>
  mutate(color = ifelse(Participant %in% outliers, "red", "blue")) |>
  ggplot(aes(x = x, y = y)) +
  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
  geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block)) +
  geom_vline(xintercept = 2500, linetype = "dashed", color = "red") +
  scale_color_manual(values=c("red"="red", "blue"="blue"), guide = "none") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(xlim = c(0, 3000)) +
  theme_modern() +
  theme(axis.text.y = element_blank()) +
  facet_wrap(~Participant) +
  labs(y = "", x = "Reaction Time (ms)")
p
# ggsave("figures/outliers.png", p, width=20, height=15)

# Filter out
df <- filter(df, !Participant %in% outliers)
```

### Reaction Time per Trial

```{r message=FALSE, warning=FALSE}
p1 <- estimate_density(df, select = "RT", at = "Participant") |>
  group_by(Participant) |>
  normalize(select = "y") |>
  ungroup() |>
  ggplot(aes(x = x, y = y)) +
  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
  geom_line(aes(color = Participant, group = Participant)) +
  geom_vline(xintercept = c(150, 3000), linetype = "dashed", color = "red") +
  scale_color_material_d("rainbow", guide = "none") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(xlim = c(0, 3500)) +
  theme_modern() +
  theme(axis.text.y = element_blank()) +
  # facet_wrap(~Participant) +
  labs(y = "", x = "Reaction Time (ms)")

df$Outlier <- df$RT < 150 | df$RT > 3000

p2 <- df |>
  group_by(Participant) |>
  summarize(Outlier = sum(Outlier) / n()) |>
  mutate(Participant = fct_reorder(Participant, Outlier)) |>
  ggplot(aes(x = Participant, y = Outlier)) +
  geom_bar(stat = "identity", aes(fill = Participant)) +
  scale_fill_material_d("rainbow", guide = "none") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
  see::theme_modern() +
  theme(axis.text.x = element_blank())

p1 | p2
```

We removed `r sum(df$Outlier)` (`r insight::format_value(sum(df$Outlier) / nrow(df), as_percent=TRUE)`) outlier trials (150 ms < RT < 3000 ms).

```{r message=FALSE, warning=FALSE}
df <- filter(df, Outlier == FALSE)
```

## Participants

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  group_by(Participant) |>
  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS) |>
  slice(1) |>
  ungroup()
```

The final sample included `r report::report_participants(dfsub, age="Age", sex="Sex")`.

```{r }
plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", fill = "orange") {
  dfsub |>
    ggplot(aes_string(x = what)) +
    geom_density(fill = fill) +
    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype = "dashed") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggtitle(title, subtitle = subtitle) +
    theme_modern() +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      plot.subtitle = element_text(face = "italic", hjust = 0.5),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank()
    )
}

plot_waffle <- function(dfsub, what = "Nationality") {
  ggwaffle::waffle_iron(dfsub, what) |>
    # mutate(label = emojifont::fontawesome('fa-twitter')) |>
    ggplot(aes(x, y, fill = group)) +
    ggwaffle::geom_waffle() +
    # geom_point() +
    # geom_text(aes(label=label), family='fontawesome-webfont', size=4) +
    coord_equal() +
    ggtitle(what) +
    labs(fill = "") +
    theme_void() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}
```

```{r fig.width=20, fig.height=15}
p1 <- plot_distribution(dfsub, "Age", fill = "#FF9800")
p2 <- plot_distribution(dfsub, "Duration", title = "Total Duration", subtitle = "in minutes", fill = "#F44336")
p3 <- plot_distribution(dfsub, "Break_Duration", title = "Break Duration", subtitle = "in minutes", fill = "#3F51B5")

p4 <- plot_waffle(dfsub, "Sex") +
  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))

p5 <- plot_waffle(dfsub, "Education") +
  scale_fill_viridis_d()

p6 <- plot_waffle(dfsub, "Nationality") +
  scale_fill_metro_d()

p7 <- plot_waffle(dfsub, "Ethnicity") +
  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0"))

p8 <- plot_waffle(dfsub, "Screen_Resolution") +
  scale_fill_pizza_d()

p9 <- plot_waffle(dfsub, "Device_OS") +
  scale_fill_bluebrown_d()

# p10 <- plot_waffle(dfsub, "Screen_Refresh") +
#   scale_fill_viridis_d()

(p1 / p2 / p3) | (p4 / p5 / p6) | (p7 / p8 / p9)
```

# Results

## Delboeuf

### Descriptive

```{r message=FALSE, warning=FALSE, cache=cache}
plot_descriptive <- function(data, side="leftright") {

  if(side == "leftright") {
    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
    x <- tools::toTitleCase(gsub("arrow", "", x))
    if(x == "Left") data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
  }

  dodge1 <- 0.1 * diff(range(data$Illusion_Difference))
  dodge2 <- -0.1 * diff(range(data$Illusion_Strength))

  colors <- colorRampPalette(c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0"))(length(unique(data$Illusion_Strength)))

  p1 <- data |> 
    group_by(Illusion_Difference, Illusion_Strength, Answer) |> 
    summarize(Error = mean(Error)) |> 
    mutate(Illusion_Strength = as.factor(Illusion_Strength)) |> 
    ggplot(aes(x = Illusion_Difference, y = Error)) +
    geom_bar(aes(fill=Illusion_Strength), position = position_dodge(width=dodge1), stat="identity") +
    geom_line(aes(color = Illusion_Strength), position = position_dodge(width=dodge1)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_manual(values = colors) +
    scale_fill_manual(values = colors) +
    theme_modern() +
    labs(
      color = "Illusion Strength", fill = "Illusion Strength",
      y = "Probability of Error",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data$Illusion_Difference)))

  p2 <- data |> 
    group_by(Illusion_Difference, Illusion_Strength, Answer) |> 
    summarize(Error = mean(Error)) |> 
    mutate(Illusion_Difference = as.factor(round(Illusion_Difference, 2))) |> 
    ggplot(aes(x = Illusion_Strength, y = Error)) +
    geom_bar(aes(fill=Illusion_Difference), position = position_dodge(width=dodge2), stat="identity") +
    geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_manual(values = colors) +
    scale_fill_manual(values = colors) +
    theme_modern() +
    labs(
      color = "Task Difficulty", fill = "Task Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  if(side == "leftright") {
    p <- ((p1 + facet_wrap(~Answer, ncol=2, labeller = "label_both")) /
      (p2 + facet_wrap(~Answer, ncol=2, labeller = "label_both"))) + 
    plot_annotation(title = "Delboeuf Illusion", 
                    theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
  }
  p
}

data <- filter(df, Illusion_Type == "Delboeuf")

plot_descriptive(data)
```

### Model Selection

```{r message=FALSE, warning=FALSE, cache=cache}
models <- list()
for(i in 1:3) {
  for(j in 1:3) {
    m <- glmmTMB::glmmTMB(
      Error ~ poly(Illusion_Difference, i) * poly(Illusion_Strength, j) + 
        (1 | Participant),
      data = data,
      family = "binomial"
      )
    if(performance::check_convergence(m)) {
      models[[paste0("dif", i, "-str", j)]] <- m
    }
    # m <- glmmTMB::glmmTMB(
    #   Error ~ poly(log(Illusion_Difference), i) * poly(Illusion_Strength, j) + 
    #     (1 | Participant),
    #   data = data,
    #   family = "binomial"
    #   )
    # if(performance::check_convergence(m)) {
    #   models[[paste0("dif(log(", i, "))-str", j)]] <- m
    # }
    m <- glmmTMB::glmmTMB(
      Error ~ Illusion_Side * poly(Illusion_Difference, i) * poly(Illusion_Strength, j) + 
        (1 | Participant),
      data = data,
      family = "binomial"
      )
    if(performance::check_convergence(m)) {
      models[[paste0("side-dif", i, "-str", j)]] <- m
    }
  }
}

test <- test_performance(models, reference=3)
perf <- compare_performance(models, metrics = c("BIC", "R2"))

merge(perf, test) |> 
  arrange(BIC) |> 
  select(Name, BIC, R2_marginal, BF) |> 
  mutate(BF = insight::format_bf(BF, name=""))
```

### Model

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
    (1 | Participant),
  family = "bernoulli"
)

model <- brms::brm(formula,
  data = data,
  algorithm = "meanfield",
  refresh = 0
)
```

```{r message=FALSE, warning=FALSE, cache=cache}
plot_model <- function(data, model) {
  data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))

  n_lines <- 5
  pred <- estimate_relation(model,
                            at = c("Illusion_Strength", "Illusion_Difference"),
                            length = c(5, n_lines)) |>
    mutate(Illusion_Difference = as.factor(round(Illusion_Difference, 2)))

  # Set colors for lines
  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)
  diffvals <- as.numeric(as.character(unique(sort(pred$Illusion_Difference))))
  names(colors) <- diffvals

  # Assign color from the same palette to every observation of data (for geom_dots)
  data$color <- colors[as.character(diffvals[max.col(-abs(outer(data$Illusion_Difference, diffvals, "-")))])]

  # Manual jittering
  xrange <- 0.05*diff(range(data$Illusion_Strength))
  data$x <- data$Illusion_Strength 
  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)

  pred |>
    ggplot(aes(x = Illusion_Strength, y = Predicted)) +
    geom_dots(
      data = mutate(data, 
                    Illusion_Strength = round(Illusion_Strength, 1)),
      aes(x=x, y = Error, group = Error, side = .dots_side), 
      fill = data$color,
      color = NA, 
      alpha=0.5) +
    geom_slab(data=data, aes(y = Error)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference, group = Illusion_Difference), alpha = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
    geom_line(aes(color = Illusion_Difference, group = Illusion_Difference)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_manual(values = colors) +
    scale_fill_manual(values = colors) +
    coord_cartesian(xlim=c(min(data$Illusion_Strength), max(data$Illusion_Strength))) +
    theme_modern() +
    labs(
      title = paste0(data$Illusion_Type[1], " Illusion"),
      color = "Difficulty", fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}

plot_model(data, model)
```

### Parameter

## Manipulation Checks {.tabset}

### Illusion Side

We tested whether at illusion strength = 0, no effect of right / left (no bias of illusion at 0).

```{r message=FALSE, warning=FALSE}
model <- glmmTMB::glmmTMB(Error ~ Illusion_Type / Illusion_Side + (1 | Participant),
  data = filter(df, Illusion_Strength == 0),
  family = "binomial"
)

parameters(model, effects = "fixed") |>
  tail(10) |> 
  arrange(p)

# select(filter(df, Error == 0, Illusion_Type == "Müller-Lyer"), Illusion_Side, Answer)
# select(filter(df, Error == 0, Illusion_Type == "Vertical-Horizontal"), Illusion_Side, Answer)
# select(filter(df, Error == 0, Illusion_Type == "Zöllner"), Illusion_Side, Answer)
```

Müller-Lyer: when no illusion, it is easier to answer that the upper line is longer.

### ISI

```{r message=FALSE, warning=FALSE}
model <- glmmTMB::glmmTMB(Error ~ Illusion_Type / poly(ISI, 2) + (1 | Participant),
  data = df,
  family = "binomial"
)

parameters(model, effects = "fixed") |>
  tail(20)

# select(filter(df, Error == 0, Illusion_Type == "Müller-Lyer"), Illusion_Side, Answer)
# select(filter(df, Error == 0, Illusion_Type == "Vertical-Horizontal"), Illusion_Side, Answer)
# select(filter(df, Error == 0, Illusion_Type == "Zöllner"), Illusion_Side, Answer)
```

## Full Models

### Errors {.tabset}

##### Ebbinghaus

###### Model Selection

```{r message=FALSE, warning=FALSE, cache=cache}
data <- filter(df, Illusion_Type == "Delboeuf")

models <- list()
for(i in 1:3) {
  for(j in 1:3) {
    m <- glmmTMB::glmmTMB(
      Error ~ poly(Illusion_Difference, i) * poly(Illusion_Strength, j) + 
        (1 | Participant),
      data = data,
      family = "binomial"
      )
    if(performance::check_convergence(m)) {
      models[[paste0("dif", i, "-str", j)]] <- m
    }
    m <- glmmTMB::glmmTMB(
      Error ~ poly(log(Illusion_Difference), i) * poly(Illusion_Strength, j) + 
        (1 | Participant),
      data = data,
      family = "binomial"
      )
    if(performance::check_convergence(m)) {
      models[[paste0("dif(log(", i, "))-str", j)]] <- m
    }
    m <- glmmTMB::glmmTMB(
      Error ~ poly(sqrt(Illusion_Difference), i) * poly(Illusion_Strength, j) + 
        (1 | Participant),
      data = data,
      family = "binomial"
      )
    if(performance::check_convergence(m)) {
      models[[paste0("dif(sqrt(", i, "))-str", j)]] <- m
    }
  }
}

compare_performance(models, metrics = c("BIC", "R2")) |> 
  merge(test_performance(models, reference=6)) |> 
  arrange(BIC) |> 
  select(Name, BIC, R2_marginal, BF) |> 
  mutate(BF = insight::format_bf(BF, name=""))
```

```{r message=FALSE, warning=FALSE}
model_errors <- function(df, illusion = "Delboeuf") {
  data <- filter(df, Illusion_Type == illusion) |>
    mutate(.dots_side = ifelse(Error == 1, "bottom", "top"))

  if (illusion == "Müller-Lyer") {
    formula <- brms::bf(
      # Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4) + (1 | Participant),
      # Error ~ s(Illusion_Strength, by=Illusion_Difference, k = 7) + (1 | Participant),
      Error ~ poly(Illusion_Difference, 3) * poly(Illusion_Strength, 3) * Illusion_Side + 
        (1 | Participant),
      family = "bernoulli"
    )
  } else {
    formula <- brms::bf(
      # Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4) + (1 | Participant),
      # Error ~ s(Illusion_Strength, by=Illusion_Difference, k = 7) + (1 | Participant),
      Error ~ poly(Illusion_Difference, 3) * poly(Illusion_Strength, 3) + 
        (1 | Participant),
      family = "bernoulli"
    )
  }

  model <- brms::brm(formula,
    data = data,
    algorithm = "meanfield",
    refresh = 0
  )

  # model <- glmmTMB::glmmTMB(
  #   Error ~ Illusion_Difference * Illusion_Strength + (1 | Participant),
  #   data = data,
  #   family = "binomial"
  # )
  # model <- mgcv::gamm(
  #   # Error ~ s(Illusion_Difference, by = Illusion_Strength, k = 4),
  #   Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4),
  #   random = list(Participant = ~1),
  #   data = data,
  #   family = "binomial"
  # )
  # model <- glmmTMB::glmmTMB(
  #   # Error ~ s(Illusion_Difference, by = Illusion_Strength, k = 4),
  #   Error ~ poly(Illusion_Difference, 2) * poly(Illusion_Strength, 2) * Illusion_Side + (1 | Participant),
  #   data = data,
  #   family = "binomial"
  # )

  n_lines <- 5
  pred <- estimate_relation(model,
                            at = c("Illusion_Strength", "Illusion_Difference"),
                            # at = list(Illusion_Strength = unique(data$Illusion_Strength), Illusion_Difference = unique(data$Illusion_Difference)), 
                            length = c(5, n_lines)) |>
    mutate(
      Illusion_Difference = as.factor(Illusion_Difference),
      Illusion_Type = illusion
    )

  # Set colors for lines
  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)
  diffvals <- as.numeric(as.character(unique(sort(pred$Illusion_Difference))))
  names(colors) <- diffvals

  # Assign color from the same palette to every observation of data (for geom_dots)
  data$color <- colors[as.character(diffvals[max.col(-abs(outer(data$Illusion_Difference, diffvals, "-")))])]

  # Manual jittering
  xrange <- 0.05*diff(range(data$Illusion_Strength))
  data$x <- data$Illusion_Strength 
  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)

  p <- pred |>
    ggplot(aes(x = Illusion_Strength, y = Predicted)) +
    geom_dots(
      data = mutate(data, Illusion_Strength = round(Illusion_Strength, 1)),
      aes(x=x, y = Error, group = Error, side = .dots_side), fill = data$color,
      color = NA, alpha=0.5) +
    # geom_dots(
    #   data = mutate(data, Illusion_Strength = round(Illusion_Strength, 1)),
    #   aes(y = Error, group = Error, side = .dots_side),
    #   fill = ifelse(data$Error == 1, "red", "green"), color = NA, alpha=0.2,
    #   position = position_jitter(width=0.05*diff(range(data$Illusion_Strength)), height=0)
    # ) +
    geom_slab(data=data, aes(y = Error)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference, group = Illusion_Difference), alpha = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
    geom_line(aes(color = Illusion_Difference, group = Illusion_Difference)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_manual(values = colors) +
    scale_fill_manual(values = colors) +
    theme_modern() +
    labs(
      title = paste0(illusion, " Illusion"),
      color = "Difficulty", fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))

  list(p = p, model = model, pred = pred)
}
```

## References

### 2022-08-02 preprocessing.R

@@ -83,14 +83,18 @@ preprocess_raw <- function(file) {
 # This is a local folder containing raw data from unzipped pavlovia
 # It has been added to .gitignore to NOT be published on github
 # (it contains the subject ID of the participants)
-participants <- list.files("rawdata/")
+participants <- list.files("rawdata/study1/")
 
 
 df <- data.frame()
 for (ppt in participants) {
-  df <- rbind(df, preprocess_raw(file = paste0("rawdata/", ppt)))
+  df <- rbind(df, preprocess_raw(file = paste0("rawdata/study1/", ppt)))
 }
-df$Study <- 1 # Study 1
+
+
+# Study 1
+df$Study <- 1
+df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"])
 
 # Save anonmized data
 write.csv(df, "data/study1.csv", row.names = FALSE)

### 2022-08-03 study1.Rmd
move files

@@
 ```{r, echo = FALSE, warning=FALSE, message=FALSE}
-options(digits = 3)
+options(digits = 3,
+        mc.cores = 2)
 
@@
 
-# Results
+# Results {.tabset}
 
@@
     x <- tools::toTitleCase(gsub("arrow", "", x))
-    if(x == "Left") data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
+    if(x == "Left") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
+    } else {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
+    }
+  } else {
+    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
+    x <- tools::toTitleCase(gsub("arrow", "", x))
+    if(x == "Up") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
+    } else {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
+    }
+    data$Answer <- fct_rev(data$Answer)
   }
@@
     summarize(Error = mean(Error)) |> 
-    mutate(Illusion_Strength = as.factor(Illusion_Strength)) |> 
+    mutate(Illusion_Strength = as.factor(round(Illusion_Strength, 2))) |> 
     ggplot(aes(x = Illusion_Difference, y = Error)) +
@@
     labs(
-      color = "Illusion Strength", fill = "Illusion Strength",
+      color = "Illusion Strength", 
+      fill = "Illusion Strength",
       y = "Probability of Error",
@@
     ggplot(aes(x = Illusion_Strength, y = Error)) +
+    geom_vline(xintercept=0, linetype="dotted", alpha=0.6) +
     geom_bar(aes(fill=Illusion_Difference), position = position_dodge(width=dodge2), stat="identity") +
@@
     labs(
-      color = "Task Difficulty", fill = "Task Difficulty",
+      color = "Task Difficulty", 
+      fill = "Task Difficulty",
       y = "Probability of Error",
@@
       (p2 + facet_wrap(~Answer, ncol=2, labeller = "label_both"))) + 
-    plot_annotation(title = "Delboeuf Illusion", 
+    plot_annotation(title = paste(data$Illusion_Type[1], "Illusion"), 
+                    theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
+  } else {
+    p <- ((p1 + facet_wrap(~Answer, nrow=2, labeller = "label_both")) |
+      (p2 + facet_wrap(~Answer, nrow=2, labeller = "label_both"))) + 
+    plot_annotation(title = paste(data$Illusion_Type[1], "Illusion"), 
                     theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models <- function(data) {
   models <- list()
-for(i in 1:3) {
-  for(j in 1:3) {
-    m <- glmmTMB::glmmTMB(
-      Error ~ poly(Illusion_Difference, i) * poly(Illusion_Strength, j) + 
-        (1 | Participant),
-      data = data,
-      family = "binomial"
-      )
+  for(i in 1:2) {
+    for(j in 1:2) {
+      for(k1 in c("", "_log")) {  # c("", "_log")
+        for(k2 in c("", "_log")) {  # c("", "_log")
+          for(side in c("", "-side")) {
+            name <- paste0("dif", k1, i, "-", "str", k2, j, side)
+            print(name)
+            f <- paste0("poly(Illusion_Difference", 
+                        k1,
+                        ", ",
+                        i,
+                        ") * poly(Illusion_Strength",
+                        k2, 
+                        ", ",
+                        j, 
+                        ") + (1|Participant)")
+
+            if(side == "-side") f <- paste0("Illusion_Side * ", f)
+
+            m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)), 
+                                  data = data, family = "binomial")
             if(performance::check_convergence(m)) {
-      models[[paste0("dif", i, "-str", j)]] <- m
+              models[[name]] <- m
             }
-    # m <- glmmTMB::glmmTMB(
-    #   Error ~ poly(log(Illusion_Difference), i) * poly(Illusion_Strength, j) + 
-    #     (1 | Participant),
-    #   data = data,
-    #   family = "binomial"
-    #   )
-    # if(performance::check_convergence(m)) {
-    #   models[[paste0("dif(log(", i, "))-str", j)]] <- m
-    # }
-    m <- glmmTMB::glmmTMB(
-      Error ~ Illusion_Side * poly(Illusion_Difference, i) * poly(Illusion_Strength, j) + 
-        (1 | Participant),
-      data = data,
-      family = "binomial"
-      )
-    if(performance::check_convergence(m)) {
-      models[[paste0("side-dif", i, "-str", j)]] <- m
+          }
+        }
       }
@@
 
-test <- test_performance(models, reference=3)
-perf <- compare_performance(models, metrics = c("BIC", "R2"))
+  to_keep <- compare_performance(models, metrics = c("BIC")) |> 
+    arrange(BIC) |> 
+    slice(1:10) |> 
+    pull(Name)
+
+  test <- test_performance(models[to_keep], reference=1)
+  perf <- compare_performance(models[to_keep], metrics = c("BIC", "R2"))
 
@@
     mutate(BF = insight::format_bf(BF, name=""))
+}
+
+best_models(data)
 ```
 
-### Model
+### Model Visualization
 
@@
   data = data,
-  algorithm = "meanfield",
-  refresh = 0
+  algorithm = "sampling",
+  refresh = 200
 )
@@
                             at = c("Illusion_Strength", "Illusion_Difference"),
-                            length = c(5, n_lines)) |>
+                            length = c(25, n_lines)) |>
     mutate(Illusion_Difference = as.factor(round(Illusion_Difference, 2)))
@@
 
-### Parameter
-
-## Manipulation Checks {.tabset}
-
-### Illusion Side
-
-We tested whether at illusion strength = 0, no effect of right / left (no bias of illusion at 0).
-
-```{r message=FALSE, warning=FALSE}
-model <- glmmTMB::glmmTMB(Error ~ Illusion_Type / Illusion_Side + (1 | Participant),
-  data = filter(df, Illusion_Strength == 0),
-  family = "binomial"
-)
-
-parameters(model, effects = "fixed") |>
-  tail(10) |> 
-  arrange(p)
-
-# select(filter(df, Error == 0, Illusion_Type == "Müller-Lyer"), Illusion_Side, Answer)
-# select(filter(df, Error == 0, Illusion_Type == "Vertical-Horizontal"), Illusion_Side, Answer)
-# select(filter(df, Error == 0, Illusion_Type == "Zöllner"), Illusion_Side, Answer)
-```
-
-Müller-Lyer: when no illusion, it is easier to answer that the upper line is longer.
-
-### ISI
-
-```{r message=FALSE, warning=FALSE}
-model <- glmmTMB::glmmTMB(Error ~ Illusion_Type / poly(ISI, 2) + (1 | Participant),
-  data = df,
-  family = "binomial"
-)
-
-parameters(model, effects = "fixed") |>
-  tail(20)
-
-# select(filter(df, Error == 0, Illusion_Type == "Müller-Lyer"), Illusion_Side, Answer)
-# select(filter(df, Error == 0, Illusion_Type == "Vertical-Horizontal"), Illusion_Side, Answer)
-# select(filter(df, Error == 0, Illusion_Type == "Zöllner"), Illusion_Side, Answer)
-```
-
-## Full Models
-
-### Errors {.tabset}
-
-##### Ebbinghaus
-
-###### Model Selection
+### GAM
 
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Delboeuf")
+make_gam <- function(data) {
 
-models <- list()
-for(i in 1:3) {
-  for(j in 1:3) {
-    m <- glmmTMB::glmmTMB(
-      Error ~ poly(Illusion_Difference, i) * poly(Illusion_Strength, j) + 
-        (1 | Participant),
-      data = data,
-      family = "binomial"
-      )
-    if(performance::check_convergence(m)) {
-      models[[paste0("dif", i, "-str", j)]] <- m
-    }
-    m <- glmmTMB::glmmTMB(
-      Error ~ poly(log(Illusion_Difference), i) * poly(Illusion_Strength, j) + 
-        (1 | Participant),
-      data = data,
-      family = "binomial"
-      )
-    if(performance::check_convergence(m)) {
-      models[[paste0("dif(log(", i, "))-str", j)]] <- m
-    }
-    m <- glmmTMB::glmmTMB(
-      Error ~ poly(sqrt(Illusion_Difference), i) * poly(Illusion_Strength, j) + 
-        (1 | Participant),
-      data = data,
-      family = "binomial"
-      )
-    if(performance::check_convergence(m)) {
-      models[[paste0("dif(sqrt(", i, "))-str", j)]] <- m
-    }
-  }
-}
-
-compare_performance(models, metrics = c("BIC", "R2")) |> 
-  merge(test_performance(models, reference=6)) |> 
-  arrange(BIC) |> 
-  select(Name, BIC, R2_marginal, BF) |> 
-  mutate(BF = insight::format_bf(BF, name=""))
-```
-
-```{r message=FALSE, warning=FALSE}
-model_errors <- function(df, illusion = "Delboeuf") {
-  data <- filter(df, Illusion_Type == illusion) |>
-    mutate(.dots_side = ifelse(Error == 1, "bottom", "top"))
-
-  if (illusion == "Müller-Lyer") {
   formula <- brms::bf(
-      # Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4) + (1 | Participant),
-      # Error ~ s(Illusion_Strength, by=Illusion_Difference, k = 7) + (1 | Participant),
-      Error ~ poly(Illusion_Difference, 3) * poly(Illusion_Strength, 3) * Illusion_Side + 
-        (1 | Participant),
+    Error ~ t2(Illusion_Difference, Illusion_Strength, bs = "cr", k=4) + 
+      ((1 + Illusion_Difference + Illusion_Strength) | Participant),
     family = "bernoulli"
   )
-  } else {
-    formula <- brms::bf(
-      # Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4) + (1 | Participant),
-      # Error ~ s(Illusion_Strength, by=Illusion_Difference, k = 7) + (1 | Participant),
-      Error ~ poly(Illusion_Difference, 3) * poly(Illusion_Strength, 3) + 
-        (1 | Participant),
-      family = "bernoulli"
-    )
-  }
 
@@
 
+  list(p = plot_model(data, model), model = model)
+}
+
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+std_params <- function(model, min=0, max=1) {
+  estimate_relation(
+  model,
+    at = list(Illusion_Strength = c(0), 
+              Illusion_Difference = seq(min, max, length.out=500)),
+    ) |> 
+    select(Illusion_Strength, Illusion_Difference, Error = Predicted) |> 
+    slice(c(which.min(abs(Error - 0.005)), 
+            which.min(abs(Error - 0.025)), 
+            which.min(abs(Error - 0.25)))) |> 
+    mutate(Error = insight::format_value(Error, as_percent=TRUE))
+}
+
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+
+std_params(model)
+std_params(gam$model)
+```
+
+## Ebbinghaus
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ebbinghaus")
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
+    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+
+std_params(model)
+std_params(gam$model)
+```
+
+## Rod and Frame
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Rod-Frame")
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
+    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+unique(data$Illusion_Strength)
+
+std_params(model, min=0.01, max=12)
+std_params(gam$model, min=0.01, max=12)
+```
+
+## Vertical-Horizontal
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Vertical-Horizontal")
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
+    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+unique(data$Illusion_Strength)
+
+std_params(model, min=0, max=0.40)
+std_params(gam$model, min=0, max=0.40)
+```
+
+## Zöllner
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Zöllner")
+
+plot_descriptive(data)
+
+data <- filter(data, abs(Illusion_Strength) < 45)
+
+plot_descriptive(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
 # model <- glmmTMB::glmmTMB(
-  #   Error ~ Illusion_Difference * Illusion_Strength + (1 | Participant),
-  #   data = data,
-  #   family = "binomial"
-  # )
-  # model <- mgcv::gamm(
-  #   # Error ~ s(Illusion_Difference, by = Illusion_Strength, k = 4),
-  #   Error ~ t2(Illusion_Difference, Illusion_Strength, k = 4),
-  #   random = list(Participant = ~1),
-  #   data = data,
-  #   family = "binomial"
-  # )
-  # model <- glmmTMB::glmmTMB(
-  #   # Error ~ s(Illusion_Difference, by = Illusion_Strength, k = 4),
-  #   Error ~ poly(Illusion_Difference, 2) * poly(Illusion_Strength, 2) * Illusion_Side + (1 | Participant),
-  #   data = data,
-  #   family = "binomial"
+#   Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
+#     ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+#   family = "binomial",
+#   data=data
 # )
 
-  n_lines <- 5
-  pred <- estimate_relation(model,
-                            at = c("Illusion_Strength", "Illusion_Difference"),
-                            # at = list(Illusion_Strength = unique(data$Illusion_Strength), Illusion_Difference = unique(data$Illusion_Difference)), 
-                            length = c(5, n_lines)) |>
-    mutate(
-      Illusion_Difference = as.factor(Illusion_Difference),
-      Illusion_Type = illusion
+formula <- brms::bf(
+  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
+    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
 )
 
-  # Set colors for lines
-  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)
-  diffvals <- as.numeric(as.character(unique(sort(pred$Illusion_Difference))))
-  names(colors) <- diffvals
-
-  # Assign color from the same palette to every observation of data (for geom_dots)
-  data$color <- colors[as.character(diffvals[max.col(-abs(outer(data$Illusion_Difference, diffvals, "-")))])]
-
-  # Manual jittering
-  xrange <- 0.05*diff(range(data$Illusion_Strength))
-  data$x <- data$Illusion_Strength 
-  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
-  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
-  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)
-
-  p <- pred |>
-    ggplot(aes(x = Illusion_Strength, y = Predicted)) +
-    geom_dots(
-      data = mutate(data, Illusion_Strength = round(Illusion_Strength, 1)),
-      aes(x=x, y = Error, group = Error, side = .dots_side), fill = data$color,
-      color = NA, alpha=0.5) +
-    # geom_dots(
-    #   data = mutate(data, Illusion_Strength = round(Illusion_Strength, 1)),
-    #   aes(y = Error, group = Error, side = .dots_side),
-    #   fill = ifelse(data$Error == 1, "red", "green"), color = NA, alpha=0.2,
-    #   position = position_jitter(width=0.05*diff(range(data$Illusion_Strength)), height=0)
-    # ) +
-    geom_slab(data=data, aes(y = Error)) +
-    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference, group = Illusion_Difference), alpha = 0.2) +
-    geom_vline(xintercept = 0, linetype = "dashed") +
-    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
-    geom_line(aes(color = Illusion_Difference, group = Illusion_Difference)) +
-    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_color_manual(values = colors) +
-    scale_fill_manual(values = colors) +
-    theme_modern() +
-    labs(
-      title = paste0(illusion, " Illusion"),
-      color = "Difficulty", fill = "Difficulty",
-      y = "Probability of Error",
-      x = "Illusion Strength"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-
-  list(p = p, model = model, pred = pred)
-}
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
 ```
 
-## References
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+
+std_params(model, min=0, max=5)
+```
+
+## White
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "White")
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
+    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+unique(data$Illusion_Strength)
+
+std_params(model, min=0, max=20)
+std_params(gam$model, min=0, max=20)
+```
+
+## Ponzo
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ponzo")
+
+plot_descriptive(data, side = "updown")
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
+    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+
+std_params(model, min=0, max=20)
+std_params(gam$model, min=0, max=20)
+```
+
+## Müller-Lyer
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Müller-Lyer")
+
+plot_descriptive(data, side = "updown")
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
+    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+
+std_params(model, min=0, max=0.6)
+std_params(gam$model, min=0, max=0.6)
+```
+
+## Poggendorff
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Poggendorff")
+
+plot_descriptive(data, side = "updown")
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Side * Illusion_Difference * poly(Illusion_Strength, 2) + 
+    ((1 + Illusion_Side + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
+
+parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+
+std_params(model, min=0, max=0.5)
+std_params(gam$model, min=0, max=0.5)
+```
+
+## Contrast
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Contrast")
+
+plot_descriptive(data, side = "updown")
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
+    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  algorithm = "meanfield",
+  refresh = 0
+)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+range(data$Illusion_Difference)
+range(data$Illusion_Strength)
+
+std_params(model, min=0, max=25)
+std_params(gam$model, min=0, max=25)
+```
+
+## Discussion
+
+This study allowed to elaborate a more precise prior idea regarding the magnitude of the effects at stake and the type of interaction between them.
+Crucially, it allowed us to refine the range of task difficulty and illusion strength for the stimuli of the next study to maximize the information gain.

### 2022-08-03 preprocessing.R

@@ -83,18 +83,30 @@ preprocess_raw <- function(file) {
 # This is a local folder containing raw data from unzipped pavlovia
 # It has been added to .gitignore to NOT be published on github
 # (it contains the subject ID of the participants)
-participants <- list.files("rawdata/study1/")
+participants <- list.files("study1/rawdata/")
 
 
 df <- data.frame()
 for (ppt in participants) {
-  df <- rbind(df, preprocess_raw(file = paste0("rawdata/study1/", ppt)))
+  df <- rbind(df, preprocess_raw(file = paste0("study1/rawdata/", ppt)))
 }
-
 
 # Study 1
 df$Study <- 1
+df$Pyllusion <- "1.1"
 df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"])
+df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"])
+df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"])
+df[df$Illusion_Type == "Zöllner", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Zöllner", "Illusion_Strength"])
+
+
+
+
+
+
+# Transformation
+df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
+df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
 
 # Save anonmized data
 write.csv(df, "data/study1.csv", row.names = FALSE)

### 2022-08-03 study1/make_stimuli.py

@@ -229,10 +229,6 @@ data = generate_images(
     name="Poggendorff",
 )
 
-# =============================================================================
-# STUDY 2
-# =============================================================================
-
 
 # -------------------------- Save data --------------------------
 # 1. Save data to a javascript file

### 2022-08-03 study2/experiment.js
duplicate folder for study 2

@@ -1,0 +1,311 @@
+/* ----------------- Internal Functions ----------------- */
+function get_results(illusion_mean, illusion_sd, illusion_type) {
+    if (typeof illusion_type != "undefined") {
+        var trials = jsPsych.data
+            .get()
+            .filter({ screen: "Trial", type: illusion_type }) // results by block
+    } else {
+        var trials = jsPsych.data.get().filter({ screen: "Trial" }) // overall results
+    }
+    var correct_trials = trials.filter({ correct: true })
+    var proportion_correct = correct_trials.count() / trials.count()
+    var rt_mean = trials.select("rt").mean()
+    if (correct_trials.count() > 0) {
+        var rt_mean_correct = correct_trials.select("rt").mean()
+        var ies = rt_mean_correct / proportion_correct // compute inverse efficiency score
+        var score_to_display = 100 - (ies/50)  // change max score to 100 instead of 5000
+        var percentile =
+            100 - cumulative_probability(ies, illusion_mean, illusion_sd) * 100
+    } else {
+        var rt_mean_correct = ""
+        var ies = ""
+        var percentile = 0
+        var score_to_display = 0
+    }
+    return {
+        accuracy: proportion_correct,
+        mean_reaction_time: rt_mean,
+        mean_reaction_time_correct: rt_mean_correct,
+        inverse_efficiency: ies,
+        percentage: percentile,
+        score: score_to_display,
+    }
+}
+
+function get_debrief_display(results, type = "Block") {
+    if (type === "Block") {
+        // Debrief at end of each block
+        var score =
+            "<p>Your score for this illusion is " +
+            '<p style="color: black; font-size: 48px; font-weight: bold;">' +
+            Math.round(results.score) +
+            "</p>"
+    } else if (type === "Final") {
+        // Final debriefing at end of game
+        var score =
+            "<p><strong>Your final score is</strong> " +
+            '<p style="color: black; font-size: 48px; font-weight: bold;">&#127881; ' +
+            Math.round(results.score) +
+            " &#127881;</p>"
+    }
+
+    return {
+        display_score: score,
+        display_accuracy:
+            "<p style='color:rgb(76,175,80);'>You responded correctly on <b>" +
+            round_digits(results.accuracy * 100) +
+            "" +
+            "%</b> of the trials.</p>",
+        display_rt:
+            "<p style='color:rgb(233,30,99);'>Your average response time was <b>" +
[…]

### 2022-08-03 study2/make_stimuli.py

@@ -1,0 +1,242 @@
+# -*- coding: utf-8 -*-
+import glob
+import json
+import os
+
+import neurokit2 as nk
+import numpy as np
+import pandas as pd
+import pyllusion as ill
+
+# Parameters
+width = 800
+height = 800
+n = 8
+data = []
+
+# Delete all existing stimuli
+for f in glob.glob("stimuli/*"):
+    os.remove(f)
+
+# Convenience functions
+def save_mosaic(strengths, differences, function, name="Delboeuf", **kwargs):
+    imgs = []
+    for strength in [abs(min(strengths, key=abs)), max(strengths)]:
+        for difference in [abs(min(differences, key=abs)), max(differences)]:
+
+            img = function(illusion_strength=strength, difference=difference, **kwargs).to_image(
+                width=width, height=height
+            )
+            img = ill.image_text(
+                "Difference: "
+                + str(np.round(difference, 2))
+                + ", Strength: "
+                + str(np.round(strength, 2)),
+                y=0.88,
+                size=40,
+                image=img,
+            )
+            imgs.append(img)
+    img = ill.image_mosaic(imgs, ncols=2)
+    img = ill.image_line(length=2, rotate=0, image=img)
+    img = ill.image_line(length=2, rotate=90, image=img)
+    img.save("utils/stimuli_examples/" + name + "_Mosaic.png")
+    return img
+
+
+def generate_images(data, strengths, differences, function, name="Delboeuf", **kwargs):
+
+    for strength in strengths:
+        for difference in differences:
+
+            img = function(illusion_strength=strength, difference=difference, **kwargs).to_image(
+                width=width, height=height
+            )
+            path = (
+                name
+                + "_str"
+                + str(np.round(strength, 2))
+                + "_diff"
+                + str(np.round(difference, 2))
+                + ".png"
+            )
+            img.save("stimuli/" + path)
+
+            # Compute expected response
+            if name in ["Delboeuf", "Ebbinghaus", "VerticalHorizontal", "White"]:
+                if difference > 0:
+                    correct = "arrowleft"
+                else:
+                    correct = "arrowright"
+            elif name in ["MullerLyer", "Contrast", "Poggendorff", "Ponzo"]:
[…]

### 2022-08-03 study2/stimuli/stimuli.js

@@ -1,0 +1,1 @@
+var stimuli = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": -0.8, "stimulus": "stimuli/Delboeuf_str-2.1_diff-0.8.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": -0.56, "stimulus": "stimuli/Delboeuf_str-2.1_diff-0.56.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": -0.32, "stimulus": "stimuli/Delboeuf_str-2.1_diff-0.32.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": -0.08, "stimulus": "stimuli/Delboeuf_str-2.1_diff-0.08.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": 0.08, "stimulus": "stimuli/Delboeuf_str-2.1_diff0.08.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": 0.32, "stimulus": "stimuli/Delboeuf_str-2.1_diff0.32.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": 0.56, "stimulus": "stimuli/Delboeuf_str-2.1_diff0.56.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": 0.8, "stimulus": "stimuli/Delboeuf_str-2.1_diff0.8.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -1.4,
[…]

### 2022-08-03 study1.Rmd
script

@@
 options(digits = 3,
-        mc.cores = 2)
+        mc.cores = 4,
+        brms.algorithm = "sampling",
+        brms.backend = "cmdstanr")
 
@@
 
-Behavioural findings regarding the Illusion Game.
+The goal of study 1 was to...
 
@@
 outliers <- c(
-  # Half of the trials are of very short RT
+  # Half of the trials have very short RT
   # Prolific Status: REJECTED
   "60684f29dbfe1bb2059e5e27_rkqoy",
-  # Error rate of 47.9%
-  # Prolific Status: RETURN REQUESTED
-  "61280140171ec546e87ed8cb_qdlgy",
-  # Error rate of 46.2%
-  # Prolific Status: RETURN REQUESTED
-  "614f36fd81c78b7a125c4262_6ax4g",
-  # Error rate of 42.1% and very large RT SD
-  # Prolific Status: RETURN REQUESTED
-  "5d398380b37ab1000111fac3_2nxxh",
   # Block n2 with very short RTs
@@
   # Error rate of 46.2% and short RTs
+  # Prolific Status: RETURNED
+  "615f319bb341cf2f306d858d_qsaq5",
+  # Error rate of 46.2%
+  # Prolific Status: RETURNED
+  "614f36fd81c78b7a125c4262_6ax4g",
+  # Error rate of 47.9%
   # Prolific Status: RETURN REQUESTED
-  "615f319bb341cf2f306d858d_qsaq5"
+  "61280140171ec546e87ed8cb_qdlgy",
+  # Error rate of 42.1% and very large RT SD
+  # Prolific Status: RETURN REQUESTED
+  "5d398380b37ab1000111fac3_2nxxh"
 )
@@
 
+## Data Analysis
+
+The analysis of study focused on the probability of errors as the main outcome variable.
+For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some insight on the underlying generative model.
+Next, we tested the performance of various models differing in their specifications, such as: with or without a log-transform on the task difficulty, with or without a 2nd order polynomial term of the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
+We then fitted the best performing model under a Bayesian model, and compared its predicted links with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships at the expense of model simplicity.
+Finally, we used these models to estimate different values of task difficulty that would lead to similar error rates (2.5, 5 and 25%) when the illusion strength is null.
+
 # Results {.tabset}
@@
       data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
-    } else {
+    } else if(x == "Right") {
       data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
@@
       data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
-    } else {
+    } else if(x == "Down") {
       data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
@@
   models <- list()
-  for(i in 1:2) {
+  for(i in 1:1) {
     for(j in 1:2) {
-      for(k1 in c("", "_log")) {  # c("", "_log")
-        for(k2 in c("", "_log")) {  # c("", "_log")
+      for(k1 in c("", "_log")) { 
+        for(k2 in c("")) {  
           for(side in c("", "-side")) {
@@
 formula <- brms::bf(
-  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
     (1 | Participant),
@@
   algorithm = "sampling",
-  refresh = 200
-)
+  refresh = 0
+)
+
+parameters::parameters(model)
 ```
@@
     Error ~ t2(Illusion_Difference, Illusion_Strength, bs = "cr", k=4) + 
-      ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+      (1 | Participant),
     family = "bernoulli"
@@
     data = data,
-    algorithm = "meanfield",
     refresh = 0
@@
 
-range(data$Illusion_Difference)
-range(data$Illusion_Strength)
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
 
@@
 formula <- brms::bf(
-  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
-    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
   family = "bernoulli"
@@
   data = data,
-  algorithm = "meanfield",
   refresh = 0
 )
+
+parameters::parameters(model)
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-range(data$Illusion_Difference)
-range(data$Illusion_Strength)
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
 
@@
 plot_descriptive(data)
+
+data <- filter(data, abs(Illusion_Strength) < 15)
+
+plot_descriptive(data)
 ```
@@
 formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
-    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  Error ~ log(Illusion_Difference) * Illusion_Strength + 
+    (1 | Participant),
   family = "bernoulli"
@@
   data = data,
-  algorithm = "meanfield",
   refresh = 0
 )
+
+parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0.01, max=12)
+std_params(gam$model, min=0.01, max=12)
+```
+
+## Vertical-Horizontal
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Vertical-Horizontal")
+
+plot_descriptive(data)
+
+data <- filter(data, abs(Illusion_Strength) < 90)
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+parameters::parameters(model)
 ```
@@
 
-std_params(model, min=0.01, max=12)
-std_params(gam$model, min=0.01, max=12)
-```
-
-## Vertical-Horizontal
+std_params(model, min=0, max=0.40)
+std_params(gam$model, min=0, max=0.40)
+```
+
+## Zöllner
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Vertical-Horizontal")
+data <- filter(df, Illusion_Type == "Zöllner")
+
+plot_descriptive(data)
+
+data <- filter(data, abs(Illusion_Strength) < 45)
 
@@
 formula <- brms::bf(
-  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
-    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  Error ~ log(Illusion_Difference) * Illusion_Strength + 
+    (1 | Participant),
   family = "bernoulli"
@@
   data = data,
-  algorithm = "meanfield",
   refresh = 0
 )
+
+parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0, max=5)
+```
+
+## White
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "White")
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+parameters::parameters(model)
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+# unique(data$Illusion_Strength)
+
+std_params(model, min=0, max=20)
+std_params(gam$model, min=0, max=20)
+```
+
+## Ponzo
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ponzo")
+
+plot_descriptive(data, side = "updown")
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
 range(data$Illusion_Difference)
 range(data$Illusion_Strength)
-unique(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.40)
-std_params(gam$model, min=0, max=0.40)
-```
-
-## Zöllner
+
+std_params(model, min=0, max=20)
+std_params(gam$model, min=0, max=20)
+```
+
+## Müller-Lyer
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Zöllner")
-
-plot_descriptive(data)
-
-data <- filter(data, abs(Illusion_Strength) < 45)
-
-plot_descriptive(data)
+data <- filter(df, Illusion_Type == "Müller-Lyer")
+
+plot_descriptive(data, side = "updown")
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-# model <- glmmTMB::glmmTMB(
-#   Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
-#     ((1 + Illusion_Difference + Illusion_Strength) | Participant),
-#   family = "binomial",
-#   data=data
-# )
-
 formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
-    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
   family = "bernoulli"
@@
 plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
 ```
@@
 
-std_params(model, min=0, max=5)
-```
-
-## White
+std_params(model, min=0, max=0.6)
+std_params(gam$model, min=0, max=0.6)
+```
+
+## Poggendorff
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "White")
-
-plot_descriptive(data)
+data <- filter(df, Illusion_Type == "Poggendorff")
+
+plot_descriptive(data, side = "updown")
+
+data <- filter(data, abs(Illusion_Strength) < 45)
+
+plot_descriptive(data, side = "updown")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-range(data$Illusion_Difference)
-range(data$Illusion_Strength)
-unique(data$Illusion_Strength)
-
-std_params(model, min=0, max=20)
-std_params(gam$model, min=0, max=20)
-```
-
-## Ponzo
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0, max=0.5)
+std_params(gam$model, min=0, max=0.5)
+```
+
+## Contrast
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Ponzo")
+data <- filter(df, Illusion_Type == "Contrast")
 
@@
   Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
-    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
+    (1 | Participant),
   family = "bernoulli"
@@
   data = data,
-  algorithm = "meanfield",
   refresh = 0
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-range(data$Illusion_Difference)
-range(data$Illusion_Strength)
-
-std_params(model, min=0, max=20)
-std_params(gam$model, min=0, max=20)
-```
-
-## Müller-Lyer
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Müller-Lyer")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
-    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  algorithm = "meanfield",
-  refresh = 0
-)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-range(data$Illusion_Difference)
-range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.6)
-std_params(gam$model, min=0, max=0.6)
-```
-
-## Poggendorff
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Poggendorff")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ Illusion_Side * Illusion_Difference * poly(Illusion_Strength, 2) + 
-    ((1 + Illusion_Side + Illusion_Difference + Illusion_Strength) | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  algorithm = "meanfield",
-  refresh = 0
-)
-
-parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-range(data$Illusion_Difference)
-range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.5)
-std_params(gam$model, min=0, max=0.5)
-```
-
-## Contrast
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Contrast")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
-    ((1 + Illusion_Difference + Illusion_Strength) | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  algorithm = "meanfield",
-  refresh = 0
-)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-range(data$Illusion_Difference)
-range(data$Illusion_Strength)
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
 

### 2022-08-03 preprocessing.R

@@ -97,7 +97,7 @@ df$Pyllusion <- "1.1"
 df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"])
 df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"])
 df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"])
-df[df$Illusion_Type == "Zöllner", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Zöllner", "Illusion_Strength"])
+df[df$Illusion_Type == "Zöllner", "Illusion_Strength"] <- -1 * round(df[df$Illusion_Type == "Zöllner", "Illusion_Strength"], 1)
 
 
 

### 2022-08-03 study1.Rmd
update figs

@@
   # Assign color from the same palette to every observation of data (for geom_dots)
-  data$color <- colors[as.character(diffvals[max.col(-abs(outer(data$Illusion_Difference, diffvals, "-")))])]
+  closest <- diffvals[max.col(-abs(outer(data$Illusion_Difference, diffvals, "-")))]
+  data$color <- colors[as.character(closest)]
+  data$color <- fct_reorder(data$color, closest)
 
@@
                     Illusion_Strength = round(Illusion_Strength, 1)),
-      aes(x=x, y = Error, group = Error, side = .dots_side), 
+      aes(x=x, y = Error, group = Error, side = .dots_side, order=color), 
       fill = data$color,
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-std_params <- function(model, min=0, max=1) {
+std_params <- function(model, min=0, max=2) {
   estimate_relation(
@@
 
-std_params(model)
-std_params(gam$model)
+std_params(model, min=0, max=2)
+std_params(gam$model, min=0, max=2)
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-range(data$Illusion_Difference)
-range(data$Illusion_Strength)
-unique(data$Illusion_Strength)
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
 

### 2022-08-03 study2/experiment.js
update instructions and add practice block

@@ -87,10 +87,10 @@ var make_break = {
     type: jsPsychHtmlButtonResponse,
     choices: ["I am ready to continue!"],
     stimulus:
-        "<p><b>BREAK</b></p>" +
-        "<p>You have reached the middle of the game.</p>" +
-        "<p>In the second part, you will see all the illusions once again. <b>Try to beat your previous score!</b></p>" +
-        "<p>We know it's long and challenging, so take a moment to rest. When you are ready, press the button below.</p>",
+        "<p><b>CONGRATULATIONS!</b></p>" +
+        "<p>You have finished one block of visual illusions.</p>" +
+        "<p>In the second part, you will see all the illusions once again. <b>Try to beat your previous score!</b></p>" ,
+        // "<p>We know it's long and challenging, so take a moment to rest. When you are ready, press the button below.</p>",
     save_trial_parameters: {
         trial_duration: true,
     },
@@ -308,4 +308,19 @@ function make_trial(instructions, illusion_name, type) {
     // Debriefing Information
     timeline.push(create_debrief((illusion_name = illusion_name)))
     return timeline
+}
+
+// Get random subset of elements from an array
+function getRandom(arr, n) {
+    var result = new Array(n),
+        len = arr.length,
+        taken = new Array(len);
+    if (n > len)
+        throw new RangeError("getRandom: more elements taken than available");
+    while (n--) {
+        var x = Math.floor(Math.random() * len);
+        result[n] = arr[x in taken ? taken[x] : x];
+        taken[x] = --len in taken ? taken[len] : len;
+    }
+    return result;
 }

### 2022-08-03 study1.Rmd
upload new stimuli

@@
 # Results {.tabset}
+
+<!-- ```{r message=FALSE, warning=FALSE, cache=cache} -->
+<!-- model <- glmmTMB::glmmTMB(Error ~ Illusion_Type * poly(ISI, 2) + (1|Participant), -->
+<!--                     data=mutate(df, Illusion_Type = as.factor(Illusion_Type)), -->
+<!--                     family="binomial") -->
+
+<!-- estimate_relation(model, at = c("ISI", "Illusion_Type")) |> -->
+<!--   ggplot(aes(x = ISI, y = Predicted, color="Illusion_Type")) + -->
+<!--   geom_line() -->
+
+<!-- model <- glmmTMB::glmmTMB(RT ~ Illusion_Type * poly(ISI, 2) + (1|Participant), -->
+<!--                     data=mutate(df, Illusion_Type = as.factor(Illusion_Type)) |>  -->
+<!--                       filter(Error == 0)) -->
+
+<!-- estimate_relation(model, at = c("ISI", "Illusion_Type")) |> -->
+<!--   ggplot(aes(x = ISI, y = Predicted, color=Illusion_Type)) + -->
+<!--   geom_line() -->
+<!-- ``` -->
 
@@
 
-## Discussion
+# Discussion
 

### 2022-08-03 preprocessing.R

@@ -47,6 +47,9 @@ preprocess_raw <- function(file) {
     RT = as.numeric(trials$rt)
   )
 
+  # Correct duration
+  df$Duration <- df$Duration - df$Break_Duration
+
   # Format names
   df$Illusion_Type <- ifelse(df$Illusion_Type == "MullerLyer", "Müller-Lyer", df$Illusion_Type)
   df$Illusion_Type <- ifelse(df$Illusion_Type == "Zollner", "Zöllner", df$Illusion_Type)

### 2022-08-03 study2/make_stimuli.py

@@ -11,8 +11,7 @@ import pyllusion as ill
 # Parameters
 width = 800
 height = 800
-n = 8
-data = []
+n = 16
 
 # Delete all existing stimuli
 for f in glob.glob("stimuli/*"):
@@ -94,149 +93,420 @@ def generate_images(data, strengths, differences, function, name="Delboeuf", **k
     return data
 
 
+def save_json(data, name="stimuli"):
+    file = "stimuli/" + name + ".js"
+    # 1. Save data to a javascript file
+    with open(file, "w") as fp:
+        json.dump(data, fp)
+
+    # 2. Re-read and add "var test_stimuli ="
+    with open(file) as f:
+        updatedfile = "var " + name + " = " + f.read()
+    with open(file, "w") as f:
+        f.write(updatedfile)
+
+
 def sqrtspace(mini=0.1, maxi=1, size=6):
     x = np.linspace(np.sqrt(0), np.sqrt(1), int(size / 2) + 1, endpoint=True) ** 2
     x = nk.rescale(x[1::], [mini, maxi])
     return np.concatenate((-1 * x[::-1], x))
 
 
-def doublelinspace(mini=0.1, maxi=1, size=6):
-    x = np.linspace(mini, maxi, int(size / 2), endpoint=True)
+def doublelinspace(mini=0.1, maxi=1, size=6, exp=False):
+    if exp is False:
+        x = np.linspace(mini, maxi, int(size / 2), endpoint=True)
+    else:
+        x = nk.expspace(mini, maxi, int(size / 2), out=float, base=2)
     return np.concatenate((-1 * x[::-1], x))
 
 
 # =============================================================================
-# STUDY 1
+# Make Stimuli
 # =============================================================================
-
+data_example = []
+data_block1 = []
+data_block2 = []
 
 # Left-right ======================================================================================
 # -------------------------- Delboeuf Illusion --------------------------
-ill.Delboeuf(illusion_strength=0.8, difference=1.2).to_image(width=800, height=600).save(
+ill.Delboeuf(illusion_strength=-0.4, difference=1.20).to_image(width=800, height=600).save(
     "utils/stimuli_demo/Delboeuf_Demo.png"
 )
-data = generate_images(
-    data,
[…]

### 2022-08-03 study2/stimuli/stimuli.js

@@ -1,1 +1,0 @@
-var stimuli = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": -0.8, "stimulus": "stimuli/Delboeuf_str-2.1_diff-0.8.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": -0.56, "stimulus": "stimuli/Delboeuf_str-2.1_diff-0.56.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": -0.32, "stimulus": "stimuli/Delboeuf_str-2.1_diff-0.32.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": -0.08, "stimulus": "stimuli/Delboeuf_str-2.1_diff-0.08.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": 0.08, "stimulus": "stimuli/Delboeuf_str-2.1_diff0.08.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": 0.32, "stimulus": "stimuli/Delboeuf_str-2.1_diff0.32.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": 0.56, "stimulus": "stimuli/Delboeuf_str-2.1_diff0.56.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -2.1, "Difference": 0.8, "stimulus": "stimuli/Delboeuf_str-2.1_diff0.8.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -1.4,
[…]

### 2022-08-03 study2/stimuli/stimuli_example.js

@@ -1,0 +1,1 @@
+var stimuli_example = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": -0.5, "Difference": -1, "stimulus": "stimuli/Delboeuf_str-0.5_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -0.5, "Difference": 1, "stimulus": "stimuli/Delboeuf_str-0.5_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.0_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.0_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.5, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.5_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.5, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.5_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -0.5, "Difference": -0.8, "stimulus": "stimuli/Ebbinghaus_str-0.5_diff-0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowright"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -0.5, "Difference": 0.8, "stimulus": "stimuli/Ebbinghaus_str-0.5_diff0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_S
[…]

### 2022-08-03 study2/stimuli/stimuli_part1.js

@@ -1,0 +1,1 @@
+var stimuli_part1 = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": -0.5, "Difference": -1, "stimulus": "stimuli/Delboeuf_str-0.5_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -0.5, "Difference": 1, "stimulus": "stimuli/Delboeuf_str-0.5_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.0_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.0_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.5, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.5_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.5, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.5_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -0.5, "Difference": -0.8, "stimulus": "stimuli/Ebbinghaus_str-0.5_diff-0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowright"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -0.5, "Difference": 0.8, "stimulus": "stimuli/Ebbinghaus_str-0.5_diff0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusio
[…]

### 2022-08-03 study2/stimuli/stimuli_part2.js

@@ -1,0 +1,1 @@
+var stimuli_part2 = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": -0.5, "Difference": -1, "stimulus": "stimuli/Delboeuf_str-0.5_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -0.5, "Difference": 1, "stimulus": "stimuli/Delboeuf_str-0.5_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.0_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.0_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.5, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.5_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.5, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.5_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -0.5, "Difference": -0.8, "stimulus": "stimuli/Ebbinghaus_str-0.5_diff-0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowright"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -0.5, "Difference": 0.8, "stimulus": "stimuli/Ebbinghaus_str-0.5_diff0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusio
[…]

### 2022-08-03 study2/make_stimuli.py
rename data_example -> data_training

@@ -123,7 +123,7 @@ def doublelinspace(mini=0.1, maxi=1, size=6, exp=False):
 # =============================================================================
 # Make Stimuli
 # =============================================================================
-data_example = []
+data_training = []
 data_block1 = []
 data_block2 = []
 
@@ -138,8 +138,8 @@ strengths = np.linspace(-0.7, 0.7, n - 1)
 diff1 = doublelinspace(mini=0.3, maxi=0.9, size=n // 2)
 diff2 = doublelinspace(mini=0.2, maxi=0.8, size=n // 2)
 
-data_example = generate_images(
-    data_example,
+data_training = generate_images(
+    data_training,
     # strengths=strengths[0 : : (n // 2) - 1],
     strengths=strengths[2 : -2 : (n // 2) - 3],
     differences=[-1, 1],
@@ -179,8 +179,8 @@ diff1 = doublelinspace(mini=0.3, maxi=0.9, size=n // 2)
 diff2 = doublelinspace(mini=0.2, maxi=0.8, size=n // 2)
 
 
-data_example = generate_images(
-    data_example,
+data_training = generate_images(
+    data_training,
     # strengths=strengths[0 : : (n // 2) - 1],
     strengths=strengths[2 : -2 : (n // 2) - 3],
     differences=[-0.8, 0.8],
@@ -219,8 +219,8 @@ diff1 = doublelinspace(mini=1.1, maxi=7.1, size=n // 2)
 diff2 = doublelinspace(mini=0.1, maxi=6.1, size=n // 2)
 
 
-data_example = generate_images(
-    data_example,
+data_training = generate_images(
+    data_training,
     # strengths=strengths[0 : : (n // 2) - 1],
     strengths=strengths[2 : -2 : (n // 2) - 3],
     differences=[-12, 12],
@@ -255,8 +255,8 @@ diff1 = doublelinspace(mini=0.06, maxi=0.24, size=n // 2)
 diff2 = doublelinspace(mini=0.03, maxi=0.21, size=n // 2)
 
 
-data_example = generate_images(
-    data_example,
+data_training = generate_images(
+    data_training,
     # strengths=strengths[0 : : (n // 2) - 1],
     strengths=strengths[2 : -2 : (n // 2) - 3],
     differences=[-0.4, 0.4],
@@ -291,8 +291,8 @@ diff1 = doublelinspace(mini=0.7, maxi=4.3, size=n // 2)
[…]

### 2022-08-04 study1.Rmd
study0

--- 2022-08-03 study1.Rmd
@@
 # Results {.tabset}
+
+```{r eval=FALSE, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE}
+# Exploratory 
+# ============
+
+# ISI
+model <- glmmTMB::glmmTMB(Error ~ Illusion_Type * poly(ISI, 2) + (1|Participant),
+                    data=mutate(df, Illusion_Type = as.factor(Illusion_Type)),
+                    family="binomial")
+
+estimate_relation(model, at = c("ISI", "Illusion_Type")) |>
+  ggplot(aes(x = ISI, y = Predicted, color="Illusion_Type")) +
+  geom_line()
+
+model <- glmmTMB::glmmTMB(RT ~ Illusion_Type * poly(ISI, 2) + (1|Participant),
+                    data=mutate(df, Illusion_Type = as.factor(Illusion_Type)) |>
+                      filter(Error == 0))
+
+estimate_relation(model, at = c("ISI", "Illusion_Type")) |>
+  ggplot(aes(x = ISI, y = Predicted, color=Illusion_Type)) +
+  geom_line()
+
+# Difficulty - Link type
+ps <- list()
+for(illusion in c("Delboeuf", "Ebbinghaus", "Rod-Frame", "Poggendorff", 
+"Vertical-Horizontal", "Zöllner", "Müller-Lyer", "Ponzo", 
+"Contrast", "White")) {
+  data <- filter(df, Illusion_Type == illusion, Illusion_Strength == 0)
+  model <- mgcv::gamm(Error ~ s(Illusion_Difference, k=4),
+                      random = list(Participant = ~1),
+                      data=data,
+                      family="binomial")
+
+  ps[[illusion]] <- estimate_relation(model, 
+                                     at = list(Illusion_Difference = unique(data$Illusion_Difference))) |>
+    ggplot(aes(x = Illusion_Difference, y = Predicted)) +
+    geom_bar(stat="identity", alpha=0.3) +
+    geom_line() +
+    ggtitle(illusion)
+}
+
+see::plots(ps)
+
+```
 

### 2022-08-04 study0/make_stimuli.py

@@ -1,0 +1,227 @@
+# -*- coding: utf-8 -*-
+import glob
+import json
+import os
+
+import neurokit2 as nk
+import numpy as np
+import pandas as pd
+import pyllusion as ill
+
+# Parameters
+width = 800
+height = 800
+n = 32
+
+# Delete all existing stimuli
+for f in glob.glob("stimuli/*"):
+    os.remove(f)
+
+
+def generate_images(data, strengths, differences, function, name="Delboeuf", **kwargs):
+
+    for strength in strengths:
+        for difference in differences:
+
+            img = function(illusion_strength=strength, difference=difference, **kwargs).to_image(
+                width=width, height=height, target_only=True
+            )
+            path = (
+                name
+                + "_str"
+                + str(np.round(strength, 5))
+                + "_diff"
+                + str(np.round(difference, 5))
+                + ".png"
+            )
+            img.save("stimuli/" + path)
+
+            # Compute expected response
+            if name in ["Delboeuf", "Ebbinghaus", "VerticalHorizontal", "White"]:
+                if difference > 0:
+                    correct = "arrowleft"
+                else:
+                    correct = "arrowright"
+            elif name in ["MullerLyer", "Contrast", "Poggendorff", "Ponzo"]:
+                if difference > 0:
+                    correct = "arrowup"
+                else:
+                    correct = "arrowdown"
+            elif name in ["Zollner", "RodFrame"]:
+                if difference < 0:
+                    correct = "arrowleft"
+                else:
+                    correct = "arrowright"
+
+            # Save parameters for Delboeuf Illusion
+            data.append(
+                {
+                    "Illusion_Type": name,
+                    "Illusion_Strength": strength,
+                    "Difference": difference,
+                    "stimulus": "stimuli/" + path,
+                    "data": {"screen": "Trial", "block": name, "correct_response": correct},
+                }
+            )
+
+    return data
+
+
+def save_json(data, name="stimuli"):
+    file = "stimuli/" + name + ".js"
+    # 1. Save data to a javascript file
+    with open(file, "w") as fp:
+        json.dump(data, fp)
+
+    # 2. Re-read and add "var test_stimuli ="
+    with open(file) as f:
[…]

### 2022-08-04 study0/stimuli/stimuli.js

@@ -1,0 +1,1 @@
+var stimuli = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -1.0, "stimulus": "stimuli/Delboeuf_str0_diff-1.0.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.9334000000000001, "stimulus": "stimuli/Delboeuf_str0_diff-0.9334.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.8668000000000001, "stimulus": "stimuli/Delboeuf_str0_diff-0.8668.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.8002000000000001, "stimulus": "stimuli/Delboeuf_str0_diff-0.8002.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.7336, "stimulus": "stimuli/Delboeuf_str0_diff-0.7336.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.667, "stimulus": "stimuli/Delboeuf_str0_diff-0.667.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.6004, "stimulus": "stimuli/Delboeuf_str0_diff-0.6004.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.5338, "stimulus": "stimuli/Delboeuf_str0_diff-0.5338.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.46720000000000006, "
[…]

### 2022-08-04 study0/experiment.js
update study 0

@@ -1,0 +1,311 @@
+/* ----------------- Internal Functions ----------------- */
+function get_results(illusion_mean, illusion_sd, illusion_type) {
+    if (typeof illusion_type != "undefined") {
+        var trials = jsPsych.data
+            .get()
+            .filter({ screen: "Trial", type: illusion_type }) // results by block
+    } else {
+        var trials = jsPsych.data.get().filter({ screen: "Trial" }) // overall results
+    }
+    var correct_trials = trials.filter({ correct: true })
+    var proportion_correct = correct_trials.count() / trials.count()
+    var rt_mean = trials.select("rt").mean()
+    if (correct_trials.count() > 0) {
+        var rt_mean_correct = correct_trials.select("rt").mean()
+        var ies = rt_mean_correct / proportion_correct // compute inverse efficiency score
+        var score_to_display = 100 - (ies/50)
+        var percentile =
+            100 - cumulative_probability(ies, illusion_mean, illusion_sd) * 100
+    } else {
+        var rt_mean_correct = ""
+        var ies = ""
+        var percentile = 0
+        var score_to_display = 0
+    }
+    return {
+        accuracy: proportion_correct,
+        mean_reaction_time: rt_mean,
+        mean_reaction_time_correct: rt_mean_correct,
+        inverse_efficiency: ies,
+        percentage: percentile,
+        score: score_to_display,
+    }
+}
+
+function get_debrief_display(results, type = "Block") {
+    if (type === "Block") {
+        // Debrief at end of each block
+        var score =
+            "<p>Your score for this illusion is " +
+            '<p style="color: black; font-size: 48px; font-weight: bold;">' +
+            Math.round(results.score) +
+            "</p>"
+    } else if (type === "Final") {
+        // Final debriefing at end of game
+        var score =
+            "<p><strong>Your final score is</strong> " +
+            '<p style="color: black; font-size: 48px; font-weight: bold;">&#127881; ' +
+            Math.round(results.score) +
+            " &#127881;</p>"
+    }
+
+    return {
+        display_score: score,
+        display_accuracy:
+            "<p style='color:rgb(76,175,80);'>You responded correctly on <b>" +
+            round_digits(results.accuracy * 100) +
+            "" +
+            "%</b> of the trials.</p>",
+        display_rt:
+            "<p style='color:rgb(233,30,99);'>Your average response time was <b>" +
+            round_digits(results.mean_reaction_time) +
[…]

### 2022-08-04 study0/experiment.js
update instructions

@@ -252,7 +252,7 @@ function create_debrief(illusion_name = "Ponzo") {
                 // "<hr>" +
                 // //
                 // show_screen.display_comparison +
-                "<hr><p>Can you do better in the next illusion?</p>"
+                "<hr><p>Can you do better in the next block?</p>"
             )
         },
         data: { screen: "block_results" },

### 2022-08-04 study1.Rmd
push new stimuli

@@
 ```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+# plot(seq(0, 1, length.out=100), seq(0, 1, length.out=100)**(1/3), "l", col="green")
+# lines(seq(0, 1, length.out=100), sqrt(seq(0, 1, length.out=100)), "l", col="red")
+# lines(seq(0, 1, length.out=100), log(1+seq(0, 1, length.out=100)), "l", col="blue")
+# lines(seq(0, 1, length.out=100), seq(0, 1, length.out=100), "l", col="black")
+
 best_models <- function(data) {
@@
   for(i in 1:1) {
-    for(j in 1:2) {
-      for(k1 in c("", "_log")) { 
-        for(k2 in c("")) {  
+    for(j in 1:1) {
+      for(k1 in c("", "_log", "_sqrt", "_cbrt")) { 
+        for(k2 in c("", "_log", "_sqrt", "_cbrt")) {  
           for(side in c("", "-side")) {
@@
 formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+  Error ~ Illusion_Difference_cbrt * Illusion_Strength + 
     (1 | Participant),
@@
   data = data,
-  algorithm = "sampling",
   refresh = 0
@@
 
-  n_lines <- 5
+  vars <- insight::find_predictors(model)$conditional
+  n_lines <- length(unique(data[[vars[1]]]))
   pred <- estimate_relation(model,
-                            at = c("Illusion_Strength", "Illusion_Difference"),
-                            length = c(25, n_lines)) |>
-    mutate(Illusion_Difference = as.factor(round(Illusion_Difference, 2)))
+                            at = vars,
+                            length = c(NA, 25))
+  pred$Illusion_Difference <- as.factor(round(pred[[vars[1]]], 2))
 
@@
   # Assign color from the same palette to every observation of data (for geom_dots)
-  closest <- diffvals[max.col(-abs(outer(data$Illusion_Difference, diffvals, "-")))]
+  closest <- diffvals[max.col(-abs(outer(data[[vars[1]]], diffvals, "-")))]
   data$color <- colors[as.character(closest)]

### 2022-08-04 preprocessing.R

@@ -97,8 +97,8 @@ for (ppt in participants) {
 # Study 1
 df$Study <- 1
 df$Pyllusion <- "1.1"
-df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"])
-df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"])
+# df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"])
+# df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"])
 df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"])
 df[df$Illusion_Type == "Zöllner", "Illusion_Strength"] <- -1 * round(df[df$Illusion_Type == "Zöllner", "Illusion_Strength"], 1)
 
@@ -109,7 +109,11 @@ df[df$Illusion_Type == "Zöllner", "Illusion_Strength"] <- -1 * round(df[df$Illu
 
 # Transformation
 df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
+df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
+df$Illusion_Difference_cbrt <- df$Illusion_Difference**(1/3)
 df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
+df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
+df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))
 
 # Save anonmized data
 write.csv(df, "data/study1.csv", row.names = FALSE)

### 2022-08-04 study0/make_stimuli.py

@@ -11,7 +11,7 @@ import pyllusion as ill
 # Parameters
 width = 800
 height = 800
-n = 32
+n = 64
 
 # Delete all existing stimuli
 for f in glob.glob("stimuli/*"):
@@ -29,9 +29,9 @@ def generate_images(data, strengths, differences, function, name="Delboeuf", **k
             path = (
                 name
                 + "_str"
-                + str(np.round(strength, 5))
+                + str(np.round(strength, 6))
                 + "_diff"
-                + str(np.round(difference, 5))
+                + str(np.round(difference, 6))
                 + ".png"
             )
             img.save("stimuli/" + path)

### 2022-08-04 study0/stimuli/stimuli.js

@@ -1,1 +1,1 @@
-var stimuli = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -1.0, "stimulus": "stimuli/Delboeuf_str0_diff-1.0.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.9334000000000001, "stimulus": "stimuli/Delboeuf_str0_diff-0.9334.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.8668000000000001, "stimulus": "stimuli/Delboeuf_str0_diff-0.8668.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.8002000000000001, "stimulus": "stimuli/Delboeuf_str0_diff-0.8002.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.7336, "stimulus": "stimuli/Delboeuf_str0_diff-0.7336.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.667, "stimulus": "stimuli/Delboeuf_str0_diff-0.667.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.6004, "stimulus": "stimuli/Delboeuf_str0_diff-0.6004.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.5338, "stimulus": "stimuli/Delboeuf_str0_diff-0.5338.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.46720000000000006, "
[…]

### 2022-08-04 study0/make_stimuli.py
decrease number of stimuli

@@ -11,7 +11,7 @@ import pyllusion as ill
 # Parameters
 width = 800
 height = 800
-n = 64
+n = 4
 
 # Delete all existing stimuli
 for f in glob.glob("stimuli/*"):

### 2022-08-04 study0/stimuli/stimuli.js

@@ -1,1 +1,1 @@
-var stimuli = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -1.0, "stimulus": "stimuli/Delboeuf_str0_diff-1.0.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.9677741935483871, "stimulus": "stimuli/Delboeuf_str0_diff-0.967774.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.9355483870967741, "stimulus": "stimuli/Delboeuf_str0_diff-0.935548.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.9033225806451612, "stimulus": "stimuli/Delboeuf_str0_diff-0.903323.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.8710967741935484, "stimulus": "stimuli/Delboeuf_str0_diff-0.871097.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.8388709677419355, "stimulus": "stimuli/Delboeuf_str0_diff-0.838871.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.8066451612903226, "stimulus": "stimuli/Delboeuf_str0_diff-0.806645.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.7744193548387096, "stimulus": "stimuli/Delboeuf_str0_diff-0.774419.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion
[…]

### 2022-08-04 study0/preprocessing_raw.R
preprocessing for study0

@@ -1,0 +1,66 @@
+#Preprocess for Study 0
+preprocess_raw <- function(file){
+  
+  # Install required packages
+  install.packages("pacman")
+  
+  pacman::p_load(jsonlite, tidyverse)
+  
+
+  data <- fromJSON(file, flatten=TRUE)
+  
+  if (!"final_results" %in% data$screen) {
+    print(paste0("Warning: Incomplete data for ", file))
+    return(data.frame())
+  }
+  
+  data <- data|> drop_na(screen)
+  trials <- data[data$screen == 'Trial', ]
+  
+  df <- data.frame(
+    Participant = trials$participant_id,
+    Duration = as.numeric(data[data$screen == "final_results", "time_elapsed"]) / 1000 / 60,
+    Illusion_Type = trials$type,
+    Block_Order = as.numeric(trials$block_number),
+    Trial = as.numeric(trials$trial_number),
+    Stimulus = gsub(".png", "", gsub("stimuli/", "", trials$stimulus)),
+    Illusion_Strength = as.numeric(trials$illusion_strength),
+    Illusion_Side = as.factor(sign(as.numeric(trials$illusion_difference))),
+    Illusion_Difference = abs(as.numeric(trials$illusion_difference)),
+    Answer = trials$response,
+    Error = as.integer(!as.logical(trials$correct)),
+    ISI = as.numeric(data[data$screen == "fixation" & !is.na(data$screen), "trial_duration"]),
+    RT = as.numeric(trials$rt)
+  )
+  
+  # Format names
+  df$Illusion_Type <- ifelse(df$Illusion_Type == "MullerLyer", "Müller-Lyer", df$Illusion_Type)
+  df$Illusion_Type <- ifelse(df$Illusion_Type == "Zollner", "Zöllner", df$Illusion_Type)
+  df$Illusion_Type <- ifelse(df$Illusion_Type == "RodFrame", "Rod-Frame", df$Illusion_Type)
+  df$Illusion_Type <- ifelse(df$Illusion_Type == "VerticalHorizontal", "Vertical-Horizontal", df$Illusion_Type)
+  
+  df
+}
+  
+  # Run ---------------------------------------------------------------------
+  participants <- list.files("study0/data/")
+  
+  
+  df <- data.frame()
+  for (ppt in participants) {
+    df <- rbind(df, preprocess_raw(file = paste0("study0/data/", ppt)))
[…]

### 2022-08-04 study1.Rmd
re-make stimuli

@@
         mc.cores = 4,
-        brms.algorithm = "sampling",
+        brms.algorithm = "meanfield",
         brms.backend = "cmdstanr")
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+```{r message=FALSE, warning=FALSE, cache=cache}
 # plot(seq(0, 1, length.out=100), seq(0, 1, length.out=100)**(1/3), "l", col="green")
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
+cbrt <- function(x) sign(x) * abs(x)^(1/3)
+
 formula <- brms::bf(
-  Error ~ Illusion_Difference_cbrt * Illusion_Strength + 
+  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
     (1 | Participant),
@@
 
+  # Get variables
   vars <- insight::find_predictors(model)$conditional
-  n_lines <- length(unique(data[[vars[1]]]))
+  vardiff <- vars[1]
+  varstrength <- vars[2]
+
+  n_lines <- length(unique(data[[vardiff]]))
   pred <- estimate_relation(model,
@@
                             length = c(NA, 25))
-  pred$Illusion_Difference <- as.factor(round(pred[[vars[1]]], 2))
+  pred$Illusion_Difference <- as.factor(round(pred[[vardiff]], 2))
 
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+```{r message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 formula <- brms::bf(
-  Error ~ Illusion_Difference_cbrt * Illusion_Strength + 
+  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
     (1 | Participant),
@@
 formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * Illusion_Strength + 
+  Error ~ poly(Illusion_Difference, 2) * poly(Illusion_Strength, 2) + 
     (1 | Participant),
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * Illusion_Strength + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0, max=0.40)
+std_params(gam$model, min=0, max=0.40)
+```
+
+## Zöllner
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Zöllner")
+
+plot_descriptive(data)
+
+data <- filter(data, abs(Illusion_Strength) < 45)
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0, max=5)
+```
+
+## White
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "White")
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 # range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.40)
-std_params(gam$model, min=0, max=0.40)
-```
-
-## Zöllner
+# unique(data$Illusion_Strength)
+
+std_params(model, min=0, max=20)
+std_params(gam$model, min=0, max=20)
+```
+
+## Ponzo
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Zöllner")
-
-plot_descriptive(data)
-
-data <- filter(data, abs(Illusion_Strength) < 45)
-
-plot_descriptive(data)
+data <- filter(df, Illusion_Type == "Ponzo")
+
+plot_descriptive(data, side = "updown")
 ```
@@
 
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
 ### Parameter Standardization
@@
 
-std_params(model, min=0, max=5)
-```
-
-## White
+std_params(model, min=0, max=20)
+std_params(gam$model, min=0, max=20)
+```
+
+## Müller-Lyer
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "White")
-
-plot_descriptive(data)
+data <- filter(df, Illusion_Type == "Müller-Lyer")
+
+plot_descriptive(data, side = "updown")
 ```
@@
 # range(data$Illusion_Strength)
-# unique(data$Illusion_Strength)
-
-std_params(model, min=0, max=20)
-std_params(gam$model, min=0, max=20)
-```
-
-## Ponzo
+
+std_params(model, min=0, max=0.6)
+std_params(gam$model, min=0, max=0.6)
+```
+
+## Poggendorff
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Ponzo")
+data <- filter(df, Illusion_Type == "Poggendorff")
 
 plot_descriptive(data, side = "updown")
+
+data <- filter(data, abs(Illusion_Strength) < 45)
+
+plot_descriptive(data, side = "updown")
 ```
@@
 
-std_params(model, min=0, max=20)
-std_params(gam$model, min=0, max=20)
-```
-
-## Müller-Lyer
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Müller-Lyer")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.6)
-std_params(gam$model, min=0, max=0.6)
-```
-
-## Poggendorff
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Poggendorff")
-
-plot_descriptive(data, side = "updown")
-
-data <- filter(data, abs(Illusion_Strength) < 45)
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache,results='hide'}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
 std_params(model, min=0, max=0.5)

### 2022-08-04 study0/make_stimuli.py

@@ -11,7 +11,7 @@ import pyllusion as ill
 # Parameters
 width = 800
 height = 800
-n = 4
+n = 64
 
 # Delete all existing stimuli
 for f in glob.glob("stimuli/*"):
@@ -156,7 +156,7 @@ ill.Zollner(illusion_strength=-40, difference=8).to_image(
 data = generate_images(
     data,
     strengths=[0],
-    differences=doublelinspace(mini=0.001, maxi=4.3, size=n),
+    differences=doublelinspace(mini=0.001, maxi=5, size=n),
     function=ill.Zollner,
     name="Zollner",
 )

### 2022-08-04 study0/preprocessing.R

@@ -1,0 +1,60 @@
+# Preprocess for Study 0
+preprocess <- function(file){
+
+  data <- jsonlite::fromJSON(file, flatten=TRUE)
+
+  if (!"final_results" %in% data$screen) {
+    print(paste0("Warning: Incomplete data for ", file))
+    return(data.frame())
+  }
+
+  data <- data[~is.na(data$screen)]
+  trials <- data[data$screen == 'Trial', ]
+
+  df <- data.frame(
+    Participant = trials$participant_id,
+    Duration = as.numeric(data[data$screen == "final_results", "time_elapsed"]) / 1000 / 60,
+    Illusion_Type = trials$type,
+    Block_Order = as.numeric(trials$block_number),
+    Trial = as.numeric(trials$trial_number),
+    Stimulus = gsub(".png", "", gsub("stimuli/", "", trials$stimulus)),
+    Illusion_Strength = as.numeric(trials$illusion_strength),
+    Illusion_Side = as.factor(sign(as.numeric(trials$illusion_difference))),
+    Illusion_Difference = abs(as.numeric(trials$illusion_difference)),
+    Answer = trials$response,
+    Error = as.integer(!as.logical(trials$correct)),
+    ISI = as.numeric(data[data$screen == "fixation" & !is.na(data$screen), "trial_duration"]),
+    RT = as.numeric(trials$rt)
+  )
+
+  # Format names
+  df$Illusion_Type <- ifelse(df$Illusion_Type == "MullerLyer", "Müller-Lyer", df$Illusion_Type)
+  df$Illusion_Type <- ifelse(df$Illusion_Type == "Zollner", "Zöllner", df$Illusion_Type)
+  df$Illusion_Type <- ifelse(df$Illusion_Type == "RodFrame", "Rod-Frame", df$Illusion_Type)
+  df$Illusion_Type <- ifelse(df$Illusion_Type == "VerticalHorizontal", "Vertical-Horizontal", df$Illusion_Type)
+
+  df
+}
+
+  # Run ---------------------------------------------------------------------
+  participants <- list.files("study0/data/")
+
+
+  df <- data.frame()
+  for (ppt in participants) {
+    df <- rbind(df, preprocess_raw(file = paste0("study0/data/", ppt)))
+  }
+
+  # Study 0
+  df$Study <- 0
+  df$Pyllusion <- "1.2"
+
+
+  # Transformation
[…]

### 2022-08-04 study0/preprocessing_raw.R

@@ -1,66 +1,0 @@
-#Preprocess for Study 0
-preprocess_raw <- function(file){
-  
-  # Install required packages
-  install.packages("pacman")
-  
-  pacman::p_load(jsonlite, tidyverse)
-  
-
-  data <- fromJSON(file, flatten=TRUE)
-  
-  if (!"final_results" %in% data$screen) {
-    print(paste0("Warning: Incomplete data for ", file))
-    return(data.frame())
-  }
-  
-  data <- data|> drop_na(screen)
-  trials <- data[data$screen == 'Trial', ]
-  
-  df <- data.frame(
-    Participant = trials$participant_id,
-    Duration = as.numeric(data[data$screen == "final_results", "time_elapsed"]) / 1000 / 60,
-    Illusion_Type = trials$type,
-    Block_Order = as.numeric(trials$block_number),
-    Trial = as.numeric(trials$trial_number),
-    Stimulus = gsub(".png", "", gsub("stimuli/", "", trials$stimulus)),
-    Illusion_Strength = as.numeric(trials$illusion_strength),
-    Illusion_Side = as.factor(sign(as.numeric(trials$illusion_difference))),
-    Illusion_Difference = abs(as.numeric(trials$illusion_difference)),
-    Answer = trials$response,
-    Error = as.integer(!as.logical(trials$correct)),
-    ISI = as.numeric(data[data$screen == "fixation" & !is.na(data$screen), "trial_duration"]),
-    RT = as.numeric(trials$rt)
-  )
-  
-  # Format names
-  df$Illusion_Type <- ifelse(df$Illusion_Type == "MullerLyer", "Müller-Lyer", df$Illusion_Type)
-  df$Illusion_Type <- ifelse(df$Illusion_Type == "Zollner", "Zöllner", df$Illusion_Type)
-  df$Illusion_Type <- ifelse(df$Illusion_Type == "RodFrame", "Rod-Frame", df$Illusion_Type)
-  df$Illusion_Type <- ifelse(df$Illusion_Type == "VerticalHorizontal", "Vertical-Horizontal", df$Illusion_Type)
-  
-  df
-}
-  
-  # Run ---------------------------------------------------------------------
-  participants <- list.files("study0/data/")
-  
-  
-  df <- data.frame()
-  for (ppt in participants) {
-    df <- rbind(df, preprocess_raw(file = paste0("study0/data/", ppt)))
[…]

### 2022-08-04 study0/stimuli/stimuli.js

@@ -1,1 +1,1 @@
-var stimuli = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -1.0, "stimulus": "stimuli/Delboeuf_str0_diff-1.0.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": -0.001, "stimulus": "stimuli/Delboeuf_str0_diff-0.001.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": 0.001, "stimulus": "stimuli/Delboeuf_str0_diff0.001.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0, "Difference": 1.0, "stimulus": "stimuli/Delboeuf_str0_diff1.0.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "RodFrame", "Illusion_Strength": 0, "Difference": -7.0, "stimulus": "stimuli/RodFrame_str0_diff-7.0.png", "data": {"screen": "Trial", "block": "RodFrame", "correct_response": "arrowleft"}}, {"Illusion_Type": "RodFrame", "Illusion_Strength": 0, "Difference": -0.001, "stimulus": "stimuli/RodFrame_str0_diff-0.001.png", "data": {"screen": "Trial", "block": "RodFrame", "correct_response": "arrowleft"}}, {"Illusion_Type": "RodFrame", "Illusion_Strength": 0, "Difference": 0.001, "stimulus": "stimuli/RodFrame_str0_diff0.001.png", "data": {"screen": "Trial", "block": "RodFrame", "correct_response": "arrowright"}}, {"Illusion_Type": "RodFrame", "Illusion_Strength": 0, "Difference": 7.0, "stimulus": "stimuli/RodFrame_str0_diff7.0.png", "data": {"screen": "Trial", "block": "RodFrame", "correct_response": "arrowright"}}, {"Illusion_Type": "VerticalHorizontal", "Illusion_Strength": 0, "Difference": -0.25, "stimulus": "stimuli/VerticalHorizontal_str0_diff-
[…]

### 2022-08-04 study0/analysis.Rmd
update stim

---
title: "Illusion Game Validation (Study 1)"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3,
        mc.cores = 4,
        brms.algorithm = "meanfield",
        brms.backend = "cmdstanr")

cache <- FALSE
fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

source("preprocessing.R")

cbrt <- function(x) x^(1/3)

df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
df$Illusion_Difference_cbrt <- cbrt(df$Illusion_Difference)

df <- filter(df, RT < 1000)
```

## Results {.tabset}

### Delboeuf

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Delboeuf")
```

```{r message=FALSE, warning=FALSE}
best_models <- function(data) {
  models_err <- list()
  models_rt <- list()
  for(k1 in c("", "_log", "_sqrt", "_cbrt")) { 
    for(side in c("", "-side")) {
      name <- paste0("dif", k1, side)
      print(name)
      f <- paste0("Illusion_Difference", 
                  k1,
                  " + (1|Participant)"
                  )

      if(side == "-side") f <- paste0("Illusion_Side * ", f)

      m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)), 
                            data = data, family = "binomial")
      if(performance::check_convergence(m)) {
        models_err[[name]] <- m
      }
      m <- glmmTMB::glmmTMB(as.formula(paste0("RT ~ ", f)), 
                            data = data)
      if(performance::check_convergence(m)) {
        models_rt[[name]] <- m
      }
    }
  }

  to_keep <- compare_performance(models_err, metrics = c("BIC")) |> 
    arrange(BIC) |> 
    slice(1:5) |> 
    pull(Name)

  test <- test_performance(models_err[to_keep], reference=1)
  perf <- compare_performance(models_err[to_keep], metrics = c("BIC", "R2"))

  rez <- merge(perf, test) |> 
    arrange(BIC) |> 
    select(Name, BIC, R2_marginal, BF) |> 
    mutate(BF = insight::format_bf(BF, name=""),
           Model = "Error")

  to_keep <- compare_performance(models_rt, metrics = c("BIC")) |> 
    arrange(BIC) |> 
    slice(1:5) |> 
    pull(Name)

  test <- test_performance(models_rt[to_keep], reference=1)
  perf <- compare_performance(models_rt[to_keep], metrics = c("BIC", "R2"))

  merge(perf, test) |> 
    arrange(BIC) |> 
    select(Name, BIC, R2_marginal, BF) |> 
    mutate(BF = insight::format_bf(BF, name=""),
           Model = "RT") |> 
    rbind(rez)
}

best_models(data)
```

```{r message=FALSE, warning=FALSE}
visualize <- function(data, model_err, model_rt) {
  model <- mgcv::gam(Error ~ s(Illusion_Difference, by = Illusion_Side),
                     data = data,
                     family = "binomial")

  p1 <- data |> 
    group_by(Illusion_Difference, Illusion_Side) |> 
    summarize(Error = mean(Error)) |> 
    ggplot(aes(x = Illusion_Difference, y = Error)) +
    geom_bar(stat="identity") +
    geom_line(data=estimate_relation(model), aes(y=Predicted), color="blue") +
    geom_line(data=estimate_relation(model_err), aes(y=Predicted), color="red") +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_modern() +
    labs(
      color = "Illusion Strength", 
      fill = "Illusion Strength",
      y = "Probability of Error",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
    facet_wrap(~Illusion_Side, ncol=2, labeller = "label_both")

  model <- mgcv::gam(RT ~ s(Illusion_Difference, by = Illusion_Side),
                     data = filter(data, Error == 0, RT < 1000))

  p2 <- data |> 
    filter(RT < 1000) |> 
    ggplot(aes(x = Illusion_Difference, y = RT)) +
    geom_point() +
    geom_line(data=estimate_relation(model), aes(y=Predicted), color="blue") +
    geom_line(data=estimate_relation(model_rt), aes(y=Predicted), color="red") +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_modern() +
    labs(
      color = "Illusion Strength", 
      fill = "Illusion Strength",
      y = "RT",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
    facet_wrap(~Illusion_Side, ncol=2, labeller = "label_both")

  p1 / p2
}

model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
                              data = data)

visualize(data, model_err, model_rt)
```

### 2022-08-04 study1.Rmd
update stim

@@
 
-cache <- FALSE
+cache <- TRUE
 fig.width <- see::golden_ratio(7)
@@
             name <- paste0("dif", k1, i, "-", "str", k2, j, side)
-            print(name)
+            # print(name)
             f <- paste0("poly(Illusion_Difference", 
@@
 
-parameters::parameters(model)
+# parameters::parameters(model)
 ```

### 2022-08-04 study0/make_stimuli.py

@@ -11,7 +11,7 @@ import pyllusion as ill
 # Parameters
 width = 800
 height = 800
-n = 64
+n = 80
 
 # Delete all existing stimuli
 for f in glob.glob("stimuli/*"):
@@ -156,7 +156,7 @@ ill.Zollner(illusion_strength=-40, difference=8).to_image(
 data = generate_images(
     data,
     strengths=[0],
-    differences=doublelinspace(mini=0.001, maxi=5, size=n),
+    differences=doublelinspace(mini=0.001, maxi=8, size=n),
     function=ill.Zollner,
     name="Zollner",
 )

### 2022-08-04 study0/preprocessing.R

@@ -8,7 +8,7 @@ preprocess <- function(file){
     return(data.frame())
   }
 
-  data <- data[~is.na(data$screen)]
+  data <- data[!is.na(data$screen), ]
   trials <- data[data$screen == 'Trial', ]
 
   df <- data.frame(
@@ -36,25 +36,12 @@ preprocess <- function(file){
   df
 }
 
-  # Run ---------------------------------------------------------------------
-  participants <- list.files("study0/data/")
+# Run ---------------------------------------------------------------------
+participants <- list.files("data/")
 
 
-  df <- data.frame()
-  for (ppt in participants) {
-    df <- rbind(df, preprocess_raw(file = paste0("study0/data/", ppt)))
-  }
-
-  # Study 0
-  df$Study <- 0
-  df$Pyllusion <- "1.2"
-
-
-  # Transformation
-  df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
-  df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
-  df$Illusion_Difference_cbrt <- df$Illusion_Difference**(1/3)
-
-  # Save anonmized data
-  write.csv(df, "data/study0.csv", row.names = FALSE)
-
+df <- data.frame()
+for (ppt in participants) {
+  df <- rbind(df, preprocess(file = paste0("data/", ppt)))
+}
+df$Illusion_Side <- as.factor(df$Illusion_Side)

### 2022-08-04 study0/experiment.js
update instructions and remove scores

@@ -72,7 +72,7 @@ var fixation = {
     type: jsPsychHtmlKeyboardResponse,
     stimulus: '<div style="font-size:60px;">+</div>',
     choices: "NO_KEYS" /* no responses will be accepted as a valid response */,
-    //trial_duration: 0, // (for testing)
+    // trial_duration: 0, // (for testing)
     trial_duration: function () {
         return randomInteger(500, 1000) // Function from RealityBending/JSmisc
     },
@@ -243,7 +243,7 @@ function create_debrief(illusion_name = "Ponzo") {
             )
             var show_screen = get_debrief_display(results)
             return (
-                show_screen.display_score +
+                // show_screen.display_score +
                 // "<hr>" +
                 // // For debugging purposes, show the raw data.
                 // show_screen.display_accuracy +
@@ -252,7 +252,8 @@ function create_debrief(illusion_name = "Ponzo") {
                 // "<hr>" +
                 // //
                 // show_screen.display_comparison +
-                "<hr><p>Can you do better in the next block?</p>"
+                "<p>Great job!</p>" +
+                "<p>Can you do better in the next block?</p>"
             )
         },
         data: { screen: "block_results" },

### 2022-08-05 study0/analysis.Rmd
update stim

@@
 
-df <- filter(df, RT < 1000)
+df <- filter(df, RT < 900)
 ```
@@
 visualize <- function(data, model_err, model_rt) {
-  model <- mgcv::gam(Error ~ s(Illusion_Difference, by = Illusion_Side),
+
+  at <- list(Illusion_Difference=seq(0, max(data$Illusion_Difference), length.out=30),
+             Illusion_Side = levels(unique(data$Illusion_Side)))
+
+  model <- mgcv::gamm(Error ~ s(Illusion_Difference, by = Illusion_Side),
+                      random = list(Participant = ~1) ,
+                      data = data,
+                      family = "binomial")
+  model_lin <- glmmTMB::glmmTMB(Error ~ Illusion_Difference + (1|Participant) ,
                       data = data,
@@
     geom_bar(stat="identity") +
-    geom_line(data=estimate_relation(model), aes(y=Predicted), color="blue") +
-    geom_line(data=estimate_relation(model_err), aes(y=Predicted), color="red") +
+    geom_hline(yintercept=0.5, linetype="dotted") +
+    geom_line(data=estimate_relation(model_lin, at=at[1]), aes(y=Predicted), color="black") +
+    geom_line(data=estimate_relation(model, at=at), aes(y=Predicted), color="blue") +
+    geom_line(data=estimate_relation(model_err, at=at[1]), aes(y=Predicted), color="red") +
     scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
@@
 
-  model <- mgcv::gam(RT ~ s(Illusion_Difference, by = Illusion_Side),
+  model <- mgcv::gamm(RT ~ s(Illusion_Difference, by = Illusion_Side),
+                     random = list(Participant = ~1) ,
                      data = filter(data, Error == 0, RT < 1000))
+    model_lin <- glmmTMB::glmmTMB(RT ~ Illusion_Difference + (1|Participant) ,
+                      data = data)
 
@@
     filter(RT < 1000) |> 
+    # group_by(Illusion_Difference, Illusion_Side) |> 
+    # summarize(RT = mean(RT)) |> 
     ggplot(aes(x = Illusion_Difference, y = RT)) +
-    geom_point() +
-    geom_line(data=estimate_relation(model), aes(y=Predicted), color="blue") +
-    geom_line(data=estimate_relation(model_rt), aes(y=Predicted), color="red") +
+    ggdist::stat_pointinterval(aes(group=Illusion_Difference)) +
+    geom_line(data=estimate_relation(model_lin, at=at[1]), aes(y=Predicted), color="black") +
+    geom_line(data=estimate_relation(model, at=at), aes(y=Predicted), color="blue") +
+    geom_line(data=estimate_relation(model_rt, at=at[1]), aes(y=Predicted), color="red") +
     scale_y_continuous(expand = c(0, 0)) +
@@
 
+model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
+                              data = data, family = "binomial")
+model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
+                              data = data)
+
+visualize(data, model_err, model_rt)
+```
+
+```{r message=FALSE, warning=FALSE}
+optimal <- function(model_err) {
+  pred <- estimate_relation(model_err, length=500)
+
+  bounds <- pred |> 
+      slice(c(which.min(abs(Predicted - 0.25)), 
+              which.min(abs(Predicted - 0.025))))
+
+  rez <- bounds |> 
+    transmute(Illusion_Difference_log = Illusion_Difference_log,
+              Illusion_Difference = exp(Illusion_Difference_log) - 1,
+              Error = insight::format_value(Predicted, as_percent=TRUE))
+
+  p <- pred |> 
+    ggplot(aes(y = Predicted, x = Illusion_Difference_log)) +
+    geom_line() +
+    geom_hline(yintercept = c(0.25, 0.025), linetype="dotted", color=c("red", "green")) +
+    geom_vline(xintercept = bounds$Illusion_Difference_log, linetype="dotted", color=c("red", "green")) +
+    theme_modern()
+  list(rez=rez, p=p)
+}
+
+model_err <- glmmTMB::glmmTMB(Error ~ Illusion_Difference_log + (1|Participant),  
+                              data = data, family = "binomial")
+
+rez <- optimal(model_err)
+rez$rez
+rez$p
+```
+
+### Müller-Lyer
+
+```{r message=FALSE, warning=FALSE}
+data <- dplyr::filter(df, Illusion_Type == "Müller-Lyer")
+```
+
+```{r message=FALSE, warning=FALSE}
+best_models(data)
+```
+
+```{r message=FALSE, warning=FALSE}
+model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
+                              data = data, family = "binomial")
+model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
+                              data = data)
+
+visualize(data, model_err, model_rt)
+```
+
+```{r message=FALSE, warning=FALSE}
+model_err <- glmmTMB::glmmTMB(Error ~ Illusion_Difference_log + (1|Participant),  
+                              data = data, family = "binomial")
+
+rez <- optimal(model_err)
+rez$rez
+rez$p
+```
+
+### Vertical-Horizontal
+
+```{r message=FALSE, warning=FALSE}
+data <- dplyr::filter(df, Illusion_Type == "Vertical-Horizontal")
+```
+
+```{r message=FALSE, warning=FALSE}
+best_models(data)
+```
+
+```{r message=FALSE, warning=FALSE}
+model_err <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) + (1|Participant),  
+                              data = data, family = "binomial")
+model_rt <- glmmTMB::glmmTMB(RT ~ sqrt(Illusion_Difference) + (1|Participant),  
+                              data = data)
+
+visualize(data, model_err, model_rt)
+```
+
+### Poggendorff
+
+```{r message=FALSE, warning=FALSE}
+data <- dplyr::filter(df, Illusion_Type == "Poggendorff")
+```
+
+```{r message=FALSE, warning=FALSE}
+best_models(data)
+```
+
+```{r message=FALSE, warning=FALSE}
 model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
@@
 ```
+
+### Zöllner
+
+```{r message=FALSE, warning=FALSE}
+data <- dplyr::filter(df, Illusion_Type == "Zöllner")
+```
+
+```{r message=FALSE, warning=FALSE}
+best_models(data)
+```
+
+```{r message=FALSE, warning=FALSE}
+model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
+                              data = data, family = "binomial")
+model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
+                              data = data)
+
+visualize(data, model_err, model_rt)
+```
+
+### Rod-Frame
+
+```{r message=FALSE, warning=FALSE}
+data <- dplyr::filter(df, Illusion_Type == "Rod-Frame")
+```
+
+```{r message=FALSE, warning=FALSE}
+best_models(data)
+```
+
+```{r message=FALSE, warning=FALSE}
+model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
+                              data = data, family = "binomial")
+model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
+                              data = data)
+
+visualize(data, model_err, model_rt)
+```
+
+### Contrast
+
+```{r message=FALSE, warning=FALSE}
+data <- dplyr::filter(df, Illusion_Type == "Contrast")
+```
+
+```{r message=FALSE, warning=FALSE}
+best_models(data)
+```
+
+```{r message=FALSE, warning=FALSE}
+model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
+                              data = data, family = "binomial")
+model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
+                              data = data)
+
+visualize(data, model_err, model_rt)
+```
+
+### White
+
+```{r message=FALSE, warning=FALSE}
+data <- dplyr::filter(df, Illusion_Type == "White")
+```
+
+```{r message=FALSE, warning=FALSE}
+best_models(data)
+```
+
+```{r message=FALSE, warning=FALSE}
+model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
+                              data = data, family = "binomial")
+model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
+                              data = data)
+
+visualize(data, model_err, model_rt)
+```

### 2022-08-05 study1.Rmd
update stim

--- 2022-08-04 study1.Rmd
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-# plot(seq(0, 1, length.out=100), seq(0, 1, length.out=100)**(1/3), "l", col="green")
-# lines(seq(0, 1, length.out=100), sqrt(seq(0, 1, length.out=100)), "l", col="red")
-# lines(seq(0, 1, length.out=100), log(1+seq(0, 1, length.out=100)), "l", col="blue")
-# lines(seq(0, 1, length.out=100), seq(0, 1, length.out=100), "l", col="black")
-
 best_models <- function(data) {
@@
       for(k1 in c("", "_log", "_sqrt", "_cbrt")) { 
-        for(k2 in c("", "_log", "_sqrt", "_cbrt")) {  
+        for(k2 in c("")) {  
           for(side in c("", "-side")) {
             name <- paste0("dif", k1, i, "-", "str", k2, j, side)
-            print(name)
+            # print(name)
             f <- paste0("poly(Illusion_Difference", 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-cbrt <- function(x) sign(x) * abs(x)^(1/3)
+cbrt <- function(x) sign(x) * abs(x)**(1/3)
 
 formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
+  Error ~ cbrt(Illusion_Difference) * Illusion_Strength + 
     (1 | Participant),
@@
 
-  n_lines <- length(unique(data[[vardiff]]))
+  # Get predicted
   pred <- estimate_relation(model,
@@
                             length = c(NA, 25))
-  pred$Illusion_Difference <- as.factor(round(pred[[vardiff]], 2))
+  pred[[vardiff]] <- as.factor(pred[[vardiff]])
 
   # Set colors for lines
-  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)
-  diffvals <- as.numeric(as.character(unique(sort(pred$Illusion_Difference))))
+  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
+  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
   names(colors) <- diffvals
@@
   # Manual jittering
-  xrange <- 0.05*diff(range(data$Illusion_Strength))
-  data$x <- data$Illusion_Strength 
+  xrange <- 0.05*diff(range(data[[varstrength]]))
+  data$x <- data[[varstrength]]
   data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
@@
   pred |>
-    ggplot(aes(x = Illusion_Strength, y = Predicted)) +
+    ggplot(aes_string(x = varstrength, y = "Predicted")) +
     geom_dots(
-      data = mutate(data, 
-                    Illusion_Strength = round(Illusion_Strength, 1)),
+      data = data,
       aes(x=x, y = Error, group = Error, side = .dots_side, order=color), 
@@
     geom_slab(data=data, aes(y = Error)) +
-    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference, group = Illusion_Difference), alpha = 0.2) +
+    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
     geom_vline(xintercept = 0, linetype = "dashed") +
     geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
-    geom_line(aes(color = Illusion_Difference, group = Illusion_Difference)) +
+    geom_line(aes_string(color = vardiff, group = vardiff)) +
     scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
@@
     scale_fill_manual(values = colors) +
-    coord_cartesian(xlim=c(min(data$Illusion_Strength), max(data$Illusion_Strength))) +
+    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
     theme_modern() +
@@
 formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
+  Error ~ cbrt(Illusion_Difference) * Illusion_Strength + 
     (1 | Participant),
@@
 std_params(gam$model, min=0, max=20)
+```
+
+## Müller-Lyer
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Müller-Lyer")
+
+plot_descriptive(data, side = "updown")
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0, max=0.6)
+std_params(gam$model, min=0, max=0.6)
 ```
@@
 
-## Müller-Lyer
+## Poggendorff
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Müller-Lyer")
+data <- filter(df, Illusion_Type == "Poggendorff")
+
+plot_descriptive(data, side = "updown")
+
+data <- filter(data, abs(Illusion_Strength) < 45)
 
@@
 
-parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.6)
-std_params(gam$model, min=0, max=0.6)
-```
-
-## Poggendorff
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Poggendorff")
-
-plot_descriptive(data, side = "updown")
-
-data <- filter(data, abs(Illusion_Strength) < 45)
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-parameters::parameters(model)
+# parameters::parameters(model)
 ```

### 2022-08-05 preprocessing.R

@@ -106,11 +106,10 @@ df[df$Illusion_Type == "Zöllner", "Illusion_Strength"] <- -1 * round(df[df$Illu
 
 
 
-
 # Transformation
 df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
 df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
-df$Illusion_Difference_cbrt <- df$Illusion_Difference**(1/3)
+df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1/3), 4)
 df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
 df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
 df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))

### 2022-08-05 study2/experiment.js

@@ -13,7 +13,7 @@ function get_results(illusion_mean, illusion_sd, illusion_type) {
     if (correct_trials.count() > 0) {
         var rt_mean_correct = correct_trials.select("rt").mean()
         var ies = rt_mean_correct / proportion_correct // compute inverse efficiency score
-        var score_to_display = 100 - (ies/50)  // change max score to 100 instead of 5000
+        var score_to_display = 100 - ies / 50 // change max score to 100 instead of 5000
         var percentile =
             100 - cumulative_probability(ies, illusion_mean, illusion_sd) * 100
     } else {
@@ -88,9 +88,8 @@ var make_break = {
     choices: ["I am ready to continue!"],
     stimulus:
         "<p><b>CONGRATULATIONS!</b></p>" +
-        "<p>You have finished one block of visual illusions.</p>" +
-        "<p>In the second part, you will see all the illusions once again. <b>Try to beat your previous score!</b></p>" ,
-        // "<p>We know it's long and challenging, so take a moment to rest. When you are ready, press the button below.</p>",
+        "<p>You have finished half of the game. We know it's long and challenging, so we thank you in advance for staying focused 'til the end!</p>" +
+        "<p>You will now see all the illusions once again. <b>Try to beat your previous score!</b></p>",
     save_trial_parameters: {
         trial_duration: true,
     },
@@ -306,10 +305,9 @@ function make_trial(stimuli, instructions, illusion_name, type) {
     })
 
     // Debriefing Information
-    if (stimuli != stimuli_training){
-        timeline.push(create_debrief(illusion_name= illusion_name))
-    }
-    else if (stimuli=stimuli_training){
+    if (stimuli != stimuli_training) {
+        timeline.push(create_debrief((illusion_name = illusion_name)))
+    } else if ((stimuli = stimuli_training)) {
         timeline.push({
             type: jsPsychHtmlButtonResponse,
             choices: ["Continue"],
@@ -320,111 +318,111 @@ function make_trial(stimuli, instructions, illusion_name, type) {
                         "#jspsych-progressbar-container"
                     ).style.display = "inline")
             },
-            stimulus: "<p>Great job!</p>" +
-            "<p>Click NEXT when you're ready for the next illusion practice.</p>",
-            data: { screen: "practice_block" }
[…]

### 2022-08-05 study2/make_stimuli.py

@@ -3,6 +3,7 @@ import glob
 import json
 import os
 
+import matplotlib.pyplot as plt
 import neurokit2 as nk
 import numpy as np
 import pandas as pd
@@ -54,9 +55,9 @@ def generate_images(data, strengths, differences, function, name="Delboeuf", **k
             path = (
                 name
                 + "_str"
-                + str(np.round(strength, 2))
+                + str(np.round(strength, 4))
                 + "_diff"
-                + str(np.round(difference, 2))
+                + str(np.round(difference, 4))
                 + ".png"
             )
             img.save("stimuli/" + path)
@@ -112,12 +113,29 @@ def sqrtspace(mini=0.1, maxi=1, size=6):
     return np.concatenate((-1 * x[::-1], x))
 
 
-def doublelinspace(mini=0.1, maxi=1, size=6, exp=False):
-    if exp is False:
-        x = np.linspace(mini, maxi, int(size / 2), endpoint=True)
-    else:
-        x = nk.expspace(mini, maxi, int(size / 2), out=float, base=2)
-    return np.concatenate((-1 * x[::-1], x))
+def doublelinspace(mini=0.1, maxi=1, size=6, transformation="lin", show=True):
+    lin = np.linspace(mini, maxi, int(size / 2), endpoint=True)
+    exp = nk.expspace(mini, maxi, int(size / 2), out=float)
+    sq = (np.linspace(mini ** (1 / 2), maxi ** (1 / 2), int(size / 2), endpoint=True)) ** 2
+    cb = (np.linspace(mini ** (1 / 3), maxi ** (1 / 3), int(size / 2), endpoint=True)) ** 3
+
+    if show is True:
+        plt.plot(lin, [0] * len(lin), "o", label="linear")
+        plt.plot(exp, [0.2] * len(lin), "o", label="exp")
+        plt.plot(sq, [0.4] * len(lin), "o", label="square")
+        plt.plot(cb, [0.6] * len(lin), "o", label="cube")
+        plt.legend()
+
+    if transformation == "lin":
+        x = lin
+    elif transformation == "exp":
+        x = exp
[…]

### 2022-08-05 study2/stimuli/stimuli_training.js

@@ -1,1 +1,650 @@
-var stimuli_training = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": -0.5, "Difference": -1, "stimulus": "stimuli/Delboeuf_str-0.5_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -0.5, "Difference": 1, "stimulus": "stimuli/Delboeuf_str-0.5_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.0_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.0_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.5, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.5_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.5, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.5_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -0.5, "Difference": -0.8, "stimulus": "stimuli/Ebbinghaus_str-0.5_diff-0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowright"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -0.5, "Difference": 0.8, "stimulus": "stimuli/Ebbinghaus_str-0.5_diff0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_S
[…]

### 2022-08-05 study2/experiment.js
update instructions

@@ -88,7 +88,7 @@ var make_break = {
     choices: ["I am ready to continue!"],
     stimulus:
         "<p><b>CONGRATULATIONS!</b></p>" +
-        "<p>You have finished half of the game. We know it's long and challenging, so we thank you in advance for staying focused 'til the end!</p>" +
+        "<p>You have finished half of the game. We know it's long and challenging, so we appreciate you staying focused until the end!</p>" +
         "<p>You will now see all the illusions once again. <b>Try to beat your previous score!</b></p>",
     save_trial_parameters: {
         trial_duration: true,

### 2022-08-05 study1.Rmd
update stim

@@
 Crucially, it allowed us to refine the range of task difficulty and illusion strength for the stimuli of the next study to maximize the information gain.
+
+We generated stimuli with the difficulty levels spaced non-linearly depending on the best underlying model (i.e., with an exponential, square or cubic spacing if the best model was *log*, *square root* or *cube root*).
+For instance, a linear space of [0.1, 0.2, 0.3, 0.4] can be transformed to an exponential space of [[0.1       , 0.19207744, 0.29186238, 0.4       ]]

### 2022-08-05 study2/make_stimuli.py

@@ -107,12 +107,6 @@ def save_json(data, name="stimuli"):
         f.write(updatedfile)
 
 
-def sqrtspace(mini=0.1, maxi=1, size=6):
-    x = np.linspace(np.sqrt(0), np.sqrt(1), int(size / 2) + 1, endpoint=True) ** 2
-    x = nk.rescale(x[1::], [mini, maxi])
-    return np.concatenate((-1 * x[::-1], x))
-
-
 def doublelinspace(mini=0.1, maxi=1, size=6, transformation="lin", show=True):
     lin = np.linspace(mini, maxi, int(size / 2), endpoint=True)
     exp = nk.expspace(mini, maxi, int(size / 2), out=float)
@@ -380,7 +374,7 @@ ill.MullerLyer(illusion_strength=-10, difference=0.7).to_image(width=800, height
 )
 
 strengths = np.linspace(-49, 49, n - 1)
-diffs = doublelinspace(mini=0.04, maxi=0.46, size=n, transformation="square")
+diffs = doublelinspace(mini=0.04, maxi=0.46, size=n, transformation="cube")
 diff1 = np.concatenate((diffs[0 : n // 2 : 2], diffs[n // 2 + 1 :: 2]))
 diff2 = np.concatenate((diffs[1 : n // 2 : 2], diffs[n // 2 :: 2]))
 
@@ -418,7 +412,7 @@ ill.Ponzo(illusion_strength=5, difference=0.7).to_image(width=800, height=600).s
 
 
 strengths = np.linspace(-25.2, 25.2, n - 1)
-diffs = doublelinspace(mini=0.04, maxi=0.46, size=n, transformation="square")
+diffs = doublelinspace(mini=0.04, maxi=0.46, size=n, transformation="cube")
 diff1 = np.concatenate((diffs[0 : n // 2 : 2], diffs[n // 2 + 1 :: 2]))
 diff2 = np.concatenate((diffs[1 : n // 2 : 2], diffs[n // 2 :: 2]))
 

### 2022-08-05 study2/stimuli/stimuli_training.js

@@ -1,650 +1,1 @@
-var stimuli_training = [
-    {
-        Illusion_Type: "Delboeuf",
-        Illusion_Strength: -1.5499999999999998,
-        Difference: -1,
-        stimulus: "stimuli/Delboeuf_str-1.55_diff-1.png",
-        data: {
-            screen: "Trial",
-            block: "Delboeuf",
-            correct_response: "arrowright",
-        },
-    },
-    {
-        Illusion_Type: "Delboeuf",
-        Illusion_Strength: -1.5499999999999998,
-        Difference: 1,
-        stimulus: "stimuli/Delboeuf_str-1.55_diff1.png",
-        data: {
-            screen: "Trial",
-            block: "Delboeuf",
-            correct_response: "arrowleft",
-        },
-    },
-    {
-        Illusion_Type: "Delboeuf",
-        Illusion_Strength: 0.0,
-        Difference: -1,
-        stimulus: "stimuli/Delboeuf_str0.0_diff-1.png",
-        data: {
-            screen: "Trial",
-            block: "Delboeuf",
-            correct_response: "arrowright",
-        },
-    },
-    {
-        Illusion_Type: "Delboeuf",
-        Illusion_Strength: 0.0,
-        Difference: 1,
-        stimulus: "stimuli/Delboeuf_str0.0_diff1.png",
-        data: {
-            screen: "Trial",
-            block: "Delboeuf",
-            correct_response: "arrowleft",
-        },
-    },
-    {
-        Illusion_Type: "Delboeuf",
-        Illusion_Strength: 1.5499999999999998,
-        Difference: -1,
-        stimulus: "stimuli/Delboeuf_str1.55_diff-1.png",
-        data: {
-            screen: "Trial",
-            block: "Delboeuf",
-            correct_response: "arrowright",
-        },
-    },
-    {
-        Illusion_Type: "Delboeuf",
-        Illusion_Strength: 1.5499999999999998,
-        Difference: 1,
-        stimulus: "stimuli/Delboeuf_str1.55_diff1.png",
-        data: {
-            screen: "Trial",
-            block: "Delboeuf",
-            correct_response: "arrowleft",
-        },
-    },
-    {
-        Illusion_Type: "Ebbinghaus",
-        Illusion_Strength: -1.4499999999999997,
-        Difference: -0.8,
-        stimulus: "stimuli/Ebbinghaus_str-1.45_diff-0.8.png",
-        data: {
-            screen: "Trial",
-            block: "Ebbinghaus",
[…]

### 2022-08-05 study1.Rmd
edit instructions

--- 2022-08-05 study1.Rmd
@@
 Crucially, it allowed us to refine the range of task difficulty and illusion strength for the stimuli of the next study to maximize the information gain.
+
+Perceptual decisions often exhibit a non-linear relationship with the objective physical values (e.g., a logarithmic relationship.
+In order to spread the difficulty levels based on the behavioral performance (to maximize the diversity of the stimuli in regards to perceptual processes), we generated, for study 2, stimuli with the difficulty levels spaced non-linearly, depending on the best underlying model (i.e., with an exponential, square or cubic spacing if the best model was *log*, *square root* or *cube root*).
+For instance, a linear space of [0.1, 0.3, 0.5, 0.7] can be transformed to an exponential space of [0.1, 0.27, 0.47, 0.7].

### 2022-08-05 study2/experiment.js

@@ -318,9 +318,7 @@ function make_trial(stimuli, instructions, illusion_name, type) {
                         "#jspsych-progressbar-container"
                     ).style.display = "inline")
             },
-            stimulus:
-                "<p>Great job!</p>" +
-                "<p>Click NEXT when you're ready for the next illusion practice.</p>",
+            stimulus: "<p><b>Great job!</b></p>",
             data: { screen: "practice_block" },
         })
     }
@@ -401,7 +399,7 @@ const rodframe_instructions =
 const poggendorff_instructions =
     "<p>In this part, two parallel red lines will appear, but they are partially hidden by a grey rectangle.</p>" +
     "<p>Your task is to tell if the <b>red line to the right</b> of the rectangle is above or below the line to the left. You need to answer as fast as you can, without making errors.</p>" +
-    "<p>Press <b>the UP or the DOWN arrow</b> to indicate whether the <b>right red line is actually above or below the left red line.</b></p>" +
+    "<p>Press <b>the UP or the DOWN arrow</b> to indicate whether the <b>right red line</b> is actually <b>above or below the left red line.</b></p>" +
     "<div style='float: center'><img src='utils/stimuli_demo/Poggendorff_Demo.png' height='300'></img>" +
     "<p><img src='utils/answer/answer_updown_keyboard.PNG' height='150'></img></p>" +
     "<p class='small'>In this example, the correct answer is the <b>UP arrow</b>.</p></div>" +

### 2022-08-05 study2/experiment.js
minor change to score display
@anshu-97 I changed so that the display is a percentage

@@ -13,7 +13,10 @@ function get_results(illusion_mean, illusion_sd, illusion_type) {
     if (correct_trials.count() > 0) {
         var rt_mean_correct = correct_trials.select("rt").mean()
         var ies = rt_mean_correct / proportion_correct // compute inverse efficiency score
-        var score_to_display = 100 - ies / 50 // change max score to 100 instead of 5000
+        var score_to_display = 100 - ies / 50
+        if (score_to_display < 0) {
+            score_to_display = 0
+        }
         var percentile =
             100 - cumulative_probability(ies, illusion_mean, illusion_sd) * 100
     } else {
@@ -39,7 +42,7 @@ function get_debrief_display(results, type = "Block") {
             "<p>Your score for this illusion is " +
             '<p style="color: black; font-size: 48px; font-weight: bold;">' +
             Math.round(results.score) +
-            "</p>"
+            " %</p>"
     } else if (type === "Final") {
         // Final debriefing at end of game
         var score =

### 2022-08-05 study2/experiment.js
display 1 decimal place

@@ -41,7 +41,7 @@ function get_debrief_display(results, type = "Block") {
         var score =
             "<p>Your score for this illusion is " +
             '<p style="color: black; font-size: 48px; font-weight: bold;">' +
-            Math.round(results.score) +
+            Math.round(results.score*10)/10 +
             " %</p>"
     } else if (type === "Final") {
         // Final debriefing at end of game

### 2022-08-05 study1.Rmd
update stim

@@
 
+# IES
+df |> 
+  group_by(Participant, Illusion_Type) |>
+  summarize(Correct = 1-mean(Error)) |> 
+  ungroup() |> 
+  estimate_density(select="Correct") |> 
+  plot()
+
+df |> 
+  filter(Error == 0) |> 
+  group_by(Participant, Illusion_Type) |>
+  summarize(RT = mean(RT)) |> 
+  ungroup() |> 
+  estimate_density(select="RT") |> 
+  plot()
 ```
@@
 
+One thing that stood out was Zöller, which seems to exhibit a non-linear relationship of illusion effect.
+Next study, by generating a wider range of illusion strength, will attempt at clarifying this point.
+
 Perceptual decisions often exhibit a non-linear relationship with the objective physical values (e.g., a logarithmic relationship.

### 2022-08-05 study2/make_stimuli.py

@@ -63,7 +63,14 @@ def generate_images(data, strengths, differences, function, name="Delboeuf", **k
             img.save("stimuli/" + path)
 
             # Compute expected response
-            if name in ["Delboeuf", "Ebbinghaus", "VerticalHorizontal", "White"]:
+            if name in [
+                "Delboeuf",
+                "Ebbinghaus",
+                "VerticalHorizontal",
+                "White",
+                "Zollner",
+                "RodFrame",
+            ]:
                 if difference > 0:
                     correct = "arrowleft"
                 else:
@@ -73,11 +80,6 @@ def generate_images(data, strengths, differences, function, name="Delboeuf", **k
                     correct = "arrowup"
                 else:
                     correct = "arrowdown"
-            elif name in ["Zollner", "RodFrame"]:
-                if difference < 0:
-                    correct = "arrowleft"
-                else:
-                    correct = "arrowright"
 
             # Save parameters for Delboeuf Illusion
             data.append(
@@ -294,12 +296,12 @@ data_block2 = generate_images(
 )
 
 # -------------------------- Zollner Illusion --------------------------
-ill.Zollner(illusion_strength=-40, difference=8).to_image(width=800, height=600).save(
+ill.Zollner(illusion_strength=-45, difference=8).to_image(width=800, height=600).save(
     "utils/stimuli_demo/Zollner_Demo.png"
 )
 
-strengths = np.linspace(-42, 42, n - 1)
-diffs = doublelinspace(mini=0.15, maxi=5, size=n, transformation="cube")
+strengths = np.linspace(-77, 77, n - 1)
+diffs = doublelinspace(mini=0.15, maxi=4.15, size=n, transformation="cube")
 diff1 = np.concatenate((diffs[0 : n // 2 : 2], diffs[n // 2 + 1 :: 2]))
 diff2 = np.concatenate((diffs[1 : n // 2 : 2], diffs[n // 2 :: 2]))
 

### 2022-08-05 study2/stimuli/stimuli_training.js

@@ -1,1 +1,1 @@
-var stimuli_training = [{"Illusion_Type": "Delboeuf", "Illusion_Strength": -1.5499999999999998, "Difference": -1, "stimulus": "stimuli/Delboeuf_str-1.55_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": -1.5499999999999998, "Difference": 1, "stimulus": "stimuli/Delboeuf_str-1.55_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": -1, "stimulus": "stimuli/Delboeuf_str0.0_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 0.0, "Difference": 1, "stimulus": "stimuli/Delboeuf_str0.0_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 1.5499999999999998, "Difference": -1, "stimulus": "stimuli/Delboeuf_str1.55_diff-1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowright"}}, {"Illusion_Type": "Delboeuf", "Illusion_Strength": 1.5499999999999998, "Difference": 1, "stimulus": "stimuli/Delboeuf_str1.55_diff1.png", "data": {"screen": "Trial", "block": "Delboeuf", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -1.4499999999999997, "Difference": -0.8, "stimulus": "stimuli/Ebbinghaus_str-1.45_diff-0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowright"}}, {"Illusion_Type": "Ebbinghaus", "Illusion_Strength": -1.4499999999999997, "Difference": 0.8, "stimulus": "stimuli/Ebbinghaus_str-1.45_diff0.8.png", "data": {"screen": "Trial", "block": "Ebbinghaus", "correct_response": "arrowleft"}}, {"Illusion_Type": "Ebbinghaus", "Illusion
[…]

### 2022-08-05 study1.Rmd
init study2

@@
 # Methods
+
+```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+# PREPROCESSING ================================================================
+source("preprocessing.R")
+
+# This is a local folder containing raw data from unzipped pavlovia
+# It has been added to .gitignore to NOT be published on github
+# (it contains the subject ID of the participants)
+participants <- list.files("study1/data/")
+
+df <- data.frame()
+for (ppt in participants) {
+  df <- rbind(df, preprocess_raw(file = paste0("study1/rawdata/", ppt)))
+}
+
+df$Pyllusion <- "1.1"
+df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"])
+df[df$Illusion_Type == "Zöllner", "Illusion_Strength"] <- -1 * round(df[df$Illusion_Type == "Zöllner", "Illusion_Strength"], 1)
+
+# Save anonymized data
+write.csv(df, "data/study1.csv", row.names = FALSE)
+```
 
@@
 # Dear participant, thank you for participating in our study.
-Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which make the data unusable for us.
-We understand that you might have been in a hurry, or had some other issues, and kindly ask you to return your participation.
+Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes your data unusable.
+We understand that you might have been in a hurry or had some other issues, and kindly ask you to return your participation.
 We hope to open-up more slots in the future would you be interested to participate again.
+
+# Dear participant, thank you for participating in our study.
+Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes it unusable.
+We understand that you might have been in a hurry or had some other issues; we hope to open-up more slots in the future would you be interested to participate again.
 
@@
   # Error rate of 47.9%
-  # Prolific Status: RETURN REQUESTED
+  # Prolific Status: REJECTED
   "61280140171ec546e87ed8cb_qdlgy",
   # Error rate of 42.1% and very large RT SD
-  # Prolific Status: RETURN REQUESTED
+  # Prolific Status: REJECTED
   "5d398380b37ab1000111fac3_2nxxh"

### 2022-08-05 study2.Rmd
init study2

--- 2022-08-05 study1.Rmd
@@
 ---
-title: "Illusion Game Validation (Study 1)"
+title: "Illusion Game Validation"
 output: html_document
@@
 ```{r, echo = FALSE, warning=FALSE, message=FALSE}
-options(digits = 3,
-        mc.cores = 4,
-        brms.algorithm = "meanfield",
-        brms.backend = "cmdstanr")
+options(digits = 3)
 
@@
 
+```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+# PREPROCESSING ================================================================
+source("preprocessing.R")
+
+# This is a local folder containing raw data from unzipped pavlovia
+# It has been added to .gitignore to NOT be published on github
+# (it contains the subject ID of the participants)
+participants <- list.files("study2/data/")
+
+df <- data.frame()
+for (ppt in participants) {
+  df <- rbind(df, preprocess_raw(file = paste0("study2/data/", ppt)))
+}
+
+df$Pyllusion <- "1.2"
+
+# Save anonymized data
+write.csv(df, "data/study2.csv", row.names = FALSE)
+```
+
 ```{r message=FALSE, warning=FALSE}
@@
 
-df <- read.csv("data/study1.csv") |>
+df <- read.csv("data/study2.csv") |>
   mutate(
@@
     Illusion_Side = as.factor(Illusion_Side),
-    Block = as.factor(Block),
-    Education = fct_relevel(Education, "Master", "Bachelor", "High School", "Other")
+    Block = as.factor(Block)
+    # Education = fct_relevel(Education, "Master", "Bachelor", "High School", "Other")
   )
@@
 # Dear participant, thank you for participating in our study.
-Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which make the data unusable for us.
-We understand that you might have been in a hurry, or had some other issues, and kindly ask you to return your participation.
+Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes your data unusable.
+We understand that you might have been in a hurry or had some other issues, and kindly ask you to return your participation.
 We hope to open-up more slots in the future would you be interested to participate again.
 
-outliers <- c(
-  # Half of the trials have very short RT
-  # Prolific Status: REJECTED
-  "60684f29dbfe1bb2059e5e27_rkqoy",
-  # Block n2 with very short RTs
-  # Prolific Status: RETURNED
-  "5e860198a846e30497df5189_6e43s",
-  # Error rate of 46.2% and short RTs
-  # Prolific Status: RETURNED
-  "615f319bb341cf2f306d858d_qsaq5",
-  # Error rate of 46.2%
-  # Prolific Status: RETURNED
-  "614f36fd81c78b7a125c4262_6ax4g",
-  # Error rate of 47.9%
-  # Prolific Status: RETURN REQUESTED
-  "61280140171ec546e87ed8cb_qdlgy",
-  # Error rate of 42.1% and very large RT SD
-  # Prolific Status: RETURN REQUESTED
-  "5d398380b37ab1000111fac3_2nxxh"
-)
+# Dear participant, thank you for participating in our study.
+Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes it unusable.
+We understand that you might have been in a hurry or had some other issues; we hope to open-up more slots in the future would you be interested to participate again.
+
+outliers <- c()
 ```
@@
 ```
-
-## Participants
-
-```{r message=FALSE, warning=FALSE}
-dfsub <- df |>
-  group_by(Participant) |>
-  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS) |>
-  slice(1) |>
-  ungroup()
-```
-
-The final sample included `r report::report_participants(dfsub, age="Age", sex="Sex")`.
-
-```{r }
-plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", fill = "orange") {
-  dfsub |>
-    ggplot(aes_string(x = what)) +
-    geom_density(fill = fill) +
-    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype = "dashed") +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_y_continuous(expand = c(0, 0)) +
-    ggtitle(title, subtitle = subtitle) +
-    theme_modern() +
-    theme(
-      plot.title = element_text(face = "bold", hjust = 0.5),
-      plot.subtitle = element_text(face = "italic", hjust = 0.5),
-      axis.title.x = element_blank(),
-      axis.title.y = element_blank(),
-      axis.text.y = element_blank()
-    )
-}
-
-plot_waffle <- function(dfsub, what = "Nationality") {
-  ggwaffle::waffle_iron(dfsub, what) |>
-    # mutate(label = emojifont::fontawesome('fa-twitter')) |>
-    ggplot(aes(x, y, fill = group)) +
-    ggwaffle::geom_waffle() +
-    # geom_point() +
-    # geom_text(aes(label=label), family='fontawesome-webfont', size=4) +
-    coord_equal() +
-    ggtitle(what) +
-    labs(fill = "") +
-    theme_void() +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-}
-```
-
-```{r fig.width=20, fig.height=15}
-p1 <- plot_distribution(dfsub, "Age", fill = "#FF9800")
-p2 <- plot_distribution(dfsub, "Duration", title = "Total Duration", subtitle = "in minutes", fill = "#F44336")
-p3 <- plot_distribution(dfsub, "Break_Duration", title = "Break Duration", subtitle = "in minutes", fill = "#3F51B5")
-
-p4 <- plot_waffle(dfsub, "Sex") +
-  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))
-
-p5 <- plot_waffle(dfsub, "Education") +
-  scale_fill_viridis_d()
-
-p6 <- plot_waffle(dfsub, "Nationality") +
-  scale_fill_metro_d()
-
-p7 <- plot_waffle(dfsub, "Ethnicity") +
-  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0"))
-
-p8 <- plot_waffle(dfsub, "Screen_Resolution") +
-  scale_fill_pizza_d()
-
-p9 <- plot_waffle(dfsub, "Device_OS") +
-  scale_fill_bluebrown_d()
-
-# p10 <- plot_waffle(dfsub, "Screen_Refresh") +
-#   scale_fill_viridis_d()
-
-(p1 / p2 / p3) | (p4 / p5 / p6) | (p7 / p8 / p9)
-```
-
-## Data Analysis
-
-The analysis of study focused on the probability of errors as the main outcome variable.
-For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some insight on the underlying generative model.
-Next, we tested the performance of various models differing in their specifications, such as: with or without a log-transform on the task difficulty, with or without a 2nd order polynomial term of the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
-We then fitted the best performing model under a Bayesian model, and compared its predicted links with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships at the expense of model simplicity.
-Finally, we used these models to estimate different values of task difficulty that would lead to similar error rates (2.5, 5 and 25%) when the illusion strength is null.
-
-# Results {.tabset}
-
-```{r eval=FALSE, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE}
-# Exploratory 
-# ============
-
-# ISI
-model <- glmmTMB::glmmTMB(Error ~ Illusion_Type * poly(ISI, 2) + (1|Participant),
-                    data=mutate(df, Illusion_Type = as.factor(Illusion_Type)),
-                    family="binomial")
-
-estimate_relation(model, at = c("ISI", "Illusion_Type")) |>
-  ggplot(aes(x = ISI, y = Predicted, color="Illusion_Type")) +
-  geom_line()
-
-model <- glmmTMB::glmmTMB(RT ~ Illusion_Type * poly(ISI, 2) + (1|Participant),
-                    data=mutate(df, Illusion_Type = as.factor(Illusion_Type)) |>
-                      filter(Error == 0))
-
-estimate_relation(model, at = c("ISI", "Illusion_Type")) |>
-  ggplot(aes(x = ISI, y = Predicted, color=Illusion_Type)) +
-  geom_line()
-
-# Difficulty - Link type
-ps <- list()
-for(illusion in c("Delboeuf", "Ebbinghaus", "Rod-Frame", "Poggendorff", 
-"Vertical-Horizontal", "Zöllner", "Müller-Lyer", "Ponzo", 
-"Contrast", "White")) {
-  data <- filter(df, Illusion_Type == illusion, Illusion_Strength == 0)
-  model <- mgcv::gamm(Error ~ s(Illusion_Difference, k=4),
-                      random = list(Participant = ~1),
-                      data=data,
-                      family="binomial")
-
-  ps[[illusion]] <- estimate_relation(model, 
-                                     at = list(Illusion_Difference = unique(data$Illusion_Difference))) |>
-    ggplot(aes(x = Illusion_Difference, y = Predicted)) +
-    geom_bar(stat="identity", alpha=0.3) +
-    geom_line() +
-    ggtitle(illusion)
-}
-
-see::plots(ps)
-
-```
-
-## Delboeuf
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_descriptive <- function(data, side="leftright") {
-
-  if(side == "leftright") {
-    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
-    x <- tools::toTitleCase(gsub("arrow", "", x))
-    if(x == "Left") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
-    } else if(x == "Right") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
-    }
-  } else {
-    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
-    x <- tools::toTitleCase(gsub("arrow", "", x))
-    if(x == "Up") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
-    } else if(x == "Down") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
-    }
-    data$Answer <- fct_rev(data$Answer)
-  }
-
-  dodge1 <- 0.1 * diff(range(data$Illusion_Difference))
-  dodge2 <- -0.1 * diff(range(data$Illusion_Strength))
-
-  colors <- colorRampPalette(c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0"))(length(unique(data$Illusion_Strength)))
-
-  p1 <- data |> 
-    group_by(Illusion_Difference, Illusion_Strength, Answer) |> 
-    summarize(Error = mean(Error)) |> 
-    mutate(Illusion_Strength = as.factor(round(Illusion_Strength, 2))) |> 
-    ggplot(aes(x = Illusion_Difference, y = Error)) +
-    geom_bar(aes(fill=Illusion_Strength), position = position_dodge(width=dodge1), stat="identity") +
-    geom_line(aes(color = Illusion_Strength), position = position_dodge(width=dodge1)) +
-    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_color_manual(values = colors) +
-    scale_fill_manual(values = colors) +
-    theme_modern() +
-    labs(
-      color = "Illusion Strength", 
-      fill = "Illusion Strength",
-      y = "Probability of Error",
-      x = "Task Difficulty"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-
-  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data$Illusion_Difference)))
-
-  p2 <- data |> 
-    group_by(Illusion_Difference, Illusion_Strength, Answer) |> 
-    summarize(Error = mean(Error)) |> 
-    mutate(Illusion_Difference = as.factor(round(Illusion_Difference, 2))) |> 
-    ggplot(aes(x = Illusion_Strength, y = Error)) +
-    geom_vline(xintercept=0, linetype="dotted", alpha=0.6) +
-    geom_bar(aes(fill=Illusion_Difference), position = position_dodge(width=dodge2), stat="identity") +
-    geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
-    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_color_manual(values = colors) +
-    scale_fill_manual(values = colors) +
-    theme_modern() +
-    labs(
-      color = "Task Difficulty", 
-      fill = "Task Difficulty",
-      y = "Probability of Error",
-      x = "Illusion Strength"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-
-  if(side == "leftright") {
-    p <- ((p1 + facet_wrap(~Answer, ncol=2, labeller = "label_both")) /
-      (p2 + facet_wrap(~Answer, ncol=2, labeller = "label_both"))) + 
-    plot_annotation(title = paste(data$Illusion_Type[1], "Illusion"), 
-                    theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
-  } else {
-    p <- ((p1 + facet_wrap(~Answer, nrow=2, labeller = "label_both")) |
-      (p2 + facet_wrap(~Answer, nrow=2, labeller = "label_both"))) + 
-    plot_annotation(title = paste(data$Illusion_Type[1], "Illusion"), 
-                    theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
-  }
-  p
-}
-
-data <- filter(df, Illusion_Type == "Delboeuf")
-
-plot_descriptive(data)
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models <- function(data) {
-  models <- list()
-  for(i in 1:1) {
-    for(j in 1:1) {
-      for(k1 in c("", "_log", "_sqrt", "_cbrt")) { 
-        for(k2 in c("")) {  
-          for(side in c("", "-side")) {
-            name <- paste0("dif", k1, i, "-", "str", k2, j, side)
-            # print(name)
-            f <- paste0("poly(Illusion_Difference", 
-                        k1,
-                        ", ",
-                        i,
-                        ") * poly(Illusion_Strength",
-                        k2, 
-                        ", ",
-                        j, 
-                        ") + (1|Participant)")
-
-            if(side == "-side") f <- paste0("Illusion_Side * ", f)
-
-            m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)), 
-                                  data = data, family = "binomial")
-            if(performance::check_convergence(m)) {
-              models[[name]] <- m
-            }
-          }
-        }
-      }
-    }
-  }
-
-  to_keep <- compare_performance(models, metrics = c("BIC")) |> 
-    arrange(BIC) |> 
-    slice(1:10) |> 
-    pull(Name)
-
-  test <- test_performance(models[to_keep], reference=1)
-  perf <- compare_performance(models[to_keep], metrics = c("BIC", "R2"))
-
-  merge(perf, test) |> 
-    arrange(BIC) |> 
-    select(Name, BIC, R2_marginal, BF) |> 
-    mutate(BF = insight::format_bf(BF, name=""))
-}
-
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-cbrt <- function(x) sign(x) * abs(x)**(1/3)
-
-formula <- brms::bf(
-  Error ~ cbrt(Illusion_Difference) * Illusion_Strength + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model <- function(data, model) {
-  data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))
-
-  # Get variables
-  vars <- insight::find_predictors(model)$conditional
-  vardiff <- vars[1]
-  varstrength <- vars[2]
-
-  # Get predicted
-  pred <- estimate_relation(model,
-                            at = vars,
-                            length = c(NA, 25))
-  pred[[vardiff]] <- as.factor(pred[[vardiff]])
-
-  # Set colors for lines
-  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
-  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
-  names(colors) <- diffvals
-
-  # Assign color from the same palette to every observation of data (for geom_dots)
-  closest <- diffvals[max.col(-abs(outer(data[[vars[1]]], diffvals, "-")))]
-  data$color <- colors[as.character(closest)]
-  data$color <- fct_reorder(data$color, closest)
-
-  # Manual jittering
-  xrange <- 0.05*diff(range(data[[varstrength]]))
-  data$x <- data[[varstrength]]
-  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
-  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
-  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)
-
-  pred |>
-    ggplot(aes_string(x = varstrength, y = "Predicted")) +
-    geom_dots(
-      data = data,
-      aes(x=x, y = Error, group = Error, side = .dots_side, order=color), 
-      fill = data$color,
-      color = NA, 
-      alpha=0.5) +
-    geom_slab(data=data, aes(y = Error)) +
-    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
-    geom_vline(xintercept = 0, linetype = "dashed") +
-    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
-    geom_line(aes_string(color = vardiff, group = vardiff)) +
-    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_color_manual(values = colors) +
-    scale_fill_manual(values = colors) +
-    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
-    theme_modern() +
-    labs(
-      title = paste0(data$Illusion_Type[1], " Illusion"),
-      color = "Difficulty", fill = "Difficulty",
-      y = "Probability of Error",
-      x = "Illusion Strength"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-}
-
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-make_gam <- function(data) {
-
-  formula <- brms::bf(
-    Error ~ t2(Illusion_Difference, Illusion_Strength, bs = "cr", k=4) + 
-      (1 | Participant),
-    family = "bernoulli"
-  )
-
-  model <- brms::brm(formula,
-    data = data,
-    refresh = 0
-  )
-
-  list(p = plot_model(data, model), model = model)
-}
-
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-std_params <- function(model, min=0, max=2) {
-  estimate_relation(
-  model,
-    at = list(Illusion_Strength = c(0), 
-              Illusion_Difference = seq(min, max, length.out=500)),
-    ) |> 
-    select(Illusion_Strength, Illusion_Difference, Error = Predicted) |> 
-    slice(c(which.min(abs(Error - 0.005)), 
-            which.min(abs(Error - 0.025)), 
-            which.min(abs(Error - 0.25)))) |> 
-    mutate(Error = insight::format_value(Error, as_percent=TRUE))
-}
-
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=2)
-std_params(gam$model, min=0, max=2)
-```
-
-## Ebbinghaus
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Ebbinghaus")
-
-plot_descriptive(data)
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ cbrt(Illusion_Difference) * Illusion_Strength + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=2)
-std_params(gam$model, min=0, max=2)
-```
-
-## Rod and Frame
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Rod-Frame")
-
-plot_descriptive(data)
-
-data <- filter(data, abs(Illusion_Strength) < 15)
-
-plot_descriptive(data)
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0.01, max=12)
-std_params(gam$model, min=0.01, max=12)
-```
-
-## Vertical-Horizontal
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Vertical-Horizontal")
-
-plot_descriptive(data)
-
-data <- filter(data, abs(Illusion_Strength) < 90)
-
-plot_descriptive(data)
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * Illusion_Strength + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.40)
-std_params(gam$model, min=0, max=0.40)
-```
-
-## Zöllner
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Zöllner")
-
-plot_descriptive(data)
-
-data <- filter(data, abs(Illusion_Strength) < 45)
-
-plot_descriptive(data)
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=5)
-```
-
-## White
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "White")
-
-plot_descriptive(data)
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-# unique(data$Illusion_Strength)
-
-std_params(model, min=0, max=20)
-std_params(gam$model, min=0, max=20)
-```
-
-## Müller-Lyer
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Müller-Lyer")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.6)
-std_params(gam$model, min=0, max=0.6)
-```
-
-## Ponzo
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Ponzo")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=20)
-std_params(gam$model, min=0, max=20)
-```
-
-## Poggendorff
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Poggendorff")
-
-plot_descriptive(data, side = "updown")
-
-data <- filter(data, abs(Illusion_Strength) < 45)
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.5)
-std_params(gam$model, min=0, max=0.5)
-```
-
-## Contrast
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Contrast")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ Illusion_Difference * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=25)
-std_params(gam$model, min=0, max=25)
-```
-
-# Discussion
-
-This study allowed to elaborate a more precise prior idea regarding the magnitude of the effects at stake and the type of interaction between them.
-Crucially, it allowed us to refine the range of task difficulty and illusion strength for the stimuli of the next study to maximize the information gain.
-
-Perceptual decisions often exhibit a non-linear relationship with the objective physical values (e.g., a logarithmic relationship.
-In order to spread the difficulty levels based on the behavioral performance (to maximize the diversity of the stimuli in regards to perceptual processes), we generated, for study 2, stimuli with the difficulty levels spaced non-linearly, depending on the best underlying model (i.e., with an exponential, square or cubic spacing if the best model was *log*, *square root* or *cube root*).
-For instance, a linear space of [0.1, 0.3, 0.5, 0.7] can be transformed to an exponential space of [0.1, 0.27, 0.47, 0.7].

### 2022-08-05 preprocessing.R

@@ -50,6 +50,14 @@ preprocess_raw <- function(file) {
   # Correct duration
   df$Duration <- df$Duration - df$Break_Duration
 
+  # Transformation
+  df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
+  df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
+  df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1/3), 4)
+  df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
+  df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
+  df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))
+
   # Format names
   df$Illusion_Type <- ifelse(df$Illusion_Type == "MullerLyer", "Müller-Lyer", df$Illusion_Type)
   df$Illusion_Type <- ifelse(df$Illusion_Type == "Zollner", "Zöllner", df$Illusion_Type)
@@ -80,39 +88,4 @@ preprocess_raw <- function(file) {
 
 
 
-# Run ---------------------------------------------------------------------
 
-
-# This is a local folder containing raw data from unzipped pavlovia
-# It has been added to .gitignore to NOT be published on github
-# (it contains the subject ID of the participants)
-participants <- list.files("study1/rawdata/")
-
-
-df <- data.frame()
-for (ppt in participants) {
-  df <- rbind(df, preprocess_raw(file = paste0("study1/rawdata/", ppt)))
-}
-
-# Study 1
-df$Study <- 1
-df$Pyllusion <- "1.1"
-# df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Delboeuf", "Illusion_Difference"])
-# df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"] <- sqrt(df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"])
-df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame", "Illusion_Strength"])
[…]

### 2022-08-05 study1.Rmd
update RodFrame stim

@@
 df[df$Illusion_Type == "Zöllner", "Illusion_Strength"] <- -1 * round(df[df$Illusion_Type == "Zöllner", "Illusion_Strength"], 1)
+
+# Transformation
+df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
+df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
+df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1/3), 4)
+df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
+df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
+df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))
 
@@
 ```{r message=FALSE, warning=FALSE}
-# plot(estimate_density(dfraw[dfraw$Participant == "60684f29dbfe1bb2059e5e27_rkqoy", "RT"]))
-
-# Dear participant, thank you for participating in our study.
-Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes your data unusable.
-We understand that you might have been in a hurry or had some other issues, and kindly ask you to return your participation.
-We hope to open-up more slots in the future would you be interested to participate again.
-
-# Dear participant, thank you for participating in our study.
-Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes it unusable.
-We understand that you might have been in a hurry or had some other issues; we hope to open-up more slots in the future would you be interested to participate again.
-
 outliers <- c(

### 2022-08-05 study2.Rmd
update RodFrame stim

@@
 ---
-title: "Illusion Game Validation"
+title: "Illusion Game Validation (Study 2)"
 output: html_document
@@
 ```{r, echo = FALSE, warning=FALSE, message=FALSE}
-options(digits = 3)
+options(digits = 3,
+        mc.cores = 4,
+        brms.algorithm = "sampling",
+        brms.backend = "cmdstanr")
 
@@
 
-The goal of study 1 was to...
+The goal of study 2 was to...
 
@@
 df$Pyllusion <- "1.2"
+
+# Fix 
+df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"])
+
+# Transformation
+df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
+df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
+df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1/3), 4)
+df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
+df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
+df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))
 
@@
 # Dear participant, thank you for participating in our study.
-Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes your data unusable.
-We understand that you might have been in a hurry or had some other issues, and kindly ask you to return your participation.
-We hope to open-up more slots in the future would you be interested to participate again.
-
-# Dear participant, thank you for participating in our study.
 Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes it unusable.
@@
 
-outliers <- c()
+outliers <- c(
+  # Error rate of 48.8% Very short RT
+  # Prolific Status: NA
+  "5e66ed53de7896486f9a71f8_p8ckk"
+  )
 ```
@@
 
+### Reaction Time Distribution
+
+```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=20}
+# RT distribution
+p <- estimate_density(df, select = "RT", at = c("Participant", "Block")) |>
+  group_by(Participant) |>
+  normalize(select = "y") |>
+  ungroup() |>
+  mutate(color = ifelse(Participant %in% outliers, "red", "blue")) |>
+  ggplot(aes(x = x, y = y)) +
+  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
+  geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block)) +
+  geom_vline(xintercept = 150, linetype = "dashed", color = "red") +
+  scale_color_manual(values = c("red" = "red", "blue" = "blue"), guide = "none") +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  coord_cartesian(xlim = c(0, 3000)) +
+  theme_modern() +
+  theme(axis.text.y = element_blank()) +
+  facet_wrap(~Participant) +
+  labs(y = "", x = "Reaction Time (ms)")
+# p
+# ggsave("figures/outliers.png", p, width=20, height=15)
+
+# Filter out
+df <- filter(df, !Participant %in% outliers)
+```
+
 ### Error Rate per Illusion Block
@@
   ungroup() |>
-  mutate(n_trials = cumsum(n * 56),
-         p_trials = n_trials / nrow(df))
+  mutate(
+    n_trials = cumsum(n * 56),
+    p_trials = n_trials / nrow(df)
+  )
 
@@
   ggplot(aes(x = Illusion_Type, y = p_trials)) +
-  geom_bar(stat="identity", aes(fill = Block)) +
+  geom_bar(stat = "identity", aes(fill = Block)) +
   scale_x_discrete(expand = c(0, 0)) +
@@
   theme_modern() +
-  theme(axis.text.x = element_text(angle=45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 
-### Reaction Time Distribution
-
-```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=20}
-# RT distribution
-p <- estimate_density(df, select = "RT", at = c("Participant", "Block")) |>
-  group_by(Participant) |>
-  normalize(select = "y") |>
-  ungroup() |>
-  mutate(color = ifelse(Participant %in% outliers, "red", "blue")) |>
-  ggplot(aes(x = x, y = y)) +
-  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
-  geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block)) +
-  geom_vline(xintercept = 2500, linetype = "dashed", color = "red") +
-  scale_color_manual(values=c("red"="red", "blue"="blue"), guide = "none") +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  coord_cartesian(xlim = c(0, 3000)) +
-  theme_modern() +
-  theme(axis.text.y = element_blank()) +
-  facet_wrap(~Participant) +
-  labs(y = "", x = "Reaction Time (ms)")
-p
-# ggsave("figures/outliers.png", p, width=20, height=15)
-
-# Filter out
-df <- filter(df, !Participant %in% outliers)
-```
-
 ### Reaction Time per Trial
@@
 ```{r message=FALSE, warning=FALSE}
+df <- df |> 
+  group_by(Participant, Error) |> 
+  mutate(Outlier = ifelse(Error == 0 & (RT < 150 | standardize(RT) > 4), TRUE, FALSE)) |> 
+  ungroup()
+
 p1 <- estimate_density(df, select = "RT", at = "Participant") |>
@@
   ungroup() |>
+  merge(df |> 
+    filter(Error == 0) |> 
+    group_by(Participant) |> 
+    summarize(Outlier = mean(RT) + 4 * sd(RT))) |> 
+  mutate(Outlier = ifelse(x >= Outlier, TRUE, FALSE)) |> 
   ggplot(aes(x = x, y = y)) +
   geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
-  geom_line(aes(color = Participant, group = Participant)) +
-  geom_vline(xintercept = c(150, 3000), linetype = "dashed", color = "red") +
+  geom_line(aes(color = Participant, linetype=Outlier)) +
+  geom_vline(xintercept = c(150), linetype = "dashed", color = "red") +
   scale_color_material_d("rainbow", guide = "none") +
@@
   scale_y_continuous(expand = c(0, 0)) +
-  coord_cartesian(xlim = c(0, 3500)) +
+  guides(linetype = "none") +
+  coord_cartesian(xlim = c(0, 4000)) +
   theme_modern() +
   theme(axis.text.y = element_blank()) +
-  # facet_wrap(~Participant) +
   labs(y = "", x = "Reaction Time (ms)")
-
-df$Outlier <- df$RT < 150 | df$RT > 3000
 
@@
 
-We removed `r sum(df$Outlier)` (`r insight::format_value(sum(df$Outlier) / nrow(df), as_percent=TRUE)`) outlier trials (150 ms < RT < 3000 ms).
+We removed `r sum(df$Outlier)` (`r insight::format_value(sum(df$Outlier) / nrow(df), as_percent=TRUE)`) outlier trials (150 ms < RT < 4 SD above mean).
 
@@
 ```
+
+## Participants
+
+```{r message=FALSE, warning=FALSE}
+dfsub <- df |>
+  group_by(Participant) |>
+  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS) |>
+  slice(1) |>
+  ungroup()
+```
+
+The final sample included `r report::report_participants(dfsub, age="Age", sex="Sex")`.
+
+```{r }
+plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", fill = "orange") {
+  dfsub |>
+    ggplot(aes_string(x = what)) +
+    geom_density(fill = fill) +
+    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype = "dashed") +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_y_continuous(expand = c(0, 0)) +
+    ggtitle(title, subtitle = subtitle) +
+    theme_modern() +
+    theme(
+      plot.title = element_text(face = "bold", hjust = 0.5),
+      plot.subtitle = element_text(face = "italic", hjust = 0.5),
+      axis.title.x = element_blank(),
+      axis.title.y = element_blank(),
+      axis.text.y = element_blank()
+    )
+}
+
+plot_waffle <- function(dfsub, what = "Nationality", title = what) {
+  ggwaffle::waffle_iron(dfsub, what) |>
+    # mutate(label = emojifont::fontawesome('fa-twitter')) |>
+    ggplot(aes(x, y, fill = group)) +
+    ggwaffle::geom_waffle() +
+    # geom_point() +
+    # geom_text(aes(label=label), family='fontawesome-webfont', size=4) +
+    coord_equal() +
+    ggtitle(what) +
+    labs(fill = title) +
+    theme_void() +
+    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+}
+```
+
+```{r fig.width=20, fig.height=15}
+p1 <- plot_distribution(dfsub, "Age", fill = "#FF9800")
+p2 <- plot_distribution(dfsub, "Duration", title = "Total Duration", subtitle = "in minutes", fill = "#F44336")
+p3 <- plot_distribution(dfsub, "Break_Duration", title = "Break Duration", subtitle = "in minutes", fill = "#3F51B5")
+
+p4 <- plot_waffle(dfsub, "Sex") +
+  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))
+
+p5 <- plot_waffle(dfsub, "Education") +
+  scale_fill_viridis_d()
+
+p6 <- plot_waffle(dfsub, "Nationality") +
+  scale_fill_metro_d()
+
+p7 <- plot_waffle(dfsub, "Ethnicity") +
+  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0"))
+
+p8 <- plot_waffle(dfsub, "Screen_Resolution", title = "Screen Resolution") +
+  scale_fill_pizza_d()
+
+p9 <- plot_waffle(dfsub, "Device_OS", title = "Device OS") +
+  scale_fill_bluebrown_d()
+
+# p10 <- plot_waffle(dfsub, "Screen_Refresh") +
+#   scale_fill_viridis_d()
+
+(p1 / p2 / p3) | (p4 / p5 / p6) | (p7 / p8 / p9)
+```
+
+## Models {.tabset}
+
+### Delboeuf
+
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models <- function(data) {
+  models_err <- list()
+  models_rt <- list()
+  for (k1 in c("", "_log", "_sqrt", "_cbrt")) {
+    for (k2 in c("", "_log", "_sqrt", "_cbrt")) {
+      for (side in c("", "--SIDE")) {
+        name <- paste0("DIFF", k1, "--", "STRENGTH", k2, side)
+        message(name)
+        f <- paste0(
+          "Illusion_Difference",
+          k1,
+          " * Illusion_Strength",
+          k2,
+          " + (1|Participant)"
+        )
+
+        if (side == "--SIDE") f <- paste0("Illusion_Side * ", f)
+
+        m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)),
+          data = data, family = "binomial"
+        )
+        if (performance::check_convergence(m)) {
+          models_err[[name]] <- m
+        }
+
+        m <- glmmTMB::glmmTMB(as.formula(paste0("RT ~ ", f)),
+          data = data
+        )
+        if (performance::check_convergence(m)) {
+          models_rt[[name]] <- m
+        }
+      }
+    }
+  }
+
+  # RT
+  to_keep <- compare_performance(models_rt, metrics = c("BIC")) |>
+    arrange(BIC) |>
+    slice(1:5) |>
+    pull(Name)
+
+  test <- test_performance(models_rt[to_keep], reference = 1)
+  perf <- compare_performance(models_rt[to_keep], metrics = c("BIC", "R2"))
+  rt <- merge(perf, test) |>
+    arrange(BIC) |>
+    select(Name, BIC, R2_marginal, BF) |>
+    mutate(
+      BF = insight::format_bf(BF, name = ""),
+      Model = "RT"
+    )
+
+  # Errors
+  to_keep <- compare_performance(models_err, metrics = c("BIC")) |>
+    arrange(BIC) |>
+    slice(1:5) |>
+    pull(Name)
+
+  test <- test_performance(models_err[to_keep], reference = 1)
+  perf <- compare_performance(models_err[to_keep], metrics = c("BIC", "R2"))
+
+  merge(perf, test) |>
+    arrange(BIC) |>
+    select(Name, BIC, R2_marginal, BF) |>
+    mutate(
+      BF = insight::format_bf(BF, name = ""),
+      Model = "Errors"
+    ) |>
+    rbind(rt)
+}
+
+data <- filter(df, Illusion_Type == "Delboeuf")
+
+best_models(data)
+```
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_descriptive_err <- function(data, side = "leftright") {
+  # Sanity checks
+  if (length(unique(data$Illusion_Difference)) != 8) stop("Illusion_Difference values != 8")
+  if (length(unique(data$Illusion_Strength)) != 15) stop("Illusion_Strength values != 15")
+
+  if (side == "leftright") {
+    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
+    x <- tools::toTitleCase(gsub("arrow", "", x))
+    if (x == "Left") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
+    } else if (x == "Right") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
+    }
+  } else {
+    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
+    x <- tools::toTitleCase(gsub("arrow", "", x))
+    if (x == "Up") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
+    } else if (x == "Down") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
+    }
+    data$Answer <- fct_rev(data$Answer)
+  }
+
+  dodge1 <- 0.1 * diff(range(data$Illusion_Difference))
+  dodge2 <- -0.1 * diff(range(data$Illusion_Strength))
+
+  p1 <- data |>
+    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
+    summarize(Error = mean(Error), .groups = "drop") |>
+    ungroup() |>
+    ggplot(aes(x = Illusion_Difference, y = Error)) +
+    geom_bar(aes(fill = Illusion_Strength, group = Illusion_Strength), position = position_dodge(width = dodge1, preserve = "single"), stat = "identity", width=0.03) +
+    # geom_line(aes(color = Illusion_Strength, group=Illusion_Strength), position = position_dodge(width=dodge1)) +
+    geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.3) +
+    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_fill_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
+    theme_modern() +
+    labs(
+      color = "Illusion Strength",
+      fill = "Illusion Strength",
+      y = "Probability of Error",
+      x = "Task Difficulty"
+    ) +
+    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+
+  p2 <- data |>
+    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
+    summarize(Error = mean(Error), .groups = "drop") |>
+    ungroup() |>
+    # mutate(Illusion_Difference = as.factor(round(Illusion_Difference, 4))) |>
+    ggplot(aes(x = Illusion_Strength, y = Error)) +
+    geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.3) +
+    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
+    geom_bar(aes(fill = Illusion_Difference, group = Illusion_Difference), position = position_dodge(width = dodge2, preserve = "single"), stat = "identity", width=0.1) +
+    # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
+    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
+    theme_modern() +
+    labs(
+      color = "Task Difficulty",
+      fill = "Task Difficulty",
+      y = "Probability of Error",
+      x = "Illusion Strength"
+    ) +
+    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+
+  if (side == "leftright") {
+    p <- ((p1 + facet_wrap(~Answer, ncol = 2, labeller = "label_both")) /
+      (p2 + facet_wrap(~Answer, ncol = 2, labeller = "label_both"))) +
+      plot_annotation(
+        title = paste(data$Illusion_Type[1], "Illusion"),
+        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
+      )
+  } else {
+    p <- ((p1 + facet_wrap(~Answer, nrow = 2, labeller = "label_both")) |
+      (p2 + facet_wrap(~Answer, nrow = 2, labeller = "label_both"))) +
+      plot_annotation(
+        title = paste(data$Illusion_Type[1], "Illusion"),
+        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
+      )
+  }
+  p
+}
+
+plot_descriptive_err(data)
+```
+
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Difference * Illusion_Strength + 
+    (1 + (Illusion_Difference * Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_delboeuf_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model_err <- function(data, model) {
+  data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))
+
+  # Get variables
+  vars <- insight::find_predictors(model)$conditional
+  vardiff <- vars[1]
+  varstrength <- vars[2]
+
+  # Get predicted
+  pred <- estimate_relation(model,
+                            at = vars,
+                            length = c(NA, 25))
+
+  # Set colors for lines
+  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
+  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
+  names(colors) <- diffvals
+
+  # Assign color from the same palette to every observation of data (for geom_dots)
+  closest <- diffvals[max.col(-abs(outer(data[[vars[1]]], diffvals, "-")))]
+  data$color <- colors[as.character(closest)]
+  data$color <- fct_reorder(data$color, closest)
+
+  # Manual jittering
+  xrange <- 0.05*diff(range(data[[varstrength]]))
+  data$x <- data[[varstrength]]
+  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
+  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
+  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)
+
+  pred |>
+    ggplot(aes_string(x = varstrength, y = "Predicted")) +
+    geom_dots(
+      data = data,
+      aes(x=x, 
+          y = Error, 
+          group = Error,
+          side = .dots_side, 
+          # order=as.numeric(color) * ifelse(Error == 1, 1, -1)), 
+          order=color), 
+      fill = data$color,
+      color = NA, 
+      alpha=0.5) +
+    geom_slab(data=data, aes(y = Error)) +
+    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
+    geom_vline(xintercept = 0, linetype = "dashed") +
+    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
+    geom_line(aes_string(color = vardiff, group = vardiff)) +
+    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
+    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
+    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
+    theme_modern() +
+    labs(
+      title = paste0(data$Illusion_Type[1], " Illusion"),
+      color = "Difficulty", fill = "Difficulty",
+      y = "Probability of Error",
+      x = "Illusion Strength"
+    ) +
+    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+}
+
+plot_model_err(data, model_delboeuf_err)
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_descriptive_rt <- function(data, side = "leftright") {
+  if (side == "leftright") {
+    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
+    x <- tools::toTitleCase(gsub("arrow", "", x))
+    if (x == "Left") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
+    } else if (x == "Right") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
+    }
+  } else {
+    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
+    x <- tools::toTitleCase(gsub("arrow", "", x))
+    if (x == "Up") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
+    } else if (x == "Down") {
+      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
+    }
+    data$Answer <- fct_rev(data$Answer)
+  }
+
+  p1 <- data |>
+    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
+    summarize(RT = mean(RT), .groups = "drop") |>
+    ggplot(aes(x = Illusion_Difference, y = RT)) +
+    ggdist::stat_slabinterval(aes(fill = Illusion_Strength, group = Illusion_Strength, color = Illusion_Strength), alpha = 0.5, normalize = "groups") +
+    # geom_line(aes(color = Illusion_Strength), position = position_dodge(width=dodge1)) +
+    scale_y_continuous(expand = c(0, 0)) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_color_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
+    scale_fill_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
+    theme_modern() +
+    labs(
+      color = "Illusion Strength",
+      fill = "Illusion Strength",
+      y = "Reaction Time (ms)",
+      x = "Task Difficulty"
+    ) +
+    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+
+  p2 <- data |>
+    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
+    summarize(RT = mean(RT), .groups = "drop") |>
+    ggplot(aes(x = Illusion_Strength, y = RT)) +
+    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
+    ggdist::stat_slabinterval(aes(fill = Illusion_Difference, group = Illusion_Difference, color = Illusion_Difference), alpha = 0.5, normalize = "groups") +
+    # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
+    scale_y_continuous(expand = c(0, 0)) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
+    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
+    theme_modern() +
+    labs(
+      color = "Task Difficulty",
+      fill = "Task Difficulty",
+      y = "Reaction Time (ms)",
+      x = "Illusion Strength"
+    ) +
+    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+
+  if (side == "leftright") {
+    p <- ((p1 + facet_wrap(~Answer, ncol = 2, labeller = "label_both")) /
+      (p2 + facet_wrap(~Answer, ncol = 2, labeller = "label_both"))) +
+      plot_annotation(
+        title = paste(data$Illusion_Type[1], "Illusion"),
+        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
+      )
+  } else {
+    p <- ((p1 + facet_wrap(~Answer, nrow = 2, labeller = "label_both")) |
+      (p2 + facet_wrap(~Answer, nrow = 2, labeller = "label_both"))) +
+      plot_annotation(
+        title = paste(data$Illusion_Type[1], "Illusion"),
+        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
+      )
+  }
+  p
+}
+
+data <- filter(data, Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ Illusion_Difference * Illusion_Strength + 
+    (1 + (Illusion_Difference * Illusion_Strength) | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_delboeuf_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model_err <- function(data, model) {
+  data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))
+
+  # Get variables
+  vars <- insight::find_predictors(model)$conditional
+  vardiff <- vars[1]
+  varstrength <- vars[2]
+
+  # Get predicted
+  pred <- estimate_relation(model,
+                            at = vars,
+                            length = c(NA, 25))
+
+  # Set colors for lines
+  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
+  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
+  names(colors) <- diffvals
+
+  # Assign color from the same palette to every observation of data (for geom_dots)
+  closest <- diffvals[max.col(-abs(outer(data[[vars[1]]], diffvals, "-")))]
+  data$color <- colors[as.character(closest)]
+  data$color <- fct_reorder(data$color, closest)
+
+  # Manual jittering
+  xrange <- 0.05*diff(range(data[[varstrength]]))
+  data$x <- data[[varstrength]]
+  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
+  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
+  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)
+
+  pred |>
+    ggplot(aes_string(x = varstrength, y = "Predicted")) +
+    geom_dots(
+      data = data,
+      aes(x=x, 
+          y = Error, 
+          group = Error,
+          side = .dots_side, 
+          # order=as.numeric(color) * ifelse(Error == 1, 1, -1)), 
+          order=color), 
+      fill = data$color,
+      color = NA, 
+      alpha=0.5) +
+    geom_slab(data=data, aes(y = Error)) +
+    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
+    geom_vline(xintercept = 0, linetype = "dashed") +
+    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
+    geom_line(aes_string(color = vardiff, group = vardiff)) +
+    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
+    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
+    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
+    theme_modern() +
+    labs(
+      title = paste0(data$Illusion_Type[1], " Illusion"),
+      color = "Difficulty", fill = "Difficulty",
+      y = "Probability of Error",
+      x = "Illusion Strength"
+    ) +
+    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+}
+
+plot_model_err(data, model_delboeuf_err)
+```
+
+### Ebbinghaus
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ebbinghaus")
+
+plot_descriptive_err(data)
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ebbinghaus", Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+### Rod and Frame
+
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Rod-Frame")
+
+best_models(data)
+```
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_descriptive_err(data)
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(data, Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+### Vertical-Horizontal
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Vertical-Horizontal")
+
+plot_descriptive_err(data)
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+### Zöllner
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Zöllner")
+
+plot_descriptive_err(data)
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Zöllner", Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+### White
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "White")
+
+plot_descriptive_err(data)
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "White", Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+### Müller-Lyer
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Müller-Lyer")
+
+plot_descriptive_err(data, side = "updown")
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Müller-Lyer", Error == 0)
+
+plot_descriptive_rt(data, side = "updown")
+```
+
+### Ponzo
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ponzo")
+
+plot_descriptive_err(data, side = "updown")
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ponzo", Error == 0)
+
+plot_descriptive_rt(data, side = "updown")
+```
+
+### Poggendorff
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Poggendorff")
+
+plot_descriptive_err(data, side = "updown")
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Poggendorff", Error == 0)
+
+plot_descriptive_rt(data, side = "updown")
+```
+
+### Contrast
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Contrast")
+
+plot_descriptive_err(data, side = "updown")
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Contrast", Error == 0)
+
+plot_descriptive_rt(data, side = "updown")
+```
+
+## Individual Scores
+
+### Delboeuf
+
+#### Reliability
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+random <- model_delboeuf_err 
+random <- as.data.frame(model_delboeuf_err)
+random <- random[str_detect(names(random), "Participant\\[")]
+random <- random |> 
+  mutate(Draw = 1:nrow(random)) |>
+  pivot_longer(-Draw, names_to="Parameter", values_to = "Value") |> 
+  separate(Parameter, into = c("Participant", "Parameter"), sep = ",") |> 
+  mutate(Participant = str_remove(Participant, "r_Participant\\["),
+         Parameter = str_remove(Parameter, "]"))
+
+random |> 
+  ggplot(aes(y= Participant, x = Value)) +
+  ggdist::stat_slabinterval() +
+  facet_wrap(~Parameter)
+
+data <- random |> 
+  mutate(Parameter = paste0("Delboeuf_", Parameter)) |> 
+  group_by(Participant, Parameter) |> 
+  summarize(Score = as.numeric(bayestestR::map_estimate(Value)), .groups="drop") |>
+  ungroup() |> 
+  pivot_wider(names_from = Parameter, values_from = Score)
+
+```
+
+#### Approximation
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+dat <- data.frame()
+for(participant in unique(data$Participant)) {
+  freq <- glm(Error ~ Illusion_Difference * Illusion_Strength, 
+              family = "binomial",
+              data=filter(df, Participant == participant, 
+                          Illusion_Type == "Delboeuf"))
+
+  parameters::parameters(freq)[1:2] |> 
+    as.data.frame() |> 
+    mutate(Participant = participant, 
+           Parameter = paste0("freq_Delboeuf_", Parameter))
+}
+```
+
+## Structure
+
+## Variability

### 2022-08-05 preprocessing.R

@@ -13,8 +13,14 @@ preprocess_raw <- function(file) {
   # Info
   info <- data[data$screen == "browser_info" & !is.na(data$screen), ]
 
+  # Filter out practice trials
+  if ("practice_debrief" %in% data$screen) {
+    data <- data[which(data$screen == "practice_debrief"):nrow(data), ]
+  }
 
+  # Trial data
   trials <- data[data$screen == "Trial", ]
+
   df <- data.frame(
     Participant = trials$participant_id,
     Age = as.numeric(jsonlite::fromJSON(dem[1])$age),
@@ -50,13 +56,8 @@ preprocess_raw <- function(file) {
   # Correct duration
   df$Duration <- df$Duration - df$Break_Duration
 
-  # Transformation
-  df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
-  df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
-  df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1/3), 4)
-  df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
-  df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
-  df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))
+  # Manual fixes
+  df[df$Participant == "5d3c6e745602310001bca8aa_6bdtb", "Sex"] <- "Female"
 
   # Format names
   df$Illusion_Type <- ifelse(df$Illusion_Type == "MullerLyer", "Müller-Lyer", df$Illusion_Type)
@@ -76,9 +77,10 @@ preprocess_raw <- function(file) {
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latin"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White Middle European (Slavic)", "Greek"), "Caucasian", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White - Caucasian", " Caucasian"), "Caucasian", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("White", "Caucasian "), "Caucasian", df$Ethnicity)
[…]

### 2022-08-06 study2/experiment.js
add personality scales during break

@@ -427,3 +427,78 @@ const verticalhorizontal_instructions =
     "<p><img src='utils/answer/answer_leftright_keyboard.PNG' height='150'></img></p>" +
     "<p class='small'>In this example, the correct answer is the <b>LEFT arrow</b>.</p></div>" +
     "<p>Are you ready? <b>Press ENTER to start</b></p>"
+
+
+
+/* Psychometric scales---------------------------------------------------------------------*/
+
+// Mini IPIP scale
+var IPIP = [
+    "Am the life of the party",
+    "Sympathize with others' feelings",
+    "Get chores done right away",
+    "Have frequent mood swings",
+    "Have a vivid imagination",
+    "Do not talk a lot",
+    "Am not interested in other people's problems",
+    "Have difficulty understanding abstract ideas",
+    "Like order",
+    "Make a mess of things",
+    "Deserve more things in life",
+    "Do not have a good imagination",
+    "Feel other's emotions",
+    "Am relaxed most of the time",
+    "Get upset easily",
+    "Seldom feel blue",
+    "would like to be seen driving around in a very expensive car",
+    "Keep in the background",
+    "Am not really interested in others",
+    "Am not interested in abstract ideas",
+    "Often forget to put things back in their proper place",
+    "Talk to a lot of different people at parties",
+    "Would get a lot of pleasure from owning expensive luxury goods"
+]
+
+var IPIP_dim = [
+    "Ext_1", "Agr_2", "Con_3", "Neu_4", "Open_5",
+    "HH_6_R", "Ext_7_R", "Agr_8_R", "Open_9_R", "Con_10",
+    "Con_11_R", "HH_12_R", "Open_13_R", "Agr_14", "Neu_15_R",
+    "Neu_16", "Neu_17_R", "HH_18_R", "Ext_19_R", "Agr_20_R",
+    "Open_21_R", "Con_22_R", "Ext_23", "HH_24_R"
+]
+
+var PID = [
+    "People would describe me as reckless",
+    "I feel like I act totally on impulse",
+    "Even though I know better, I can't stop making rash decisions",
+    "I often feel like nothing I do really matters",
+    "Others see me as irresponsible",
+    "I'm not good at planning ahead",
+    "My thoughts often don't make sense to others",
+    "I worry about almost everything",
+    "I get emotional easily, often for very little reason",
+    "I fear being alone in life more than anything else",
+    "I get stuck on one way of doing things,even when it's clear it won't work",
[…]

### 2022-08-06 study2/experiment.js
format

@@ -75,7 +75,7 @@ var fixation = {
     type: jsPsychHtmlKeyboardResponse,
     stimulus: '<div style="font-size:60px;">+</div>',
     choices: "NO_KEYS" /* no responses will be accepted as a valid response */,
-    //trial_duration: 0, // (for testing)
+    // trial_duration: 0, // (for testing)
     trial_duration: function () {
         return randomInteger(500, 1000) // Function from RealityBending/JSmisc
     },
@@ -434,29 +434,29 @@ const verticalhorizontal_instructions =
 
 // Mini IPIP scale
 var IPIP = [
-    "Am the life of the party",
-    "Sympathize with others' feelings",
-    "Get chores done right away",
-    "Have frequent mood swings",
-    "Have a vivid imagination",
-    "Do not talk a lot",
-    "Am not interested in other people's problems",
-    "Have difficulty understanding abstract ideas",
-    "Like order",
-    "Make a mess of things",
-    "Deserve more things in life",
-    "Do not have a good imagination",
-    "Feel other's emotions",
-    "Am relaxed most of the time",
-    "Get upset easily",
-    "Seldom feel blue",
-    "would like to be seen driving around in a very expensive car",
-    "Keep in the background",
-    "Am not really interested in others",
-    "Am not interested in abstract ideas",
-    "Often forget to put things back in their proper place",
-    "Talk to a lot of different people at parties",
-    "Would get a lot of pleasure from owning expensive luxury goods"
+    "Am the life of the party<br>",
+    "Sympathize with others' feelings<br>",
+    "Get chores done right away<br>",
+    "Have frequent mood swings<br>",
+    "Have a vivid imagination<br>",
+    "Do not talk a lot<br>",
+    "Am not interested in other people's problems<br>",
+    "Have difficulty understanding abstract ideas<br>",
+    "Like order<br>",
+    "Make a mess of things<br>",
+    "Deserve more things in life<br>",
+    "Do not have a good imagination<br>",
+    "Feel other's emotions<br>",
+    "Am relaxed most of the time<br>",
+    "Get upset easily<br>",
+    "Seldom feel blue<br>",
+    "would like to be seen driving around in a very expensive car<br>",
+    "Keep in the background<br>",
+    "Am not really interested in others<br>",
+    "Am not interested in abstract ideas<br>",
+    "Often forget to put things back in their proper place<br>",
[…]

### 2022-08-06 manuscript.Rmd
update stim

---
title             : "**Something**"
shorttitle        : "Optimal Tolerance"
author: 
  - name          : "Dominique Makowski"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
    address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
    email         : "dmakowski@ntu.edu.sg"
    orcid         : 0000-0001-5375-9967

affiliation:
  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
authornote: |
  Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (dom.makowski@gmail.com).
abstract: |
  Abstract abstract.

keywords          : "chaos, complexity, fractal, physiology, tolerance"
wordcount         : "1156"
bibliography      : "bibliography.bib"
floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no
figurelist        : yes
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
csl: utils/apa.csl
header-includes:
  - \usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# options and parameters
options(digits = 3)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.width = see::golden_ratio(9),
  fig.height = 9,
  fig.path = "figures/"
)

cache <- TRUE
```

## Introduction

<!-- Intro  & Gaps-->
Visual illusions are blabla. Despite the historical interest, several gaps remained unanswered:

This could be caused by the relative difficulty to adapt visual illusions to an experimental setting, which typically requires the parametric modulation of variables at stake.

<!-- Target, strength, difficulty -->
Many visual illusions can be seen as made of *targets* (e.g., parallel lines, same-size circles, same-length lines), which perception is biased by the *context* (e.g., parallel lines are seen as angled, circle or segments appear of different lengths). To date, most of the paradigms prompt the subjective extent the context biases the perception ("how much does the circle appear bigger", e.g., **REFS**).

Although visual illusions are inherently tied to subjective perception, we tested a reversal of the traditional paradigm to access the "objective" effect of illusions.

To achieve this, we first developed a parametric framework to generate visual illusions that we implemented and made accessible in the open-source Pyllusion software (REF). This allows us to generate stimuli ...

In the present study, we will test this novel paradigm to ...

\newpage

## References

::: {#refs custom-style="Bibliography"}
:::

### 2022-08-06 study1/study1.Rmd
update stim

--- 2022-08-05 study1.Rmd
@@
 # (it contains the subject ID of the participants)
-participants <- list.files("study1/data/")
+participants <- list.files("data/")
 
@@
 for (ppt in participants) {
-  df <- rbind(df, preprocess_raw(file = paste0("study1/rawdata/", ppt)))
+  df <- rbind(df, preprocess_raw(file = paste0("data/", ppt)))
 }
@@
 # Save anonymized data
-write.csv(df, "data/study1.csv", row.names = FALSE)
+write.csv(df, "../data/study1.csv", row.names = FALSE)
 ```
@@
 
-df <- read.csv("data/study1.csv") |>
+df <- read.csv("../data/study1.csv") |>
   mutate(

### 2022-08-06 study2/study2.Rmd
update stim

--- 2022-08-05 study2.Rmd
@@
 # (it contains the subject ID of the participants)
-participants <- list.files("study2/data/")
+participants <- list.files("data/")
 
@@
 for (ppt in participants) {
-  df <- rbind(df, preprocess_raw(file = paste0("study2/data/", ppt)))
+  df <- rbind(df, preprocess_raw(file = paste0("data/", ppt)))
 }
@@
 # Save anonymized data
-write.csv(df, "data/study2.csv", row.names = FALSE)
+write.csv(df, "../data/study2.csv", row.names = FALSE)
 ```
@@
 
-df <- read.csv("data/study2.csv") |>
+df <- read.csv("../data/study2.csv") |>
   mutate(
@@
 
-### Model Visualization
+##### Model Visualization
 
@@
   # Assign color from the same palette to every observation of data (for geom_dots)
-  closest <- diffvals[max.col(-abs(outer(data[[vars[1]]], diffvals, "-")))]
+  closest <- diffvals[max.col(-abs(outer(data[[vardiff]], diffvals, "-")))]
   data$color <- colors[as.character(closest)]
@@
     geom_vline(xintercept = 0, linetype = "dashed") +
-    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
+    geom_hline(yintercept = c(0.5), linetype = "dotted", alpha = 0.5) +
     geom_line(aes_string(color = vardiff, group = vardiff)) +
@@
 
-plot_model_err(data, model_delboeuf_err)
+p_delboeuf_err <- plot_model_err(data, model_delboeuf_err)
+p_delboeuf_err
 ```
@@
 formula <- brms::bf(
-  Error ~ Illusion_Difference * Illusion_Strength + 
+  RT ~ Illusion_Difference * Illusion_Strength + 
     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
-  family = "bernoulli"
-)
+  family="shifted_lognormal"
+)
+
+# prior = c(
+#     prior(normal(6, 1.5), class = Intercept),
+#     prior(normal(0, 1), class = sigma),
+#     prior(normal(0, .01), class = b, coef = c_trial)
+#   )
 
@@
 
-model_delboeuf_err <- brms::brm(formula,
+model_delboeuf_rt <- brms::brm(formula,
   data = data,
@@
 
+modelbased::estimate_prediction(model_delboeuf_rt, keep_iterations = 10) |> 
+  as.data.frame() |> 
+  head()
+  bayestestR::reshape_iterations()
+nrow(as.data.frame(pred))
+performance::check_model(model_delboeuf_rt)
 # parameters::parameters(model)
@@
 
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model_err <- function(data, model) {
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model_rt <- function(data, model) {
   data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))
@@
 
-  # Manual jittering
-  xrange <- 0.05*diff(range(data[[varstrength]]))
-  data$x <- data[[varstrength]]
-  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
-  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
-  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)
-
-  pred |>
+  p <- pred |>
     ggplot(aes_string(x = varstrength, y = "Predicted")) +
-    geom_dots(
-      data = data,
-      aes(x=x, 
-          y = Error, 
-          group = Error,
-          side = .dots_side, 
-          # order=as.numeric(color) * ifelse(Error == 1, 1, -1)), 
-          order=color), 
-      fill = data$color,
-      color = NA, 
-      alpha=0.5) +
-    geom_slab(data=data, aes(y = Error)) +
+    ggdist::stat_slab(data=data, aes_string(y="RT", group = vardiff, fill=vardiff), alpha=0.5) +
     geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
     geom_vline(xintercept = 0, linetype = "dashed") +
-    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
     geom_line(aes_string(color = vardiff, group = vardiff)) +
-    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
+    scale_y_continuous(expand = c(0, 0)) +
     scale_x_continuous(expand = c(0, 0)) +
@@
     scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
+    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]])),
+                    ylim=c(150, 2000)) +
     theme_modern() +
@@
       color = "Difficulty", fill = "Difficulty",
-      y = "Probability of Error",
+      y = "Reaction Time (ms)",
       x = "Illusion Strength"
@@
     theme(plot.title = element_text(face = "bold", hjust = 0.5))
-}
-
-plot_model_err(data, model_delboeuf_err)
+
+}
+
+p_delboeuf_rt <- plot_model_rt(data, model_delboeuf_rt)
+p_delboeuf_rt
+```
+
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2}
+plot_all <- function(data, p_err, p_rt) {
+  # Get stimuli
+  dat <- df |>
+    filter(Error == 0) |> 
+    filter(Illusion_Type == unique(data$Illusion_Type),
+           Answer %in% c("arrowleft", "arrowup")) |>
+    select(Stimulus, Illusion_Strength, Illusion_Difference)
+
+  dat <- rbind(
+    filter(dat, Illusion_Strength == min(Illusion_Strength)) |> 
+    filter(Illusion_Difference %in% c(min(Illusion_Difference), max(Illusion_Difference))),
+    filter(dat, Illusion_Strength == max(Illusion_Strength)) |> 
+    filter(Illusion_Difference %in% c(min(Illusion_Difference), max(Illusion_Difference))),
+    filter(dat, Illusion_Difference == min(Illusion_Difference)) |> 
+    filter(Illusion_Strength %in% c(min(Illusion_Strength), max(Illusion_Strength))),
+    filter(dat, Illusion_Difference == max(Illusion_Difference)) |> 
+    filter(Illusion_Strength %in% c(min(Illusion_Strength), max(Illusion_Strength))))
+
+  img_leftdown <- filter(dat, Illusion_Difference == max(Illusion_Difference)) |> 
+    filter(Illusion_Strength == min(Illusion_Strength)) |> 
+    pull(Stimulus) |> 
+    unique()
+  img_rightdown <- filter(dat, Illusion_Difference == max(Illusion_Difference)) |> 
+    filter(Illusion_Strength == max(Illusion_Strength)) |> 
+    pull(Stimulus) |> 
+    unique()
+  img_leftup <- filter(dat, Illusion_Strength == min(Illusion_Strength)) |> 
+    filter(Illusion_Difference == min(Illusion_Difference)) |> 
+    pull(Stimulus) |> 
+    unique()
+  img_rightup <- filter(dat, Illusion_Strength == max(Illusion_Strength)) |> 
+    filter(Illusion_Difference == min(Illusion_Difference)) |> 
+    pull(Stimulus) |> 
+    unique()
+
+  img_leftdown <- paste0("study2/stimuli/", img_leftdown, ".png") |> 
+    png::readPNG() |> 
+    grid::rasterGrob(interpolate=TRUE) |> 
+    patchwork::wrap_elements()
+  img_rightdown <- paste0("study2/stimuli/", img_rightdown, ".png") |> 
+    png::readPNG() |> 
+    grid::rasterGrob(interpolate=TRUE) |> 
+    patchwork::wrap_elements()
+  img_leftup <- paste0("study2/stimuli/", img_leftup, ".png") |> 
+    png::readPNG() |>  
+    grid::rasterGrob(interpolate=TRUE) |> 
+    patchwork::wrap_elements()
+  img_rightup <- paste0("study2/stimuli/", img_rightup, ".png") |> 
+    png::readPNG() |> 
+    grid::rasterGrob(interpolate=TRUE) |> 
+    patchwork::wrap_elements()
+
+  p <- ((p_err + labs(x = "")) / (p_rt + labs(title = "")) + patchwork::plot_layout(guides="collect"))
+
+  # (
+  #   (img_leftup / patchwork::plot_spacer() / img_leftdown + patchwork::plot_layout(heights = c(0.5, 2, 0.5))) | 
+  #   (patchwork::plot_spacer() / p / patchwork::plot_spacer()  + patchwork::plot_layout(heights = c(0.5, 2, 0.5))) | 
+  #   (img_rightup / patchwork::plot_spacer() / img_rightdown  + patchwork::plot_layout(heights = c(0.5, 2, 0.5)))
+  #   ) + 
+  #   patchwork::plot_layout(widths = c(0.5, 1, 0.5))
+
+  (img_leftup | patchwork::plot_spacer() | img_rightup) / p / (img_leftdown | patchwork::plot_spacer() | img_rightdown) + 
+    patchwork::plot_layout(heights = c(0.5, 1.5, 0.5))
+}
+
+p_delboeuf <- plot_all(data, p_delboeuf_err, p_delboeuf_rt)
+p_delboeuf
 ```

### 2022-08-06 study1/preprocessing.R

@@ -1,0 +1,85 @@
+# Preprocessing Function
+preprocess_raw <- function(file) {
+  data <- read.csv(file)
+
+  if (!"final_results" %in% data$screen) {
+    print(paste0("Warning: Incomplete data for ", file))
+    return(data.frame())
+  }
+
+  # Demographics
+  dem <- data[data$screen == "demographics" & !is.na(data$screen), "response"]
+
+  # Info
+  info <- data[data$screen == "browser_info" & !is.na(data$screen), ]
+
+  # Trial data
+  trials <- data[data$screen == "Trial", ]
+
+  df <- data.frame(
+    Participant = trials$participant_id,
+    Age = as.numeric(jsonlite::fromJSON(dem[1])$age),
+    Sex = jsonlite::fromJSON(dem[2])$sex,
+    Education = jsonlite::fromJSON(dem[2])$education,
+    Nationality = tools::toTitleCase(jsonlite::fromJSON(dem[1])$nationality),
+    Ethnicity = tools::toTitleCase(jsonlite::fromJSON(dem[1])$ethnicity),
+    Date = ifelse(is.null(info$date), NA, info$date),
+    Time = ifelse(is.null(info$time), NA, info$time),
+    Duration = as.numeric(data[data$screen == "final_results", "time_elapsed"]) / 1000 / 60,
+    Break_Duration = as.numeric(data[data$screen == "break" & !is.na(data$screen), "rt"]) / 1000 / 60,
+    Screen_Resolution = paste0(trials$screen_width, "x", trials$screen_height),
+    Screen_Size = (as.numeric(trials$screen_width) / 1000) * (as.numeric(trials$screen_height) / 1000),
+    Screen_Refresh = trials$vsync_rate,
+    Browser = trials$browser,
+    Browser_Version = trials$browser_version,
+    Device = ifelse(trials$mobile == TRUE, "Mobile", "Desktop"),
+    Device_OS = trials$os,
+    Illusion_Type = trials$type,
+    Block = ifelse(trials$block_number > 10, 2, 1),
+    Block_Order = as.numeric(trials$block_number),
+    Trial = as.numeric(trials$trial_number),
+    Stimulus = gsub(".png", "", gsub("stimuli/", "", trials$stimulus)),
[…]

### 2022-08-06 study2/make_stimuli.py

@@ -55,9 +55,9 @@ def generate_images(data, strengths, differences, function, name="Delboeuf", **k
             path = (
                 name
                 + "_str"
-                + str(np.round(strength, 4))
+                + str(np.round(strength, 5))
                 + "_diff"
-                + str(np.round(difference, 4))
+                + str(np.round(difference, 5))
                 + ".png"
             )
             img.save("stimuli/" + path)
@@ -175,13 +175,12 @@ data_block1 = generate_images(
 
 data_block2 = generate_images(
     data_block2,
-    strengths=strengths[0::2],
+    strengths=np.concatenate((strengths[0::2], [0])),
     differences=diff2,
     function=ill.Delboeuf,
     name="Delboeuf",
     distance=0.9,  # Distance between circles
 )
-
 
 # -------------------------- Ebbinghaus Illusion --------------------------
 ill.Ebbinghaus(illusion_strength=-1.4, difference=1.4).to_image(width=800, height=600).save(
@@ -216,7 +215,7 @@ data_block1 = generate_images(
 
 data_block2 = generate_images(
     data_block2,
-    strengths=strengths[0::2],
+    strengths=np.concatenate((strengths[0::2], [0])),
     differences=diff2,
     function=ill.Ebbinghaus,
     name="Ebbinghaus",
@@ -254,7 +253,7 @@ data_block1 = generate_images(
 
 data_block2 = generate_images(
     data_block2,
-    strengths=strengths[0::2],
+    strengths=np.concatenate((strengths[0::2], [0])),
     differences=diff2,
     function=ill.RodFrame,
     name="RodFrame",
@@ -289,7 +288,7 @@ data_block1 = generate_images(
 
 data_block2 = generate_images(
     data_block2,
-    strengths=strengths[0::2],
+    strengths=np.concatenate((strengths[0::2], [0])),
     differences=diff2,
     function=ill.VerticalHorizontal,
     name="VerticalHorizontal",
@@ -325,7 +324,7 @@ data_block1 = generate_images(
 
 data_block2 = generate_images(
     data_block2,
-    strengths=strengths[0::2],
+    strengths=np.concatenate((strengths[0::2], [0])),
     differences=diff2,
     function=ill.Zollner,
     name="Zollner",
@@ -362,7 +361,7 @@ data_block1 = generate_images(
 
 data_block2 = generate_images(
     data_block2,
-    strengths=strengths[0::2],
[…]

### 2022-08-06 study2/experiment.js
update dimensions

@@ -41,7 +41,7 @@ function get_debrief_display(results, type = "Block") {
         var score =
             "<p>Your score for this illusion is " +
             '<p style="color: black; font-size: 48px; font-weight: bold;">' +
-            Math.round(results.score*10)/10 +
+            Math.round(results.score * 10) / 10 +
             " %</p>"
     } else if (type === "Final") {
         // Final debriefing at end of game
@@ -86,17 +86,28 @@ var fixation = {
 }
 
 // Break
-var make_break = {
+var make_break1 = {
     type: jsPsychHtmlButtonResponse,
     choices: ["I am ready to continue!"],
     stimulus:
         "<p><b>CONGRATULATIONS!</b></p>" +
         "<p>You have finished half of the game. We know it's long and challenging, so we appreciate you staying focused until the end!</p>" +
-        "<p>You will now see all the illusions once again. <b>Try to beat your previous score!</b></p>",
+        "<p>Before you see all the illusions once again, let's take a break by answering a few questions about yourself.</p>",
     save_trial_parameters: {
         trial_duration: true,
     },
-    data: { screen: "break" },
+    data: { screen: "break1" },
+}
+var make_break2 = {
+    type: jsPsychHtmlButtonResponse,
+    choices: ["I am ready to continue!"],
+    stimulus:
+        "<p><b>Back to the illusions</b></p>" +
+        "<p>Try to improve your previous score!</p>",
+    save_trial_parameters: {
+        trial_duration: true,
+    },
+    data: { screen: "break2" },
 }
 
 // Marker
@@ -428,43 +439,60 @@ const verticalhorizontal_instructions =
     "<p class='small'>In this example, the correct answer is the <b>LEFT arrow</b>.</p></div>" +
     "<p>Are you ready? <b>Press ENTER to start</b></p>"
 
-
-
 /* Psychometric scales---------------------------------------------------------------------*/
 
 // Mini IPIP scale
 var IPIP = [
-    "Am the life of the party<br>",
-    "Sympathize with others' feelings<br>",
-    "Get chores done right away<br>",
-    "Have frequent mood swings<br>",
-    "Have a vivid imagination<br>",
-    "Do not talk a lot<br>",
[…]

### 2022-08-06 study2/study2.Rmd
Update study2.Rmd

@@
 ```{r, echo = FALSE, warning=FALSE, message=FALSE}
-options(digits = 3,
+options(
+  digits = 3,
   mc.cores = 4,
   brms.algorithm = "sampling",
-        brms.backend = "cmdstanr")
+  brms.backend = "cmdstanr"
+)
 
@@
 df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
-df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1/3), 4)
+df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1 / 3), 4)
 df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
 df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
-df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))
+df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1 / 3))
 
@@
   geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block)) +
-  geom_vline(xintercept = 150, linetype = "dashed", color = "red") +
+  geom_vline(xintercept = 125, linetype = "dashed", color = "red") +
   scale_color_manual(values = c("red" = "red", "blue" = "blue"), guide = "none") +
@@
   group_by(Participant, Error) |>
-  mutate(Outlier = ifelse(Error == 0 & (RT < 150 | standardize(RT) > 4), TRUE, FALSE)) |> 
+  mutate(Outlier = ifelse(Error == 0 & (RT < 125 | standardize(RT) > 4), TRUE, FALSE)) |>
   ungroup()
@@
   geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
-  geom_line(aes(color = Participant, linetype=Outlier)) +
-  geom_vline(xintercept = c(150), linetype = "dashed", color = "red") +
+  geom_line(aes(color = Participant, linetype = Outlier)) +
+  geom_vline(xintercept = c(125), linetype = "dashed", color = "red") +
   scale_color_material_d("rainbow", guide = "none") +
@@
 
-We removed `r sum(df$Outlier)` (`r insight::format_value(sum(df$Outlier) / nrow(df), as_percent=TRUE)`) outlier trials (150 ms < RT < 4 SD above mean).
+We removed `r sum(df$Outlier)` (`r insight::format_value(sum(df$Outlier) / nrow(df), as_percent=TRUE)`) outlier trials (125 ms < RT < 4 SD above mean).
 
@@
     ggplot(aes(x = Illusion_Difference, y = Error)) +
-    geom_bar(aes(fill = Illusion_Strength, group = Illusion_Strength), position = position_dodge(width = dodge1, preserve = "single"), stat = "identity", width=0.03) +
+    geom_bar(aes(fill = Illusion_Strength, group = Illusion_Strength), position = position_dodge(width = dodge1, preserve = "single"), stat = "identity", width = 0.03) +
     # geom_line(aes(color = Illusion_Strength, group=Illusion_Strength), position = position_dodge(width=dodge1)) +
@@
     geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
-    geom_bar(aes(fill = Illusion_Difference, group = Illusion_Difference), position = position_dodge(width = dodge2, preserve = "single"), stat = "identity", width=0.1) +
+    geom_bar(aes(fill = Illusion_Difference, group = Illusion_Difference), position = position_dodge(width = dodge2, preserve = "single"), stat = "identity", width = 0.1) +
     # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
@@
     at = vars,
-                            length = c(NA, 25))
+    length = c(NA, 25)
+  )
 
@@
   # Manual jittering
-  xrange <- 0.05*diff(range(data[[varstrength]]))
+  xrange <- 0.05 * diff(range(data[[varstrength]]))
   data$x <- data[[varstrength]]
@@
   data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
-  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)
+  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange / 2, xrange / 2)
 
@@
       data = data,
-      aes(x=x, 
+      aes(
+        x = x,
         y = Error,
@@
         # order=as.numeric(color) * ifelse(Error == 1, 1, -1)),
-          order=color), 
+        order = color
+      ),
       fill = data$color,
       color = NA,
-      alpha=0.5) +
-    geom_slab(data=data, aes(y = Error)) +
+      alpha = 0.5
+    ) +
+    geom_slab(data = data, aes(y = Error)) +
     geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
@@
     scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
+    coord_cartesian(xlim = c(min(data[[varstrength]]), max(data[[varstrength]]))) +
     theme_modern() +
@@
   RT ~ Illusion_Difference * Illusion_Strength +
-    (1 + (Illusion_Difference * Illusion_Strength) | Participant),
-  family="shifted_lognormal"
+    (1 + (Illusion_Difference * Illusion_Strength) | G | Participant),
+  sigma ~ Illusion_Difference * Illusion_Strength +
+    (1 + (Illusion_Difference * Illusion_Strength) | G | Participant),
+  family = "shifted_lognormal"
 )
@@
 
-modelbased::estimate_prediction(model_delboeuf_rt, keep_iterations = 10) |> 
-  as.data.frame() |> 
-  head()
-  bayestestR::reshape_iterations()
-nrow(as.data.frame(pred))
-performance::check_model(model_delboeuf_rt)
+plot_ppcheck <- function(model) {
+  pred <- modelbased::estimate_prediction(model_delboeuf_rt, keep_iterations = 100)
+  estimate_density(pred$RT) |>
+    ggplot(aes(x = x, y = y)) +
+    geom_area(fill = "#9E9E9E") +
+    geom_line(
+      data = pred |>
+        bayestestR::reshape_iterations() |>
+        mutate(iter_group = as.factor(iter_group)) |>
+        estimate_density(select = "iter_value", at = "iter_group"),
+      aes(group = iter_group), color = "#2196F3", size = 0.1, alpha = 0.5
+    ) +
+    scale_y_continuous(expand = c(0, 0)) +
+    scale_x_continuous(expand = c(0, 0)) +
+    coord_cartesian(xlim = c(100, 3000)) +
+    theme_modern() +
+    labs(x = "Reaction Time (ms)", y = "", title = "Posterior Predictive Check") +
+    theme(
+      plot.title = element_text(face = "bold", hjust = 0.5),
+      axis.text.y = element_blank()
+    )
+}
+
+plot_ppcheck(model_delboeuf_rt)
+# performance::check_model(model_delboeuf_rt)
 # parameters::parameters(model)
@@
     ggplot(aes_string(x = varstrength, y = "Predicted")) +
-    ggdist::stat_slab(data=data, aes_string(y="RT", group = vardiff, fill=vardiff), alpha=0.5) +
+    ggdist::stat_slab(data = data, aes_string(y = "RT", group = vardiff, fill = vardiff), alpha = 0.5) +
     geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
@@
     scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]])),
-                    ylim=c(150, 2000)) +
+    coord_cartesian(
+      xlim = c(min(data[[varstrength]]), max(data[[varstrength]])),
+      ylim = c(125, 2000)
+    ) +
     theme_modern() +
@@
     theme(plot.title = element_text(face = "bold", hjust = 0.5))
-
 }
@@
     filter(Error == 0) |>
-    filter(Illusion_Type == unique(data$Illusion_Type),
-           Answer %in% c("arrowleft", "arrowup")) |>
+    filter(
+      Illusion_Type == unique(data$Illusion_Type),
+      Answer %in% c("arrowleft", "arrowup")
+    ) |>
     select(Stimulus, Illusion_Strength, Illusion_Difference)
@@
     filter(dat, Illusion_Difference == max(Illusion_Difference)) |>
-    filter(Illusion_Strength %in% c(min(Illusion_Strength), max(Illusion_Strength))))
+      filter(Illusion_Strength %in% c(min(Illusion_Strength), max(Illusion_Strength)))
+  )
 
@@
 
-  img_leftdown <- paste0("study2/stimuli/", img_leftdown, ".png") |> 
+  img_leftdown <- paste0("stimuli/", img_leftdown, ".png") |>
     png::readPNG() |>
-    grid::rasterGrob(interpolate=TRUE) |> 
+    grid::rasterGrob(interpolate = TRUE) |>
     patchwork::wrap_elements()
-  img_rightdown <- paste0("study2/stimuli/", img_rightdown, ".png") |> 
+  img_rightdown <- paste0("stimuli/", img_rightdown, ".png") |>
     png::readPNG() |>
-    grid::rasterGrob(interpolate=TRUE) |> 
+    grid::rasterGrob(interpolate = TRUE) |>
     patchwork::wrap_elements()
-  img_leftup <- paste0("study2/stimuli/", img_leftup, ".png") |> 
+  img_leftup <- paste0("stimuli/", img_leftup, ".png") |>
     png::readPNG() |>
-    grid::rasterGrob(interpolate=TRUE) |> 
+    grid::rasterGrob(interpolate = TRUE) |>
     patchwork::wrap_elements()
-  img_rightup <- paste0("study2/stimuli/", img_rightup, ".png") |> 
+  img_rightup <- paste0("stimuli/", img_rightup, ".png") |>
     png::readPNG() |>
-    grid::rasterGrob(interpolate=TRUE) |> 
+    grid::rasterGrob(interpolate = TRUE) |>
     patchwork::wrap_elements()
 
-  p <- ((p_err + labs(x = "")) / (p_rt + labs(title = "")) + patchwork::plot_layout(guides="collect"))
+  p <- ((p_err + labs(x = "")) / (p_rt + labs(title = "")) + patchwork::plot_layout(guides = "collect"))
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-random <- model_delboeuf_err 
-random <- as.data.frame(model_delboeuf_err)
-random <- random[str_detect(names(random), "Participant\\[")]
+extract_random <- function(model, illusion = "Delboeuf") {
+  random <- as.data.frame(model_delboeuf_rt)
+  random <- random[str_detect(names(random), regex("^r_Participant"))]
+
+  if (insight::model_info(model)$is_logit) {
+    param <- "Prob,"
+  } else {
+    param <- "Loc,"
+  }
+
+  random |>
+    mutate(Draw = 1:nrow(random)) |>
+    pivot_longer(-Draw, names_to = "Parameter", values_to = "Value") |>
+    mutate(
+      Parameter = str_remove(Parameter, "r_Participant"),
+      Parameter = str_remove(Parameter, "\\]"),
+      Parameter = str_remove(Parameter, "\\["),
+      Parameter = paste0(param, Parameter),
+      Parameter = str_replace(Parameter, paste0(param, "__sigma"), "Disp,")
+    ) |>
+    separate(Parameter, into = c("Component", "Participant", "Parameter"), sep = ",") |>
+    mutate(
+      Parameter = paste0(Parameter, "_", Component),
+      Illusion_Type = illusion
+    ) |>
+    mutate(
+      Parameter = str_replace(Parameter, "Illusion_Strength", "Illu"),
+      Parameter = str_replace(Parameter, "Illusion_Difference", "Diff")
+    )
+}
+
+random <- rbind(
+  extract_random(model_delboeuf_rt, "Delboeuf"),
+  extract_random(model_delboeuf_err, "Delboeuf")
+)
+
+random |>
+  ggplot(aes(y = Participant, x = Value)) +
+  ggdist::stat_slabinterval(aes(fill = Illusion_Type)) +
+  facet_wrap(~Parameter, scales = "free_x")
+
 random <- random |>
-  mutate(Draw = 1:nrow(random)) |>
-  pivot_longer(-Draw, names_to="Parameter", values_to = "Value") |> 
-  separate(Parameter, into = c("Participant", "Parameter"), sep = ",") |> 
-  mutate(Participant = str_remove(Participant, "r_Participant\\["),
-         Parameter = str_remove(Parameter, "]"))
-
-random |> 
-  ggplot(aes(y= Participant, x = Value)) +
-  ggdist::stat_slabinterval() +
-  facet_wrap(~Parameter)
-
-data <- random |> 
-  mutate(Parameter = paste0("Delboeuf_", Parameter)) |> 
+  mutate(Parameter = paste0(Illusion_Type, "_", Parameter)) |>
   group_by(Participant, Parameter) |>
-  summarize(Score = as.numeric(bayestestR::map_estimate(Value)), .groups="drop") |>
+  summarize(Score = as.numeric(bayestestR::map_estimate(Value)), .groups = "drop") |>
   ungroup() |>
   pivot_wider(names_from = Parameter, values_from = Score)
-
 ```
@@
 dat <- data.frame()
-for(participant in unique(data$Participant)) {
+for (participant in unique(data$Participant)) {
   freq <- glm(Error ~ Illusion_Difference * Illusion_Strength,
     family = "binomial",
-              data=filter(df, Participant == participant, 
-                          Illusion_Type == "Delboeuf"))
+    data = filter(
+      df, Participant == participant,
+      Illusion_Type == "Delboeuf"
+    )
+  )
 
@@
     as.data.frame() |>
-    mutate(Participant = participant, 
-           Parameter = paste0("freq_Delboeuf_", Parameter))
+    mutate(
+      Participant = participant,
+      Parameter = paste0("freq_Delboeuf_", Parameter)
+    )
 }
@@
 
-## Variability
+### Correlation
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+clean_name <- function(x) {
+  x |>
+    str_replace("_Loc", " (loc)") |>
+    str_replace("_Disp", " (disp)") |>
+    str_replace("_Prob", " (prob)") |>
+    str_replace("_", " - ")
+}
+
+cor <- correlation::correlation(random, redundant = TRUE)
+
+p_data <- cor |>
+  mutate(
+    Parameter1 = clean_name(Parameter1),
+    Parameter2 = clean_name(Parameter2)
+  ) |>
+  correlation::cor_sort(hclust_method = "ward.D2") |>
+  cor_lower() |>
+  mutate(
+    Text = insight::format_value(r, zap_small = TRUE, digits = 3),
+    Text = str_replace(str_remove(Text, "^0+"), "^-0+", "-"),
+    Parameter2 = fct_rev(Parameter2)
+  )
+
+p_data |>
+  ggplot(aes(x = Parameter2, y = Parameter1)) +
+  geom_tile(aes(fill = r)) +
+  # geom_text(aes(label = Text), size = 2) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  scale_x_discrete(expand = c(0, 0)) +
+  scale_y_discrete(expand = c(0, 0)) +
+  labs(title = "Correlation Matrix", x = NULL, y = NULL) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 90, hjust = 1),
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank()
+  )
+```
+
+### Factor Analysis
+
+#### First Level
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- select(random, -Participant)
+
+rez <- parameters::n_factors(data)
+plot(rez)
+
+fa <- parameters::factor_analysis(data,
+  cor = as.matrix(cor),
+  n = 3,
+  rotation = "oblimin",
+  sort = TRUE
+)
+
+# psych::omega(data, nfactors=3)
+# fa <- psych::fa.multi(as.matrix(cor), nfactors = 3, nfact2 = 1, n.obs = nrow(random))
+# psych::fa.multi.diagram(fa)
+```
+
+#### Second Level
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- predict(fa)
+
+rez <- parameters::n_factors(data, rotation = "oblimin")
+plot(rez)
+
+fa2 <- parameters::factor_analysis(data,
+  n = 1,
+  sort = TRUE
+)
+
+# psych::schmid(attributes(fa)$model, nfactors = 1)
+```
+
+#### Visualization
+```{r message=FALSE, warning=FALSE, cache=cache}
+library(ggraph)
+
+dat <- rbind(
+  attributes(fa)$loadings_long,
+  attributes(fa2)$loadings_long |>
+    mutate(Component = str_replace(Component, "MR1", "G"))
+) |>
+  tidygraph::as_tbl_graph() |>
+  tidygraph::activate("nodes")
+
+# Get order of indices
+idx <- correlation::cor_sort(cor, hclust_method = "ward.D2") |>
+  summary() |>
+  pull(Parameter)
+
+dat |>
+  mutate(
+    x = sapply(as.data.frame(dat)$name, function(x) {
+      if (x %in% idx) {
+        (which(idx == x) - 1) / (length(idx) - 1)
+      } else {
+        NA
+      }
+    })
+  ) |>
+  mutate(
+    x = case_when(
+    name == "G" ~ 0.5,
+    name == "MR1" ~ 0,
+    name == "MR2" ~ 0.5,
+    name == "MR3" ~ 1,
+    TRUE ~ x),
+    y = case_when(
+    name == "G" ~ 2,
+    name == "MR1" ~ 1,
+    name == "MR2" ~ 1,
+    name == "MR3" ~ 1,
+    TRUE ~ 0)) |> 
+  ggraph(layout = "nicely") +
+  geom_edge_link(aes(edge_width = Loading, edge_colour = as.factor(sign(Loading)))) +
+  geom_node_point() +
+  geom_node_text(aes(label = name)) +
+  guides(edge_colour = "none") +
+  ggraph::theme_graph()
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+# Save
+dfsub <- dfsub |> 
+  merge(random, by = "Participant") |> 
+  cbind(data, predict(fa2, names="G"))
+
+write.csv("../data/study3.csv")
+```

### 2022-08-07 study2/study2.Rmd
script

@@
   mc.cores = 4,
-  brms.algorithm = "sampling",
+  brms.algorithm = "meanfield",
   brms.backend = "cmdstanr"
@@
 # Dear participant, thank you for participating in our study.
+Unfortunately, we didn't receive your data :( did something happen?
+Some technical issue?
+We would like to kindly ask you to return your participation so that we can open up more slots.
+Thank you in advance, and apologies for the inconvenience!
+
+# Dear participant, thank you for participating in our study.
 Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers), which makes it unusable.
@@
 
+# Just received the results: in your case, the three most prominent issues were your response pattern that was equivalent to random (in the sequence of keystrokes) that led to a chance-level performance (that was also significantly different from the rest of the population).
+Moreover, your reaction time distribution was also very different from the norm, with a vast majority of implausibly short responses (i.e., that are faster than the time it takes the brain to process a visual input).
+This issue was even more prominent in the second block (after the break), which typically happens when participants are in a rush to finish.
+Finally, your overall completion time was also significantly below the average.
+Again, we apologize, we understand that your time is valuable, but unfortunately we run too on limited funds and can hardly spare more on unusable data.
+Sorry for the inconvenience, we will try to open-up more slots soon.
+
 outliers <- c(
   # Error rate of 48.8% Very short RT
-  # Prolific Status: NA
+  # Prolific Status: REJECTED (06/08)
   "5e66ed53de7896486f9a71f8_p8ckk"
@@
     RT_SD = sd(RT),
+    IPIP6_SD = mean(IPIP6_SD),
+    PID5_SD = mean(PID5_SD),
   ) |>
@@
 
-knitr::kable(dfsub) |>
-  kableExtra::row_spec(which(dfsub$Participant %in% outliers), background = "#EF9A9A")
+data.frame(Participant = c("Total"), t(sapply(dfsub[2:ncol(dfsub)], mean, na.rm=TRUE))) |> 
+  rbind(dfsub) |> 
+  knitr::kable() |>
+  kableExtra::row_spec(1, italic = TRUE) |> 
+  kableExtra::row_spec(which(dfsub$Participant %in% outliers) + 1, background = "#EF9A9A")
 ```
@@
 
+# Drop also participant with bad block
+df <- filter(
+  df,
+  !(Participant == "5c43cd414fe4f800016e4983_0zp37" & df$Block == 2))
+
 rm(temp, temp2)
@@
   group_by(Participant) |>
-  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS) |>
+  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS, starts_with("IPIP6"), starts_with("PID5")) |>
   slice(1) |>
@@
     coord_equal() +
-    ggtitle(what) +
-    labs(fill = title) +
+    ggtitle(title) +
+    labs(fill = "") +
     theme_void() +
@@
       for (side in c("", "--SIDE")) {
+        for (effect in c("", "--EFFECT")) {
           name <- paste0("DIFF", k1, "--", "STRENGTH", k2, side)
@@
     }
+  }
 
@@
     ggplot(aes(x = Illusion_Difference, y = Error)) +
-    geom_bar(aes(fill = Illusion_Strength, group = Illusion_Strength), position = position_dodge(width = dodge1, preserve = "single"), stat = "identity", width = 0.03) +
+    geom_bar(aes(fill = Illusion_Strength, group = Illusion_Strength), position = position_dodge(width = dodge1, preserve = "single"), stat = "identity", width = dodge1) +
     # geom_line(aes(color = Illusion_Strength, group=Illusion_Strength), position = position_dodge(width=dodge1)) +
@@
     geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
-    geom_bar(aes(fill = Illusion_Difference, group = Illusion_Difference), position = position_dodge(width = dodge2, preserve = "single"), stat = "identity", width = 0.1) +
+    geom_bar(aes(fill = Illusion_Difference, group = Illusion_Difference), position = position_dodge(width = dodge2, preserve = "single"), stat = "identity", width = dodge2) +
     # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
+logmod <- function(x) sign(x) * log(1 + abs(x))
+
 formula <- brms::bf(
-  Error ~ Illusion_Difference * Illusion_Strength +
-    (1 + (Illusion_Difference * Illusion_Strength) | Participant),
+  Error ~ logmod(1 + Illusion_Difference) * logmod(1 + Illusion_Strength) +
+    (1 | Participant),
   family = "bernoulli"
@@
     ) +
-    geom_slab(data = data, aes(y = Error)) +
+    # geom_slab(data = data, aes(y = Error)) +
     geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
@@
   p1 <- data |>
-    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
-    summarize(RT = mean(RT), .groups = "drop") |>
     ggplot(aes(x = Illusion_Difference, y = RT)) +
-    ggdist::stat_slabinterval(aes(fill = Illusion_Strength, group = Illusion_Strength, color = Illusion_Strength), alpha = 0.5, normalize = "groups") +
+    ggdist::stat_slabinterval(aes(fill = Illusion_Strength, group = Illusion_Strength, color = Illusion_Strength), alpha = 0.5, normalize = "groups", position="dodge") +
     # geom_line(aes(color = Illusion_Strength), position = position_dodge(width=dodge1)) +
@@
     scale_fill_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
+    coord_cartesian(ylim = c(125, 3000)) +
     theme_modern() +
@@
   p2 <- data |>
-    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
-    summarize(RT = mean(RT), .groups = "drop") |>
     ggplot(aes(x = Illusion_Strength, y = RT)) +
     geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
-    ggdist::stat_slabinterval(aes(fill = Illusion_Difference, group = Illusion_Difference, color = Illusion_Difference), alpha = 0.5, normalize = "groups") +
+    ggdist::stat_slabinterval(aes(fill = Illusion_Difference, group = Illusion_Difference, color = Illusion_Difference), alpha = 0.5, normalize = "groups", position="dodge") +
     # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
@@
     scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
+    coord_cartesian(ylim = c(125, 3000)) +
     theme_modern() +
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
+# formula <- brms::bf(
+#   RT ~ Illusion_Difference * Illusion_Strength +
+#     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
+#   sigma ~ Illusion_Difference * Illusion_Strength +
+#     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
+#   # ndt ~ Illusion_Difference * Illusion_Strength +
+#   #   (1 | G | Participant),
+#   family = "shifted_lognormal"
+# )
+# formula <- brms::bf(
+#   RT ~ Illusion_Difference * Illusion_Strength +
+#     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
+#   sigma ~ Illusion_Difference * Illusion_Strength +
+#     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
+#   # beta ~ Illusion_Difference * Illusion_Strength +
+#   #   (1 + (Illusion_Difference * Illusion_Strength) | Participant),
+#   family = "exgaussian"
+# )
 formula <- brms::bf(
-  RT ~ Illusion_Difference * Illusion_Strength +
-    (1 + (Illusion_Difference * Illusion_Strength) | G | Participant),
-  sigma ~ Illusion_Difference * Illusion_Strength +
-    (1 + (Illusion_Difference * Illusion_Strength) | G | Participant),
-  family = "shifted_lognormal"
-)
-
+  RT ~ t2(Illusion_Difference, Illusion_Strength, k=4) +
+    (1 | Participant)
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  # family = "exgaussian"
+)
 # prior = c(

### 2022-08-07 study2/preprocessing.R

@@ -18,6 +18,7 @@ preprocess_raw <- function(file) {
     data <- data[which(data$screen == "practice_debrief"):nrow(data), ]
   }
 
+
   # Trial data
   trials <- data[data$screen == "Trial", ]
 
@@ -31,7 +32,7 @@ preprocess_raw <- function(file) {
     Date = ifelse(is.null(info$date), NA, info$date),
     Time = ifelse(is.null(info$time), NA, info$time),
     Duration = as.numeric(data[data$screen == "final_results", "time_elapsed"]) / 1000 / 60,
-    Break_Duration = as.numeric(data[data$screen == "break" & !is.na(data$screen), "rt"]) / 1000 / 60,
+    Break_Duration = as.numeric(data[data$screen %in% c("break", "break2") & !is.na(data$screen), "rt"]) / 1000 / 60,
     Screen_Resolution = paste0(trials$screen_width, "x", trials$screen_height),
     Screen_Size = (as.numeric(trials$screen_width) / 1000) * (as.numeric(trials$screen_height) / 1000),
     Screen_Refresh = trials$vsync_rate,
@@ -45,6 +46,7 @@ preprocess_raw <- function(file) {
     Trial = as.numeric(trials$trial_number),
     Stimulus = gsub(".png", "", gsub("stimuli/", "", trials$stimulus)),
     Illusion_Strength = as.numeric(trials$illusion_strength),
+    Illusion_Effect = as.factor(sign(as.numeric(trials$illusion_strength))),
     Illusion_Side = as.factor(sign(as.numeric(trials$illusion_difference))),
     Illusion_Difference = abs(as.numeric(trials$illusion_difference)),
     Answer = trials$response,
@@ -53,11 +55,50 @@ preprocess_raw <- function(file) {
     RT = as.numeric(trials$rt)
   )
 
+
+  if("IPIP6" %in% data$screen) {
+    # IPIP6
+    ipip6 <- as.data.frame(jsonlite::fromJSON(data[data$screen == "IPIP6", "response"]))
+    ipip6[grepl("_R", names(ipip6))] <- 100 - ipip6[grepl("_R", names(ipip6))]
+    df$IPIP6_Extraversion <- rowMeans(ipip6[grepl("Extraversion", names(ipip6))])
[…]

### 2022-08-07 study2/study2.Rmd
model

@@
 cache <- FALSE
+runModels <- FALSE
 fig.width <- see::golden_ratio(7)
@@
 df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"])
+df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"] <- round(df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"], 5)
 
@@
 library(brms)
+source("functions.R")
 
@@
     Illusion_Side = as.factor(Illusion_Side),
-    Block = as.factor(Block)
-    # Education = fct_relevel(Education, "Master", "Bachelor", "High School", "Other")
+    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
+    Block = as.factor(Block),
+    Education = fct_relevel(Education, "Doctorate", "Master", "Bachelor", "High School", "Other")
   )
@@
 We understand that you might have been in a hurry or had some other issues; we hope to open-up more slots in the future would you be interested to participate again.
+
+# Dear participant, thank you for participating in our study.
+Unfortunately, our system detected multiple issues in your data (such as implausibly short responses, and random-like pattern of answers - in particular in the 2nd part of the study), which makes it unusable.
+We understand that you might have been in a hurry or had some other issues, and so we kindly ask you to return your participation; we hope to open-up more slots in the future would you be interested to participate again.
 
@@
   # Prolific Status: REJECTED (06/08)
-  "5e66ed53de7896486f9a71f8_p8ckk"
-)
+  "5e66ed53de7896486f9a71f8_p8ckk",
+  # 2nd block of responses very fast
+  # Prolific Status: RETURN REQUESTED (07/08)
+  "61443e47c248a375c899a0cf_9qz3u"
+)
+
+partial_outliers <- c(
+  # 2nd block a bit bad
+  "5c43cd414fe4f800016e4983_0zp37", 
+  # Entire 2nd block bad
+  "61564e378e974bdbe42763a2_hhufm")
 ```
@@
   kableExtra::row_spec(1, italic = TRUE) |> 
-  kableExtra::row_spec(which(dfsub$Participant %in% outliers) + 1, background = "#EF9A9A")
+  kableExtra::row_spec(which(dfsub$Participant %in% outliers) + 1, background = "#EF9A9A") |> 
+  kableExtra::row_spec(which(dfsub$Participant %in% partial_outliers) + 1, background = "#FFECB3")
 ```
@@
 # RT distribution
-p <- estimate_density(df, select = "RT", at = c("Participant", "Block")) |>
+p <- df |> 
+  filter(RT < 10000) |> 
+  estimate_density(select = "RT", at = c("Participant", "Block")) |>
   group_by(Participant) |>
@@
   ungroup() |>
-  mutate(color = ifelse(Participant %in% outliers, "red", "blue")) |>
+  mutate(color = ifelse(Participant %in% outliers, "red", ifelse(Participant %in% partial_outliers, "orange", "blue"))) |>
   ggplot(aes(x = x, y = y)) +
-  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
+  geom_area(data = normalize(estimate_density(filter(df, RT < 10000), select = "RT"), select = "y"), alpha = 0.2) +
   geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block)) +
   geom_vline(xintercept = 125, linetype = "dashed", color = "red") +
-  scale_color_manual(values = c("red" = "red", "blue" = "blue"), guide = "none") +
+  scale_color_manual(values = c("red" = "#F44336", "orange" = "#FF9800", "blue" = "blue"), guide = "none") +
   scale_x_continuous(expand = c(0, 0)) +
@@
 
-# Drop also participant with bad block
+# Drop also participant with bad second block
 df <- filter(
   df,
-  !(Participant == "5c43cd414fe4f800016e4983_0zp37" & df$Block == 2))
+  !(Participant %in% partial_outliers & df$Block == 2))
 
@@
 df <- filter(df, Outlier == FALSE)
+df$RT <- df$RT / 1000  # Convert to second for better model convergence
 ```
@@
 p8 <- plot_waffle(dfsub, "Screen_Resolution", title = "Screen Resolution") +
-  scale_fill_pizza_d()
+  scale_fill_pizza_d() +
+  guides(fill = "none")
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-best_models <- function(data) {
-  models_err <- list()
-  models_rt <- list()
-  for (k1 in c("", "_log", "_sqrt", "_cbrt")) {
-    for (k2 in c("", "_log", "_sqrt", "_cbrt")) {
-      for (side in c("", "--SIDE")) {
-        for (effect in c("", "--EFFECT")) {
-          name <- paste0("DIFF", k1, "--", "STRENGTH", k2, side)
-          message(name)
-          f <- paste0(
-            "Illusion_Difference",
-            k1,
-            " * Illusion_Strength",
-            k2,
-            " + (1|Participant)"
-          )
-
-          if (side == "--SIDE") f <- paste0("Illusion_Side * ", f)
-
-          m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)),
-            data = data, family = "binomial"
-          )
-          if (performance::check_convergence(m)) {
-            models_err[[name]] <- m
-          }
-
-          m <- glmmTMB::glmmTMB(as.formula(paste0("RT ~ ", f)),
-            data = data
-          )
-          if (performance::check_convergence(m)) {
-            models_rt[[name]] <- m
-          }
-        }
-      }
-    }
-  }
-
-  # RT
-  to_keep <- compare_performance(models_rt, metrics = c("BIC")) |>
-    arrange(BIC) |>
-    slice(1:5) |>
-    pull(Name)
-
-  test <- test_performance(models_rt[to_keep], reference = 1)
-  perf <- compare_performance(models_rt[to_keep], metrics = c("BIC", "R2"))
-  rt <- merge(perf, test) |>
-    arrange(BIC) |>
-    select(Name, BIC, R2_marginal, BF) |>
-    mutate(
-      BF = insight::format_bf(BF, name = ""),
-      Model = "RT"
-    )
-
-  # Errors
-  to_keep <- compare_performance(models_err, metrics = c("BIC")) |>
-    arrange(BIC) |>
-    slice(1:5) |>
-    pull(Name)
-
-  test <- test_performance(models_err[to_keep], reference = 1)
-  perf <- compare_performance(models_err[to_keep], metrics = c("BIC", "R2"))
-
-  merge(perf, test) |>
-    arrange(BIC) |>
-    select(Name, BIC, R2_marginal, BF) |>
-    mutate(
-      BF = insight::format_bf(BF, name = ""),
-      Model = "Errors"
-    ) |>
-    rbind(rt)
-}
-
 data <- filter(df, Illusion_Type == "Delboeuf")
-
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 best_models(data)
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-plot_descriptive_err <- function(data, side = "leftright") {
-  # Sanity checks
-  if (length(unique(data$Illusion_Difference)) != 8) stop("Illusion_Difference values != 8")
-  if (length(unique(data$Illusion_Strength)) != 15) stop("Illusion_Strength values != 15")
-
-  if (side == "leftright") {
-    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
-    x <- tools::toTitleCase(gsub("arrow", "", x))
-    if (x == "Left") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
-    } else if (x == "Right") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
-    }
-  } else {
-    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
-    x <- tools::toTitleCase(gsub("arrow", "", x))
-    if (x == "Up") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
-    } else if (x == "Down") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
-    }
-    data$Answer <- fct_rev(data$Answer)
-  }
-
-  dodge1 <- 0.1 * diff(range(data$Illusion_Difference))
-  dodge2 <- -0.1 * diff(range(data$Illusion_Strength))
-
-  p1 <- data |>
-    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
-    summarize(Error = mean(Error), .groups = "drop") |>
-    ungroup() |>
-    ggplot(aes(x = Illusion_Difference, y = Error)) +
-    geom_bar(aes(fill = Illusion_Strength, group = Illusion_Strength), position = position_dodge(width = dodge1, preserve = "single"), stat = "identity", width = dodge1) +
-    # geom_line(aes(color = Illusion_Strength, group=Illusion_Strength), position = position_dodge(width=dodge1)) +
-    geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.3) +
-    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_fill_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
-    theme_modern() +
-    labs(
-      color = "Illusion Strength",
-      fill = "Illusion Strength",
-      y = "Probability of Error",
-      x = "Task Difficulty"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-
-  p2 <- data |>
-    group_by(Illusion_Difference, Illusion_Strength, Answer) |>
-    summarize(Error = mean(Error), .groups = "drop") |>
-    ungroup() |>
-    # mutate(Illusion_Difference = as.factor(round(Illusion_Difference, 4))) |>
-    ggplot(aes(x = Illusion_Strength, y = Error)) +
-    geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.3) +
-    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
-    geom_bar(aes(fill = Illusion_Difference, group = Illusion_Difference), position = position_dodge(width = dodge2, preserve = "single"), stat = "identity", width = dodge2) +
-    # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
-    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    theme_modern() +
-    labs(
-      color = "Task Difficulty",
-      fill = "Task Difficulty",
-      y = "Probability of Error",
-      x = "Illusion Strength"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-
-  if (side == "leftright") {
-    p <- ((p1 + facet_wrap(~Answer, ncol = 2, labeller = "label_both")) /
-      (p2 + facet_wrap(~Answer, ncol = 2, labeller = "label_both"))) +
-      plot_annotation(
-        title = paste(data$Illusion_Type[1], "Illusion"),
-        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
-      )
-  } else {
-    p <- ((p1 + facet_wrap(~Answer, nrow = 2, labeller = "label_both")) |
-      (p2 + facet_wrap(~Answer, nrow = 2, labeller = "label_both"))) +
-      plot_annotation(
-        title = paste(data$Illusion_Type[1], "Illusion"),
-        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
-      )
-  }
-  p
-}
-
 plot_descriptive_err(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-logmod <- function(x) sign(x) * log(1 + abs(x))
-
-formula <- brms::bf(
-  Error ~ logmod(1 + Illusion_Difference) * logmod(1 + Illusion_Strength) +
-    (1 | Participant),
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect | Participant),
   family = "bernoulli"
@@
 
-# parameters::parameters(model)
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model_err <- function(data, model) {
-  data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))
-
-  # Get variables
-  vars <- insight::find_predictors(model)$conditional
-  vardiff <- vars[1]
-  varstrength <- vars[2]
-
-  # Get predicted
-  pred <- estimate_relation(model,
-    at = vars,
-    length = c(NA, 25)
-  )
-
-  # Set colors for lines
-  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
-  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
-  names(colors) <- diffvals
-
-  # Assign color from the same palette to every observation of data (for geom_dots)
-  closest <- diffvals[max.col(-abs(outer(data[[vardiff]], diffvals, "-")))]
-  data$color <- colors[as.character(closest)]
-  data$color <- fct_reorder(data$color, closest)
-
-  # Manual jittering
-  xrange <- 0.05 * diff(range(data[[varstrength]]))
-  data$x <- data[[varstrength]]
-  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
-  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
-  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange / 2, xrange / 2)
-
-  pred |>
-    ggplot(aes_string(x = varstrength, y = "Predicted")) +
-    geom_dots(
-      data = data,
-      aes(
-        x = x,
-        y = Error,
-        group = Error,
-        side = .dots_side,
-        # order=as.numeric(color) * ifelse(Error == 1, 1, -1)),
-        order = color
-      ),
-      fill = data$color,
-      color = NA,
-      alpha = 0.5
-    ) +
-    # geom_slab(data = data, aes(y = Error)) +
-    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
-    geom_vline(xintercept = 0, linetype = "dashed") +
-    geom_hline(yintercept = c(0.5), linetype = "dotted", alpha = 0.5) +
-    geom_line(aes_string(color = vardiff, group = vardiff)) +
-    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    coord_cartesian(xlim = c(min(data[[varstrength]]), max(data[[varstrength]]))) +
-    theme_modern() +
-    labs(
-      title = paste0(data$Illusion_Type[1], " Illusion"),
-      color = "Difficulty", fill = "Difficulty",
-      y = "Probability of Error",
-      x = "Illusion Strength"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-}
-
+# parameters::parameters(model_delboeuf_err)
+
+save(model_delboeuf_err, file="models/model_delboeuf_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_delboeuf_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
 p_delboeuf_err <- plot_model_err(data, model_delboeuf_err)
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-plot_descriptive_rt <- function(data, side = "leftright") {
-  if (side == "leftright") {
-    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
-    x <- tools::toTitleCase(gsub("arrow", "", x))
-    if (x == "Left") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Left", "Right")
-    } else if (x == "Right") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Right", "Left")
-    }
-  } else {
-    x <- data[data$Error == 0 & data$Illusion_Side == 1, ]$Answer[1]
-    x <- tools::toTitleCase(gsub("arrow", "", x))
-    if (x == "Up") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Up", "Down")
-    } else if (x == "Down") {
-      data$Answer <- ifelse(data$Illusion_Side == 1, "Down", "Up")
-    }
-    data$Answer <- fct_rev(data$Answer)
-  }
-
-  p1 <- data |>
-    ggplot(aes(x = Illusion_Difference, y = RT)) +
-    ggdist::stat_slabinterval(aes(fill = Illusion_Strength, group = Illusion_Strength, color = Illusion_Strength), alpha = 0.5, normalize = "groups", position="dodge") +
-    # geom_line(aes(color = Illusion_Strength), position = position_dodge(width=dodge1)) +
-    scale_y_continuous(expand = c(0, 0)) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_color_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
-    scale_fill_gradientn(colours = c("#4CAF50", "#009688", "#00BCD4", "#2196F3", "#3F51B5", "#673AB7", "#9C27B0")) +
-    coord_cartesian(ylim = c(125, 3000)) +
-    theme_modern() +
-    labs(
-      color = "Illusion Strength",
-      fill = "Illusion Strength",
-      y = "Reaction Time (ms)",
-      x = "Task Difficulty"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-
-  p2 <- data |>
-    ggplot(aes(x = Illusion_Strength, y = RT)) +
-    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
-    ggdist::stat_slabinterval(aes(fill = Illusion_Difference, group = Illusion_Difference, color = Illusion_Difference), alpha = 0.5, normalize = "groups", position="dodge") +
-    # geom_line(aes(color = Illusion_Difference), position = position_dodge(width=dodge2)) +
-    scale_y_continuous(expand = c(0, 0)) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    coord_cartesian(ylim = c(125, 3000)) +
-    theme_modern() +
-    labs(
-      color = "Task Difficulty",
-      fill = "Task Difficulty",
-      y = "Reaction Time (ms)",
-      x = "Illusion Strength"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-
-  if (side == "leftright") {
-    p <- ((p1 + facet_wrap(~Answer, ncol = 2, labeller = "label_both")) /
-      (p2 + facet_wrap(~Answer, ncol = 2, labeller = "label_both"))) +
-      plot_annotation(
-        title = paste(data$Illusion_Type[1], "Illusion"),
-        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
-      )
-  } else {
-    p <- ((p1 + facet_wrap(~Answer, nrow = 2, labeller = "label_both")) |
-      (p2 + facet_wrap(~Answer, nrow = 2, labeller = "label_both"))) +
-      plot_annotation(
-        title = paste(data$Illusion_Type[1], "Illusion"),
-        theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))
-      )
-  }
-  p
-}
-
-data <- filter(data, Error == 0)
+
+data <- filter(df, Illusion_Type == "Delboeuf", Error == 0)
 
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
 # formula <- brms::bf(
@@
 # )
-formula <- brms::bf(
-  RT ~ t2(Illusion_Difference, Illusion_Strength, k=4) +
-    (1 | Participant)
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
-  # family = "exgaussian"
+
+formula <- brms::bf(
+  RT ~ logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  sigma ~ 1 + (1 | Participant),
+  family = "exgaussian"
 )
@@
 
-plot_ppcheck <- function(model) {
-  pred <- modelbased::estimate_prediction(model_delboeuf_rt, keep_iterations = 100)
-  estimate_density(pred$RT) |>
-    ggplot(aes(x = x, y = y)) +
-    geom_area(fill = "#9E9E9E") +
-    geom_line(
-      data = pred |>
-        bayestestR::reshape_iterations() |>
-        mutate(iter_group = as.factor(iter_group)) |>
-        estimate_density(select = "iter_value", at = "iter_group"),
-      aes(group = iter_group), color = "#2196F3", size = 0.1, alpha = 0.5
-    ) +
-    scale_y_continuous(expand = c(0, 0)) +
-    scale_x_continuous(expand = c(0, 0)) +
-    coord_cartesian(xlim = c(100, 3000)) +
-    theme_modern() +
-    labs(x = "Reaction Time (ms)", y = "", title = "Posterior Predictive Check") +
-    theme(
-      plot.title = element_text(face = "bold", hjust = 0.5),
-      axis.text.y = element_blank()
-    )
-}
-
 plot_ppcheck(model_delboeuf_rt)
@@
 # parameters::parameters(model)
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model_rt <- function(data, model) {
-  data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))
-
-  # Get variables
-  vars <- insight::find_predictors(model)$conditional
-  vardiff <- vars[1]
-  varstrength <- vars[2]
-
-  # Get predicted
-  pred <- estimate_relation(model,
-    at = vars,
-    length = c(NA, 25)
-  )
-
-  # Set colors for lines
-  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
-  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
-  names(colors) <- diffvals
-
-  # Assign color from the same palette to every observation of data (for geom_dots)
-  closest <- diffvals[max.col(-abs(outer(data[[vardiff]], diffvals, "-")))]
-  data$color <- colors[as.character(closest)]
-  data$color <- fct_reorder(data$color, closest)
-
-  p <- pred |>
-    ggplot(aes_string(x = varstrength, y = "Predicted")) +
-    ggdist::stat_slab(data = data, aes_string(y = "RT", group = vardiff, fill = vardiff), alpha = 0.5) +
-    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
-    geom_vline(xintercept = 0, linetype = "dashed") +
-    geom_line(aes_string(color = vardiff, group = vardiff)) +
-    scale_y_continuous(expand = c(0, 0)) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
-    coord_cartesian(
-      xlim = c(min(data[[varstrength]]), max(data[[varstrength]])),
-      ylim = c(125, 2000)
-    ) +
-    theme_modern() +
-    labs(
-      title = paste0(data$Illusion_Type[1], " Illusion"),
-      color = "Difficulty", fill = "Difficulty",
-      y = "Reaction Time (ms)",
-      x = "Illusion Strength"
-    ) +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-}
-
+
+save(model_delboeuf_rt, file="models/model_delboeuf_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_delboeuf_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
 p_delboeuf_rt <- plot_model_rt(data, model_delboeuf_rt)
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2}
-plot_all <- function(data, p_err, p_rt) {
-  # Get stimuli
-  dat <- df |>
-    filter(Error == 0) |>
-    filter(
-      Illusion_Type == unique(data$Illusion_Type),
-      Answer %in% c("arrowleft", "arrowup")
-    ) |>
-    select(Stimulus, Illusion_Strength, Illusion_Difference)
-
-  dat <- rbind(
-    filter(dat, Illusion_Strength == min(Illusion_Strength)) |>
-      filter(Illusion_Difference %in% c(min(Illusion_Difference), max(Illusion_Difference))),
-    filter(dat, Illusion_Strength == max(Illusion_Strength)) |>
-      filter(Illusion_Difference %in% c(min(Illusion_Difference), max(Illusion_Difference))),
-    filter(dat, Illusion_Difference == min(Illusion_Difference)) |>
-      filter(Illusion_Strength %in% c(min(Illusion_Strength), max(Illusion_Strength))),
-    filter(dat, Illusion_Difference == max(Illusion_Difference)) |>
-      filter(Illusion_Strength %in% c(min(Illusion_Strength), max(Illusion_Strength)))
-  )
-
-  img_leftdown <- filter(dat, Illusion_Difference == max(Illusion_Difference)) |>
-    filter(Illusion_Strength == min(Illusion_Strength)) |>
-    pull(Stimulus) |>
-    unique()
-  img_rightdown <- filter(dat, Illusion_Difference == max(Illusion_Difference)) |>
-    filter(Illusion_Strength == max(Illusion_Strength)) |>
-    pull(Stimulus) |>
-    unique()
-  img_leftup <- filter(dat, Illusion_Strength == min(Illusion_Strength)) |>
-    filter(Illusion_Difference == min(Illusion_Difference)) |>
-    pull(Stimulus) |>
-    unique()
-  img_rightup <- filter(dat, Illusion_Strength == max(Illusion_Strength)) |>
-    filter(Illusion_Difference == min(Illusion_Difference)) |>
-    pull(Stimulus) |>
-    unique()
-
-  img_leftdown <- paste0("stimuli/", img_leftdown, ".png") |>
-    png::readPNG() |>
-    grid::rasterGrob(interpolate = TRUE) |>
-    patchwork::wrap_elements()
-  img_rightdown <- paste0("stimuli/", img_rightdown, ".png") |>
-    png::readPNG() |>
-    grid::rasterGrob(interpolate = TRUE) |>
-    patchwork::wrap_elements()
-  img_leftup <- paste0("stimuli/", img_leftup, ".png") |>
-    png::readPNG() |>
-    grid::rasterGrob(interpolate = TRUE) |>
-    patchwork::wrap_elements()
-  img_rightup <- paste0("stimuli/", img_rightup, ".png") |>
-    png::readPNG() |>
-    grid::rasterGrob(interpolate = TRUE) |>
-    patchwork::wrap_elements()
-
-  p <- ((p_err + labs(x = "")) / (p_rt + labs(title = "")) + patchwork::plot_layout(guides = "collect"))
-
-  # (
-  #   (img_leftup / patchwork::plot_spacer() / img_leftdown + patchwork::plot_layout(heights = c(0.5, 2, 0.5))) |
-  #   (patchwork::plot_spacer() / p / patchwork::plot_spacer()  + patchwork::plot_layout(heights = c(0.5, 2, 0.5))) |
-  #   (img_rightup / patchwork::plot_spacer() / img_rightdown  + patchwork::plot_layout(heights = c(0.5, 2, 0.5)))
-  #   ) +
-  #   patchwork::plot_layout(widths = c(0.5, 1, 0.5))
-
-  (img_leftup | patchwork::plot_spacer() | img_rightup) / p / (img_leftdown | patchwork::plot_spacer() | img_rightdown) +
-    patchwork::plot_layout(heights = c(0.5, 1.5, 0.5))
-}
-
 p_delboeuf <- plot_all(data, p_delboeuf_err, p_delboeuf_rt)
@@
 
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ebbinghaus") 
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+best_models(data)
+```
+
 #### Error Rate
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Ebbinghaus")
-
 plot_descriptive_err(data)
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_ebbinghaus_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_delboeuf_err)
+
+save(model_ebbinghaus_err, file="models/model_ebbinghaus_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_ebbinghaus_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_ebbinghaus_err <- plot_model_err(data, model_ebbinghaus_err)
+p_ebbinghaus_err
+```
+
 #### Reaction Time
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  family = "exgaussian"
+)
+
+model_ebbinghaus_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_ebbinghaus_rt)
+# performance::check_model(model_delboeuf_rt)
+# parameters::parameters(model)
+
+save(model_ebbinghaus_rt, file="models/model_ebbinghaus_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_ebbinghaus_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_ebbinghaus_rt <- plot_model_rt(data, model_ebbinghaus_rt)
+p_ebbinghaus_rt
+```
+
 ### Rod and Frame
@@
 data <- filter(df, Illusion_Type == "Rod-Frame")
-
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 best_models(data)
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_rodframe_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_delboeuf_err)
+
+save(model_rodframe_err, file="models/model_rodframe_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_rodframe_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_rodframe_err <- plot_model_err(data, model_rodframe_err)
+p_rodframe_err
+```
+
 #### Reaction Time
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(data, Error == 0)
+data <- filter(df, Illusion_Type == "Rod-Frame", Error == 0)
 
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  family = "exgaussian"
+)
+
+model_rodframe_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_rodframe_rt)
+# performance::check_model(model_delboeuf_rt)
+# parameters::parameters(model)
+
+save(model_rodframe_rt, file="models/model_rodframe_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_rodframe_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_rodframe_rt <- plot_model_rt(data, model_rodframe_rt)
+p_rodframe_rt
+```
+
 ### Vertical-Horizontal
 
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Vertical-Horizontal") 
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+best_models(data)
+```
+
 #### Error Rate
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Vertical-Horizontal")
-
 plot_descriptive_err(data)
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_verticalhorizontal_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_verticalhorizontal_err)
+
+save(model_verticalhorizontal_err, file="models/model_verticalhorizontal_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_verticalhorizontal_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_verticalhorizontal_err <- plot_model_err(data, model_verticalhorizontal_err)
+p_verticalhorizontal_err
+```
+
 #### Reaction Time
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  family = "exgaussian"
+)
+
+model_verticalhorizontal_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_verticalhorizontal_rt)
+# performance::check_model(model_verticalhorizontal_rt)
+# parameters::parameters(model)
+
+save(model_verticalhorizontal_rt, file="models/model_verticalhorizontal_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_verticalhorizontal_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_verticalhorizontal_rt <- plot_model_rt(data, model_verticalhorizontal_rt)
+p_verticalhorizontal_rt
+```
+
 ### Zöllner
 
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Zöllner")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+best_models(data)
+```
+
 #### Error Rate
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Zöllner")
-
 plot_descriptive_err(data)
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ logmod(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_zollner_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_zollner_err)
+
+save(model_zollner_err, file="models/model_zollner_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_zollner_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_zollner_err <- plot_model_err(data, model_zollner_err)
+p_zollner_err
+```
+
 #### Reaction Time
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ logmod(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  family = "exgaussian"
+)
+
+model_zollner_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_zollner_rt)
+# performance::check_model(model_zollner_rt)
+# parameters::parameters(model)
+
+save(model_zollner_rt, file="models/model_zollner_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_zollner_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_zollner_rt <- plot_model_rt(data, model_zollner_rt)
+p_zollner_rt
+```
+
 ### White
 
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "White")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+best_models(data)
+```
+
 #### Error Rate
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "White")
-
 plot_descriptive_err(data)
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_white_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_white_err)
+
+save(model_white_err, file="models/model_white_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_white_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_white_err <- plot_model_err(data, model_white_err)
+p_white_err
+```
+
 #### Reaction Time
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  family = "exgaussian"
+)
+
+model_white_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_white_rt)
+# performance::check_model(model_white_rt)
+# parameters::parameters(model)
+
+save(model_white_rt, file="models/model_white_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_white_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_white_rt <- plot_model_rt(data, model_white_rt)
+p_white_rt
+```
+
 ### Müller-Lyer
 
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Müller-Lyer")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+best_models(data)
+```
+
 #### Error Rate
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Müller-Lyer")
-
 plot_descriptive_err(data, side = "updown")
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ logmod(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + logmod(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_mullerlyer_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_white_err)
+
+save(model_mullerlyer_err, file="models/model_mullerlyer_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_mullerlyer_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_mullerlyer_err <- plot_model_err(data, model_mullerlyer_err)
+p_mullerlyer_err
+```
+
 #### Reaction Time
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  family = "exgaussian"
+)
+
+model_mullerlyer_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_mullerlyer_rt)
+# performance::check_model(model_mullerlyer_rt)
+# parameters::parameters(model)
+
+save(model_mullerlyer_rt, file="models/model_mullerlyer_rt.Rdata")
+# parameters::parameters(model_white_err)
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_mullerlyer_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_mullerlyer_rt <- plot_model_rt(data, model_mullerlyer_rt)
+p_mullerlyer_rt
+```
+
 ### Ponzo
 
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ponzo")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+best_models(data)
+```
+
 #### Error Rate
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Ponzo")
-
 plot_descriptive_err(data, side = "updown")
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_ponzo_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_ponzo_err)
+
+save(model_ponzo_err, file="models/model_ponzo_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_ponzo_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_ponzo_err <- plot_model_err(data, model_ponzo_err)
+p_ponzo_err
+```
+
 #### Reaction Time
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  family = "exgaussian"
+)
+
+model_ponzo_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_ponzo_rt)
+# performance::check_model(model_ponzo_rt)
+# parameters::parameters(model)
+
+save(model_ponzo_rt, file="models/model_ponzo_rt.Rdata")
+# parameters::parameters(model_white_err)
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_ponzo_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_ponzo_rt <- plot_model_rt(data, model_ponzo_rt)
+p_ponzo_rt
+```
+
 ### Poggendorff
 
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Poggendorff")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+best_models(data)
+```
+
 #### Error Rate
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Poggendorff")
-
 plot_descriptive_err(data, side = "updown")
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_poggendorff_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_poggendorff_err)
+
+save(model_poggendorff_err, file="models/model_poggendorff_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_poggendorff_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_poggendorff_err <- plot_model_err(data, model_poggendorff_err)
+p_poggendorff_err
+```
+
 #### Reaction Time
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ 1 + (1 | Participant),
+  family = "exgaussian"
+)
+
+model_poggendorff_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_poggendorff_rt)
+# performance::check_model(model_poggendorff_rt)
+# parameters::parameters(model)
+
+save(model_ponzo_rt, file="models/model_poggendorff_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_poggendorff_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_poggendorff_rt <- plot_model_rt(data, model_poggendorff_rt)
+p_poggendorff_rt
+```
+
 ### Contrast
 
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Contrast")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+best_models(data)
+```
+
 #### Error Rate
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_contrast_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_contrast_err)
+
+save(model_contrast_err, file="models/model_contrast_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_contrast_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_contrast_err <- plot_model_err(data, model_contrast_err)
+p_contrast_err
+```
+
 #### Reaction Time
@@
 
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ 1 + (1 | Participant),
+  family = "exgaussian"
+)
+
+model_contrast_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_contrast_rt)
+# performance::check_model(model_contrast_rt)
+# parameters::parameters(model)
+
+save(model_contrast_rt, file="models/model_contrast_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_contrast_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_contrast_rt <- plot_model_rt(data, model_contrast_rt)
+p_contrast_rt
+```
+
 ## Individual Scores
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-extract_random <- function(model, illusion = "Delboeuf") {
-  random <- as.data.frame(model_delboeuf_rt)
-  random <- random[str_detect(names(random), regex("^r_Participant"))]
-
-  if (insight::model_info(model)$is_logit) {
-    param <- "Prob,"
-  } else {
-    param <- "Loc,"
-  }
+random <- rbind(
+  extract_random(model_delboeuf_err, "Delboeuf"),
+  # extract_random(model_delboeuf_rt, "Delboeuf"),
+  extract_random(model_ebbinghaus_err, "Ebbinghaus"),
+  # extract_random(model_ebbinghaus_rt, "Ebbinghaus"),
+  extract_random(model_rodframe_err, "Rod-Frame"),
+  # extract_random(model_rodframe_rt, "Rod-Frame"),
+  extract_random(model_verticalhorizontal_err, "Vertical-Horizontal"),
+  # extract_random(model_verticalhorizontal_rt, "Vertical-Horizontal"),
+  extract_random(model_zollner_err, "Zöllner"),
+  # extract_random(model_zollner_rt, "Zöllner"),
+  extract_random(model_white_err, "White"),
+  # extract_random(model_white_rt, "White"),
+  extract_random(model_mullerlyer_err, "Müller-Lyer"),
+  # extract_random(model_mullerlyer_rt, "Müller-Lyer"),
+  extract_random(model_ponzo_err, "Ponzo"),
+  # extract_random(model_ponzo_rt, "Ponzo"),
+  extract_random(model_poggendorff_err, "Poggendorff"),
+  # extract_random(model_poggendorff_rt, "Poggendorff"),
+  # extract_random(model_contrast_err, "Contrast")
+  # extract_random(model_contrast_rt, "Contrast")
+) |> 
+  filter(
+    # !str_detect(Parameter, "Intercept_"),
+    !str_detect(Parameter, "Cong_"),
+    !str_detect(Parameter, "Null_"))
 
 random |>
-    mutate(Draw = 1:nrow(random)) |>
-    pivot_longer(-Draw, names_to = "Parameter", values_to = "Value") |>
-    mutate(
-      Parameter = str_remove(Parameter, "r_Participant"),
-      Parameter = str_remove(Parameter, "\\]"),
-      Parameter = str_remove(Parameter, "\\["),
-      Parameter = paste0(param, Parameter),
-      Parameter = str_replace(Parameter, paste0(param, "__sigma"), "Disp,")
-    ) |>
-    separate(Parameter, into = c("Component", "Participant", "Parameter"), sep = ",") |>
-    mutate(
-      Parameter = paste0(Parameter, "_", Component),
-      Illusion_Type = illusion
-    ) |>
-    mutate(
-      Parameter = str_replace(Parameter, "Illusion_Strength", "Illu"),
-      Parameter = str_replace(Parameter, "Illusion_Difference", "Diff")
-    )
-}
-
-random <- rbind(
-  extract_random(model_delboeuf_rt, "Delboeuf"),
-  extract_random(model_delboeuf_err, "Delboeuf")
-)
-
-random |>
+  # group_by(Illusion_Type, Parameter) |> 
+  # standardize(select="Value") |> 
+  # ungroup() |> 
+  # mutate(order = mean(Value)) |> 
+  # ungroup() |> 
+  # mutate(Participant = fct_reorder(Participant, order)) |> 
   ggplot(aes(y = Participant, x = Value)) +
-  ggdist::stat_slabinterval(aes(fill = Illusion_Type)) +
+  ggdist::stat_slabinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
+  # coord_cartesian(xlim=c(-4, 4)) +
   facet_wrap(~Parameter, scales = "free_x")
@@
 random <- random |>
-  mutate(Parameter = paste0(Illusion_Type, "_", Parameter)) |>
+  mutate(Parameter = paste0(Illusion_Type, "_", Parameter),
+         Parameter = str_replace(Parameter, "Rod-Frame", "RodFrame"),
+         Parameter = str_replace(Parameter, "Vertical-Horizontal", "VerticalHorizontal"),
+         Parameter = str_replace(Parameter, "Zöllner", "Zollner")) |>
   group_by(Participant, Parameter) |>
@@
   ungroup() |>
+  group_by(Parameter) |> 
+  mutate(Score = Score / sd(Score)) |>  # Scale
+  ungroup()  |> 
   pivot_wider(names_from = Parameter, values_from = Score)
-```
-
-#### Approximation
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-dat <- data.frame()
-for (participant in unique(data$Participant)) {
-  freq <- glm(Error ~ Illusion_Difference * Illusion_Strength,
-    family = "binomial",
-    data = filter(
-      df, Participant == participant,
-      Illusion_Type == "Delboeuf"
-    )
-  )
-
-  parameters::parameters(freq)[1:2] |>
-    as.data.frame() |>
-    mutate(
-      Participant = participant,
-      Parameter = paste0("freq_Delboeuf_", Parameter)
-    )
-}
-```
+
+# plot(estimate_density(select(random, -Participant)))
+```
+
+<!-- #### Approximation -->
+
+<!-- ```{r message=FALSE, warning=FALSE, cache=cache} -->
+<!-- dat <- data.frame() -->
+<!-- for (participant in unique(data$Participant)) { -->
+<!--   freq <- glm(Error ~ Illusion_Difference * Illusion_Strength, -->
+<!--     family = "binomial", -->
+<!--     data = filter( -->
+<!--       df, Participant == participant, -->
+<!--       Illusion_Type == "Delboeuf" -->
+<!--     ) -->
+<!--   ) -->
+
+<!--   parameters::parameters(freq)[1:2] |> -->
+<!--     as.data.frame() |> -->
+<!--     mutate( -->
+<!--       Participant = participant, -->
+<!--       Parameter = paste0("freq_Delboeuf_", Parameter) -->
+<!--     ) -->
+<!-- } -->
+<!-- ``` -->
 
@@
   cor = as.matrix(cor),
-  n = 3,
+  n = 5,
   rotation = "oblimin",
+  fm = "mle",
   sort = TRUE
@@
 fa2 <- parameters::factor_analysis(data,
-  n = 1,
+  n = 2,
+  rotation = "varimax",
+  fm = "mle",
   sort = TRUE
@@
   attributes(fa2)$loadings_long |>
-    mutate(Component = str_replace(Component, "MR1", "G"))
+    mutate(Component = str_replace(Component, "ML", "G"))
 ) |>
@@
 # Get order of indices
-idx <- correlation::cor_sort(cor, hclust_method = "ward.D2") |>
-  summary() |>
-  pull(Parameter)
+# idx <- correlation::cor_sort(cor, hclust_method = "ward.D2") |>
+#   summary() |>
+#   pull(Parameter)
+# idx <- sort(fa$Variable)
+idx <- fa$Variable
+
+get_x <- function(name, x) {
+  case_when(
+    name == "G1" ~ 0/1,
+    name == "G2" ~ 1/1,
+    name == "ML1" ~ 0/4,
+    name == "ML2" ~ 1/4,
+    name == "ML3" ~ 2/4,
+    name == "ML4" ~ 3/4,
+    name == "ML5" ~ 4/4,
+    TRUE ~ x)
+}
 
@@
   mutate(
-    x = case_when(
-    name == "G" ~ 0.5,
-    name == "MR1" ~ 0,
-    name == "MR2" ~ 0.5,
-    name == "MR3" ~ 1,
-    TRUE ~ x),
+    x = get_x(name, x),
     y = case_when(
-    name == "G" ~ 2,
-    name == "MR1" ~ 1,
-    name == "MR2" ~ 1,
-    name == "MR3" ~ 1,
-    TRUE ~ 0)) |> 
+    str_detect(name, "G") ~ 2,
+    str_detect(name, "ML") ~ 1,
+    TRUE ~ 0),
+    angle = ifelse(name %in% idx, 90, 0),
+    hjust = ifelse(name %in% idx, 1.01, 0),
+    name = clean_name(name),
+    Illusion_Type = str_split_fixed(name, " - ", n=2)[, 1]) |> 
+  tidygraph::activate("edges") |> 
+  filter(abs(Loading) > 0.2) |> 
   ggraph(layout = "nicely") +
   geom_edge_link(aes(edge_width = Loading, edge_colour = as.factor(sign(Loading)))) +
-  geom_node_point() +
-  geom_node_text(aes(label = name)) +
+  geom_node_point(aes(color = Illusion_Type), size=3) +
+  geom_node_text(aes(label = name, angle = angle, hjust=hjust)) +
+  scale_y_continuous(expand = expansion(c(.6, .1))) +
+  scale_edge_width_continuous(range=c(0.1, 1)) +
+  scale_edge_color_manual(values=c("#F44336", "#2196F3")) +
+  scale_color_manual(values=c("Delboeuf" = "#F44336", Ebbinghaus = "#E91E63",
+                              "Rod-Frame" = "#9C27B0")) +
   guides(edge_colour = "none") +
   ggraph::theme_graph()
+```
+
+### SEM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+library(lavaan)
+library(tidySEM)
+
+names <- names(select(random, -Participant))
+data <- setNames(as.data.frame(str_split_fixed(names, "_", 3)), c("Illusion_Type", "Parameter", "Model"))
+data$Name <- names
+
+# Model 1
+lvl1 <- paste(data$Illusion_Type, "=~", data$Name)
+lvl2 <- paste("G =~", unique(data$Illusion_Type))
+resid <- as.data.frame(t(utils::combn(x=unique(data$Illusion_Type), 2)))
+resid <- paste(resid$V1, "~~", resid$V2)[1:2]
+
+m1 <- sem(model = paste0(c(lvl1, lvl2, resid), collapse = "\n"),
+           data  = random)
+
+# Model 2
+lvl1 <- paste(data$Model, "=~", data$Name)
+lvl2 <- paste("G =~", unique(data$Model))
+
+m2 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
+           data  = random)
+
+# Model 3
+lvl1 <- paste(data$Parameter, "=~", data$Name)
+lvl2 <- paste("G =~", unique(data$Parameter))
+
+m3 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
+           data  = random)
+
+anova(m1, m2, m3)
+
+summary(m1, standardize = TRUE)
+
+tidySEM::graph_sem(model = m1)
+
+# as_tbl_graph(fit, standardize = TRUE)
 ```
@@
   merge(random, by = "Participant") |> 
-  cbind(data, predict(fa2, names="G"))
+  cbind(data, predict(fa2, names=c("G1", "G2")))
 

### 2022-08-07 study2/functions.R

@@ -1,0 +1,540 @@
+logmod <- function(x) sign(x) * log(1 + abs(x))
+sqrtmod <- function(x) sign(x) * sqrt(abs(x))
+cbrtmod <- function(x) sign(x) * (abs(x)**(1/3))
+
+
+
+
+
+
+best_models <- function(data) {
+  models_err <- list()
+  models_rt <- list()
+  for (k1 in c("", "_log", "_sqrt", "_cbrt")) {
+    for (k2 in c("", "_log", "_sqrt", "_cbrt")) {
+      for (side in c("", "--SIDE")) {
+        for (effect in c("", "--EFFECT")) {
+          name <- paste0("DIFF", k1, "--", "STRENGTH", k2, side, effect)
+          message(name)
+          f <- paste0(
+            "Illusion_Difference",
+            k1,
+            " * Illusion_Strength",
+            k2,
+            " + (1|Participant)"
+          )
+
+          if (side == "--SIDE") f <- paste0("Illusion_Side * ", f)
+          if (effect == "--EFFECT") f <- paste0("Illusion_Effect * ", str_replace(str_replace(f, "Illusion_Strength", "abs(Illusion_Strength"), " \\+ \\(1", "\\) \\+ \\(1"))
+
+          m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)),
+                                data = data, family = "binomial"
+          )
+          if (performance::check_convergence(m)) {
+            models_err[[name]] <- m
+          }
+
+          m <- glmmTMB::glmmTMB(as.formula(paste0("RT ~ ", f)),
+                                data = filter(data, Error == 0)
+          )
+          if (performance::check_convergence(m)) {
+            models_rt[[name]] <- m
+          }
+        }
+      }
+    }
+  }
+
+  # RT
+  to_keep <- compare_performance(models_rt, metrics = c("BIC")) |>
+    arrange(BIC) |>
+    slice(1:5) |>
+    pull(Name)
+
+  test <- test_performance(models_rt[to_keep], reference = 1)
+  perf <- compare_performance(models_rt[to_keep], metrics = c("BIC", "R2"))
+  rt <- merge(perf, test) |>
+    arrange(BIC) |>
+    select(Name, BIC, R2_marginal, BF) |>
+    mutate(
+      BF = insight::format_bf(BF, name = ""),
+      Model = "RT"
[…]

### 2022-08-07 study2/preprocessing.R

@@ -46,7 +46,7 @@ preprocess_raw <- function(file) {
     Trial = as.numeric(trials$trial_number),
     Stimulus = gsub(".png", "", gsub("stimuli/", "", trials$stimulus)),
     Illusion_Strength = as.numeric(trials$illusion_strength),
-    Illusion_Effect = as.factor(sign(as.numeric(trials$illusion_strength))),
+    Illusion_Effect = ifelse(sign(as.numeric(trials$illusion_strength)) == -1, "Congruent", ifelse(sign(as.numeric(trials$illusion_strength)) == 0, "Null", "Incongruent")),
     Illusion_Side = as.factor(sign(as.numeric(trials$illusion_difference))),
     Illusion_Difference = abs(as.numeric(trials$illusion_difference)),
     Answer = trials$response,
@@ -116,17 +116,25 @@ preprocess_raw <- function(file) {
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Hispanic", "Hisapanic"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Mixed/Latino", "Mexicano"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Hispanic/Latino "), "Latino", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latin"), "Latino", df$Ethnicity)
+  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latin", "Latina"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White Middle European (Slavic)", "Greek"), "Caucasian", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White - Caucasian", " Caucasian"), "Caucasian", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("White", "Caucasian ", "German"), "Caucasian", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("European/White"), "Caucasian", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Black", "Black African "), "African", df$Ethnicity)
[…]

### 2022-08-16 manuscript.Rmd
script

@@
 <!-- Intro  & Gaps-->
-Visual illusions are blabla. Despite the historical interest, several gaps remained unanswered:
+Visual illusions are *blabla*.
+Despite the historical interest, several gaps remained unanswered: *blabla*.
+One key challenge hindering the further development of the field is the relative difficulty to adapt visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
 
-This could be caused by the relative difficulty to adapt visual illusions to an experimental setting, which typically requires the parametric modulation of variables at stake.
+<!-- Pyllusion -->
+To address this issue, we first developed a parametric framework to generate visual illusions, that we implemented and made accessible in the open-source *Pyllusion* software (REF).
 
@@
 Many visual illusions can be seen as made of *targets* (e.g., parallel lines, same-size circles, same-length lines), which perception is biased by the *context* (e.g., parallel lines are seen as angled, circle or segments appear of different lengths).
+
+This framework allows the continuous and independent modulation of *illusion strength* and *task difficulty*.
+
 To date, most of the paradigms prompt the subjective extent the context biases the perception ("how much does the circle appear bigger", e.g., **REFS**).
 
-Although visual illusions are inherently tied to subjective perception, we tested a reversal of the traditional paradigm to access the "objective" effect of illusions.
+Although visual illusions are inherently tied to subjective perception, we tested a reversal of the traditional paradigm to access the "objective" effect of illusions. By generating stimuli that
 
-To achieve this, we first developed a parametric framework to generate visual illusions that we implemented and made accessible in the open-source Pyllusion software (REF).
-This allows us to generate stimuli ...
+In the present set of studies, we will test this novel paradigm to investigate 1) if the effect of illusion can be manipulated continuously, 2) whether 10 different illusions share a common latent factor, 3) how the the inter-individual sensitivity to illusions relates to other characteristics, such as sex or personality.
 
-In the present study, we will test this novel paradigm to ...
+## Methods
 

### 2022-08-16 study2/study2.Rmd
script

@@
   mc.cores = 4,
-  brms.algorithm = "meanfield",
+  brms.algorithm = "sampling",
   brms.backend = "cmdstanr"
@@
 
-cache <- FALSE
-runModels <- FALSE
+cache <- TRUE
+runModels <- TRUE
+bestModels <- TRUE
 fig.width <- see::golden_ratio(7)
@@
 df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"])
-df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"] <- round(df[df$Illusion_Type == "Ebbinghaus", "Illusion_Difference"], 5)
-
-# Transformation
-df$Illusion_Difference_log <- log(1 + df$Illusion_Difference)
-df$Illusion_Difference_sqrt <- sqrt(df$Illusion_Difference)
-df$Illusion_Difference_cbrt <- round(df$Illusion_Difference**(1 / 3), 4)
-df$Illusion_Strength_log <- sign(df$Illusion_Strength) * log(1 + abs(df$Illusion_Strength))
-df$Illusion_Strength_sqrt <- sign(df$Illusion_Strength) * sqrt(abs(df$Illusion_Strength))
-df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1 / 3))
 
@@
   mutate(
-    Date = as.Date(Date),
+    Date = lubridate::dmy(Date),
     Participant = fct_reorder(Participant, Date),
@@
     Illusion_Side = as.factor(Illusion_Side),
+    # Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Null", "Congruent"),
     Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
@@
   )
+
+# Fix precision
+for(ill in unique(df$Illusion_Type)) {
+  data <- df[df$Illusion_Type == ill, ] 
+  i <- 10
+  while (length(sort(unique(signif(data$Illusion_Difference, i)))) != 8) {
+    i <- i - 1
+  }
+  df[df$Illusion_Type == ill, "Illusion_Difference"] <- signif(df[df$Illusion_Type == ill, "Illusion_Difference"], i)
+  if (i != 10) {
+    message(ill, ": Illusion_Difference values != 8. Rounded to ", i, ".")
+  }
+}
+
+# Transformation
+df <- df |> 
+  mutate(    
+    Illusion_Difference_log = log(1 + Illusion_Difference),
+    Illusion_Difference_sqrt = sqrt(Illusion_Difference),
+    Illusion_Difference_cbrt = Illusion_Difference**(1 / 3),
+    Illusion_Strength_log = sign(Illusion_Strength) * log(1 + abs(Illusion_Strength)),
+    Illusion_Strength_sqrt = sign(Illusion_Strength) * sqrt(abs(Illusion_Strength)),
+    Illusion_Strength_cbrt = sign(Illusion_Strength) * (abs(Illusion_Strength)**(1 / 3))
+    )
 ```
@@
 
-# Dear participant, thank you for participating in our study.
+# Dear participant, thank you for participating in our studfy.
 Unfortunately, we didn't receive your data :( did something happen?
@@
   # 2nd block of responses very fast
-  # Prolific Status: RETURN REQUESTED (07/08)
+  # Prolific Status: REJECTED (15/08)
   "61443e47c248a375c899a0cf_9qz3u"
@@
 p <- df |> 
+  # filter(Date > "2022-08-12") |> 
   filter(RT < 10000) |> 
@@
   ungroup() |>
+  mutate(Participant = factor(Participant, levels=levels(df$Participant))) |> 
   mutate(color = ifelse(Participant %in% outliers, "red", ifelse(Participant %in% partial_outliers, "orange", "blue"))) |>
@@
 p1 <- temp |>
-  estimate_density(at = c("Illusion_Type", "Block")) |>
+  estimate_density(at = c("Illusion_Type", "Block"), method="KernSmooth") |>
   ggplot(aes(x = x, y = y)) +
@@
   ggplot(aes(x = x, y = y)) +
-  geom_area(data = normalize(estimate_density(df, select = "RT"), select = "y"), alpha = 0.2) +
-  geom_line(aes(color = Participant, linetype = Outlier)) +
+  geom_area(data = normalize(estimate_density(filter(df, RT < 10000), select = "RT"), select = "y"), alpha = 0.2) +
+  geom_line(aes(color = Participant, linetype = Outlier), alpha=0.2) +
   geom_vline(xintercept = c(125), linetype = "dashed", color = "red") +
@@
   see::theme_modern() +
-  theme(axis.text.x = element_blank())
-
-p1 | p2
+  theme(axis.text.x = element_blank()) +
+  labs(y = "Percentage of outlier trials")
+
+p1 / p2
 ```
@@
 p6 <- plot_waffle(dfsub, "Nationality") +
-  scale_fill_metro_d()
+  scale_fill_manual(values = c("South American" = "#FF5722", "Middle Eastern" = "#FF9800", "Western" = "#2196F3", "African" = "#4CAF50", "South African" = "#009688"))
 
 p7 <- plot_waffle(dfsub, "Ethnicity") +
-  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0"))
+  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0", "Mixed" = "#795548"))
 
@@
 
-(p1 / p2 / p3) | (p4 / p5 / p6) | (p7 / p8 / p9)
+(p1 / p2 / p3) | (p4 / p5 / p9) | (p6 / p7 / p8)
 ```
@@
 data <- filter(df, Illusion_Type == "Delboeuf")
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+
+# mm <- model.matrix(Error ~ abs(Illusion_Strength) * Illusion_Effect, data=data)
+# head(as.data.frame(mm))
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 formula <- brms::bf(
-  Error ~ logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
   family = "bernoulli"
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-
 data <- filter(df, Illusion_Type == "Delboeuf", Error == 0)
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
-# formula <- brms::bf(
-#   RT ~ Illusion_Difference * Illusion_Strength +
-#     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
-#   sigma ~ Illusion_Difference * Illusion_Strength +
-#     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
-#   # ndt ~ Illusion_Difference * Illusion_Strength +
-#   #   (1 | G | Participant),
-#   family = "shifted_lognormal"
-# )
-# formula <- brms::bf(
-#   RT ~ Illusion_Difference * Illusion_Strength +
-#     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
-#   sigma ~ Illusion_Difference * Illusion_Strength +
-#     (1 + (Illusion_Difference * Illusion_Strength) | Participant),
-#   # beta ~ Illusion_Difference * Illusion_Strength +
-#   #   (1 + (Illusion_Difference * Illusion_Strength) | Participant),
-#   family = "exgaussian"
-# )
-
-formula <- brms::bf(
-  RT ~ logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant),
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
   sigma ~ 1 + (1 | Participant),
+  beta ~ 1 + (1 | Participant),
   family = "exgaussian"
 )
-# prior = c(
-#     prior(normal(6, 1.5), class = Intercept),
-#     prior(normal(0, 1), class = sigma),
-#     prior(normal(0, .01), class = b, coef = c_trial)
-#   )
 
 # brms::get_prior(formula, data = data)
-# brms::validate_prior(formula)
+# plot(seq(-5, 5, length.out=100), brms::dstudent_t(seq(-5, 5, length.out=100), df = 3, 0.1, 0.5))
+priors <- c(
+  # Intercept
+  prior("student_t(3, 0, 0.1)", class = "Intercept"),
+  prior("student_t(3, 0.35, 0.1)", class = "sd", coef="Intercept", group="Participant"),
+  # Default parameters
+  prior("student_t(3, 0, 3)", class = "b"),
+  prior("student_t(3, 0.35, 0.5)", class = "sd", group="Participant")
+
+  # prior("student_t(3, 0, 1)", class="Intercept", dpar = "sigma"),
+  # prior("student_t(3, 0.1, 1)", class="sd", dpar = "sigma")
+
+  # prior("student_t(3, 0.1, 0.5)", class = "ndt")
+  ) |> 
+  brms::validate_prior(formula, data = data)
+
+# bayestestR::model_to_priors(model_delboeuf_rt, scale_multiply=3)
 
@@
   data = data,
-  refresh = 0
+  refresh = 50,
+  init=0
 )
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2}
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=FALSE}
 p_delboeuf <- plot_all(data, p_delboeuf_err, p_delboeuf_rt)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+best_models(data)
+```
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_descriptive_err(data)
+```
+
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_ebbinghaus_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_delboeuf_err)
+
+save(model_ebbinghaus_err, file="models/model_ebbinghaus_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_ebbinghaus_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_ebbinghaus_err <- plot_model_err(data, model_ebbinghaus_err)
+p_ebbinghaus_err
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ebbinghaus", Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 + sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  sigma ~ 1 + (1 | Participant),
+  beta ~ 1 + (1 | Participant),
+  family = "exgaussian"
+)
+
+model_ebbinghaus_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0,
+  init = 0
+)
+
+plot_ppcheck(model_ebbinghaus_rt)
+# performance::check_model(model_delboeuf_rt)
+# parameters::parameters(model)
+
+save(model_ebbinghaus_rt, file="models/model_ebbinghaus_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_ebbinghaus_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_ebbinghaus_rt <- plot_model_rt(data, model_ebbinghaus_rt)
+p_ebbinghaus_rt
+```
+
+### Rod and Frame
+
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Rod-Frame")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-model_ebbinghaus_err <- brms::brm(formula,
+model_rodframe_err <- brms::brm(formula,
   data = data,
@@
 
-save(model_ebbinghaus_err, file="models/model_ebbinghaus_err.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_ebbinghaus_err.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_ebbinghaus_err <- plot_model_err(data, model_ebbinghaus_err)
-p_ebbinghaus_err
+save(model_rodframe_err, file="models/model_rodframe_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_rodframe_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_rodframe_err <- plot_model_err(data, model_rodframe_err)
+p_rodframe_err
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Ebbinghaus", Error == 0)
+data <- filter(df, Illusion_Type == "Rod-Frame", Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  sigma ~ 1 + (1 | Participant),
+  beta ~ 1 + (1 | Participant),
+  family = "exgaussian"
+)
+
+model_rodframe_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0,
+  init = 0
+)
+
+plot_ppcheck(model_rodframe_rt)
+# performance::check_model(model_delboeuf_rt)
+# parameters::parameters(model)
+
+save(model_rodframe_rt, file="models/model_rodframe_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_rodframe_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_rodframe_rt <- plot_model_rt(data, model_rodframe_rt)
+p_rodframe_rt
+```
+
+### Vertical-Horizontal
+
+#### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Vertical-Horizontal") 
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+best_models(data)
+```
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_descriptive_err(data)
+```
+
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_verticalhorizontal_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_verticalhorizontal_err)
+
+save(model_verticalhorizontal_err, file="models/model_verticalhorizontal_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_verticalhorizontal_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_verticalhorizontal_err <- plot_model_err(data, model_verticalhorizontal_err)
+p_verticalhorizontal_err
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0)
 
@@
 
-model_ebbinghaus_rt <- brms::brm(formula,
+model_verticalhorizontal_rt <- brms::brm(formula,
   data = data,
@@
 
-plot_ppcheck(model_ebbinghaus_rt)
-# performance::check_model(model_delboeuf_rt)
+plot_ppcheck(model_verticalhorizontal_rt)
+# performance::check_model(model_verticalhorizontal_rt)
 # parameters::parameters(model)
 
-save(model_ebbinghaus_rt, file="models/model_ebbinghaus_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_ebbinghaus_rt.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_ebbinghaus_rt <- plot_model_rt(data, model_ebbinghaus_rt)
-p_ebbinghaus_rt
-```
-
-### Rod and Frame
+save(model_verticalhorizontal_rt, file="models/model_verticalhorizontal_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_verticalhorizontal_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_verticalhorizontal_rt <- plot_model_rt(data, model_verticalhorizontal_rt)
+p_verticalhorizontal_rt
+```
+
+### Zöllner
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Rod-Frame")
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+data <- filter(df, Illusion_Type == "Zöllner")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+best_models(data)
+```
+
+#### Error Rate
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_descriptive_err(data)
+```
+
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  Error ~ logmod(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+# brms::get_prior(formula, data = data)
+# brms::validate_prior(formula)
+
+model_zollner_err <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model_zollner_err)
+
+save(model_zollner_err, file="models/model_zollner_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_zollner_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_zollner_err <- plot_model_err(data, model_zollner_err)
+p_zollner_err
+```
+
+#### Reaction Time
+
+##### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Zöllner", Error == 0)
+
+plot_descriptive_rt(data)
+```
+
+##### Model Specification
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+formula <- brms::bf(
+  RT ~ logmod(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
+    (1 | Participant),
+  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
+  #   (1 | Participant),
+  family = "exgaussian"
+)
+
+model_zollner_rt <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+plot_ppcheck(model_zollner_rt)
+# performance::check_model(model_zollner_rt)
+# parameters::parameters(model)
+
+save(model_zollner_rt, file="models/model_zollner_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_zollner_rt.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_zollner_rt <- plot_model_rt(data, model_zollner_rt)
+p_zollner_rt
+```
+
+### White
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "White")
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-model_rodframe_err <- brms::brm(formula,
+model_white_err <- brms::brm(formula,
   data = data,
@@
 
-# parameters::parameters(model_delboeuf_err)
-
-save(model_rodframe_err, file="models/model_rodframe_err.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_rodframe_err.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_rodframe_err <- plot_model_err(data, model_rodframe_err)
-p_rodframe_err
+# parameters::parameters(model_white_err)
+
+save(model_white_err, file="models/model_white_err.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache}
+load("models/model_white_err.RData")
+```
+
+##### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+p_white_err <- plot_model_err(data, model_white_err)
+p_white_err
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Rod-Frame", Error == 0)
+data <- filter(df, Illusion_Type == "White", Error == 0)
 
@@
 
-model_rodframe_rt <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-plot_ppcheck(model_rodframe_rt)
-# performance::check_model(model_delboeuf_rt)
-# parameters::parameters(model)
-
-save(model_rodframe_rt, file="models/model_rodframe_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_rodframe_rt.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_rodframe_rt <- plot_model_rt(data, model_rodframe_rt)
-p_rodframe_rt
-```
-
-### Vertical-Horizontal
-
-#### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Vertical-Horizontal") 
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-best_models(data)
-```
-
-#### Error Rate
-
-##### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_descriptive_err(data)
-```
-
-##### Model Specification
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
-formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-# brms::get_prior(formula, data = data)
-# brms::validate_prior(formula)
-
-model_verticalhorizontal_err <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model_verticalhorizontal_err)
-
-save(model_verticalhorizontal_err, file="models/model_verticalhorizontal_err.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_verticalhorizontal_err.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_verticalhorizontal_err <- plot_model_err(data, model_verticalhorizontal_err)
-p_verticalhorizontal_err
-```
-
-#### Reaction Time
-
-##### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0)
-
-plot_descriptive_rt(data)
-```
-
-##### Model Specification
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
-formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant),
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
-  family = "exgaussian"
-)
-
-model_verticalhorizontal_rt <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-plot_ppcheck(model_verticalhorizontal_rt)
-# performance::check_model(model_verticalhorizontal_rt)
-# parameters::parameters(model)
-
-save(model_verticalhorizontal_rt, file="models/model_verticalhorizontal_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_verticalhorizontal_rt.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_verticalhorizontal_rt <- plot_model_rt(data, model_verticalhorizontal_rt)
-p_verticalhorizontal_rt
-```
-
-### Zöllner
-
-#### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Zöllner")
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-best_models(data)
-```
-
-#### Error Rate
-
-##### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_descriptive_err(data)
-```
-
-##### Model Specification
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
-formula <- brms::bf(
-  Error ~ logmod(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-# brms::get_prior(formula, data = data)
-# brms::validate_prior(formula)
-
-model_zollner_err <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model_zollner_err)
-
-save(model_zollner_err, file="models/model_zollner_err.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_zollner_err.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_zollner_err <- plot_model_err(data, model_zollner_err)
-p_zollner_err
-```
-
-#### Reaction Time
-
-##### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Zöllner", Error == 0)
-
-plot_descriptive_rt(data)
-```
-
-##### Model Specification
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
-formula <- brms::bf(
-  RT ~ logmod(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
-    (1 | Participant),
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
-  family = "exgaussian"
-)
-
-model_zollner_rt <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-plot_ppcheck(model_zollner_rt)
-# performance::check_model(model_zollner_rt)
-# parameters::parameters(model)
-
-save(model_zollner_rt, file="models/model_zollner_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_zollner_rt.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_zollner_rt <- plot_model_rt(data, model_zollner_rt)
-p_zollner_rt
-```
-
-### White
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "White")
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-best_models(data)
-```
-
-#### Error Rate
-
-##### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_descriptive_err(data)
-```
-
-##### Model Specification
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
-formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
-  family = "bernoulli"
-)
-
-# brms::get_prior(formula, data = data)
-# brms::validate_prior(formula)
-
-model_white_err <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model_white_err)
-
-save(model_white_err, file="models/model_white_err.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_white_err.RData")
-```
-
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_white_err <- plot_model_err(data, model_white_err)
-p_white_err
-```
-
-#### Reaction Time
-
-##### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "White", Error == 0)
-
-plot_descriptive_rt(data)
-```
-
-##### Model Specification
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
-formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant),
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
-  family = "exgaussian"
-)
-
 model_white_rt <- brms::brm(formula,
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-### Delboeuf
-
-#### Reliability
+### Reliability
 
@@
   extract_random(model_delboeuf_err, "Delboeuf"),
-  # extract_random(model_delboeuf_rt, "Delboeuf"),
-  extract_random(model_ebbinghaus_err, "Ebbinghaus"),
+  extract_random(model_delboeuf_rt, "Delboeuf")
+  # extract_random(model_ebbinghaus_err, "Ebbinghaus"),
   # extract_random(model_ebbinghaus_rt, "Ebbinghaus"),
-  extract_random(model_rodframe_err, "Rod-Frame"),
+  # extract_random(model_rodframe_err, "Rod-Frame"),
   # extract_random(model_rodframe_rt, "Rod-Frame"),
-  extract_random(model_verticalhorizontal_err, "Vertical-Horizontal"),
+  # extract_random(model_verticalhorizontal_err, "Vertical-Horizontal"),
   # extract_random(model_verticalhorizontal_rt, "Vertical-Horizontal"),
-  extract_random(model_zollner_err, "Zöllner"),
+  # extract_random(model_zollner_err, "Zöllner"),
   # extract_random(model_zollner_rt, "Zöllner"),
-  extract_random(model_white_err, "White"),
+  # extract_random(model_white_err, "White"),
   # extract_random(model_white_rt, "White"),
-  extract_random(model_mullerlyer_err, "Müller-Lyer"),
+  # extract_random(model_mullerlyer_err, "Müller-Lyer"),
   # extract_random(model_mullerlyer_rt, "Müller-Lyer"),
-  extract_random(model_ponzo_err, "Ponzo"),
+  # extract_random(model_ponzo_err, "Ponzo"),
   # extract_random(model_ponzo_rt, "Ponzo"),
-  extract_random(model_poggendorff_err, "Poggendorff"),
+  # extract_random(model_poggendorff_err, "Poggendorff"),
   # extract_random(model_poggendorff_rt, "Poggendorff"),
@@
   ggplot(aes(y = Participant, x = Value)) +
-  ggdist::stat_slabinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
+  # ggdist::stat_slabinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
+  ggdist::stat_pointinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
   # coord_cartesian(xlim=c(-4, 4)) +
-  facet_wrap(~Parameter, scales = "free_x")
+  facet_wrap(~Parameter, scales = "free_x") +
+  theme(axis.text.y = element_blank())
 
@@
 
-<!-- #### Approximation -->
+### Relationship with empirical descriptors
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+empirical <- df |>
+  group_by(Participant, Illusion_Type) |>
+  summarize(
+    # n = n(),
+    Error = sum(Error, na.rm=TRUE) / n(),
+    RT_Mean = mean(RT, na.rm=TRUE),
+    RT_SD = sd(RT, na.rm=TRUE),
+  ) |>
+  ungroup()
+
+r <- correlation::correlation(
+  data = random |> 
+    arrange(Participant) |> 
+    data_select(select="Delboeuf", regex=TRUE),
+  data2 = empirical |>
+    filter(Illusion_Type %in% c("Delboeuf")) |> 
+    pivot_wider(names_from = c("Illusion_Type"), values_from = c("Error", "RT_Mean", "RT_SD")) |> 
+    arrange(Participant)
+  )
+
+plot(summary(r))
+```
+<!-- ### Approximation -->
 

### 2022-08-16 study2/functions.R

@@ -1,6 +1,6 @@
 logmod <- function(x) sign(x) * log(1 + abs(x))
 sqrtmod <- function(x) sign(x) * sqrt(abs(x))
-cbrtmod <- function(x) sign(x) * (abs(x)**(1/3))
+cbrtmod <- function(x) sign(x) * (abs(x)**(1 / 3))
 
 
 
@@ -12,36 +12,33 @@ best_models <- function(data) {
   models_rt <- list()
   for (k1 in c("", "_log", "_sqrt", "_cbrt")) {
     for (k2 in c("", "_log", "_sqrt", "_cbrt")) {
-      for (side in c("", "--SIDE")) {
-        for (effect in c("", "--EFFECT")) {
-          name <- paste0("DIFF", k1, "--", "STRENGTH", k2, side, effect)
-          message(name)
-          f <- paste0(
-            "Illusion_Difference",
-            k1,
-            " * Illusion_Strength",
-            k2,
-            " + (1|Participant)"
-          )
+      # for (side in c("", "--SIDE")) {
+      # for (effect in c("", "--EFFECT")) {
+      name <- paste0("DIFF", k1, "--", "STRENGTH", k2)
+      message(name)
+      f <- paste0(
+        "Illusion_Side + Illusion_Effect * Illusion_Difference",
+        k1,
+        " * abs(Illusion_Strength",
+        k2,
+        ") + (1|Participant)"
+      )
 
-          if (side == "--SIDE") f <- paste0("Illusion_Side * ", f)
-          if (effect == "--EFFECT") f <- paste0("Illusion_Effect * ", str_replace(str_replace(f, "Illusion_Strength", "abs(Illusion_Strength"), " \\+ \\(1", "\\) \\+ \\(1"))
-
-          m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)),
-                                data = data, family = "binomial"
-          )
-          if (performance::check_convergence(m)) {
-            models_err[[name]] <- m
-          }
-
-          m <- glmmTMB::glmmTMB(as.formula(paste0("RT ~ ", f)),
-                                data = filter(data, Error == 0)
-          )
-          if (performance::check_convergence(m)) {
-            models_rt[[name]] <- m
-          }
-        }
+      m <- glmmTMB::glmmTMB(as.formula(paste0("Error ~ ", f)),
+        data = data, family = "binomial"
+      )
+      if (performance::check_convergence(m)) {
[…]

### 2022-08-16 study2/preprocessing.R

@@ -46,7 +46,8 @@ preprocess_raw <- function(file) {
     Trial = as.numeric(trials$trial_number),
     Stimulus = gsub(".png", "", gsub("stimuli/", "", trials$stimulus)),
     Illusion_Strength = as.numeric(trials$illusion_strength),
-    Illusion_Effect = ifelse(sign(as.numeric(trials$illusion_strength)) == -1, "Congruent", ifelse(sign(as.numeric(trials$illusion_strength)) == 0, "Null", "Incongruent")),
+    # Illusion_Effect = ifelse(sign(as.numeric(trials$illusion_strength)) == -1, "Congruent", ifelse(sign(as.numeric(trials$illusion_strength)) == 0, "Null", "Incongruent")),
+    Illusion_Effect = ifelse(sign(as.numeric(trials$illusion_strength)) == -1, "Congruent", "Incongruent"),
     Illusion_Side = as.factor(sign(as.numeric(trials$illusion_difference))),
     Illusion_Difference = abs(as.numeric(trials$illusion_difference)),
     Answer = trials$response,
@@ -115,26 +116,35 @@ preprocess_raw <- function(file) {
   # unique(df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Hispanic", "Hisapanic"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Mixed/Latino", "Mexicano"), "Latino", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Hispanic/Latino "), "Latino", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latin", "Latina"), "Latino", df$Ethnicity)
+  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Hispanic/Latino ", "Latin American "), "Latino", df$Ethnicity)
+  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latin", "Latina", "HISPANIC"), "Latino", df$Ethnicity)
+  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latino American", "Latino/Hispanic"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White Middle European (Slavic)", "Greek"), "Caucasian", df$Ethnicity)
[…]

### 2022-08-27 manuscript.Rmd
update

--- 2022-08-06 manuscript.Rmd
@@
 ---
-title             : "**Something**"
-shorttitle        : "Optimal Tolerance"
+title             : "**The Illusion Game: A Novel Experimental Paradigm to study Visual Illusions**"
+shorttitle        : "Illusion Game"
 author: 
@@
 
-keywords          : "chaos, complexity, fractal, physiology, tolerance"
+keywords          : "visual illusions, illusion game, pyllusion, illusion effect, personality"
 wordcount         : "1156"
 bibliography      : "bibliography.bib"
-floatsintext      : no
+floatsintext      : yes
 linenumbers       : yes
@@
 
+<!-- Title: -->
+<!-- The Illusion Game: Validation of a Novel Experimental Paradigm to study Visual Illusions  -->
+
 ## Introduction
@@
 <!-- Intro  & Gaps-->
-Visual illusions are blabla. Despite the historical interest, several gaps remained unanswered:
+Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems.
+They eloquently show that our brains did not evolve to be perfect perceptual devices, but to take into account contextual information and prior knowledge that manifests in our conscious experience **(REF)**.
+Despite the historical interest within the fields of visual perception **(REF)**, consciousness science **(REF)**, or psychiatry **(REF)**, several gaps remained unanswered.
+*mention common factor for illusions and link with disorders*.
 
-This could be caused by the relative difficulty to adapt visual illusions to an experimental setting, which typically requires the parametric modulation of variables at stake.
+<!-- Pyllusion -->
+One key challenge hindering the further development of illusion research is the relative difficulty to adapt visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+To address this issue, we first developed a parametric framework to manipulate visual illusions, that we implemented and made accessible in the open-source software *Pyllusion* **(REF)**.
+This software allows us to generate different types of historical visual illusions with a continuous and independent modulation of two parameters, *illusion strength* and *task difficulty* (see **Figure 1**).
+
+```{r message=FALSE, warning=FALSE, fig.cap="Explanation of the parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+Examples of stimuli showcasing the manipulation of the two main parameters, the task difficulty and the illusion strength (below)."}
+knitr::include_graphics("figures/Figure1.png")
+```
 
 <!-- Target, strength, difficulty -->
-Many visual illusions can be seen as made of *targets* (e.g., parallel lines, same-size circles, same-length lines), which perception is biased by the *context* (e.g., parallel lines are seen as angled, circle or segments appear of different lengths).
-To date, most of the paradigms prompt the subjective extent the context biases the perception ("how much does the circle appear bigger", e.g., **REFS**).
+Indeed, many visual illusions can be seen as made of *targets* (e.g., same-length lines), which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the segments appear of different lengths when they end with inwards or outwards pointing arrows).
+While most of the paradigms used in the literature prompt the subjective extent the context biases the perception ("how much does the circle appear bigger", e.g., **REFS**), *Pyllusions* allows to create illusions in which the targets are actually more or less different (i.e., one segment is longer than the other), and in which the illusion is of varying strength (the arrows are more or less pointy).
 
-Although visual illusions are inherently tied to subjective perception, we tested a reversal of the traditional paradigm to access the "objective" effect of illusions.
+<!-- Paradigm -->
+This opens the door of an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+Moreover, the illusion effect can be either "incongruent" (making the task even harder by biasing the perception in the opposite way) or "congruent" (making the task easier).
+Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm that could quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
 
-To achieve this, we first developed a parametric framework to generate visual illusions that we implemented and made accessible in the open-source Pyllusion software (REF).
-This allows us to generate stimuli ...
+<!-- Goals -->
+In the present set of studies, we will first test this novel paradigm by investigate if the effect of illusion and task difficulty can be manipulated continuously and separated statistically.
+Then, we will use the paradigm to assess whether 10 different illusions share a common latent factor (a long-standing debate).
+Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
 
-In the present study, we will test this novel paradigm to ...
+In line with our full open-access standards, all the material (stimuli generation code, experiment code, )
+
+## Study 1
+
+### Aim
+
+### Participants
+
+### Procedure
+
+### Results
+
+### Discussion
+
+## General Discussion
+
+## Future Directions
+
+We strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding further.
 

### 2022-08-27 manuscript.tex
update

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{visual illusions, illusion game, pyllusion, illusion effect, personality\newline\indent Word count: 1156}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Illusion Game: A Novel Experimental Paradigm to study Visual Illusions},
  pdfauthor={Dominique Makowski1},
  pdflang={en-EN},
  pdfkeywords={visual illusions, illusion game, pyllusion, illusion effect, personality},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{The Illusion Game: A Novel Experimental Paradigm to study Visual Illusions}}
\author{Dominique Makowski\textsuperscript{1}}
\date{}

\shorttitle{Illusion Game}

\authornote{

Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (\href{mailto:dom.makowski@gmail.com}{\nolinkurl{dom.makowski@gmail.com}}).

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Social Sciences, Nanyang Technological University, Singapore}

\abstract{%
Abstract abstract.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems. They eloquently show that our brains did not evolve to be perfect perceptual devices, but to take into account contextual information and prior knowledge that manifests in our conscious experience \textbf{(REF)}. Despite the historical interest within the fields of visual perception \textbf{(REF)}, consciousness science \textbf{(REF)}, or psychiatry \textbf{(REF)}, several gaps remained unanswered. \emph{mention common factor for illusions and link with disorders}.

One key challenge hindering the further development of illusion research is the relative difficulty to adapt visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest. To address this issue, we first developed a parametric framework to manipulate visual illusions, that we implemented and made accessible in the open-source software \emph{Pyllusion} \textbf{(REF)}. This software allows us to generate different types of historical visual illusions with a continuous and independent modulation of two parameters, \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).

\begin{figure}
\includegraphics[width=4.2in]{figures/Figure1} \caption{Explanation of the parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above). Examples of stimuli showcasing the manipulation of the two main parameters, the task difficulty and the illusion strength (below).}\label{fig:unnamed-chunk-2}
\end{figure}

Indeed, many visual illusions can be seen as made of \emph{targets} (e.g., same-length lines), which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the segments appear of different lengths when they end with inwards or outwards pointing arrows). While most of the paradigms used in the literature prompt the subjective extent the context biases the perception (``how much does the circle appear bigger'', e.g., \textbf{REFS}), \emph{Pyllusions} allows to create illusions in which the targets are actually more or less different (i.e., one segment is longer than the other), and in which the illusion is of varying strength (the arrows are more or less pointy).

This opens the door of an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength. Moreover, the illusion effect can be either ``incongruent'' (making the task even harder by biasing the perception in the opposite way) or ``congruent'' (making the task easier). Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm that could quantify the ``objective'' effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.

In the present set of studies, we will first test this novel paradigm by investigate if the effect of illusion can be manipulated continuously. Then, we will use it to assess whether 10 different illusions share a common latent factor (a long-standing debate), 3) how the the inter-individual sensitivity to illusions relates to other characteristics, such as sex or personality.

\hypertarget{study-1}{%
\subsection{Study 1}\label{study-1}}

\hypertarget{aim}{%
\subsubsection{Aim}\label{aim}}

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

\hypertarget{procedure}{%
\subsubsection{Procedure}\label{procedure}}

\hypertarget{results}{%
\subsubsection{Results}\label{results}}

\hypertarget{discussion}{%
\subsubsection{Discussion}\label{discussion}}

\hypertarget{general-discussion}{%
\subsection{General Discussion}\label{general-discussion}}

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\end{CSLReferences}

\clearpage
\renewcommand{\listfigurename}{Figure captions}

\end{document}

### 2022-08-27 study2/study2.Rmd
update

@@
 cache <- TRUE
-runModels <- TRUE
-bestModels <- TRUE
+runModels <- FALSE
+bestModels <- FALSE
 fig.width <- see::golden_ratio(7)
@@
 
-# Fix
+# Fix due to technical issue in the paradigm on the first day
 df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"] <- -1 * (df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"])
+df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Effect"] <- ifelse(sign(df[df$Illusion_Type == "Rod-Frame" & df$Date == "05/08/2022", "Illusion_Strength"]) == -1, "Congruent", "Incongruent")
 
@@
     Block = as.factor(Block),
-    Education = fct_relevel(Education, "Doctorate", "Master", "Bachelor", "High School", "Other")
+    Education = fct_relevel(Education, "High School", "Bachelor", "Master", "Doctorate", "Other", "Prefer not to Say")
   )
@@
 
+dfsub$Participant[grepl("60be8d", dfsub$Participant)]
+
 outliers <- c(
@@
   # Prolific Status: REJECTED (15/08)
-  "61443e47c248a375c899a0cf_9qz3u"
+  "61443e47c248a375c899a0cf_9qz3u",
+  # Error rate of 44% and very short RTs
+  # Prolific Status: RETURN REQUESTED (22/08)
+  "60dd626e212aa26c16d953b7_vw35k",
+  # 2nd block bad, first block 1/3 bad
+  # Prolific Status: RETURN REQUESTED (26/08)
+  "5eed37c13661d91555c38308_hpgnt",
+  # Prolific Status: RETURN REQUESTED (26/08)
+  "6159f8a9a2d80ecf1a5acd30_zc8rt",
+  # Prolific status: accepted (not enough proof)
+  "6136aa846e80ebd7daf6e148_aggf6"
 )
@@
   # Entire 2nd block bad
-  "61564e378e974bdbe42763a2_hhufm")
+  "61564e378e974bdbe42763a2_hhufm",
+  # Entire 2nd block bad
+  "5ff8b21dd2ca6f12bdaa5be8_c5rs6",
+  # Half of 2nd block bad
+  "613a9193ed68e49209b471b7_wrymn",
+  # 2nd block not good
+  "60be8dd61cec210d627483de_cmfx4")
 ```
@@
 p <- df |> 
-  # filter(Date > "2022-08-12") |> 
+  # filter(Date > "2022-08-24") |>
   filter(RT < 10000) |> 
@@
   ungroup() |>
-  mutate(Participant = factor(Participant, levels=levels(df$Participant))) |> 
+  mutate(Participant = fct_relevel(Participant, as.character(dfsub$Participant))) |> 
   mutate(color = ifelse(Participant %in% outliers, "red", ifelse(Participant %in% partial_outliers, "orange", "blue"))) |>
@@
 # p
-# ggsave("figures/outliers.png", p, width=20, height=15)
+# ggsave("figures/outliers.png", p, width=25, height=25)
 
@@
 
-p1 <- estimate_density(df, select = "RT", at = "Participant") |>
+p1 <- df |> 
+  filter(RT < 10000) |> 
+  estimate_density(select = "RT", at = "Participant") |>
   group_by(Participant) |>
@@
   group_by(Participant) |>
-  select(Participant, Age, Sex, Education, Nationality, Ethnicity, Duration, Break_Duration, Screen_Resolution, Screen_Refresh, Device_OS, starts_with("IPIP6"), starts_with("PID5")) |>
+  select(Participant, Age, Sex, Ethnicity, Education, Nationality, Duration, Break_Duration, Device_OS, starts_with("Screen"), starts_with("IPIP6"), starts_with("PID5")) |>
   slice(1) |>
@@
 p6 <- plot_waffle(dfsub, "Nationality") +
-  scale_fill_manual(values = c("South American" = "#FF5722", "Middle Eastern" = "#FF9800", "Western" = "#2196F3", "African" = "#4CAF50", "South African" = "#009688"))
+  scale_fill_manual(values = c("South American" = "#FF5722", "Middle Eastern" = "#FF9800", "Western" = "#2196F3", "Eastern" = "#673AB7", "African" = "#4CAF50", "South African" = "#009688"))
 
@@
 
+## Data Analysis
+
+In the first part of the study focuses on modelling the effect of illusion strength and task difficulty, on errors and reaction time, within each illusion.
+In order to achieve that, we will start by fitting a General Additive Models (GAM), which can accommodate possible non-linear effects and interactions.
+Using them as a "ground-truth", we will attempt at approximating them using general linear models, in which we can specify a participant-level variability of the effects (via random effects).
+
+We will then extract this participant-level variability in the effect of illusion strength, task difficulty and their interaction and use it as inter-individual features.
+We will estimate the relationship across different illusions using factor analysis (FA) and structural equation modelling (SEM).
+
+```{r make_GAMs, message=FALSE, warning=FALSE, eval=runModels}
+for(ill in c("Delboeuf", 
+             "Ebbinghaus", 
+             "Rod-Frame",
+             "Vertical-Horizontal", 
+             "Zöllner",
+             "White",
+             "Müller-Lyer", 
+             "Ponzo", 
+             "Poggendorff",
+             "Contrast")) {
+  print(ill)
+  model <- brms::brm(
+    Error ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + (1 | Participant),
+    data = filter(df, Illusion_Type == ill),
+    family = "bernoulli",
+    algorithm="sampling"
+  )
+
+  name <- paste0("gam_", tolower(clean_illusionName(ill)), "_err")
+  assign(name, model)  # Rename with string
+  save(list=name, file=paste0("models/", name, ".Rdata"))
+
+  model <- brms::brm(
+    brms::bf(
+      RT ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + (1 | Participant),
+      sigma ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + (1 | Participant),
+      beta ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + (1 | Participant)
+    ),
+    data = filter(df, Illusion_Type == ill, Error == 0),
+    family = "exgaussian",
+    init=0,
+    algorithm="sampling"
+  )
+
+  name <- paste0("gam_", tolower(clean_illusionName(ill)), "_rt")
+  assign(name, model)  # Rename with string
+  save(list=name, file=paste0("models/", name, ".Rdata"))
+}
+```
+
+# Results
+
 ## Models {.tabset}
@@
 best_models(data)
+
+m1 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect + (1 | Participant), data = data, family = "binomial")
+m2 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * sqrtmod(Illusion_Strength) + (1 | Participant), data = data, family = "binomial")
+
+compare_performance(m1, m2, metrics = c("R2", "BIC"))
+test_performance(m1, m2)
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_delboeuf, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  Error ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
   family = "bernoulli"
@@
 load("models/model_delboeuf_err.RData")
+load("models/gam_delboeuf_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_delboeuf_err <- plot_model_err(data, model_delboeuf_err)
+p_delboeuf_err <- plot_model_err(data, model_delboeuf_err, gam_delboeuf_err)
 p_delboeuf_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_delboeuf, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
-  sigma ~ 1 + (1 | Participant),
-  beta ~ 1 + (1 | Participant),
+  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)))  +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + (1 | Participant),
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + (1 | Participant),
   family = "exgaussian"
@@
   # Intercept
-  prior("student_t(3, 0, 0.1)", class = "Intercept"),
-  prior("student_t(3, 0.35, 0.1)", class = "sd", coef="Intercept", group="Participant"),
+  # prior("student_t(3, 0, 0.1)", class = "Intercept"),
+  # prior("student_t(3, 0.35, 0.1)", class = "sd", coef="Intercept", group="Participant"),
   # Default parameters
-  prior("student_t(3, 0, 3)", class = "b"),
-  prior("student_t(3, 0.35, 0.5)", class = "sd", group="Participant")
+  # prior("student_t(3, 0, 3)", class = "b"),
+  # prior("student_t(3, 0.35, 0.5)", class = "sd", group="Participant")
 
@@
 
+save(model_delboeuf_rt, file="models/model_delboeuf_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_delboeuf_rt.RData")
+load("models/gam_delboeuf_rt.RData")
+
 plot_ppcheck(model_delboeuf_rt)
@@
 # parameters::parameters(model)
-
-save(model_delboeuf_rt, file="models/model_delboeuf_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_delboeuf_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_delboeuf_rt <- plot_model_rt(data, model_delboeuf_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_delboeuf_rt <- plot_model_rt(data, model_delboeuf_rt, gam_delboeuf_rt)
 p_delboeuf_rt
@@
 best_models(data)
+
+m1 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant), data = data, family = "binomial")
+m2 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * logmod(Illusion_Strength) +
+    (1 | Participant), data = data, family = "binomial")
+
+test_performance(m1, m2)
+compare_performance(m1, m2)
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_ebbinghaus, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  Error ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) | Participant),
   family = "bernoulli"
@@
 load("models/model_ebbinghaus_err.RData")
+load("models/gam_ebbinghaus_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_ebbinghaus_err <- plot_model_err(data, model_ebbinghaus_err)
+p_ebbinghaus_err <- plot_model_err(data, model_ebbinghaus_err, gam_ebbinghaus_err)
 p_ebbinghaus_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_ebbinghaus, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect | Participant),
-  sigma ~ 1 + (1 | Participant),
-  beta ~ 1 + (1 | Participant),
+  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)))  +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) + (1 | Participant),
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) + (1 | Participant),
   family = "exgaussian"
@@
 
+save(model_ebbinghaus_rt, file="models/model_ebbinghaus_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_ebbinghaus_rt.RData")
+load("models/gam_ebbinghaus_rt.RData")
+
 plot_ppcheck(model_ebbinghaus_rt)
@@
 # parameters::parameters(model)
-
-save(model_ebbinghaus_rt, file="models/model_ebbinghaus_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_ebbinghaus_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_ebbinghaus_rt <- plot_model_rt(data, model_ebbinghaus_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_ebbinghaus_rt <- plot_model_rt(data, model_ebbinghaus_rt, gam_ebbinghaus_rt)
 p_ebbinghaus_rt
@@
 best_models(data)
+
+m1 <- glmmTMB::glmmTMB(Error ~ logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+    (1 | Participant), data = data, family = "binomial")
+m2 <- glmmTMB::glmmTMB(Error ~ logmod(Illusion_Difference) * logmod(Illusion_Strength) +
+    (1 | Participant), data = data, family = "binomial")
+
+compare_performance(m1, m2)
+test_performance(m1, m2)
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_rodframe, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  Error ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(Illusion_Strength)) +
+    (1 + Illusion_Effect / (logmod(Illusion_Difference) * logmod(Illusion_Strength)) | Participant),
   family = "bernoulli"
@@
 load("models/model_rodframe_err.RData")
+load("models/gam_rodframe_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_rodframe_err <- plot_model_err(data, model_rodframe_err)
+p_rodframe_err <- plot_model_err(data, model_rodframe_err, gam_rodframe_err)
 p_rodframe_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_rodframe, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant),
-  sigma ~ 1 + (1 | Participant),
-  beta ~ 1 + (1 | Participant),
+  RT ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) + (1 | Participant),
+  beta ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) + (1 | Participant),
   family = "exgaussian"
@@
 
+save(model_rodframe_rt, file="models/model_rodframe_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_rodframe_rt.RData")
+load("models/gam_rodframe_rt.RData")
+
 plot_ppcheck(model_rodframe_rt)
@@
 # parameters::parameters(model)
-
-save(model_rodframe_rt, file="models/model_rodframe_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_rodframe_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_rodframe_rt <- plot_model_rt(data, model_rodframe_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_rodframe_rt <- plot_model_rt(data, model_rodframe_rt, gam_rodframe_rt)
 p_rodframe_rt
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_verticalhorizontal, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant),
+  Error ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
   family = "bernoulli"
@@
 load("models/model_verticalhorizontal_err.RData")
+load("models/gam_verticalhorizontal_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_verticalhorizontal_err <- plot_model_err(data, model_verticalhorizontal_err)
+p_verticalhorizontal_err <- plot_model_err(data, model_verticalhorizontal_err, gam_verticalhorizontal_err)
 p_verticalhorizontal_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_verticalhorizontal, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant),
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
+  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + (1 | Participant),
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + (1 | Participant),
   family = "exgaussian"
@@
   data = data,
-  refresh = 0
-)
+  refresh = 100,
+  init = 0
+)
+
+save(model_verticalhorizontal_rt, file="models/model_verticalhorizontal_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_verticalhorizontal_rt.RData")
+load("models/gam_verticalhorizontal_rt.RData")
 
@@
 # parameters::parameters(model)
-
-save(model_verticalhorizontal_rt, file="models/model_verticalhorizontal_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_verticalhorizontal_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_verticalhorizontal_rt <- plot_model_rt(data, model_verticalhorizontal_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_verticalhorizontal_rt <- plot_model_rt(data, model_verticalhorizontal_rt, gam_verticalhorizontal_rt)
 p_verticalhorizontal_rt
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_zollner, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ logmod(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
-    (1 | Participant),
+  Error ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
+    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
   family = "bernoulli"
@@
 load("models/model_zollner_err.RData")
+load("models/gam_zollner_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_zollner_err <- plot_model_err(data, model_zollner_err)
+p_zollner_err <- plot_model_err(data, model_zollner_err, gam_zollner_err)
 p_zollner_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_zollner, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ logmod(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
+  RT ~  Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
+    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
+  sigma ~  Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
     (1 | Participant),
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
+  beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
+    (1 | Participant),
   family = "exgaussian"
@@
   data = data,
-  refresh = 0
-)
+  refresh = 100,
+  init = 0
+)
+
+save(model_zollner_rt, file="models/model_zollner_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_zollner_rt.RData")
+load("models/gam_zollner_rt.RData")
 
@@
 
-save(model_zollner_rt, file="models/model_zollner_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_zollner_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_zollner_rt <- plot_model_rt(data, model_zollner_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_zollner_rt <- plot_model_rt(data, model_zollner_rt, gam_zollner_rt)
 p_zollner_rt
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_white, essage=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  Error ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) | Participant),
   family = "bernoulli"
@@
   data = data,
-  refresh = 0
+  refresh = 100
 )
@@
 load("models/model_white_err.RData")
+load("models/gam_white_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_white_err <- plot_model_err(data, model_white_err)
+p_white_err <- plot_model_err(data, model_white_err, gam_white_err)
 p_white_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_white, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+  RT ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
     (1 | Participant),
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
+  beta ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
+    (1 | Participant),
   family = "exgaussian"
@@
   data = data,
-  refresh = 0
-)
+  refresh = 100,
+  init = 0
+)
+
+save(model_white_rt, file="models/model_white_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_white_rt.RData")
+load("models/gam_white_rt.RData")
 
@@
 # parameters::parameters(model)
-
-save(model_white_rt, file="models/model_white_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_white_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_white_rt <- plot_model_rt(data, model_white_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_white_rt <- plot_model_rt(data, model_white_rt, gam_white_rt)
 p_white_rt
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_mullerlyer, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ logmod(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + logmod(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  Error ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
   family = "bernoulli"
@@
 load("models/model_mullerlyer_err.RData")
+load("models/gam_mullerlyer_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_mullerlyer_err <- plot_model_err(data, model_mullerlyer_err)
+p_mullerlyer_err <- plot_model_err(data, model_mullerlyer_err, gam_mullerlyer_err)
 p_mullerlyer_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_mullerlyer, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ logmod(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
     (1 | Participant),
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
+    (1 | Participant),
   family = "exgaussian"
@@
   data = data,
-  refresh = 0
-)
+  refresh = 100,
+  init = 0
+)
+
+save(model_mullerlyer_rt, file="models/model_mullerlyer_rt.Rdata")
+# parameters::parameters(model_white_err)
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_mullerlyer_rt.RData")
+load("models/gam_mullerlyer_rt.RData")
 
@@
 # parameters::parameters(model)
-
-save(model_mullerlyer_rt, file="models/model_mullerlyer_rt.Rdata")
-# parameters::parameters(model_white_err)
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_mullerlyer_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_mullerlyer_rt <- plot_model_rt(data, model_mullerlyer_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_mullerlyer_rt <- plot_model_rt(data, model_mullerlyer_rt, gam_mullerlyer_rt)
 p_mullerlyer_rt
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r message=FALSE, warning=FALSE, cache=FALSE, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_ponzo, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
-    (1 | Participant),
+  Error ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
+    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
   family = "bernoulli"
@@
 load("models/model_ponzo_err.RData")
+load("models/gam_ponzo_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_ponzo_err <- plot_model_err(data, model_ponzo_err)
+p_ponzo_err <- plot_model_err(data, model_ponzo_err, gam_ponzo_err)
 p_ponzo_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_ponzo, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * abs(Illusion_Strength) * Illusion_Effect +
+  RT ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
+    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
+  sigma ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
     (1 | Participant),
-  # sigma ~ t2(Illusion_Difference * Illusion_Strength) +
-  #   (1 | Participant),
+  beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
+    (1 | Participant),
   family = "exgaussian"
@@
   data = data,
-  refresh = 0
-)
+  refresh = 100,
+  init = 0
+)
+
+save(model_ponzo_rt, file="models/model_ponzo_rt.Rdata")
+# parameters::parameters(model_white_err)
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_ponzo_rt.RData")
+load("models/gam_ponzo_rt.RData")
 
@@
 # parameters::parameters(model)
-
-save(model_ponzo_rt, file="models/model_ponzo_rt.Rdata")
-# parameters::parameters(model_white_err)
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_ponzo_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_ponzo_rt <- plot_model_rt(data, model_ponzo_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_ponzo_rt <- plot_model_rt(data, model_ponzo_rt, gam_ponzo_rt)
 p_ponzo_rt
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_poggendorff, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant),
+  Error ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
   family = "bernoulli"
@@
 load("models/model_poggendorff_err.RData")
+load("models/gam_poggendorff_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_poggendorff_err <- plot_model_err(data, model_poggendorff_err)
+p_poggendorff_err <- plot_model_err(data, model_poggendorff_err, gam_poggendorff_err)
 p_poggendorff_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_poggendorff, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+  RT ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
     (1 | Participant),
-  # sigma ~ 1 + (1 | Participant),
+  beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
+    (1 | Participant),
   family = "exgaussian"
@@
   data = data,
-  refresh = 0
-)
+  refresh = 100,
+  init = 0
+)
+
+save(model_ponzo_rt, file="models/model_poggendorff_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_poggendorff_rt.RData")
+load("models/gam_poggendorff_rt.RData")
 
@@
 # parameters::parameters(model)
-
-save(model_ponzo_rt, file="models/model_poggendorff_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_poggendorff_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_poggendorff_rt <- plot_model_rt(data, model_poggendorff_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_poggendorff_rt <- plot_model_rt(data, model_poggendorff_rt, gam_poggendorff_rt)
 p_poggendorff_rt
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_err_contrast, message=FALSE, warning=FALSE, eval=runModels}
 formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 + sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect | Participant),
+  Error ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
   family = "bernoulli"
@@
   data = data,
-  refresh = 0
+  refresh = 50
 )
@@
 load("models/model_contrast_err.RData")
+load("models/gam_contrast_err.RData")
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-p_contrast_err <- plot_model_err(data, model_contrast_err)
+p_contrast_err <- plot_model_err(data, model_contrast_err, gam_contrast_err)
 p_contrast_err
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=runModels}
+```{r m_rt_contrast, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 formula <- brms::bf(
-  RT ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect +
+  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
     (1 | Participant),
-  # sigma ~ 1 + (1 | Participant),
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
+    (1 | Participant),
   family = "exgaussian"
@@
   data = data,
-  refresh = 0
-)
+  refresh = 50,
+  init = 0
+)
+
+save(model_contrast_rt, file="models/model_contrast_rt.Rdata")
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_contrast_rt.RData")
+load("models/gam_contrast_rt.RData")
 
@@
 # parameters::parameters(model)
-
-save(model_contrast_rt, file="models/model_contrast_rt.Rdata")
-```
-```{r message=FALSE, warning=FALSE, cache=cache}
-load("models/model_contrast_rt.RData")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-p_contrast_rt <- plot_model_rt(data, model_contrast_rt)
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_contrast_rt <- plot_model_rt(data, model_contrast_rt, gam_contrast_rt)
 p_contrast_rt
@@
 
-### Reliability
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-random <- rbind(
+### Model-based Descriptors
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+data <- rbind(
   extract_random(model_delboeuf_err, "Delboeuf"),
-  extract_random(model_delboeuf_rt, "Delboeuf")
-  # extract_random(model_ebbinghaus_err, "Ebbinghaus"),
+  # extract_random(model_delboeuf_rt, "Delboeuf"),
+  extract_random(model_ebbinghaus_err, "Ebbinghaus"),
   # extract_random(model_ebbinghaus_rt, "Ebbinghaus"),
-  # extract_random(model_rodframe_err, "Rod-Frame"),
+  extract_random(model_rodframe_err, "Rod-Frame"),
   # extract_random(model_rodframe_rt, "Rod-Frame"),
-  # extract_random(model_verticalhorizontal_err, "Vertical-Horizontal"),
+  extract_random(model_verticalhorizontal_err, "Vertical-Horizontal"),
   # extract_random(model_verticalhorizontal_rt, "Vertical-Horizontal"),
-  # extract_random(model_zollner_err, "Zöllner"),
+  extract_random(model_zollner_err, "Zöllner"),
   # extract_random(model_zollner_rt, "Zöllner"),
-  # extract_random(model_white_err, "White"),
+  extract_random(model_white_err, "White"),
   # extract_random(model_white_rt, "White"),
-  # extract_random(model_mullerlyer_err, "Müller-Lyer"),
+  extract_random(model_mullerlyer_err, "Müller-Lyer"),
   # extract_random(model_mullerlyer_rt, "Müller-Lyer"),
-  # extract_random(model_ponzo_err, "Ponzo"),
+  extract_random(model_ponzo_err, "Ponzo"),
   # extract_random(model_ponzo_rt, "Ponzo"),
-  # extract_random(model_poggendorff_err, "Poggendorff"),
+  extract_random(model_poggendorff_err, "Poggendorff"),
   # extract_random(model_poggendorff_rt, "Poggendorff"),
-  # extract_random(model_contrast_err, "Contrast")
+  extract_random(model_contrast_err, "Contrast")
   # extract_random(model_contrast_rt, "Contrast")
@@
   filter(
-    # !str_detect(Parameter, "Intercept_"),
-    !str_detect(Parameter, "Cong_"),
+    !str_detect(Parameter, "Intercept_"),
+    !str_detect(Parameter, "Cong"),
+    !str_detect(Parameter, "Cong_Prob"),
     !str_detect(Parameter, "Null_"))
-
-random |>
-  # group_by(Illusion_Type, Parameter) |> 
-  # standardize(select="Value") |> 
-  # ungroup() |> 
-  # mutate(order = mean(Value)) |> 
-  # ungroup() |> 
-  # mutate(Participant = fct_reorder(Participant, order)) |> 
-  ggplot(aes(y = Participant, x = Value)) +
-  # ggdist::stat_slabinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
-  ggdist::stat_pointinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
-  # coord_cartesian(xlim=c(-4, 4)) +
-  facet_wrap(~Parameter, scales = "free_x") +
-  theme(axis.text.y = element_blank())
-
-random <- random |>
+# unique(data$Parameter)
+
+# data |>
+#   # group_by(Illusion_Type, Parameter) |> 
+#   # standardize(select="Value") |> 
+#   # ungroup() |> 
+#   # mutate(order = mean(Value)) |> 
+#   # ungroup() |> 
+#   # mutate(Participant = fct_reorder(Participant, order)) |> 
+#   ggplot(aes(y = Participant, x = Value)) +
+#   # ggdist::stat_slabinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
+#   ggdist::stat_pointinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
+#   # coord_cartesian(xlim=c(-4, 4)) +
+#   facet_wrap(~Parameter, scales = "free_x") +
+#   theme(axis.text.y = element_blank())
+
+# df[df$Illusion_Type == "Poggendorff" & df$Participant == "5cbb77bf418c540001641e67_sp53j", ]
+
+random <- data |>
   mutate(Parameter = paste0(Illusion_Type, "_", Parameter),
-         Parameter = str_replace(Parameter, "Rod-Frame", "RodFrame"),
-         Parameter = str_replace(Parameter, "Vertical-Horizontal", "VerticalHorizontal"),
-         Parameter = str_replace(Parameter, "Zöllner", "Zollner")) |>
+         Parameter = clean_illusionName(Parameter)) |>
   group_by(Participant, Parameter) |>
-  summarize(Score = as.numeric(bayestestR::map_estimate(Value)), .groups = "drop") |>
+  summarize(Score = median(Value, na.rm = TRUE), .groups = "drop") |>
   ungroup() |>
   group_by(Parameter) |>
-  mutate(Score = Score / sd(Score)) |>  # Scale
+  mutate(Score = Score / sd(Score, na.rm = TRUE)) |>  # Scale
   ungroup()  |>
@@
 
-### Relationship with empirical descriptors
-
-```{r message=FALSE, warning=FALSE, cache=cache}
+### Empirical Descriptors
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Average for 4 higher level of task difficulty
 empirical <- df |>
+  filter(Illusion_Effect == "Incongruent") |> 
+  group_by(Participant, Illusion_Type, Illusion_Strength) |>
+  arrange(Illusion_Difference) |> 
+  slice(0:4) |> 
+  # select(Illusion_Strength, Illusion_Difference)
   group_by(Participant, Illusion_Type) |>
   summarize(
-    # n = n(),
-    Error = sum(Error, na.rm=TRUE) / n(),
-    RT_Mean = mean(RT, na.rm=TRUE),
-    RT_SD = sd(RT, na.rm=TRUE),
+    TaskHard_Error = sum(Error, na.rm=TRUE) / n(),
+    TaskHard_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
+    # Illu_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
   ) |>
-  ungroup()
-
+  ungroup() |> 
+  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
+  pivot_wider(names_from = c("Illusion_Type"), values_from = c("TaskHard_Error", "TaskHard_RTMean"), names_glue = "{Illusion_Type}_{.value}")
+
+# Average for 4 higher level of illusion strength
+empirical <- df |>
+  filter(Illusion_Effect == "Incongruent") |> 
+  group_by(Participant, Illusion_Type, Illusion_Difference) |>
+  arrange(desc(Illusion_Strength)) |> 
+  slice(0:4) |> 
+  # select(Illusion_Strength, Illusion_Difference)
+  group_by(Participant, Illusion_Type) |>
+  summarize(
+    IlluStrong_Error = sum(Error, na.rm=TRUE) / n(),
+    IlluStrong_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
+    # Diff_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
+  ) |>
+  ungroup() |> 
+  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
+  pivot_wider(names_from = c("Illusion_Type"), values_from = c("IlluStrong_Error", "IlluStrong_RTMean"), names_glue = "{Illusion_Type}_{.value}") |> 
+  full_join(empirical, by = "Participant") 
+```
+
+<!-- ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE} -->
+<!-- rez <- parameters::n_factors(select(empirical, -Participant), n_max=30) -->
+<!-- plot(rez) -->
+
+<!-- r <- correlation::correlation(empirical, p_adjust = "none", sort=TRUE) -->
+
+<!-- r |> -->
+<!--   ggplot(aes(x = Parameter2, y = Parameter1)) + -->
+<!--   geom_tile(aes(fill = r)) + -->
+<!--   # geom_text(aes(label = Text), size = 2) + -->
+<!--   scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") + -->
+<!--   scale_x_discrete(expand = c(0, 0)) + -->
+<!--   scale_y_discrete(expand = c(0, 0)) + -->
+<!--   labs(title = "Correlation Matrix", x = NULL, y = NULL) + -->
+<!--   theme_minimal() + -->
+<!--   theme( -->
+<!--     axis.text.x = element_text(angle = 90, hjust = 1), -->
+<!--     plot.title = element_text(hjust = 0.5, face = "bold"), -->
+<!--     panel.grid.major = element_blank(), -->
+<!--     panel.grid.minor = element_blank() -->
+<!--   ) -->
+
+<!-- fa <- parameters::factor_analysis(select(empirical, -Participant), -->
+<!--   n = 13, -->
+<!--   rotation = "oblimin", -->
+<!--   fm = "mle", -->
+<!--   sort = TRUE -->
+<!-- ) -->
+
+<!-- print(tibble::as_tibble(fa), n=30) -->
+<!-- plot(fa) -->
+<!-- ``` -->
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Correlation
+r <- data.frame()
+for(ill in c("Delboeuf", 
+             "Ebbinghaus",
+             "RodFrame",
+             "VerticalHorizontal",
+             "Zollner",
+             "White",
+             "MullerLyer",
+             "Ponzo",
+             "Poggendorff",
+             "Contrast"
+             )) {
   r <- correlation::correlation(
@@
       arrange(Participant) |> 
-    data_select(select="Delboeuf", regex=TRUE),
+      data_select(select=ill, regex=TRUE),
     data2 = empirical |>
-    filter(Illusion_Type %in% c("Delboeuf")) |> 
-    pivot_wider(names_from = c("Illusion_Type"), values_from = c("Error", "RT_Mean", "RT_SD")) |> 
-    arrange(Participant)
-  )
-
-plot(summary(r))
-```
+      data_select(select=ill, regex=TRUE),
+    p_adjust="none"
+    ) |> 
+    mutate(Illusion_Type = ill,
+           Parameter1 = prettify_parameterName(Parameter1),
+           Parameter2 = prettify_parameterName(Parameter2)) |> 
+    select(Illusion_Type, Parameter1, Parameter2, r, p) |> 
+    rbind(r)
+}
+
+r |>
+  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
+  ggplot(aes(x = Parameter2, y = Parameter1)) +
+  geom_tile(aes(fill = r)) +
+  geom_text(aes(label=label)) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  # facet_wrap(~Illusion_Type, scales = "free") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+```
+
 <!-- ### Approximation -->
@@
 <!-- ```{r message=FALSE, warning=FALSE, cache=cache} -->
+<!-- individual <- data.frame() -->
+<!-- for (participant in unique(dfsub$Participant)) { -->
+<!--   cat(".") -->
 <!--   dat <- data.frame() -->
-<!-- for (participant in unique(data$Participant)) { -->
-<!--   freq <- glm(Error ~ Illusion_Difference * Illusion_Strength, -->
+
+<!--   dat <- glm(Error ~ sqrtmod(Illusion_Difference) * sqrtmod(Illusion_Strength), -->
 <!--     family = "binomial", -->
 <!--     data = filter( -->
-<!--       df, Participant == participant, -->
+<!--       df,  -->
+<!--       Participant == participant, -->
 <!--       Illusion_Type == "Delboeuf" -->
 <!--     ) -->
+<!--   ) |>  -->
+<!--     parameters::parameters() |>  -->
+<!--     mutate(Parameter = paste0("Delboeuf_Individual_", Parameter)) |>  -->
+<!--     rbind(dat) -->
+
+<!--   dat <- glm(Error ~ sqrt(Illusion_Difference) * logmod(Illusion_Strength), -->
+<!--     family = "binomial", -->
+<!--     data = filter( -->
+<!--       df,  -->
+<!--       Participant == participant, -->
+<!--       Illusion_Type == "Ebbinghaus" -->
 <!--     ) -->
-
-<!--   parameters::parameters(freq)[1:2] |> -->
+<!--   ) |>  -->
+<!--     parameters::parameters() |>  -->
+<!--     mutate(Parameter = paste0("Ebbinghaus_Individual_", Parameter)) |>  -->
+<!--     rbind(dat) -->
+
+<!--   individual <- dat |> -->
+<!--     select(Parameter, Coefficient) |>  -->
 <!--     as.data.frame() |> -->
@@
 <!--       Participant = participant, -->
-<!--       Parameter = paste0("freq_Delboeuf_", Parameter) -->
-<!--     ) -->
+<!--       Parameter = clean_parameterName(Parameter) -->
+<!--     ) |>  -->
+<!--     rbind(individual) -->
 <!-- } -->
+
+<!-- individual <- individual |>  -->
+<!--   pivot_wider(names_from = c("Parameter"), values_from = c("Coefficient")) |>  -->
+<!--   arrange(Participant) -->
 <!-- ``` -->
 
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Correlation
+r <- data.frame()
+for(ill in c("Delboeuf", 
+             "Ebbinghaus")) {
+  r <- correlation::correlation(
+    # data = random |> 
+    #   arrange(Participant) |> 
+    #   data_select(select=ill, regex=TRUE),
+    data = empirical |> 
+      data_select(select=ill, regex=TRUE),
+    data2 = individual |>
+      data_select(select=ill, regex=TRUE),
+    p_adjust="none"
+    ) |> 
+    mutate(Illusion_Type = ill,
+           Parameter1 = prettify_parameterName(Parameter1),
+           Parameter2 = prettify_parameterName(Parameter2)) |> 
+    select(Illusion_Type, Parameter1, Parameter2, r, p) |> 
+    rbind(r)
+}
+
+r |>
+  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
+  ggplot(aes(x = Parameter2, y = Parameter1)) +
+  geom_tile(aes(fill = r)) +
+  geom_text(aes(label=label)) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  facet_wrap(~Illusion_Type, scales = "free")
+```
+
 ## Structure
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-clean_name <- function(x) {
-  x |>
-    str_replace("_Loc", " (loc)") |>
-    str_replace("_Disp", " (disp)") |>
-    str_replace("_Prob", " (prob)") |>
-    str_replace("_", " - ")
-}
-
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 cor <- correlation::correlation(random, redundant = TRUE)
@@
   mutate(
-    Parameter1 = clean_name(Parameter1),
-    Parameter2 = clean_name(Parameter2)
+    Parameter1 = prettify_parameterName(Parameter1),
+    Parameter2 = prettify_parameterName(Parameter2)
   ) |>
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 data <- select(random, -Participant)
 
-rez <- parameters::n_factors(data)
+rez <- parameters::n_factors(data, n_max=15)
 plot(rez)
@@
   cor = as.matrix(cor),
-  n = 5,
+  n = 9,
   rotation = "oblimin",
@@
 
-# psych::omega(data, nfactors=3)
+print(tibble::as_tibble(fa), n=30)
+plot(fa)
+# psych::omega(data, fm = "mle", nfactors=9)
 # fa <- psych::fa.multi(as.matrix(cor), nfactors = 3, nfact2 = 1, n.obs = nrow(random))
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 data <- predict(fa)
@@
 fa2 <- parameters::factor_analysis(data,
-  n = 2,
+  n = 1,
   rotation = "varimax",
@@
 #### Visualization
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 library(ggraph)
@@
   case_when(
-    name == "G1" ~ 0/1,
-    name == "G2" ~ 1/1,
-    name == "ML1" ~ 0/4,
-    name == "ML2" ~ 1/4,
-    name == "ML3" ~ 2/4,
-    name == "ML4" ~ 3/4,
-    name == "ML5" ~ 4/4,
+    name == "G1" ~ 0.5,
+    # name == "G1" ~ 0/3,
+    # name == "G2" ~ 1/3,
+    # name == "G3" ~ 2/3,
+    # name == "G4" ~ 3/3,
+    name == "ML1" ~ 0/(ncol(fa)-3),
+    name == "ML2" ~ 1/(ncol(fa)-3),
+    name == "ML3" ~ 2/(ncol(fa)-3),
+    name == "ML4" ~ 3/(ncol(fa)-3),
+    name == "ML5" ~ 4/(ncol(fa)-3),
+    name == "ML6" ~ 5/(ncol(fa)-3),
+    name == "ML7" ~ 6/(ncol(fa)-3),
+    name == "ML8" ~ 7/(ncol(fa)-3),
+    name == "ML9" ~ 8/(ncol(fa)-3),
+    name == "ML10" ~ 9/(ncol(fa)-3),
+    name == "ML11" ~ 10/(ncol(fa)-3),
+    name == "ML12" ~ 11/(ncol(fa)-3),
     TRUE ~ x)
@@
     hjust = ifelse(name %in% idx, 1.01, 0),
-    name = clean_name(name),
+    name = prettify_parameterName(name),
     Illusion_Type = str_split_fixed(name, " - ", n=2)[, 1]) |> 
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 library(lavaan)
 library(tidySEM)
+library(ggraph)
 
@@
 
-# Model 1
+# Model 1 - by Illusion
 lvl1 <- paste(data$Illusion_Type, "=~", data$Name)
-lvl2 <- paste("G =~", unique(data$Illusion_Type))
+lvl2 <- paste("I =~", unique(data$Illusion_Type))
 resid <- as.data.frame(t(utils::combn(x=unique(data$Illusion_Type), 2)))
-resid <- paste(resid$V1, "~~", resid$V2)[1:2]
-
-m1 <- sem(model = paste0(c(lvl1, lvl2, resid), collapse = "\n"),
-           data  = random)
-
-# Model 2
-lvl1 <- paste(data$Model, "=~", data$Name)
-lvl2 <- paste("G =~", unique(data$Model))
+resid <- paste(resid$V1, "~~", resid$V2)
+
+m1 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
+          data  = random,
+          estimator = "ML",
+          missing = "ml",
+          bounds = TRUE)  # robust.two.stage
+
+# Model 2 - by Parameter
+lvl1 <- paste(data$Parameter, "=~", data$Name)
+lvl2 <- paste("I =~", unique(data$Parameter))
+resid <- as.data.frame(t(utils::combn(x=unique(data$Parameter), 2)))
+resid <- paste(resid$V1, "~~", resid$V2)
 
@@
 
-# Model 3
-lvl1 <- paste(data$Parameter, "=~", data$Name)
-lvl2 <- paste("G =~", unique(data$Parameter))
-
-m3 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
-           data  = random)
-
-anova(m1, m2, m3)
+# # Model 3
+# lvl1 <- paste(data$Model, "=~", data$Name)
+# lvl2 <- paste("G =~", unique(data$Model))
+# 
+# m2 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
+#            data  = random)
+
+anova(m1, m2)
 
 summary(m1, standardize = TRUE)
-
-tidySEM::graph_sem(model = m1)
-
-# as_tbl_graph(fit, standardize = TRUE)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# https://cjvanlissa.github.io/tidySEM/articles/sem_graph.html
+edges <- tidySEM::get_edges(m1) |> 
+  mutate(color = ifelse(sign(as.numeric(est)) >= 0, "1", "-1"),
+         width = abs(as.numeric(est))) |> 
+  filter(lhs != rhs)
+
+nodes <- tidySEM::get_nodes(m1) |> 
+  mutate(label = prettify_parameterName(name),
+         angle = ifelse(grepl("_", name), 90, 0),
+         hjust = ifelse(grepl("_", name), 1.01, 0))
+
+tidygraph::tbl_graph(nodes = nodes,
+                     edges = edges) |> 
+  ggraph(layout = "sugiyama") +
+  geom_edge_link(aes(edge_width = width, 
+                             color = color,
+                             label=est_sig_std), angle_calc="along", label_dodge=unit(-0.015, "npc")) +
+  geom_node_point() +
+  geom_node_text(aes(label = label, angle = angle, hjust=hjust)) +
+  scale_edge_color_manual(values=c("#F44336", "#2196F3"))  +
+  scale_y_continuous(expand = expansion(c(.6, .1))) +
+  scale_edge_width_continuous(range=c(0.1, 1.5)) +
+  guides(edge_colour = "none", edge_width = "none") +
+  theme_graph()
+
+dfsem <- mutate(as.data.frame(predict(m1)), Participant = random$Participant)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+r <- correlation(arrange(dfsem, Participant), arrange(empirical, Participant), sort=TRUE, p_adjust="none")
+
+r |>
+  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
+  ggplot(aes(x = Parameter1, y = Parameter2)) +
+  geom_tile(aes(fill = r)) +
+  geom_text(aes(label=label)) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 # Save
-dfsub <- dfsub |> 
-  merge(random, by = "Participant") |> 
-  cbind(data, predict(fa2, names=c("G1", "G2")))
-
-write.csv("../data/study3.csv")
-```
+dfsub |> 
+  merge(dfsem, by = "Participant") |> 
+  merge(empirical, by = "Participant") |> 
+  write.csv("../data/study3.csv", row.names = FALSE)
+```

### 2022-08-27 study3/study3.Rmd
update

---
title: "Illusion Game Validation (Study 3)"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(
  digits = 3,
  mc.cores = 4,
  brms.algorithm = "sampling",
  brms.backend = "cmdstanr"
)

cache <- TRUE
runModels <- FALSE
bestModels <- FALSE
fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

# Introduction

The goal of study 3 was to...

We have to keep in mind that individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of ...), 2) the structural equation model from which they are obtained is also constructed on layers of abstractions. Though, as we have seen, these indices to relate to empirical scores, such as the average error rate and RT in the task.

# Methods

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

# Variables
DVs <- c("Delboeuf" = "#2196F3", 
         "Ebbinghaus" = "#3F51B5", 
         "RodFrame" = "#F44336", 
         "VerticalHorizontal" = "#FF5722", 
         "Zollner" = "#FF9800", 
         "White" = "#9E9E9E", 
         "MullerLyer" = "#4CAF50", 
         "Ponzo" = "#009688", 
         "Poggendorff" = "#795548", 
         "Contrast" = "#607D8B", 
         "I" = "#9C27B0")

# Load data
dfsub <- read.csv("../data/study3.csv") |>
  mutate(
    Screen_Refresh = as.numeric(Screen_Refresh),
    Education = fct_relevel(Education, "High School", "Bachelor", "Master", "Doctorate", "Other", "Prefer not to Say")
  ) |> 
  datawizard::standardise(select=names(DVs))

dflong <- dfsub |> 
  pivot_longer(all_of(names(DVs)), names_to = "Index", values_to = "Score")
# plot(estimate_density(dfsub[DVs]))
```

## Results

### Contextual

#### Screen Size

```{r message=FALSE, warning=FALSE, cache=cache}
preds <- data.frame()
for(i in names(DVs)) {
  model <- brms::brm(paste0(i, " ~ s(Screen_Size)"), data=dfsub, algorithm = "meanfield", refresh=0)
  param <- bayestestR::describe_posterior(as.data.frame(model))

  pred <- estimate_relation(model, length=50)
  pred$Index <- i
  pred$sig <- estimate_slopes(model, trend = "Screen_Size")$pd
  preds <- rbind(preds, pred)

  if(any(pred$sig > .95)) {
    print(i)
  }
}

preds |> 
  mutate(sig = format_pd(sig, stars_only = TRUE),
         sig = ifelse(sig == "", "NS", sig)) |> 
  ggplot(aes(x = Screen_Size, y = Predicted)) +
  geom_point2(data=dflong, aes(y = Score, color= Index), alpha=0.1, size=2) +
  geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid")) +
  scale_color_manual(values = DVs) +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  theme_modern() +
  labs(y = "Illusion Sensitivity", x = "Screen Size") +
  ggside::geom_xsidedensity(data=dfsub, fill ="grey", color = "white") +
  ggside::geom_ysidedensity(data=dflong, aes(y=Score, color = Index)) +
  ggside::theme_ggside_void() 
```

#### Refresh Rate

```{r message=FALSE, warning=FALSE, cache=cache}
sig <- list()
preds <- data.frame()
for(i in DVs) {
  model <- brms::brm(paste0(i, " ~ poly(Screen_Refresh, 2)"), data=dfsub, algorithm = "meanfield", refresh=0)
  param <- bayestestR::describe_posterior(as.data.frame(model))

  pred <- estimate_relation(model, length=50)
  pred$Index <- i
  pred$sig <- max(param$pd[2:3])
  preds <- rbind(preds, pred)

  if(any(max(param$pd[2:3]) > .95)) {
    print(i)
  }
}

preds |> 
  mutate(sig = format_pd(sig, stars_only = TRUE),
         sig = ifelse(sig == "", "NS", sig)) |> 
  ggplot(aes(x = Screen_Refresh, y = Predicted)) +
  geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
```

### Demographics

#### Sex

```{r message=FALSE, warning=FALSE, cache=cache}
sig <- list()
params <- data.frame()
for(i in DVs) {
  model <- lm(paste0(i, " ~ Sex"), data=dfsub)
  param <- parameters::parameters(model)
  param$Index <- i
  param$ymiddle <- param$Coefficient[1] + diff(param$Coefficient) / 2
  params <- rbind(params, as.data.frame(param[2, ]))
  if(param$p[2] < .05) {
    print(i)
  }
}

data <- dfsub |> 
  pivot_longer(all_of(DVs), names_to = "Index", values_to = "Score") |> 
  mutate(Index = fct_relevel(Index, arrange(params, p)$Index))

data |> 
  ggplot(aes(x = Index, y = Score)) +
  stat_slab(data=filter(data, Sex == "Male"), aes(fill = Sex), side = "right", scale = 0.5, position = "dodge") +
  stat_slab(data=filter(data, Sex == "Female"), aes(fill = Sex), side = "left", scale = 0.5, position = "dodge") +
  stat_pointinterval(aes(group = Sex), point_interval = "mean_qi",, position = "dodge") +
  geom_text(data = mutate(params, sig = insight::format_p(p, stars_only = TRUE)), aes(label = sig, y = ymiddle))
```

#### Ethnicity

```{r message=FALSE, warning=FALSE, cache=cache}
sig <- list()
params <- data.frame()
for(i in DVs) {
  model <- lm(paste0(i, " ~ Ethnicity"), data=dfsub)
  param <- estimate_contrasts(model, contrast = "Ethnicity")
  param$Index <- i
  params <- rbind(params, as.data.frame(param[2, ]))
  if(any(param$p < .05)) {
    print(i)
  }
}
```

#### Education

```{r message=FALSE, warning=FALSE, cache=cache}
preds <- data.frame()
for(i in DVs) {
  model <- brms::brm(paste0(i, " ~ mo(as.numeric(Education))"), data=filter(dfsub, !Education %in% c("Other", "Prefer not to Say")), algorithm = "meanfield", refresh=0)
  param <- bayestestR::describe_posterior(as.data.frame(model)["bsp_moas.numericEducation"])

  pred <- estimate_relation(model)
  pred$Index <- i
  pred$sig <- param$pd
  preds <- rbind(preds, pred)

  if(any(param$pd > .95)) {
    print(i)
  }
}

preds |> 
  mutate(sig = format_pd(sig, stars_only = TRUE),
         sig = ifelse(sig == "", "NS", sig)) |> 
  ggplot(aes(x = Education, y = Predicted)) +
  geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
```

#### Age

```{r message=FALSE, warning=FALSE, cache=cache}
sig <- list()
preds <- data.frame()
for(i in DVs) {
  model <- brms::brm(paste0(i, " ~ poly(Age, 2)"), data=dfsub, algorithm = "meanfield", refresh=0)
  param <- bayestestR::describe_posterior(as.data.frame(model))

  pred <- estimate_relation(model, length=50)
  pred$Index <- i
  pred$sig <- max(param$pd[2:3])
  preds <- rbind(preds, pred)

  if(any(max(param$pd[2:3]) > .95)) {
    print(i)
  }
}

preds |> 
  mutate(sig = format_pd(sig, stars_only = TRUE),
         sig = ifelse(sig == "", "NS", sig)) |> 
  ggplot(aes(x = Age, y = Predicted)) +
  geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
```

### Personality

```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
r <- correlation::correlation(
  datawizard::data_select(dfsub, DVs),
  select(dfsub, starts_with("IPIP") & !ends_with("_SD")),
  p_adjust = "none",
  sort=TRUE
)

r |>
  mutate(label = paste0(format_value(r, zap_small = TRUE), insight::format_p(p, stars_only=TRUE))) |> 
  ggplot(aes(x = Parameter2, y = Parameter1)) +
  geom_tile(aes(fill = r)) +
  geom_text(aes(label=label), size=3) +
  scale_alpha_continuous(range = c(1, 0.5)) +
  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
  # facet_wrap(~Illusion_Type, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

### 2022-08-27 figures/make_figure1.py

@@ -1,0 +1,17 @@
+import pyllusion as ill
+
+ill.MullerLyer(illusion_strength=45, difference=0).to_image(width=500, height=500).save(
+    "MullerLyer_0.png"
+)
+ill.MullerLyer(illusion_strength=60, difference=1).to_image(width=500, height=500).save(
+    "MullerLyer_1.png"
+)
+ill.MullerLyer(illusion_strength=-60, difference=1).to_image(width=500, height=500).save(
+    "MullerLyer_2.png"
+)
+ill.MullerLyer(illusion_strength=10, difference=0.25).to_image(width=500, height=500).save(
+    "MullerLyer_3.png"
+)
+ill.MullerLyer(illusion_strength=-10, difference=0.25).to_image(width=500, height=500).save(
+    "MullerLyer_4.png"
+)

### 2022-08-27 study2/functions.R

@@ -206,13 +206,13 @@ plot_descriptive_err <- function(data, side = "leftright") {
 
 
 
-plot_model_err <- function(data, model) {
+plot_model_err <- function(data, model, gam=NULL) {
   data <- mutate(data, .dots_side = ifelse(Error == 1, "bottom", "top"))
 
   # Get variables
   vars <- insight::find_predictors(model)$conditional
-  vardiff <- vars[1]
-  varstrength <- vars[2]
+  vardiff <- vars[2]
+  varstrength <- vars[3]
 
   # Get predicted
   pred <- estimate_relation(model,
@@ -240,6 +240,8 @@ plot_model_err <- function(data, model) {
   # Remove first stim of strength 0
   data <- filter(data, !(Block == 1 & Illusion_Strength == 0))
 
+  gamdata <- estimate_relation(gam, length = c(NA, 100))
+
   pred |>
     ggplot(aes_string(x = varstrength, y = "Predicted")) +
     geom_dots(
@@ -256,19 +258,24 @@ plot_model_err <- function(data, model) {
       color = NA,
       alpha = 0.5
     ) +
-    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
+    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff),
+                data=gamdata,
+                alpha = 0.2) +
     geom_vline(xintercept = 0, linetype = "dashed") +
     geom_hline(yintercept = c(0.5), linetype = "dotted", alpha = 0.5) +
-    geom_line(aes_string(color = vardiff, group = vardiff)) +
+    geom_line(aes_string(color = vardiff, group = vardiff), data = gamdata) +
+    geom_line(aes_string(color = vardiff, group = vardiff), linetype="dashed", size=0.5) +
     scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
     scale_x_continuous(expand = c(0, 0)) +
     scale_color_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
     scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
     coord_cartesian(xlim = c(min(data[[varstrength]]), max(data[[varstrength]]))) +
     theme_modern() +
[…]

### 2022-08-27 study2/preprocessing.R

@@ -95,10 +95,14 @@ preprocess_raw <- function(file) {
   # Correct duration
   df$Duration <- df$Duration - df$Break_Duration
 
-  # Manual fixes
-  # df[df$Sex == "Prefer not to say", "Participant"]
+  # Manual fixes for reported errors (using demographic info from prolific)
+  # df[df$Sex == "Other", "Participant"]
   df[df$Participant == "5d3c6e745602310001bca8aa_6bdtb", "Sex"] <- "Female"
   df[df$Participant == "60f5fa488f7b1381d175ecb5_qwvfs", "Sex"] <- "Male"
+  df[df$Participant == "5d558c262d8a6200172a8b7b_deegh", "Sex"] <- "Male"
+  df[df$Participant == "614a4c9d8172751921452626_mc4l4", "Sex"] <- "Female"
+  df[df$Participant == "61323c38f727b9f84e45a9ef_ls925", "Sex"] <- "Female"
+  df[df$Participant == "61568a59110d8f1cc99558af_8tqre", "Sex"] <- "Female"
 
 
   # Format names
@@ -117,14 +121,15 @@ preprocess_raw <- function(file) {
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Hispanic", "Hisapanic"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Mixed/Latino", "Mexicano"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Hispanic/Latino ", "Latin American "), "Latino", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latin", "Latina", "HISPANIC"), "Latino", df$Ethnicity)
+  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latin", "Latina", "HISPANIC", "Mexican"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latino American", "Latino/Hispanic"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White Middle European (Slavic)", "Greek"), "Caucasian", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White - Caucasian", " Caucasian"), "Caucasian", df$Ethnicity)
[…]

### 2022-08-31 manuscript.Rmd
update

@@
 ---
-title             : "**The Illusion Game: A Novel Experimental Paradigm to study Visual Illusions**"
+title             : "**The Illusion Game: A Novel Experimental Paradigm to Study Visual Illusions**"
 shorttitle        : "Illusion Game"
@@
     affiliation   : "1"
-    corresponding : no    # Define only one corresponding author
+    corresponding : yes    # Define only one corresponding author
     address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
-    email         : "dmakowski@ntu.edu.sg"
+    email         : "dom.makowski@gmail.com"
     orcid         : 0000-0001-5375-9967
@@
 abstract: |
-  Abstract abstract.
+  Abstract abstract. We invite researchers to re-analyze our open-access data to provide complimentary evidence (or absence thereof) on the effect and structure of illusion sensitivity.
 
@@
 
-## Introduction
+# Introduction
 
@@
 <!-- Goals -->
-In the present set of studies, we will first test this novel paradigm by investigate if the effect of illusion and task difficulty can be manipulated continuously and separated statistically.
-Then, we will use the paradigm to assess whether 10 different illusions share a common latent factor (a long-standing debate).
+In the present set of preregistered studies, we will first test this novel paradigm by investigate if the effect of illusion and task difficulty can be manipulated continuously and separated statistically.
+Then, we will use the paradigm to assess whether 10 different historical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate).
 Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
 
-In line with our full open-access standards, all the material (stimuli generation code, experiment code, )
+In line with our full open-access standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at **https://github.com/RealityBending/IllusionGameValidation**.
 
-## Study 1
+# Study 1
 
-### Aim
+## Aim
 
-### Participants
+Study 1 can be seen as a pilot study which goal was to gather some preliminary data to 1) assess whether the stimuli generated by *Pyllusion* behave as expected for each of the 10 illusion types (i.e., increase of task difficulty and illusion strength leading to increased error rate) and 2) develop an intuition of the magnitude of effects, to then refine the stimuli parameters to a sensible range (i.e., not overly easy and not impossible hard).
 
-### Procedure
+## Procedure
 
-### Results
+We generated 56 stimuli for each of the 10 illusion types.
+These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where the negative values correspond to congruent illusion strengths).
 
-### Discussion
+The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the block was also randomized.
+After the first series of 10 blocks, another series was done (with new randomized blocks and trials order).
+In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
+The task was implemented using *jsPsych* (*REF*).
+See instructions for each illusion type in the experiment code.
 
-## General Discussion
+## Participants
 
-## Future Directions
+Fifty-two participants were recruited via *Prolific* (**REF** about prolific and their data quality) with no restrictions on age or country (the only requisite was a "fluent" level of English to match the language of instructions), with a reward of about \textsterling 7.5.
+
+We removed 6 participants upon inspection of the average error rage (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
+For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that instructions got mixed up; e.g., participants were selecting the shorter line instead of the longer one).
+Finally, we removed 692 (1.37%) trials based on an implausibly short or wrong response time (< 150 ms or > 3000 ms).
+
+The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: [19, 60]; Sex: 39.1% females, 56.5% males).
+
+## Data Analysis
+
+The analysis of study 1 focused on the probability of errors as the main outcome variable.
+For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some intuition on the underlying generative model.
+Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
+We then fitted the best performing model under a Bayesian model, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships (at the expense of model simplicity).
+
+The analysis was carried out using *R 4.2* (**ref**), *brms* (*REF*), the *tidyverse* (*JOSS REF*) and the *easystats* collection of packages (*REFS*).
+
+## Results
+
+The statistical models suggested that the effect of task difficulty had a cube relationship with error rate for Delboeuf and Ebbinghaus illusions (both based on circle size), square for Rod and Frame, Vertical-Horizontal, cube for Zöllner and Poggendorff, exponential for White, cube for Müller-Lyer and Ponzo (both based on lie lengths), and linear for Contrast.
+See details and figures in the analysis script.
+
+## Discussion
+
+This study allowed to elaborate a more precise prior idea regarding the magnitude of the effects at stake and the type of interaction between them.
+It allowed us to better understand and test-out the stimuli generated by *Pyllusion*.
+It allowed us to uncover a few bugs and issues (for instance, the specification direction of the illusion strength was reversed in the software), which were fixed in a new release.
+
+Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize the information gain.
+In most illusions, the task difficulty exhibited a non-linear effect, which is in line with the psychophysics literature on perceptual decisions.
+One notable result was the illusion effect for the Zöller illusion, which suggested a non-linear pattern.
+By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.
+
+# Study 2
+
+## Aim
+
+The aim of study 2 was two-fold.
+In the first part, we will carefully model the error rate and the reaction time of each illusion type in order to validate our novel paradigm and show that the effect of illusions can be manipulated parametrically.
+In the second part, we will derive the participant-level scores from the models (i.e., the effect of illusion strength, task difficulty and their interaction for each individual) and analyze their latent factors structure.
+
+## Procedure
+
+The paradigm was similar to that of study 1, with a few changes: stimuli were re-generating with within a refined space of parameters.
+Moreover, based on study 1, we used non-linearly spaced difficulty levels, depending on the best underlying model (i.e., with an exponential, square or cubic spacing depending on the relationship).
+For instance, a linear space of [0.1, 0.3, 0.5, 0.7] can be transformed to an exponential space of [0.1, 0.27, 0.47, 0.7].
+
+One additional difference was that instead of repeating two times each stimulus, we generated more versions (with more levels of difficulty and illusion strength).
+For each illusion type, we generated 134 stimuli, that were split into two groups (67 stimuli per illusion block).
+Additionally, instead of a simple break screen, we added two personality questionnaires between the two series of 10 blocks of illusions (see study 3).
+
+## Participants
+
+Using the same recruitment procedure, we recruited 256 participants, from which 6 were identified as outliers and discarded, leading to a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
+See the next study for the full demographic breakdown.
+We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials based on response time (< 125 ms or > 4 SD above mean).
+
+## Data Analysis
+
+The first part of the analysis focuses on modelling the effect of illusion strength and task difficulty, on errors and reaction time, within each illusion.
+In order to achieve that, we will start by fitting a General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
+Errors were analyzed using Bayesian binomial mixed models, and reaction times (RT) of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].
+
+Using the GAMs as a "ground-truth", we will attempt at approximating them using general linear models, which have the advantage  which we can specify a participant-level variability of the effects (via random effects).
+Firstly, we will do a model comparison between a combination of transformations (raw, log, square root or cubic root) on the main predictors (task difficulty *diff* and illusion *strength*).
+We will then select and fit the best models (best on their indices of fit), and compare their predictions visually (see **Figure 2**).
+
+We will then extract the inter-individual variability (specified as random slopes) in the effect of illusion strength, task difficulty and their interaction and use it as participant-level scores.
+Finally, We will explore the relationship across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
+
+## Results
+
+The best models were $log(diff)*strength$ for Delboeuf;
+
+Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, dynamics, we decided not to use add them for the individual scores analysis.
+Additionally, it would have increased the parameter space from 30 to 90 variables, which was deemed as statistically unreasonable given our sample size).
+
+## Discussion
+
+Future studies should investigate reaction times to find the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
+
+# Study 3
+
+## Aim
+
+## Procedure
+
+This study was based on the data collected in study 2.
+However, the variables of interest were here taken from the questionnaires inserted in between the two series of illusion blocks.
+We used the *IPIP6* (x items; *REF*) to measure 6 normal personality traits (extraversion, **complete**).
+The PID-5 (x items; *REF*) was used to measure psychopathological personality traits (..., **complete**).
+The participants were the same as in study 2, however due to a technical issue the first height participants do not have data for the personality questionnaires.
+
+# General Discussion
+
+# Future Directions
 
@@
 
-## References
+# References
 

### 2022-08-31 manuscript.tex
update

@@
 \hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}}
 
@@
 
-In the present set of studies, we will first test this novel paradigm by investigate if the effect of illusion can be manipulated continuously.
-Then, we will use it to assess whether 10 different illusions share a common latent factor (a long-standing debate), 3) how the the inter-individual sensitivity to illusions relates to other characteristics, such as sex or personality.
+In the present set of preregistered studies, we will first test this novel paradigm by investigate if the effect of illusion and task difficulty can be manipulated continuously and separated statistically.
+Then, we will use the paradigm to assess whether 10 different historical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate).
+Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
+
+In line with our full open-access standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at \textbf{\url{https://github.com/RealityBending/IllusionGameValidation}}.
 
 \hypertarget{study-1}{%
-\subsection{Study 1}\label{study-1}}
+\section{Study 1}\label{study-1}}
 
 \hypertarget{aim}{%
-\subsubsection{Aim}\label{aim}}
+\subsection{Aim}\label{aim}}
+
+Study 1 can be seen as a pilot study which goal was to gather some preliminary data to 1) assess whether the stimuli generated by \emph{Pyllusion} behave as expected for each of the 10 illusion types (i.e., increase of task difficulty and illusion strength leading to increased error rate) and 2) develop an intuition of the magnitude of effects, to then refine the stimuli parameters to a sensible range (i.e., not overly easy and not impossible hard).
+
+\hypertarget{procedure}{%
+\subsection{Procedure}\label{procedure}}
+
+We generated 56 stimuli for each of the 10 illusion types.
+These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., {[}1, 2, 3, 4, 5, 6, 7{]}, where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., {[}-3, -2, -1, 0, 1, 2, 3{]}, where the negative values correspond to congruent illusion strengths).
+
+The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the block was also randomized.
+After the first series of 10 blocks, another series was done (with new randomized blocks and trials order).
+In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
+See instructions for each illusion type in the experiment code.
 
 \hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
+\subsection{Participants}\label{participants}}
 
-\hypertarget{procedure}{%
-\subsubsection{Procedure}\label{procedure}}
+Fifty-two participants were recruited via \emph{Prolific} (\textbf{REF} about prolific and their data quality) with no restrictions on age or country (the only requisite was a ``fluent'' level of English to match the language of instructions), with a reward of about \textsterling 7.5.
+
+We removed 6 participants upon inspection of the average error rage (when close to 50\%, suggesting random answers), and when the reaction time distribution was implausibly fast.
+For the remaining participants, we discarded blocks where the error rate was higher than 50\% (possibly indicating that instructions got mixed up; e.g., participants were selecting the shorter line instead of the longer one).
+Finally, we removed 692 (1.37\%) trials based on an implausibly short or wrong response time (150 ms \textless{} RT \textless{} 3000 ms).
+
+The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: {[}19, 60{]}; Sex: 39.1\% females, 56.5\% males).
+
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
+
+The analysis of study 1 focused on the probability of errors as the main outcome variable.
+For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some intuition on the underlying generative model.
+Next, we tested the performance of various models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up \emph{vs.} down or left \emph{vs.} right) as an additional predictor.
+We then fitted the best performing model under a Bayesian model, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships (at the expense of model simplicity).
+
+The analysis was carried out using \emph{R 4.2} (\textbf{ref}), \emph{brms} (\emph{REF}), the \emph{tidyverse} (\emph{JOSS REF}) and the \emph{easystats} collection of packages (\emph{REFS}).
 
@@
 
+The statistical models suggested that the effect of task difficulty had a cube relationship with error rate for Delboeuf and Ebbinghaus illusions (both based on circle size), square for Rod and Frame, Vertical-Horizontal, cube for Zöllner and Poggendorff, exponential for White, cube for Müller-Lyer and Ponzo (both based on lie lengths), and linear for Contrast.
+See details and figures in the analysis script.
+
 \hypertarget{discussion}{%
-\subsubsection{Discussion}\label{discussion}}
+\subsection{Discussion}\label{discussion}}
+
+This study allowed to elaborate a more precise prior idea regarding the magnitude of the effects at stake and the type of interaction between them.
+It allowed us to better understand and test-out the stimuli generated by \emph{Pyllusion}.
+It allowed us to uncover a few bugs and issues (for instance, the specification direction of the illusion strength was reversed in the software), which were fixed in a new release.
+
+Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize the information gain.
+One notable result was the illusion effect for the Zöller illusion, which suggested a non-linear pattern.
+By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.
+
+\hypertarget{study-2}{%
+\section{Study 2}\label{study-2}}
+
+\hypertarget{aim-1}{%
+\subsection{Aim}\label{aim-1}}
+
+\hypertarget{procedure-1}{%
+\subsection{Procedure}\label{procedure-1}}
+
+Perceptual decisions often exhibit a non-linear relationship with the objective physical values (e.g., a logarithmic relationship.
+In order to spread the difficulty levels based on the behavioral performance (to maximize the diversity of the stimuli in regards to perceptual processes), we generated, for study 2, stimuli with the difficulty levels spaced non-linearly, depending on the best underlying model (i.e., with an exponential, square or cubic spacing if the best model was \emph{log}, \emph{square root} or \emph{cube root}).
+For instance, a linear space of {[}0.1, 0.3, 0.5, 0.7{]} can be transformed to an exponential space of {[}0.1, 0.27, 0.47, 0.7{]}.
+
+\hypertarget{participants-1}{%
+\subsection{Participants}\label{participants-1}}
 
 \hypertarget{general-discussion}{%
-\subsection{General Discussion}\label{general-discussion}}
+\section{General Discussion}\label{general-discussion}}
+
+\hypertarget{future-directions}{%
+\section{Future Directions}\label{future-directions}}
+
+We strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding further.
 
@@
 \hypertarget{references}{%
-\subsection{References}\label{references}}
+\section{References}\label{references}}
 

### 2022-08-31 study1/study1.Rmd
model
update

@@
         mc.cores = 4,
-        brms.algorithm = "meanfield",
+        brms.algorithm = "sampling",
         brms.backend = "cmdstanr")
 
-cache <- FALSE
+cache <- TRUE
 fig.width <- see::golden_ratio(7)
@@
 
-# Introduction
-
-The goal of study 1 was to...
-
-# Methods
+# Preprocessing
 
@@
 df$Illusion_Strength_cbrt <- sign(df$Illusion_Strength) * (abs(df$Illusion_Strength)**(1/3))
+
+# Anonymize
+replacement <- paste0("S", str_pad(1:length(unique(df$Participant)), 2, pad = "0"))
+names(replacement) <- unique(df$Participant)
+df$Participant <- replacement[df$Participant]
 
@@
   # Prolific Status: REJECTED
-  "60684f29dbfe1bb2059e5e27_rkqoy",
+  "S33",
   # Block n2 with very short RTs
   # Prolific Status: RETURNED
-  "5e860198a846e30497df5189_6e43s",
+  "S20",
   # Error rate of 46.2% and short RTs
   # Prolific Status: RETURNED
-  "615f319bb341cf2f306d858d_qsaq5",
+  "S51",
   # Error rate of 46.2%
   # Prolific Status: RETURNED
-  "614f36fd81c78b7a125c4262_6ax4g",
+  "S49",
   # Error rate of 47.9%
   # Prolific Status: REJECTED
-  "61280140171ec546e87ed8cb_qdlgy",
+  "S47",
   # Error rate of 42.1% and very large RT SD
   # Prolific Status: REJECTED
-  "5d398380b37ab1000111fac3_2nxxh"
+  "S12"
 )
@@
 ```
-
-## Data Analysis
-
-The analysis of study focused on the probability of errors as the main outcome variable.
-For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some insight on the underlying generative model.
-Next, we tested the performance of various models differing in their specifications, such as: with or without a log-transform on the task difficulty, with or without a 2nd order polynomial term of the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
-We then fitted the best performing model under a Bayesian model, and compared its predicted links with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships at the expense of model simplicity.
-Finally, we used these models to estimate different values of task difficulty that would lead to similar error rates (2.5, 5 and 25%) when the illusion strength is null.
 
@@
 
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Parameter Standardization
+
 std_params <- function(model, min=0, max=2) {
@@
 
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Parameter Standardization
+
 # range(data$Illusion_Difference)
@@
 plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Parameter Standardization
+
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0.01, max=12)
+std_params(gam$model, min=0.01, max=12)
+```
+
+## Vertical-Horizontal
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Vertical-Horizontal")
+
+plot_descriptive(data)
+
+data <- filter(data, abs(Illusion_Strength) < 90)
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ sqrt(Illusion_Difference) * Illusion_Strength + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Parameter Standardization
+
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0, max=0.40)
+std_params(gam$model, min=0, max=0.40)
+```
+
+## Zöllner
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Zöllner")
+
+plot_descriptive(data)
+
+data <- filter(data, abs(Illusion_Strength) < 45)
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ cbrt(Illusion_Difference) * Illusion_Strength + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### Parameter Standardization
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Parameter Standardization
+
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0, max=5)
+```
+
+## White
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "White")
+
+plot_descriptive(data)
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+# unique(data$Illusion_Strength)
+
+std_params(model, min=0, max=20)
+std_params(gam$model, min=0, max=20)
+```
+
+## Müller-Lyer
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Müller-Lyer")
+
+plot_descriptive(data, side = "updown")
+```
+
+### Model Selection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+best_models(data)
+```
+
+### Model Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+formula <- brms::bf(
+  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
+    (1 | Participant),
+  family = "bernoulli"
+)
+
+model <- brms::brm(formula,
+  data = data,
+  refresh = 0
+)
+
+# parameters::parameters(model)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+plot_model(data, model)
+```
+
+### GAM
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+gam <- make_gam(data)
+gam$p
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# range(data$Illusion_Difference)
+# range(data$Illusion_Strength)
+
+std_params(model, min=0, max=0.6)
+std_params(gam$model, min=0, max=0.6)
+```
+
+## Ponzo
+
+### Descriptive
+
+```{r message=FALSE, warning=FALSE, cache=cache}
+data <- filter(df, Illusion_Type == "Ponzo")
+
+plot_descriptive(data, side = "updown")
 ```
@@
 
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 # range(data$Illusion_Difference)
@@
 
-std_params(model, min=0.01, max=12)
-std_params(gam$model, min=0.01, max=12)
-```
-
-## Vertical-Horizontal
+std_params(model, min=0, max=20)
+std_params(gam$model, min=0, max=20)
+```
+
+## Poggendorff
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Vertical-Horizontal")
-
-plot_descriptive(data)
-
-data <- filter(data, abs(Illusion_Strength) < 90)
-
-plot_descriptive(data)
+data <- filter(df, Illusion_Type == "Poggendorff")
+
+plot_descriptive(data, side = "updown")
+
+data <- filter(data, abs(Illusion_Strength) < 45)
+
+plot_descriptive(data, side = "updown")
 ```
@@
 formula <- brms::bf(
-  Error ~ sqrt(Illusion_Difference) * Illusion_Strength + 
+  Error ~ cbrt(Illusion_Difference) * poly(Illusion_Strength, 2) + 
     (1 | Participant),
@@
 
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.40)
-std_params(gam$model, min=0, max=0.40)
-```
-
-## Zöllner
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Zöllner")
-
-plot_descriptive(data)
-
-data <- filter(data, abs(Illusion_Strength) < 45)
-
-plot_descriptive(data)
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=5)
-```
-
-## White
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "White")
-
-plot_descriptive(data)
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-# unique(data$Illusion_Strength)
-
-std_params(model, min=0, max=20)
-std_params(gam$model, min=0, max=20)
-```
-
-## Müller-Lyer
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Müller-Lyer")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=0.6)
-std_params(gam$model, min=0, max=0.6)
-```
-
-## Ponzo
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Ponzo")
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ poly(Illusion_Difference, 2) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-# range(data$Illusion_Difference)
-# range(data$Illusion_Strength)
-
-std_params(model, min=0, max=20)
-std_params(gam$model, min=0, max=20)
-```
-
-## Poggendorff
-
-### Descriptive
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-data <- filter(df, Illusion_Type == "Poggendorff")
-
-plot_descriptive(data, side = "updown")
-
-data <- filter(data, abs(Illusion_Strength) < 45)
-
-plot_descriptive(data, side = "updown")
-```
-
-### Model Selection
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-best_models(data)
-```
-
-### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-formula <- brms::bf(
-  Error ~ log(Illusion_Difference) * poly(Illusion_Strength, 2) + 
-    (1 | Participant),
-  family = "bernoulli"
-)
-
-model <- brms::brm(formula,
-  data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-plot_model(data, model)
-```
-
-### GAM
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-gam <- make_gam(data)
-gam$p
-```
-
-### Parameter Standardization
-
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 # range(data$Illusion_Difference)
@@
 ```
-
-# Discussion
-
-This study allowed to elaborate a more precise prior idea regarding the magnitude of the effects at stake and the type of interaction between them.
-Crucially, it allowed us to refine the range of task difficulty and illusion strength for the stimuli of the next study to maximize the information gain.
-
-One thing that stood out was Zöller, which seems to exhibit a non-linear relationship of illusion effect.
-Next study, by generating a wider range of illusion strength, will attempt at clarifying this point.
-
-Perceptual decisions often exhibit a non-linear relationship with the objective physical values (e.g., a logarithmic relationship.
-In order to spread the difficulty levels based on the behavioral performance (to maximize the diversity of the stimuli in regards to perceptual processes), we generated, for study 2, stimuli with the difficulty levels spaced non-linearly, depending on the best underlying model (i.e., with an exponential, square or cubic spacing if the best model was *log*, *square root* or *cube root*).
-For instance, a linear space of [0.1, 0.3, 0.5, 0.7] can be transformed to an exponential space of [0.1, 0.27, 0.47, 0.7].

### 2022-08-31 study2/study2.Rmd
update

@@
 runModels <- FALSE
-bestModels <- FALSE
+bestModels <- TRUE
 fig.width <- see::golden_ratio(7)
@@
 
-# Introduction
-
-The goal of study 2 was to...
-
-# Methods
+# Preprocessing
 
@@
 
+# Anonymize
+replacement <- paste0("S", str_pad(1:length(unique(df$Participant)), 2, pad = "0"))
+names(replacement) <- unique(df$Participant)
+df$Participant <- replacement[df$Participant]
+
 # Save anonymized data
-write.csv(df, "../data/study2.csv", row.names = FALSE)
+write.csv(df[1:150000, ], "../data/study2_part1.csv", row.names = FALSE)
+write.csv(df[150001:nrow(df), ], "../data/study2_part2.csv", row.names = FALSE)
 ```
@@
 
-df <- read.csv("../data/study2.csv") |>
+df <- read.csv("../data/study2_part1.csv") |>
+  rbind(read.csv("../data/study2_part2.csv")) |> 
   mutate(
@@
 
-dfsub$Participant[grepl("60be8d", dfsub$Participant)]
-
 outliers <- c(
@@
   # Prolific Status: REJECTED (06/08)
-  "5e66ed53de7896486f9a71f8_p8ckk",
+  "S46",
   # 2nd block of responses very fast
   # Prolific Status: REJECTED (15/08)
-  "61443e47c248a375c899a0cf_9qz3u",
+  "S221",
   # Error rate of 44% and very short RTs
   # Prolific Status: RETURN REQUESTED (22/08)
-  "60dd626e212aa26c16d953b7_vw35k",
+  "S154",
   # 2nd block bad, first block 1/3 bad
   # Prolific Status: RETURN REQUESTED (26/08)
-  "5eed37c13661d91555c38308_hpgnt",
+  "S68",
   # Prolific Status: RETURN REQUESTED (26/08)
-  "6159f8a9a2d80ecf1a5acd30_zc8rt",
+  "S238",
   # Prolific status: accepted (not enough proof)
-  "6136aa846e80ebd7daf6e148_aggf6"
+  "S201"
 )
@@
   # 2nd block a bit bad
-  "5c43cd414fe4f800016e4983_0zp37", 
+  "S22", 
   # Entire 2nd block bad
-  "61564e378e974bdbe42763a2_hhufm",
+  "S235",
   # Entire 2nd block bad
-  "5ff8b21dd2ca6f12bdaa5be8_c5rs6",
+  "S107",
   # Half of 2nd block bad
-  "613a9193ed68e49209b471b7_wrymn",
+  "S204",
   # 2nd block not good
-  "60be8dd61cec210d627483de_cmfx4")
+  "S140")
 ```
@@
 
-### Error Rate
+### Summary
 
@@
   arrange(desc(Error))
-
+```
+
+See the next tab for the descriptive table.
+
+### Descriptive Table
+
+```{r message=FALSE, warning=FALSE}
 data.frame(Participant = c("Total"), t(sapply(dfsub[2:ncol(dfsub)], mean, na.rm=TRUE))) |> 
@@
 
-```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=20}
+```{r outliers_RT, message=FALSE, warning=FALSE, fig.width=20, fig.height=20, cache=cache}
 # RT distribution
 p <- df |> 
-  # filter(Date > "2022-08-24") |>
   filter(RT < 10000) |> 
@@
   labs(y = "", x = "Reaction Time (ms)")
-# p
+p
 # ggsave("figures/outliers.png", p, width=25, height=25)
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r outliers_errorblock, message=FALSE, warning=FALSE}
 temp <- df |>
@@
 # Drop
+n <- nrow(df)
 df <- df |>
@@
 
+We removed `r n-nrow(df)` (`r insight::format_value((n-nrow(df)) / n, as_percent=TRUE)`) trials belonging to bad blocks.
+
 ### Reaction Time per Trial
 
-```{r message=FALSE, warning=FALSE}
+```{r outliers_RTtrials, message=FALSE, warning=FALSE}
 df <- df |>
@@
 
-```{r }
-plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", fill = "orange") {
-  dfsub |>
-    ggplot(aes_string(x = what)) +
-    geom_density(fill = fill) +
-    geom_vline(xintercept = mean(dfsub[[what]]), color = "red", linetype = "dashed") +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_y_continuous(expand = c(0, 0)) +
-    ggtitle(title, subtitle = subtitle) +
-    theme_modern() +
-    theme(
-      plot.title = element_text(face = "bold", hjust = 0.5),
-      plot.subtitle = element_text(face = "italic", hjust = 0.5),
-      axis.title.x = element_blank(),
-      axis.title.y = element_blank(),
-      axis.text.y = element_blank()
-    )
-}
-
-plot_waffle <- function(dfsub, what = "Nationality", title = what) {
-  ggwaffle::waffle_iron(dfsub, what) |>
-    # mutate(label = emojifont::fontawesome('fa-twitter')) |>
-    ggplot(aes(x, y, fill = group)) +
-    ggwaffle::geom_waffle() +
-    # geom_point() +
-    # geom_text(aes(label=label), family='fontawesome-webfont', size=4) +
-    coord_equal() +
-    ggtitle(title) +
-    labs(fill = "") +
-    theme_void() +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
-}
-```
-
-```{r fig.width=20, fig.height=15}
+```{r demographics, fig.width=20, fig.height=15}
 p1 <- plot_distribution(dfsub, "Age", fill = "#FF9800")
 p2 <- plot_distribution(dfsub, "Duration", title = "Total Duration", subtitle = "in minutes", fill = "#F44336")
-p3 <- plot_distribution(dfsub, "Break_Duration", title = "Break Duration", subtitle = "in minutes", fill = "#3F51B5")
+p3 <- plot_distribution(dfsub[dfsub$Break_Duration < 3, ], "Break_Duration", title = "Break Duration", subtitle = "in minutes", fill = "#3F51B5")
 
 p4 <- plot_waffle(dfsub, "Sex") +
-  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))
+  scale_color_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))
 
 p5 <- plot_waffle(dfsub, "Education") +
-  scale_fill_viridis_d()
+  scale_color_viridis_d()
 
 p6 <- plot_waffle(dfsub, "Nationality") +
-  scale_fill_manual(values = c("South American" = "#FF5722", "Middle Eastern" = "#FF9800", "Western" = "#2196F3", "Eastern" = "#673AB7", "African" = "#4CAF50", "South African" = "#009688"))
+  scale_color_material_d()
+  # scale_fill_manual(values = c("South American" = "#FF5722", "Middle Eastern" = "#FF9800", "Western" = "#2196F3", "Eastern" = "#673AB7", "African" = "#4CAF50", "South African" = "#009688"))
 
 p7 <- plot_waffle(dfsub, "Ethnicity") +
-  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0", "Mixed" = "#795548"))
+  scale_color_manual(values = c("Latino" = "#FF5722", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Other" = "#795548"))
 
 p8 <- plot_waffle(dfsub, "Screen_Resolution", title = "Screen Resolution") +
-  scale_fill_pizza_d() +
+  scale_color_pizza_d() +
   guides(fill = "none")
@@
 p9 <- plot_waffle(dfsub, "Device_OS", title = "Device OS") +
-  scale_fill_bluebrown_d()
+  scale_color_bluebrown_d()
 
@@
 
-(p1 / p2 / p3) | (p4 / p5 / p9) | (p6 / p7 / p8)
-```
-
-## Data Analysis
-
-In the first part of the study focuses on modelling the effect of illusion strength and task difficulty, on errors and reaction time, within each illusion.
-In order to achieve that, we will start by fitting a General Additive Models (GAM), which can accommodate possible non-linear effects and interactions.
-Using them as a "ground-truth", we will attempt at approximating them using general linear models, in which we can specify a participant-level variability of the effects (via random effects).
-
-We will then extract this participant-level variability in the effect of illusion strength, task difficulty and their interaction and use it as inter-individual features.
-We will estimate the relationship across different illusions using factor analysis (FA) and structural equation modelling (SEM).
-
-```{r make_GAMs, message=FALSE, warning=FALSE, eval=runModels}
+(p1 / p2) | (p4 / p5 / p9) | (p6 / p7 / p8)
+```
+
+# Results
+
+```{r test_ISI, eval=FALSE, message=FALSE, warning=FALSE}
+# Test where ISI needs to be included
+rez <- data.frame()
+for(ill in c("Delboeuf", 
+             "Ebbinghaus", 
+             "Rod-Frame",
+             "Vertical-Horizontal", 
+             "Zöllner",
+             "White",
+             "Müller-Lyer", 
+             "Ponzo", 
+             "Poggendorff",
+             "Contrast")) {
+  print(ill)
+  print("===================")
+  m1 <- brms::brm(
+    Error ~ poly(ISI, 2) + (poly(ISI, 2) | Participant),
+    data = filter(df, Illusion_Type == ill),
+    family = "bernoulli",
+    algorithm="meanfield"
+  )
+  m2 <- brms::brm(
+    brms::bf(
+      RT ~ poly(ISI, 2) + (1 | Participant),
+      sigma ~ poly(ISI, 2),
+      beta ~ poly(ISI, 2)
+    ),
+    data = filter(df, Illusion_Type == ill, Error == 0),
+    family = "exgaussian",
+    init=0,
+    iter = 4000,
+    algorithm="meanfield"
+  )
+  rez <- rbind(
+    bayestestR::describe_posterior(as.data.frame(m1))[2:3, ],
+    bayestestR::describe_posterior(as.data.frame(m2))[4:9, ]
+  ) |> 
+    mutate(Illusion = ill) |> 
+    rbind(rez)
+}
+```
+
+```{r make_GAMs, eval=FALSE, message=FALSE, warning=FALSE}
 for(ill in c("Delboeuf", 
@@
     brms::bf(
-      RT ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + (1 | Participant),
-      sigma ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + (1 | Participant),
-      beta ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + (1 | Participant)
+      RT ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + poly(ISI, 2) + (1 | Participant),
+      sigma ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + poly(ISI, 2) + (1 | Participant),
+      beta ~ t2(Illusion_Difference, Illusion_Strength, k=8, bs="cr") + poly(ISI, 2) + (1 | Participant)
     ),
@@
 
-# Results
-
 ## Models {.tabset}
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_delboeuf, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
 
-m1 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect + (1 | Participant), data = data, family = "binomial")
-m2 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * sqrtmod(Illusion_Strength) + (1 | Participant), data = data, family = "binomial")
-
-compare_performance(m1, m2, metrics = c("R2", "BIC"))
-test_performance(m1, m2)
+# m1 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)) * Illusion_Effect + (1 | Participant), data = data, family = "binomial")
+# m2 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * sqrtmod(Illusion_Strength) + (1 | Participant), data = data, family = "binomial")
+# 
+# compare_performance(m1, m2, metrics = c("R2", "BIC"))
+# test_performance(m1, m2)
 ```
@@
   data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model_delboeuf_err)
+  refresh = 100
+)
 
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r message=FALSE, warning=FALSE, eval=TRUE}
 load("models/model_delboeuf_err.RData")
 load("models/gam_delboeuf_err.RData")
+
+# parameters::parameters(model_delboeuf_err)
 ```
@@
 formula <- brms::bf(
-  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength)))  +
-    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
-  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + (1 | Participant),
-  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + (1 | Participant),
+  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) +  
+    poly(ISI, 2) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) +  
+    poly(ISI, 2) + (1 | Participant),
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) +  
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
 # plot(seq(-5, 5, length.out=100), brms::dstudent_t(seq(-5, 5, length.out=100), df = 3, 0.1, 0.5))
-priors <- c(
+# priors <- c(
   # Intercept
@@
   # prior("student_t(3, 0.1, 0.5)", class = "ndt")
-  ) |> 
-  brms::validate_prior(formula, data = data)
+  # ) |> 
+  # brms::validate_prior(formula, data = data)
 
@@
   data = data,
-  refresh = 50,
+  refresh = 100,
   init=0
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_delboeuf_rt.RData")
@@
 
-plot_ppcheck(model_delboeuf_rt)
 # performance::check_model(model_delboeuf_rt)
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+plot_ppcheck(model_delboeuf_rt, gam_delboeuf_rt)
+
 p_delboeuf_rt <- plot_model_rt(data, model_delboeuf_rt, gam_delboeuf_rt)
@@
 
-m1 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
-    (1 | Participant), data = data, family = "binomial")
-m2 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * logmod(Illusion_Strength) +
-    (1 | Participant), data = data, family = "binomial")
-
-test_performance(m1, m2)
-compare_performance(m1, m2)
+# m1 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)) * Illusion_Effect +
+#     (1 | Participant), data = data, family = "binomial")
+# m2 <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) * logmod(Illusion_Strength) +
+#     (1 | Participant), data = data, family = "binomial")
+# 
+# test_performance(m1, m2)
+# compare_performance(m1, m2)
 ```
@@
 formula <- brms::bf(
-  Error ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) +
-    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) | Participant),
+  Error ~ Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
   family = "bernoulli"
@@
 load("models/gam_ebbinghaus_err.RData")
+
+# parameters::parameters(model_delboeuf_err)
 ```
@@
 formula <- brms::bf(
-  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength)))  +
-    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) | Participant),
-  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) + (1 | Participant),
-  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * logmod(abs(Illusion_Strength))) + (1 | Participant),
+  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) +
+    poly(ISI, 2) + 
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) + 
+    poly(ISI, 2) + (1 | Participant),
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) + 
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
   data = data,
-  refresh = 0,
+  refresh = 100,
   init = 0
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_ebbinghaus_rt.RData")
@@
 
-plot_ppcheck(model_ebbinghaus_rt)
 # performance::check_model(model_delboeuf_rt)
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+plot_ppcheck(model_ebbinghaus_rt, gam_ebbinghaus_rt)
+
 p_ebbinghaus_rt <- plot_model_rt(data, model_ebbinghaus_rt, gam_ebbinghaus_rt)
 p_ebbinghaus_rt
+```
+
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=FALSE}
+p_ebbinghaus <- plot_all(data, p_ebbinghaus_err, p_ebbinghaus_rt)
+p_ebbinghaus
 ```
@@
   RT ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) +
+    poly(ISI, 2) + 
     (1 + Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) | Participant),
-  sigma ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) + (1 | Participant),
-  beta ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) + (1 | Participant),
+  sigma ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) + 
+    poly(ISI, 2) + (1 | Participant),
+  beta ~ Illusion_Effect / (logmod(Illusion_Difference) * logmod(abs(Illusion_Strength))) + 
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_rodframe_rt.RData")
@@
 
-plot_ppcheck(model_rodframe_rt)
 # performance::check_model(model_delboeuf_rt)
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+plot_ppcheck(model_rodframe_rt, gam_rodframe_rt)
+
 p_rodframe_rt <- plot_model_rt(data, model_rodframe_rt, gam_rodframe_rt)
 p_rodframe_rt
+```
+
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=FALSE}
+p_rodframe <- plot_all(data, p_rodframe_err, p_rodframe_rt)
+p_rodframe
 ```
@@
   data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model_verticalhorizontal_err)
+  refresh = 100
+)
 
@@
 load("models/gam_verticalhorizontal_err.RData")
+
+# parameters::parameters(model_verticalhorizontal_err)
+
 ```
@@
   RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    poly(ISI, 2) + 
     (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
-  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + (1 | Participant),
-  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + (1 | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
+    poly(ISI, 2) + (1 | Participant),
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
   RT ~  Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
+    poly(ISI, 2) + 
     (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
   sigma ~  Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
   RT ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
+    poly(ISI, 2) + 
     (1 + Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) | Participant),
   sigma ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   beta ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
   RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    poly(ISI, 2) + 
     (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
   sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
   RT ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
+    poly(ISI, 2) + 
     (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
   sigma ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
   RT ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    poly(ISI, 2) + 
     (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
   sigma ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
   RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    poly(ISI, 2) + 
     (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
   sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
-    (1 | Participant),
+    poly(ISI, 2) + (1 | Participant),
   family = "exgaussian"
@@
     !str_detect(Parameter, "Cong_Prob"),
-    !str_detect(Parameter, "Null_"))
+    !str_detect(Parameter, "Null_"),
+    !str_detect(Parameter, "Diff_"))
 # unique(data$Parameter)
@@
 # plot(estimate_density(select(random, -Participant)))
-```
-
-### Empirical Descriptors
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-# Average for 4 higher level of task difficulty
-empirical <- df |>
-  filter(Illusion_Effect == "Incongruent") |> 
-  group_by(Participant, Illusion_Type, Illusion_Strength) |>
-  arrange(Illusion_Difference) |> 
-  slice(0:4) |> 
-  # select(Illusion_Strength, Illusion_Difference)
-  group_by(Participant, Illusion_Type) |>
-  summarize(
-    TaskHard_Error = sum(Error, na.rm=TRUE) / n(),
-    TaskHard_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
-    # Illu_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
-  ) |>
-  ungroup() |> 
-  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
-  pivot_wider(names_from = c("Illusion_Type"), values_from = c("TaskHard_Error", "TaskHard_RTMean"), names_glue = "{Illusion_Type}_{.value}")
-
-# Average for 4 higher level of illusion strength
-empirical <- df |>
-  filter(Illusion_Effect == "Incongruent") |> 
-  group_by(Participant, Illusion_Type, Illusion_Difference) |>
-  arrange(desc(Illusion_Strength)) |> 
-  slice(0:4) |> 
-  # select(Illusion_Strength, Illusion_Difference)
-  group_by(Participant, Illusion_Type) |>
-  summarize(
-    IlluStrong_Error = sum(Error, na.rm=TRUE) / n(),
-    IlluStrong_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
-    # Diff_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
-  ) |>
-  ungroup() |> 
-  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
-  pivot_wider(names_from = c("Illusion_Type"), values_from = c("IlluStrong_Error", "IlluStrong_RTMean"), names_glue = "{Illusion_Type}_{.value}") |> 
-  full_join(empirical, by = "Participant") 
-```
-
-<!-- ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE} -->
-<!-- rez <- parameters::n_factors(select(empirical, -Participant), n_max=30) -->
-<!-- plot(rez) -->
-
-<!-- r <- correlation::correlation(empirical, p_adjust = "none", sort=TRUE) -->
-
-<!-- r |> -->
-<!--   ggplot(aes(x = Parameter2, y = Parameter1)) + -->
-<!--   geom_tile(aes(fill = r)) + -->
-<!--   # geom_text(aes(label = Text), size = 2) + -->
-<!--   scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") + -->
-<!--   scale_x_discrete(expand = c(0, 0)) + -->
-<!--   scale_y_discrete(expand = c(0, 0)) + -->
-<!--   labs(title = "Correlation Matrix", x = NULL, y = NULL) + -->
-<!--   theme_minimal() + -->
-<!--   theme( -->
-<!--     axis.text.x = element_text(angle = 90, hjust = 1), -->
-<!--     plot.title = element_text(hjust = 0.5, face = "bold"), -->
-<!--     panel.grid.major = element_blank(), -->
-<!--     panel.grid.minor = element_blank() -->
-<!--   ) -->
-
-<!-- fa <- parameters::factor_analysis(select(empirical, -Participant), -->
-<!--   n = 13, -->
-<!--   rotation = "oblimin", -->
-<!--   fm = "mle", -->
-<!--   sort = TRUE -->
-<!-- ) -->
-
-<!-- print(tibble::as_tibble(fa), n=30) -->
-<!-- plot(fa) -->
-<!-- ``` -->
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-# Correlation
-r <- data.frame()
-for(ill in c("Delboeuf", 
-             "Ebbinghaus",
-             "RodFrame",
-             "VerticalHorizontal",
-             "Zollner",
-             "White",
-             "MullerLyer",
-             "Ponzo",
-             "Poggendorff",
-             "Contrast"
-             )) {
-  r <- correlation::correlation(
-    data = random |> 
-      arrange(Participant) |> 
-      data_select(select=ill, regex=TRUE),
-    data2 = empirical |>
-      data_select(select=ill, regex=TRUE),
-    p_adjust="none"
-    ) |> 
-    mutate(Illusion_Type = ill,
-           Parameter1 = prettify_parameterName(Parameter1),
-           Parameter2 = prettify_parameterName(Parameter2)) |> 
-    select(Illusion_Type, Parameter1, Parameter2, r, p) |> 
-    rbind(r)
-}
-
-r |>
-  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
-  ggplot(aes(x = Parameter2, y = Parameter1)) +
-  geom_tile(aes(fill = r)) +
-  geom_text(aes(label=label)) +
-  scale_alpha_continuous(range = c(1, 0.5)) +
-  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
-  # facet_wrap(~Illusion_Type, scales = "free") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
 ```
@@
 
+### Empirical Descriptors
+
+We have to keep in mind that individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of ...), 2) the structural equation model from which they are obtained is also constructed on layers of abstractions.
+Thus, we wanted to test wether these indices relate to some simple empirical scores, such as the average error rate and RT in the task.
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Average for 4 higher level of task difficulty
+empirical <- df |>
+  filter(Illusion_Effect == "Incongruent") |> 
+  group_by(Participant, Illusion_Type, Illusion_Strength) |>
+  arrange(Illusion_Difference) |> 
+  slice(0:4) |> 
+  # select(Illusion_Strength, Illusion_Difference)
+  group_by(Participant, Illusion_Type) |>
+  summarize(
+    TaskHard_Error = sum(Error, na.rm=TRUE) / n(),
+    TaskHard_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
+    # Illu_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
+  ) |>
+  ungroup() |> 
+  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
+  pivot_wider(names_from = c("Illusion_Type"), values_from = c("TaskHard_Error", "TaskHard_RTMean"), names_glue = "{Illusion_Type}_{.value}")
+
+# Average for 4 higher level of illusion strength
+empirical <- df |>
+  filter(Illusion_Effect == "Incongruent") |> 
+  group_by(Participant, Illusion_Type, Illusion_Difference) |>
+  arrange(desc(Illusion_Strength)) |> 
+  slice(0:4) |> 
+  # select(Illusion_Strength, Illusion_Difference)
+  group_by(Participant, Illusion_Type) |>
+  summarize(
+    IlluStrong_Error = sum(Error, na.rm=TRUE) / n(),
+    IlluStrong_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
+    # Diff_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
+  ) |>
+  ungroup() |> 
+  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
+  pivot_wider(names_from = c("Illusion_Type"), values_from = c("IlluStrong_Error", "IlluStrong_RTMean"), names_glue = "{Illusion_Type}_{.value}") |> 
+  full_join(empirical, by = "Participant") 
+```
+
+<!-- ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE} -->
+<!-- rez <- parameters::n_factors(select(empirical, -Participant), n_max=30) -->
+<!-- plot(rez) -->
+
+<!-- r <- correlation::correlation(empirical, p_adjust = "none", sort=TRUE) -->
+
+<!-- r |> -->
+<!--   ggplot(aes(x = Parameter2, y = Parameter1)) + -->
+<!--   geom_tile(aes(fill = r)) + -->
+<!--   # geom_text(aes(label = Text), size = 2) + -->
+<!--   scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") + -->
+<!--   scale_x_discrete(expand = c(0, 0)) + -->
+<!--   scale_y_discrete(expand = c(0, 0)) + -->
+<!--   labs(title = "Correlation Matrix", x = NULL, y = NULL) + -->
+<!--   theme_minimal() + -->
+<!--   theme( -->
+<!--     axis.text.x = element_text(angle = 90, hjust = 1), -->
+<!--     plot.title = element_text(hjust = 0.5, face = "bold"), -->
+<!--     panel.grid.major = element_blank(), -->
+<!--     panel.grid.minor = element_blank() -->
+<!--   ) -->
+
+<!-- fa <- parameters::factor_analysis(select(empirical, -Participant), -->
+<!--   n = 13, -->
+<!--   rotation = "oblimin", -->
+<!--   fm = "mle", -->
+<!--   sort = TRUE -->
+<!-- ) -->
+
+<!-- print(tibble::as_tibble(fa), n=30) -->
+<!-- plot(fa) -->
+<!-- ``` -->
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Correlation
+r <- data.frame()
+for(ill in c("Delboeuf", 
+             "Ebbinghaus",
+             "RodFrame",
+             "VerticalHorizontal",
+             "Zollner",
+             "White",
+             "MullerLyer",
+             "Ponzo",
+             "Poggendorff",
+             "Contrast"
+             )) {
+  r <- correlation::correlation(
+    data = random |> 
+      arrange(Participant) |> 
+      data_select(select=ill, regex=TRUE),
+    data2 = empirical |>
+      data_select(select=ill, regex=TRUE),
+    p_adjust="none"
+    ) |> 
+    mutate(Illusion_Type = ill,
+           Parameter1 = prettify_parameterName(Parameter1),
+           Parameter2 = prettify_parameterName(Parameter2)) |> 
+    select(Illusion_Type, Parameter1, Parameter2, r, p) |> 
+    rbind(r)
+}
+
+r |>
+  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
+  ggplot(aes(x = Parameter2, y = Parameter1)) +
+  geom_tile(aes(fill = r)) +
+  geom_text(aes(label=label)) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  # facet_wrap(~Illusion_Type, scales = "free") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+```
+
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}

### 2022-08-31 study3/study3.Rmd
update

@@
 
-We have to keep in mind that individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of ...), 2) the structural equation model from which they are obtained is also constructed on layers of abstractions.
-Though, as we have seen, these indices to relate to empirical scores, such as the average error rate and RT in the task.
-
 # Methods
@@
 library(brms)
+
+source("../study2/functions.R")
 
@@
   ) |> 
-  datawizard::standardise(select=names(DVs))
+  datawizard::standardise(select=names(DVs)) |> 
+  datawizard::change_scale(select=starts_with("IPIP"), to = c(-1, 1), range = c(0, 100)) |> 
+  datawizard::change_scale(select=starts_with("PID"), to = c(-1, 1), range = c(0, 100))
 
@@
   pivot_longer(all_of(names(DVs)), names_to = "Index", values_to = "Score")
-# plot(estimate_density(dfsub[DVs]))
 ```
 
-## Results
+## Participants
 
-### Contextual
+`r sum(is.na(dfsub$IPIP6_Agreeableness))` participants did not do the personality scales.
 
-#### Screen Size
+```{r demographics, fig.width=20, fig.height=15}
+p_age <- estimate_density(dfsub$Age) |> 
+  ggplot(aes(x = x, y = y)) +
+  geom_area(fill = "#607D8B") +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "Age", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(face = "italic", hjust = 0.5),
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.title.x = element_blank()
+  )
+
+p_sex <- plot_waffle(dfsub, what="Sex", rows = 12) +
+  scale_color_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63"))
+
+p_edu <- plot_waffle(dfsub, "Education", rows = 12) +
+  scale_color_viridis_d()
+
+p_race <- plot_waffle(dfsub, "Ethnicity", rows = 12) +
+  scale_color_manual(values = c("Latino" = "#FF5722", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Other" = "#795548"))
+
+p_nat <- dfsub |> 
+  group_by(Nationality) |> 
+  summarize(n = n()) |> 
+  mutate(Nationality = fct_reorder(Nationality, desc(n))) |> 
+  ggplot(aes(x = Nationality, y = n, fill=Nationality)) +
+  geom_bar(stat = "identity") +
+  labs(y = "Number", title = "Nationality") +
+  scale_fill_material_d(guide = "none") +
+  scale_x_discrete(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(face = "italic", hjust = 0.5),
+    axis.line.y = element_blank(),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size=rel(0.8), angle = 45, hjust=1)
+  )
+
+p_ipip <- dfsub |> 
+  select(starts_with("IPIP") & !ends_with("_SD")) |> 
+  estimate_density() |> 
+  mutate(Parameter = str_remove_all(Parameter, "IPIP6_"),
+         Parameter = str_replace(Parameter, "HonestyHumility", "Honesty-Humility")) |> 
+  ggplot(aes(x = x, y = y, color = Parameter)) +
+  geom_line(size = 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  scale_color_manual(values=c("Agreeableness" = "#FFC107", "Honesty-Humility" = "#00BCD4", "Extraversion" = "#9C27B0", "Conscientiousness" = "#3F51B5", "Openness" = "#4CAF50", "Neuroticism" = "#E91E63")) +
+  labs(x = "Score", title = "Normal Personality (IPIP-6)", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(face = "italic", hjust = 0.5),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.title.x = element_blank(),
+    legend.position = "top"
+  )
+
+p_pid <- dfsub |> 
+  select(starts_with("PID") & !ends_with("_SD")) |> 
+  estimate_density() |> 
+  mutate(Parameter = str_remove_all(Parameter, "PID5_"),
+         Parameter = str_replace(Parameter, "NegativeAffect", "Negative Affect")) |> 
+  ggplot(aes(x = x, y = y, color = Parameter)) +
+  geom_line(size = 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  scale_color_manual(values=c("Antagonism" = "#F44336", "Detachment" = "#03A9F4", "Disinhibition" = "#FF5722", "Negative Affect" = "#FF9800", "Psychoticism" = "#673AB7")) +
+  labs(x = "Score", title = "Pathological Personality (PID-5)", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(face = "italic", hjust = 0.5),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.title.x = element_blank(),
+    legend.position = "top"
+  )
+
+p <- (p_age / p_nat) | (p_sex / p_edu / p_race) | (p_ipip / p_pid)
+p
+# ggsave("figures/figure_demo.png", width=fig.width, height = fig.height)
+```
+
+# Results
+
+## Contextual
+
+### Screen Size
 
@@
   ggside::geom_ysidedensity(data=dflong, aes(y=Score, color = Index)) +
-  ggside::theme_ggside_void() 
+  ggside::theme_ggside_void() +
+  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
+  ggside::scale_xsidey_continuous(expand = c(0, 0))
 ```
 
-#### Refresh Rate
+### Refresh Rate
 
@@
 
-### Demographics
+## Demographics
 
-#### Sex
+### Sex
 
@@
 
-#### Ethnicity
+### Ethnicity
 
@@
 
-#### Education
+### Education
 
@@
 
-#### Age
+### Age
 
@@
 
-### Personality
+## Personality
+
+### IPIP6
 

### 2022-08-31 presentations/make_figures.R

@@ -1,0 +1,27 @@
+library(ggplot2)
+library(gganimate)
+library(mgcv)
+
+df <- MASS::mcycle
+
+m_lm <- lm(accel ~ times, data=df)
+m_poly2 <- lm(accel ~ poly(times, 2), data=df)
+m_poly3 <- lm(accel ~ poly(times, 3), data=df)
+m_poly5 <- lm(accel ~ poly(times, 5), data=df)
+m_poly7 <- lm(accel ~ poly(times, 7), data=df)
+m_gam <- gam(accel ~ s(times), data=df)
+
+# series of measurements of head acceleration in a simulated motorcycle accident, used to test crash helmets.
+p <- ggplot(df, aes(x=times, y=Predicted)) +
+  geom_point(aes(y=accel)) +
+  see::theme_modern() +
+  labs(x = "Time after impact (ms)", y = "Acceleration (g)", color="Models") +
+  scale_color_manual(values=c("Linear" = "black", "Poly (2)" = "#FFC107", "Poly (3)" = "#FF9800", "Poly (5)" = "#FF5722", "Poly (7)" = "#F44336", "GAM" = "#2196F3")) +
+  geom_line(data=modelbased::estimate_relation(m_lm), aes(color = "Linear")) +
+  geom_line(data=modelbased::estimate_relation(m_poly2, length=100), aes(color = "Poly (2)")) +
+  geom_line(data=modelbased::estimate_relation(m_poly3, length=100), aes(color = "Poly (3)")) +
+  geom_line(data=modelbased::estimate_relation(m_poly5, length=100), aes(color = "Poly (5)")) +
+  geom_line(data=modelbased::estimate_relation(m_poly7, length=100), aes(color = "Poly (7)")) +
+  geom_line(data=modelbased::estimate_relation(m_gam, length=100), aes(color = "GAM"), size=2) +
+  transition_layers() + enter_fade()
+

### 2022-08-31 study2/functions.R

@@ -2,8 +2,43 @@ logmod <- function(x) sign(x) * log(1 + abs(x))
 sqrtmod <- function(x) sign(x) * sqrt(abs(x))
 cbrtmod <- function(x) sign(x) * (abs(x)**(1 / 3))
 
+plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", fill = "orange") {
+  dfsub |>
+    ggplot(aes_string(x = what)) +
+    geom_density(fill = fill) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_y_continuous(expand = c(0, 0)) +
+    ggtitle(title, subtitle = subtitle) +
+    theme_modern() +
+    theme(
+      plot.title = element_text(face = "bold", hjust = 0.5),
+      plot.subtitle = element_text(face = "italic", hjust = 0.5),
+      axis.title.x = element_blank(),
+      axis.title.y = element_blank(),
+      axis.text.y = element_blank()
+    )
+}
 
-
+plot_waffle <- function(dfsub, what = "Nationality", title = what, rows = 8, size=3) {
+  library(emojifont)
+  ggwaffle::waffle_iron(dfsub, what, rows = rows) |>
+    # mutate(label = emojifont::fontawesome('fa-smiley')) |>
+    # mutate(label = emojifont::emoji('smiley')) |>
+    ggplot(aes(x, y)) +
+    geom_point(aes(color = group), shape = "square", size=size) +
+    # ggwaffle::geom_waffle(color = "white") +
+    # geom_point() +
+    # geom_text(aes(color=group ,label=label), family='fontawesome-webfont', size=4) +
+    # geom_text(aes(color=group ,label=label), family='EmojiOne', size=4) +
+    coord_equal() +
+    ggtitle(title) +
+    labs(fill = "", color = "") +
+    # scale_x_continuous(expand = c(0, 0)) +
+    # scale_y_continuous(expand = c(0, 0)) +
+    theme_void() +
+    # ggwaffle::theme_waffle() +
+    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+}
 
 
 
@@ -231,10 +266,12 @@ plot_model_err <- function(data, model, gam=NULL) {
   data$color <- fct_reorder(data$color, closest)
 
   # Manual jittering
-  xrange <- 0.03 * diff(range(data[[varstrength]]))
+  xrange <- 0.04 * diff(range(data[[varstrength]]))
   data$x <- data[[varstrength]]
[…]

### 2022-08-31 study2/preprocessing.R

@@ -96,7 +96,6 @@ preprocess_raw <- function(file) {
   df$Duration <- df$Duration - df$Break_Duration
 
   # Manual fixes for reported errors (using demographic info from prolific)
-  # df[df$Sex == "Other", "Participant"]
   df[df$Participant == "5d3c6e745602310001bca8aa_6bdtb", "Sex"] <- "Female"
   df[df$Participant == "60f5fa488f7b1381d175ecb5_qwvfs", "Sex"] <- "Male"
   df[df$Participant == "5d558c262d8a6200172a8b7b_deegh", "Sex"] <- "Male"
@@ -122,38 +121,43 @@ preprocess_raw <- function(file) {
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Mixed/Latino", "Mexicano"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Hispanic/Latino ", "Latin American "), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latin", "Latina", "HISPANIC", "Mexican"), "Latino", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latino American", "Latino/Hispanic"), "Latino", df$Ethnicity)
+  df$Ethnicity <- ifelse(df$Ethnicity %in% c("Latino American", "Latino/Hispanic", "Latinx"), "Latino", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White Middle European (Slavic)", "Greek"), "Caucasian", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White - Caucasian", " Caucasian"), "Caucasian", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White", "Caucasian ", "German", "English"), "Caucasian", df$Ethnicity)
-  df$Ethnicity <- ifelse(df$Ethnicity %in% c("European/White", "WHITE", "White Ethnic"), "Caucasian", df$Ethnicity)
+  df$Ethnicity <- ifelse(df$Ethnicity %in% c("European/White", "WHITE", "White Ethnic", "American"), "Caucasian", df$Ethnicity)
   df$Ethnicity <- ifelse(df$Ethnicity %in% c("White/Caucasian", "British", "White/European"), "Caucasian", df$Ethnicity)
[…]

### 2022-08-31 manuscript.Rmd
Update manuscript.Rmd

@@
 
-In line with our full open-access standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at **https://github.com/RealityBending/IllusionGameValidation**.
+In line with our open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available and open-source at **https://github.com/RealityBending/IllusionGameValidation**.
 

### 2022-08-31 manuscript.Rmd
edit

@@
 
-Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize the information gain.
-In most illusions, the task difficulty exhibited a non-linear effect, which is in line with the psychophysics literature on perceptual decisions.
+Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
+In most illusions, the task difficulty exhibited a non-linear effect, which is in line with the psychophysics literature on perceptual decisions (*REF*).
 One notable result was the illusion effect for the Zöller illusion, which suggested a non-linear pattern.
@@
 See the next study for the full demographic breakdown.
-We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials based on response time (< 125 ms or > 4 SD above mean).
+We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials with extreme response times (< 125 ms or > 4 SD above mean).
 
@@
 
-The best models were $log(diff)*strength$ for Delboeuf;
+The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; ....
+In all of these models, the effect of illusion strength, task difficulty and their interaction were significant.
 
-Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, dynamics, we decided not to use add them for the individual scores analysis.
-Additionally, it would have increased the parameter space from 30 to 90 variables, which was deemed as statistically unreasonable given our sample size).
+For errors, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0.
+
+However, for reaction times, the GAMs suggested a consistent non-linear relationship between the mean RT and illusion strength: as the illusion strength gets very strong, the reaction times are faster.
+While this is is not surprising (strong illusions are so biasing that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this pattern and hence are not good representatives of the underlying dynamics.
+Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, relationship, we decided not to use add them for the individual scores analysis.
+<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size). -->
+
+Though imperfect, we believe that these models capture inter-individual differences more accurately than basic empirical scores (e.g., the total amount of errors, or the average RT).
+Thus, for each illusion, we have extracted the effect of illusion strength and its interaction with task difficulty (when the illusion incongruent) within each participants.
+These twenty participant-level scores were subjected to exploratory factor analysis (EFA), which suggested that ...
 
@@
 
+<!-- This is likely to be caused by the outline. It will be fixed in the next release of the Pyllusion software. -->
 Future studies should investigate reaction times to find the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
@@
 
+Thus opening the door for new illusions-based paradigms and tasks, we have shown that illusions can be manipulated continuously across several different visual illusions with the parametric illusion generation framework introduced with *Pyllusion* (*REF*).
+By applying this approach to a reaction-time task, we were able to measure inter-individual scores of objective illusion sensitivity.
+
 # Future Directions

### 2022-09-01 manuscript.Rmd
update

@@
 Indeed, many visual illusions can be seen as made of *targets* (e.g., same-length lines), which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the segments appear of different lengths when they end with inwards or outwards pointing arrows).
-While most of the paradigms used in the literature prompt the subjective extent the context biases the perception ("how much does the circle appear bigger", e.g., **REFS**), *Pyllusions* allows to create illusions in which the targets are actually more or less different (i.e., one segment is longer than the other), and in which the illusion is of varying strength (the arrows are more or less pointy).
+While most of the paradigms used in the literature prompt the subjective extent the context biases the perception (e.g."how much doe the line appear different"; **REFS**), *Pyllusions* allows to create illusions in which the targets are actually more or less different (e.g., one segment is truly longer than the other), and in which the illusion varies in strength (the arrows are more or less pointy).
 
@@
 This opens the door of an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can be either "incongruent" (making the task even harder by biasing the perception in the opposite way) or "congruent" (making the task easier).
-Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm that could quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
+Moreover, the illusion effect can be either "incongruent" (making the task even harder by biasing the perception in the opposite way) or "congruent" (making the task easier by making for instance the longer segment look even longer).
+Although visual illusions are inherently tied to subjective and conscious perception (**REF about whether illusion can work on a subconscious level, in which case we need to remove "conscious"**), this framework allows a reversal of the traditional paradigm that could quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
 
 <!-- Goals -->
-In the present set of preregistered studies, we will first test this novel paradigm by investigate if the effect of illusion and task difficulty can be manipulated continuously and separated statistically.
-Then, we will use the paradigm to assess whether 10 different historical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate).
-Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
+In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously and statistically discriminated.
+Then, we will use the paradigm to assess whether 10 different historical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate; *REF*).
+Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality (*REF*).
 
@@
 
-The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; ....
+The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; $cbrt(diff)*strength$ for Zöllner; $diff*sqrt(strength)$ and $log(diff)*strength$ respectively for errors and RT in White; $sqrt(diff)*sqrt(strength)$ and $sqrt(diff)*strength$ respectively for errors and RT in Müller-Lyer; $cbrt(diff)*strength$ for Ponzo; $cbrt(diff)*sqrt(strength)$ and $cbrt(diff)*strength$ respectively for errors and RT in Poggendorff; $sqrt(diff)*sqrt(strength)$ for Contrast.
 In all of these models, the effect of illusion strength, task difficulty and their interaction were significant.
@@
 For errors, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0.
-
-However, for reaction times, the GAMs suggested a consistent non-linear relationship between the mean RT and illusion strength: as the illusion strength gets very strong, the reaction times are faster.
-While this is is not surprising (strong illusions are so biasing that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this pattern and hence are not good representatives of the underlying dynamics.
-Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, relationship, we decided not to use add them for the individual scores analysis.
+However, for RTs, the GAMs suggested a consistent non-linear relationship between the mean RT and illusion strength: as the illusion strength gets very strong, the participants respond faster.
+While this is is not surprising (strong illusions are likely so biasing that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this pattern and hence are not good representatives of the underlying dynamics.
+Because these random-slopes models for RT did not appear to capture well the underlying - likely quadratic - relationship, we decided not to use add them for the individual scores analysis.
 <!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size). -->
 
-Though imperfect, we believe that these models capture inter-individual differences more accurately than basic empirical scores (e.g., the total amount of errors, or the average RT).
+Though imperfect, we believe that the logistic mixed models capture inter-individual differences more accurately than basic empirical scores (e.g., the total amount of errors).
 Thus, for each illusion, we have extracted the effect of illusion strength and its interaction with task difficulty (when the illusion incongruent) within each participants.
@@
 
-<!-- This is likely to be caused by the outline. It will be fixed in the next release of the Pyllusion software. -->
+<!-- This is likely an artifact caused by the outline. It will be fixed in the next release of the Pyllusion software. -->
+
 Future studies should investigate reaction times to find the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
@@
 
-We strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding further.
+We strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+The task, data and analysis script are available in open-access at **https://github.com/RealityBending/IllusionGameValidation**.
 

### 2022-09-01 study2/study2.Rmd
edit
update

@@
 runModels <- FALSE
-bestModels <- TRUE
+bestModels <- FALSE
 fig.width <- see::golden_ratio(7)
@@
   labs(y = "", x = "Reaction Time (ms)")
-p
+# p
 # ggsave("figures/outliers.png", p, width=25, height=25)
@@
 ```
+
+##### Model Inspection
+
 ```{r message=FALSE, warning=FALSE, eval=TRUE}
@@
 
-##### Model Visualization
-
 ```{r message=FALSE, warning=FALSE, cache=cache}
 p_delboeuf_err <- plot_model_err(data, model_delboeuf_err, gam_delboeuf_err)
-p_delboeuf_err
 ```
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
+
+##### Model Inspection
+
+```{r message=FALSE, warning=FALSE, cache=cache}
 load("models/model_delboeuf_rt.RData")
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache}
 plot_ppcheck(model_delboeuf_rt, gam_delboeuf_rt)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2}
 p_delboeuf <- plot_all(data, p_delboeuf_err, p_delboeuf_rt)
@@
 ```
+
+##### Model Inspection
+
 ```{r message=FALSE, warning=FALSE, cache=cache}
@@
 
-##### Model Visualization
-
 ```{r message=FALSE, warning=FALSE, cache=cache}
 p_ebbinghaus_err <- plot_model_err(data, model_ebbinghaus_err, gam_ebbinghaus_err)
-p_ebbinghaus_err
 ```
@@
 
-```{r m_rt_ebbinghaus, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r m_rt_ebbinghaus, message=FALSE, warning=FALSE, cache=cache, eval=runModels}
 formula <- brms::bf(
@@
 ```
+
+##### Model Inspection
+
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 plot_ppcheck(model_ebbinghaus_rt, gam_ebbinghaus_rt)
@@
 p_ebbinghaus_rt <- plot_model_rt(data, model_ebbinghaus_rt, gam_ebbinghaus_rt)
-p_ebbinghaus_rt
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
 p_ebbinghaus <- plot_all(data, p_ebbinghaus_err, p_ebbinghaus_rt)
@@
 
-##### Model Visualization
-
 ```{r message=FALSE, warning=FALSE, cache=cache}
 p_rodframe_err <- plot_model_err(data, model_rodframe_err, gam_rodframe_err)
-p_rodframe_err
 ```
@@
 
-```{r m_rt_rodframe, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r m_rt_rodframe, message=FALSE, warning=FALSE, cache=cache, eval=runModels}
 formula <- brms::bf(
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 plot_ppcheck(model_rodframe_rt, gam_rodframe_rt)
@@
 p_rodframe_rt <- plot_model_rt(data, model_rodframe_rt, gam_rodframe_rt)
-p_rodframe_rt
 ```
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
 p_rodframe <- plot_all(data, p_rodframe_err, p_rodframe_rt)
@@
 # parameters::parameters(model_verticalhorizontal_err)
-
-```
-
-##### Model Visualization
+```
 
@@
 p_verticalhorizontal_err <- plot_model_err(data, model_verticalhorizontal_err, gam_verticalhorizontal_err)
-p_verticalhorizontal_err
 ```
@@
 
-```{r m_rt_verticalhorizontal, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r m_rt_verticalhorizontal, message=FALSE, warning=FALSE, cache=cache, eval=runModels}
 formula <- brms::bf(
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+
+##### Model Inspection
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_verticalhorizontal_rt.RData")
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 p_verticalhorizontal_rt <- plot_model_rt(data, model_verticalhorizontal_rt, gam_verticalhorizontal_rt)
-p_verticalhorizontal_rt
+```
+
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
+p_verticalhorizontal <- plot_all(data, p_verticalhorizontal_err, p_verticalhorizontal_rt)
+p_verticalhorizontal
 ```
@@
   data = data,
-  refresh = 0
-)
-
-# parameters::parameters(model_zollner_err)
+  refresh = 100
+)
 
@@
 load("models/gam_zollner_err.RData")
-```
-
-##### Model Visualization
+
+# parameters::parameters(model_zollner_err)
+```
 
@@
 p_zollner_err <- plot_model_err(data, model_zollner_err, gam_zollner_err)
-p_zollner_err
 ```
@@
 
-```{r m_rt_zollner, message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r m_rt_zollner, message=FALSE, warning=FALSE, cache=cache, eval=runModels}
 formula <- brms::bf(
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+
+##### Model Inspection
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_zollner_rt.RData")
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 p_zollner_rt <- plot_model_rt(data, model_zollner_rt, gam_zollner_rt)
-p_zollner_rt
+```
+
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
+p_zollner <- plot_all(data, p_zollner_err, p_zollner_rt)
+p_zollner
 ```
@@
 
-##### Model Visualization
-
 ```{r message=FALSE, warning=FALSE, cache=cache}
 p_white_err <- plot_model_err(data, model_white_err, gam_white_err)
-p_white_err
 ```
@@
 formula <- brms::bf(
-  RT ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
+  RT ~ Illusion_Effect / (logmod(Illusion_Difference) * abs(Illusion_Strength)) +
     poly(ISI, 2) + 
-    (1 + Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) | Participant),
-  sigma ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (logmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
+  sigma ~ Illusion_Effect / (logmod(Illusion_Difference) * abs(Illusion_Strength)) +
     poly(ISI, 2) + (1 | Participant),
-  beta ~ Illusion_Effect / (Illusion_Difference * sqrt(abs(Illusion_Strength))) +
+  beta ~ Illusion_Effect / (logmod(Illusion_Difference) * abs(Illusion_Strength)) +
     poly(ISI, 2) + (1 | Participant),
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+
+##### Model Inspection
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_white_rt.RData")
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 p_white_rt <- plot_model_rt(data, model_white_rt, gam_white_rt)
-p_white_rt
+```
+
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
+p_white <- plot_all(data, p_white_err, p_white_rt)
+p_white
 ```
@@
 
-##### Model Visualization
-
 ```{r message=FALSE, warning=FALSE, cache=cache}
 p_mullerlyer_err <- plot_model_err(data, model_mullerlyer_err, gam_mullerlyer_err)
-p_mullerlyer_err
 ```
@@
 formula <- brms::bf(
-  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+  RT ~ Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) +
     poly(ISI, 2) + 
-    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
-  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+    (1 + Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
+  sigma ~ Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) +
     poly(ISI, 2) + (1 | Participant),
-  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
+  beta ~ Illusion_Effect / (sqrt(Illusion_Difference) * abs(Illusion_Strength)) + 
     poly(ISI, 2) + (1 | Participant),
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+
+##### Model Inspection
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_mullerlyer_rt.RData")
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 p_mullerlyer_rt <- plot_model_rt(data, model_mullerlyer_rt, gam_mullerlyer_rt)
-p_mullerlyer_rt
 ```
@@
 
-##### Model Visualization
-
 ```{r message=FALSE, warning=FALSE, cache=cache}
 p_ponzo_err <- plot_model_err(data, model_ponzo_err, gam_ponzo_err)
-p_ponzo_err
 ```
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+
+##### Model Inspection
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_ponzo_rt.RData")
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 p_ponzo_rt <- plot_model_rt(data, model_ponzo_rt, gam_ponzo_rt)
-p_ponzo_rt
 ```
@@
 
-##### Model Visualization
-
 ```{r message=FALSE, warning=FALSE, cache=cache}
 p_poggendorff_err <- plot_model_err(data, model_poggendorff_err, gam_poggendorff_err)
-p_poggendorff_err
 ```
@@
 formula <- brms::bf(
-  RT ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) +
+  RT ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) +
     poly(ISI, 2) + 
-    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) | Participant),
-  sigma ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
+    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
+  sigma ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) + 
     poly(ISI, 2) + (1 | Participant),
-  beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrt(abs(Illusion_Strength))) + 
+  beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * abs(Illusion_Strength)) + 
     poly(ISI, 2) + (1 | Participant),
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+
+##### Model Inspection
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_poggendorff_rt.RData")
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 p_poggendorff_rt <- plot_model_rt(data, model_poggendorff_rt, gam_poggendorff_rt)
-p_poggendorff_rt
 ```
@@
   data = data,
-  refresh = 50
+  refresh = 100
 )
@@
 
-##### Model Visualization
-
 ```{r message=FALSE, warning=FALSE, cache=cache}
 p_contrast_err <- plot_model_err(data, model_contrast_err, gam_contrast_err)
-p_contrast_err
 ```
@@
 ```
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+
+##### Model Inspection
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 load("models/model_contrast_rt.RData")
@@
 
-##### Model Visualization
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 p_contrast_rt <- plot_model_rt(data, model_contrast_rt, gam_contrast_rt)
-p_contrast_rt
 ```

### 2022-09-01 study2/functions.R

@@ -411,12 +411,12 @@ plot_ppcheck <- function(model, gam) {
   #   estimate_density(select = "iter_value", at = "iter_group") |>
   #   normalize(select = "y")
 
-  pred <- modelbased::estimate_prediction(model, keep_iterations = 100) |>
+  pred <- modelbased::estimate_prediction(model, keep_iterations = 50) |>
     bayestestR::reshape_iterations() |>
     mutate(iter_group = as.factor(iter_group)) |>
     estimate_density(select = "iter_value", at = "iter_group")
 
-  predgam <- modelbased::estimate_prediction(gam, keep_iterations = 100) |>
+  predgam <- modelbased::estimate_prediction(gam, keep_iterations = 50) |>
     bayestestR::reshape_iterations() |>
     mutate(iter_group = as.factor(iter_group)) |>
     estimate_density(select = "iter_value", at = "iter_group")

### 2022-09-01 manuscript.Rmd
update manu

--- 2022-08-31 manuscript.Rmd
@@
   chunk_output_type: console
+bibliography: references.bib
 ---
@@
 <!-- Title: -->
+
 <!-- The Illusion Game: Validation of a Novel Experimental Paradigm to study Visual Illusions  -->
@@
 <!-- Intro  & Gaps-->
+
 Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems.
-They eloquently show that our brains did not evolve to be perfect perceptual devices, but to take into account contextual information and prior knowledge that manifests in our conscious experience **(REF)**.
-Despite the historical interest within the fields of visual perception **(REF)**, consciousness science **(REF)**, or psychiatry **(REF)**, several gaps remained unanswered.
-*mention common factor for illusions and link with disorders*.
+They eloquently show that our brains do not simply map veridical accounts of physical reality; we generate perceptions of the environment by taking into consideration the contextual information and the prior knowledge that manifests in our subjective conscious experience [@carbon2014].
+Despite the withstanding interest in illusions within various fields of study, including that of visual perception [@day1972; @eagleman2001; @gomez-villa2022 ], consciousness science [@caporuscio2022, @lamme2020] and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015re], several gaps still remain in the present research.
+Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions [@mylniec2016].
+For instance, while some researchers have tried to explain our sensitivity to illusory effects as a result of deficits in the low-level visual processing system [@cretenoud2019; @gori2016], others have provided a compelling case using a top-down approach, suggesting that such visual phenomena occur as a result of a conflict between our visual input and our prior beliefs[@teufel2018; @caporuscio2022].
+Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
+Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder (ASD)) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
 
 <!-- Pyllusion -->
-One key challenge hindering the further development of illusion research is the relative difficulty to adapt visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-To address this issue, we first developed a parametric framework to manipulate visual illusions, that we implemented and made accessible in the open-source software *Pyllusion* **(REF)**.
-This software allows us to generate different types of historical visual illusions with a continuous and independent modulation of two parameters, *illusion strength* and *task difficulty* (see **Figure 1**).
+
+One key challenge hindering the further development of illusion research is the relative difficulty in adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+To address this issue, we first developed a parametric framework to manipulate visual illusions, of which we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
+This software allows us to generate different types of classic visual illusions and more importantly, dissociate the parameters of an illusion from its rendered output.
+Specifically, it enables us to modulate continuously and independently two parameters, *illusion strength* and *task difficulty*, that could then be manipulated systematically to conduct objective assessments on visual illusions (see **Figure 1**).
 
@@
 <!-- Target, strength, difficulty -->
-Indeed, many visual illusions can be seen as made of *targets* (e.g., same-length lines), which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the segments appear of different lengths when they end with inwards or outwards pointing arrows).
-While most of the paradigms used in the literature prompt the subjective extent the context biases the perception ("how much does the circle appear bigger", e.g., **REFS**), *Pyllusions* allows to create illusions in which the targets are actually more or less different (i.e., one segment is longer than the other), and in which the illusion is of varying strength (the arrows are more or less pointy).
+
+Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
+In past studies, the implemented paradigms have mostly involved prompting the subjective extent the context biases the perception (for instance, by getting participants to adjust the parameters of one comparison stimulus (the size of a circle) to another (circle) e.g., @mylniec2016a; @grzeczkowski2018).
+Alternatively, *Pyllusion* allows us to create illusions while manipulating specific features of interest (e.g., size, length etc., ) and thereby the difficulty in distinguishing them, so that the targets are objectively distinct (i.e., one line segment is physically longer than the other), and the strength of the generated illusion are variable (i.e., the biasing angle of the arrows are more or less acute).
 
 <!-- Paradigm -->
-This opens the door of an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can be either "incongruent" (making the task even harder by biasing the perception in the opposite way) or "congruent" (making the task easier).
-Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm that could quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
+
+This thereby opens the door to a standardizable experimental task in which participants can be asked to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+Moreover, the illusory effect can be set to either be "incongruent" (increasing the difficulty of the task by biasing the perception in the opposite way) or "congruent" (making the task easier).
+Although visual illusions are inherently tied to subjective perception, this framework radically reverses the traditional paradigm in favor of an approach that could quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
 
 <!-- Goals -->
-In the present set of preregistered studies, we will first test this novel paradigm by investigate if the effect of illusion and task difficulty can be manipulated continuously and separated statistically.
-Then, we will use the paradigm to assess whether 10 different historical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate).
+
+In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separated statistically.
+Then, we will further utilize the paradigm to assess whether 10 different cannonical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate, [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]).
 Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
 
-In line with our open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available and open-source at **https://github.com/RealityBending/IllusionGameValidation**.
+In line with our open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available and open-source at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
 
@@
 
-Study 1 can be seen as a pilot study which goal was to gather some preliminary data to 1) assess whether the stimuli generated by *Pyllusion* behave as expected for each of the 10 illusion types (i.e., increase of task difficulty and illusion strength leading to increased error rate) and 2) develop an intuition of the magnitude of effects, to then refine the stimuli parameters to a sensible range (i.e., not overly easy and not impossible hard).
+The objective of Study 1 is to act as a pilot study in order to gather some preliminary data to 1) assess whether the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., increase of task difficulty and illusion strength leads to an increased error rate) and 2) develop a gauged understanding of the magnitude of effects, to then refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard).
 
@@
 We generated 56 stimuli for each of the 10 illusion types.
-These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where the negative values correspond to congruent illusion strengths).
+These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where values correspond to the objective difference between targets, i.e., 1 corresponds to the smallest objective difference and thereby the highest level of difficulty in the task) and 7 levels of illusion strength (3 values (i.e, degrees) of strength on the side of illusion congruence, 3 on the side of illusion incongruence, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3].
+Notably, the magnitude of these values correspond to illusion strength (greater values represent greater strength) whereas the valence indicates the illusion congruence (negative sign represents an illusion congruent effect).
 
-The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the block was also randomized.
-After the first series of 10 blocks, another series was done (with new randomized blocks and trials order).
-In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
-The task was implemented using *jsPsych* (*REF*).
-See instructions for each illusion type in the experiment code.
+The 10 types of illusion stimuli were randomly presented in distinct blocks, within which the 56 stimuli were also displayed in a randomized order.
+After the first series of 10 blocks was presented, another series was displayed (with the same set of stimuli but with new randomized block and trial orders).
+All in all, each participant saw 56 different trials per illusion type (10 blocks in total), in 2 series of blocks (amounting to a total of 1120 trials).
+In each illusion trial, participants were required to evaluate the targets comprising the illusions without being distracted by its biasing contextual information.
+For example, in a Müller-Lyer illusion trial, participants were instructed to evaluate the length of line segments without being distracted by the direction in which the arrows at its ends are pointing towards.
+Importantly, participants were told to respond "as fast as possible without making errors" before each block of trials (i.e., an explicit double constraint incorporated to mitigate the inter-individual variability in the speed-accuracy trade off in responses).
+This task was implemented using *jsPsych* [@deleeuw2014].
+For specific details of the instructions given for each illusion type, please refer to the experiment code.
 
@@
 
-Fifty-two participants were recruited via *Prolific* (**REF** about prolific and their data quality) with no restrictions on age or country (the only requisite was a "fluent" level of English to match the language of instructions), with a reward of about \textsterling 7.5.
+Fifty-two participants were recruited via *Prolific* ([www.prolificacademic.co.uk](www.prolificacademic.co.uk)), a popular crowd-sourcing platform to recruit subjects for online behavioural research.
+In a study comparing across different research recruitment platforms (e.g., MTurks, Qualtrics etc.,,), participants from *Prolific* were found to provide the highest data quality across all examined measures, including that of attention (the extent to which participants actually read the questions prior to responding), comprehension (the extent to which participants were able to understand the task instructions or items), honesty (the degree to which participants responded to questions truthfully) and reliability (the extent to which participants' responses are internally consistent within the same scale) [@peer2022].
+For the purposes of our study, no restrictions on age or country were imposed on participant recruitment; the only requisite was a "fluent" proficiency in English to ensure that the task instructions, given in English, would be well-comprehended.
+Participants were incentivized with a reward of about \textsterling 7.5 for completing the task, which took an average of an hour to finish.
 
-We removed 6 participants upon inspection of the average error rage (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
-For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that instructions got mixed up; e.g., participants were selecting the shorter line instead of the longer one).
-Finally, we removed 692 (1.37%) trials based on an implausibly short or wrong response time (< 150 ms or > 3000 ms).
+Average error rate and reaction time distributions were analysed in the pre-processing of the data.
+We removed 6 participants upon inspection of their average error rage (which was close to 50%, suggesting random answers), and because their reaction time distributions were implausibly fast.
+For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that participants misunderstood the instructions given; e.g., participants were selecting the shorter line instead of the longer one).
+Finally, we removed 692 (1.37%) trials based on an implausibly short or lomg response time (\< 150 ms or \> 3000 ms).
 
@@
 The analysis of study 1 focused on the probability of errors as the main outcome variable.
-For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some intuition on the underlying generative model.
+For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some perspective on the underlying generative model.
 Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
-We then fitted the best performing model under a Bayesian model, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships (at the expense of model simplicity).
+We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships (at the expense of model simplicity).
 
-The analysis was carried out using *R 4.2* (**ref**), *brms* (*REF*), the *tidyverse* (*JOSS REF*) and the *easystats* collection of packages (*REFS*).
+The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019] and the *easystats* collection of packages [@easystatsPackage].
 
@@
 
-The statistical models suggested that the effect of task difficulty had a cube relationship with error rate for Delboeuf and Ebbinghaus illusions (both based on circle size), square for Rod and Frame, Vertical-Horizontal, cube for Zöllner and Poggendorff, exponential for White, cube for Müller-Lyer and Ponzo (both based on lie lengths), and linear for Contrast.
-See details and figures in the analysis script.
+The statistical models suggested that the effect of task difficulty had a cubic relationship with error rate for the Delboeuf and Ebbinghaus illusions (both composed of circular shapes), square (\*\*quadratic/inverse?\*\*) relationship for the Rod and Frame and Vertical-Horizontal illusions, cubic relationship for the Zöllner and Poggendorff illusions, exponential relationship for the White illusion, cubic relationship for the Müller-Lyer and Ponzo illusions (both based on line lengths), and linear relationship for the Contrast illusion.
+Please see details and figures in the analysis script.
 
@@
 
-This study allowed to elaborate a more precise prior idea regarding the magnitude of the effects at stake and the type of interaction between them.
-It allowed us to better understand and test-out the stimuli generated by *Pyllusion*.
-It allowed us to uncover a few bugs and issues (for instance, the specification direction of the illusion strength was reversed in the software), which were fixed in a new release.
+This study enabled a more precise elaboration, and therefore clearer understanding, of the exact magnitude of the parametric effects at stake and the type of interaction between them.
+Furthermore, it allowed us to better understand and test the stimuli generated by *Pyllusion*, as well as unintentionally uncover a few technological bugs and issues (for instance, the specification direction of the illusion strength was reversed in the software), which were fixed in a new release.
 
-Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize the information gain.
-In most illusions, the task difficulty exhibited a non-linear effect, which is in line with the psychophysics literature on perceptual decisions.
-One notable result was the illusion effect for the Zöller illusion, which suggested a non-linear pattern.
+Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
+In most illusions, the task difficulty exhibited a non-linear effect, which is in line with the psychophysics literature on perceptual decisions ([@ditzinger2010; @shekhar2021; @bogacz2006]).
+In particular, one notable result was the illusion effect for the Zöller illusion, which suggested a non-linear relationship.
 By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.
@@
 
-The paradigm was similar to that of study 1, with a few changes: stimuli were re-generating with within a refined space of parameters.
-Moreover, based on study 1, we used non-linearly spaced difficulty levels, depending on the best underlying model (i.e., with an exponential, square or cubic spacing depending on the relationship).
+The paradigm of study 2 was similar to that of study 1, with a few changes: namely, the illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
+Moreover, taking into account the findings of study 1, we used non-linearly spaced difficulty levels, depending on the best underlying model (i.e., with an exponential, square or cubic spacing depending on the relationship).
 For instance, a linear space of [0.1, 0.3, 0.5, 0.7] can be transformed to an exponential space of [0.1, 0.27, 0.47, 0.7].
 
-One additional difference was that instead of repeating two times each stimulus, we generated more versions (with more levels of difficulty and illusion strength).
-For each illusion type, we generated 134 stimuli, that were split into two groups (67 stimuli per illusion block).
-Additionally, instead of a simple break screen, we added two personality questionnaires between the two series of 10 blocks of illusions (see study 3).
+Additionally, instead of repeating each series of illusion blocks once so that each illusion stimuli is seen two times, we generated more stimuli by increasing the degree of variation in the manipulated parameters (i.e., with more levels of difficulty and illusion strength).
+As such, for each illusion type, we generated a total of 134 stimuli that were split into two groups (67 stimuli per illusion block).
+Furthermore, instead of a simple break screen, we added two personality questionnaires between the two series of 10 illusion blocks (see study 3).
 
@@
 
-Using the same recruitment procedure, we recruited 256 participants, from which 6 were identified as outliers and discarded, leading to a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
-See the next study for the full demographic breakdown.
-We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials based on response time (< 125 ms or > 4 SD above mean).
+Using the same recruitment procedure as in study 1, we recruited 256 participants, within which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
+Please see the next study (study 3) for the full demographic breakdown.
+We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean).
 
@@
 
-The first part of the analysis focuses on modelling the effect of illusion strength and task difficulty, on errors and reaction time, within each illusion.
+The first part of the analysis focuses on modelling the effect of illusion strength and task difficulty on errors and reaction time, within each illusion.
 In order to achieve that, we will start by fitting a General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
@@
 
-Using the GAMs as a "ground-truth", we will attempt at approximating them using general linear models, which have the advantage  which we can specify a participant-level variability of the effects (via random effects).
+Using the GAMs as a "ground-truth", we will attempt at approximating them using general linear models, which have the advantage of specifying a participant-level variability of the effects (via random effects).
 Firstly, we will do a model comparison between a combination of transformations (raw, log, square root or cubic root) on the main predictors (task difficulty *diff* and illusion *strength*).
@@
 
-We will then extract the inter-individual variability (specified as random slopes) in the effect of illusion strength, task difficulty and their interaction and use it as participant-level scores.
+We will then extract the inter-individual variability (specified as random slopes) in the effects of illusion strength, task difficulty and their interaction and use it as participant-level scores.
 Finally, We will explore the relationship across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
@@
 
-The best models were $log(diff)*strength$ for Delboeuf;
+The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; ....
+For all of the specified models, the effect of illusion strength, task difficulty and their interaction were significant.
 
-Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, dynamics, we decided not to use add them for the individual scores analysis.
-Additionally, it would have increased the parameter space from 30 to 90 variables, which was deemed as statistically unreasonable given our sample size).
+With regards to its effect on errors, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0.
+
+However, regarding the effect on reaction times, the GAMs suggested a consistent non-linear relationship between the mean RT and illusion strength: as the illusion strength increases, the reaction times are faster.
+While this is is not entirely surprising (for instance, it is conceptually possible to theorize that strong illusions are so effective at biasing our visual perceptions that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this pattern and hence are not good representatives of the underlying dynamics.
+Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, relationship, we decided not to include them for the individual scores analysis.
+<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size).
+-->
+
+Though imperfect, we believe that these models capture inter-individual differences more accurately than basic empirical scores (e.g., the total number of errors, or the average RT).
+Thus, for each illusion, we extracted the effect of illusion strength and its interaction with task difficulty (when the illusion effect was incongruent) within each participant.
+These twenty participant-level scores were subjected to exploratory factor analysis (EFA), which suggested that ...
 
 ## Discussion
+
+<!-- This is likely to be caused by the outline. It will be fixed in the next release of the Pyllusion software. -->
 
@@
 This study was based on the data collected in study 2.
-However, the variables of interest were here taken from the questionnaires inserted in between the two series of illusion blocks.
-We used the *IPIP6* (x items; *REF*) to measure 6 normal personality traits (extraversion, **complete**).
-The PID-5 (x items; *REF*) was used to measure psychopathological personality traits (..., **complete**).
-The participants were the same as in study 2, however due to a technical issue the first height participants do not have data for the personality questionnaires.
+However, the variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
+We used the *IPIP6* (24 items; [@sibley2011]) to measure 6 personality traits (namely, extraversion, openness, conscientiousness, agreeableness, neuroticism and honesty-humility).
+On the other hand, the PID-5 (25 items; [@hopwood2012]) was used to measure the degree to which participants inhibit psychopathological personality traits (Disinhibition, antagonism, detachment, negative affect and psychoticism).
+The participants were the same as in study 2.
+However, due to a technical issue, no data on the personality questions are available for the first eight participants.
 
 # General Discussion
+
+Using the parametric illusion generation framework we developed, *Pyllusion* [@makowski2021a], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
+This thus opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.
+The validation of this novel framework also affords future illusion scientists a standardized measure of illusion susceptibility, instead of relying on conventional methods that depend upon participants' subjective perceptions.
+In our paradigm, in which we apply this approach to a reaction-time task, we were able to measure inter-individual scores of objective illusion sensitivity.
 

### 2022-09-01 manuscript.Rmd

@@
 <!-- Target, strength, difficulty -->
+Indeed, many visual illusions can be seen as made of *targets* (e.g., same-length lines), which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the segments appear of different lengths when they end with inwards or outwards pointing arrows).
+While most of the paradigms used in the literature prompt the subjective extent the context biases the perception (e.g."how much doe the line appear different"; **REFS**), *Pyllusions* allows to create illusions in which the targets are actually more or less different (e.g., one segment is truly longer than the other), and in which the illusion varies in strength (the arrows are more or less pointy).
 
@@
 <!-- Paradigm -->
-
-This thereby opens the door to a standardizable experimental task in which participants can be asked to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusory effect can be set to either be "incongruent" (increasing the difficulty of the task by biasing the perception in the opposite way) or "congruent" (making the task easier).
-Although visual illusions are inherently tied to subjective perception, this framework radically reverses the traditional paradigm in favor of an approach that could quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
+This opens the door of an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+Moreover, the illusion effect can be either "incongruent" (making the task even harder by biasing the perception in the opposite way) or "congruent" (making the task easier).
+Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm that could quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
 
 <!-- Goals -->
+In the present set of preregistered studies, we will first test this novel paradigm by investigate if the effect of illusion and task difficulty can be manipulated continuously and separated statistically.
+Then, we will use the paradigm to assess whether 10 different historical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate).
+Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
 
@@
 We generated 56 stimuli for each of the 10 illusion types.
-These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where values correspond to the objective difference between targets, i.e., 1 corresponds to the smallest objective difference and thereby the highest level of difficulty in the task) and 7 levels of illusion strength (3 values (i.e, degrees) of strength on the side of illusion congruence, 3 on the side of illusion incongruence, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3].
-Notably, the magnitude of these values correspond to illusion strength (greater values represent greater strength) whereas the valence indicates the illusion congruence (negative sign represents an illusion congruent effect).
+These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where the negative values correspond to congruent illusion strengths).
 
-The 10 types of illusion stimuli were randomly presented in distinct blocks, within which the 56 stimuli were also displayed in a randomized order.
-After the first series of 10 blocks was presented, another series was displayed (with the same set of stimuli but with new randomized block and trial orders).
-All in all, each participant saw 56 different trials per illusion type (10 blocks in total), in 2 series of blocks (amounting to a total of 1120 trials).
-In each illusion trial, participants were required to evaluate the targets comprising the illusions without being distracted by its biasing contextual information.
-For example, in a Müller-Lyer illusion trial, participants were instructed to evaluate the length of line segments without being distracted by the direction in which the arrows at its ends are pointing towards.
-Importantly, participants were told to respond "as fast as possible without making errors" before each block of trials (i.e., an explicit double constraint incorporated to mitigate the inter-individual variability in the speed-accuracy trade off in responses).
-This task was implemented using *jsPsych* [@deleeuw2014].
-For specific details of the instructions given for each illusion type, please refer to the experiment code.
+The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the block was also randomized.
+After the first series of 10 blocks, another series was done (with new randomized blocks and trials order).
+In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
+The task was implemented using *jsPsych* (*REF*).
+See instructions for each illusion type in the experiment code.
 
@@
 The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; ....
-For all of the specified models, the effect of illusion strength, task difficulty and their interaction were significant.
+In all of these models, the effect of illusion strength, task difficulty and their interaction were significant.
+
+For errors, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0.
+
+However, for reaction times, the GAMs suggested a consistent non-linear relationship between the mean RT and illusion strength: as the illusion strength gets very strong, the reaction times are faster.
+While this is is not surprising (strong illusions are so biasing that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this pattern and hence are not good representatives of the underlying dynamics.
+Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, relationship, we decided not to use add them for the individual scores analysis.
+<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size). -->
+
+Though imperfect, we believe that these models capture inter-individual differences more accurately than basic empirical scores (e.g., the total amount of errors, or the average RT).
+Thus, for each illusion, we have extracted the effect of illusion strength and its interaction with task difficulty (when the illusion incongruent) within each participants.
+These twenty participant-level scores were subjected to exploratory factor analysis (EFA), which suggested that ...
 
@@
 <!-- This is likely to be caused by the outline. It will be fixed in the next release of the Pyllusion software. -->
-
 Future studies should investigate reaction times to find the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).

### 2022-09-01 manuscript.Rmd
update manu

@@
 <!-- Target, strength, difficulty -->
-Indeed, many visual illusions can be seen as made of *targets* (e.g., same-length lines), which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the segments appear of different lengths when they end with inwards or outwards pointing arrows).
-While most of the paradigms used in the literature prompt the subjective extent the context biases the perception (e.g."how much doe the line appear different"; **REFS**), *Pyllusions* allows to create illusions in which the targets are actually more or less different (e.g., one segment is truly longer than the other), and in which the illusion varies in strength (the arrows are more or less pointy).
-
 Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
@@
 <!-- Goals -->
-In the present set of preregistered studies, we will first test this novel paradigm by investigate if the effect of illusion and task difficulty can be manipulated continuously and separated statistically.
-Then, we will use the paradigm to assess whether 10 different historical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate).
-Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
-
 In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separated statistically.
@@
 
-For errors, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0.
-
-However, for reaction times, the GAMs suggested a consistent non-linear relationship between the mean RT and illusion strength: as the illusion strength gets very strong, the reaction times are faster.
-While this is is not surprising (strong illusions are so biasing that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this pattern and hence are not good representatives of the underlying dynamics.
-Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, relationship, we decided not to use add them for the individual scores analysis.
-<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size). -->
-
-Though imperfect, we believe that these models capture inter-individual differences more accurately than basic empirical scores (e.g., the total amount of errors, or the average RT).
-Thus, for each illusion, we have extracted the effect of illusion strength and its interaction with task difficulty (when the illusion incongruent) within each participants.
-These twenty participant-level scores were subjected to exploratory factor analysis (EFA), which suggested that ...
-
 With regards to its effect on errors, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0.

### 2022-09-02 manuscript.Rmd
update

@@
 
+We have to keep in mind that individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of ...), 2) the structural equation model from which they are obtained is also constructed on layers of abstractions.
+Thus, we wanted to test wether these indices relate to some simple empirical scores, such as the average error rate and RT in the task.
+
 ## Discussion

### 2022-09-02 study2/study2.Rmd
update

@@
 
-### Model-based Descriptors
+### Extraction
 
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+load("models/model_delboeuf_err.RData")
+load("models/model_ebbinghaus_err.RData")
+load("models/model_rodframe_err.RData")
+load("models/model_verticalhorizontal_err.RData")
+load("models/model_zollner_err.RData")
+load("models/model_white_err.RData")
+load("models/model_mullerlyer_err.RData")
+load("models/model_ponzo_err.RData")
+load("models/model_poggendorff_err.RData")
+load("models/model_contrast_err.RData")
+
 data <- rbind(
@@
     !str_detect(Parameter, "Null_"),
-    !str_detect(Parameter, "Diff_"))
-# unique(data$Parameter)
-
-# data |>
-#   # group_by(Illusion_Type, Parameter) |> 
-#   # standardize(select="Value") |> 
-#   # ungroup() |> 
-#   # mutate(order = mean(Value)) |> 
-#   # ungroup() |> 
-#   # mutate(Participant = fct_reorder(Participant, order)) |> 
-#   ggplot(aes(y = Participant, x = Value)) +
-#   # ggdist::stat_slabinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
-#   ggdist::stat_pointinterval(aes(fill = Illusion_Type, color=Illusion_Type), alpha=0.5) +
-#   # coord_cartesian(xlim=c(-4, 4)) +
-#   facet_wrap(~Parameter, scales = "free_x") +
-#   theme(axis.text.y = element_blank())
-
-# df[df$Illusion_Type == "Poggendorff" & df$Participant == "5cbb77bf418c540001641e67_sp53j", ]
+    !str_detect(Parameter, "Diff_"))  |> 
+  mutate(Parameter = str_remove(Parameter, "_Prob"))
+
+data |>
+  ggplot(aes(y = Median, x = Participant)) +
+  geom_pointrange(aes(ymin=CI_low, ymax=CI_high, color=Parameter), size=0.5) +
+  geom_hline(yintercept = 0, linetype = "dotted") +
+  facet_wrap(Parameter~Illusion_Type, scales = "free") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  coord_flip()
 
@@
          Parameter = clean_illusionName(Parameter)) |>
-  group_by(Participant, Parameter) |>
-  summarize(Score = median(Value, na.rm = TRUE), .groups = "drop") |>
+  select(Participant, Parameter, Median) |> 
+  pivot_wider(names_from = Parameter, values_from = Median) |> 
+  mutate_if(is.numeric, function(x) x / sd(x, na.rm=TRUE))
+
+random[is.na(random)] <- 0
+```
+
+### Correlation
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+cor <- correlation::correlation(random, redundant = TRUE, p_adjust = "none")
+
+p_data <- cor |>
+  mutate(
+    Parameter1 = prettify_parameterName(Parameter1),
+    Parameter2 = prettify_parameterName(Parameter2)
+  ) |>
+  correlation::cor_sort(hclust_method = "ward.D2") |>
+  cor_lower() |>
+  mutate(
+    Text = insight::format_value(r, zap_small = TRUE, digits = 3),
+    Text = str_replace(str_remove(Text, "^0+"), "^-0+", "-"),
+    Text = paste0(Text, insight::format_p(p, stars_only = TRUE)),
+    Parameter2 = fct_rev(Parameter2)
+  )
+
+p_data |>
+  ggplot(aes(x = Parameter2, y = Parameter1)) +
+  geom_tile(aes(fill = r)) +
+  geom_text(aes(label = Text), size = rel(2), alpha=2/3) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  scale_x_discrete(expand = c(0, 0)) +
+  scale_y_discrete(expand = c(0, 0)) +
+  labs(title = "Correlation Matrix", x = NULL, y = NULL) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 90, hjust = 1),
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank()
+  )
+
+# estimate_density(p_data$r, at=as.character(sign(p_data$r))) |> 
+#   ggplot(aes(x=x, y=y, color=Group)) +
+#   geom_line()
+```
+
+### Factor Analysis
+
+#### First Level
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+data <- select(random, -Participant)
+
+rez <- parameters::n_factors(data, n_max=15)
+plot(rez)
+
+fa <- parameters::factor_analysis(data,
+  cor = as.matrix(cor),
+  n = 7,
+  rotation = "oblimin",
+  fm = "mle",
+  sort = TRUE
+)
+
+insight::print_md(fa)
+
+plot(fa) +
+  theme_minimal()
+# psych::omega(data, fm = "mle", nfactors=7)
+# fa <- psych::fa.multi(as.matrix(cor), nfactors = 3, nfact2 = 1, n.obs = nrow(random))
+# psych::fa.multi.diagram(fa)
+```
+
+#### Second Level
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+data <- predict(fa)
+
+rez <- parameters::n_factors(data, rotation = "varimax")
+plot(rez)
+
+fa2 <- parameters::factor_analysis(data,
+  n = 2,
+  rotation = "varimax",
+  fm = "mle",
+  sort = TRUE
+)
+
+insight::print_md(fa2)
+```
+
+### Sructural Equation Model
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+library(lavaan)
+
+names <- names(select(random, -Participant))
+data <- setNames(as.data.frame(str_split_fixed(names, "_", 2)), c("Illusion_Type", "Parameter"))
+data$Name <- names
+
+# Model 0 - EFA
+model <- "
+I2 =~ RodFrame_Strength + RodFrame_Interaction + VerticalHorizontal_Strength + VerticalHorizontal_Interaction + MullerLyer_Strength + MullerLyer_Interaction + Ponzo_Strength + Ponzo_Interaction
+Zollner =~ Zollner_Strength + Zollner_Interaction
+White =~ White_Strength + White_Interaction
+Ebbinghaus =~ Ebbinghaus_Strength + Ebbinghaus_Interaction
+Poggendorff =~ Poggendorff_Strength + Poggendorff_Interaction
+Contrast =~ Contrast_Strength + Contrast_Interaction
+Delboeuf =~ Delboeuf_Strength + Delboeuf_Interaction
+I =~ I2 + Zollner + White + Ebbinghaus + Poggendorff + Contrast
+"
+
+m0 <- sem(model = model,
+          data  = random,
+          bounds = "standard",
+          estimator = "ML",
+          )
+
+# Model 1 - by Illusion
+lvl1 <- paste(data$Illusion_Type, "=~", data$Name)
+lvl2 <- paste("I =~", unique(data$Illusion_Type))
+resid <- as.data.frame(t(utils::combn(x=unique(data$Illusion_Type), 2)))
+resid <- paste(resid$V1, "~~", resid$V2)
+
+m1 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
+          data  = random,
+          bounds = "standard",
+          estimator = "ML",
+          )
+
+# Model 1b - without Delboeuf
+m1b <- sem(model = paste0(c(lvl1, lvl2[lvl2 != "I =~ Delboeuf"]), collapse = "\n"),
+          data  = random,
+          bounds = "standard",
+          estimator = "ML",
+          )
+
+# Model 2 - by type
+model <- "
+Lines =~ MullerLyer_Strength + MullerLyer_Interaction + Ponzo_Strength + Ponzo_Interaction + VerticalHorizontal_Strength + VerticalHorizontal_Interaction
+Circles =~ Ebbinghaus_Strength + Ebbinghaus_Interaction + Delboeuf_Strength + Delboeuf_Interaction
+Contrast =~ Contrast_Strength + Contrast_Interaction + White_Strength + White_Interaction
+Angle =~ Zollner_Strength + Zollner_Interaction + Poggendorff_Strength + Poggendorff_Interaction  + RodFrame_Strength + RodFrame_Interaction
+I =~ Lines + Circles + Contrast + Angle 
+"
+
+m2 <- sem(model = model,
+          data  = random,
+          bounds = "standard",
+          estimator = "ML"
+          )
+
+# Model 3 - by Parameter
+lvl1 <- paste(data$Parameter, "=~", data$Name)
+lvl2 <- paste("I =~", unique(data$Parameter))
+resid <- as.data.frame(t(utils::combn(x=unique(data$Parameter), 2)))
+resid <- paste(resid$V1, "~~", resid$V2)
+
+m3 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
+          data  = random,
+          bounds = "standard",
+          estimator = "ML"
+          )
+
+anova(m0, m1, m2)
+
+summary(m1, standardize = TRUE, fit.measures = TRUE)
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# https://cjvanlissa.github.io/tidySEM/articles/sem_graph.html
+library(ggraph)
+library(tidySEM)
+
+colors <- c("I" = "#9C27B0", 
+            "Delboeuf"="#F44336", "Ebbinghaus"="#E91E63",
+            "Rod-Frame"="#4CAF50", "Rod\nFrame"="#4CAF50", 
+            "Vertical-Horizontal"="#FF9800", "Vertical\nHorizontal"="#FF9800",
+            "Zöllner"="#FF5722",
+            "White"="#616161", 
+             "Müller-Lyer" = "#3F51B5", "Müller\n-Lyer" = "#3F51B5",
+            "Ponzo" = "#2196F3",
+            "Poggendorff"="#795548",
+            "Contrast"="#607D8B", 
+            "black"="black")
+
+edges <- tidySEM::get_edges(m1) |> 
+  mutate(color = ifelse(sign(as.numeric(est)) >= 0, "1", "-1"),
+         width = abs(as.numeric(est)),
+         type = ifelse(from == "I", "curved", "straight")) |> 
+  filter(lhs != rhs)
+
+nodes <- tidySEM::get_nodes(m1) |> 
+  mutate(label = prettify_parameterName(name),
+         angle = ifelse(grepl("_", name), 90, 0),
+         hjust = case_when(
+           grepl("_", name) ~ 1.15,
+           name == "I" ~ 0.5,
+           TRUE ~ 0.5),
+         size = case_when(
+           grepl("_", name) ~ 1,
+           name == "I" ~ 7,
+           TRUE ~ 6),
+         textsize = case_when(
+           grepl("_", name) ~ 1,
+           name == "I" ~ 3,
+           TRUE ~ 1.1),
+         face = case_when(
+           grepl("_", name) ~ "italic",
+           name == "I" ~ "bold",
+           TRUE ~ "plain"),
+         color = str_remove_all(label, " - .*"),
+         textcolor = ifelse(grepl("_", name), "black", "white"),
+         label = str_remove_all(label, ".* - "),
+         label = str_replace(label, "Müller-Lyer", "Müller\n-Lyer"),
+         label = str_replace(label, "Rod-Frame", "Rod\nFrame"),
+         label = str_replace(label, "Vertical-Horizontal", "Vertical\nHorizontal"),
+         family = ifelse(name == "I", "serif", "sans"))
+
+p <- tidygraph::tbl_graph(nodes = nodes, edges = edges) |> 
+  ggraph(layout = "sugiyama") +
+  geom_edge_arc(aes(filter=type!="straight",
+                    edge_width = width,
+                    color = color,
+                    label=est_std),
+                strength = 0.03 * c(-1, -1, 1, 1, 1, 1, -1, -1, -1, -1),
+                angle_calc="along",
+                label_dodge=unit(-0.015, "npc")) +
+  geom_edge_link(aes(filter=type=="straight",
+                     edge_width = width, 
+                     color = color,
+                     label=est_std), 
+                 angle_calc="along", 
+                 label_dodge=unit(-0.015, "npc")) +
+  geom_node_point(aes(shape = shape, color=color, size=size)) +
+  geom_node_text(aes(label = label, angle = angle, hjust=hjust),
+                 color=nodes$textcolor, 
+                 fontface=nodes$face,
+                 size = rel(nodes$textsize*3),
+                 family=nodes$family) +
+  scale_y_continuous(expand = expansion(c(.6, .1))) +
+  scale_edge_width_continuous(range=c(0.1, 1.5)) +
+  scale_edge_color_manual(values=c("1"="#263238", "-1"="#B71C1C")) +
+  scale_color_manual(values=colors, guide="none") +
+  scale_shape_manual(values=c("oval"="circle", "rect"="square"), guide="none") +
+  scale_size_continuous(range = c(2, 27), guide="none") +
+  guides(edge_colour = "none", edge_width = "none") + 
+  labs(title = "Structural Equation Model")  +
+  theme_graph() +
+  theme(plot.title = element_text(hjust=0.5))
+p
+save(p, file="models/sem.Rdata")
+
+dfsem <- mutate(as.data.frame(predict(m1)), Participant = random$Participant)
+```
+
+### Empirical Descriptors
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Average for 4 higher level of task difficulty
+empirical <- df |>
+  filter(Illusion_Effect == "Incongruent") |> 
+  group_by(Participant, Illusion_Type, Illusion_Strength) |>
+  arrange(Illusion_Difference) |> 
+  slice(0:2) |> 
+  # select(Illusion_Strength, Illusion_Difference)
+  group_by(Participant, Illusion_Type) |>
+  summarize(
+    TaskHard_Error = sum(Error, na.rm=TRUE) / n(),
+    TaskHard_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
+    # Illu_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
+  ) |>
   ungroup() |> 
-  group_by(Parameter) |>
-  mutate(Score = Score / sd(Score, na.rm = TRUE)) |>  # Scale
+  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
+  pivot_wider(names_from = c("Illusion_Type"), values_from = c("TaskHard_Error", "TaskHard_RTMean"), names_glue = "{Illusion_Type}_{.value}")
+
+# Average for 4 higher level of illusion strength
+empirical <- df |>
+  filter(Illusion_Effect == "Incongruent") |> 
+  group_by(Participant, Illusion_Type, Illusion_Difference) |>
+  arrange(desc(Illusion_Strength)) |> 
+  slice(0:2) |> 
+  # select(Illusion_Strength, Illusion_Difference)
+  group_by(Participant, Illusion_Type) |>
+  summarize(
+    IlluStrong_Error = sum(Error, na.rm=TRUE) / n(),
+    IlluStrong_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
+    # Diff_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
+  ) |>
   ungroup() |> 
-  pivot_wider(names_from = Parameter, values_from = Score)
-
-# plot(estimate_density(select(random, -Participant)))
-```
+  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
+  pivot_wider(names_from = c("Illusion_Type"), values_from = c("IlluStrong_Error", "IlluStrong_RTMean"), names_glue = "{Illusion_Type}_{.value}") |> 
+  full_join(empirical, by = "Participant") 
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+r <- correlation(arrange(dfsem, Participant), 
+                 arrange(empirical, Participant), 
+                 sort=TRUE, p_adjust="none")
+
+r |>
+  mutate(label = insight::format_p(p, stars_only=TRUE),
+         Type = ifelse(str_detect(Parameter2, "_Error"), "Error", "RT")) |> 
+  ggplot(aes(x = Parameter1, y = Parameter2)) +
+  geom_tile(aes(fill = r)) +
+  geom_text(aes(label=label)) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  facet_wrap(~Type, scales="free") +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Correlation
+r <- data.frame()
+for(ill in c("Delboeuf", 
+             "Ebbinghaus",
+             "RodFrame",
+             "VerticalHorizontal",
+             "Zollner",
+             "White",
+             "MullerLyer",
+             "Ponzo",
+             "Poggendorff",
+             "Contrast"
+             )) {
+  r <- correlation::correlation(
+    data = dfsem |> 
+      arrange(Participant),
+    data2 = empirical |>
+      data_select(select=ill, regex=TRUE),
+    p_adjust="none"
+    ) |> 
+    mutate(Illusion_Type = ill,
+           Parameter1 = prettify_parameterName(Parameter1),
+           Parameter2 = prettify_parameterName(Parameter2)) |> 
+    select(Illusion_Type, Parameter1, Parameter2, r, p) |> 
+    rbind(r)
+}
+
+r |>
+  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
+  ggplot(aes(x = Parameter1, y = Parameter2)) +
+  geom_tile(aes(fill = r)) +
+  geom_text(aes(label=label)) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  # facet_wrap(~Illusion_Type, scales = "free") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+```
+
+### GARBAGE
+<!-- ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE} -->
+<!-- rez <- parameters::n_factors(select(empirical, -Participant), n_max=30) -->
+<!-- plot(rez) -->
+
+<!-- r <- correlation::correlation(empirical, p_adjust = "none", sort=TRUE) -->
+
+<!-- r |> -->
+<!--   ggplot(aes(x = Parameter2, y = Parameter1)) + -->
+<!--   geom_tile(aes(fill = r)) + -->
+<!--   # geom_text(aes(label = Text), size = 2) + -->
+<!--   scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") + -->
+<!--   scale_x_discrete(expand = c(0, 0)) + -->
+<!--   scale_y_discrete(expand = c(0, 0)) + -->
+<!--   labs(title = "Correlation Matrix", x = NULL, y = NULL) + -->
+<!--   theme_minimal() + -->
+<!--   theme( -->
+<!--     axis.text.x = element_text(angle = 90, hjust = 1), -->
+<!--     plot.title = element_text(hjust = 0.5, face = "bold"), -->
+<!--     panel.grid.major = element_blank(), -->
+<!--     panel.grid.minor = element_blank() -->
+<!--   ) -->
+
+<!-- fa <- parameters::factor_analysis(select(empirical, -Participant), -->
+<!--   n = 13, -->
+<!--   rotation = "oblimin", -->
+<!--   fm = "mle", -->
+<!--   sort = TRUE -->
+<!-- ) -->
+
+<!-- print(tibble::as_tibble(fa), n=30) -->
+<!-- plot(fa) -->
+<!-- ``` -->
 
@@
 
-## Structure
-
-### Correlation
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-cor <- correlation::correlation(random, redundant = TRUE)
-
-p_data <- cor |>
-  mutate(
-    Parameter1 = prettify_parameterName(Parameter1),
-    Parameter2 = prettify_parameterName(Parameter2)
-  ) |>
-  correlation::cor_sort(hclust_method = "ward.D2") |>
-  cor_lower() |>
-  mutate(
-    Text = insight::format_value(r, zap_small = TRUE, digits = 3),
-    Text = str_replace(str_remove(Text, "^0+"), "^-0+", "-"),
-    Parameter2 = fct_rev(Parameter2)
-  )
-
-p_data |>
-  ggplot(aes(x = Parameter2, y = Parameter1)) +
-  geom_tile(aes(fill = r)) +
-  # geom_text(aes(label = Text), size = 2) +
-  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
-  scale_x_discrete(expand = c(0, 0)) +
-  scale_y_discrete(expand = c(0, 0)) +
-  labs(title = "Correlation Matrix", x = NULL, y = NULL) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 90, hjust = 1),
-    plot.title = element_text(hjust = 0.5, face = "bold"),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank()
-  )
-```
-
-### Factor Analysis
-
-#### First Level
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-data <- select(random, -Participant)
-
-rez <- parameters::n_factors(data, n_max=15)
-plot(rez)
-
-fa <- parameters::factor_analysis(data,
-  cor = as.matrix(cor),
-  n = 9,
-  rotation = "oblimin",
-  fm = "mle",
-  sort = TRUE
-)
-
-print(tibble::as_tibble(fa), n=30)
-plot(fa)
-# psych::omega(data, fm = "mle", nfactors=9)
-# fa <- psych::fa.multi(as.matrix(cor), nfactors = 3, nfact2 = 1, n.obs = nrow(random))
-# psych::fa.multi.diagram(fa)
-```
-
-#### Second Level
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-data <- predict(fa)
-
-rez <- parameters::n_factors(data, rotation = "oblimin")
-plot(rez)
-
-fa2 <- parameters::factor_analysis(data,
-  n = 1,
-  rotation = "varimax",
-  fm = "mle",
-  sort = TRUE
-)
-
-# psych::schmid(attributes(fa)$model, nfactors = 1)
-```
-
-#### Visualization
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
@@
 
-### SEM
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-library(lavaan)
-library(tidySEM)
-library(ggraph)
-
-names <- names(select(random, -Participant))
-data <- setNames(as.data.frame(str_split_fixed(names, "_", 3)), c("Illusion_Type", "Parameter", "Model"))
-data$Name <- names
-
-# Model 1 - by Illusion
-lvl1 <- paste(data$Illusion_Type, "=~", data$Name)
-lvl2 <- paste("I =~", unique(data$Illusion_Type))
-resid <- as.data.frame(t(utils::combn(x=unique(data$Illusion_Type), 2)))
-resid <- paste(resid$V1, "~~", resid$V2)
-
-m1 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
-          data  = random,
-          estimator = "ML",
-          missing = "ml",
-          bounds = TRUE)  # robust.two.stage
-
-# Model 2 - by Parameter
-lvl1 <- paste(data$Parameter, "=~", data$Name)
-lvl2 <- paste("I =~", unique(data$Parameter))
-resid <- as.data.frame(t(utils::combn(x=unique(data$Parameter), 2)))
-resid <- paste(resid$V1, "~~", resid$V2)
-
-m2 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
-          data  = random)
-
-# # Model 3
-# lvl1 <- paste(data$Model, "=~", data$Name)
-# lvl2 <- paste("G =~", unique(data$Model))
-# 
-# m2 <- sem(model = paste0(c(lvl1, lvl2), collapse = "\n"),
-#            data  = random)
-
-anova(m1, m2)
-
-summary(m1, standardize = TRUE)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-# https://cjvanlissa.github.io/tidySEM/articles/sem_graph.html
-edges <- tidySEM::get_edges(m1) |> 
-  mutate(color = ifelse(sign(as.numeric(est)) >= 0, "1", "-1"),
-         width = abs(as.numeric(est))) |> 
-  filter(lhs != rhs)
-
-nodes <- tidySEM::get_nodes(m1) |> 
-  mutate(label = prettify_parameterName(name),
-         angle = ifelse(grepl("_", name), 90, 0),
-         hjust = ifelse(grepl("_", name), 1.01, 0))
-
-tidygraph::tbl_graph(nodes = nodes,
-                     edges = edges) |> 
-  ggraph(layout = "sugiyama") +
-  geom_edge_link(aes(edge_width = width, 
-                             color = color,
-                             label=est_sig_std), angle_calc="along", label_dodge=unit(-0.015, "npc")) +
-  geom_node_point() +
-  geom_node_text(aes(label = label, angle = angle, hjust=hjust)) +
-  scale_edge_color_manual(values=c("#F44336", "#2196F3"))  +
-  scale_y_continuous(expand = expansion(c(.6, .1))) +
-  scale_edge_width_continuous(range=c(0.1, 1.5)) +
-  guides(edge_colour = "none", edge_width = "none") +
-  theme_graph()
-
-dfsem <- mutate(as.data.frame(predict(m1)), Participant = random$Participant)
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-r <- correlation(arrange(dfsem, Participant), arrange(empirical, Participant), sort=TRUE, p_adjust="none")
-
-r |>
-  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
-  ggplot(aes(x = Parameter1, y = Parameter2)) +
-  geom_tile(aes(fill = r)) +
-  geom_text(aes(label=label)) +
-  scale_alpha_continuous(range = c(1, 0.5)) +
-  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
-```
-
-### Empirical Descriptors
-
-We have to keep in mind that individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of ...), 2) the structural equation model from which they are obtained is also constructed on layers of abstractions.
-Thus, we wanted to test wether these indices relate to some simple empirical scores, such as the average error rate and RT in the task.
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-# Average for 4 higher level of task difficulty
-empirical <- df |>
-  filter(Illusion_Effect == "Incongruent") |> 
-  group_by(Participant, Illusion_Type, Illusion_Strength) |>
-  arrange(Illusion_Difference) |> 
-  slice(0:4) |> 
-  # select(Illusion_Strength, Illusion_Difference)
-  group_by(Participant, Illusion_Type) |>
-  summarize(
-    TaskHard_Error = sum(Error, na.rm=TRUE) / n(),
-    TaskHard_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
-    # Illu_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
-  ) |>
-  ungroup() |> 
-  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
-  pivot_wider(names_from = c("Illusion_Type"), values_from = c("TaskHard_Error", "TaskHard_RTMean"), names_glue = "{Illusion_Type}_{.value}")
-
-# Average for 4 higher level of illusion strength
-empirical <- df |>
-  filter(Illusion_Effect == "Incongruent") |> 
-  group_by(Participant, Illusion_Type, Illusion_Difference) |>
-  arrange(desc(Illusion_Strength)) |> 
-  slice(0:4) |> 
-  # select(Illusion_Strength, Illusion_Difference)
-  group_by(Participant, Illusion_Type) |>
-  summarize(
-    IlluStrong_Error = sum(Error, na.rm=TRUE) / n(),
-    IlluStrong_RTMean = mean(ifelse(Error == 0, RT, NA), na.rm=TRUE)
-    # Diff_RTSpread = sd(ifelse(Error == 0, RT, NA), na.rm=TRUE),
-  ) |>
-  ungroup() |> 
-  mutate(Illusion_Type = clean_illusionName(Illusion_Type))  |> 
-  pivot_wider(names_from = c("Illusion_Type"), values_from = c("IlluStrong_Error", "IlluStrong_RTMean"), names_glue = "{Illusion_Type}_{.value}") |> 
-  full_join(empirical, by = "Participant") 
-```
-
-<!-- ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE} -->
-<!-- rez <- parameters::n_factors(select(empirical, -Participant), n_max=30) -->
-<!-- plot(rez) -->
-
-<!-- r <- correlation::correlation(empirical, p_adjust = "none", sort=TRUE) -->
-
-<!-- r |> -->
-<!--   ggplot(aes(x = Parameter2, y = Parameter1)) + -->
-<!--   geom_tile(aes(fill = r)) + -->
-<!--   # geom_text(aes(label = Text), size = 2) + -->
-<!--   scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") + -->
-<!--   scale_x_discrete(expand = c(0, 0)) + -->
-<!--   scale_y_discrete(expand = c(0, 0)) + -->
-<!--   labs(title = "Correlation Matrix", x = NULL, y = NULL) + -->
-<!--   theme_minimal() + -->
-<!--   theme( -->
-<!--     axis.text.x = element_text(angle = 90, hjust = 1), -->
-<!--     plot.title = element_text(hjust = 0.5, face = "bold"), -->
-<!--     panel.grid.major = element_blank(), -->
-<!--     panel.grid.minor = element_blank() -->
-<!--   ) -->
-
-<!-- fa <- parameters::factor_analysis(select(empirical, -Participant), -->
-<!--   n = 13, -->
-<!--   rotation = "oblimin", -->
-<!--   fm = "mle", -->
-<!--   sort = TRUE -->
-<!-- ) -->
-
-<!-- print(tibble::as_tibble(fa), n=30) -->
-<!-- plot(fa) -->
-<!-- ``` -->
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-# Correlation
-r <- data.frame()
-for(ill in c("Delboeuf", 
-             "Ebbinghaus",
-             "RodFrame",
-             "VerticalHorizontal",
-             "Zollner",
-             "White",
-             "MullerLyer",
-             "Ponzo",
-             "Poggendorff",
-             "Contrast"
-             )) {
-  r <- correlation::correlation(
-    data = random |> 
-      arrange(Participant) |> 
-      data_select(select=ill, regex=TRUE),
-    data2 = empirical |>
-      data_select(select=ill, regex=TRUE),
-    p_adjust="none"
-    ) |> 
-    mutate(Illusion_Type = ill,
-           Parameter1 = prettify_parameterName(Parameter1),
-           Parameter2 = prettify_parameterName(Parameter2)) |> 
-    select(Illusion_Type, Parameter1, Parameter2, r, p) |> 
-    rbind(r)
-}
-
-r |>
-  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
-  ggplot(aes(x = Parameter2, y = Parameter1)) +
-  geom_tile(aes(fill = r)) +
-  geom_text(aes(label=label)) +
-  scale_alpha_continuous(range = c(1, 0.5)) +
-  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
-  # facet_wrap(~Illusion_Type, scales = "free") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
-```
-
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
@@
   merge(dfsem, by = "Participant") |> 
-  merge(empirical, by = "Participant") |> 
   write.csv("../data/study3.csv", row.names = FALSE)

### 2022-09-02 study2/functions.R

@@ -584,7 +584,7 @@ plot_all <- function(data, p_err, p_rt) {
 
 extract_random <- function(model, illusion = "Delboeuf") {
   random <- as.data.frame(model)
-  random <- random[sample(nrow(random), 300), ]
+  # random <- random[sample(nrow(random), 300), ]
   random <- random[str_detect(names(random), regex("^r_Participant"))]
 
   if (insight::model_info(model)$is_logit) {
@@ -593,9 +593,24 @@ extract_random <- function(model, illusion = "Delboeuf") {
     param <- "Loc,"
   }
 
+  # All draws
+  # random |>
+  #   mutate(Draw = 1:nrow(random)) |>
+  #   pivot_longer(-Draw, names_to = "Parameter", values_to = "Value") |>
+  #   mutate(
+  #     Parameter = clean_parameterName(Parameter),
+  #     Parameter = paste0(param, Parameter),
+  #     Parameter = str_replace(Parameter, paste0(param, "__sigma"), "Disp,")
+  #   ) |>
+  #   separate(Parameter, into = c("Component", "Participant", "Parameter"), sep = ",") |>
+  #   mutate(
+  #     Parameter = paste0(Parameter, "_", Component),
+  #     Illusion_Type = illusion
+  #   )
+
+  # Summary
   random |>
-    mutate(Draw = 1:nrow(random)) |>
-    pivot_longer(-Draw, names_to = "Parameter", values_to = "Value") |>
+    describe_posterior(test=NULL) |>
     mutate(
       Parameter = clean_parameterName(Parameter),
       Parameter = paste0(param, Parameter),
@@ -610,11 +625,11 @@ extract_random <- function(model, illusion = "Delboeuf") {
 
 
 clean_illusionName <- function(x) {
-  x <- str_replace(x, "Rod-Frame", "RodFrame")
-  x <- str_replace(x, "Vertical-Horizontal", "VerticalHorizontal")
-  x <- str_replace(x, "Zöllner", "Zollner")
-  x <- str_replace(x, "Müller-Lyer", "MullerLyer")
-  x
+  x |>
+    str_replace("Rod-Frame", "RodFrame") |>
+    str_replace("Vertical-Horizontal", "VerticalHorizontal") |>
+    str_replace("Zöllner", "Zollner") |>
+    str_replace("Müller-Lyer", "MullerLyer")
 }
 
[…]

### 2022-09-02 manuscript.Rmd

@@
 Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems.
-They eloquently show that our brains do not simply map veridical accounts of physical reality; we generate perceptions of the environment by taking into consideration the contextual information and the prior knowledge that manifests in our subjective conscious experience [@carbon2014].
+They elegantly demonstrate that our brains do not simply map veridical accounts of physical reality; we generate perceptions of the environment by taking the contextual information into consideration, and the prior knowledge that manifests in our subjective conscious experience [@carbon2014].
 Despite the withstanding interest in illusions within various fields of study, including that of visual perception [@day1972; @eagleman2001; @gomez-villa2022 ], consciousness science [@caporuscio2022, @lamme2020] and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015re], several gaps still remain in the present research.
@@
 Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
-Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder (ASD)) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
+Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder [ASD]) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
 
@@
 <!-- Target, strength, difficulty -->
+
 Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-In past studies, the implemented paradigms have mostly involved prompting the subjective extent the context biases the perception (for instance, by getting participants to adjust the parameters of one comparison stimulus (the size of a circle) to another (circle) e.g., @mylniec2016a; @grzeczkowski2018).
-Alternatively, *Pyllusion* allows us to create illusions while manipulating specific features of interest (e.g., size, length etc., ) and thereby the difficulty in distinguishing them, so that the targets are objectively distinct (i.e., one line segment is physically longer than the other), and the strength of the generated illusion are variable (i.e., the biasing angle of the arrows are more or less acute).
+Past studies have implemented paradigms emphasizing the subjective perception of the participant, to examine the extent to which individual context biases perceptual variance (for instance, having participants adjust the parametres of the target stimulus to a reference stimulus relying only on their perception, eg., @mylniec2016a; @grzeczkowski2018).
+Alternatively, Pyllusion allows us to create illusions while manipulating specific features of interest (e.g., size, length etc.), and thereby the difficulty in distinguishing them, so that the targets are objectively distinct (i.e., one line segment is physically longer than the other), and the strength of the generated illusion are variable (i.e., the biasing angle of the arrows are more or less acute).
 
 <!-- Paradigm -->
+
 This opens the door of an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
@@
 <!-- Goals -->
+
 In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separated statistically.
-Then, we will further utilize the paradigm to assess whether 10 different cannonical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate, [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]).
+Then, we will further utilize the paradigm to assess whether 10 different canonical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate, [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]).
 Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
@@
 
-The objective of Study 1 is to act as a pilot study in order to gather some preliminary data to 1) assess whether the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., increase of task difficulty and illusion strength leads to an increased error rate) and 2) develop a gauged understanding of the magnitude of effects, to then refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard).
+The objective of Study 1 is to act as a pilot study in order to gather some preliminary data to: 1) assess whether the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., increase of task difficulty and illusion strength leads to an increased error rate) and 2) develop a gauged understanding of the magnitude of effects, to then refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard).
 
@@
 Fifty-two participants were recruited via *Prolific* ([www.prolificacademic.co.uk](www.prolificacademic.co.uk)), a popular crowd-sourcing platform to recruit subjects for online behavioural research.
-In a study comparing across different research recruitment platforms (e.g., MTurks, Qualtrics etc.,,), participants from *Prolific* were found to provide the highest data quality across all examined measures, including that of attention (the extent to which participants actually read the questions prior to responding), comprehension (the extent to which participants were able to understand the task instructions or items), honesty (the degree to which participants responded to questions truthfully) and reliability (the extent to which participants' responses are internally consistent within the same scale) [@peer2022].
+In a study comparing across different research recruitment platforms (e.g., MTurks, Qualtrics etc.), participants from *Prolific* were found to provide the highest data quality across all examined measures, including that of attention (the extent to which participants actually read the questions prior to responding), comprehension (the extent to which participants were able to understand the task instructions or items), honesty (the degree to which participants responded to questions truthfully) and reliability (the extent to which participants' responses are internally consistent within the same scale) [@peer2022].
 For the purposes of our study, no restrictions on age or country were imposed on participant recruitment; the only requisite was a "fluent" proficiency in English to ensure that the task instructions, given in English, would be well-comprehended.
-Participants were incentivized with a reward of about \textsterling 7.5 for completing the task, which took an average of an hour to finish.
+Participants were incentivised with a reward of about \textsterling 7.5 for completing the task, which took an average of an hour to finish.
 
@@
 Using the same recruitment procedure as in study 1, we recruited 256 participants, within which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
-Please see the next study (study 3) for the full demographic breakdown.
+Please see study 3 for the full demographic breakdown.
 We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean).
@@
 The first part of the analysis focuses on modelling the effect of illusion strength and task difficulty on errors and reaction time, within each illusion.
-In order to achieve that, we will start by fitting a General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
+In order to achieve that, we will start by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
 Errors were analyzed using Bayesian binomial mixed models, and reaction times (RT) of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].
@@
 <!-- This is likely to be caused by the outline. It will be fixed in the next release of the Pyllusion software. -->
+
 Future studies should investigate reaction times to find the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
@@
 We used the *IPIP6* (24 items; [@sibley2011]) to measure 6 personality traits (namely, extraversion, openness, conscientiousness, agreeableness, neuroticism and honesty-humility).
-On the other hand, the PID-5 (25 items; [@hopwood2012]) was used to measure the degree to which participants inhibit psychopathological personality traits (Disinhibition, antagonism, detachment, negative affect and psychoticism).
+On the other hand, the PID-5 (25 items; [@hopwood2012]) was used to measure the degree to which participants inhibit psychopathological personality traits (disinhibition, antagonism, detachment, negative affect and psychoticism).
 The participants were the same as in study 2.
@@
 
-Using the parametric illusion generation framework we developed, *Pyllusion* [@makowski2021a], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
-This thus opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.
+Using the parametric illusion generation framework we developed, Pyllusion [@makowski2021a], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
+This opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.
 The validation of this novel framework also affords future illusion scientists a standardized measure of illusion susceptibility, instead of relying on conventional methods that depend upon participants' subjective perceptions.
@@
 We strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
-The task, data and analysis script are available in open-access at **https://github.com/RealityBending/IllusionGameValidation**.
+The task, data and analysis script are available in open-access at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
 

### 2022-09-05 study2/study2.Rmd
update

@@
 
-```{r demographics, fig.width=20, fig.height=15}
-p1 <- plot_distribution(dfsub, "Age", fill = "#FF9800")
-p2 <- plot_distribution(dfsub, "Duration", title = "Total Duration", subtitle = "in minutes", fill = "#F44336")
-p3 <- plot_distribution(dfsub[dfsub$Break_Duration < 3, ], "Break_Duration", title = "Break Duration", subtitle = "in minutes", fill = "#3F51B5")
-
-p4 <- plot_waffle(dfsub, "Sex") +
-  scale_color_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))
-
-p5 <- plot_waffle(dfsub, "Education") +
-  scale_color_viridis_d()
-
-p6 <- plot_waffle(dfsub, "Nationality") +
-  scale_color_material_d()
-  # scale_fill_manual(values = c("South American" = "#FF5722", "Middle Eastern" = "#FF9800", "Western" = "#2196F3", "Eastern" = "#673AB7", "African" = "#4CAF50", "South African" = "#009688"))
-
-p7 <- plot_waffle(dfsub, "Ethnicity") +
-  scale_color_manual(values = c("Latino" = "#FF5722", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Other" = "#795548"))
-
-p8 <- plot_waffle(dfsub, "Screen_Resolution", title = "Screen Resolution") +
-  scale_color_pizza_d() +
-  guides(fill = "none")
-
-p9 <- plot_waffle(dfsub, "Device_OS", title = "Device OS") +
-  scale_color_bluebrown_d()
-
-# p10 <- plot_waffle(dfsub, "Screen_Refresh") +
-#   scale_fill_viridis_d()
-
-(p1 / p2) | (p4 / p5 / p9) | (p6 / p7 / p8)
-```
-
 # Results
@@
 p_delboeuf_rt <- plot_model_rt(data, model_delboeuf_rt, gam_delboeuf_rt)
-p_delboeuf_rt
 ```
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
-p_white_rt <- plot_model_rt(data, model_white_rt, gam_white_rt)
+p_white_rt <- plot_model_rt(data, model_white_rt, gam_white_rt) + 
+  coord_cartesian(ylim=c(0.125, 1.7), xlim = c(-17.5, 17.5))
 ```
@@
 
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
+p_mullerlyer <- plot_all(data, p_mullerlyer_err, p_mullerlyer_rt)
+p_mullerlyer
+```
+
 ### Ponzo
@@
 
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
+p_ponzo <- plot_all(data, p_ponzo_err, p_ponzo_rt)
+p_ponzo
+```
+
 ### Poggendorff
@@
 
-save(model_ponzo_rt, file="models/model_poggendorff_rt.Rdata")
+save(model_poggendorff_rt, file="models/model_poggendorff_rt.Rdata")
 ```
@@
 p_poggendorff_rt <- plot_model_rt(data, model_poggendorff_rt, gam_poggendorff_rt)
+```
+
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
+p_poggendorff <- plot_all(data, p_poggendorff_err, p_poggendorff_rt)
+p_poggendorff
 ```
@@
 
-plot_ppcheck(model_contrast_rt)
+plot_ppcheck(model_contrast_rt, gam_contrast_rt)
 # performance::check_model(model_contrast_rt)
@@
 
+#### Visualization
+
+```{r message=FALSE, warning=FALSE, cache=cache, fig.height=fig.height*2, eval=TRUE}
+p_contrast <- plot_all(data, p_contrast_err, p_contrast_rt)
+p_contrast
+```
+
 ## Individual Scores
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-load("models/model_delboeuf_err.RData")
-load("models/model_ebbinghaus_err.RData")
-load("models/model_rodframe_err.RData")
-load("models/model_verticalhorizontal_err.RData")
-load("models/model_zollner_err.RData")
-load("models/model_white_err.RData")
-load("models/model_mullerlyer_err.RData")
-load("models/model_ponzo_err.RData")
-load("models/model_poggendorff_err.RData")
-load("models/model_contrast_err.RData")
-
 data <- rbind(
@@
 random[is.na(random)] <- 0
+
+# x <- ranef(model_delboeuf_err, robust=TRUE)$Participant
+# random2 <- x[1:250, 
+#              "Estimate",
+#              c("Illusion_EffectIncongruent:sqrtabsIllusion_Strength",
+#                "Illusion_EffectIncongruent:sqrtIllusion_Difference:sqrtabsIllusion_Strength")] |> 
+#   as.data.frame()
+# cor(arrange(random2, row.names(random2)), arrange(random[2:3], random$Participant))
 ```
@@
 model <- "
-I2 =~ RodFrame_Strength + RodFrame_Interaction + VerticalHorizontal_Strength + VerticalHorizontal_Interaction + MullerLyer_Strength + MullerLyer_Interaction + Ponzo_Strength + Ponzo_Interaction
+i2 =~ RodFrame_Strength + RodFrame_Interaction + VerticalHorizontal_Strength + VerticalHorizontal_Interaction + MullerLyer_Strength + MullerLyer_Interaction + Ponzo_Strength + Ponzo_Interaction
 Zollner =~ Zollner_Strength + Zollner_Interaction
@@
 Delboeuf =~ Delboeuf_Strength + Delboeuf_Interaction
-I =~ I2 + Zollner + White + Ebbinghaus + Poggendorff + Contrast
+i =~ i2 + Zollner + White + Ebbinghaus + Poggendorff + Contrast
 "
@@
 lvl1 <- paste(data$Illusion_Type, "=~", data$Name)
-lvl2 <- paste("I =~", unique(data$Illusion_Type))
+lvl2 <- paste("i =~", unique(data$Illusion_Type))
 resid <- as.data.frame(t(utils::combn(x=unique(data$Illusion_Type), 2)))
@@
 # Model 1b - without Delboeuf
-m1b <- sem(model = paste0(c(lvl1, lvl2[lvl2 != "I =~ Delboeuf"]), collapse = "\n"),
+m1b <- sem(model = paste0(c(lvl1, lvl2[lvl2 != "i =~ Delboeuf"]), collapse = "\n"),
           data  = random,
@@
 Angle =~ Zollner_Strength + Zollner_Interaction + Poggendorff_Strength + Poggendorff_Interaction  + RodFrame_Strength + RodFrame_Interaction
-I =~ Lines + Circles + Contrast + Angle 
+i =~ Lines + Circles + Contrast + Angle 
 "
@@
 lvl1 <- paste(data$Parameter, "=~", data$Name)
-lvl2 <- paste("I =~", unique(data$Parameter))
+lvl2 <- paste("i =~", unique(data$Parameter))
 resid <- as.data.frame(t(utils::combn(x=unique(data$Parameter), 2)))
@@
 
-anova(m0, m1, m2)
+anova(m0, m1, m1b, m2)
 
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, fig.width=30, fig.height=15}
 # https://cjvanlissa.github.io/tidySEM/articles/sem_graph.html
@@
 
-colors <- c("I" = "#9C27B0", 
-            "Delboeuf"="#F44336", "Ebbinghaus"="#E91E63",
-            "Rod-Frame"="#4CAF50", "Rod\nFrame"="#4CAF50", 
-            "Vertical-Horizontal"="#FF9800", "Vertical\nHorizontal"="#FF9800",
-            "Zöllner"="#FF5722",
-            "White"="#616161", 
-             "Müller-Lyer" = "#3F51B5", "Müller\n-Lyer" = "#3F51B5",
-            "Ponzo" = "#2196F3",
+colors <- c("i" = "#9C27B0", 
+            "Delboeuf"="#2196F3", 
+            "Ebbinghaus"="#3F51B5",
+            "Rod-Frame"="#F44336", "Rod\nFrame"="#F44336", 
+            "Vertical-Horizontal"="#FF5722", "Vertical\nHorizontal"="#FF5722",
+            "Zöllner"="#FF9800",
+            "White"="#9E9E9E", 
+            "Müller-Lyer" = "#4CAF50", "Müller\n-Lyer" = "#4CAF50",
+            "Ponzo" = "#009688",
             "Poggendorff"="#795548",
@@
          width = abs(as.numeric(est)),
-         type = ifelse(from == "I", "curved", "straight")) |> 
+         type = ifelse(from == "i", "curved", "straight")) |> 
   filter(lhs != rhs)
@@
            grepl("_", name) ~ 1.15,
-           name == "I" ~ 0.5,
+           name == "i" ~ 0.5,
            TRUE ~ 0.5),
@@
            grepl("_", name) ~ 1,
-           name == "I" ~ 7,
+           name == "i" ~ 7,
            TRUE ~ 6),
@@
            grepl("_", name) ~ 1,
-           name == "I" ~ 3,
+           name == "i" ~ 3,
            TRUE ~ 1.1),
@@
            grepl("_", name) ~ "italic",
-           name == "I" ~ "bold",
+           name == "i" ~ "bold.italic",
            TRUE ~ "plain"),
@@
          label = str_replace(label, "Vertical-Horizontal", "Vertical\nHorizontal"),
-         family = ifelse(name == "I", "serif", "sans"))
-
-p <- tidygraph::tbl_graph(nodes = nodes, edges = edges) |> 
+         family = ifelse(name == "i", "serif", "sans"))
+
+p_sem <- tidygraph::tbl_graph(nodes = nodes, edges = edges) |> 
   ggraph(layout = "sugiyama") +
@@
                     label=est_std),
-                strength = 0.03 * c(-1, -1, 1, 1, 1, 1, -1, -1, -1, -1),
+                strength = 0.03 * c(-1, -1, -1, -1, -1, 1, 1, 1, 1, 1),
                 angle_calc="along",
-                label_dodge=unit(-0.015, "npc")) +
+                label_dodge=unit(-0.015, "npc"),
+                label_size = rel(8)) +
   geom_edge_link(aes(filter=type=="straight",
@@
                  angle_calc="along", 
-                 label_dodge=unit(-0.015, "npc")) +
+                 label_dodge=unit(-0.015, "npc"),
+                 label_size = rel(8)) +
   geom_node_point(aes(shape = shape, color=color, size=size)) +
@@
                  fontface=nodes$face,
-                 size = rel(nodes$textsize*3),
+                 size = rel(nodes$textsize*10),
                  family=nodes$family) +
-  scale_y_continuous(expand = expansion(c(.6, .1))) +
-  scale_edge_width_continuous(range=c(0.1, 1.5)) +
+  scale_y_continuous(expand = expansion(add=c(0.42, 0.32))) +
+  scale_edge_width_continuous(range=c(0.3, 6)) +
   scale_edge_color_manual(values=c("1"="#263238", "-1"="#B71C1C")) +
@@
   scale_shape_manual(values=c("oval"="circle", "rect"="square"), guide="none") +
-  scale_size_continuous(range = c(2, 27), guide="none") +
+  scale_size_continuous(range = c(10, 95), guide="none") +
   guides(edge_colour = "none", edge_width = "none") + 
@@
   theme(plot.title = element_text(hjust=0.5))
-p
-save(p, file="models/sem.Rdata")
+
+p_sem
+ggsave("figures/p_sem.png", p_sem, width=30, height=15)
+save(p_sem, file="models/p_sem.Rdata")
 
@@
 empirical <- df |>
+  # filter(Participant == "S01", Illusion_Type == "Poggendorff", Illusion_Difference <= 0.03) |> 
   filter(Illusion_Effect == "Incongruent") |>
-  group_by(Participant, Illusion_Type, Illusion_Strength) |>
-  arrange(Illusion_Difference) |> 
-  slice(0:2) |> 
+  arrange(Participant, Illusion_Type, Illusion_Difference) |>
+  group_by(Participant, Illusion_Type) |>
+  # select(Participant, Block, Illusion_Effect, Illusion_Type, Illusion_Strength, Illusion_Difference, Error) |>
+  # print(n=20)
+  slice(1:(8*2)) |> 
   # select(Illusion_Strength, Illusion_Difference)
-  group_by(Participant, Illusion_Type) |>
+  # group_by(Participant, Illusion_Type) |>
   summarize(
+    # Illusion_Strength= mean(Illusion_Strength),
+    # Illusion_Difference=mean(Illusion_Difference),
     TaskHard_Error = sum(Error, na.rm=TRUE) / n(),
@@
   filter(Illusion_Effect == "Incongruent") |> 
-  group_by(Participant, Illusion_Type, Illusion_Difference) |>
-  arrange(desc(Illusion_Strength)) |> 
-  slice(0:2) |> 
-  # select(Illusion_Strength, Illusion_Difference)
+  arrange(Participant, Illusion_Type, desc(Illusion_Strength)) |>
   group_by(Participant, Illusion_Type) |>
+  slice(1:(8*2)) |> 
+  # group_by(Participant, Illusion_Type) |>
   summarize(
+    # Illusion_Strength= mean(Illusion_Strength),
+    # Illusion_Difference=mean(Illusion_Difference),
     IlluStrong_Error = sum(Error, na.rm=TRUE) / n(),
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-r <- correlation(arrange(dfsem, Participant), 
-                 arrange(empirical, Participant), 
+r <- correlation(arrange(dfsem, as.character(Participant)), 
+                 arrange(empirical, as.character(Participant)), 
                  sort=TRUE, p_adjust="none")
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-# Correlation
-r <- data.frame()
-for(ill in c("Delboeuf", 
-             "Ebbinghaus",
-             "RodFrame",
-             "VerticalHorizontal",
-             "Zollner",
-             "White",
-             "MullerLyer",
-             "Ponzo",
-             "Poggendorff",
-             "Contrast"
-             )) {
-  r <- correlation::correlation(
-    data = dfsem |> 
-      arrange(Participant),
-    data2 = empirical |>
-      data_select(select=ill, regex=TRUE),
-    p_adjust="none"
-    ) |> 
-    mutate(Illusion_Type = ill,
-           Parameter1 = prettify_parameterName(Parameter1),
-           Parameter2 = prettify_parameterName(Parameter2)) |> 
-    select(Illusion_Type, Parameter1, Parameter2, r, p) |> 
-    rbind(r)
-}
-
-r |>
-  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
-  ggplot(aes(x = Parameter1, y = Parameter2)) +
-  geom_tile(aes(fill = r)) +
-  geom_text(aes(label=label)) +
-  scale_alpha_continuous(range = c(1, 0.5)) +
-  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
-  # facet_wrap(~Illusion_Type, scales = "free") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
-```
-
-### GARBAGE
-<!-- ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE} -->
-<!-- rez <- parameters::n_factors(select(empirical, -Participant), n_max=30) -->
-<!-- plot(rez) -->
-
-<!-- r <- correlation::correlation(empirical, p_adjust = "none", sort=TRUE) -->
-
-<!-- r |> -->
-<!--   ggplot(aes(x = Parameter2, y = Parameter1)) + -->
-<!--   geom_tile(aes(fill = r)) + -->
-<!--   # geom_text(aes(label = Text), size = 2) + -->
-<!--   scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") + -->
-<!--   scale_x_discrete(expand = c(0, 0)) + -->
-<!--   scale_y_discrete(expand = c(0, 0)) + -->
-<!--   labs(title = "Correlation Matrix", x = NULL, y = NULL) + -->
-<!--   theme_minimal() + -->
-<!--   theme( -->
-<!--     axis.text.x = element_text(angle = 90, hjust = 1), -->
-<!--     plot.title = element_text(hjust = 0.5, face = "bold"), -->
-<!--     panel.grid.major = element_blank(), -->
-<!--     panel.grid.minor = element_blank() -->
-<!--   ) -->
-
-<!-- fa <- parameters::factor_analysis(select(empirical, -Participant), -->
-<!--   n = 13, -->
-<!--   rotation = "oblimin", -->
-<!--   fm = "mle", -->
-<!--   sort = TRUE -->
-<!-- ) -->
-
-<!-- print(tibble::as_tibble(fa), n=30) -->
-<!-- plot(fa) -->
-<!-- ``` -->
-
-<!-- ### Approximation -->
-
-<!-- ```{r message=FALSE, warning=FALSE, cache=cache} -->
-<!-- individual <- data.frame() -->
-<!-- for (participant in unique(dfsub$Participant)) { -->
-<!--   cat(".") -->
-<!--   dat <- data.frame() -->
-
-<!--   dat <- glm(Error ~ sqrtmod(Illusion_Difference) * sqrtmod(Illusion_Strength), -->
-<!--     family = "binomial", -->
-<!--     data = filter( -->
-<!--       df,  -->
-<!--       Participant == participant, -->
-<!--       Illusion_Type == "Delboeuf" -->
-<!--     ) -->
-<!--   ) |>  -->
-<!--     parameters::parameters() |>  -->
-<!--     mutate(Parameter = paste0("Delboeuf_Individual_", Parameter)) |>  -->
-<!--     rbind(dat) -->
-
-<!--   dat <- glm(Error ~ sqrt(Illusion_Difference) * logmod(Illusion_Strength), -->
-<!--     family = "binomial", -->
-<!--     data = filter( -->
-<!--       df,  -->
-<!--       Participant == participant, -->
-<!--       Illusion_Type == "Ebbinghaus" -->
-<!--     ) -->
-<!--   ) |>  -->
-<!--     parameters::parameters() |>  -->
-<!--     mutate(Parameter = paste0("Ebbinghaus_Individual_", Parameter)) |>  -->
-<!--     rbind(dat) -->
-
-<!--   individual <- dat |> -->
-<!--     select(Parameter, Coefficient) |>  -->
-<!--     as.data.frame() |> -->
-<!--     mutate( -->
-<!--       Participant = participant, -->
-<!--       Parameter = clean_parameterName(Parameter) -->
-<!--     ) |>  -->
-<!--     rbind(individual) -->
-<!-- } -->
-
-<!-- individual <- individual |>  -->
-<!--   pivot_wider(names_from = c("Parameter"), values_from = c("Coefficient")) |>  -->
-<!--   arrange(Participant) -->
-<!-- ``` -->
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-# Correlation
-r <- data.frame()
-for(ill in c("Delboeuf", 
-             "Ebbinghaus")) {
-  r <- correlation::correlation(
-    # data = random |> 
-    #   arrange(Participant) |> 
-    #   data_select(select=ill, regex=TRUE),
-    data = empirical |> 
-      data_select(select=ill, regex=TRUE),
-    data2 = individual |>
-      data_select(select=ill, regex=TRUE),
-    p_adjust="none"
-    ) |> 
-    mutate(Illusion_Type = ill,
-           Parameter1 = prettify_parameterName(Parameter1),
-           Parameter2 = prettify_parameterName(Parameter2)) |> 
-    select(Illusion_Type, Parameter1, Parameter2, r, p) |> 
-    rbind(r)
-}
-
-r |>
-  mutate(label = insight::format_p(p, stars_only=TRUE)) |> 
-  ggplot(aes(x = Parameter2, y = Parameter1)) +
-  geom_tile(aes(fill = r)) +
-  geom_text(aes(label=label)) +
-  scale_alpha_continuous(range = c(1, 0.5)) +
-  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  facet_wrap(~Illusion_Type, scales = "free")
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-library(ggraph)
-
-dat <- rbind(
-  attributes(fa)$loadings_long,
-  attributes(fa2)$loadings_long |>
-    mutate(Component = str_replace(Component, "ML", "G"))
-) |>
-  tidygraph::as_tbl_graph() |>
-  tidygraph::activate("nodes")
-
-# Get order of indices
-# idx <- correlation::cor_sort(cor, hclust_method = "ward.D2") |>
-#   summary() |>
-#   pull(Parameter)
-# idx <- sort(fa$Variable)
-idx <- fa$Variable
-
-get_x <- function(name, x) {
-  case_when(
-    name == "G1" ~ 0.5,
-    # name == "G1" ~ 0/3,
-    # name == "G2" ~ 1/3,
-    # name == "G3" ~ 2/3,
-    # name == "G4" ~ 3/3,
-    name == "ML1" ~ 0/(ncol(fa)-3),
-    name == "ML2" ~ 1/(ncol(fa)-3),
-    name == "ML3" ~ 2/(ncol(fa)-3),
-    name == "ML4" ~ 3/(ncol(fa)-3),
-    name == "ML5" ~ 4/(ncol(fa)-3),
-    name == "ML6" ~ 5/(ncol(fa)-3),
-    name == "ML7" ~ 6/(ncol(fa)-3),
-    name == "ML8" ~ 7/(ncol(fa)-3),
-    name == "ML9" ~ 8/(ncol(fa)-3),
-    name == "ML10" ~ 9/(ncol(fa)-3),
-    name == "ML11" ~ 10/(ncol(fa)-3),
-    name == "ML12" ~ 11/(ncol(fa)-3),
-    TRUE ~ x)
-}
-
-dat |>
-  mutate(
-    x = sapply(as.data.frame(dat)$name, function(x) {
-      if (x %in% idx) {
-        (which(idx == x) - 1) / (length(idx) - 1)
-      } else {
-        NA
-      }
-    })
-  ) |>
-  mutate(
-    x = get_x(name, x),
-    y = case_when(
-    str_detect(name, "G") ~ 2,
-    str_detect(name, "ML") ~ 1,
-    TRUE ~ 0),
-    angle = ifelse(name %in% idx, 90, 0),
-    hjust = ifelse(name %in% idx, 1.01, 0),
-    name = prettify_parameterName(name),
-    Illusion_Type = str_split_fixed(name, " - ", n=2)[, 1]) |> 
-  tidygraph::activate("edges") |> 
-  filter(abs(Loading) > 0.2) |> 
-  ggraph(layout = "nicely") +
-  geom_edge_link(aes(edge_width = Loading, edge_colour = as.factor(sign(Loading)))) +
-  geom_node_point(aes(color = Illusion_Type), size=3) +
-  geom_node_text(aes(label = name, angle = angle, hjust=hjust)) +
-  scale_y_continuous(expand = expansion(c(.6, .1))) +
-  scale_edge_width_continuous(range=c(0.1, 1)) +
-  scale_edge_color_manual(values=c("#F44336", "#2196F3")) +
-  scale_color_manual(values=c("Delboeuf" = "#F44336", Ebbinghaus = "#E91E63",
-                              "Rod-Frame" = "#9C27B0")) +
-  guides(edge_colour = "none") +
-  ggraph::theme_graph()
-```
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+### Figure
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 # Save
@@
   write.csv("../data/study3.csv", row.names = FALSE)
-```
+
+# Figure 1a
+# -----------
+for(i in c("delboeuf", "ebbinghaus", "rodframe", "verticalhorizontal", "zollner", "white", "mullerlyer", "ponzo", "poggendorff", "contrast")) {
+  load(paste0("models/model_", i, "_err.RData"))
+  load(paste0("models/gam_", i, "_err.RData"))
+  load(paste0("models/model_", i, "_rt.RData"))
+  load(paste0("models/gam_", i, "_rt.RData"))
+}
+load("models/p_sem.RData")
+
+p_delboeuf <- plot_all(
+  filter(df, Illusion_Type == "Delboeuf"), 
+  plot_model_err(filter(df, Illusion_Type == "Delboeuf"), model_delboeuf_err, gam_delboeuf_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Delboeuf", Error == 0), model_delboeuf_rt, gam_delboeuf_rt))
+p_ebbinghaus <- plot_all(
+  filter(df, Illusion_Type == "Ebbinghaus"), 
+  plot_model_err(filter(df, Illusion_Type == "Ebbinghaus"), model_ebbinghaus_err, gam_ebbinghaus_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Ebbinghaus", Error == 0), model_ebbinghaus_rt, gam_ebbinghaus_rt))
+p_rodframe <- plot_all(
+  filter(df, Illusion_Type == "Rod-Frame"), 
+  plot_model_err(filter(df, Illusion_Type == "Rod-Frame"), model_rodframe_err, gam_rodframe_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Rod-Frame", Error == 0), model_rodframe_rt, gam_rodframe_rt))
+p_verticalhorizontal <- plot_all(
+  filter(df, Illusion_Type == "Vertical-Horizontal"), 
+  plot_model_err(filter(df, Illusion_Type == "Vertical-Horizontal"), model_verticalhorizontal_err, gam_verticalhorizontal_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0), model_verticalhorizontal_rt, gam_verticalhorizontal_rt))
+p_zollner <- plot_all(
+  filter(df, Illusion_Type == "Zöllner"), 
+  plot_model_err(filter(df, Illusion_Type == "Zöllner"), model_zollner_err, gam_zollner_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Zöllner", Error == 0), model_zollner_rt, gam_zollner_rt))
+p_white <- plot_all(
+  filter(df, Illusion_Type == "White"), 
+  plot_model_err(filter(df, Illusion_Type == "White"), model_white_err, gam_white_err), 
+  plot_model_rt(filter(df, Illusion_Type == "White", Error == 0), model_white_rt, gam_white_rt) + coord_cartesian(ylim=c(0.125, 1.7), xlim = c(-17.5, 17.5)))
+p_mullerlyer <- plot_all(
+  filter(df, Illusion_Type == "Müller-Lyer"), 
+  plot_model_err(filter(df, Illusion_Type == "Müller-Lyer"), model_mullerlyer_err, gam_mullerlyer_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Müller-Lyer", Error == 0), model_mullerlyer_rt, gam_mullerlyer_rt))
+p_ponzo <- plot_all(
+  filter(df, Illusion_Type == "Ponzo"), 
+  plot_model_err(filter(df, Illusion_Type == "Ponzo"), model_ponzo_err, gam_ponzo_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Ponzo", Error == 0), model_ponzo_rt, gam_ponzo_rt))
+p_poggendorff <- plot_all(
+  filter(df, Illusion_Type == "Poggendorff"), 
+  plot_model_err(filter(df, Illusion_Type == "Poggendorff"), model_poggendorff_err, gam_poggendorff_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Poggendorff", Error == 0), model_poggendorff_rt, gam_poggendorff_rt))
+p_contrast <- plot_all(
+  filter(df, Illusion_Type == "Contrast"), 
+  plot_model_err(filter(df, Illusion_Type == "Contrast"), model_contrast_err, gam_contrast_err), 
+  plot_model_rt(filter(df, Illusion_Type == "Contrast", Error == 0), model_contrast_rt, gam_contrast_rt))
+
+p_ill <- 
+  (p_delboeuf + p_ebbinghaus + 
+     p_rodframe + p_verticalhorizontal +
+     p_zollner + p_white +
+     p_mullerlyer + p_ponzo +
+     p_poggendorff + p_contrast) + 
+    patchwork::plot_layout(ncol= 5, nrow=2)
+
+ggsave("figures/figure_ill.png", p_ill, height = 20, width = 30, dpi=300)
+
+ggsave("../figures/figure2.png", ((p_ill) / p_sem) + plot_layout(heights=c(2, 1.5)), height = 35, width = 30, dpi=300)
+```

### 2022-09-05 study3/study3.Rmd
update

--- 2022-08-27 study3/study3.Rmd
@@
   collapse = TRUE,
-  dpi = 450,
+  dpi = 300,
   fig.path = "./figures/",
@@
     Screen_Refresh = as.numeric(Screen_Refresh),
+    Ethnicity = ifelse(Ethnicity == "Latino", "Hispanic", Ethnicity),
     Education = fct_relevel(Education, "High School", "Bachelor", "Master", "Doctorate", "Other", "Prefer not to Say")
   ) |> 
-  datawizard::standardise(select=names(DVs))
+  datawizard::standardise(select=names(DVs)) |> 
+  datawizard::change_scale(select=starts_with("IPIP"), to = c(0, 1), range = c(0, 100)) |> 
+  datawizard::change_scale(select=starts_with("PID"), to = c(0, 1), range = c(0, 100))
+
+# Remove outliers (p < 0.0001)
+dfsub[names(DVs)][abs(dfsub[names(DVs)]) > qnorm(0.9999)] <- NA
 
@@
   pivot_longer(all_of(names(DVs)), names_to = "Index", values_to = "Score")
-# plot(estimate_density(dfsub[DVs]))
 ```
 
-## Results
+## Participants
 
-### Contextual
+`r sum(is.na(dfsub$IPIP6_Agreeableness))` participants did not do the personality scales.
 
-#### Screen Size
+```{r demographics}
+p_age <- estimate_density(dfsub$Age) |> 
+  ggplot(aes(x = x, y = y)) +
+  geom_area(fill = "#607D8B") +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "Age", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(size=rel(1), face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(face = "italic", hjust = 0.5),
+    # axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size=rel(0.6)),
+    axis.title.x = element_blank()
+  )
+
+p_nat <- dfsub |> 
+  group_by(Nationality) |> 
+  summarize(n = n()) |> 
+  mutate(Nationality = fct_reorder(Nationality, desc(n))) |> 
+  ggplot(aes(x = Nationality, y = n, fill=Nationality)) +
+  geom_bar(stat = "identity") +
+  labs(y = "Number", title = "Nationality") +
+  scale_fill_material_d(guide = "none") +
+  scale_x_discrete(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(size=rel(1), face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(face = "italic", hjust = 0.5),
+    # axis.line.y = element_blank(),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_text(size=rel(0.6)),
+    axis.text.x = element_text(size=rel(0.5), angle = 45, hjust=1)
+  )
+
+ggsave("figures/figure_dem1.png", 
+       patchwork::wrap_elements(p_age / p_nat), 
+       width=0.5*(fig.height), 
+       height=1*(fig.height), dpi=1000, bg="white")
+
+p_sex <- plot_waffle(dfsub, what="Sex", rows = 10, size=5) +
+  scale_color_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63"))
+
+p_edu <- plot_waffle(dfsub, "Education", rows = 10, size=5) +
+  scale_color_viridis_d()
+
+p_race <- plot_waffle(dfsub, "Ethnicity", rows = 10, size=5) +
+  scale_color_manual(values = c("Hispanic" = "#FF5722", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Other" = "#795548"))
+
+ggsave("figures/figure_dem2.png", 
+       patchwork::wrap_elements(p_sex / p_edu / p_race), width=1*(fig.height), height=1*(fig.height), dpi=1000, bg="white")
+
+p_ipip <- dfsub |> 
+  select(starts_with("IPIP") & !ends_with("_SD")) |> 
+  estimate_density() |> 
+  mutate(Parameter = str_remove_all(Parameter, "IPIP6_"),
+         Parameter = str_replace(Parameter, "HonestyHumility", "Honesty-Humility")) |> 
+  ggplot(aes(x = x, y = y, color = Parameter)) +
+  geom_line(size = 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  scale_color_manual(values=c("Agreeableness" = "#FFC107", "Honesty-Humility" = "#00BCD4", "Extraversion" = "#9C27B0", "Conscientiousness" = "#3F51B5", "Openness" = "#4CAF50", "Neuroticism" = "#E91E63")) +
+  labs(x = "Score", title = "Normal Personality", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(size=rel(1), face = "bold", hjust = 0.5),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size=rel(0.6)),
+    axis.title.x = element_blank(),
+    legend.text = element_text(size=rel(0.8))
+    # legend.position = "top",
+  )
+
+p_pid <- dfsub |> 
+  select(starts_with("PID") & !ends_with("_SD")) |> 
+  estimate_density() |> 
+  mutate(Parameter = str_remove_all(Parameter, "PID5_"),
+         Parameter = str_replace(Parameter, "NegativeAffect", "Negative Affect")) |> 
+  ggplot(aes(x = x, y = y, color = Parameter)) +
+  geom_line(size = 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  scale_color_manual(values=c("Antagonism" = "#FF9800", "Detachment" = "#03A9F4", "Disinhibition" = "#FF5722", "Negative Affect" = "#F44336", "Psychoticism" = "#673AB7")) +
+  labs(x = "Score", title = "Pathological Personality", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(size=rel(1), face = "bold", hjust = 0.5),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size=rel(0.6)),
+    axis.title.x = element_blank(),
+    legend.text = element_text(size=rel(0.8))
+    # legend.position = "top",
+  )
+
+ggsave("figures/figure_dem3.png", 
+       patchwork::wrap_elements(p_ipip / p_pid), width=2/3*(fig.height), height=1*(fig.height), dpi=1000, bg="white")
+
+p_dem <- (
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem1.png"), interpolate = TRUE)) | 
+    patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem2.png"), interpolate = TRUE)) | 
+    patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem3.png"), interpolate = TRUE))
+) + 
+  patchwork::plot_layout(widths=c(1/2, 1, 2/3)) + 
+  patchwork::plot_annotation(title = "Participants (n = 250)", 
+                             theme = theme(plot.title = element_text(size = rel(1.5), face="bold", hjust=0.5, vjust=-5)))
+
+ggsave("figures/figure_dem.png", p_dem, dpi=1000)
+
+p_dem
+```
+
+# Results
+
+## Contextual
+
+### Screen Size
 
@@
 for(i in names(DVs)) {
-  model <- brms::brm(paste0(i, " ~ s(Screen_Size)"), data=dfsub, algorithm = "meanfield", refresh=0)
-  param <- bayestestR::describe_posterior(as.data.frame(model))
+  # model <- brms::brm(paste0(i, " ~ s(Screen_Size)"), data=dfsub, algorithm = "meanfield", refresh=0)
+  # param <- bayestestR::describe_posterior(as.data.frame(model))
+  model <- lm(as.formula(paste0(i, " ~ Screen_Size")), 
+              data=dfsub)
+  param <- parameters::parameters(model)
 
@@
   pred$Index <- i
-  pred$sig <- estimate_slopes(model, trend = "Screen_Size")$pd
+  pred$sig <- param$p[2]
   preds <- rbind(preds, pred)
 
-  if(any(pred$sig > .95)) {
-    print(i)
-  }
 }
@@
 preds |> 
-  mutate(sig = format_pd(sig, stars_only = TRUE),
+  mutate(sig = format_p(sig, stars_only = TRUE),
          sig = ifelse(sig == "", "NS", sig)) |> 
@@
   scale_x_continuous(expand = c(0, 0)) +
-  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
-  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid")) +
+  scale_size_manual(values = c("NS" = 0.5, "*" = 1)) +
+  scale_linetype_manual(values = c("NS" = "dotted", "*" = "solid")) +
   scale_color_manual(values = DVs) +
@@
   theme_modern() +
-  labs(y = "Illusion Sensitivity", x = "Screen Size") +
+  labs(y = "Illusion Sensitivity", x = "Screen Size", linetype="Significance", size="Significance") +
   ggside::geom_xsidedensity(data=dfsub, fill ="grey", color = "white") +
   ggside::geom_ysidedensity(data=dflong, aes(y=Score, color = Index)) +
-  ggside::theme_ggside_void() 
+  ggside::theme_ggside_void() +
+  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
+  ggside::scale_xsidey_continuous(expand = c(0, 0))
 ```
 
-#### Refresh Rate
+### Refresh Rate
 
 ```{r message=FALSE, warning=FALSE, cache=cache}
-sig <- list()
 preds <- data.frame()
-for(i in DVs) {
-  model <- brms::brm(paste0(i, " ~ poly(Screen_Refresh, 2)"), data=dfsub, algorithm = "meanfield", refresh=0)
-  param <- bayestestR::describe_posterior(as.data.frame(model))
+for(i in names(DVs)) {
+  # model <- brms::brm(paste0(i, " ~ s(Screen_Refresh)"), data=dfsub, algorithm = "meanfield", refresh=0)
+  # param <- bayestestR::describe_posterior(as.data.frame(model))
+  model <- lm(as.formula(paste0(i, " ~ Screen_Refresh")), 
+              data=dfsub)
+  param <- parameters::parameters(model)
 
@@
   pred$Index <- i
-  pred$sig <- max(param$pd[2:3])
+  pred$sig <- param$p[2]
   preds <- rbind(preds, pred)
 
-  if(any(max(param$pd[2:3]) > .95)) {
-    print(i)
-  }
 }
@@
 preds |> 
-  mutate(sig = format_pd(sig, stars_only = TRUE),
+  mutate(sig = format_p(sig, stars_only = TRUE),
          sig = ifelse(sig == "", "NS", sig)) |> 
   ggplot(aes(x = Screen_Refresh, y = Predicted)) +
+  geom_point2(data=dflong, aes(y = Score, color= Index), alpha=0.1, size=2) +
   geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
-  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
-  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_size_manual(values = c("NS" = 0.5, "*" = 1)) +
+  scale_linetype_manual(values = c("NS" = "dotted", "*" = "solid")) +
+  scale_color_manual(values = DVs) +
+  coord_cartesian(ylim = c(-2.5, 2.5)) +
+  theme_modern() +
+  labs(y = "Illusion Sensitivity", x = "Screen Refresh Rate", linetype="Significance", size="Significance") +
+  ggside::geom_xsidedensity(data=dfsub, fill ="grey", color = "white") +
+  ggside::geom_ysidedensity(data=dflong, aes(y=Score, color = Index)) +
+  ggside::theme_ggside_void() +
+  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
+  ggside::scale_xsidey_continuous(expand = c(0, 0))
 ```
 
-### Demographics
+## Demographics
 
-#### Sex
+### Sex
 
@@
 params <- data.frame()
-for(i in DVs) {
+for(i in names(DVs)) {
   model <- lm(paste0(i, " ~ Sex"), data=dfsub)
@@
   param$ymiddle <- param$Coefficient[1] + diff(param$Coefficient) / 2
+  param$BF <- parameters::parameters(BayesFactor::ttestBF(formula=as.formula(paste0(i, " ~ Sex")), data=dfsub[!is.na(dfsub[[i]]), ]))$BF
   params <- rbind(params, as.data.frame(param[2, ]))
-  if(param$p[2] < .05) {
-    print(i)
-  }
 }
@@
 data <- dfsub |> 
-  pivot_longer(all_of(DVs), names_to = "Index", values_to = "Score") |> 
-  mutate(Index = fct_relevel(Index, arrange(params, p)$Index))
+  pivot_longer(all_of(names(DVs)), names_to = "Index", values_to = "Score") |> 
+  mutate(Index = fct_relevel(Index, names(DVs)),
+         Index = fct_relabel(Index, ~prettify_itemName(.x)))  # arrange(params, p)$Index
 
-data |> 
+p_gender <- data |> 
   ggplot(aes(x = Index, y = Score)) +
-  stat_slab(data=filter(data, Sex == "Male"), aes(fill = Sex), side = "right", scale = 0.5, position = "dodge") +
-  stat_slab(data=filter(data, Sex == "Female"), aes(fill = Sex), side = "left", scale = 0.5, position = "dodge") +
-  stat_pointinterval(aes(group = Sex), point_interval = "mean_qi",, position = "dodge") +
-  geom_text(data = mutate(params, sig = insight::format_p(p, stars_only = TRUE)), aes(label = sig, y = ymiddle))
+  stat_slab(data=filter(data, Sex == "Male"), aes(fill = Sex), side = "left", scale = 0.5, position = "dodge") +
+  stat_slab(data=filter(data, Sex == "Female"), aes(fill = Sex), side = "right", scale = 0.5, position = "dodge") +
+  stat_pointinterval(aes(group = fct_rev(Sex)), point_interval = "mean_qi", position = "dodge") +
+  geom_text(data = mutate(params, sig = insight::format_p(p, stars_only = TRUE), Index = prettify_itemName(Index)), aes(label = sig, y = ymiddle)) +
+    geom_label(data = mutate(params, label = insight::format_bf(BF), Index = prettify_itemName(Index)), 
+             aes(label = label), y=2.7) +
+  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63")) +
+  theme_minimal() +
+  labs(title = "Illusion Sensitivity and Demographics", fill = "") +
+  theme(legend.position = "top",
+        axis.title = element_blank(),
+        plot.title = element_text(face = "bold", hjust=0.5))
+
+p_gender
+ggsave("figures/figure_sex.png", p_gender, width=2*(fig.height), height=0.8*(fig.height), dpi=600, bg="white")
 ```
 
-#### Ethnicity
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-sig <- list()
-params <- data.frame()
-for(i in DVs) {
-  model <- lm(paste0(i, " ~ Ethnicity"), data=dfsub)
-  param <- estimate_contrasts(model, contrast = "Ethnicity")
-  param$Index <- i
-  params <- rbind(params, as.data.frame(param[2, ]))
-  if(any(param$p < .05)) {
-    print(i)
-  }
-}
-```
-
-#### Education
+### Education
 
@@
 preds <- data.frame()
-for(i in DVs) {
-  model <- brms::brm(paste0(i, " ~ mo(as.numeric(Education))"), data=filter(dfsub, !Education %in% c("Other", "Prefer not to Say")), algorithm = "meanfield", refresh=0)
+for(i in names(DVs)) {
+  model <- brms::brm(paste0(i, " ~ mo(as.numeric(Education)) + Sex"), data=filter(dfsub, !Education %in% c("Other", "Prefer not to Say")), algorithm = "meanfield", refresh=0)
   param <- bayestestR::describe_posterior(as.data.frame(model)["bsp_moas.numericEducation"])
@@
   preds <- rbind(preds, pred)
-
-  if(any(param$pd > .95)) {
-    print(i)
-  }
 }
@@
 preds |> 
+  filter(Sex == "Male") |> 
   mutate(sig = format_pd(sig, stars_only = TRUE),
@@
   geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
-  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
-  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
+  scale_size_manual(values = c("NS" = 0.5, "*" = 1)) +
+  scale_linetype_manual(values = c("NS" = "dotted", "*" = "solid")) + 
+  theme_modern()
 ```
 
-#### Age
+### Age
 
 ```{r message=FALSE, warning=FALSE, cache=cache}
-sig <- list()
 preds <- data.frame()
-for(i in DVs) {
-  model <- brms::brm(paste0(i, " ~ poly(Age, 2)"), data=dfsub, algorithm = "meanfield", refresh=0)
-  param <- bayestestR::describe_posterior(as.data.frame(model))
+for(i in names(DVs)) {
+  # model <- brms::brm(paste0(i, " ~ poly(Age, 2)"), data=dfsub, algorithm = "meanfield", refresh=0)
+  # param <- bayestestR::describe_posterior(as.data.frame(model))
+  model <- lm(paste0(i, " ~ poly(Age, 2) + Sex + Education"), data=dfsub)
+  param <- parameters::parameters(model)
 
-  pred <- estimate_relation(model, length=50)
+  pred <- get_datagrid(model, at = "Age", length=50)
+  pred$Predicted <- get_predicted(model, data=pred)
+
   pred$Index <- i
@@
 
-  if(any(max(param$pd[2:3]) > .95)) {
-    print(i)
-  }
 }
@@
   geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
-  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
-  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
+  scale_size_manual(values = c("NS" = 0.5, "*" = 1)) +
+  scale_linetype_manual(values = c("NS" = "dotted", "*" = "solid")) +
+  theme_modern()
 ```
 
-### Personality
+## Personality
+
+### IPIP6
 
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-r <- correlation::correlation(
-  datawizard::data_select(dfsub, DVs),
+ripip <- correlation::correlation(
+  datawizard::data_select(dfsub, names(DVs)),
   select(dfsub, starts_with("IPIP") & !ends_with("_SD")),
   p_adjust = "none",
+  bayesian=TRUE,
   sort=TRUE
@@
 
-r |>
-  mutate(label = paste0(format_value(r, zap_small = TRUE), insight::format_p(p, stars_only=TRUE))) |> 
+ripip |>
+  mutate(label = paste0(format_value(rho, zap_small = TRUE), insight::format_bf(BF, stars_only=TRUE))) |> 
   ggplot(aes(x = Parameter2, y = Parameter1)) +
-  geom_tile(aes(fill = r)) +
+  geom_tile(aes(fill = rho)) +
+  geom_text(aes(label=label), size=3) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+
+p_ipip1 <- plot_correlation(dfsub, x="I", y="IPIP6_Agreeableness", fill="#FFC107") + 
+  labs(x = "") +
+  theme(axis.title.y = element_text(face = "bold"))
+p_ipip2 <- plot_correlation(dfsub, x="I", y="IPIP6_HonestyHumility", fill="#00BCD4") + 
+  labs(x = "") +
+  theme(axis.title.y = element_text(face = "bold"))
+
+plot_correlation(dfsub, x="Zollner", y="IPIP6_HonestyHumility", fill="#00BCD4")
+plot_correlation(dfsub, x="VerticalHorizontal", y="IPIP6_Agreeableness", fill="#FFC107")
+plot_correlation(dfsub, x="VerticalHorizontal", y="IPIP6_HonestyHumility", fill="#00BCD4")
+```
+
+### PID-5
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+rpid <- correlation::correlation(
+  datawizard::data_select(dfsub, names(DVs)),
+  select(dfsub, starts_with("PID") & !ends_with("_SD")),
+  p_adjust = "none",
+  bayesian=TRUE,
+  sort=TRUE
+)
+
+rpid |>
+  mutate(label = paste0(format_value(rho, zap_small = TRUE), insight::format_bf(BF, stars_only=TRUE))) |> 
+  ggplot(aes(x = Parameter2, y = Parameter1)) +
+  geom_tile(aes(fill = rho)) +
   geom_text(aes(label=label), size=3) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust=1))
+
+p_pid1 <- plot_correlation(dfsub, x="I", y="PID5_Antagonism", fill="#FF9800") + 
+  labs(x = "") +
+  theme(axis.title.y = element_text(face = "bold"))
+p_pid2 <- plot_correlation(dfsub, x="I", y="PID5_Disinhibition", fill="#FF5722") +
+  theme(axis.title.y = element_text(face = "bold"))
+p_pid3 <- plot_correlation(dfsub, x="I", y="PID5_Psychoticism", fill="#673AB7") + 
+  labs(x = "") +
+  theme(axis.title.y = element_text(face = "bold"))
+p_pid4 <- plot_correlation(dfsub, x="I", y="PID5_NegativeAffect", fill="#F44336") + 
+  labs(x = "") +
+  theme(axis.title.y = element_text(face = "bold"))
+
+plot_correlation(dfsub, x="VerticalHorizontal", y="PID5_Antagonism", fill="#FF9800")
+plot_correlation(dfsub, x="Ponzo", y="PID5_Antagonism", fill="#FF9800")
 ```
+
+## Table and Figure
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+table1 <- ripip |> 
+  rbind(rpid) |> 
+  data_remove(c("ROPE_Percentage", "Prior_Distribution", "Prior_Location", "Prior_Scale", "pd")) |>  
+  separate(Parameter2, "_", into = c("Scale", "Dimension")) |> 
+  arrange(Scale, desc(BF)) |> 
+  filter(BF > 3)
+
+table1 |> 
+  print_md()
+
+table1 |> 
+  format_table() |> 
+  select(-Method, -n_Obs) |> 
+  rename(Factor=Parameter1) |> 
+  write.csv("table1.csv", row.names = FALSE)
+```
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p_pers <- wrap_elements((p_ipip1 / p_ipip2) | (p_pid1 / p_pid2) | (p_pid3 / p_pid4)) +
+  patchwork::plot_annotation(title = "Illusion Sensitivity and Personality", 
+                             theme = theme(plot.title = element_text(size = rel(1.5), face="bold", hjust=0.5, vjust=0)))
+
+ggsave("figures/figure_pers.png", p_pers, dpi=600, width=2*(fig.height), height=1.3*(fig.height), bg="white")
+
+p <- (patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem.png"), interpolate = TRUE)) /
+        patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_sex.png"), interpolate = TRUE)) / 
+        patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_pers.png"), interpolate = TRUE)))
+
+p <- gridExtra::grid.arrange(
+  grid::rasterGrob(png::readPNG("figures/figure_dem.png"), interpolate = TRUE, width=10),
+  grid::rasterGrob(png::readPNG("figures/figure_sex.png"), interpolate = TRUE, width=10),
+  grid::rasterGrob(png::readPNG("figures/figure_pers.png"), interpolate = TRUE, width=10), ncol=1)
+p <- cowplot::plot_grid(
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem.png"), interpolate = TRUE)),
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_sex.png"), interpolate = TRUE)),
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_pers.png"), interpolate = TRUE)), 
+  ncol=1)
+p <- patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem.png"), interpolate = TRUE)) /
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_sex.png"), interpolate = TRUE)) /
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_pers.png"), interpolate = TRUE)) +
+  plot_layout(widths = c(1, 1, 1))
+
+ggsave("../figures/figure3.png", p, dpi=600, height=1.1*fig.width, width=fig.height)
+```

### 2022-09-05 study2/functions.R

@@ -20,7 +20,7 @@ plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", 
 }
 
 plot_waffle <- function(dfsub, what = "Nationality", title = what, rows = 8, size=3) {
-  library(emojifont)
+  # library(emojifont)
   ggwaffle::waffle_iron(dfsub, what, rows = rows) |>
     # mutate(label = emojifont::fontawesome('fa-smiley')) |>
     # mutate(label = emojifont::emoji('smiley')) |>
@@ -37,7 +37,10 @@ plot_waffle <- function(dfsub, what = "Nationality", title = what, rows = 8, siz
     # scale_y_continuous(expand = c(0, 0)) +
     theme_void() +
     # ggwaffle::theme_waffle() +
-    theme(plot.title = element_text(face = "bold", hjust = 0.5))
+    theme(plot.title = element_text(face = "bold", hjust = 0.5),
+          # legend.margin=margin(l=-5),
+          legend.key.height = unit(1, "mm"),
+          legend.key.width = unit(1, "mm"))
 }
 
 
@@ -490,7 +493,7 @@ plot_model_rt <- function(data, model, gam) {
     scale_fill_gradientn(colours = c("#F44336", "#FFC107", "#4CAF50")) +
     coord_cartesian(
       xlim = c(min(data[[varstrength]]), max(data[[varstrength]])),
-      ylim = c(125, 1500) / 1000
+      ylim = c(125, 1400) / 1000
     ) +
     theme_modern() +
     guides(color = "none") +
@@ -510,11 +513,12 @@ plot_model_rt <- function(data, model, gam) {
 
 
 plot_all <- function(data, p_err, p_rt) {
+  illname <- unique(data$Illusion_Type)
   # Get stimuli
   dat <- df |>
     filter(Error == 0) |>
     filter(
-      Illusion_Type == unique(data$Illusion_Type),
+      Illusion_Type == illname,
       Answer %in% c("arrowleft", "arrowup")
     ) |>
     select(Stimulus, Illusion_Strength, Illusion_Difference)
@@ -566,7 +570,10 @@ plot_all <- function(data, p_err, p_rt) {
     grid::rasterGrob(interpolate = TRUE) |>
     patchwork::wrap_elements()
 
-  p <- ((p_err + labs(x = "")) / (p_rt + labs(title = "")) + patchwork::plot_layout(guides = "collect"))
[…]

### 2022-09-05 manuscript.Rmd
update

@@
 ---
-title             : "**The Illusion Game: A Novel Experimental Paradigm to Study Visual Illusions**"
+title             : "**The Illusion Game: A Novel Experimental Paradigm provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Pathological Personality Correlates**"
 shorttitle        : "Illusion Game"
@@
 wordcount         : "1156"
-bibliography      : "bibliography.bib"
+bibliography      : "references.bib"
 floatsintext      : yes
@@
   chunk_output_type: console
-bibliography: references.bib
 ---
@@
 Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems. 
-They elegantly demonstrate that our brains do not simply map veridical accounts of physical reality; we generate perceptions of the environment by taking the contextual information into consideration, and the prior knowledge that manifests in our subjective conscious experience [@carbon2014].
-Despite the withstanding interest in illusions within various fields of study, including that of visual perception [@day1972; @eagleman2001; @gomez-villa2022 ], consciousness science [@caporuscio2022, @lamme2020] and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015re], several gaps still remain in the present research.
-Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions [@mylniec2016].
-For instance, while some researchers have tried to explain our sensitivity to illusory effects as a result of deficits in the low-level visual processing system [@cretenoud2019; @gori2016], others have provided a compelling case using a top-down approach, suggesting that such visual phenomena occur as a result of a conflict between our visual input and our prior beliefs[@teufel2018; @caporuscio2022].
-Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
-Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder [ASD]) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
+They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but to integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
+Despite the historical and intensive interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022, @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
+
+Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
+The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014] or **Insert some demographic / personality feature** [**REF**] - is another area of controversy.
 
@@
 One key challenge hindering the further development of illusion research is the relative difficulty in adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-To address this issue, we first developed a parametric framework to manipulate visual illusions, of which we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
-This software allows us to generate different types of classic visual illusions and more importantly, dissociate the parameters of an illusion from its rendered output.
-Specifically, it enables us to modulate continuously and independently two parameters, *illusion strength* and *task difficulty*, that could then be manipulated systematically to conduct objective assessments on visual illusions (see **Figure 1**).
+To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021]. 
+This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus) with a continuous and independent modulation of two parameters, *illusion strength* and *task difficulty* (see **Figure 1**).
 
-```{r message=FALSE, warning=FALSE, fig.cap="Explanation of the parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
-Examples of stimuli showcasing the manipulation of the two main parameters, the task difficulty and the illusion strength (below)."}
+```{r message=FALSE, warning=FALSE, fig.cap="The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength."}
 knitr::include_graphics("figures/Figure1.png")
@@
 Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-Past studies have implemented paradigms emphasizing the subjective perception of the participant, to examine the extent to which individual context biases perceptual variance (for instance, having participants adjust the parametres of the target stimulus to a reference stimulus relying only on their perception, eg., @mylniec2016a; @grzeczkowski2018).
-Alternatively, Pyllusion allows us to create illusions while manipulating specific features of interest (e.g., size, length etc.), and thereby the difficulty in distinguishing them, so that the targets are objectively distinct (i.e., one line segment is physically longer than the other), and the strength of the generated illusion are variable (i.e., the biasing angle of the arrows are more or less acute).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (*REF*), or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016a; @grzeczkowski2018].
+Alternatively, *Pyllusions* allows to create illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
@@
 
-This opens the door of an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+This opens the door for an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
 Moreover, the illusion effect can be either "incongruent" (making the task even harder by biasing the perception in the opposite way) or "congruent" (making the task easier).
-Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm that could quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
+Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
 
@@
 In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separated statistically.
-Then, we will further utilize the paradigm to assess whether 10 different canonical illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor (a long-standing debate, [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]).
-Finally, we will investigate how the the inter-individual sensitivity to illusions relates to other characteristics, such as demographic variables and personality.
+Then, we will further utilize the paradigm to assess whether 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor.
+Finally, we will investigate how the the inter-individual sensitivity to illusions relates to dispositional variables, such as demographic characteristics and personality.
 
-In line with our open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available and open-source at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+In line with total open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
 
@@
 
-The objective of Study 1 is to act as a pilot study in order to gather some preliminary data to: 1) assess whether the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., increase of task difficulty and illusion strength leads to an increased error rate) and 2) develop a gauged understanding of the magnitude of effects, to then refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard).
+Study 1 can be seen as a pilot experiment aiming at gathering some preliminary data to assess if the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors); and develop an intuition about the magnitude of effects, to refine for the next study the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard).
 
@@
 We generated 56 stimuli for each of the 10 illusion types.
-These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where the negative values correspond to congruent illusion strengths).
+These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where negative values correspond to congruent illusion strengths).
 
-The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the block was also randomized.
-After the first series of 10 blocks, another series was done (with new randomized blocks and trials order).
+The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the blocks was also randomized.
+After the first series of 10 blocks, another series was done (with new randomized order of blocks and trials).
 In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
-The task was implemented using *jsPsych* (*REF*).
-See instructions for each illusion type in the experiment code.
+The task was implemented using *jsPsych* [@de2015jspsych].
+The instructions for each illusion type are available in the experiment code.
 
@@
 
-Fifty-two participants were recruited via *Prolific* ([www.prolificacademic.co.uk](www.prolificacademic.co.uk)), a popular crowd-sourcing platform to recruit subjects for online behavioural research.
-In a study comparing across different research recruitment platforms (e.g., MTurks, Qualtrics etc.), participants from *Prolific* were found to provide the highest data quality across all examined measures, including that of attention (the extent to which participants actually read the questions prior to responding), comprehension (the extent to which participants were able to understand the task instructions or items), honesty (the degree to which participants responded to questions truthfully) and reliability (the extent to which participants' responses are internally consistent within the same scale) [@peer2022].
-For the purposes of our study, no restrictions on age or country were imposed on participant recruitment; the only requisite was a "fluent" proficiency in English to ensure that the task instructions, given in English, would be well-comprehended.
-Participants were incentivised with a reward of about \textsterling 7.5 for completing the task, which took an average of an hour to finish.
+Fifty-two participants were recruited via *Prolific* ([www.prolificacademic.co.uk](www.prolificacademic.co.uk)), a crowd-sourcing platform providing high data quality [@peer2022].
+The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
+Participants were incentivised with a reward of about \textsterling 7.5 for completing the task, which took about 50 minutes to finish.
 
-Average error rate and reaction time distributions were analysed in the pre-processing of the data.
-We removed 6 participants upon inspection of their average error rage (which was close to 50%, suggesting random answers), and because their reaction time distributions were implausibly fast.
-For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that participants misunderstood the instructions given; e.g., participants were selecting the shorter line instead of the longer one).
-Finally, we removed 692 (1.37%) trials based on an implausibly short or lomg response time (\< 150 ms or \> 3000 ms).
+We removed 6 participants upon inspection of the average error rage (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
+For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that instructions got misunderstood; e.g., participants were selecting the shorter line instead of the longer one).
+Finally, we removed 692 (1.37%) trials based on an implausibly short or long response time (< 150 ms or > 3000 ms).
 
@@
 The analysis of study 1 focused on the probability of errors as the main outcome variable.
-For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some perspective on the underlying generative model.
+For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some intuition on the underlying generative model.
 Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
@@
 
-The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019] and the *easystats* collection of packages [@easystatsPackage].
+The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle].
 
@@
 
-The statistical models suggested that the effect of task difficulty had a cubic relationship with error rate for the Delboeuf and Ebbinghaus illusions (both composed of circular shapes), square (\*\*quadratic/inverse?\*\*) relationship for the Rod and Frame and Vertical-Horizontal illusions, cubic relationship for the Zöllner and Poggendorff illusions, exponential relationship for the White illusion, cubic relationship for the Müller-Lyer and Ponzo illusions (both based on line lengths), and linear relationship for the Contrast illusion.
-Please see details and figures in the analysis script.
+The statistical models suggested that the effect of task difficulty had a cubic relationship with error rate for the Delboeuf and Ebbinghaus illusions (both composed of circular shapes), square relationship for the Rod and Frame and Vertical-Horizontal illusions, cubic relationship for the Zöllner and Poggendorff illusions, exponential relationship for the White illusion, cubic relationship for the Müller-Lyer and Ponzo illusions (both based on line lengths), and linear relationship for the Contrast illusion.
+All models suggested a significant effect of illusion strength and task difficulty.
+See details and figures in the analysis script.
 
@@
 
-This study enabled a more precise elaboration, and therefore clearer understanding, of the exact magnitude of the parametric effects at stake and the type of interaction between them.
-Furthermore, it allowed us to better understand and test the stimuli generated by *Pyllusion*, as well as unintentionally uncover a few technological bugs and issues (for instance, the specification direction of the illusion strength was reversed in the software), which were fixed in a new release.
+This study provided a clearer understanding of the magnitude of the parametric effects at stake and the type of interaction between them.
+Furthermore, it allowed us to better understand and test the stimuli generated by *Pyllusion*, as well as uncover technical bugs and issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed by a new software release.
+Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
 
-Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
-In most illusions, the task difficulty exhibited a non-linear effect, which is in line with the psychophysics literature on perceptual decisions ([@ditzinger2010; @shekhar2021; @bogacz2006]).
-In particular, one notable result was the illusion effect for the Zöller illusion, which suggested a non-linear relationship.
+In most illusions, the task difficulty exhibited monotonic power-law scaled effects, which is in line with the psychophysics literature on perceptual decisions [@ditzinger2010; @shekhar2021; @bogacz2006].
+One notable result was the illusion effect pattern for the Zöllner illusion, which suggested a non-linear relationship.
 By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.
@@
 The aim of study 2 was two-fold.
-In the first part, we will carefully model the error rate and the reaction time of each illusion type in order to validate our novel paradigm and show that the effect of illusions can be manipulated parametrically.
-In the second part, we will derive the participant-level scores from the models (i.e., the effect of illusion strength, task difficulty and their interaction for each individual) and analyze their latent factors structure.
+In the first part, we carefully modeled the error rate and the reaction time of each illusion type in order to validate our novel paradigm and show that the effect of illusions can be manipulated continuously.
+In the second part, we derived the participant-level scores from the models (i.e., the effect of illusion strength for each individual) and analyzed their latent factors structure.
 
@@
 
-The paradigm of study 2 was similar to that of study 1, with a few changes: namely, the illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
+The paradigm of study 2 was similar to that of study 1, with the following changes.
+The illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
 Moreover, taking into account the findings of study 1, we used non-linearly spaced difficulty levels, depending on the best underlying model (i.e., with an exponential, square or cubic spacing depending on the relationship).
-For instance, a linear space of [0.1, 0.3, 0.5, 0.7] can be transformed to an exponential space of [0.1, 0.27, 0.47, 0.7].
+For instance, a linear space of [0.1, 0.4, 0.7, 1.0] can be transformed to an exponential space of [0.1, 0.34, 0.64, 1.0].
 
-Additionally, instead of repeating each series of illusion blocks once so that each illusion stimuli is seen two times, we generated more stimuli by increasing the degree of variation in the manipulated parameters (i.e., with more levels of difficulty and illusion strength).
+Additionally, instead of repeating each stimulus two times, we generated illusions using more levels of difficulty and illusion strength.
 As such, for each illusion type, we generated a total of 134 stimuli that were split into two groups (67 stimuli per illusion block).
@@
 
-The first part of the analysis focuses on modelling the effect of illusion strength and task difficulty on errors and reaction time, within each illusion.
-In order to achieve that, we will start by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
-Errors were analyzed using Bayesian binomial mixed models, and reaction times (RT) of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT), within each illusion.
+In order to achieve that, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
+Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].
 
-Using the GAMs as a "ground-truth", we will attempt at approximating them using general linear models, which have the advantage of specifying a participant-level variability of the effects (via random effects).
-Firstly, we will do a model comparison between a combination of transformations (raw, log, square root or cubic root) on the main predictors (task difficulty *diff* and illusion *strength*).
-We will then select and fit the best models (best on their indices of fit), and compare their predictions visually (see **Figure 2**).
+Using the GAMs as a "ground-truth", we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we selected and fitted the best model (best on their indices of fit), and compared their output visually (see **Figure 2**).
 
-We will then extract the inter-individual variability (specified as random slopes) in the effects of illusion strength, task difficulty and their interaction and use it as participant-level scores.
-Finally, We will explore the relationship across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
+We then extracted the inter-individual variability in the effect of illusion strength and its interaction with task difficulty, and used it as participant-level scores.
+Finally, We explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
 
@@
 
-The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; ....
-In all of these models, the effect of illusion strength, task difficulty and their interaction were significant.
+The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; $cbrt(diff)*strength$ for Zöllner; $diff*sqrt(strength)$ and $log(diff)*strength$ respectively for errors and RT in White; $sqrt(diff)*sqrt(strength)$ and $sqrt(diff)*strength$ respectively for errors and RT in Müller-Lyer; $cbrt(diff)*strength$ for Ponzo; $cbrt(diff)*sqrt(strength)$ and $cbrt(diff)*strength$ respectively for errors and RT in Poggendorff; $sqrt(diff)*sqrt(strength)$ for Contrast.
+In all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
 
-With regards to its effect on errors, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0.
+For errors, most of the models closely matched their GAMs counterpart (see **Figure 2**), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner, in which theoretically congruent illusion effects were related to increased error rate.
 
-However, regarding the effect on reaction times, the GAMs suggested a consistent non-linear relationship between the mean RT and illusion strength: as the illusion strength increases, the reaction times are faster.
-While this is is not entirely surprising (for instance, it is conceptually possible to theorize that strong illusions are so effective at biasing our visual perceptions that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this pattern and hence are not good representatives of the underlying dynamics.
-Because the random-slopes models for RT did not appear to capture well the underlying, likely quadratic, relationship, we decided not to include them for the individual scores analysis.
-<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size).
--->
+```{r message=FALSE, warning=FALSE, fig.cap="CAPTION."}
+knitr::include_graphics("figures/Figure2.png")
+```
 
-Though imperfect, we believe that these models capture inter-individual differences more accurately than basic empirical scores (e.g., the total number of errors, or the average RT).
-Thus, for each illusion, we extracted the effect of illusion strength and its interaction with task difficulty (when the illusion effect was incongruent) within each participant.
-These twenty participant-level scores were subjected to exploratory factor analysis (EFA), which suggested that ...
+For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increase beyond a certain threshold, the participants respond faster.
+While this is is not surprising (strong illusions are likely so effective in biasing perception that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
+As such, we decided not to use add them for the individual scores analysis.
+<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size). -->
 
-We have to keep in mind that individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of ...), 2) the structural equation model from which they are obtained is also constructed on layers of abstractions.
-Thus, we wanted to test wether these indices relate to some simple empirical scores, such as the average error rate and RT in the task.
+Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy than basic empirical scores, such as the total number of errors, or the average RT.
+Thus, for each illusion and within each participant, we extracted the effect of illusion strength and its interaction with task difficulty when the illusion effect was incongruent.
+These twenty participant-level scores were subjected to exploratory factor analysis (EFA).
+The Method Agreement Procedure [@parametersArticle] suggested the presence of 7 latent factors.
+An oblique (*oblimin* rotation) factor solution explaining 66.69% of variance suggested separate dimensions for the effect of Zöllner, White, Poggendorff, Contrast, Ebbinghaus, Delboeuf, and a common factor for the parameters related to Müller-Lyer, Vertical-Horizontal, Ponzo and Rod and Frame.
+We submitted these factors to a second-level analysis and extracted two orthogonal (*varimax* rotation) factors.
+The first factor was loaded by all the previous dimensions with the exception of Delboeuf, which formed its own separate factor.
+
+Finally, we tested this data-driven model (*m0*) against four other structural models using structural equation modelling (SEM): one in which the two parameters of each of the 10 illusions (illusion strength and interaction with task difficulty) loaded on separate factors, which then all loaded on a common factor (*m1*); one which the parameters were grouped by illusion type (lines, circles, contrast and angle) before loading on a common factor (*m2*); one in which all the parameters related to strength, and all the parameters related to the interaction loaded onto two respective factors, which then loaded on a common factor (*m3*); and one in which there was no intermediate level: all 20 parameters loaded directly on a common factor (*m4*).
+
+The model *m1*, in which the parameters loaded on a first level of 10 illusion-specific factors, which then all loaded on a common factor significantly outperformed the other models.
+Its indices of fit were ranging from acceptable to satisfactory (CFI = .92; SRMR = .08; NNFI = .91; PNFI = .74; RMSEA = .08), and all the specified effects were significant.
+The illusion-specific latent factors were loaded positively by the sensitivity to illusion strength, and positively by the interaction effect with task difficulty (with the exception of Delboeuf, Ebbinghaus, Vertical-Horizontal, Müller-Lyer and Contrast, for which the loading was negative).
+The general factor of illusion sensitivity, labelled Factor *i*  (i- for illusion), explained 48.02\% of the total variance of the initial dataset, and was strongly related to Vertical-Horizontal ($\beta_{std.}=0.83$), Müller-Lyer ($\beta_{std.}=0.76$), Ponzo ($\beta_{std.}=0.65$), Ebbinghaus ($\beta_{std.}=0.64$); moderately to Zöllner ($\beta_{std.}=0.53$), Poggendorff ($\beta_{std.}=0.44$), Rod and Frame ($\beta_{std.}=0.42$), Contrast ($\beta_{std.}=0.40$) and White ($\beta_{std.}=0.35$); and weakly to Delboeuf ($\beta_{std.}=0.19$).
+We then computed, for each participant, its score for the 10 illusion-specific factors and for the general Factor *i*.
+
+We have to keep in mind that these individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of Delboeuf and Zöllner); 2) we only used the models on error rate,  which could be biased by the speed-accuracy decision criterion used by participants; 3) the structural equation model used to compute the scores also incorporated multiple levels of abstractions.
+Thus, in order to validate the individual scores, we computed the correlation between them and simple empirical scores, such as the average error rate and the mean RT in the task.
+This analysis revealed strong and significant correlations between each illusion-specific factor and the average amount of errors in its respective task.
+Moreover, each individual score was strongly associated with the average RT across multiple illusion types.
+This suggests that the individual scores obtained from the structural equation model do capture the sensitivity of each participant to visual illusions, manifesting in both the amount of errors and high reaction times.
 
@@
 
-<!-- This is likely to be caused by the outline. It will be fixed in the next release of the Pyllusion software. -->
+This study confirmed that it was possible to manipulate continuously the effect of illusion strength for 10 classical illusions.
+Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
+Future studies are needed to explore reaction times and try to identify the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
 
-Future studies should investigate reaction times to find the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
+The effect on errors was monotonic for most illusions, but Delboeuf and Zöllner.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a slight and strong increase of errors, respectively.
+For Delboeuf, we believe that it was an artifact caused by illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
+This was fixed in latest release of *Pyllusion* (v1.2), which now generate outlines of the same size as the target circle.
+For Zöllner, however, we did not find a good explanation of the pattern.
+**TODO: is there some explanation that we can find in the literature?**
+
+Finally, this study provided evidence for both the existence of illusion-specific factors, as well as for a common latent factor (labelled Factor *i*) that explained about half of the total variance.
+These participant-level scores were related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
 
@@
 
+<!-- Dump  -->
+Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions [@mylniec2016].
+For instance, while some researchers have tried to explain our sensitivity to illusory effects as a result of deficits in the low-level visual processing system [@cretenoud2019; @gori2016], others have provided a compelling case using a top-down approach, suggesting that such visual phenomena occur as a result of a conflict between our visual input and our prior beliefs[@teufel2018; @caporuscio2022].
+
+Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
+Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder [ASD]) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
+
 # Future Directions
@@
 The task, data and analysis script are available in open-access at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+
+# Acknowledgments
+
+We would like to thank Tam Pham and Zen J. Lau for their contribution to *Pyllusion*, as well as Prof Dólos for the inspiration.
 

### 2022-09-05 manuscript.tex
update

--- 2022-08-27 manuscript.tex
@@
 \hypersetup{
-  pdftitle={The Illusion Game: A Novel Experimental Paradigm to study Visual Illusions},
+  pdftitle={The Illusion Game: A Novel Experimental Paradigm provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Pathological Personality Correlates},
   pdfauthor={Dominique Makowski1},
@@
 
-\title{\textbf{The Illusion Game: A Novel Experimental Paradigm to study Visual Illusions}}
+\title{\textbf{The Illusion Game: A Novel Experimental Paradigm provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Pathological Personality Correlates}}
 \author{Dominique Makowski\textsuperscript{1}}
@@
 
+Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore.
+E-mail: \href{mailto:dom.makowski@gmail.com}{\nolinkurl{dom.makowski@gmail.com}}
+
 }
@@
 \abstract{%
-Abstract abstract.
+Abstract abstract. We invite researchers to re-analyze our open-access data to provide complimentary evidence (or absence thereof) on the effect and structure of illusion sensitivity.
 }
@@
 \hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}}
 
 Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems.
-They eloquently show that our brains did not evolve to be perfect perceptual devices, but to take into account contextual information and prior knowledge that manifests in our conscious experience \textbf{(REF)}.
-Despite the historical interest within the fields of visual perception \textbf{(REF)}, consciousness science \textbf{(REF)}, or psychiatry \textbf{(REF)}, several gaps remained unanswered.
-\emph{mention common factor for illusions and link with disorders}.
+They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but to integrate prior knowledge and contextual information - blended together in our subjective conscious experience (Carbon, 2014).
+Despite the historical and intensive interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science Lamme (2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
 
-One key challenge hindering the further development of illusion research is the relative difficulty to adapt visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-To address this issue, we first developed a parametric framework to manipulate visual illusions, that we implemented and made accessible in the open-source software \emph{Pyllusion} \textbf{(REF)}.
-This software allows us to generate different types of historical visual illusions with a continuous and independent modulation of two parameters, \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
+Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud et al., 2020, 2019a; Hamburger, 2016; Teufel et al., 2018b); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019b; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018a) are strongly debated.
+The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022) or \textbf{Insert some demographic / personality feature} {[}\textbf{REF}{]} - is another area of controversy.
+
+One key challenge hindering the further development of illusion research is the relative difficulty in adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021a).
+This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus) with a continuous and independent modulation of two parameters, \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
 
 \begin{figure}
-\includegraphics[width=4.2in]{figures/Figure1} \caption{Explanation of the parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
-Examples of stimuli showcasing the manipulation of the two main parameters, the task difficulty and the illusion strength (below).}\label{fig:unnamed-chunk-2}
+\includegraphics[width=4.2in]{figures/Figure1} \caption{The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.}\label{fig:unnamed-chunk-2}
 \end{figure}
 
-Indeed, many visual illusions can be seen as made of \emph{targets} (e.g., same-length lines), which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the segments appear of different lengths when they end with inwards or outwards pointing arrows).
-While most of the paradigms used in the literature prompt the subjective extent the context biases the perception (``how much does the circle appear bigger'', e.g., \textbf{REFS}), \emph{Pyllusions} allows to create illusions in which the targets are actually more or less different (i.e., one segment is longer than the other), and in which the illusion is of varying strength (the arrows are more or less pointy).
+Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (\emph{REF}), or having them adjust the targets to a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016a).
+Alternatively, \emph{Pyllusions} allows to create illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
-This opens the door of an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+This opens the door for an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
 Moreover, the illusion effect can be either ``incongruent'' (making the task even harder by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
-Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm that could quantify the ``objective'' effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
+Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the ``objective'' effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
 
-In the present set of studies, we will first test this novel paradigm by investigate if the effect of illusion can be manipulated continuously.
-Then, we will use it to assess whether 10 different illusions share a common latent factor (a long-standing debate), 3) how the the inter-individual sensitivity to illusions relates to other characteristics, such as sex or personality.
+In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separated statistically.
+Then, we will further utilize the paradigm to assess whether 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor.
+Finally, we will investigate how the the inter-individual sensitivity to illusions relates to dispositional variables, such as demographic characteristics and personality.
+
+In line with total open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
 
 \hypertarget{study-1}{%
-\subsection{Study 1}\label{study-1}}
+\section{Study 1}\label{study-1}}
 
 \hypertarget{aim}{%
-\subsubsection{Aim}\label{aim}}
+\subsection{Aim}\label{aim}}
+
+Study 1 can be seen as a pilot experiment aiming at gathering some preliminary data to assess if the stimuli generated by \emph{Pyllusion} behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors); and develop an intuition about the magnitude of effects, to refine for the next study the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard).
+
+\hypertarget{procedure}{%
+\subsection{Procedure}\label{procedure}}
+
+We generated 56 stimuli for each of the 10 illusion types.
+These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., {[}1, 2, 3, 4, 5, 6, 7{]}, where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., {[}-3, -2, -1, 0, 1, 2, 3{]}, where negative values correspond to congruent illusion strengths).
+
+The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the blocks was also randomized.
+After the first series of 10 blocks, another series was done (with new randomized order of blocks and trials).
+In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
+The task was implemented using \emph{jsPsych} (De Leeuw, 2015).
+The instructions for each illusion type are available in the experiment code.
 
 \hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
+\subsection{Participants}\label{participants}}
 
-\hypertarget{procedure}{%
-\subsubsection{Procedure}\label{procedure}}
+Fifty-two participants were recruited via \emph{Prolific} (\url{www.prolificacademic.co.uk}), a crowd-sourcing platform providing high data quality (Peer et al., 2022).
+The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
+Participants were incentivised with a reward of about \textsterling 7.5 for completing the task, which took about 50 minutes to finish.
+
+We removed 6 participants upon inspection of the average error rage (when close to 50\%, suggesting random answers), and when the reaction time distribution was implausibly fast.
+For the remaining participants, we discarded blocks where the error rate was higher than 50\% (possibly indicating that instructions got misunderstood; e.g., participants were selecting the shorter line instead of the longer one).
+Finally, we removed 692 (1.37\%) trials based on an implausibly short or long response time (\textless{} 150 ms or \textgreater{} 3000 ms).
+
+The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: {[}19, 60{]}; Sex: 39.1\% females, 56.5\% males).
+
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
+
+The analysis of study 1 focused on the probability of errors as the main outcome variable.
+For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some intuition on the underlying generative model.
+Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up \emph{vs.} down or left \emph{vs.} right) as an additional predictor.
+We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships (at the expense of model simplicity).
+
+The analysis was carried out using \emph{R 4.2} (R Core Team, 2022), \emph{brms} (Bürkner, 2017), the \emph{tidyverse} (Wickham et al., 2019), and the \emph{easystats} collection of packages (Lüdecke et al., 2021, 2019; Makowski et al., 2020, 2019).
 
 \hypertarget{results}{%
-\subsubsection{Results}\label{results}}
+\subsection{Results}\label{results}}
+
+The statistical models suggested that the effect of task difficulty had a cubic relationship with error rate for the Delboeuf and Ebbinghaus illusions (both composed of circular shapes), square relationship for the Rod and Frame and Vertical-Horizontal illusions, cubic relationship for the Zöllner and Poggendorff illusions, exponential relationship for the White illusion, cubic relationship for the Müller-Lyer and Ponzo illusions (both based on line lengths), and linear relationship for the Contrast illusion.
+All models suggested a significant effect of illusion strength and task difficulty.
+See details and figures in the analysis script.
 
 \hypertarget{discussion}{%
-\subsubsection{Discussion}\label{discussion}}
+\subsection{Discussion}\label{discussion}}
+
+This study provided a clearer understanding of the magnitude of the parametric effects at stake and the type of interaction between them.
+Furthermore, it allowed us to better understand and test the stimuli generated by \emph{Pyllusion}, as well as uncover technical bugs and issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed by a new software release.
+Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
+
+In most illusions, the task difficulty exhibited monotonic power-law scaled effects, which is in line with the psychophysics literature on perceptual decisions (Bogacz et al., 2006; Ditzinger, 2010; Shekhar \& Rahnev, 2021).
+One notable result was the illusion effect pattern for the Zöllner illusion, which suggested a non-linear relationship.
+By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.
+
+\hypertarget{study-2}{%
+\section{Study 2}\label{study-2}}
+
+\hypertarget{aim-1}{%
+\subsection{Aim}\label{aim-1}}
+
+The aim of study 2 was two-fold.
+In the first part, we carefully modeled the error rate and the reaction time of each illusion type in order to validate our novel paradigm and show that the effect of illusions can be manipulated continuously.
+In the second part, we derived the participant-level scores from the models (i.e., the effect of illusion strength for each individual) and analyzed their latent factors structure.
+
+\hypertarget{procedure-1}{%
+\subsection{Procedure}\label{procedure-1}}
+
+The paradigm of study 2 was similar to that of study 1, with the following changes.
+The illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
+Moreover, taking into account the findings of study 1, we used non-linearly spaced difficulty levels, depending on the best underlying model (i.e., with an exponential, square or cubic spacing depending on the relationship).
+For instance, a linear space of {[}0.1, 0.4, 0.7, 1.0{]} can be transformed to an exponential space of {[}0.1, 0.34, 0.64, 1.0{]}.
+
+Additionally, instead of repeating each stimulus two times, we generated illusions using more levels of difficulty and illusion strength.
+As such, for each illusion type, we generated a total of 134 stimuli that were split into two groups (67 stimuli per illusion block).
+Furthermore, instead of a simple break screen, we added two personality questionnaires between the two series of 10 illusion blocks (see study 3).
+
+\hypertarget{participants-1}{%
+\subsection{Participants}\label{participants-1}}
+
+Using the same recruitment procedure as in study 1, we recruited 256 participants, within which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: {[}18, 69{]}; Sex: 48\% females, 52\% males).
+Please see study 3 for the full demographic breakdown.
+We discarded blocks with more than 50\% of errors (2.16\% of trials) and 0.76\% trials with extreme response times (\textless{} 125 ms or \textgreater{} 4 SD above mean).
+
+\hypertarget{data-analysis-1}{%
+\subsection{Data Analysis}\label{data-analysis-1}}
+
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT), within each illusion.
+In order to achieve that, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
+Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location \(\mu\) (mean), scale \(\sigma\) (spread) and tail-dominance \(\tau\) of the RT distribution (Balota \& Yap, 2011; Matzke \& Wagenmakers, 2009).
+
+Using the GAMs as a ``ground-truth'', we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task \emph{difficulty} and illusion \emph{strength}), we selected and fitted the best model (best on their indices of fit), and compared their output visually (see \textbf{Figure 2}).
+
+We then extracted the inter-individual variability in the effect of illusion strength and its interaction with task difficulty, and used it as participant-level scores.
+Finally, We explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
+
+\hypertarget{results-1}{%
+\subsection{Results}\label{results-1}}
+
+The best models were \(log(diff)*strength\) for Delboeuf; \(sqrt(diff)*strength\) for Ebbinghaus; \(log(diff)*log(strength)\) for Rod and Frame; \(sqrt(diff)*sqrt(strength)\) for Vertical-Horizontal; \(cbrt(diff)*strength\) for Zöllner; \(diff*sqrt(strength)\) and \(log(diff)*strength\) respectively for errors and RT in White; \(sqrt(diff)*sqrt(strength)\) and \(sqrt(diff)*strength\) respectively for errors and RT in Müller-Lyer; \(cbrt(diff)*strength\) for Ponzo; \(cbrt(diff)*sqrt(strength)\) and \(cbrt(diff)*strength\) respectively for errors and RT in Poggendorff; \(sqrt(diff)*sqrt(strength)\) for Contrast.
+In all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
+
+For errors, most of the models closely matched their GAMs counterpart (see \textbf{Figure 2}), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner, in which theoretically congruent illusion effects were related to increased error rate.
+
+\begin{figure}
+\includegraphics[width=20in]{figures/Figure2} \caption{CAPTION.}\label{fig:unnamed-chunk-3}
+\end{figure}
+
+For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increase beyond a certain threshold, the participants respond faster.
+While this is is not surprising (strong illusions are likely so effective in biasing perception that it is ``easier'', i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
+As such, we decided not to use add them for the individual scores analysis.
+
+Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy than basic empirical scores, such as the total number of errors, or the average RT.
+Thus, for each illusion and within each participant, we extracted the effect of illusion strength and its interaction with task difficulty when the illusion effect was incongruent.
+These twenty participant-level scores were subjected to exploratory factor analysis (EFA).
+The Method Agreement Procedure (Lüdecke et al., 2020) suggested the presence of 7 latent factors.
+An oblique (\emph{oblimin} rotation) factor solution explaining 66.69\% of variance suggested separate dimensions for the effect of Zöllner, White, Poggendorff, Contrast, Ebbinghaus, Delboeuf, and a common factor for the parameters related to Müller-Lyer, Vertical-Horizontal, Ponzo and Rod and Frame.
+We submitted these factors to a second-level analysis and extracted two orthogonal (\emph{varimax} rotation) factors.
+The first factor was loaded by all the previous dimensions with the exception of Delboeuf, which formed its own separate factor.
+
+Finally, we tested this data-driven model (\emph{m0}) against four other structural models using structural equation modelling (SEM): one in which the two parameters of each of the 10 illusions (illusion strength and interaction with task difficulty) loaded on separate factors, which then all loaded on a common factor (\emph{m1}); one which the parameters were grouped by illusion type (lines, circles, contrast and angle) before loading on a common factor (\emph{m2}); one in which all the parameters related to strength, and all the parameters related to the interaction loaded onto two respective factors, which then loaded on a common factor (\emph{m3}); and one in which there was no intermediate level: all 20 parameters loaded directly on a common factor (\emph{m4}).
+
+The model \emph{m1}, in which the parameters loaded on a first level of 10 illusion-specific factors, which then all loaded on a common factor significantly outperformed the other models.
+Its indices of fit were ranging from acceptable to satisfactory (CFI = .92; SRMR = .08; NNFI = .91; PNFI = .74; RMSEA = .08), and all the specified effects were significant.
+The illusion-specific latent factors were loaded positively by the sensitivity to illusion strength, and positively by the interaction effect with task difficulty (with the exception of Delboeuf, Ebbinghaus, Vertical-Horizontal, Müller-Lyer and Contrast, for which the loading was negative).
+The general factor of illusion sensitivity, labelled Factor \emph{i} (i- for illusion), explained 48.02\% of the total variance of the initial dataset, and was strongly related to Vertical-Horizontal (\(\beta_{std.}=0.83\)), Müller-Lyer (\(\beta_{std.}=0.76\)), Ponzo (\(\beta_{std.}=0.65\)), Ebbinghaus (\(\beta_{std.}=0.64\)); moderately to Zöllner (\(\beta_{std.}=0.53\)), Poggendorff (\(\beta_{std.}=0.44\)), Rod and Frame (\(\beta_{std.}=0.42\)), Contrast (\(\beta_{std.}=0.40\)) and White (\(\beta_{std.}=0.35\)); and weakly to Delboeuf (\(\beta_{std.}=0.19\)).
+We then computed, for each participant, its score for the 10 illusion-specific factors and for the general Factor \emph{i}.
+
+We have to keep in mind that these individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of Delboeuf and Zöllner); 2) we only used the models on error rate, which could be biased by the speed-accuracy decision criterion used by participants; 3) the structural equation model used to compute the scores also incorporated multiple levels of abstractions.
+Thus, in order to validate the individual scores, we computed the correlation between them and simple empirical scores, such as the average error rate and the mean RT in the task.
+This analysis revealed strong and significant correlations between each illusion-specific factor and the average amount of errors in its respective task.
+Moreover, each individual score was strongly associated with the average RT across multiple illusion types.
+This suggests that the individual scores obtained from the structural equation model do capture the sensitivity of each participant to visual illusions, manifesting in both the amount of errors and high reaction times.
+
+\hypertarget{discussion-1}{%
+\subsection{Discussion}\label{discussion-1}}
+
+This study confirmed that it was possible to manipulate continuously the effect of illusion strength for 10 classical illusions.
+Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
+Future studies are needed to explore reaction times and try to identify the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
+
+The effect on errors was monotonic for most illusions, but Delboeuf and Zöllner.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a slight and strong increase of errors, respectively.
+For Delboeuf, we believe that it was an artifact caused by illusion generation algorithm: the outline of the target circles was always created as slitghly bigger, which made the difference between them more obvious at an illusion strength of 0.
+This was fixed in latest release of \emph{Pyllusion} (v1.2), which now generate outlines of the same size as the target circle.
+For Zöllner, however, we did not find a good explanation of the pattern.
+\textbf{TODO: is there some explanation that we can find in the literature?}
+
+Finally, this study provided evidence for both the existence of illusion-specific factors, as well as for a common latent factor (labelled Factor \emph{i}) that explained about half of the total variance.
+These participant-level scores were related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
+
+\hypertarget{study-3}{%
+\section{Study 3}\label{study-3}}
+
+\hypertarget{aim-2}{%
+\subsection{Aim}\label{aim-2}}
+
+\hypertarget{procedure-2}{%
+\subsection{Procedure}\label{procedure-2}}
+
+This study was based on the data collected in study 2.
+However, the variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
+We used the \emph{IPIP6} (24 items; (Sibley et al., 2011)) to measure 6 personality traits (namely, extraversion, openness, conscientiousness, agreeableness, neuroticism and honesty-humility).
+On the other hand, the PID-5 (25 items; (Hopwood et al., 2012)) was used to measure the degree to which participants inhibit psychopathological personality traits (disinhibition, antagonism, detachment, negative affect and psychoticism).
+The participants were the same as in study 2.
+However, due to a technical issue, no data on the personality questions are available for the first eight participants.
 
 \hypertarget{general-discussion}{%
-\subsection{General Discussion}\label{general-discussion}}
+\section{General Discussion}\label{general-discussion}}
+
+Using the parametric illusion generation framework we developed, Pyllusion (Makowski et al., 2021b), we have hence shown that illusions can be manipulated continuously across several different visual illusions.
+This opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.
+The validation of this novel framework also affords future illusion scientists a standardized measure of illusion susceptibility, instead of relying on conventional methods that depend upon participants' subjective perceptions.
+In our paradigm, in which we apply this approach to a reaction-time task, we were able to measure inter-individual scores of objective illusion sensitivity.
+
+Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions (Mylniec \& Bednarek, 2016b).
+For instance, while some researchers have tried to explain our sensitivity to illusory effects as a result of deficits in the low-level visual processing system (Cretenoud et al., 2019b; Gori et al., 2016), others have provided a compelling case using a top-down approach, suggesting that such visual phenomena occur as a result of a conflict between our visual input and our prior beliefs(Caporuscio et al., 2022; Teufel et al., 2018a).
+
+Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
+Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder {[}ASD{]}) (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022), other studies have found no significant differences between such individuals and healthy controls (Kaliuzhna et al., 2019; Spencer \& Ghorashi, 2014; Tibber et al., 2013; Yang et al., 2012) or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects (Grzeczkowski et al., 2018; Manning et al., 2017).
+
+\hypertarget{future-directions}{%
+\section{Future Directions}\label{future-directions}}
+
+We strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+The task, data and analysis script are available in open-access at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+We would like to thank Tam Pham and Zen J. Lau for their contribution to \emph{Pyllusion}, as well as Prof Dólos for the inspiration.
 
@@
 \hypertarget{references}{%
-\subsection{References}\label{references}}
+\section{References}\label{references}}
 
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-balota2011moving}{}}%
+Balota, D.
+A., \& Yap, M.
+J.
+(2011).
+Moving beyond the mean in studies of mental chronometry: The power of response time distributional analyses.
+\emph{Current Directions in Psychological Science}, \emph{20}(3), 160--166.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bogacz2006}{}}%
+Bogacz, R., Brown, E., Moehlis, J., Holmes, P., \& Cohen, J.
+D.
+(2006).
+The physics of optimal decision making: A formal analysis of models of performance in two-alternative forced-choice tasks.
+\emph{Psychological Review}, \emph{113}(4), 700--765.
+\url{https://doi.org/10.1037/0033-295X.113.4.700}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
+Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel models using {Stan}. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
+
+\leavevmode\vadjust pre{\hypertarget{ref-caporuscio2022}{}}%
+Caporuscio, C., Fink, S.
+B., Sterzer, P., \& Martin, J.
+M.
+(2022).
+When seeing is not believing: A mechanistic basis for predictive divergence.
+\emph{Consciousness and Cognition}, \emph{102}, 103334.
+\url{https://doi.org/10.1016/j.concog.2022.103334}
+
+\leavevmode\vadjust pre{\hypertarget{ref-carbon2014}{}}%
+Carbon, C.-C. (2014). Understanding human perception by human-made illusions. \emph{Frontiers in Human Neuroscience}, \emph{8}. \url{https://www.frontiersin.org/articles/10.3389/fnhum.2014.00566}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020}{}}%
+Cretenoud, A. F., Francis, G., \& Herzog, M. H. (2020). When illusions merge. \emph{Journal of Vision}, \emph{20}(8), 12. \url{https://doi.org/10.1167/jov.20.8.12}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2019}{}}%
+Cretenoud, A.
+F., Karimpur, H., Grzeczkowski, L., Francis, G., Hamburger, K., \& Herzog, M.
+H.
+(2019b).
+Factors underlying visual illusions are illusion-specific but not feature-specific.
+\emph{Journal of Vision}, \emph{19}(14), 12.
+\url{https://doi.org/10.1167/19.14.12}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2019a}{}}%
+Cretenoud, A.
+F., Karimpur, H., Grzeczkowski, L., Francis, G., Hamburger, K., \& Herzog, M.
+H.
+(2019a).
+Factors underlying visual illusions are illusion-specific but not feature-specific.
+\emph{Journal of Vision}, \emph{19}(14), 12.
+\url{https://doi.org/10.1167/19.14.12}
+
+\leavevmode\vadjust pre{\hypertarget{ref-day1972}{}}%
+Day, R.
+H.
+(1972).
+Visual Spatial Illusions: A General Explanation: A wide range of visual illusions, including geometrical distortions, can be explained by a single principle.
+\emph{Science}, \emph{175}(4028), 1335--1340.
+\url{https://doi.org/10.1126/science.175.4028.1335}
+
+\leavevmode\vadjust pre{\hypertarget{ref-de2015jspsych}{}}%
+De Leeuw, J. R. (2015). jsPsych: A JavaScript library for creating behavioral experiments in a web browser. \emph{Behavior Research Methods}, \emph{47}(1), 1--12.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ditzinger2010}{}}%
+Ditzinger, T. (2010). \emph{Optical illusions: Examples for nonlinear dynamics in perception} (Vol. 328, pp. 179--191).
+
+\leavevmode\vadjust pre{\hypertarget{ref-eagleman2001}{}}%
+Eagleman, D. M. (2001). Visual illusions and neurobiology. \emph{Nature Reviews Neuroscience}, \emph{2}(12), 920--926. \url{https://doi.org/10.1038/35104092}
+
+\leavevmode\vadjust pre{\hypertarget{ref-giaouri2011}{}}%
+Giaouri, S., \& Alevriadou, A.
+(2011).
+Are children with down syndrome susceptible to visual illusions?
+\emph{Procedia - Social and Behavioral Sciences}, \emph{15}, 1988--1992.
+\url{https://doi.org/10.1016/j.sbspro.2011.04.040}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gomez-villa2022}{}}%
+Gomez-Villa, A., Martín, A., Vazquez-Corral, J., Bertalmío, M., \& Malo, J.
+(2022).
+On the synthesis of visual illusions using deep generative models.
+\emph{Journal of Vision}, \emph{22}(8), 2.
+\url{https://doi.org/10.1167/jov.22.8.2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gori2016}{}}%
+Gori, S., Molteni, M., \& Facoetti, A.
+(2016).
+Visual illusions: An interesting tool to investigate developmental dyslexia and autism spectrum disorder.
+\emph{Frontiers in Human Neuroscience}, \emph{10}, 175.
+\url{https://doi.org/10.3389/fnhum.2016.00175}
+
+\leavevmode\vadjust pre{\hypertarget{ref-grzeczkowski2018}{}}%
+Grzeczkowski, L., Roinishvili, M., Chkonia, E., Brand, A., Mast, F.
+W., Herzog, M.
+H., \& Shaqiri, A.
+(2018).
+Is the perception of illusions abnormal in schizophrenia?
+\emph{Psychiatry Research}, \emph{270}, 929--939.
+\url{https://doi.org/10.1016/j.psychres.2018.10.063}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hamburger2016}{}}%
+Hamburger, K. (2016). Visual Illusions Based on Processes: New Classification System Needed: \emph{Perception}. \url{https://doi.org/10.1177/0301006616629038}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hopwood2012}{}}%
+Hopwood, C.
+J., Thomas, K.
+M., Markon, K.
+E., Wright, A.
+G.
+C., \& Krueger, R.
+F.
+(2012).
+DSM-5 personality traits and DSM{\textendash}IV personality disorders.
+\emph{Journal of Abnormal Psychology}, \emph{121}(2), 424--432.
+\url{https://doi.org/10.1037/a0026656}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kaliuzhna2019}{}}%
+Kaliuzhna, M., Stein, T., Rusch, T., Sekutowicz, M., Sterzer, P., \& Seymour, K.
+J.
+(2019).
+No evidence for abnormal priors in early vision in schizophrenia.
+\emph{Schizophrenia Research}, \emph{210}, 245--254.
+\url{https://doi.org/10.1016/j.schres.2018.12.027}
+
+\leavevmode\vadjust pre{\hypertarget{ref-keane2014}{}}%
+Keane, B.
+P., Joseph, J., \& Silverstein, S.
+M.
+(2014).
+Late, not early, stages of Kanizsa shape perception are compromised in schizophrenia.
+\emph{Neuropsychologia}, \emph{56}, 302--311.
+\url{https://doi.org/10.1016/j.neuropsychologia.2014.02.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lamme2020}{}}%
+Lamme, V. A. F. (2020). Visual functions generating conscious seeing. \emph{Frontiers in Psychology}, \emph{11}. \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2020.00083}
+
+\leavevmode\vadjust pre{\hypertarget{ref-parametersArticle}{}}%
+Lüdecke, D., Ben-Shachar, M., Patil, I., \& Makowski, D.
+(2020).
+Extracting, computing and exploring the parameters of statistical models using {R}.
+\emph{Journal of Open Source Software}, \emph{5}(53), 2445.
+\url{https://doi.org/10.21105/joss.02445}
+
+\leavevmode\vadjust pre{\hypertarget{ref-performanceArticle}{}}%
+Lüdecke, D., Ben-Shachar, M., Patil, I., Waggoner, P., \& Makowski, D.
+(2021).
+{performance}: An {R} package for assessment, comparison and testing of statistical models.
+\emph{Journal of Open Source Software}, \emph{6}(60), 3139.
+\url{https://doi.org/10.21105/joss.03139}
+
+\leavevmode\vadjust pre{\hypertarget{ref-insightArticle}{}}%
+Lüdecke, D., Waggoner, P., \& Makowski, D.
+(2019).
+Insight: A unified interface to access information from model objects in {R}.
+\emph{Journal of Open Source Software}, \emph{4}(38), 1412.
+\url{https://doi.org/10.21105/joss.01412}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bayestestRArticle}{}}%
+Makowski, D., Ben-Shachar, M., \& Lüdecke, D.
+(2019).
+{bayestestR}: Describing effects and their uncertainty, existence and significance within the {Bayesian} framework.
+\emph{Journal of Open Source Software}, \emph{4}(40), 1541.
+\url{https://doi.org/10.21105/joss.01541}
+
+\leavevmode\vadjust pre{\hypertarget{ref-correlationArticle}{}}%
+Makowski, D., Ben-Shachar, M., Patil, I., \& Lüdecke, D.
+(2020).
+Methods and algorithms for correlation analysis in {R}.
+\emph{Journal of Open Source Software}, \emph{5}(51), 2306.
+\url{https://doi.org/10.21105/joss.02306}
+
+\leavevmode\vadjust pre{\hypertarget{ref-makowski2021}{}}%
+Makowski, D., Lau, Z.
+J., Pham, T., Paul Boyce, W., \& Annabel Chen, S.
+H.
+(2021a).
+A Parametric Framework to Generate Visual Illusions Using Python.
+\emph{Perception}, \emph{50}(11), 950--965.
+\url{https://doi.org/10.1177/03010066211057347}
+
+\leavevmode\vadjust pre{\hypertarget{ref-makowski2021a}{}}%
+Makowski, D., Lau, Z.
+J., Pham, T., Paul Boyce, W., \& Annabel Chen, S.
+H.
+(2021b).
+A Parametric Framework to Generate Visual Illusions Using Python.
+\emph{Perception}, \emph{50}(11), 950--965.
+\url{https://doi.org/10.1177/03010066211057347}
+
+\leavevmode\vadjust pre{\hypertarget{ref-manning2017}{}}%
+Manning, C., Morgan, M.
+J., Allen, C.
+T.
+W., \& Pellicano, E.
+(2017).
+Susceptibility to ebbinghaus and müller-lyer illusions in autistic children: A comparison of three different methods.
+\emph{Molecular Autism}, \emph{8}(1), 16.
+\url{https://doi.org/10.1186/s13229-017-0127-y}
+
+\leavevmode\vadjust pre{\hypertarget{ref-matzke2009psychological}{}}%
+Matzke, D., \& Wagenmakers, E.-J.
+(2009).
+Psychological interpretation of the ex-gaussian and shifted wald parameters: A diffusion model analysis.
+\emph{Psychonomic Bulletin \& Review}, \emph{16}(5), 798--817.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mylniec2016}{}}%
+Mylniec, A., \& Bednarek, H.
+(2016b).
+Field dependence, efficiency of information processing in working memory and susceptibility to orientation illusions among architects.
+\emph{Polish Psychological Bulletin}, \emph{47}(1), 112--122.
+\url{https://doi.org/10.1515/ppb-2016-0012}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mylniec2016a}{}}%
+Mylniec, A., \& Bednarek, H.
+(2016a).
+Field dependence, efficiency of information processing in working memory and susceptibility to orientation illusions among architects.
+\emph{Polish Psychological Bulletin}, \emph{47}(1), 112--122.
+\url{https://doi.org/10.1515/ppb-2016-0012}
+
+\leavevmode\vadjust pre{\hypertarget{ref-notredame2014}{}}%
+Notredame, C.-E., Pins, D., Deneve, S., \& Jardri, R.
+(2014).
+What visual illusions teach us about schizophrenia.
+\emph{Frontiers in Integrative Neuroscience}, \emph{8}, 63.
+\url{https://doi.org/10.3389/fnint.2014.00063}
+
+\leavevmode\vadjust pre{\hypertarget{ref-park2022}{}}%
+Park, S., Zikopoulos, B., \& Yazdanbakhsh, A.
+(2022).
+Visual illusion susceptibility in autism: A neural model.
+\emph{European Journal of Neuroscience}, \emph{56}.
+\url{https://doi.org/10.1111/ejn.15739}
+
+\leavevmode\vadjust pre{\hypertarget{ref-peer2022}{}}%
+Peer, E., Rothschild, D., Gordon, A., Evernden, Z., \& Damer, E.
+(2022).
+Data quality of platforms and panels for online behavioral research.
+\emph{Behavior Research Methods}, \emph{54}(4), 1643--1662.
+\url{https://doi.org/10.3758/s13428-021-01694-3}
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
+R Core Team. (2022). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-razeghi2022}{}}%
+Razeghi, R., Arsham, S., Movahedi, A., \& Sammaknejad, N.
+(2022).
+The effect of visual illusion on performance and quiet eye in autistic children.
+\emph{Early Child Development and Care}, \emph{192}(5), 807--815.
+\url{https://doi.org/10.1080/03004430.2020.1802260}
+
+\leavevmode\vadjust pre{\hypertarget{ref-shekhar2021}{}}%
+Shekhar, M., \& Rahnev, D. (2021). The nature of metacognitive inefficiency in perceptual decision making. \emph{Psychological Review}, \emph{128}(1), 45--70. \url{https://doi.org/10.1037/rev0000249}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sibley2011}{}}%
+Sibley, C., Luyten, N., Wolfman, M., Mobberley, A., Wootton, L.
+W., Hammond, M., Sengupta, N., Perry, R., West-Newman, T., Wilson, M., McLellan, L., Hoverd, W.
+J., \& Robertson, A.
+(2011).
+The mini-IPIP6: Validation and extension of a short measure of the big-six factors of personality in new zealand.
+\emph{New Zealand Journal of Psychology}, \emph{40}, 142--159.
+
+\leavevmode\vadjust pre{\hypertarget{ref-spencer2014}{}}%
+Spencer, K., \& Ghorashi, S.
+(2014).
+Oscillatory dynamics of gestalt perception in schizophrenia revisited.
+\emph{Frontiers in Psychology}, \emph{5}.
+\url{https://www.frontiersin.org/articles/10.3389/fpsyg.2014.00068}
+
+\leavevmode\vadjust pre{\hypertarget{ref-teufel2018}{}}%
+Teufel, C., Dakin, S.
+C., \& Fletcher, P.
+C.
+(2018a).
+Prior object-knowledge sharpens properties of early visual feature-detectors.
+\emph{Scientific Reports}, \emph{8}(1), 10853.
+\url{https://doi.org/10.1038/s41598-018-28845-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-teufel2018a}{}}%
+Teufel, C., Dakin, S., \& Fletcher, P.
+(2018b).
+Prior object-knowledge sharpens properties of early visual feature-detectors.
+\emph{Scientific Reports}, \emph{8}.
+\url{https://doi.org/10.1038/s41598-018-28845-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-teufel2015}{}}%
+Teufel, C., Subramaniam, N., Dobler, V., Perez, J., Finnemann, J., Mehta, P.
+R., Goodyer, I.
+M., \& Fletcher, P.
+C.
+(2015).
+Shift toward prior knowledge confers a perceptual advantage in early psychosis and psychosis-prone healthy individuals.
+\emph{Proceedings of the National Academy of Sciences}, \emph{112}(43), 13401--13406.
+\url{https://doi.org/10.1073/pnas.1503916112}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tibber2013}{}}%
+Tibber, M., Anderson, E., Bobin, T., Antonova, E., Seabright, A., Wright, B., Carlin, P., Shergill, S., \& Dakin, S.
+(2013).
+Visual surround suppression in schizophrenia.
+\emph{Frontiers in Psychology}, \emph{4}.
+\url{https://www.frontiersin.org/articles/10.3389/fpsyg.2013.00088}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T., Miller, E., Bache, S., Müller, K., Ooms, J., Robinson, D., Seidel, D., Spinu, V., \ldots{} Yutani, H.
+(2019).
+Welcome to the tidyverse.
+\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
+\url{https://doi.org/10.21105/joss.01686}
+
+\leavevmode\vadjust pre{\hypertarget{ref-yang2012}{}}%
+Yang, E., Tadin, D., Glasser, D.
+M., Hong, S.
+W., Blake, R., \& Park, S.
+(2012).
+Visual Context Processing in Schizophrenia: \emph{Clinical Psychological Science}.
+\url{https://doi.org/10.1177/2167702612464618}
+
 \end{CSLReferences}

### 2022-09-05 study2/study2.Rmd
update

@@
 
-plot_ppcheck(model_verticalhorizontal_rt)
+plot_ppcheck(model_verticalhorizontal_rt, gam_verticalhorizontal_rt)
 # performance::check_model(model_verticalhorizontal_rt)
@@
 
-plot_ppcheck(model_zollner_rt)
+plot_ppcheck(model_zollner_rt, gam_zollner_rt)
 # performance::check_model(model_zollner_rt)
@@
 
-plot_ppcheck(model_white_rt)
+plot_ppcheck(model_white_rt, gam_white_rt)
 # performance::check_model(model_white_rt)
@@
 
-plot_ppcheck(model_mullerlyer_rt)
+plot_ppcheck(model_mullerlyer_rt, gam_mullerlyer_rt)
 # performance::check_model(model_mullerlyer_rt)
@@
 
-plot_ppcheck(model_ponzo_rt)
+plot_ppcheck(model_ponzo_rt, gam_ponzo_rt)
 # performance::check_model(model_ponzo_rt)
@@
 
-plot_ppcheck(model_poggendorff_rt)
+plot_ppcheck(model_poggendorff_rt, gam_poggendorff_rt)
 # performance::check_model(model_poggendorff_rt)
@@
 rez <- parameters::n_factors(data, n_max=15)
+rez
 plot(rez)
@@
 
-anova(m0, m1, m1b, m2)
+# Model 4 - No middle level
+lvl1 <- paste("i =~", data$Name)
+
+m4 <- sem(model = paste0(c(lvl1), collapse = "\n"),
+          data  = random,
+          bounds = "standard",
+          estimator = "ML"
+          )
+
+anova(m0, m1, m1b, m2, m4)
 
@@
 ```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+# Variance Explained
+SumCy <- sum(cov(random[-1]), na.rm = TRUE)
+
+get_r2 <- function(data, dat, SumCy) {
+  Cx <- cov(data)
+  Cxy <- cov(data, dat)
+  solved <- tryCatch(solve(Cx), error = function(e) NULL)
+  if (is.null(solved)) {
+    return(NA)
+    next
+  } else {
+    VAF <- t(Cxy) %*% solved %*% Cxy
+    return(sum(VAF, na.rm = TRUE) / SumCy)
+  }
+}
+var <- get_r2(data=dfsem["i"], dat=random[-1], SumCy)
+```
+
+Factor *i* explained `r insight::format_percent(var)` of the variance of the dataset.
 

### 2022-09-06 manuscript.Rmd
edits by @Stephmkirk

@@
 ---
-title             : "**The Illusion Game: A Novel Experimental Paradigm provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Pathological Personality Correlates**"
+title             : "**The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates**"
 shorttitle        : "Illusion Game"
@@
 
-Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems. 
+Visual illusions are fascinating stimuli, capturing a key feature of our neurocognitive systems. 
 They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but to integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
-Despite the historical and intensive interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022, @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
+Despite the historical and intensive interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022; @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
 
@@
 To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021]. 
-This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus) with a continuous and independent modulation of two parameters, *illusion strength* and *task difficulty* (see **Figure 1**).
+This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus) with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (see **Figure 1**).
 
@@
 Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (*REF*), or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016a; @grzeczkowski2018].
-Alternatively, *Pyllusions* allows to create illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
+Alternatively, *Pyllusions* allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
@@
 
-Study 1 can be seen as a pilot experiment aiming at gathering some preliminary data to assess if the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors); and develop an intuition about the magnitude of effects, to refine for the next study the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard).
+Study 1 can be seen as a pilot experiment aiming to gather some preliminary data to assess if the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors); and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
 
@@
 
-Using the same recruitment procedure as in study 1, we recruited 256 participants, within which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
+Using the same recruitment procedure as in study 1, we recruited 256 participants, out of which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
 Please see study 3 for the full demographic breakdown.
@@
 While this is is not surprising (strong illusions are likely so effective in biasing perception that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
-As such, we decided not to use add them for the individual scores analysis.
+As such, we decided not to use them for the individual scores analysis.
 <!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size). -->
 
-Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy than basic empirical scores, such as the total number of errors, or the average RT.
+Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy (and are also more conservative estimates due shrinkage) than basic empirical scores, such as the total number of errors, or the average RT.
 Thus, for each illusion and within each participant, we extracted the effect of illusion strength and its interaction with task difficulty when the illusion effect was incongruent.
@@
 
-This study confirmed that it was possible to manipulate continuously the effect of illusion strength for 10 classical illusions.
+This study confirmed that it was possible to continuously manipulate the effect of illusion strength for 10 classical illusions.
 Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
@@
 
-The effect on errors was monotonic for most illusions, but Delboeuf and Zöllner.
-For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a slight and strong increase of errors, respectively.
+The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a small and strong increase of errors, respectively.
 For Delboeuf, we believe that it was an artifact caused by illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
@@
 
+Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
+
 ## Procedure
@@
 This study was based on the data collected in study 2.
-However, the variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
-We used the *IPIP6* (24 items; [@sibley2011]) to measure 6 personality traits (namely, extraversion, openness, conscientiousness, agreeableness, neuroticism and honesty-humility).
-On the other hand, the PID-5 (25 items; [@hopwood2012]) was used to measure the degree to which participants inhibit psychopathological personality traits (disinhibition, antagonism, detachment, negative affect and psychoticism).
-The participants were the same as in study 2.
-However, due to a technical issue, no data on the personality questions are available for the first eight participants.
+The variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
+We used the *IPIP6* [24 items, @sibley2011] to measure 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-humility), and the PID-5 [25 items, @hopwood2012] to measure "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+The participants were the same as in study 2 (see **Figure 3**).
+However, due to a technical issue, no personality data was recorded for the first eight participants.
+
+```{r message=FALSE, warning=FALSE, fig.cap="CAPTION."}
+knitr::include_graphics("figures/Figure2.png")
+```
+
+## Data Analysis
+
+## Results
+
+## Discussion
 

### 2022-09-06 manuscript.tex
edits by @Stephmkirk

@@
 \hypersetup{
-  pdftitle={The Illusion Game: A Novel Experimental Paradigm provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Pathological Personality Correlates},
+  pdftitle={The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates},
   pdfauthor={Dominique Makowski1},
@@
 
-\title{\textbf{The Illusion Game: A Novel Experimental Paradigm provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Pathological Personality Correlates}}
+\title{\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates}}
 \author{Dominique Makowski\textsuperscript{1}}
@@
 
-Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems.
+Visual illusions are fascinating stimuli, capturing a key feature of our neurocognitive systems.
 They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but to integrate prior knowledge and contextual information - blended together in our subjective conscious experience (Carbon, 2014).
-Despite the historical and intensive interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science Lamme (2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
+Despite the historical and intensive interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science (Caporuscio et al., 2022; Lamme, 2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
 
@@
 To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021a).
-This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus) with a continuous and independent modulation of two parameters, \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
+This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus) with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
 
@@
 Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (\emph{REF}), or having them adjust the targets to a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016a).
-Alternatively, \emph{Pyllusions} allows to create illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
+Alternatively, \emph{Pyllusions} allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
@@
 
-Study 1 can be seen as a pilot experiment aiming at gathering some preliminary data to assess if the stimuli generated by \emph{Pyllusion} behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors); and develop an intuition about the magnitude of effects, to refine for the next study the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard).
+Study 1 can be seen as a pilot experiment aiming to gather some preliminary data to assess if the stimuli generated by \emph{Pyllusion} behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors); and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
 
@@
 
-Using the same recruitment procedure as in study 1, we recruited 256 participants, within which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: {[}18, 69{]}; Sex: 48\% females, 52\% males).
+Using the same recruitment procedure as in study 1, we recruited 256 participants, out of which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: {[}18, 69{]}; Sex: 48\% females, 52\% males).
 Please see study 3 for the full demographic breakdown.
@@
 While this is is not surprising (strong illusions are likely so effective in biasing perception that it is ``easier'', i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
-As such, we decided not to use add them for the individual scores analysis.
+As such, we decided not to use them for the individual scores analysis.
 
-Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy than basic empirical scores, such as the total number of errors, or the average RT.
+Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy (and are also more conservative estimates due shrinkage) than basic empirical scores, such as the total number of errors, or the average RT.
 Thus, for each illusion and within each participant, we extracted the effect of illusion strength and its interaction with task difficulty when the illusion effect was incongruent.
@@
 
-The effect on errors was monotonic for most illusions, but Delboeuf and Zöllner.
-For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a slight and strong increase of errors, respectively.
-For Delboeuf, we believe that it was an artifact caused by illusion generation algorithm: the outline of the target circles was always created as slitghly bigger, which made the difference between them more obvious at an illusion strength of 0.
+The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a small and strong increase of errors, respectively.
+For Delboeuf, we believe that it was an artifact caused by illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
 This was fixed in latest release of \emph{Pyllusion} (v1.2), which now generate outlines of the same size as the target circle.
@@
 
+Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
+
 \hypertarget{procedure-2}{%
@@
 This study was based on the data collected in study 2.
-However, the variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
-We used the \emph{IPIP6} (24 items; (Sibley et al., 2011)) to measure 6 personality traits (namely, extraversion, openness, conscientiousness, agreeableness, neuroticism and honesty-humility).
-On the other hand, the PID-5 (25 items; (Hopwood et al., 2012)) was used to measure the degree to which participants inhibit psychopathological personality traits (disinhibition, antagonism, detachment, negative affect and psychoticism).
-The participants were the same as in study 2.
-However, due to a technical issue, no data on the personality questions are available for the first eight participants.
+The variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
+We used the \emph{IPIP6} (24 items, Sibley et al., 2011) to measure 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-humility), and the PID-5 (25 items, Hopwood et al., 2012) to measure ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+The participants were the same as in study 2 (see \textbf{Figure 3}).
+However, due to a technical issue, no personality data was recorded for the first eight participants.
+
+\begin{figure}
+\includegraphics[width=20in]{figures/Figure2} \caption{CAPTION.}\label{fig:unnamed-chunk-4}
+\end{figure}
+
+\hypertarget{data-analysis-2}{%
+\subsection{Data Analysis}\label{data-analysis-2}}
+
+\hypertarget{results-2}{%
+\subsection{Results}\label{results-2}}
+
+\hypertarget{discussion-2}{%
+\subsection{Discussion}\label{discussion-2}}
 

### 2022-09-06 study2/study2.Rmd
figs

@@
 ```{r message=FALSE, warning=FALSE, cache=cache}
-plot_ppcheck(model_delboeuf_rt, gam_delboeuf_rt)
+# plot_ppcheck(model_delboeuf_rt, gam_delboeuf_rt)
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
-plot_ppcheck(model_ebbinghaus_rt, gam_ebbinghaus_rt)
+# # plot_ppcheck(model_ebbinghaus_rt, gam_ebbinghaus_rt)
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
-plot_ppcheck(model_rodframe_rt, gam_rodframe_rt)
+# plot_ppcheck(model_rodframe_rt, gam_rodframe_rt)
 
@@
 
-plot_ppcheck(model_verticalhorizontal_rt, gam_verticalhorizontal_rt)
+# plot_ppcheck(model_verticalhorizontal_rt, gam_verticalhorizontal_rt)
 # performance::check_model(model_verticalhorizontal_rt)
@@
 
-plot_ppcheck(model_zollner_rt, gam_zollner_rt)
+# plot_ppcheck(model_zollner_rt, gam_zollner_rt)
 # performance::check_model(model_zollner_rt)
@@
 
-plot_ppcheck(model_white_rt, gam_white_rt)
+# plot_ppcheck(model_white_rt, gam_white_rt)
 # performance::check_model(model_white_rt)
@@
 
-plot_ppcheck(model_mullerlyer_rt, gam_mullerlyer_rt)
+# plot_ppcheck(model_mullerlyer_rt, gam_mullerlyer_rt)
 # performance::check_model(model_mullerlyer_rt)
@@
 
-plot_ppcheck(model_ponzo_rt, gam_ponzo_rt)
+# plot_ppcheck(model_ponzo_rt, gam_ponzo_rt)
 # performance::check_model(model_ponzo_rt)
@@
 
-plot_ppcheck(model_poggendorff_rt, gam_poggendorff_rt)
+# plot_ppcheck(model_poggendorff_rt, gam_poggendorff_rt)
 # performance::check_model(model_poggendorff_rt)
@@
 
-plot_ppcheck(model_contrast_rt, gam_contrast_rt)
+# plot_ppcheck(model_contrast_rt, gam_contrast_rt)
 # performance::check_model(model_contrast_rt)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 data <- rbind(
@@
 
+# Vertical
+p_ill <- 
+  (p_delboeuf + p_ebbinghaus + 
+     p_rodframe + p_verticalhorizontal +
+     p_zollner + p_white +
+     p_mullerlyer + p_ponzo +
+     p_poggendorff + p_contrast) + 
+    patchwork::plot_layout(ncol= 2, nrow=5)
+
+ggsave("figures/figure_ill2.png", p_ill, height = 30, width = 20, dpi=300)
+
+# Horizontal
 p_ill <- 

### 2022-09-06 manuscript.Rmd
update manu

@@
 Visual illusions are fascinating stimuli, capturing a key feature of our neurocognitive systems.
-They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but to integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
+They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
 Despite the historical and intensive interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022; @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
 
-Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
-The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014] or **Insert some demographic / personality feature** [**REF**] - is another area of controversy.
+Notably, the presence of a common mechanism central to the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
+The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014]; females tend to be more susceptible to illusory effects than their male counterparts [@miller2001; @papageorgiou2020; @lo2011; @shaqiri2018]- is another area of controversy.
 
@@
 Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (*REF*), or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016a; @grzeczkowski2018].
-Alternatively, *Pyllusions* allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different [@lányi], or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016a; @grzeczkowski2018].
+Alternatively, *Pyllusion* allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
@@
 
-This opens the door for an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can be either "incongruent" (making the task even harder by biasing the perception in the opposite way) or "congruent" (making the task easier).
+This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+Moreover, the illusion effect can be either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier) (see **Figure 1**).
 Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
@@
 
-In line with total open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
 
@@
 We generated 56 stimuli for each of the 10 illusion types.
-These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where negative values correspond to congruent illusion strengths).
+These stimuli resulted from the combination of 8 linearly-spread levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where negative values correspond to congruent illusion strengths).
 
@@
 For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that instructions got misunderstood; e.g., participants were selecting the shorter line instead of the longer one).
-Finally, we removed 692 (1.37%) trials based on an implausibly short or long response time (< 150 ms or > 3000 ms).
+Finally, we removed 692 (1.37%) trials based on an implausibly short or long response time (\< 150 ms or \> 3000 ms).
 
@@
 Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
-We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships (at the expense of model simplicity).
+We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potential non-linear relationships (at the expense of model simplicity).
 
@@
 
-For errors, most of the models closely matched their GAMs counterpart (see **Figure 2**), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner, in which theoretically congruent illusion effects were related to increased error rate.
+For errors, most of the models closely matched their GAMs counterpart (see **Figure 2**), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
 
@@
 As such, we decided not to use them for the individual scores analysis.
-<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size). -->
+<!-- Additionally, it would have increased the parameter space from 20 to 40 variables, which was deemed as statistically unreasonable given our sample size).
+-->
 
@@
 The illusion-specific latent factors were loaded positively by the sensitivity to illusion strength, and positively by the interaction effect with task difficulty (with the exception of Delboeuf, Ebbinghaus, Vertical-Horizontal, Müller-Lyer and Contrast, for which the loading was negative).
-The general factor of illusion sensitivity, labelled Factor *i*  (i- for illusion), explained 48.02\% of the total variance of the initial dataset, and was strongly related to Vertical-Horizontal ($\beta_{std.}=0.83$), Müller-Lyer ($\beta_{std.}=0.76$), Ponzo ($\beta_{std.}=0.65$), Ebbinghaus ($\beta_{std.}=0.64$); moderately to Zöllner ($\beta_{std.}=0.53$), Poggendorff ($\beta_{std.}=0.44$), Rod and Frame ($\beta_{std.}=0.42$), Contrast ($\beta_{std.}=0.40$) and White ($\beta_{std.}=0.35$); and weakly to Delboeuf ($\beta_{std.}=0.19$).
+The general factor of illusion sensitivity, labelled Factor *i* (i- for illusion), explained 48.02% of the total variance of the initial dataset, and was strongly related to Vertical-Horizontal ($\beta_{std.}=0.83$), Müller-Lyer ($\beta_{std.}=0.76$), Ponzo ($\beta_{std.}=0.65$), Ebbinghaus ($\beta_{std.}=0.64$); moderately to Zöllner ($\beta_{std.}=0.53$), Poggendorff ($\beta_{std.}=0.44$), Rod and Frame ($\beta_{std.}=0.42$), Contrast ($\beta_{std.}=0.40$) and White ($\beta_{std.}=0.35$); and weakly to Delboeuf ($\beta_{std.}=0.19$).
 We then computed, for each participant, its score for the 10 illusion-specific factors and for the general Factor *i*.
@@
 Moreover, each individual score was strongly associated with the average RT across multiple illusion types.
-This suggests that the individual scores obtained from the structural equation model do capture the sensitivity of each participant to visual illusions, manifesting in both the amount of errors and high reaction times.
+This suggests that the individual scores obtained from the structural equation model do capture the sensitivity of each participant to visual illusions, manifesting in both the number of errors and high reaction times.
 
@@
 For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a small and strong increase of errors, respectively.
-For Delboeuf, we believe that it was an artifact caused by illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
+Forthe Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
 This was fixed in latest release of *Pyllusion* (v1.2), which now generate outlines of the same size as the target circle.
-For Zöllner, however, we did not find a good explanation of the pattern.
-**TODO: is there some explanation that we can find in the literature?**
+For the Zöllner illusion, however, this observation is consistent with the findings of [@kitaoka2000], who reported an acute angle contraction effect at large angles (50 to 90 degrees) between the target horizontal line and the biasing horizontal bars.
+Although the Zollner illusion is commonly thought to be driven by an acute angle expansion mechanism that is optimal at small angles (below 30 degrees), an indirect illusory effect can be observed at sufficiently large angles due to the opposing acute angle contraction mechanism invoked [@kitaoka2007; @kitaoka2000a].
 
@@
 
+Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
+Whereas some researchers report no relationship between inter-individual traits and inidividuals' susceptibility to illusions [@cretenoud2020a; @grzeczkowski2017], others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects [@zhang2017; @ohmann2016].
+Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal [@pessoa2008].
+
 # General Discussion
 
-Using the parametric illusion generation framework we developed, Pyllusion [@makowski2021a], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
+Using the parametric illusion generation framework we developed, *Pyllusion* [@makowski2021a], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
 This opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.
@@
 <!-- Dump  -->
+
 Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions [@mylniec2016].

### 2022-09-06 manuscript.Rmd
update

--- 2022-09-06 manuscript.Rmd
@@
 
-```{r message=FALSE, warning=FALSE, fig.cap="CAPTION."}
+```{r message=FALSE, warning=FALSE, out.width="100%", fig.cap="Top: the effect of illusion strength and task difficulty on the error rage and reaction time (RT) for each individual illusions.
+The solid line represent the General Additive Model (GAM), and the dashed line correspond to its approximation via linear models.
+Descriptive data is shown with stacked dots (errors are hanging from the top) and distributions for RTs.
+Negative values for illusion strength correspond to congruent (i.e., facilitating) illusion effects.
+Task difficulty (the objective difference between the targets of perceptual decision) levels are shown as colors, with lowest values corresponding to harder trials.
+Each illusion type is surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficutly (on top) and the strongest illusion (on the right for incongruent illusions).
+Bottom: We extracted the effect slope of the illusion strength and its interaction with task difficulty for each participant.
+We fitted a Structural Equation Model (SEM) suggesting that these manifest variables group to first-level illusion-specific latent factors, which then load on a general factor of illusion sensitivity (Factor \\textit{i})."}
 knitr::include_graphics("figures/Figure2.png")
@@
 Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
+**Insert a bit of litterature about why and some findings justifying what we did**.
 
@@
 ```{r message=FALSE, warning=FALSE, fig.cap="CAPTION."}
-knitr::include_graphics("figures/Figure2.png")
+knitr::include_graphics("figures/Figure3.png")
 ```
@@
 
+For each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) and personality.
+As the supplementary material contains the detailed results, we will here only report the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
+
 ## Results
+
+The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions.
+Anecdotal ($BF_{10} = 2.05$) and moderate evidence ($BF_{10} = 4.11$) was found for a negative correlation between screen size and the sensitivity to the White and the Contrast illusion, respectively.
+To test whether this result could be an artifact related to the highly skewed screen size distribution (caused by very few participants with extreme screen sizes), we re-ran a robust correlation (with rank-transformed values), which provided even stronger evidence in favor of the effect existence ($BF_{10} = 28.19$, $BF_{10} = 4.31$ for White and Contrast, respectively).
+
+The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females ($\Delta=-0.37$, 95\% CI [-0.62, -0.13], $BF_{10} = 12.74$).
+We fitted Bayesian linear models with the education level entered as a monotonic predictor [appropriate for ordinal variables, @burkner2020modelling], which yielded no significant effects.
+For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
+These consistently suggested a significant positive linear relationship between age and Factor *i* ($pd=100\%$), as well as the sensitivity to Müller-Lyer ($pd=100\%$), Vertical-Horizontal ($pd=100\%$), Zöllner ($pd=100\%$) and Ebbinghaus ($pd=99\%$) illusions.
+
+Regarding "normal" personality traits, Bayesian correlations suggested substantial evidence in favor of a positive relationship between *Honesty-Humility* and Zöllner ($BF_{10} > 100$), Vertical-Horizontal ($BF_{10} = 9.78$) and the Factor *i* ($BF_{10} = 4.00$); as well as between *Agreeableness* and Vertical-Horizontal ($BF_{10} = 25.06$), Ponzo ($BF_{10} = 4.88$) and the Factor *i* ($BF_{10} = 19.65$).
+
+Regarding "pathological" personality traits, the results yielded strong evidence in favor of a negative relationship between multiple illusion scores and multiple traits.
+*Antagonism* was associated with the sensitivity to Vertical-Horizontal ($BF_{10} > 100$), Müller-Lyer ($BF_{10} = 21.57$), Ponzo ($BF_{10} = 17.97$) illusions, and the Factor *i* ($BF_{10} = 55.45$); *Psychoticism* was associated with the sensitivity to Vertical-Horizontal ($BF_{10} = 66.63$) and Müller-Lyer ($BF_{10} = 35.59$) illusions, and the Factor *i* ($BF_{10} = 35.02$); *Disinhibition* was associated with the sensitivity to Vertical-Horizontal ($BF_{10} = 25.38$), Zöllner ($BF_{10} = 7.59$), Müller-Lyer ($BF_{10} = 5.89$) illusions, and the Factor *i* ($BF_{10} = 31.42$); and *Negative Affect* was associated with Zöllner ($BF_{10} = 62.04$), Vertical-Horizontal ($BF_{10} = 12.65$), Müller-Lyer ($BF_{10} = 3.17$), and the Factor *i* ($BF_{10} = 6.39$).
+The last remaining trait, *Detachment*, did not share any relationship with illusion sensitivity.
 
@@
 In our paradigm, in which we apply this approach to a reaction-time task, we were able to measure inter-individual scores of objective illusion sensitivity.
+
+Notably, the general sensitivity to illusions Factor *i* was negatively associated with *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
 

### 2022-09-06 manuscript.tex
update

@@
 \begin{figure}
-\includegraphics[width=20in]{figures/Figure2} \caption{CAPTION.}\label{fig:unnamed-chunk-3}
+\includegraphics[width=1\linewidth]{figures/Figure2} \caption{Top: the effect of illusion strength and task difficulty on the error rage and reaction time (RT) for each individual illusions.
+The solid line represent the General Additive Model (GAM), and the dashed line correspond to its approximation via linear models.
+Descriptive data is shown with stacked dots (errors are hanging from the top) and distributions for RTs.
+Negative values for illusion strength correspond to congruent (i.e., facilitating) illusion effects.
+Task difficulty (the objective difference between the targets of perceptual decision) levels are shown as colors, with lowest values corresponding to harder trials.
+Each illusion type is surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficutly (on top) and the strongest illusion (on the right for incongruent illusions).
+Bottom: We extracted the effect slope of the illusion strength and its interaction with task difficulty for each participant.
+We fitted a Structural Equation Model (SEM) suggesting that these manifest variables group to first-level illusion-specific latent factors, which then load on a general factor of illusion sensitivity (Factor \textit{i}).}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
+\textbf{Insert a bit of litterature about why and some findings justifying what we did}.
 
@@
 \begin{figure}
-\includegraphics[width=20in]{figures/Figure2} \caption{CAPTION.}\label{fig:unnamed-chunk-4}
+\includegraphics[width=9.33in]{figures/Figure3} \caption{CAPTION.}\label{fig:unnamed-chunk-4}
 \end{figure}

### 2022-09-06 study2/study2.Rmd
update

@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_ebbinghaus, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_rodframe, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_verticalhorizontal, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_zollner, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_white, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_mullerlyer, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=FALSE, eval=bestModels}
+```{r bestmodels_ponzo, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_poggendorff, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
+```{r bestmodels_contrast, message=FALSE, warning=FALSE, cache=cache, eval=bestModels}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 cor <- correlation::correlation(random, redundant = TRUE, p_adjust = "none")
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 data <- select(random, -Participant)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 data <- predict(fa)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 library(lavaan)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 # Variance Explained
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 # Average for 4 higher level of task difficulty
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 r <- correlation(arrange(dfsem, as.character(Participant)), 
@@
 
-### Figure
-
-```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
-# Save
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
+# Save data
 dfsub |> 
@@
   write.csv("../data/study3.csv", row.names = FALSE)
-
+```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 # Figure 1a
@@
 
-ggsave("figures/figure_ill2.png", p_ill, height = 30, width = 20, dpi=300)
+ggsave("figures/figure_ill2.png", p_ill, height = 40, width = 20, dpi=300)
 

### 2022-09-06 study3/study3.Rmd
update

--- 2022-08-27 study3/study3.Rmd
@@
 cache <- TRUE
-runModels <- FALSE
-bestModels <- FALSE
 fig.width <- see::golden_ratio(7)
@@
 The goal of study 3 was to...
-
-We have to keep in mind that individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of ...), 2) the structural equation model from which they are obtained is also constructed on layers of abstractions.
-Though, as we have seen, these indices to relate to empirical scores, such as the average error rate and RT in the task.
 
@@
 
+source("../study2/functions.R")
+
 # Variables
-DVs <- c("Delboeuf" = "#2196F3", 
+DVs <- c(
+  "Delboeuf" = "#2196F3",
   "Ebbinghaus" = "#3F51B5",
@@
   "Contrast" = "#607D8B",
-         "I" = "#9C27B0")
+  "i" = "#9C27B0"
+)
 
@@
     Screen_Refresh = as.numeric(Screen_Refresh),
+    Ethnicity = ifelse(Ethnicity == "Latino", "Hispanic", Ethnicity),
     Education = fct_relevel(Education, "High School", "Bachelor", "Master", "Doctorate", "Other", "Prefer not to Say")
   ) |>
-  datawizard::standardise(select=names(DVs))
+  datawizard::standardise(select = names(DVs)) |>
+  datawizard::change_scale(select = starts_with("IPIP"), to = c(0, 1), range = c(0, 100)) |>
+  datawizard::change_scale(select = starts_with("PID"), to = c(0, 1), range = c(0, 100))
+
+# Remove outliers (p < 0.0001)
+dfsub[names(DVs)][abs(dfsub[names(DVs)]) > qnorm(0.9999)] <- NA
 
@@
   pivot_longer(all_of(names(DVs)), names_to = "Index", values_to = "Score")
-# plot(estimate_density(dfsub[DVs]))
 ```
 
-## Results
+## Participants
 
-### Contextual
+`r sum(is.na(dfsub$IPIP6_Agreeableness))` participants did not do the personality scales.
 
-#### Screen Size
+```{r demographics}
+p_age <- estimate_density(dfsub$Age) |>
+  ggplot(aes(x = x, y = y)) +
+  geom_area(fill = "#607D8B") +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "Age", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(size = rel(1), face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(face = "italic", hjust = 0.5),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = rel(0.8)),
+    axis.title.x = element_blank()
+  )
+
+p_nat <- dfsub |>
+  group_by(Nationality) |>
+  summarize(n = n()) |>
+  mutate(Nationality = fct_reorder(Nationality, desc(n))) |>
+  ggplot(aes(x = Nationality, y = n, fill = Nationality)) +
+  geom_bar(stat = "identity") +
+  labs(y = "Number", title = "Nationality") +
+  scale_fill_material_d(guide = "none") +
+  scale_x_discrete(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(size = rel(1), face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(face = "italic", hjust = 0.5),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_text(size = rel(0.8)),
+    axis.text.x = element_text(size = rel(0.6), angle = 50, hjust = 1)
+  )
+
+# ggsave("figures/figure_dem1.png",
+#        patchwork::wrap_elements(p_age / p_nat),
+#        width=0.5*(fig.height),
+#        height=1*(fig.height), dpi=1000, bg="white")
+
+p_sex <- plot_waffle(dfsub, what = "Sex", rows = 10, size = 5) +
+  scale_color_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63"))
+
+p_edu <- plot_waffle(dfsub, "Education", rows = 10, size = 5) +
+  scale_color_viridis_d()
+
+p_race <- plot_waffle(dfsub, "Ethnicity", rows = 10, size = 5) +
+  scale_color_manual(values = c("Hispanic" = "#FF5722", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Other" = "#795548"))
+
+# ggsave("figures/figure_dem2.png",
+#        patchwork::wrap_elements(p_sex / p_edu / p_race), width=1*(fig.height), height=1*(fig.height), dpi=1000, bg="white")
+
+p_ipip <- dfsub |>
+  select(starts_with("IPIP") & !ends_with("_SD")) |>
+  estimate_density() |>
+  mutate(
+    Parameter = str_remove_all(Parameter, "IPIP6_"),
+    Parameter = str_replace(Parameter, "HonestyHumility", "Honesty-Humility")
+  ) |>
+  ggplot(aes(x = x, y = y, color = Parameter)) +
+  geom_line(size = 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  scale_color_manual(values = c("Agreeableness" = "#FFC107", "Honesty-Humility" = "#00BCD4", "Extraversion" = "#9C27B0", "Conscientiousness" = "#3F51B5", "Openness" = "#4CAF50", "Neuroticism" = "#E91E63")) +
+  labs(x = "Score", title = "Normal Personality", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(size = rel(1), face = "bold", hjust = 0.5),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = rel(0.8)),
+    axis.title.x = element_blank(),
+    legend.text = element_text(size = rel(0.8))
+  )
+
+p_pid <- dfsub |>
+  select(starts_with("PID") & !ends_with("_SD")) |>
+  estimate_density() |>
+  mutate(
+    Parameter = str_remove_all(Parameter, "PID5_"),
+    Parameter = str_replace(Parameter, "NegativeAffect", "Negative Affect")
+  ) |>
+  ggplot(aes(x = x, y = y, color = Parameter)) +
+  geom_line(size = 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  scale_color_manual(values = c("Antagonism" = "#FF9800", "Detachment" = "#03A9F4", "Disinhibition" = "#FF5722", "Negative Affect" = "#F44336", "Psychoticism" = "#673AB7")) +
+  labs(x = "Score", title = "Pathological Personality", color = NULL) +
+  theme_modern() +
+  theme(
+    plot.title = element_text(size = rel(1), face = "bold", hjust = 0.5),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = rel(0.8)),
+    axis.title.x = element_blank(),
+    legend.text = element_text(size = rel(0.8))
+    # legend.position = "top",
+  )
+
+# ggsave("figures/figure_dem3.png",
+#        patchwork::wrap_elements(p_ipip / p_pid), width=2/3*(fig.height), height=1*(fig.height), dpi=1000, bg="white")
+
+p_dem <- (
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem1.png"), interpolate = TRUE)) |
+    patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem2.png"), interpolate = TRUE)) |
+    patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem3.png"), interpolate = TRUE))
+) +
+  patchwork::plot_layout(widths = c(1 / 2, 1, 2 / 3))
+
+p_dem +
+  patchwork::plot_annotation(
+    title = "Participants (n = 250)",
+    theme = theme(plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0.5, vjust = -10))
+  )
+```
+```{r eval=FALSE}
+# Run this manually to avoid saving margins
+ggsave("figures/figure_dem.png", p_dem +
+  patchwork::plot_annotation(
+    title = "Participants (n = 250)",
+    theme = theme(plot.title = element_text(size = 50, face = "bold", hjust = 0.5, vjust = 0))
+  ), height = 7200, width = 15166, units = "px", limitsize = FALSE)
+```
+
+# Results
+
+## Contextual
 
 ```{r message=FALSE, warning=FALSE, cache=cache}
-preds <- data.frame()
-for(i in names(DVs)) {
-  model <- brms::brm(paste0(i, " ~ s(Screen_Size)"), data=dfsub, algorithm = "meanfield", refresh=0)
-  param <- bayestestR::describe_posterior(as.data.frame(model))
-
-  pred <- estimate_relation(model, length=50)
-  pred$Index <- i
-  pred$sig <- estimate_slopes(model, trend = "Screen_Size")$pd
-  preds <- rbind(preds, pred)
-
-  if(any(pred$sig > .95)) {
-    print(i)
-  }
+make_correlation <- function(x, y) {
+  cor <- correlation::correlation(x,
+    y,
+    bayesian = TRUE,
+    bayesian_prior = "medium.narrow",
+    sort = TRUE
+  ) |>
+    datawizard::data_remove(c("ROPE_Percentage"))
+  cor$`BF (Spearman)` <- format_bf(
+    correlation::correlation(
+      x, y,
+      bayesian = TRUE,
+      ranktransform = TRUE,
+      bayesian_prior = "medium.narrow"
+    )$BF,
+    name = NULL, stars = TRUE
+  )
+  cor |>
+    arrange(desc(BF))
 }
 
-preds |> 
-  mutate(sig = format_pd(sig, stars_only = TRUE),
-         sig = ifelse(sig == "", "NS", sig)) |> 
-  ggplot(aes(x = Screen_Size, y = Predicted)) +
-  geom_point2(data=dflong, aes(y = Score, color= Index), alpha=0.1, size=2) +
-  geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
+make_correlation(dfsub[names(DVs)], dfsub[c("Screen_Size", "Screen_Refresh")])
+
+dflong |>
+  ggplot(aes(x = Screen_Size, y = Score)) +
+  geom_point2(aes(color = Index), alpha = 0.1, size = 2) +
+  geom_smooth(aes(color = Index, group = Index, linetype = Index), se = FALSE, method = "lm") +
   scale_x_continuous(expand = c(0, 0)) +
-  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
-  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid")) +
+  scale_size_manual(values = c("NS" = 0.5, "*" = 1)) +
+  scale_linetype_manual(values = c("White" = "solid", "Contrast" = "solid", setNames(rep("dotted", 9), names(DVs)[!names(DVs) %in% c("White", "Contrast")])), guide = "none") +
   scale_color_manual(values = DVs) +
-  coord_cartesian(ylim = c(-2.5, 2.5)) +
   theme_modern() +
-  labs(y = "Illusion Sensitivity", x = "Screen Size") +
-  ggside::geom_xsidedensity(data=dfsub, fill ="grey", color = "white") +
-  ggside::geom_ysidedensity(data=dflong, aes(y=Score, color = Index)) +
-  ggside::theme_ggside_void() 
+  labs(y = "Illusion Sensitivity", x = "Screen Size", linetype = "Significance", size = "Significance") +
+  ggside::geom_xsidedensity(data = dfsub, fill = "grey", color = "white") +
+  ggside::geom_ysidedensity(aes(y = Score, color = Index)) +
+  ggside::theme_ggside_void() +
+  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
+  ggside::scale_xsidey_continuous(expand = c(0, 0))
 ```
 
-#### Refresh Rate
+## Demographics
 
-```{r message=FALSE, warning=FALSE, cache=cache}
-sig <- list()
-preds <- data.frame()
-for(i in DVs) {
-  model <- brms::brm(paste0(i, " ~ poly(Screen_Refresh, 2)"), data=dfsub, algorithm = "meanfield", refresh=0)
-  param <- bayestestR::describe_posterior(as.data.frame(model))
-
-  pred <- estimate_relation(model, length=50)
-  pred$Index <- i
-  pred$sig <- max(param$pd[2:3])
-  preds <- rbind(preds, pred)
-
-  if(any(max(param$pd[2:3]) > .95)) {
-    print(i)
-  }
-}
-
-preds |> 
-  mutate(sig = format_pd(sig, stars_only = TRUE),
-         sig = ifelse(sig == "", "NS", sig)) |> 
-  ggplot(aes(x = Screen_Refresh, y = Predicted)) +
-  geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
-  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
-  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
-```
-
-### Demographics
-
-#### Sex
+### Sex
 
@@
 params <- data.frame()
-for(i in DVs) {
-  model <- lm(paste0(i, " ~ Sex"), data=dfsub)
+for (i in names(DVs)) {
+  model <- lm(paste0(i, " ~ Sex"), data = dfsub)
   param <- parameters::parameters(model)
@@
   param$ymiddle <- param$Coefficient[1] + diff(param$Coefficient) / 2
+  param$BF <- parameters::parameters(BayesFactor::ttestBF(formula = as.formula(paste0(i, " ~ Sex")), data = dfsub[!is.na(dfsub[[i]]), ]))$BF
   params <- rbind(params, as.data.frame(param[2, ]))
-  if(param$p[2] < .05) {
-    print(i)
-  }
 }
@@
 data <- dfsub |>
-  pivot_longer(all_of(DVs), names_to = "Index", values_to = "Score") |> 
-  mutate(Index = fct_relevel(Index, arrange(params, p)$Index))
+  pivot_longer(all_of(names(DVs)), names_to = "Index", values_to = "Score") |>
+  mutate(
+    Index = fct_relevel(Index, names(DVs)),
+    Index = fct_relabel(Index, ~ prettify_itemName(.x))
+  )
 
-data |> 
+# parameters::parameters(BayesFactor::ttestBF(formula=Zollner ~ Sex, data=dfsub[!is.na(dfsub[[i]]), ]))
+
+p_gender <- data |>
   ggplot(aes(x = Index, y = Score)) +
-  stat_slab(data=filter(data, Sex == "Male"), aes(fill = Sex), side = "right", scale = 0.5, position = "dodge") +
-  stat_slab(data=filter(data, Sex == "Female"), aes(fill = Sex), side = "left", scale = 0.5, position = "dodge") +
-  stat_pointinterval(aes(group = Sex), point_interval = "mean_qi",, position = "dodge") +
-  geom_text(data = mutate(params, sig = insight::format_p(p, stars_only = TRUE)), aes(label = sig, y = ymiddle))
+  stat_slab(data = filter(data, Sex == "Male"), aes(fill = Sex), side = "left", scale = 0.5, position = "dodge") +
+  stat_slab(data = filter(data, Sex == "Female"), aes(fill = Sex), side = "right", scale = 0.5, position = "dodge") +
+  stat_pointinterval(aes(group = fct_rev(Sex)), point_interval = "mean_qi", position = "dodge") +
+  geom_text(data = mutate(params, sig = insight::format_bf(BF, stars_only = TRUE), Index = prettify_itemName(Index)), aes(label = sig, y = ymiddle), size = 6) +
+  geom_label(
+    data = mutate(params, label = insight::format_bf(BF, protect_ratio = TRUE), Index = prettify_itemName(Index)),
+    aes(label = label), y = 2.7, size = 3
+  ) +
+  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63")) +
+  scale_x_discrete(labels = function(x) ifelse(x == "i", "Factor i", x)) +
+  theme_minimal() +
+  labs(fill = "", x = "Sex", y = "Illusion Sensitivity", title = "Sex") +
+  theme(
+    legend.position = "top",
+    axis.title.x = element_blank(),
+    plot.title = element_text(size = rel(1), face = "bold", hjust = 0.5)
+  )
+
+p_gender
+# ggsave("figures/figure_sex.png", p_gender, width=15166, height=5000, bg="white", units = "px", limitsize = FALSE, scale = 0.3)
 ```
 
-#### Ethnicity
-
-```{r message=FALSE, warning=FALSE, cache=cache}
-sig <- list()
-params <- data.frame()
-for(i in DVs) {
-  model <- lm(paste0(i, " ~ Ethnicity"), data=dfsub)
-  param <- estimate_contrasts(model, contrast = "Ethnicity")
-  param$Index <- i
-  params <- rbind(params, as.data.frame(param[2, ]))
-  if(any(param$p < .05)) {
-    print(i)
-  }
-}
-```
-
-#### Education
+### Education
 
@@
 preds <- data.frame()
-for(i in DVs) {
-  model <- brms::brm(paste0(i, " ~ mo(as.numeric(Education))"), data=filter(dfsub, !Education %in% c("Other", "Prefer not to Say")), algorithm = "meanfield", refresh=0)
+for (i in names(DVs)) {
+  model <- brms::brm(paste0(i, " ~ mo(as.numeric(Education))"), data = filter(dfsub, !Education %in% c("Other", "Prefer not to Say")), refresh = 0)
   param <- bayestestR::describe_posterior(as.data.frame(model)["bsp_moas.numericEducation"])
@@
 preds |>
-  mutate(sig = format_pd(sig, stars_only = TRUE),
-         sig = ifelse(sig == "", "NS", sig)) |> 
+  mutate(
+    sig = format_pd(sig, stars_only = TRUE),
+    sig = ifelse(sig == "", "NS", sig)
+  ) |>
   ggplot(aes(x = Education, y = Predicted)) +
   geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
-  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
-  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
+  scale_size_manual(values = c("NS" = 0.5, "*" = 1)) +
+  scale_linetype_manual(values = c("NS" = "dotted", "*" = "solid")) +
+  theme_modern()
 ```
 
-#### Age
+### Age
 
 ```{r message=FALSE, warning=FALSE, cache=cache}
-sig <- list()
 preds <- data.frame()
-for(i in DVs) {
-  model <- brms::brm(paste0(i, " ~ poly(Age, 2)"), data=dfsub, algorithm = "meanfield", refresh=0)
-  param <- bayestestR::describe_posterior(as.data.frame(model))
+for (i in names(DVs)) {
+  gam <- brms::brm(paste0(i, " ~ s(Age, k=3)"), data = dfsub, refresh = 0, seed = 1)
+  model <- brms::brm(paste0(i, " ~ poly(Age, 2)"), data = dfsub, refresh = 0, seed = 1)
 
-  pred <- estimate_relation(model, length=50)
+  param <- parameters::parameters(model)
+  slopes <- estimate_slopes(gam, trend = "Age")
+
+  pred <- get_datagrid(gam, at = "Age", length = 50)
+  pred$Predicted <- get_predicted(gam, data = pred)
+
+  pred$ylabel <- pred$Predicted[50]
   pred$Index <- i
-  pred$sig <- max(param$pd[2:3])
+  pred$Effect <- slopes$Coefficient
+  pred$pd_gam <- slopes$pd
+  pred$pd_poly1 <- param$pd[2]
+  pred$pd_poly2 <- param$pd[3]
   preds <- rbind(preds, pred)
-
-  if(any(max(param$pd[2:3]) > .95)) {
-    print(i)
-  }
 }
@@
 preds |>
-  mutate(sig = format_pd(sig, stars_only = TRUE),
-         sig = ifelse(sig == "", "NS", sig)) |> 
+  group_by(Index) |>
+  slice(1) |>
+  select(-Predicted, -Age) |>
+  arrange(desc(pd_gam))
+
+p_age <- preds |>
+  mutate(
+    pd_gam = format_pd(pd_gam, stars_only = TRUE),
+    pd_gam = ifelse(pd_gam == "", "NS", pd_gam)
+  ) |>
   ggplot(aes(x = Age, y = Predicted)) +
-  geom_line(aes(color = Index, group = Index, linetype = sig, size = sig)) +
-  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1.5)) +
-  scale_linetype_manual(values = c("NS" = "dotted", "*" = "dashed", "**" = "solid"))
+  geom_line(aes(color = Index, group = Index, linetype = pd_gam, size = pd_gam)) +
+  ggrepel::geom_label_repel(
+    data = preds |>
+      group_by(Index) |>
+      slice(1) |>
+      mutate(label = prettify_itemName(ifelse(Index == "i", "Factor i", Index))),
+    aes(y = ylabel, label = label, x = 69, color = Index), hjust = 0, direction = "y",
+    min.segment.length = 3
+  ) +
+  scale_size_manual(values = c("NS" = 0.5, "*" = 1, "**" = 1, "***" = 1)) +
+  scale_linetype_manual(values = c("NS" = "dotted", "*" = "solid", "**" = "solid", "***" = "solid")) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_color_manual(values = DVs) +
+  guides(size = "none", linetype = "none", color = "none") +
+  theme_modern() +
+  labs(y = "Illusion Sensitivity", title = "Age") +
+  theme(
+    legend.position = "top",
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    plot.title = element_text(size = rel(1), face = "bold", hjust = 0.5)
+  )
+
+p_age
+
+ggsave("figures/figure_agesex.png",
+  (p_gender + p_age) + plot_layout(widths = c(2 / 3, 1 / 3)) +
+    patchwork::plot_annotation(
+      title = "Illusion Sensitivity and Demographic Variables",
+      theme = theme(plot.title = element_text(size = rel(1.5), face = "bold", hjust = 0.5, vjust = 0))
+    ),
+  width = 15166, height = 5000, bg = "white", units = "px", limitsize = FALSE, dpi = 800, scale = 1
+)
 ```
 
-### Personality
+## Personality
+
+### IPIP6
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
+cor <- make_correlation(
+  datawizard::data_select(dfsub, names(DVs)),
+  select(dfsub, starts_with("IPIP") & !ends_with("_SD"))
+)
+
+cor |>
+  data_filter(BF > 1) |>
+  arrange(desc(BF))
+
+cor |>
+  mutate(
+    label = paste0(format_value(rho, zap_small = TRUE), insight::format_bf(BF, stars_only = TRUE)),
+    Parameter2 = prettify_parameterName(Parameter2),
+    Parameter2 = str_remove_all(Parameter2, "IPIP6 - ")
+  ) |>
+  ggplot(aes(x = Parameter2, y = Parameter1)) +
+  geom_tile(aes(fill = rho)) +
+  geom_text(aes(label = label), size = 3) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  labs(y = "Illusion Sensitivity Scores", x = "Normal Personality (IPIP6)") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+p_ipip1 <- plot_correlation(dfsub, x = "i", y = "IPIP6_Agreeableness", fill = "#FFC107")
+p_ipip2 <- plot_correlation(dfsub, x = "i", y = "IPIP6_HonestyHumility", fill = "#00BCD4")
+
+(p_ipip1 + plot_correlation(dfsub, x = "VerticalHorizontal", y = "IPIP6_Agreeableness", fill = "#FFC107") + plot_correlation(dfsub, x = "Ponzo", y = "IPIP6_Agreeableness", fill = "#FFC107")) /
+  (p_ipip2 + plot_correlation(dfsub, x = "Zollner", y = "IPIP6_HonestyHumility", fill = "#00BCD4") + plot_correlation(dfsub, x = "VerticalHorizontal", y = "IPIP6_HonestyHumility", fill = "#00BCD4"))
+```
+
+### PID-5
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
+cor <- make_correlation(
+  datawizard::data_select(dfsub, names(DVs)),
+  select(dfsub, starts_with("PID") & !ends_with("_SD"))
+)
+
+cor |>
+  data_filter(BF > 1) |>
+  arrange(desc(BF))
+
+cor |>
+  mutate(label = paste0(format_value(rho, zap_small = TRUE), insight::format_bf(BF, stars_only = TRUE))) |>
+  ggplot(aes(x = Parameter2, y = Parameter1)) +
+  geom_tile(aes(fill = rho)) +
+  geom_text(aes(label = label), size = 3) +
+  scale_alpha_continuous(range = c(1, 0.5)) +
+  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+p_pid1 <- plot_correlation(dfsub, x = "i", y = "PID5_Antagonism", fill = "#FF9800")
+p_pid2 <- plot_correlation(dfsub, x = "i", y = "PID5_Disinhibition", fill = "#FF5722")
+p_pid3 <- plot_correlation(dfsub, x = "i", y = "PID5_Psychoticism", fill = "#673AB7")
+p_pid4 <- plot_correlation(dfsub, x = "i", y = "PID5_NegativeAffect", fill = "#F44336")
+
+(p_pid1 | plot_correlation(dfsub, x = "VerticalHorizontal", y = "PID5_Antagonism", fill = "#FF9800") | plot_correlation(dfsub, x = "Ponzo", y = "PID5_Antagonism", fill = "#FF9800")) / (p_pid2 | p_pid3 | p_pid4)
+
+p_pers <- (p_ipip1 +
+  theme(
+    axis.title.y = element_text(face = "bold"),
+    axis.title.x = element_blank()
+  )) /
+  (p_ipip2 +
+    theme(
+      axis.title.y = element_text(face = "bold"),
+      axis.title.x = element_blank()
+    )) |
+  (p_pid1 +
+    theme(
+      axis.title.y = element_text(face = "bold"),
+      axis.title.x = element_blank()
+    )) /
+    (p_pid2 +
+      labs(x = expression("Factor" ~ italic(i))) +
+      theme(
+        axis.title.y = element_text(face = "bold"),
+        axis.title.x = element_text(size = rel(1.5), face = "bold")
+      )) |
+  (p_pid3 +
+    theme(
+      axis.title.y = element_text(face = "bold"),
+      axis.title.x = element_blank()
+    )) /
+    (p_pid4 +
+      theme(
+        axis.title.y = element_text(face = "bold"),
+        axis.title.x = element_blank()
+      ))
+
+# ggsave("figures/figure_pers.png",
+#        p_pers + patchwork::plot_annotation(
+#          title = "Illusion Sensitivity and Personality",
+#          theme = theme(plot.title = element_text(size = rel(1.5), face="bold", hjust=0.5, vjust=0))),
+#        width=15166, height=8000, bg="white", units = "px", limitsize = FALSE, dpi=800, scale = 1)
+```
+
+## Table and Figure
 
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-r <- correlation::correlation(
-  datawizard::data_select(dfsub, DVs),
-  select(dfsub, starts_with("IPIP") & !ends_with("_SD")),
-  p_adjust = "none",
-  sort=TRUE
-)
+table1 <- ripip |>
+  rbind(rpid) |>
+  data_remove(c("ROPE_Percentage", "Prior_Distribution", "Prior_Location", "Prior_Scale", "pd")) |>
+  separate(Parameter2, "_", into = c("Scale", "Dimension")) |>
+  arrange(Scale, desc(BF)) |>
+  filter(BF > 3)
 
-r |>
-  mutate(label = paste0(format_value(r, zap_small = TRUE), insight::format_p(p, stars_only=TRUE))) |> 
-  ggplot(aes(x = Parameter2, y = Parameter1)) +
-  geom_tile(aes(fill = r)) +
-  geom_text(aes(label=label), size=3) +
-  scale_alpha_continuous(range = c(1, 0.5)) +
-  scale_fill_gradient2(low = "#2196F3", mid = "white", high = "#F44336", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation", guide = "legend") +
-  # facet_wrap(~Illusion_Type, scales = "free") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
+table1 |>
+  print_md()
+
+table1 |>
+  format_table() |>
+  select(-Method, -n_Obs) |>
+  rename(Factor = Parameter1) |>
+  write.csv("table1.csv", row.names = FALSE)
 ```
+
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
+p <- patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem.png"), interpolate = TRUE)) /
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_agesex.png"), interpolate = TRUE)) /
+  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_pers.png"), interpolate = TRUE)) +
+  plot_layout(heights = c(7200, 5000, 8000))
+
+ggsave("../figures/figure3.png", p, height = 20200, width = 15166, units = "px", limitsize = FALSE)
+```

### 2022-09-06 manuscript.Rmd
format

--- 2022-09-06 manuscript.Rmd
@@
 For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a small and strong increase of errors, respectively.
-Forthe Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
+For Delboeuf, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
 This was fixed in latest release of *Pyllusion* (v1.2), which now generate outlines of the same size as the target circle.
-For the Zöllner illusion, however, this observation is consistent with the findings of [@kitaoka2000], who reported an acute angle contraction effect at large angles (50 to 90 degrees) between the target horizontal line and the biasing horizontal bars.
-Although the Zollner illusion is commonly thought to be driven by an acute angle expansion mechanism that is optimal at small angles (below 30 degrees), an indirect illusory effect can be observed at sufficiently large angles due to the opposing acute angle contraction mechanism invoked [@kitaoka2007; @kitaoka2000a].
+For the Zöllner illusion, however, this observation is consistent with the findings of [@kitaoka2000], who reported an acute angle contraction effect at very small as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars.
+This underlying acute angle contraction mechanism could hence potentially explain the indirect illusory effect observed [@kitaoka2007; @kitaoka2000a].
 
@@
 
-Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
-Whereas some researchers report no relationship between inter-individual traits and inidividuals' susceptibility to illusions [@cretenoud2020a; @grzeczkowski2017], others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects [@zhang2017; @ohmann2016].
+Despite the wide interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
+Whereas some researchers report no relationship between inter-individual traits and individuals' susceptibility to illusions [@cretenoud2020a; @grzeczkowski2017], others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects [@zhang2017; @ohmann2016].
 Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal [@pessoa2008].

### 2022-09-06 manuscript.Rmd
Revert "format"
This reverts commit 11b12c954a47422c0fa46f988006c86bf6450300.

--- 2022-09-06 manuscript.Rmd
@@
 This was fixed in latest release of *Pyllusion* (v1.2), which now generate outlines of the same size as the target circle.
-For the Zöllner illusion, however, this observation is consistent with the findings of [@kitaoka2000], who reported an acute angle contraction effect at large angles (50 to 90 degrees) between the target horizontal line and the biasing horizontal bars.
-Although the Zollner illusion is commonly thought to be driven by an acute angle expansion mechanism that is optimal at small angles (below 30 degrees), an indirect illusory effect can be observed at sufficiently large angles due to the opposing acute angle contraction mechanism invoked [@kitaoka2007; @kitaoka2000a].
+For the Zöllner illusion, however, this observation is consistent with the findings of [@kitaoka2000], who reported an acute angle contraction effect at very small as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
+This invoked underlying acute angle contraction mechanism could hence potentially explain the indirect illusory effect observed [@kitaoka2007; @kitaoka2000a].
 

### 2022-09-06 manuscript.Rmd

--- 2022-09-06 manuscript.Rmd
@@
 Visual illusions are fascinating stimuli, capturing a key feature of our neurocognitive systems.
-They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but to integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
+They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
 Despite the historical and intensive interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022; @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
@@
 Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
-The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014] or **Insert some demographic / personality feature** [**REF**] - is another area of controversy.
+The existence of dispositional or demographic correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014]; as well as in males as compared to females [@miller2001; @papageorgiou2020; @lo2011; @shaqiri2018] - is another area of controversy.
 
@@
 
-This opens the door for an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can be either "incongruent" (making the task even harder by biasing the perception in the opposite way) or "congruent" (making the task easier).
+This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+Moreover, the illusion effect can be either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier).
 Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
@@
 We generated 56 stimuli for each of the 10 illusion types.
-These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where negative values correspond to congruent illusion strengths).
+These stimuli resulted from the combination of 8 linearly-spread levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where negative values correspond to congruent illusion strengths).
 
@@
 The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
-For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a small and strong increase of errors, respectively.
-For Delboeuf, we believe that it was an artifact caused by illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to small and strong increases of errors, respectively.
+For the Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
 This was fixed in latest release of *Pyllusion* (v1.2), which now generate outlines of the same size as the target circle.
-For Zöllner, however, we did not find a good explanation of the pattern.
-**TODO: is there some explanation that we can find in the literature?**
+For the Zöllner illusion, this observation is consistent with @kitaoka2000, in which they reported an acute angle contraction effect at very small as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
+This invoked underlying acute angle contraction mechanism could potentially explain the non-monotonic pattern found [@kitaoka2007; @kitaoka2000a].
 
@@
 
+Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
+Whereas some researchers report no relationship between inter-individual traits and individuals' susceptibility to illusions [@cretenoud2020a; @grzeczkowski2017], others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects [@zhang2017; @ohmann2016].
+Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal [@pessoa2008].
+
 # General Discussion
 
-Using the parametric illusion generation framework we developed, Pyllusion [@makowski2021a], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
+Using the parametric illusion generation framework we developed, *Pyllusion* [@makowski2021a], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
 This opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.

### 2022-09-06 manuscript.Rmd
manu

@@
     orcid         : 0000-0001-5375-9967
+    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+      - "Conceptualization"
+      - "Data curation"
+      - "Formal Analysis"
+      - "Funding acquisition"
+      - "Investigation"
+      - "Methodology"
+      - "Project administration"
+      - "Resources"
+      - "Software"
+      - "Supervision"
+      - "Validation"
+      - "Visualization"
+      - "Writing – original draft"
+  - name          : "An Shu Te"
+    affiliation   : "1"
+    orcid         : 0000-0001-8008-2824
+    role:
+      - "Project administration"
+      - "Resources"
+      - "Investigation"
+      - "Writing – original draft"
+  - name          : "Stephanie Kirk"
+    affiliation   : "1"
+    orcid         : 0000-0001-8008-2824
+    role:
+      - "Project administration"
+      - "Resources"
+      - "Writing – original draft"
+  - name          : "Ngoi Zi Liang"
+    affiliation   : "1"
+    orcid         : 0000-0001-8008-2824
+    role:
+      - "Project administration"
+      - "Resources"
+      - "Writing – original draft"
+  - name          : "S.H. Annabel Chen"
+    affiliation   : "1, 2, 3, 4"
+    orcid         : 0000-0002-1540-5516
+    role:
+      - "Project administration"
+      - "Supervision"
+      - "Writing – review & editing"
 
@@
     institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
+  - id            : "2"
+    institution   : "LKC Medicine, Nanyang Technological University, Singapore"
+  - id            : "3"
+    institution   : "National Institute of Education, Singapore"
+  - id            : "4"
+    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
+
 authornote: |
@@
 Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
-The existence of dispositional or demographic correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014]; as well as in males as compared to females [@miller2001; @papageorgiou2020; @lo2011; @shaqiri2018] - is another area of controversy.
+The existence of dispositional or demographic correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014]; as well as in males as compared to females [@miller2001; @papageorgiou2020; @lo2011; @shaqiri2018] (**Is there something for personality that we could use here instead of sex?**) - is another area of controversy.
 
@@
 ```{r message=FALSE, warning=FALSE, fig.cap="The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
-Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength."}
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.", out.width="100%"}
 knitr::include_graphics("figures/Figure1.png")
@@
 Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different [@lányi], or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016a; @grzeczkowski2018].
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016a; @grzeczkowski2018].
 Alternatively, *Pyllusion* allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
@@
 
-In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separated statistically.
+In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separately modeled statistically.
 Then, we will further utilize the paradigm to assess whether 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor.
@@
 
-Using the GAMs as a "ground-truth", we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
+Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
 Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we selected and fitted the best model (best on their indices of fit), and compared their output visually (see **Figure 2**).
@@
 This was fixed in latest release of *Pyllusion* (v1.2), which now generate outlines of the same size as the target circle.
-For the Zöllner illusion, this observation is consistent with @kitaoka2000, in which they reported an acute angle contraction effect at very small as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
-This invoked underlying acute angle contraction mechanism could potentially explain the non-monotonic pattern found [@kitaoka2007; @kitaoka2000a].
+For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports [@kitaoka2000; @kitaoka2007], suggesting an acute angle contraction effect at very small as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
 
 Finally, this study provided evidence for both the existence of illusion-specific factors, as well as for a common latent factor (labelled Factor *i*) that explained about half of the total variance.
-These participant-level scores were related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
+These participant-level scores were positively related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
 
@@
 Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
-**Insert a bit of litterature about why and some findings justifying what we did**.
+**TODO: Insert a bit of literature about why and some findings justifying what we did**.
 
@@
 
-```{r message=FALSE, warning=FALSE, fig.cap="CAPTION."}
+```{r message=FALSE, warning=FALSE, fig.cap="The upper plots show the distribution of demographic and dispositional variables.
+The middle plots show the relationship between illusion sensitivty scores, sex, and age (solid lines indicate significant relationships).
+Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \\textit{i}), and personality traits.", out.width="100%"}
 knitr::include_graphics("figures/Figure3.png")

### 2022-09-06 manuscript.tex
manu

@@
   pdftitle={The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates},
-  pdfauthor={Dominique Makowski1},
+  pdfauthor={Dominique Makowski1, An Shu Te1, Stephanie Kirk1, Ngoi Zi Liang1, \& S.H. Annabel Chen1, 2, 3, 4},
   pdflang={en-EN},
@@
 \title{\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates}}
-\author{Dominique Makowski\textsuperscript{1}}
+\author{Dominique Makowski\textsuperscript{1}, An Shu Te\textsuperscript{1}, Stephanie Kirk\textsuperscript{1}, Ngoi Zi Liang\textsuperscript{1}, \& S.H. Annabel Chen\textsuperscript{1, 2, 3, 4}}
 \date{}
@@
 
+The authors made the following contributions.
+Dominique Makowski: Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Software, Supervision, Validation, Visualization, Writing -- original draft; An Shu Te: Project administration, Resources, Investigation, Writing -- original draft; Stephanie Kirk: Project administration, Resources, Writing -- original draft; Ngoi Zi Liang: Project administration, Resources, Writing -- original draft; S.H.
+Annabel Chen: Project administration, Supervision, Writing -- review \& editing.
+
 Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore.
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Social Sciences, Nanyang Technological University, Singapore}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Social Sciences, Nanyang Technological University, Singapore\\\textsuperscript{2} LKC Medicine, Nanyang Technological University, Singapore\\\textsuperscript{3} National Institute of Education, Singapore\\\textsuperscript{4} Centre for Research and Development in Learning, Nanyang Technological University, Singapore}
 
@@
 Visual illusions are fascinating stimuli, capturing a key feature of our neurocognitive systems.
-They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but to integrate prior knowledge and contextual information - blended together in our subjective conscious experience (Carbon, 2014).
+They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience (Carbon, 2014).
 Despite the historical and intensive interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science (Caporuscio et al., 2022; Lamme, 2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
 
-Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud et al., 2020, 2019a; Hamburger, 2016; Teufel et al., 2018b); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019b; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018a) are strongly debated.
-The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022) or \textbf{Insert some demographic / personality feature} {[}\textbf{REF}{]} - is another area of controversy.
+Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud, Francis, et al., 2020; Cretenoud et al., 2019a; Hamburger, 2016; Teufel et al., 2018b); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019b; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018a) are strongly debated.
+The existence of dispositional or demographic correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022); as well as in males as compared to females (Lo \& Dinov, 2011; Miller, 2001; Papageorgiou et al., 2020; Shaqiri et al., 2018) (\textbf{Is there something for personality that we could use here instead of sex?}) - is another area of controversy.
 
@@
 \begin{figure}
-\includegraphics[width=4.2in]{figures/Figure1} \caption{The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+\includegraphics[width=1\linewidth]{figures/Figure1} \caption{The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
 Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.}\label{fig:unnamed-chunk-2}
@@
 Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (\emph{REF}), or having them adjust the targets to a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016a).
-Alternatively, \emph{Pyllusions} allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016a).
+Alternatively, \emph{Pyllusion} allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
-This opens the door for an experimental task in which participants have to make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can be either ``incongruent'' (making the task even harder by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
+This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+Moreover, the illusion effect can be either ``incongruent'' (making the task more difficult by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
 Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the ``objective'' effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
 
-In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separated statistically.
+In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separately modeled statistically.
 Then, we will further utilize the paradigm to assess whether 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor.
@@
 
-In line with total open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
+In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
 
@@
 We generated 56 stimuli for each of the 10 illusion types.
-These stimuli resulted from the combination of 8 linearly-spread levels levels of task difficulty (e.g., {[}1, 2, 3, 4, 5, 6, 7{]}, where 1 corresponds to the higher difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., {[}-3, -2, -1, 0, 1, 2, 3{]}, where negative values correspond to congruent illusion strengths).
+These stimuli resulted from the combination of 8 linearly-spread levels of task difficulty (e.g., {[}1, 2, 3, 4, 5, 6, 7{]}, where 1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., {[}-3, -2, -1, 0, 1, 2, 3{]}, where negative values correspond to congruent illusion strengths).
 
@@
 Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up \emph{vs.} down or left \emph{vs.} right) as an additional predictor.
-We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potentially non-linear relationships (at the expense of model simplicity).
+We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potential non-linear relationships (at the expense of model simplicity).
 
-The analysis was carried out using \emph{R 4.2} (R Core Team, 2022), \emph{brms} (Bürkner, 2017), the \emph{tidyverse} (Wickham et al., 2019), and the \emph{easystats} collection of packages (Lüdecke et al., 2021, 2019; Makowski et al., 2020, 2019).
+The analysis was carried out using \emph{R 4.2} (R Core Team, 2022), \emph{brms} (Bürkner, 2017), the \emph{tidyverse} (Wickham et al., 2019), and the \emph{easystats} collection of packages (Lüdecke et al., 2021, 2019; Makowski et al., 2020; Makowski, Ben-Shachar, \& Lüdecke, 2019).
 
@@
 
-Using the GAMs as a ``ground-truth'', we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
+Using GAMs as the ``ground-truth'' models, we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
 Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task \emph{difficulty} and illusion \emph{strength}), we selected and fitted the best model (best on their indices of fit), and compared their output visually (see \textbf{Figure 2}).
@@
 
-For errors, most of the models closely matched their GAMs counterpart (see \textbf{Figure 2}), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner, in which theoretically congruent illusion effects were related to increased error rate.
+For errors, most of the models closely matched their GAMs counterpart (see \textbf{Figure 2}), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
 
@@
 The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
-For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to a small and strong increase of errors, respectively.
-For Delboeuf, we believe that it was an artifact caused by illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to small and strong increases of errors, respectively.
+For the Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
 This was fixed in latest release of \emph{Pyllusion} (v1.2), which now generate outlines of the same size as the target circle.
-For Zöllner, however, we did not find a good explanation of the pattern.
-\textbf{TODO: is there some explanation that we can find in the literature?}
+For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports (Kitaoka, 2007; Kitaoka \& Ishihara, 2000), suggesting an acute angle contraction effect at very small as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
 
 Finally, this study provided evidence for both the existence of illusion-specific factors, as well as for a common latent factor (labelled Factor \emph{i}) that explained about half of the total variance.
-These participant-level scores were related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
+These participant-level scores were positively related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
 
@@
 Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
-\textbf{Insert a bit of litterature about why and some findings justifying what we did}.
+\textbf{TODO: Insert a bit of literature about why and some findings justifying what we did}.
 
@@
 \begin{figure}
-\includegraphics[width=9.33in]{figures/Figure3} \caption{CAPTION.}\label{fig:unnamed-chunk-4}
+\includegraphics[width=1\linewidth]{figures/Figure3} \caption{The upper plots show the distribution of demographic and dispositional variables.
+The middle plots show the relationship between illusion sensitivty scores, sex, and age (solid lines indicate significant relationships).
+Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \textit{i}), and personality traits.}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 
+For each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor \emph{i}), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) and personality.
+As the supplementary material contains the detailed results, we will here only report the significant results (based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see Makowski, Ben-Shachar, Chen, et al., 2019).
+
 \hypertarget{results-2}{%
 \subsection{Results}\label{results-2}}
+
+The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions.
+Anecdotal (\(BF_{10} = 2.05\)) and moderate evidence (\(BF_{10} = 4.11\)) was found for a negative correlation between screen size and the sensitivity to the White and the Contrast illusion, respectively.
+To test whether this result could be an artifact related to the highly skewed screen size distribution (caused by very few participants with extreme screen sizes), we re-ran a robust correlation (with rank-transformed values), which provided even stronger evidence in favor of the effect existence (\(BF_{10} = 28.19\), \(BF_{10} = 4.31\) for White and Contrast, respectively).
+
+The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females (\(\Delta=-0.37\), 95\% CI {[}-0.62, -0.13{]}, \(BF_{10} = 12.74\)).
+We fitted Bayesian linear models with the education level entered as a monotonic predictor (appropriate for ordinal variables, Bürkner \& Charpentier, 2020), which yielded no significant effects.
+For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
+These consistently suggested a significant positive linear relationship between age and Factor \emph{i} (\(pd=100\%\)), as well as the sensitivity to Müller-Lyer (\(pd=100\%\)), Vertical-Horizontal (\(pd=100\%\)), Zöllner (\(pd=100\%\)) and Ebbinghaus (\(pd=99\%\)) illusions.
+
+Regarding ``normal'' personality traits, Bayesian correlations suggested substantial evidence in favor of a positive relationship between \emph{Honesty-Humility} and Zöllner (\(BF_{10} > 100\)), Vertical-Horizontal (\(BF_{10} = 9.78\)) and the Factor \emph{i} (\(BF_{10} = 4.00\)); as well as between \emph{Agreeableness} and Vertical-Horizontal (\(BF_{10} = 25.06\)), Ponzo (\(BF_{10} = 4.88\)) and the Factor \emph{i} (\(BF_{10} = 19.65\)).
+
+Regarding ``pathological'' personality traits, the results yielded strong evidence in favor of a negative relationship between multiple illusion scores and multiple traits.
+\emph{Antagonism} was associated with the sensitivity to Vertical-Horizontal (\(BF_{10} > 100\)), Müller-Lyer (\(BF_{10} = 21.57\)), Ponzo (\(BF_{10} = 17.97\)) illusions, and the Factor \emph{i} (\(BF_{10} = 55.45\)); \emph{Psychoticism} was associated with the sensitivity to Vertical-Horizontal (\(BF_{10} = 66.63\)) and Müller-Lyer (\(BF_{10} = 35.59\)) illusions, and the Factor \emph{i} (\(BF_{10} = 35.02\)); \emph{Disinhibition} was associated with the sensitivity to Vertical-Horizontal (\(BF_{10} = 25.38\)), Zöllner (\(BF_{10} = 7.59\)), Müller-Lyer (\(BF_{10} = 5.89\)) illusions, and the Factor \emph{i} (\(BF_{10} = 31.42\)); and \emph{Negative Affect} was associated with Zöllner (\(BF_{10} = 62.04\)), Vertical-Horizontal (\(BF_{10} = 12.65\)), Müller-Lyer (\(BF_{10} = 3.17\)), and the Factor \emph{i} (\(BF_{10} = 6.39\)).
+The last remaining trait, \emph{Detachment}, did not share any relationship with illusion sensitivity.
 
@@
 
+Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
+Whereas some researchers report no relationship between inter-individual traits and individuals' susceptibility to illusions (Cretenoud, Grzeczkowski, et al., 2020; Grzeczkowski et al., 2017), others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects (Ohmann \& Burgmer, 2016; Zhang et al., 2017).
+Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal (Pessoa et al., 2008).
+
 \hypertarget{general-discussion}{%
@@
 
-Using the parametric illusion generation framework we developed, Pyllusion (Makowski et al., 2021b), we have hence shown that illusions can be manipulated continuously across several different visual illusions.
+Using the parametric illusion generation framework we developed, \emph{Pyllusion} (Makowski et al., 2021b), we have hence shown that illusions can be manipulated continuously across several different visual illusions.
 This opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.
@@
 In our paradigm, in which we apply this approach to a reaction-time task, we were able to measure inter-individual scores of objective illusion sensitivity.
+
+Notably, the general sensitivity to illusions Factor \emph{i} was negatively associated with \emph{Antagonism}, \emph{Psychoticism}, \emph{Disinhibition} and \emph{Negative Affect}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-burkner2020modelling}{}}%
+Bürkner, P.-C., \& Charpentier, E.
+(2020).
+Modelling monotonic effects of ordinal predictors in bayesian regression models.
+\emph{British Journal of Mathematical and Statistical Psychology}, \emph{73}(3), 420--451.
+
 \leavevmode\vadjust pre{\hypertarget{ref-caporuscio2022}{}}%
@@
 Cretenoud, A. F., Francis, G., \& Herzog, M. H. (2020). When illusions merge. \emph{Journal of Vision}, \emph{20}(8), 12. \url{https://doi.org/10.1167/jov.20.8.12}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020a}{}}%
+Cretenoud, A.
+F., Grzeczkowski, L., Bertamini, M., \& Herzog, M.
+H.
+(2020).
+Individual differences in the Müller-Lyer and Ponzo illusions are stable across different contexts.
+\emph{Journal of Vision}, \emph{20}(6), 4.
+\url{https://doi.org/10.1167/jov.20.6.4}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-grzeczkowski2017}{}}%
+Grzeczkowski, L., Clarke, A.
+M., Francis, G., Mast, F.
+W., \& Herzog, M.
+H.
+(2017).
+About individual differences in vision.
+\emph{Vision Research}, \emph{141}, 282--292.
+\url{https://doi.org/10.1016/j.visres.2016.10.006}
+
 \leavevmode\vadjust pre{\hypertarget{ref-grzeczkowski2018}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kitaoka2007}{}}%
+Kitaoka, A.
+(2007).
+Tilt illusions after Oyama (1960): A review {\textsuperscript{1}}: Tilt illusions: A review.
+\emph{Japanese Psychological Research}, \emph{49}(1), 7--19.
+\url{https://doi.org/10.1111/j.1468-5884.2007.00328.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kitaoka2000}{}}%
+Kitaoka, A., \& Ishihara, M. (2000). Three elemental illusions determine the zöllner illusion. \emph{Perception \& Psychophysics}, \emph{62}, 569--575.
+
 \leavevmode\vadjust pre{\hypertarget{ref-lamme2020}{}}%
 Lamme, V. A. F. (2020). Visual functions generating conscious seeing. \emph{Frontiers in Psychology}, \emph{11}. \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2020.00083}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lanyi2022can}{}}%
+Lányi, O., Keri, S., Pálffy, Z., \& Polner, B. (2022). \emph{Can you believe your eyes? Positive schizotypy is associated with increased susceptibility to the m{ü}ller-lyer illusion}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lo2011}{}}%
+Lo, K. C., \& Dinov, D. I. (2011). \emph{Investigation of Optical Illusions on the Aspects of Gender and Age}. \emph{24}, 8.
 
@@
 \url{https://doi.org/10.21105/joss.01412}
+
+\leavevmode\vadjust pre{\hypertarget{ref-makowski2019indices}{}}%
+Makowski, D., Ben-Shachar, M. S., Chen, S. A., \& Lüdecke, D. (2019). Indices of effect existence and significance in the bayesian framework. \emph{Frontiers in Psychology}, \emph{10}, 2767.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-miller2001}{}}%
+Miller, R.
+J.
+(2001).
+Gender differences in illusion response: The inffluence of spatial strategy and sex ratio.
+\emph{Sex Roles}, \emph{44}(3/4), 209--225.
+\url{https://doi.org/10.1023/A:1010907204430}
+
 \leavevmode\vadjust pre{\hypertarget{ref-mylniec2016}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ohmann2016}{}}%
+Ohmann, K., \& Burgmer, P.
+(2016).
+Nothing compares to me: How narcissism shapes comparative thinking.
+\emph{Personality and Individual Differences}, \emph{98}, 162--170.
+\url{https://doi.org/10.1016/j.paid.2016.03.069}
+
+\leavevmode\vadjust pre{\hypertarget{ref-papageorgiou2020}{}}%
+Papageorgiou, C., Stachtea, X., Papageorgiou, P., Alexandridis, A.
+T., Makris, G., Chrousos, G., \& Kosteletos, G.
+(2020).
+Gender-dependent variations in optical illusions: evidence from N400 waveforms.
+\emph{Physiological Measurement}, \emph{41}(9), 095006.
+\url{https://doi.org/10.1088/1361-6579/abb2eb}
+
 \leavevmode\vadjust pre{\hypertarget{ref-park2022}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pessoa2008}{}}%
+Pessoa, V.
+F., Monge-Fuentes, V., Simon, C.
+Y., Suganuma, E., \& Tavares, M.
+C.
+H.
+(2008).
+The miiller-lyer illusion as a tool for schizophrenia screening.
+\emph{Reviews in the Neurosciences}, \emph{19}(2-3).
+\url{https://doi.org/10.1515/REVNEURO.2008.19.2-3.91}
+
 \leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
@@
 \url{https://doi.org/10.1080/03004430.2020.1802260}
+
+\leavevmode\vadjust pre{\hypertarget{ref-shaqiri2018}{}}%
+Shaqiri, A., Roinishvili, M., Grzeczkowski, L., Chkonia, E., Pilz, K., Mohr, C., Brand, A., Kunchulia, M., \& Herzog, M.
+H.
+(2018).
+Sex-related differences in vision are heterogeneous.
+\emph{Scientific Reports}, \emph{8}(1), 7521.
+\url{https://doi.org/10.1038/s41598-018-25298-8}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-zhang2017}{}}%
+Zhang, Y., Liu, J., Wang, Y., Huang, J., Wei, L., Zhang, B., Wang, W., \& Chen, W.
+(2017).
+Personality traits and perception of Müller-Lyer illusion in male Chinese military soldiers and university students.
+\emph{Translational Neuroscience}, \emph{8}(1), 15--20.
+\url{https://doi.org/10.1515/tnsci-2017-0004}
+
 \end{CSLReferences}

### 2022-09-06 manuscript.Rmd
abstract

@@
 abstract: |
-  Abstract abstract. We invite researchers to re-analyze our open-access data to provide complimentary evidence (or absence thereof) on the effect and structure of illusion sensitivity.
+  Visual illusions strikingly highlight how the brain uses contextual and prior information to inform our perception of reality.
+Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, which ideally require a controlled and gradual modulation of the effects of interest.
+In this set of studies, we used the parametric framework for visual illusions implemented in the *Pyllusion* software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength.
+We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task.
+We then extracted the participant-level performance scores (n=250), and provide evidence in favour of the existence of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions.
+Finally, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Negative Affect and Disinhibition.
+All the materials are available in open-access (https://github.com/RealityBending/IllusionGameValidation).
+We invite researchers to re-analyze the data using alternative approaches to provide complimentary findings on the effect, structure and correlates, of illusion sensitivity.
 
-keywords          : "visual illusions, illusion game, pyllusion, illusion effect, personality"
+keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
 wordcount         : "1156"

### 2022-09-06 manuscript.tex
abstract

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, pyllusion, illusion effect, personality\newline\indent Word count: 1156}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 1156}
 \usepackage{lineno}
@@
   pdflang={en-EN},
-  pdfkeywords={visual illusions, illusion game, pyllusion, illusion effect, personality},
+  pdfkeywords={visual illusions, illusion game, Pyllusion, personality, general factor},
   hidelinks,
@@
 \abstract{%
-Abstract abstract. We invite researchers to re-analyze our open-access data to provide complimentary evidence (or absence thereof) on the effect and structure of illusion sensitivity.
+Visual illusions strikingly highlight how the brain uses contextual and prior information to inform our perception of reality.
+Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, which ideally require a controlled and gradual modulation of the effects of interest.
+In this set of studies, we used the parametric framework for visual illusions implemented in the \emph{Pyllusion} software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength.
+We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task.
+We then extracted the participant-level performance scores (n=250), and provide evidence in favour of the existence of a general factor (labelled Factor \emph{i}) underlying the sensitivity to different illusions.
+Finally, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Negative Affect and Disinhibition.
+All the materials are available in open-access (\url{https://github.com/RealityBending/IllusionGameValidation}).
+We invite researchers to re-analyze the data using alternative approaches to provide complimentary findings on the effect, structure and correlates, of illusion sensitivity.
 }

### 2022-09-06 manuscript.Rmd
duplcate ref

@@
 
-Using the parametric illusion generation framework we developed, *Pyllusion* [@makowski2021a], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
+Using the parametric illusion generation framework developed with *Pyllusion* [@makowski2021], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
 This opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.

### 2022-09-06 manuscript.Rmd
update manu

@@
 Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
-The existence of dispositional or demographic correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014]; as well as in males as compared to females [@miller2001; @papageorgiou2020; @lo2011; @shaqiri2018] (**Is there something for personality that we could use here instead of sex?**) - is another area of controversy.
+The existence of dispositional or demographic correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014]; as well as for individuals with lower trait anxiety [@zhang2017; @qadri2020; @baroun2005] - is another area of controversy.
 
@@
 Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
-**TODO: Insert a bit of literature about why and some findings justifying what we did**.
+Despite the abundance of literature on visual illusions, relatively few researchers have attempted to examine the ties between demographic variables and personality traits with regard to individuals' resistance to illusions.
+Furthermore, most studies that focused on personality differences in illusion response primarily investigated the effects of traits associated with schizophrenia and autism spectrum disorders (such as impulsivity and sensation-seeking etc; [@hlavatá2018; @zhang2017; @lanyi2022can; @pessoa2008]).
+In this study, we instead aim to further elucidate inter-individual susceptibility to illusory effects by delineating the relationship between illusion susceptibility and demographic (i.e., age and gender) and dispositional components.
 
@@
 
-The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females ($\Delta=-0.37$, 95\% CI [-0.62, -0.13], $BF_{10} = 12.74$).
+The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females ($\Delta=-0.37$, 95% CI [-0.62, -0.13], $BF_{10} = 12.74$).
 We fitted Bayesian linear models with the education level entered as a monotonic predictor [appropriate for ordinal variables, @burkner2020modelling], which yielded no significant effects.
@@
 
-Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
-Whereas some researchers report no relationship between inter-individual traits and individuals' susceptibility to illusions [@cretenoud2020a; @grzeczkowski2017], others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects [@zhang2017; @ohmann2016].
-Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal [@pessoa2008].
+Interestingly, screen size was found to have a significant negative effect on illusion sensitivity for the White and Contrast illusions.
+This could be a result of our visual systems filtering through more low spatial frequency content when the size of the target object is small [@dixon2014].
+As this filtering process excludes illumination information from visual processing, smaller screen sizes could result in an unprecedented change in brightness perception.
+As such, this could attenuate the illusory effect of the White and Contrast illusions, decreasing participants' susceptibility to the illusions since the illusory effects of both are dependent on luminance.
+
+Although the present literature regarding the effects of personality correlates on illusion susceptibility remains fairly mixed [@cretenoud2020a; @grzeczkowski2017; @zhang2017; @ohmann2016], this study provides evidence of a positive relationship between illusion sensitivity and multiple personality traits, such as *Honesty-Humility* and *Agreeableness,* in healthy populations.
+Notably, higher scores in "pathological" traits such as *Antagonism, Psychoticism, Disinhibition* as well as *Negative Affect* were strongly correlated with lower illusion sensitivity (Factor *i*).
+This observation is consistent with past studies, in which higher self-reported Aggression-Hostility and Narcissism scores were associated with lower susceptibility to illusions [@zhang2017; @ohmann2016].
+Moreover, low aggression and high self-control have also been found to correlate strongly with greater resistance to illusions [@mcgurk], conferring further support for the results we observed.
+As such, while the effects of personality on illusion sensitivity have typically only been found among clinical populations (usually in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019]), this study demonstrates that differences in personality traits can also lead to inter-individual variation in illusion susceptibility.
+Further studies are required to replicate the observed findings and validate the underlying effects of personality dimensions on illusion susceptibility.
 

### 2022-09-07 manuscript.Rmd
update

--- 2022-09-06 manuscript.Rmd
@@
     affiliation   : "1"
-    orcid         : 0000-0001-8008-2824
+    orcid         : 0000-0002-9312-5552
     role:
@@
     affiliation   : "1"
-    orcid         : 0000-0001-8008-2824
     role:
@@
       - "Resources"
-      - "Writing – original draft"
+      - "Writing – review & editing"
   - name          : "S.H. Annabel Chen"
@@
 
-Using the parametric illusion generation framework developed with *Pyllusion* [@makowski2021], we have hence shown that illusions can be manipulated continuously across several different visual illusions.
-This opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.
-The validation of this novel framework also affords future illusion scientists a standardized measure of illusion susceptibility, instead of relying on conventional methods that depend upon participants' subjective perceptions.
-In our paradigm, in which we apply this approach to a reaction-time task, we were able to measure inter-individual scores of objective illusion sensitivity.
+<!-- Parametric framework -->
+The parametric illusion generation framework developed that we have developed in @makowski2021 proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
+In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
+This important methodological step opens the door for new illusions-based paradigms and tasks, to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
+
+<!-- Factor i -->
+Our findings suggested that the participants' sensitivities to 10 different types of visual illusions shared a common part of variance, suggesting the existence of a general factor of illusion sensitivity (Factor *i*).
+This result comes in a field of mixed findings.
+**TODO: discuss some papers investigating that**.
+
+Furthermore, although the illusions were fairly different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), The possibility of our general factor being driven by inter-individual perceptual skills variability cannot be discarded.
+Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity.
+
+<!-- Personality -->
 
@@
 
+<!-- Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
+Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder [ASD]) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
+-->
+
 <!-- Dump  -->
 
-Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions [@mylniec2016].
+<!-- Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions [@mylniec2016].
 For instance, while some researchers have tried to explain our sensitivity to illusory effects as a result of deficits in the low-level visual processing system [@cretenoud2019; @gori2016], others have provided a compelling case using a top-down approach, suggesting that such visual phenomena occur as a result of a conflict between our visual input and our prior beliefs[@teufel2018; @caporuscio2022].
-
-Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
-Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder [ASD]) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
+-->
 
 # Future Directions
+
+Future study should investigate the psychometric properties of similar paradigms, including stability (e.g., test-retest reliability), consistency (in particular, the relationship with perceptual abilities), and validity.
 

### 2022-09-07 study2/study2.Rmd
update

@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 # Save data

### 2022-09-07 study2/make_shortversion.R

@@ -1,0 +1,45 @@
+library(tidyverse)
+library(easystats)
+
+
+for(i in c("delboeuf", "ebbinghaus", "rodframe", "verticalhorizontal", "zollner", "white", "mullerlyer", "ponzo", "poggendorff", "contrast")) {
+  load(paste0("models/gam_", i, "_err.RData"))
+  load(paste0("models/gam_", i, "_rt.RData"))
+}
+
+
+
+
+make_plot <- function(gam_err, gam_rt, diff, strength){
+  prederr <- estimate_relation(gam_err, at=c(diff, "Illusion_Strength"), length=c(5, 100))
+  predrt <- estimate_relation(gam_rt, at=c(diff, "Illusion_Strength"), length=c(5, 100)) |>
+    mutate(Predicted = datawizard::rescale(Predicted, to=c(min(prederr$Predicted), max(prederr$Predicted))))
+
+  ggplot(prederr, aes(x = Illusion_Strength, y=Predicted)) +
+    geom_hline(yintercept=0.2, linetype="dashed") +
+    geom_vline(xintercept=strength, linetype="dotted") +
+    geom_line(data=predrt, aes(color=Illusion_Difference, group=Illusion_Difference), linetype="dashed", size=1) +
+    geom_line(aes(color=Illusion_Difference, group=Illusion_Difference), size=1)
+
+}
+
+# Vertical Horizontal
+make_plot(gam_verticalhorizontal_err, gam_verticalhorizontal_rt,
+          diff="Illusion_Difference=c(0.05, 0.225)",
+          strength=c(-33, 33))
+
+# M??ller-Lyer
+make_plot(gam_mullerlyer_err, gam_mullerlyer_rt,
+          diff="Illusion_Difference=c(0.05, 0.35)",
+          strength=c(-23, 23))
+
+# Ebbinghaus
+make_plot(gam_ebbinghaus_err, gam_ebbinghaus_rt,
+          diff="Illusion_Difference=c(0.09, 0.28)",
+          strength=c(-1, 1))
+
+# Poggendorff
+summary(get_data(gam_poggendorff_err)$Illusion_Difference)
+make_plot(gam_poggendorff_err, gam_poggendorff_rt,
+          diff="Illusion_Difference=c(0.002, 0.17)",
+          strength=c(-28, 28))

### 2022-09-07 manuscript.Rmd

@@
 wordcount         : "1156"
-bibliography      : "references.bib"
+bibliography      : references.bib
 floatsintext      : yes
@@
 Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
-The existence of dispositional or demographic correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014]; as well as in males as compared to females [@miller2001; @papageorgiou2020; @lo2011; @shaqiri2018] (**Is there something for personality that we could use here instead of sex?**) - is another area of controversy.
+The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], as well as in individuals with stronger traits of aggression and narcissism [@zhang2017; @konrath2009seeing] - is another area of controversy.
 
@@
 To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
-This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus) with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (see **Figure 1**).
+This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus, ...) with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (see **Figure 1**).
 
@@
 Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
-**TODO: Insert a bit of literature about why and some findings justifying what we did**.
+Despite the abundant literature on visual illusions, relatively few studies investigate the ties with participants' characteristics.
+**TODO: HERE mention studies on sex or age**.
+Furthermore, most works assessing the links with personality focused on traits associated with psychopathology [such as impulsivity or sensation-seeking, @hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
+In this study, we instead propose to explore the relationship between illusion sensitivity, demographic variables and stable "general" personality traits.
 
@@
 
-Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
-Whereas some researchers report no relationship between inter-individual traits and individuals' susceptibility to illusions [@cretenoud2020a; @grzeczkowski2017], others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects [@zhang2017; @ohmann2016].
-Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal [@pessoa2008].
+<!-- Context -->
+We report significant links between inter-individual indices of illusion sensitivity and varialbes related to experimental context, demographic characteristics and personality.
+Firstly, screen size was found to have a significant negative relationship with the sensitivity to the two contrast-based illusions, namely the White and Contrast illusions.
+One possible explanation can be found in the mechanism by which visual systems filter through more low spatial frequencies when the size of the target object is small [@dixon2014].
+As this filtering process excludes illumination information from visual processing, smaller screen sizes could yield artifactual changes in brightness perception, which in turn could attenuate the illusory effect of luminance-related illusions.
+
+<!-- Demographic -->
+**TODO: talk about absence of effect on sex + effect on age**
+
+<!-- Normal personality -->
+Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*.
+Although the past literature regarding the links between illusion sensitivity and Big 5 or HEXACO personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing].
+**TODO: do we have any leads of explanations?**
+
+<!-- Pathological personality -->
+The positive relationship between illusion sensitivity and "positive" personality traits is mirrored by a negative relationship with several pathological traits, including *Antagonism*, *Psychoticism*, *Disinhibition*, and *Negative Affect*.
+These results are in general consistent with the past findings and theories, suggesting that ...**TODO: do we have a theory?**
+
+Our results, suggesting the existence of a link between illusion sensitivity and maladaptive personality traits in non-clinical populations, could be put in relation with clinical findings, which could be extreme cases where the relationship with illusion sensitivity is the most manifest.
+In line with our results on *Psychoticism* or *Disinhibition*, prior research has found greater illusion resistance in schizophrenia [**REF**], and in particular in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019].
+
+<!-- The results reported by past studies on the relationship between illusion sensitivity and personality remain fairly mixed [@cretenoud2020a; @grzeczkowski2017].  -->
+
+<!-- Moreover, low aggression and high self-control have also been found to correlate strongly with greater resistance to illusions [@mcgurk]. -->
 
@@
 Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder [ASD]) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
+-->
+
+<!-- Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
+Whereas some researchers report no relationship between inter-individual traits and individuals' susceptibility to illusions [@cretenoud2020a; @grzeczkowski2017], others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects [@zhang2017; @konrath2009seeing].
+Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal [@pessoa2008].
 -->

### 2022-09-07 manuscript.Rmd
remove duplicate ref

@@
 
-Visual illusions are fascinating stimuli, capturing a key feature of our neurocognitive systems.
+Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems.
 They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
@@
 Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
-The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], as well as in individuals with stronger traits of aggression and narcissism [@zhang2017; @konrath2009seeing] - is another area of controversy.
+The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], as well as in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing] - is another area of controversy.
 
@@
 Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016a; @grzeczkowski2018].
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016; @grzeczkowski2018].
 Alternatively, *Pyllusion* allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).

### 2022-09-07 manuscript.Rmd
update

@@
 
-The paradigm of study 2 was similar to that of study 1, with the following changes.
+The paradigm of study 2 was similar to that of study 1, with the following changes made.
 The illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
@@
 Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
-Despite the abundant literature on visual illusions, relatively few studies investigate the ties with participants' characteristics.
-**TODO: HERE mention studies on sex or age**.
-Furthermore, most works assessing the links with personality focused on traits associated with psychopathology [such as impulsivity or sensation-seeking, @hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
+Despite the abundant literature on visual illusions, relatively few studies have investigated its ties with participants' characteristics.
+Moreover, research examining the influence of demographic variables tend to be centered on gender and age-related differences, and have generally found inconsistent results [@cretenoud2020a; @grzeczkowski2017; @lo2011; @papageorgiou2020].
+Alternatively, among works assessing the links with personality, most focused on traits associated with psychopathology [such as impulsivity or sensation-seeking, @hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
 In this study, we instead propose to explore the relationship between illusion sensitivity, demographic variables and stable "general" personality traits.
@@
 <!-- Context -->
-We report significant links between inter-individual indices of illusion sensitivity and varialbes related to experimental context, demographic characteristics and personality.
+
+We report significant links between inter-individual indices of illusion sensitivity and variables related to experimental context, demographic characteristics and personality.
 Firstly, screen size was found to have a significant negative relationship with the sensitivity to the two contrast-based illusions, namely the White and Contrast illusions.
@@
 <!-- Demographic -->
-**TODO: talk about absence of effect on sex + effect on age**
+
+Contrary to past studies in which females were found to exhibit greater illusion sensitivity [@lo2011; @miller2001; @papageorgiou2020], our study finds males to be more sensitive to the effects of the Zollner illusion.
+This inconsistency in findings could be due to past results conflating the objective illusory effect with the subjective perceptual skills needed to evaluate the visual illusions.
+Indeed, gender-related differences in visuospatial strategies have often been cited as the underlying mechanism explaining greater illusion sensitivity among females [@miller2001; @lo2011].
+In this study, however, individuals' perceptual abilities and the objective illusory effect were independently modulated using our *Pyllusion* software.
+This thus allowed us to capture the variability related only to the latter, thereby mitigating the effects of subjective perceptual skills.
+As such, the greater illusion sensitivity among males observed (albeit only for the Zollner illusion) could be a more veridical reflection of the illusion effect.
 
 <!-- Normal personality -->
+
 Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*.
-Although the past literature regarding the links between illusion sensitivity and Big 5 or HEXACO personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing].
-**TODO: do we have any leads of explanations?**
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing].
+One possible explanation for this association could be drawn from the literature on field dependence, which states that individuals with a field dependent cognitive style are more reliant on external cues when faced with ambiguity [@witkin1976].
+Since personality traits such as narcissism and aggression have been correlated with relative field independence [@ohmann2016; @witkin1976a; @damour2021], this implies that individuals with high levels of such traits may be less reliant on biasing visual cues and thus display less illusion sensitivity.
 
 <!-- Pathological personality -->
+
 The positive relationship between illusion sensitivity and "positive" personality traits is mirrored by a negative relationship with several pathological traits, including *Antagonism*, *Psychoticism*, *Disinhibition*, and *Negative Affect*.
-These results are in general consistent with the past findings and theories, suggesting that ...**TODO: do we have a theory?**
+These results are, in general, consistent with past findings and theories, suggesting a potential negative correlation between egocentric cognitions and context processing.
+Specifically, pathological egocentric beliefs, which is indexed by high levels of self-focus (and low levels of other-focus) and have been related to maladaptive personality traits like *Psychoticism,* could be associated with a reduced consideration for context integration, such as a tendency to disembed objects from their surroundings when processing visual stimuli [@ohmann2016; @konrath2009; @fox2006].
+This could hence explain the positive relationship between illusion sensitivity and pathological traits observed in our study.
+Indeed, high levels of self-focus have been found to correlate with lower illusion sensitivity [@konrath2009], conferring support for the postulated links between egocentrism and decontextualized information processing.
 
-Our results, suggesting the existence of a link between illusion sensitivity and maladaptive personality traits in non-clinical populations, could be put in relation with clinical findings, which could be extreme cases where the relationship with illusion sensitivity is the most manifest.
-In line with our results on *Psychoticism* or *Disinhibition*, prior research has found greater illusion resistance in schizophrenia [**REF**], and in particular in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019].
+Furthermore, our results, which suggests the existence of a link between illusion sensitivity and maladaptive personality traits in non-clinical populations, could be put in relation with clinical findings, which could be seen as extreme cases where the relationship with illusion sensitivity most manifests.
+In line with our results on *Psychoticism* or *Disinhibition*, prior research has found greater illusion resistance in schizophrenia [@notredame2014; @pessoa2008; @grzeczkowski2018], and in particular, in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019; @lanyi2022can].
 
@@
 <!-- Parametric framework -->
-The parametric illusion generation framework developed that we have developed in @makowski2021 proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
+
+The parametric illusion generation framework that we have developed in @makowski2021 proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
@@
 <!-- Factor i -->
+
 Our findings suggested that the participants' sensitivities to 10 different types of visual illusions shared a common part of variance, suggesting the existence of a general factor of illusion sensitivity (Factor *i*).
 This result comes in a field of mixed findings.
-**TODO: discuss some papers investigating that**.
+In fact, contrary to early studies on visual illusions, recent research (over the past 10 years) have generally not found any significant evidence for a common stable factor across illusions within individuals [@grzeczkowski2017; @cretenoud2019; @cretenoud2020; @grzeczkowski2018; @yang2012].
+Instead, past findings suggests illusory effects are highly specific to the features of different illusions, albeit some attempts at grouping certain illusions in hopes of extracting a common generalizable factor have been made [@grzeczkowski2017; @cretenoud2019].
+However, the absence of significant results could have been due to the reliance of conventional paradigms, in which participants were instructed to adjust targets to be aligned with a reference stimulus.
+As such paradigms depend upon participants' subjective perceptual skills, our study hence provides a possible objective framework that could culminate in a renewed approach into the field of visual illusions.
 
-Furthermore, although the illusions were fairly different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), The possibility of our general factor being driven by inter-individual perceptual skills variability cannot be discarded.
+Furthermore, although the illusions we assessed were fairly different in terms of the perceptual task involved (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability cannot be discarded.
 Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity.

### 2022-09-07 manuscript.Rmd
abstract

--- 2022-09-07 manuscript.Rmd
@@
 In this set of studies, we used the parametric framework for visual illusions implemented in the *Pyllusion* software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength.
-We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task.
-We then extracted the participant-level performance scores (n=250), and provide evidence in favour of the existence of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions.
-Finally, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Negative Affect and Disinhibition.
+We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task, from which we extracted participant-level performance scores (n=250).
+Our results provide evidence in favour of the existence of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions.
+Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
 All the materials are available in open-access (https://github.com/RealityBending/IllusionGameValidation).

### 2022-09-07 manuscript.Rmd
update

@@
 
-The paradigm of study 2 was similar to that of study 1, with the following changes.
-The illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
+The paradigm of study 2 was similar to that of study 1, with the following changes: the illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
 Moreover, taking into account the findings of study 1, we used non-linearly spaced difficulty levels, depending on the best underlying model (i.e., with an exponential, square or cubic spacing depending on the relationship).
@@
 
-Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
-Despite the abundant literature on visual illusions, relatively few studies investigate the ties with participants' characteristics.
-**TODO: HERE mention studies on sex or age**.
-Furthermore, most works assessing the links with personality focused on traits associated with psychopathology [such as impulsivity or sensation-seeking, @hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
-In this study, we instead propose to explore the relationship between illusion sensitivity, demographic variables and stable "general" personality traits.
+Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), contextual variables (pertaining to the experiment setting), such as screen size, demographic features (such as sex and age), and stable dispositional variables such as "general" personality traits.
+Indeed, despite the abundant literature on visual illusions, relatively few studies have investigated its ties with participants' characteristics.
+Research examining the influence of demographic variables such as gender and age have generally found inconsistent results [@cretenoud2020a; @grzeczkowski2017; @lo2011; @papageorgiou2020].
+Regarding links with personality, most works focused on traits associated with psychopathology, such as impulsivity or sensation-seeking [@hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
 
@@
 <!-- Demographic -->
-**TODO: talk about absence of effect on sex + effect on age**
+Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females reported to the Zöllner illusion.
+As we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity [@lo2011; @miller2001; @papageorgiou2020].
+This inconsistency could be due to past studies using a measure of illusion sensitvity that conflates the effect of illusions *per se* with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
+On the contrary, the perceptual difficulty of the task and the illusion effect was independently modulated in our paradigm, and statistically dissociated.
+Our scores of illusion sensitivity might thus be less loaded with perceptual skills, thereby mitigating its effect.
 
-<!-- Normal personality -->
-Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*.
-Although the past literature regarding the links between illusion sensitivity and Big 5 or HEXACO personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing].
-**TODO: do we have any leads of explanations?**
+<!-- Normal personality + Antagonism -->
+Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*, and a negative link with *Antagonism*.
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing].
+While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
+Since narcissism and aggression tendencies are correlated with lower field dependence [i.e., a lesser reliance on external cues in ambiguous contexts, @witkin1976; @ohmann2016; @damour2021], opposite traits, such as *Honesty-Humility* and *Agreeableness*, could conversely be more biased by contextual cues and thus more sensitive to illusions.
 
 <!-- Pathological personality -->
-The positive relationship between illusion sensitivity and "positive" personality traits is mirrored by a negative relationship with several pathological traits, including *Antagonism*, *Psychoticism*, *Disinhibition*, and *Negative Affect*.
-These results are in general consistent with the past findings and theories, suggesting that ...**TODO: do we have a theory?**
+The positive relationship between illusion sensitivity and "positive" personality traits is mirrored by a negative relationship with several other pathological traits, including *Psychoticism*, *Disinhibition*, and *Negative Affect*.
+These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing [including illusion sensitivity, @konrath2009].
+For instance, pathological egocentric beliefs [often observed alongside *Psychoticism*, @fox2006] have been related to reduced context integration [manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, @ohmann2016; @konrath2009; @fox2006].
+As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits [**REF**].
 
-Our results, suggesting the existence of a link between illusion sensitivity and maladaptive personality traits in non-clinical populations, could be put in relation with clinical findings, which could be extreme cases where the relationship with illusion sensitivity is the most manifest.
-In line with our results on *Psychoticism* or *Disinhibition*, prior research has found greater illusion resistance in schizophrenia [**REF**], and in particular in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019].
+Furthermore, these results in favour of a link between illusion sensitivity and maladaptive personality traits in a non-clinical population could be put in relation with clinical findings, which could be seen as extreme cases where the relationship with illusion sensitivity is the most manifest.
+In line with our results (in particular on *Psychoticism* and *Disinhibition*), prior research has found greater illusion resistance in schizophrenia [@notredame2014; @pessoa2008; @grzeczkowski2018], and in particular, in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019; @lanyi2022can].
 
@@
 <!-- Parametric framework -->
-The parametric illusion generation framework developed that we have developed in @makowski2021 proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
+The parametric illusion generation framework developed in @makowski2021 proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
@@
 <!-- Factor i -->
-Our findings suggested that the participants' sensitivities to 10 different types of visual illusions shared a common part of variance, suggesting the existence of a general factor of illusion sensitivity (Factor *i*).
+The participants' sensitivities to 10 different types of visual illusions shared a common part of variance, suggesting the existence of a general factor of illusion sensitivity (Factor *i*).
 This result comes in a field of mixed findings.
-**TODO: discuss some papers investigating that**.
+In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals [@grzeczkowski2017; @cretenoud2019; @cretenoud2020; @grzeczkowski2018; @yang2012].
+Instead, past findings suggests illusory effects are highly specific to the perceptual features of the illusions at stake [@grzeczkowski2017; @cretenoud2019].
+It is to note, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
+We believe that our study presents several methodological improvements, including statistical power (high number of trials per participant), homogeneous stimuli (with minimal and highly controlled features) and tasks (decision-making reaction-time task), and more reliable participant-level score extraction method (based on random-factors models), which in our opinion contributed to the emergence of the common factor.
 
-Furthermore, although the illusions were fairly different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), The possibility of our general factor being driven by inter-individual perceptual skills variability cannot be discarded.
-Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity.
+However, although the illusions were relatively different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), The possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity, and assess the psychometric properties of similar paradigms, including stability (e.g., test-retest reliability) and validity.
 
@@
 
-Notably, the general sensitivity to illusions Factor *i* was negatively associated with *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
+Finally, we found that the general sensitivity to illusions Factor *i* was negatively associated with *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
 
@@
 
-# Future Directions
-
-Future study should investigate the psychometric properties of similar paradigms, including stability (e.g., test-retest reliability), consistency (in particular, the relationship with perceptual abilities), and validity.
-
-We strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+In conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
 The task, data and analysis script are available in open-access at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.

### 2022-09-07 manuscript.tex
update

@@
 The authors made the following contributions.
-Dominique Makowski: Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Software, Supervision, Validation, Visualization, Writing -- original draft; An Shu Te: Project administration, Resources, Investigation, Writing -- original draft; Stephanie Kirk: Project administration, Resources, Writing -- original draft; Ngoi Zi Liang: Project administration, Resources, Writing -- original draft; S.H.
+Dominique Makowski: Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Software, Supervision, Validation, Visualization, Writing -- original draft; An Shu Te: Project administration, Resources, Investigation, Writing -- original draft; Stephanie Kirk: Project administration, Resources, Writing -- original draft; Ngoi Zi Liang: Project administration, Resources, Writing -- review \& editing; S.H.
 Annabel Chen: Project administration, Supervision, Writing -- review \& editing.
@@
 In this set of studies, we used the parametric framework for visual illusions implemented in the \emph{Pyllusion} software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength.
-We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task.
-We then extracted the participant-level performance scores (n=250), and provide evidence in favour of the existence of a general factor (labelled Factor \emph{i}) underlying the sensitivity to different illusions.
-Finally, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Negative Affect and Disinhibition.
+We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task, from which we extracted participant-level performance scores (n=250).
+Our results provide evidence in favour of the existence of a general factor (labelled Factor \emph{i}) underlying the sensitivity to different illusions.
+Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
 All the materials are available in open-access (\url{https://github.com/RealityBending/IllusionGameValidation}).
@@
 
-Visual illusions are fascinating stimuli, capturing a key feature of our neurocognitive systems.
+Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems.
 They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience (Carbon, 2014).
@@
 Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud, Francis, et al., 2020; Cretenoud et al., 2019a; Hamburger, 2016; Teufel et al., 2018b); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019b; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018a) are strongly debated.
-The existence of dispositional or demographic correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022); as well as in males as compared to females (Lo \& Dinov, 2011; Miller, 2001; Papageorgiou et al., 2020; Shaqiri et al., 2018) (\textbf{Is there something for personality that we could use here instead of sex?}) - is another area of controversy.
+The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022; Pessoa et al., 2008), as well as in individuals with stronger aggression and narcissism traits (Konrath et al., 2009a; Zhang et al., 2017) - is another area of controversy.
 
 One key challenge hindering the further development of illusion research is the relative difficulty in adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021a).
-This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus) with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
+To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021).
+This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus, \ldots) with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
 
@@
 Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016a).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
 Alternatively, \emph{Pyllusion} allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
@@
 
-Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), and demographic and dispositional variables.
-\textbf{TODO: Insert a bit of literature about why and some findings justifying what we did}.
+Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), contextual variables (pertaining to the experiment setting), such as screen size, demographic features (such as sex and age), and stable dispositional variables such as ``general'' personality traits.
+Indeed, despite the abundant literature on visual illusions, relatively few studies have investigated its ties with participants' characteristics.
+Research examining the influence of demographic variables such as gender and age have generally found inconsistent results (Cretenoud, Grzeczkowski, et al., 2020; Grzeczkowski et al., 2017; Lo \& Dinov, 2011; Papageorgiou et al., 2020).
+Regarding links with personality, most works focused on traits associated with psychopathology, such as impulsivity or sensation-seeking (Hlavatá et al., 2018; Lányi et al., 2022; Pessoa et al., 2008; Zhang et al., 2017).
 
@@
 
-Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
-Whereas some researchers report no relationship between inter-individual traits and individuals' susceptibility to illusions (Cretenoud, Grzeczkowski, et al., 2020; Grzeczkowski et al., 2017), others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects (Ohmann \& Burgmer, 2016; Zhang et al., 2017).
-Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal (Pessoa et al., 2008).
+We report significant links between inter-individual indices of illusion sensitivity and varialbes related to experimental context, demographic characteristics and personality.
+Firstly, screen size was found to have a significant negative relationship with the sensitivity to the two contrast-based illusions, namely the White and Contrast illusions.
+One possible explanation can be found in the mechanism by which visual systems filter through more low spatial frequencies when the size of the target object is small (Dixon et al., 2014).
+As this filtering process excludes illumination information from visual processing, smaller screen sizes could yield artifactual changes in brightness perception, which in turn could attenuate the illusory effect of luminance-related illusions.
+
+Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females reported to the Zöllner illusion.
+As we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity (Lo \& Dinov, 2011; Miller, 2001; Papageorgiou et al., 2020).
+This inconsistency could be due to past studies using a measure of illusion sensitvity that conflates the effect of illusions \emph{per se} with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
+On the contrary, the perceptual difficulty of the task and the illusion effect was independently modulated in our paradigm, and statistically dissociated.
+Our scores of illusion sensitivity might thus be less loaded with perceptual skills, thereby mitigating its effect.
+
+Our findings also suggested a positive relationship between illusion sensitivity and two ``normal'' personality traits, namely \emph{Honesty-Humility} and \emph{Agreeableness}, and a negative link with \emph{Antagonism}.
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism (Konrath et al., 2009a; Zhang et al., 2017).
+While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
+Since narcissism and aggression tendencies are correlated with lower field dependence (D'Amour et al., 2021; Ohmann \& Burgmer, 2016; i.e., a lesser reliance on external cues in ambiguous contexts, Witkin \& Goodenough, 1976), opposite traits, such as \emph{Honesty-Humility} and \emph{Agreeableness}, could conversely be more biased by contextual cues and thus more sensitive to illusions.
+
+The positive relationship between illusion sensitivity and ``positive'' personality traits is mirrored by a negative relationship with several other pathological traits, including \emph{Psychoticism}, \emph{Disinhibition}, and \emph{Negative Affect}.
+These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing (including illusion sensitivity, Konrath et al., 2009b).
+For instance, pathological egocentric beliefs (often observed alongside \emph{Psychoticism}, Fox, 2006) have been related to reduced context integration (Fox, 2006; Konrath et al., 2009b; manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, Ohmann \& Burgmer, 2016).
+As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits {[}\textbf{REF}{]}.
+
+Furthermore, these results in favour of a link between illusion sensitivity and maladaptive personality traits in a non-clinical population could be put in relation with clinical findings, which could be seen as extreme cases where the relationship with illusion sensitivity is the most manifest.
+In line with our results (in particular on \emph{Psychoticism} and \emph{Disinhibition}), prior research has found greater illusion resistance in schizophrenia (Grzeczkowski et al., 2018; Notredame et al., 2014; Pessoa et al., 2008), and in particular, in association with schizotypal traits, such as cognitive disorganization (Cretenoud et al., 2019b; Lányi et al., 2022).
 
@@
 
-Using the parametric illusion generation framework we developed, \emph{Pyllusion} (Makowski et al., 2021b), we have hence shown that illusions can be manipulated continuously across several different visual illusions.
-This opens the door for new illusions-based paradigms and tasks, therefore making it possible for future researchers to directly manipulate specific features and parameters of the illusion that are of interest.
-The validation of this novel framework also affords future illusion scientists a standardized measure of illusion susceptibility, instead of relying on conventional methods that depend upon participants' subjective perceptions.
-In our paradigm, in which we apply this approach to a reaction-time task, we were able to measure inter-individual scores of objective illusion sensitivity.
+The parametric illusion generation framework developed in Makowski et al.
+(2021) proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
+In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
+This important methodological step opens the door for new illusions-based paradigms and tasks, to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
 
-Notably, the general sensitivity to illusions Factor \emph{i} was negatively associated with \emph{Antagonism}, \emph{Psychoticism}, \emph{Disinhibition} and \emph{Negative Affect}.
+The participants' sensitivities to 10 different types of visual illusions shared a common part of variance, suggesting the existence of a general factor of illusion sensitivity (Factor \emph{i}).
+This result comes in a field of mixed findings.
+In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals (Cretenoud, Francis, et al., 2020; Cretenoud et al., 2019b; Grzeczkowski et al., 2017, 2018; Yang et al., 2012).
+Instead, past findings suggests illusory effects are highly specific to the perceptual features of the illusions at stake (Cretenoud et al., 2019b; Grzeczkowski et al., 2017).
+It is to note, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
+We believe that our study presents several methodological improvements, including statistical power (high number of trials per participant), homogeneous stimuli (with minimal and highly controlled features) and tasks (decision-making reaction-time task), and more reliable participant-level score extraction method (based on random-factors models), which in our opinion contributed to the emergence of the common factor.
 
-Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions (Mylniec \& Bednarek, 2016b).
-For instance, while some researchers have tried to explain our sensitivity to illusory effects as a result of deficits in the low-level visual processing system (Cretenoud et al., 2019b; Gori et al., 2016), others have provided a compelling case using a top-down approach, suggesting that such visual phenomena occur as a result of a conflict between our visual input and our prior beliefs(Caporuscio et al., 2022; Teufel et al., 2018a).
+However, although the illusions were relatively different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), The possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity, and assess the psychometric properties of similar paradigms, including stability (e.g., test-retest reliability) and validity.
 
-Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
-Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder {[}ASD{]}) (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022), other studies have found no significant differences between such individuals and healthy controls (Kaliuzhna et al., 2019; Spencer \& Ghorashi, 2014; Tibber et al., 2013; Yang et al., 2012) or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects (Grzeczkowski et al., 2018; Manning et al., 2017).
-
-\hypertarget{future-directions}{%
-\section{Future Directions}\label{future-directions}}
+Finally, we found that the general sensitivity to illusions Factor \emph{i} was negatively associated with \emph{Antagonism}, \emph{Psychoticism}, \emph{Disinhibition} and \emph{Negative Affect}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-damour2021}{}}%
+D'Amour, S., Harris, L.
+R., Berti, S., \& Keshavarz, B.
+(2021).
+The role of cognitive factors and personality traits in the perception of illusory self-motion (vection).
+\emph{Attention, Perception \& Psychophysics}, \emph{83}(4), 1804--1817.
+\url{https://doi.org/10.3758/s13414-020-02228-3}
+
 \leavevmode\vadjust pre{\hypertarget{ref-day1972}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-dixon2014}{}}%
+Dixon, E., Shapiro, A., \& Lu, Z.-L.
+(2014).
+Scale-invariance in brightness illusions implicates object-level visual processing.
+\emph{Scientific Reports}, \emph{4}(1), 3900.
+\url{https://doi.org/10.1038/srep03900}
+
 \leavevmode\vadjust pre{\hypertarget{ref-eagleman2001}{}}%
 Eagleman, D. M. (2001). Visual illusions and neurobiology. \emph{Nature Reviews Neuroscience}, \emph{2}(12), 920--926. \url{https://doi.org/10.1038/35104092}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fox2006}{}}%
+Fox, A. (2006). \emph{Adolescent self-development and psychopathology: Anorexia nervosa and psychosis} {[}PhD thesis{]}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-hlavata2018}{}}%
+Hlavatá, P., Kašpárek, T., Linhartová, P., Ošlejšková, H., \& Bareš, M.
+(2018).
+Autism, impulsivity and inhibition a review of the literature.
+\emph{Basal Ganglia}, \emph{14}, 44--53.
+\url{https://doi.org/10.1016/j.baga.2018.10.002}
+
 \leavevmode\vadjust pre{\hypertarget{ref-hopwood2012}{}}%
@@
 \url{https://doi.org/10.1037/a0026656}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kaliuzhna2019}{}}%
-Kaliuzhna, M., Stein, T., Rusch, T., Sekutowicz, M., Sterzer, P., \& Seymour, K.
-J.
-(2019).
-No evidence for abnormal priors in early vision in schizophrenia.
-\emph{Schizophrenia Research}, \emph{210}, 245--254.
-\url{https://doi.org/10.1016/j.schres.2018.12.027}
 
@@
 Kitaoka, A., \& Ishihara, M. (2000). Three elemental illusions determine the zöllner illusion. \emph{Perception \& Psychophysics}, \emph{62}, 569--575.
+
+\leavevmode\vadjust pre{\hypertarget{ref-konrath2009seeing}{}}%
+Konrath, S., Bushman, B.
+J., \& Grove, T.
+(2009a).
+Seeing my world in a million little pieces: Narcissism, self-construal, and cognitive--perceptual style.
+\emph{Journal of Personality}, \emph{77}(4), 1197--1228.
+
+\leavevmode\vadjust pre{\hypertarget{ref-konrath2009}{}}%
+Konrath, S., Bushman, B., \& Grove, T.
+(2009b).
+Seeing my world in a million little pieces: Narcissism, self-construal, and cognitive-perceptual style.
+\emph{Journal of Personality}, \emph{77}, 1197--1228.
+\url{https://doi.org/10.1111/j.1467-6494.2009.00579.x}
 
@@
 H.
-(2021a).
+(2021).
 A Parametric Framework to Generate Visual Illusions Using Python.
@@
 \url{https://doi.org/10.1177/03010066211057347}
-
-\leavevmode\vadjust pre{\hypertarget{ref-makowski2021a}{}}%
-Makowski, D., Lau, Z.
-J., Pham, T., Paul Boyce, W., \& Annabel Chen, S.
-H.
-(2021b).
-A Parametric Framework to Generate Visual Illusions Using Python.
-\emph{Perception}, \emph{50}(11), 950--965.
-\url{https://doi.org/10.1177/03010066211057347}
-
-\leavevmode\vadjust pre{\hypertarget{ref-manning2017}{}}%
-Manning, C., Morgan, M.
-J., Allen, C.
-T.
-W., \& Pellicano, E.
-(2017).
-Susceptibility to ebbinghaus and müller-lyer illusions in autistic children: A comparison of three different methods.
-\emph{Molecular Autism}, \emph{8}(1), 16.
-\url{https://doi.org/10.1186/s13229-017-0127-y}
 
@@
 Mylniec, A., \& Bednarek, H.
-(2016b).
-Field dependence, efficiency of information processing in working memory and susceptibility to orientation illusions among architects.
-\emph{Polish Psychological Bulletin}, \emph{47}(1), 112--122.
-\url{https://doi.org/10.1515/ppb-2016-0012}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mylniec2016a}{}}%
-Mylniec, A., \& Bednarek, H.
-(2016a).
+(2016).
 Field dependence, efficiency of information processing in working memory and susceptibility to orientation illusions among architects.
@@
 (2008).
-The miiller-lyer illusion as a tool for schizophrenia screening.
+The müller-lyer illusion as a tool for schizophrenia screening.
 \emph{Reviews in the Neurosciences}, \emph{19}(2-3).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-shaqiri2018}{}}%
-Shaqiri, A., Roinishvili, M., Grzeczkowski, L., Chkonia, E., Pilz, K., Mohr, C., Brand, A., Kunchulia, M., \& Herzog, M.
-H.
-(2018).
-Sex-related differences in vision are heterogeneous.
-\emph{Scientific Reports}, \emph{8}(1), 7521.
-\url{https://doi.org/10.1038/s41598-018-25298-8}
-
 \leavevmode\vadjust pre{\hypertarget{ref-shekhar2021}{}}%
@@
 \emph{New Zealand Journal of Psychology}, \emph{40}, 142--159.
-
-\leavevmode\vadjust pre{\hypertarget{ref-spencer2014}{}}%
-Spencer, K., \& Ghorashi, S.
-(2014).
-Oscillatory dynamics of gestalt perception in schizophrenia revisited.
-\emph{Frontiers in Psychology}, \emph{5}.
-\url{https://www.frontiersin.org/articles/10.3389/fpsyg.2014.00068}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-tibber2013}{}}%
-Tibber, M., Anderson, E., Bobin, T., Antonova, E., Seabright, A., Wright, B., Carlin, P., Shergill, S., \& Dakin, S.
-(2013).
-Visual surround suppression in schizophrenia.
-\emph{Frontiers in Psychology}, \emph{4}.
-\url{https://www.frontiersin.org/articles/10.3389/fpsyg.2013.00088}
-
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
@@
 \url{https://doi.org/10.21105/joss.01686}
+
+\leavevmode\vadjust pre{\hypertarget{ref-witkin1976}{}}%
+Witkin, H.
+A., \& Goodenough, D.
+R.
+(1976).
+Field Dependence and Interpersonal Behavior.
+\emph{ETS Research Bulletin Series}, \emph{1976}(1), i--78.
+\url{https://doi.org/10.1002/j.2333-8504.1976.tb01098.x}
 

### 2022-09-07 manuscript.Rmd
update

@@
 title             : "**The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates**"
-shorttitle        : "Illusion Game"
+shorttitle        : "Illusion Game Validation"
 author:
@@
 
-Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @teufel2018a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
+Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
 The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], as well as in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing] - is another area of controversy.
@@
 Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*, and a negative link with *Antagonism*.
-Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing].
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing; pessoa2008].
 While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
@@
 
-Finally, we found that the general sensitivity to illusions Factor *i* was negatively associated with *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
+Finally, we found that the sensitivity to illusions was positively associated with "positive" personality traits, such as *Agreeableness* and *Honesty-Humility*, and negatively associated with maladaptive traits such as *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
+Beyond highlighting the relevance of illusions beyond the field of visual perception, these results point towards an association with high-level general-domain mechanisms.
+While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
 
-<!-- Furthermore, results from studies that have been conducted to elucidate the mechanism underlying our susceptibility towards visual illusions remain relatively mixed.
-Whereas higher resistance towards such illusions have been reported for individuals with pathologically strong prior beliefs (such as schizophrenics) and atypical sensory perception (for example, those with autism spectrum disorder [ASD]) [@notredame2014; @park2022; @giaouri2011; @keane2014], other studies have found no significant differences between such individuals and healthy controls [@spencer2014; @yang2012; @tibber2013; @kaliuzhna2019] or only a weak correlation between the magnitude of visual illusions and such individuals' susceptibility to these illusory effects [@grzeczkowski2018; @manning2017].
--->
-
-<!-- Despite the widespread interest in finding associations between personality correlates and a general sensitivity to visual illusions, the supporting literature remains relatively mixed.
-Whereas some researchers report no relationship between inter-individual traits and individuals' susceptibility to illusions [@cretenoud2020a; @grzeczkowski2017], others have found pathological traits such as higher Aggression-Hostility and Narcissism, and lower Impulsive Sensation-Seeking to be significantly correlated to a greater resistance against such illusory effects [@zhang2017; @konrath2009seeing].
-Furthermore, research on patients with schizophrenia have also generally demonstrated lower illusion sensitivity to be linked to symptoms of the disorder, such as greater hostility as well as social and emotional withdrawal [@pessoa2008].
--->
-
-<!-- Dump  -->
-
-<!-- Most notably, there is currently no universally agreed upon neurocognitive mechanism that explains individuals' susceptibility to visual illusions [@mylniec2016].
-For instance, while some researchers have tried to explain our sensitivity to illusory effects as a result of deficits in the low-level visual processing system [@cretenoud2019; @gori2016], others have provided a compelling case using a top-down approach, suggesting that such visual phenomena occur as a result of a conflict between our visual input and our prior beliefs[@teufel2018; @caporuscio2022].
--->
-
-In conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+As a conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
 The task, data and analysis script are available in open-access at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.

### 2022-09-07 manuscript.tex
update

@@
 
-Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud, Francis, et al., 2020; Cretenoud et al., 2019a; Hamburger, 2016; Teufel et al., 2018b); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019b; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018a) are strongly debated.
+Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud, Francis, et al., 2020; Cretenoud et al., 2019a; Hamburger, 2016); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019b; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018) are strongly debated.
 The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022; Pessoa et al., 2008), as well as in individuals with stronger aggression and narcissism traits (Konrath et al., 2009a; Zhang et al., 2017) - is another area of controversy.
@@
 Our findings also suggested a positive relationship between illusion sensitivity and two ``normal'' personality traits, namely \emph{Honesty-Humility} and \emph{Agreeableness}, and a negative link with \emph{Antagonism}.
-Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism (Konrath et al., 2009a; Zhang et al., 2017).
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism {[}Zhang et al.
+(2017); Konrath et al.
+(2009a); pessoa2008{]}.
 While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
@@
 
-Finally, we found that the general sensitivity to illusions Factor \emph{i} was negatively associated with \emph{Antagonism}, \emph{Psychoticism}, \emph{Disinhibition} and \emph{Negative Affect}.
+Finally, we found that the sensitivity to illusions was positively associated with ``positive'' personality traits, such as \emph{Agreeableness} and \emph{Honesty-Humility}, and negatively associated with maladaptive traits such as \emph{Antagonism}, \emph{Psychoticism}, \emph{Disinhibition} and \emph{Negative Affect}.
+Beyond highlighting the relevance of illusions beyond the field of visual perception, these results point towards an association with high-level general-domain mechanisms.
+While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
 
-We strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+As a conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
 The task, data and analysis script are available in open-access at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
@@
 C.
-(2018a).
+(2018).
 Prior object-knowledge sharpens properties of early visual feature-detectors.
 \emph{Scientific Reports}, \emph{8}(1), 10853.
-\url{https://doi.org/10.1038/s41598-018-28845-5}
-
-\leavevmode\vadjust pre{\hypertarget{ref-teufel2018a}{}}%
-Teufel, C., Dakin, S., \& Fletcher, P.
-(2018b).
-Prior object-knowledge sharpens properties of early visual feature-detectors.
-\emph{Scientific Reports}, \emph{8}.
 \url{https://doi.org/10.1038/s41598-018-28845-5}

### 2022-09-07 manuscript.Rmd
fix

@@
 abstract: |
-  Visual illusions strikingly highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, which ideally require a controlled and gradual modulation of the effects of interest.
+  Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
+Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings.
 In this set of studies, we used the parametric framework for visual illusions implemented in the *Pyllusion* software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength.
@@
 Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
-All the materials are available in open-access (https://github.com/RealityBending/IllusionGameValidation).
-We invite researchers to re-analyze the data using alternative approaches to provide complimentary findings on the effect, structure and correlates, of illusion sensitivity.
 
@@
 Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*, and a negative link with *Antagonism*.
-Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing; pessoa2008].
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing; @pessoa2008].
 While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
@@
 In line with our results (in particular on *Psychoticism* and *Disinhibition*), prior research has found greater illusion resistance in schizophrenia [@notredame2014; @pessoa2008; @grzeczkowski2018], and in particular, in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019; @lanyi2022can].
-
-<!-- The results reported by past studies on the relationship between illusion sensitivity and personality remain fairly mixed [@cretenoud2020a; @grzeczkowski2017].  -->
-
-<!-- Moreover, low aggression and high self-control have also been found to correlate strongly with greater resistance to illusions [@mcgurk]. -->
 

### 2022-09-07 manuscript.tex
fix

@@
 
-\shorttitle{Illusion Game}
+\shorttitle{Illusion Game Validation}
 
@@
 \abstract{%
-Visual illusions strikingly highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, which ideally require a controlled and gradual modulation of the effects of interest.
+Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
+Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings.
 In this set of studies, we used the parametric framework for visual illusions implemented in the \emph{Pyllusion} software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength.
@@
 Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
-All the materials are available in open-access (\url{https://github.com/RealityBending/IllusionGameValidation}).
-We invite researchers to re-analyze the data using alternative approaches to provide complimentary findings on the effect, structure and correlates, of illusion sensitivity.
 }
@@
 Our findings also suggested a positive relationship between illusion sensitivity and two ``normal'' personality traits, namely \emph{Honesty-Humility} and \emph{Agreeableness}, and a negative link with \emph{Antagonism}.
-Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism {[}Zhang et al.
-(2017); Konrath et al.
-(2009a); pessoa2008{]}.
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism (Konrath et al., 2009a; Pessoa et al., 2008; Zhang et al., 2017).
 While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.

### 2022-09-07 manuscript.Rmd
update ref

@@
 <!-- Context -->
+
 We report significant links between inter-individual indices of illusion sensitivity and varialbes related to experimental context, demographic characteristics and personality.
@@
 <!-- Demographic -->
+
 Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females reported to the Zöllner illusion.
@@
 <!-- Normal personality + Antagonism -->
+
 Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*, and a negative link with *Antagonism*.
@@
 <!-- Pathological personality -->
+
 The positive relationship between illusion sensitivity and "positive" personality traits is mirrored by a negative relationship with several other pathological traits, including *Psychoticism*, *Disinhibition*, and *Negative Affect*.
@@
 For instance, pathological egocentric beliefs [often observed alongside *Psychoticism*, @fox2006] have been related to reduced context integration [manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, @ohmann2016; @konrath2009; @fox2006].
-As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits [**REF**].
+As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits [@msetfi2009; @parkes1981field; @calamari2000field; @hoyle2006] .
 
@@
 <!-- Parametric framework -->
+
 The parametric illusion generation framework developed in @makowski2021 proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
@@
 <!-- Factor i -->
+
 The participants' sensitivities to 10 different types of visual illusions shared a common part of variance, suggesting the existence of a general factor of illusion sensitivity (Factor *i*).
@@
 Instead, past findings suggests illusory effects are highly specific to the perceptual features of the illusions at stake [@grzeczkowski2017; @cretenoud2019].
-It is to note, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
+It should be noted, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
 We believe that our study presents several methodological improvements, including statistical power (high number of trials per participant), homogeneous stimuli (with minimal and highly controlled features) and tasks (decision-making reaction-time task), and more reliable participant-level score extraction method (based on random-factors models), which in our opinion contributed to the emergence of the common factor.
@@
 
-As a conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+To conclude, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
 The task, data and analysis script are available in open-access at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.

### 2022-09-07 manuscript.Rmd
update

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "1156"
+wordcount         : "5096"
 bibliography      : references.bib
@@
 
-<!-- Title: -->
-
-<!-- The Illusion Game: Validation of a Novel Experimental Paradigm to study Visual Illusions  -->
+<!-- Significance Statement (< 120 words) -->
+<!-- A novel paradigm to study the objective effect of visual illusions is presented, which yielded evidence in favor of a common factor to visual illusions (Factor *i*) and a relationship between a low illusion sensitivity and maladaptive personality traits, such as antagonism, psychoticism and disinhibition.
+-->
 

### 2022-09-07 manuscript.tex
update

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 1156}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 5096}
 \usepackage{lineno}
@@
 For instance, pathological egocentric beliefs (often observed alongside \emph{Psychoticism}, Fox, 2006) have been related to reduced context integration (Fox, 2006; Konrath et al., 2009b; manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, Ohmann \& Burgmer, 2016).
-As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits {[}\textbf{REF}{]}.
+As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits (Calamari et al., 2000; Hoyle, 2006; Msetfi et al., 2009; Parkes, 1981) .
 
@@
 Instead, past findings suggests illusory effects are highly specific to the perceptual features of the illusions at stake (Cretenoud et al., 2019b; Grzeczkowski et al., 2017).
-It is to note, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
+It should be noted, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
 We believe that our study presents several methodological improvements, including statistical power (high number of trials per participant), homogeneous stimuli (with minimal and highly controlled features) and tasks (decision-making reaction-time task), and more reliable participant-level score extraction method (based on random-factors models), which in our opinion contributed to the emergence of the common factor.
@@
 
-As a conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+To conclude, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
 The task, data and analysis script are available in open-access at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
@@
 \emph{British Journal of Mathematical and Statistical Psychology}, \emph{73}(3), 420--451.
+
+\leavevmode\vadjust pre{\hypertarget{ref-calamari2000field}{}}%
+Calamari, E., Pini, M., \& Puleggio, A.
+(2000).
+Field dependence and verbalized strategies on the portable rod-and-frame test by depressed outpatients and normal controls.
+\emph{Perceptual and Motor Skills}, \emph{91}(3\_suppl), 1221--1229.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-hoyle2006}{}}%
+Hoyle, R.
+H.
+(2006).
+Personality and self-regulation: trait and information-processing perspectives.
+\emph{Journal of Personality}, \emph{74}(6), 1507--1525.
+\url{https://doi.org/10.1111/j.1467-6494.2006.00418.x}
+
 \leavevmode\vadjust pre{\hypertarget{ref-keane2014}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-msetfi2009}{}}%
+Msetfi, R.
+M., Murphy, R.
+A., Kornbrot, D.
+E., \& Simpson, J.
+(2009).
+Short article: Impaired context maintenance in mild to moderately depressed students.
+\emph{Quarterly Journal of Experimental Psychology}, \emph{62}(4), 653--662.
+\url{https://doi.org/10.1080/17470210802486092}
+
 \leavevmode\vadjust pre{\hypertarget{ref-mylniec2016}{}}%
@@
 \url{https://doi.org/10.1111/ejn.15739}
+
+\leavevmode\vadjust pre{\hypertarget{ref-parkes1981field}{}}%
+Parkes, K. R. (1981). Field dependence and the differentiation of affective states. \emph{The British Journal of Psychiatry}, \emph{139}(1), 52--58.
 

### 2022-09-07 manuscript.Rmd
address comments by steph

@@
 
-One key challenge hindering the further development of illusion research is the relative difficulty in adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+One key challenge hindering the further development of illusion research is the relative difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
 To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
-This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus, ...) with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (see **Figure 1**).
+This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (see **Figure 1**).
 
@@
 
-The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: [19, 60]; Sex: 39.1% females, 56.5% males).
+The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: [19, 60]; Sex: 39.1% females, 56.5% males, and 4.4% other).
 
@@
 We then extracted the inter-individual variability in the effect of illusion strength and its interaction with task difficulty, and used it as participant-level scores.
-Finally, We explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
+Finally, we explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
 
@@
 
-```{r message=FALSE, warning=FALSE, out.width="100%", fig.cap="Top: the effect of illusion strength and task difficulty on the error rage and reaction time (RT) for each individual illusions.
+```{r message=FALSE, warning=FALSE, out.width="100%", fig.cap="Top: the effect of illusion strength and task difficulty on the error rate and reaction time (RT) for each individual illusions.
 The solid line represent the General Additive Model (GAM), and the dashed line correspond to its approximation via linear models.
@@
 Task difficulty (the objective difference between the targets of perceptual decision) levels are shown as colors, with lowest values corresponding to harder trials.
-Each illusion type is surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficutly (on top) and the strongest illusion (on the right for incongruent illusions).
+Each illusion type is surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficulty (on top) and the strongest illusion (on the right for incongruent illusions).
 Bottom: We extracted the effect slope of the illusion strength and its interaction with task difficulty for each participant.
@@
 
-Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy (and are also more conservative estimates due shrinkage) than basic empirical scores, such as the total number of errors, or the average RT.
+Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy (and are also more conservative estimates due to shrinkage) than basic empirical scores, such as the total number of errors, or the average RT.
 Thus, for each illusion and within each participant, we extracted the effect of illusion strength and its interaction with task difficulty when the illusion effect was incongruent.
@@
 This was fixed in latest release of *Pyllusion* (v1.2), which now generate outlines of the same size as the target circle.
-For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports [@kitaoka2000; @kitaoka2007], suggesting an acute angle contraction effect at very small as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
+For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports [@kitaoka2000; @kitaoka2007], suggesting an acute angle contraction effect at very small - as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
 
@@
 ```{r message=FALSE, warning=FALSE, fig.cap="The upper plots show the distribution of demographic and dispositional variables.
-The middle plots show the relationship between illusion sensitivty scores, sex, and age (solid lines indicate significant relationships).
+The middle plots show the relationship between illusion sensitivity scores, sex, and age (solid lines indicate significant relationships).
 Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \\textit{i}), and personality traits.", out.width="100%"}
@@
 
-We report significant links between inter-individual indices of illusion sensitivity and varialbes related to experimental context, demographic characteristics and personality.
+We report significant links between inter-individual indices of illusion sensitivity and variables related to experimental context, demographic characteristics and personality.
 Firstly, screen size was found to have a significant negative relationship with the sensitivity to the two contrast-based illusions, namely the White and Contrast illusions.

### 2022-09-07 study2/functions.R

@@ -692,7 +692,8 @@ prettify_itemName <- function(x) {
     str_replace("VerticalHorizontal", "Vertical-Horizontal") |>
     str_replace("Zollner", "Zöllner") |>
     str_replace("MullerLyer", "Müller-Lyer") |>
-    str_replace("HonestyHumility", "Honesty-Humility")
+    str_replace("HonestyHumility", "Honesty-Humility") |>
+    str_replace("NegativeAffect", "Negative Affect")
 }
 
 

### 2022-09-07 manuscript.Rmd
Update manuscript.Rmd

@@
 Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
-Future studies are needed to explore reaction times and try to identify the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
+Future studies are needed to explore reaction times and try to identify the most appropriate models, and/or use models that integrate errors and reaction time (e.g., drift diffusion models).
 
@@
 In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals [@grzeczkowski2017; @cretenoud2019; @cretenoud2020; @grzeczkowski2018; @yang2012].
-Instead, past findings suggests illusory effects are highly specific to the perceptual features of the illusions at stake [@grzeczkowski2017; @cretenoud2019].
+Instead, past findings suggest illusory effects are highly specific to the perceptual features of the illusions at stake [@grzeczkowski2017; @cretenoud2019].
 It should be noted, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.

### 2022-09-07 manuscript.Rmd
Update manuscript.Rmd

@@
 Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
-The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], as well as in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing] - is another area of controversy.
+The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008] and in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
 

### 2022-09-07 manuscript.Rmd
fixes

@@
 They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
-Despite the historical and intensive interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022; @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
+Despite the longstanding interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022; @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
 
@@
 Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to a reference stimulus relying only on their perception [@mylniec2016; @grzeczkowski2018].
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to match a reference stimulus relying only on their perception [@mylniec2016; @grzeczkowski2018].
 Alternatively, *Pyllusion* allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
@@
 
-This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
 Moreover, the illusion effect can be either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier).
@@
 
-In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion and task difficulty can be manipulated continuously, and separately modeled statistically.
+In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously, and separately modeled statistically.
 Then, we will further utilize the paradigm to assess whether 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor.
@@
 
-Study 1 can be seen as a pilot experiment aiming to gather some preliminary data to assess if the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors); and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
+Study 1 can be seen as a pilot experiment aiming to gather some preliminary data to assess if the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors), and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
 
@@
 The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the blocks was also randomized.
-After the first series of 10 blocks, another series was done (with new randomized order of blocks and trials).
+After the first series of 10 blocks, another series was administered (with new randomized orders of blocks and trials).
 In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
-The task was implemented using *jsPsych* [@de2015jspsych].
-The instructions for each illusion type are available in the experiment code.
+The task was implemented using *jsPsych* [@de2015jspsych], and the instructions for each illusion type are available in the experiment code.
 
@@
 
-We removed 6 participants upon inspection of the average error rage (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
+We removed 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
 For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that instructions got misunderstood; e.g., participants were selecting the shorter line instead of the longer one).
@@
 This study provided a clearer understanding of the magnitude of the parametric effects at stake and the type of interaction between them.
-Furthermore, it allowed us to better understand and test the stimuli generated by *Pyllusion*, as well as uncover technical bugs and issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed by a new software release.
+Furthermore, it allowed us to better understand and test the stimuli generated by *Pyllusion*, as well as uncover incidental bugs and technical issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed in a new software release.
 Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
@@
 The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT), within each illusion.
-In order to achieve that, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
+In order to achieve this, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
 Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].
@@
 Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
-Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we selected and fitted the best model (best on their indices of fit), and compared their output visually (see **Figure 2**).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we selected and fitted the best model (based on their indices of fit), and compared their output visually (see **Figure 2**).
 
@@
 
-The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; $cbrt(diff)*strength$ for Zöllner; $diff*sqrt(strength)$ and $log(diff)*strength$ respectively for errors and RT in White; $sqrt(diff)*sqrt(strength)$ and $sqrt(diff)*strength$ respectively for errors and RT in Müller-Lyer; $cbrt(diff)*strength$ for Ponzo; $cbrt(diff)*sqrt(strength)$ and $cbrt(diff)*strength$ respectively for errors and RT in Poggendorff; $sqrt(diff)*sqrt(strength)$ for Contrast.
-In all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
+The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; $cbrt(diff)*strength$ for Zöllner; $diff*sqrt(strength)$ and $log(diff)*strength$ respectively for errors and RT in White; $sqrt(diff)*sqrt(strength)$ and $sqrt(diff)*strength$ respectively for errors and RT in Müller-Lyer; $cbrt(diff)*strength$ for Ponzo; $cbrt(diff)*sqrt(strength)$ and $cbrt(diff)*strength$ respectively for errors and RT in Poggendorff; and $sqrt(diff)*sqrt(strength)$ for Contrast.
+For all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
 
-For errors, most of the models closely matched their GAMs counterpart (see **Figure 2**), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
+For error rates, most of the models closely matched their GAMs counterpart (see **Figure 2**), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
 
@@
 The solid line represent the General Additive Model (GAM), and the dashed line correspond to its approximation via linear models.
-Descriptive data is shown with stacked dots (errors are hanging from the top) and distributions for RTs.
+Descriptive data is shown with stacked dots (for which errors start from the top) and distributions for RTs.
 Negative values for illusion strength correspond to congruent (i.e., facilitating) illusion effects.
-Task difficulty (the objective difference between the targets of perceptual decision) levels are shown as colors, with lowest values corresponding to harder trials.
-Each illusion type is surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficulty (on top) and the strongest illusion (on the right for incongruent illusions).
+Task difficulty (the objective difference between the targets of perceptual decision) levels are shown as colors, with lower values corresponding to harder trials.
+The results for each illusion are surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficulty (on top) and the strongest illusion (on the right for incongruent illusions).
 Bottom: We extracted the effect slope of the illusion strength and its interaction with task difficulty for each participant.
@@
 
-For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increase beyond a certain threshold, the participants respond faster.
+For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants respond faster.
 While this is is not surprising (strong illusions are likely so effective in biasing perception that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
@@
 
-Finally, we tested this data-driven model (*m0*) against four other structural models using structural equation modelling (SEM): one in which the two parameters of each of the 10 illusions (illusion strength and interaction with task difficulty) loaded on separate factors, which then all loaded on a common factor (*m1*); one which the parameters were grouped by illusion type (lines, circles, contrast and angle) before loading on a common factor (*m2*); one in which all the parameters related to strength, and all the parameters related to the interaction loaded onto two respective factors, which then loaded on a common factor (*m3*); and one in which there was no intermediate level: all 20 parameters loaded directly on a common factor (*m4*).
+Finally, we tested this data-driven model (*m0*) against four other structural models using structural equation modelling (SEM): one in which the two parameters of each of the 10 illusions (illusion strength and interaction with task difficulty) loaded on separate factors, which then all loaded on a common factor (*m1*); one in which the parameters were grouped by illusion type (lines, circles, contrast and angle) before loading on a common factor (*m2*); one in which all the parameters related to strength, and all the parameters related to the interaction loaded onto two respective factors, which then loaded on a common factor (*m3*); and one in which there was no intermediate level: all 20 parameters loaded directly on a common factor (*m4*).
 
-The model *m1*, in which the parameters loaded on a first level of 10 illusion-specific factors, which then all loaded on a common factor significantly outperformed the other models.
-Its indices of fit were ranging from acceptable to satisfactory (CFI = .92; SRMR = .08; NNFI = .91; PNFI = .74; RMSEA = .08), and all the specified effects were significant.
-The illusion-specific latent factors were loaded positively by the sensitivity to illusion strength, and positively by the interaction effect with task difficulty (with the exception of Delboeuf, Ebbinghaus, Vertical-Horizontal, Müller-Lyer and Contrast, for which the loading was negative).
+The model *m1*, in which the parameters loaded on a first level of 10 illusion-specific factors, which then all loaded on a common factor, significantly outperformed the other models.
+Its indices of fit ranged from acceptable to satisfactory (CFI = .92; SRMR = .08; NNFI = .91; PNFI = .74; RMSEA = .08), and all the specified effects were significant.
+The illusion-specific latent factors were loaded positively by the sensitivity to illusion strength, as well as by the interaction effect with task difficulty (with the exception of Delboeuf, Ebbinghaus, Vertical-Horizontal, Müller-Lyer and Contrast, for which the loading was negative).
 The general factor of illusion sensitivity, labelled Factor *i* (i- for illusion), explained 48.02% of the total variance of the initial dataset, and was strongly related to Vertical-Horizontal ($\beta_{std.}=0.83$), Müller-Lyer ($\beta_{std.}=0.76$), Ponzo ($\beta_{std.}=0.65$), Ebbinghaus ($\beta_{std.}=0.64$); moderately to Zöllner ($\beta_{std.}=0.53$), Poggendorff ($\beta_{std.}=0.44$), Rod and Frame ($\beta_{std.}=0.42$), Contrast ($\beta_{std.}=0.40$) and White ($\beta_{std.}=0.35$); and weakly to Delboeuf ($\beta_{std.}=0.19$).
@@
 
-We have to keep in mind that these individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of Delboeuf and Zöllner); 2) we only used the models on error rate, which could be biased by the speed-accuracy decision criterion used by participants; 3) the structural equation model used to compute the scores also incorporated multiple levels of abstractions.
+It is important to note that these individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of Delboeuf and Zöllner); 2) we only used the models on error rate, which could be biased by the speed-accuracy decision criterion used by participants; 3) the structural equation model used to compute the scores also incorporated multiple levels of abstractions.
 Thus, in order to validate the individual scores, we computed the correlation between them and simple empirical scores, such as the average error rate and the mean RT in the task.
-This analysis revealed strong and significant correlations between each illusion-specific factor and the average amount of errors in its respective task.
+This analysis revealed strong and significant correlations between each illusion-specific factor and the average amount of errors in its corresponding task.
 Moreover, each individual score was strongly associated with the average RT across multiple illusion types.
-This suggests that the individual scores obtained from the structural equation model do capture the sensitivity of each participant to visual illusions, manifesting in both the number of errors and high reaction times.
+This suggests that the individual scores obtained from the structural equation model do capture the sensitivity of each participant to visual illusions, manifesting in both the number of errors and long reaction times.
 
@@
 
-This study confirmed that it was possible to continuously manipulate the effect of illusion strength for 10 classical illusions.
+This study confirmed that it was possible to continuously manipulate the effect of illusion strength for 10 classic illusions.
 Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
-Future studies are needed to explore reaction times and try to identify the most appropriate models, and/or use models that integrate errors and reaction time (e.g., drift diffusion models).
+Future studies are needed to explore reaction times and identify the most appropriate models, and/or use models that integrate both errors and reaction time (e.g., drift diffusion models).
 
@@
 For the Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
-This was fixed in latest release of *Pyllusion* (v1.2), which now generate outlines of the same size as the target circle.
+This was fixed in the latest release of *Pyllusion* (v1.2), which now generates outlines of the same size as the target circle.
 For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports [@kitaoka2000; @kitaoka2007], suggesting an acute angle contraction effect at very small - as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
@@
 ```{r message=FALSE, warning=FALSE, fig.cap="The upper plots show the distribution of demographic and dispositional variables.
-The middle plots show the relationship between illusion sensitivity scores, sex, and age (solid lines indicate significant relationships).
-Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \\textit{i}), and personality traits.", out.width="100%"}
+The middle plots shows the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
+Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \\textit{i}) and personality traits.", out.width="100%"}
 knitr::include_graphics("figures/Figure3.png")
@@
 For each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) and personality.
-As the supplementary material contains the detailed results, we will here only report the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
+As the supplementary material contains the detailed results, we will only report the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
 
@@
 
-Regarding "pathological" personality traits, the results yielded strong evidence in favor of a negative relationship between multiple illusion scores and multiple traits.
+Regarding "pathological" personality traits, the results yielded strong evidence in favor of a negative relationship between illusion scores and multiple traits.
 *Antagonism* was associated with the sensitivity to Vertical-Horizontal ($BF_{10} > 100$), Müller-Lyer ($BF_{10} = 21.57$), Ponzo ($BF_{10} = 17.97$) illusions, and the Factor *i* ($BF_{10} = 55.45$); *Psychoticism* was associated with the sensitivity to Vertical-Horizontal ($BF_{10} = 66.63$) and Müller-Lyer ($BF_{10} = 35.59$) illusions, and the Factor *i* ($BF_{10} = 35.02$); *Disinhibition* was associated with the sensitivity to Vertical-Horizontal ($BF_{10} = 25.38$), Zöllner ($BF_{10} = 7.59$), Müller-Lyer ($BF_{10} = 5.89$) illusions, and the Factor *i* ($BF_{10} = 31.42$); and *Negative Affect* was associated with Zöllner ($BF_{10} = 62.04$), Vertical-Horizontal ($BF_{10} = 12.65$), Müller-Lyer ($BF_{10} = 3.17$), and the Factor *i* ($BF_{10} = 6.39$).
-The last remaining trait, *Detachment*, did not share any relationship with illusion sensitivity.
+The last remaining trait, *Detachment*, did not share any significant relationship with illusion sensitivity.
 
@@
 
-Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females reported to the Zöllner illusion.
-As we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity [@lo2011; @miller2001; @papageorgiou2020].
+Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females for the Zöllner illusion.
+Although we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity [@lo2011; @miller2001; @papageorgiou2020].
 This inconsistency could be due to past studies using a measure of illusion sensitvity that conflates the effect of illusions *per se* with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
@@
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
-This important methodological step opens the door for new illusions-based paradigms and tasks, to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
+This important methodological step opens the door for new illusions-based paradigms and tasks to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
 
@@
 
-The participants' sensitivities to 10 different types of visual illusions shared a common part of variance, suggesting the existence of a general factor of illusion sensitivity (Factor *i*).
+Our findings suggest that the sensitivity to 10 different types of visual illusions share a common part of variance, supporting the existence of a general factor of illusion sensitivity (Factor *i*).
 This result comes in a field of mixed findings.
@@
 It should be noted, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
-We believe that our study presents several methodological improvements, including statistical power (high number of trials per participant), homogeneous stimuli (with minimal and highly controlled features) and tasks (decision-making reaction-time task), and more reliable participant-level score extraction method (based on random-factors models), which in our opinion contributed to the emergence of the common factor.
+We believe that our study presents several methodological improvements, including statistical power (high number of trials per participant), homogeneous stimuli (with minimal and highly controlled features) and tasks (decision-making reaction-time task), and a more reliable participant-level score extraction method (based on random-factors models), which in our opinion contributed to the emergence of the common factor.
 
-However, although the illusions were relatively different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), The possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+However, although the illusions were relatively different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
 Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity, and assess the psychometric properties of similar paradigms, including stability (e.g., test-retest reliability) and validity.

### 2022-09-08 manuscript.Rmd
update ref

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "5096"
+wordcount         : "5099"
 bibliography      : references.bib
@@
 
-Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019a; @cretenoud2020]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
+Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
 The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008] and in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
@@
 
-The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT), within each illusion.
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT) within each illusion.
 In order to achieve this, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
@@
 
-```{r message=FALSE, warning=FALSE, out.width="100%", fig.cap="Top: the effect of illusion strength and task difficulty on the error rate and reaction time (RT) for each individual illusions.
-The solid line represent the General Additive Model (GAM), and the dashed line correspond to its approximation via linear models.
+```{r message=FALSE, warning=FALSE, out.width="100%", fig.cap="Top: the effect of illusion strength and task difficulty on the error rate and reaction time (RT) for each individual illusion.
+The solid line represents the General Additive Model (GAM), and the dashed line corresponds to its approximation via linear models.
 Descriptive data is shown with stacked dots (for which errors start from the top) and distributions for RTs.
@@
 The general factor of illusion sensitivity, labelled Factor *i* (i- for illusion), explained 48.02% of the total variance of the initial dataset, and was strongly related to Vertical-Horizontal ($\beta_{std.}=0.83$), Müller-Lyer ($\beta_{std.}=0.76$), Ponzo ($\beta_{std.}=0.65$), Ebbinghaus ($\beta_{std.}=0.64$); moderately to Zöllner ($\beta_{std.}=0.53$), Poggendorff ($\beta_{std.}=0.44$), Rod and Frame ($\beta_{std.}=0.42$), Contrast ($\beta_{std.}=0.40$) and White ($\beta_{std.}=0.35$); and weakly to Delboeuf ($\beta_{std.}=0.19$).
-We then computed, for each participant, its score for the 10 illusion-specific factors and for the general Factor *i*.
+We then computed, for each participant, the score for the 10 illusion-specific factors and for the general Factor *i*.
 
@@
 The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
-For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to small and strong increases of errors, respectively.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with fewer errors than incongruent effects) were related to small and strong increases of errors, respectively.
 For the Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
@@
 Indeed, despite the abundant literature on visual illusions, relatively few studies have investigated its ties with participants' characteristics.
-Research examining the influence of demographic variables such as gender and age have generally found inconsistent results [@cretenoud2020a; @grzeczkowski2017; @lo2011; @papageorgiou2020].
+Research examining the influence of demographic variables such as gender and age have generally found inconsistent results [@cretenoud2020individual; @grzeczkowski2017; @lo2011; @papageorgiou2020].
 Regarding links with personality, most works focused on traits associated with psychopathology, such as impulsivity or sensation-seeking [@hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
@@
 The variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
-We used the *IPIP6* [24 items, @sibley2011] to measure 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-humility), and the PID-5 [25 items, @hopwood2012] to measure "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+We used the *IPIP6* [24 items, @sibley2011] to measure 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the PID-5 [25 items, @hopwood2012] to measure "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
 The participants were the same as in study 2 (see **Figure 3**).
@@
 Although we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity [@lo2011; @miller2001; @papageorgiou2020].
-This inconsistency could be due to past studies using a measure of illusion sensitvity that conflates the effect of illusions *per se* with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
+This inconsistency could be due to past studies using a measure of illusion sensitivity that conflates the effect of illusions *per se* with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
 On the contrary, the perceptual difficulty of the task and the illusion effect was independently modulated in our paradigm, and statistically dissociated.
@@
 The positive relationship between illusion sensitivity and "positive" personality traits is mirrored by a negative relationship with several other pathological traits, including *Psychoticism*, *Disinhibition*, and *Negative Affect*.
-These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing [including illusion sensitivity, @konrath2009].
-For instance, pathological egocentric beliefs [often observed alongside *Psychoticism*, @fox2006] have been related to reduced context integration [manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, @ohmann2016; @konrath2009; @fox2006].
+These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing [including illusion sensitivity, @konrath2009seeing].
+For instance, pathological egocentric beliefs [often observed alongside *Psychoticism*, @fox2006] have been related to reduced context integration [manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, @ohmann2016; @konrath2009seeing; @fox2006].
 As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits [@msetfi2009; @parkes1981field; @calamari2000field; @hoyle2006] .
@@
 This result comes in a field of mixed findings.
-In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals [@grzeczkowski2017; @cretenoud2019; @cretenoud2020; @grzeczkowski2018; @yang2012].
+In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals [@grzeczkowski2017; @cretenoud2019; @cretenoud2020illusions; @grzeczkowski2018; @yang2012].
 Instead, past findings suggest illusory effects are highly specific to the perceptual features of the illusions at stake [@grzeczkowski2017; @cretenoud2019].

### 2022-09-08 manuscript.tex
update ref

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 5096}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 5099}
 \usepackage{lineno}
@@
 They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience (Carbon, 2014).
-Despite the historical and intensive interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science (Caporuscio et al., 2022; Lamme, 2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
+Despite the longstanding interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science (Caporuscio et al., 2022; Lamme, 2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
 
-Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud, Francis, et al., 2020; Cretenoud et al., 2019a; Hamburger, 2016); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019b; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018) are strongly debated.
-The existence of dispositional correlates of illusion sensitivity - for example, higher illusion resistance has been reported in schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022; Pessoa et al., 2008), as well as in individuals with stronger aggression and narcissism traits (Konrath et al., 2009a; Zhang et al., 2017) - is another area of controversy.
+Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud et al., 2019; Cretenoud, Francis, et al., 2020; Hamburger, 2016); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018) are strongly debated.
+The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022; Pessoa et al., 2008) and in individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
 
-One key challenge hindering the further development of illusion research is the relative difficulty in adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+One key challenge hindering the further development of illusion research is the relative difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
 To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021).
-This software allows us to generate different types of classic visual illusions (e.g., Müller-Lyer, Ponzo, Delboeuf, Ebbinghaus, \ldots) with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
+This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
 
@@
 Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them to what extent they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to match a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
 Alternatively, \emph{Pyllusion} allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
-This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment in the longest) under different conditions of objective difficulty and illusion strength.
+This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
 Moreover, the illusion effect can be either ``incongruent'' (making the task more difficult by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
@@
 
-Study 1 can be seen as a pilot experiment aiming to gather some preliminary data to assess if the stimuli generated by \emph{Pyllusion} behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors); and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
+Study 1 can be seen as a pilot experiment aiming to gather some preliminary data to assess if the stimuli generated by \emph{Pyllusion} behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors), and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
 
@@
 The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the blocks was also randomized.
-After the first series of 10 blocks, another series was done (with new randomized order of blocks and trials).
+After the first series of 10 blocks, another series was administered (with new randomized orders of blocks and trials).
 In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
-The task was implemented using \emph{jsPsych} (De Leeuw, 2015).
-The instructions for each illusion type are available in the experiment code.
+The task was implemented using \emph{jsPsych} (De Leeuw, 2015), and the instructions for each illusion type are available in the experiment code.
 
@@
 
-We removed 6 participants upon inspection of the average error rage (when close to 50\%, suggesting random answers), and when the reaction time distribution was implausibly fast.
+We removed 6 participants upon inspection of the average error rate (when close to 50\%, suggesting random answers), and when the reaction time distribution was implausibly fast.
 For the remaining participants, we discarded blocks where the error rate was higher than 50\% (possibly indicating that instructions got misunderstood; e.g., participants were selecting the shorter line instead of the longer one).
@@
 
-The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: {[}19, 60{]}; Sex: 39.1\% females, 56.5\% males).
+The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: {[}19, 60{]}; Sex: 39.1\% females, 56.5\% males, and 4.4\% other).
 
@@
 This study provided a clearer understanding of the magnitude of the parametric effects at stake and the type of interaction between them.
-Furthermore, it allowed us to better understand and test the stimuli generated by \emph{Pyllusion}, as well as uncover technical bugs and issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed by a new software release.
+Furthermore, it allowed us to better understand and test the stimuli generated by \emph{Pyllusion}, as well as uncover incidental bugs and technical issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed in a new software release.
 Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
@@
 
-The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT), within each illusion.
-In order to achieve that, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT) within each illusion.
+In order to achieve this, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
 Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location \(\mu\) (mean), scale \(\sigma\) (spread) and tail-dominance \(\tau\) of the RT distribution (Balota \& Yap, 2011; Matzke \& Wagenmakers, 2009).
@@
 Using GAMs as the ``ground-truth'' models, we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
-Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task \emph{difficulty} and illusion \emph{strength}), we selected and fitted the best model (best on their indices of fit), and compared their output visually (see \textbf{Figure 2}).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task \emph{difficulty} and illusion \emph{strength}), we selected and fitted the best model (based on their indices of fit), and compared their output visually (see \textbf{Figure 2}).
 
 We then extracted the inter-individual variability in the effect of illusion strength and its interaction with task difficulty, and used it as participant-level scores.
-Finally, We explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
+Finally, we explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
 
@@
 
-The best models were \(log(diff)*strength\) for Delboeuf; \(sqrt(diff)*strength\) for Ebbinghaus; \(log(diff)*log(strength)\) for Rod and Frame; \(sqrt(diff)*sqrt(strength)\) for Vertical-Horizontal; \(cbrt(diff)*strength\) for Zöllner; \(diff*sqrt(strength)\) and \(log(diff)*strength\) respectively for errors and RT in White; \(sqrt(diff)*sqrt(strength)\) and \(sqrt(diff)*strength\) respectively for errors and RT in Müller-Lyer; \(cbrt(diff)*strength\) for Ponzo; \(cbrt(diff)*sqrt(strength)\) and \(cbrt(diff)*strength\) respectively for errors and RT in Poggendorff; \(sqrt(diff)*sqrt(strength)\) for Contrast.
-In all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
+The best models were \(log(diff)*strength\) for Delboeuf; \(sqrt(diff)*strength\) for Ebbinghaus; \(log(diff)*log(strength)\) for Rod and Frame; \(sqrt(diff)*sqrt(strength)\) for Vertical-Horizontal; \(cbrt(diff)*strength\) for Zöllner; \(diff*sqrt(strength)\) and \(log(diff)*strength\) respectively for errors and RT in White; \(sqrt(diff)*sqrt(strength)\) and \(sqrt(diff)*strength\) respectively for errors and RT in Müller-Lyer; \(cbrt(diff)*strength\) for Ponzo; \(cbrt(diff)*sqrt(strength)\) and \(cbrt(diff)*strength\) respectively for errors and RT in Poggendorff; and \(sqrt(diff)*sqrt(strength)\) for Contrast.
+For all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
 
-For errors, most of the models closely matched their GAMs counterpart (see \textbf{Figure 2}), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
+For error rates, most of the models closely matched their GAMs counterpart (see \textbf{Figure 2}), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures/Figure2} \caption{Top: the effect of illusion strength and task difficulty on the error rage and reaction time (RT) for each individual illusions.
-The solid line represent the General Additive Model (GAM), and the dashed line correspond to its approximation via linear models.
-Descriptive data is shown with stacked dots (errors are hanging from the top) and distributions for RTs.
+\includegraphics[width=1\linewidth]{figures/Figure2} \caption{Top: the effect of illusion strength and task difficulty on the error rate and reaction time (RT) for each individual illusion.
+The solid line represents the General Additive Model (GAM), and the dashed line corresponds to its approximation via linear models.
+Descriptive data is shown with stacked dots (for which errors start from the top) and distributions for RTs.
 Negative values for illusion strength correspond to congruent (i.e., facilitating) illusion effects.
-Task difficulty (the objective difference between the targets of perceptual decision) levels are shown as colors, with lowest values corresponding to harder trials.
-Each illusion type is surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficutly (on top) and the strongest illusion (on the right for incongruent illusions).
+Task difficulty (the objective difference between the targets of perceptual decision) levels are shown as colors, with lower values corresponding to harder trials.
+The results for each illusion are surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficulty (on top) and the strongest illusion (on the right for incongruent illusions).
 Bottom: We extracted the effect slope of the illusion strength and its interaction with task difficulty for each participant.
@@
 
-For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increase beyond a certain threshold, the participants respond faster.
+For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants respond faster.
 While this is is not surprising (strong illusions are likely so effective in biasing perception that it is ``easier'', i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
@@
 
-Finally, we tested this data-driven model (\emph{m0}) against four other structural models using structural equation modelling (SEM): one in which the two parameters of each of the 10 illusions (illusion strength and interaction with task difficulty) loaded on separate factors, which then all loaded on a common factor (\emph{m1}); one which the parameters were grouped by illusion type (lines, circles, contrast and angle) before loading on a common factor (\emph{m2}); one in which all the parameters related to strength, and all the parameters related to the interaction loaded onto two respective factors, which then loaded on a common factor (\emph{m3}); and one in which there was no intermediate level: all 20 parameters loaded directly on a common factor (\emph{m4}).
+Finally, we tested this data-driven model (\emph{m0}) against four other structural models using structural equation modelling (SEM): one in which the two parameters of each of the 10 illusions (illusion strength and interaction with task difficulty) loaded on separate factors, which then all loaded on a common factor (\emph{m1}); one in which the parameters were grouped by illusion type (lines, circles, contrast and angle) before loading on a common factor (\emph{m2}); one in which all the parameters related to strength, and all the parameters related to the interaction loaded onto two respective factors, which then loaded on a common factor (\emph{m3}); and one in which there was no intermediate level: all 20 parameters loaded directly on a common factor (\emph{m4}).
 
-The model \emph{m1}, in which the parameters loaded on a first level of 10 illusion-specific factors, which then all loaded on a common factor significantly outperformed the other models.
-Its indices of fit were ranging from acceptable to satisfactory (CFI = .92; SRMR = .08; NNFI = .91; PNFI = .74; RMSEA = .08), and all the specified effects were significant.
-The illusion-specific latent factors were loaded positively by the sensitivity to illusion strength, and positively by the interaction effect with task difficulty (with the exception of Delboeuf, Ebbinghaus, Vertical-Horizontal, Müller-Lyer and Contrast, for which the loading was negative).
+The model \emph{m1}, in which the parameters loaded on a first level of 10 illusion-specific factors, which then all loaded on a common factor, significantly outperformed the other models.
+Its indices of fit ranged from acceptable to satisfactory (CFI = .92; SRMR = .08; NNFI = .91; PNFI = .74; RMSEA = .08), and all the specified effects were significant.
+The illusion-specific latent factors were loaded positively by the sensitivity to illusion strength, as well as by the interaction effect with task difficulty (with the exception of Delboeuf, Ebbinghaus, Vertical-Horizontal, Müller-Lyer and Contrast, for which the loading was negative).
 The general factor of illusion sensitivity, labelled Factor \emph{i} (i- for illusion), explained 48.02\% of the total variance of the initial dataset, and was strongly related to Vertical-Horizontal (\(\beta_{std.}=0.83\)), Müller-Lyer (\(\beta_{std.}=0.76\)), Ponzo (\(\beta_{std.}=0.65\)), Ebbinghaus (\(\beta_{std.}=0.64\)); moderately to Zöllner (\(\beta_{std.}=0.53\)), Poggendorff (\(\beta_{std.}=0.44\)), Rod and Frame (\(\beta_{std.}=0.42\)), Contrast (\(\beta_{std.}=0.40\)) and White (\(\beta_{std.}=0.35\)); and weakly to Delboeuf (\(\beta_{std.}=0.19\)).
-We then computed, for each participant, its score for the 10 illusion-specific factors and for the general Factor \emph{i}.
+We then computed, for each participant, the score for the 10 illusion-specific factors and for the general Factor \emph{i}.
 
-We have to keep in mind that these individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of Delboeuf and Zöllner); 2) we only used the models on error rate, which could be biased by the speed-accuracy decision criterion used by participants; 3) the structural equation model used to compute the scores also incorporated multiple levels of abstractions.
+It is important to note that these individual scores are the result of several layers of simplification: 1) the individual coefficient is that of simpler models that sometimes do not perfectly capture the underlying dynamics (especially in the case of Delboeuf and Zöllner); 2) we only used the models on error rate, which could be biased by the speed-accuracy decision criterion used by participants; 3) the structural equation model used to compute the scores also incorporated multiple levels of abstractions.
 Thus, in order to validate the individual scores, we computed the correlation between them and simple empirical scores, such as the average error rate and the mean RT in the task.
-This analysis revealed strong and significant correlations between each illusion-specific factor and the average amount of errors in its respective task.
+This analysis revealed strong and significant correlations between each illusion-specific factor and the average amount of errors in its corresponding task.
 Moreover, each individual score was strongly associated with the average RT across multiple illusion types.
-This suggests that the individual scores obtained from the structural equation model do capture the sensitivity of each participant to visual illusions, manifesting in both the number of errors and high reaction times.
+This suggests that the individual scores obtained from the structural equation model do capture the sensitivity of each participant to visual illusions, manifesting in both the number of errors and long reaction times.
 
@@
 
-This study confirmed that it was possible to continuously manipulate the effect of illusion strength for 10 classical illusions.
+This study confirmed that it was possible to continuously manipulate the effect of illusion strength for 10 classic illusions.
 Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
-Future studies are needed to explore reaction times and try to identify the most appropriate models, and / or use models that integrate errors and reaction time (e.g., drift diffusion models).
+Future studies are needed to explore reaction times and identify the most appropriate models, and/or use models that integrate both errors and reaction time (e.g., drift diffusion models).
 
 The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
-For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with less errors than incongruent effects) were related to small and strong increases of errors, respectively.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with fewer errors than incongruent effects) were related to small and strong increases of errors, respectively.
 For the Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
-This was fixed in latest release of \emph{Pyllusion} (v1.2), which now generate outlines of the same size as the target circle.
-For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports (Kitaoka, 2007; Kitaoka \& Ishihara, 2000), suggesting an acute angle contraction effect at very small as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
+This was fixed in the latest release of \emph{Pyllusion} (v1.2), which now generates outlines of the same size as the target circle.
+For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports (Kitaoka, 2007; Kitaoka \& Ishihara, 2000), suggesting an acute angle contraction effect at very small - as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
 
@@
 The variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
-We used the \emph{IPIP6} (24 items, Sibley et al., 2011) to measure 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-humility), and the PID-5 (25 items, Hopwood et al., 2012) to measure ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+We used the \emph{IPIP6} (24 items, Sibley et al., 2011) to measure 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the PID-5 (25 items, Hopwood et al., 2012) to measure ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
 The participants were the same as in study 2 (see \textbf{Figure 3}).
@@
 \includegraphics[width=1\linewidth]{figures/Figure3} \caption{The upper plots show the distribution of demographic and dispositional variables.
-The middle plots show the relationship between illusion sensitivty scores, sex, and age (solid lines indicate significant relationships).
-Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \textit{i}), and personality traits.}\label{fig:unnamed-chunk-4}
+The middle plots shows the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
+Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \textit{i}) and personality traits.}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 For each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor \emph{i}), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) and personality.
-As the supplementary material contains the detailed results, we will here only report the significant results (based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see Makowski, Ben-Shachar, Chen, et al., 2019).
+As the supplementary material contains the detailed results, we will only report the significant results (based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see Makowski, Ben-Shachar, Chen, et al., 2019).
 
@@
 
-Regarding ``pathological'' personality traits, the results yielded strong evidence in favor of a negative relationship between multiple illusion scores and multiple traits.
+Regarding ``pathological'' personality traits, the results yielded strong evidence in favor of a negative relationship between illusion scores and multiple traits.
 \emph{Antagonism} was associated with the sensitivity to Vertical-Horizontal (\(BF_{10} > 100\)), Müller-Lyer (\(BF_{10} = 21.57\)), Ponzo (\(BF_{10} = 17.97\)) illusions, and the Factor \emph{i} (\(BF_{10} = 55.45\)); \emph{Psychoticism} was associated with the sensitivity to Vertical-Horizontal (\(BF_{10} = 66.63\)) and Müller-Lyer (\(BF_{10} = 35.59\)) illusions, and the Factor \emph{i} (\(BF_{10} = 35.02\)); \emph{Disinhibition} was associated with the sensitivity to Vertical-Horizontal (\(BF_{10} = 25.38\)), Zöllner (\(BF_{10} = 7.59\)), Müller-Lyer (\(BF_{10} = 5.89\)) illusions, and the Factor \emph{i} (\(BF_{10} = 31.42\)); and \emph{Negative Affect} was associated with Zöllner (\(BF_{10} = 62.04\)), Vertical-Horizontal (\(BF_{10} = 12.65\)), Müller-Lyer (\(BF_{10} = 3.17\)), and the Factor \emph{i} (\(BF_{10} = 6.39\)).
-The last remaining trait, \emph{Detachment}, did not share any relationship with illusion sensitivity.
+The last remaining trait, \emph{Detachment}, did not share any significant relationship with illusion sensitivity.
 
@@
 
-Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females reported to the Zöllner illusion.
-As we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity (Lo \& Dinov, 2011; Miller, 2001; Papageorgiou et al., 2020).
-This inconsistency could be due to past studies using a measure of illusion sensitvity that conflates the effect of illusions \emph{per se} with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
+Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females for the Zöllner illusion.
+Although we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity (Lo \& Dinov, 2011; Miller, 2001; Papageorgiou et al., 2020).
+This inconsistency could be due to past studies using a measure of illusion sensitivity that conflates the effect of illusions \emph{per se} with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
 On the contrary, the perceptual difficulty of the task and the illusion effect was independently modulated in our paradigm, and statistically dissociated.
@@
 Our findings also suggested a positive relationship between illusion sensitivity and two ``normal'' personality traits, namely \emph{Honesty-Humility} and \emph{Agreeableness}, and a negative link with \emph{Antagonism}.
-Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism (Konrath et al., 2009a; Pessoa et al., 2008; Zhang et al., 2017).
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism (Konrath et al., 2009; Pessoa et al., 2008; Zhang et al., 2017).
 While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
@@
 The positive relationship between illusion sensitivity and ``positive'' personality traits is mirrored by a negative relationship with several other pathological traits, including \emph{Psychoticism}, \emph{Disinhibition}, and \emph{Negative Affect}.
-These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing (including illusion sensitivity, Konrath et al., 2009b).
-For instance, pathological egocentric beliefs (often observed alongside \emph{Psychoticism}, Fox, 2006) have been related to reduced context integration (Fox, 2006; Konrath et al., 2009b; manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, Ohmann \& Burgmer, 2016).
+These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing (including illusion sensitivity, Konrath et al., 2009).
+For instance, pathological egocentric beliefs (often observed alongside \emph{Psychoticism}, Fox, 2006) have been related to reduced context integration (Fox, 2006; Konrath et al., 2009; manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, Ohmann \& Burgmer, 2016).
 As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits (Calamari et al., 2000; Hoyle, 2006; Msetfi et al., 2009; Parkes, 1981) .
@@
 Furthermore, these results in favour of a link between illusion sensitivity and maladaptive personality traits in a non-clinical population could be put in relation with clinical findings, which could be seen as extreme cases where the relationship with illusion sensitivity is the most manifest.
-In line with our results (in particular on \emph{Psychoticism} and \emph{Disinhibition}), prior research has found greater illusion resistance in schizophrenia (Grzeczkowski et al., 2018; Notredame et al., 2014; Pessoa et al., 2008), and in particular, in association with schizotypal traits, such as cognitive disorganization (Cretenoud et al., 2019b; Lányi et al., 2022).
+In line with our results (in particular on \emph{Psychoticism} and \emph{Disinhibition}), prior research has found greater illusion resistance in schizophrenia (Grzeczkowski et al., 2018; Notredame et al., 2014; Pessoa et al., 2008), and in particular, in association with schizotypal traits, such as cognitive disorganization (Cretenoud et al., 2019; Lányi et al., 2022).
 
@@
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
-This important methodological step opens the door for new illusions-based paradigms and tasks, to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
+This important methodological step opens the door for new illusions-based paradigms and tasks to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
 
-The participants' sensitivities to 10 different types of visual illusions shared a common part of variance, suggesting the existence of a general factor of illusion sensitivity (Factor \emph{i}).
+Our findings suggest that the sensitivity to 10 different types of visual illusions share a common part of variance, supporting the existence of a general factor of illusion sensitivity (Factor \emph{i}).
 This result comes in a field of mixed findings.
-In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals (Cretenoud, Francis, et al., 2020; Cretenoud et al., 2019b; Grzeczkowski et al., 2017, 2018; Yang et al., 2012).
-Instead, past findings suggests illusory effects are highly specific to the perceptual features of the illusions at stake (Cretenoud et al., 2019b; Grzeczkowski et al., 2017).
+In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals (Cretenoud et al., 2019; Cretenoud, Francis, et al., 2020; Grzeczkowski et al., 2017, 2018; Yang et al., 2012).
+Instead, past findings suggest illusory effects are highly specific to the perceptual features of the illusions at stake (Cretenoud et al., 2019; Grzeczkowski et al., 2017).
 It should be noted, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
-We believe that our study presents several methodological improvements, including statistical power (high number of trials per participant), homogeneous stimuli (with minimal and highly controlled features) and tasks (decision-making reaction-time task), and more reliable participant-level score extraction method (based on random-factors models), which in our opinion contributed to the emergence of the common factor.
+We believe that our study presents several methodological improvements, including statistical power (high number of trials per participant), homogeneous stimuli (with minimal and highly controlled features) and tasks (decision-making reaction-time task), and a more reliable participant-level score extraction method (based on random-factors models), which in our opinion contributed to the emergence of the common factor.
 
-However, although the illusions were relatively different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), The possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+However, although the illusions were relatively different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
 Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity, and assess the psychometric properties of similar paradigms, including stability (e.g., test-retest reliability) and validity.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020}{}}%
-Cretenoud, A. F., Francis, G., \& Herzog, M. H. (2020). When illusions merge. \emph{Journal of Vision}, \emph{20}(8), 12. \url{https://doi.org/10.1167/jov.20.8.12}
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020illusions}{}}%
+Cretenoud, A. F., Francis, G., \& Herzog, M. H. (2020). When illusions merge. \emph{Journal of Vision}, \emph{20}(8), 12--12.
 
-\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020a}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020individual}{}}%
 Cretenoud, A.
@@
 (2020).
-Individual differences in the Müller-Lyer and Ponzo illusions are stable across different contexts.
-\emph{Journal of Vision}, \emph{20}(6), 4.
-\url{https://doi.org/10.1167/jov.20.6.4}
+Individual differences in the m{ü}ller-lyer and ponzo illusions are stable across different contexts.
+\emph{Journal of Vision}, \emph{20}(6), 4--4.
 
@@
 H.
-(2019b).
-Factors underlying visual illusions are illusion-specific but not feature-specific.
-\emph{Journal of Vision}, \emph{19}(14), 12.
-\url{https://doi.org/10.1167/19.14.12}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2019a}{}}%
-Cretenoud, A.
-F., Karimpur, H., Grzeczkowski, L., Francis, G., Hamburger, K., \& Herzog, M.
-H.
-(2019a).
+(2019).
 Factors underlying visual illusions are illusion-specific but not feature-specific.
@@
 J., \& Grove, T.
-(2009a).
+(2009).
 Seeing my world in a million little pieces: Narcissism, self-construal, and cognitive--perceptual style.
 \emph{Journal of Personality}, \emph{77}(4), 1197--1228.
-
-\leavevmode\vadjust pre{\hypertarget{ref-konrath2009}{}}%
-Konrath, S., Bushman, B., \& Grove, T.
-(2009b).
-Seeing my world in a million little pieces: Narcissism, self-construal, and cognitive-perceptual style.
-\emph{Journal of Personality}, \emph{77}, 1197--1228.
-\url{https://doi.org/10.1111/j.1467-6494.2009.00579.x}
 

### 2022-09-08 manuscript.Rmd
Update manuscript.Rmd

--- 2022-09-07 manuscript.Rmd
@@
 We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task, from which we extracted participant-level performance scores (n=250).
-Our results provide evidence in favour of the existence of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions.
+Our results provide evidence in favour of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions.
 Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
+Data and code are available in open-access.
 
@@
 Participants were incentivised with a reward of about \textsterling 7.5 for completing the task, which took about 50 minutes to finish.
+Demographic variables (age, gender, and ethnicity) were self-reported on a voluntary basis.
 
@@
 
-We would like to thank Tam Pham and Zen J. Lau for their contribution to *Pyllusion*, as well as Prof Dólos for the inspiration.
+We would like to thank Zen J. Lau, Tam Pham, and W. Paul Boyce for their contribution to *Pyllusion*, as well as Prof Dólos for the inspiration.
 

### 2022-09-08 manuscript.Rmd
knit

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "5099"
+wordcount         : "5114"
 bibliography      : references.bib

### 2022-09-08 manuscript.tex
knit

@@
 We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task, from which we extracted participant-level performance scores (n=250).
-Our results provide evidence in favour of the existence of a general factor (labelled Factor \emph{i}) underlying the sensitivity to different illusions.
+Our results provide evidence in favour of a general factor (labelled Factor \emph{i}) underlying the sensitivity to different illusions.
 Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
+Data and code are available in open-access.
 }
@@
 
-We would like to thank Tam Pham and Zen J. Lau for their contribution to \emph{Pyllusion}, as well as Prof Dólos for the inspiration.
+We would like to thank Zen J. Lau, Tam Pham, and W. Paul Boyce for their contribution to \emph{Pyllusion}, as well as Prof Dólos for the inspiration.
 

### 2022-09-08 submission/manuscript.Rmd
ms

--- 2022-09-08 manuscript.Rmd
@@
 author:
+
   - name          : "Dominique Makowski"
@@
       - "Writing – review & editing"
+affiliation:
 
-affiliation:
   - id            : "1"
@@
     institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
-
 authornote: |
@@
 Data and code are available in open-access.
-
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
@@
 Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.", out.width="100%"}
-knitr::include_graphics("figures/Figure1.png")
+knitr::include_graphics("../figures/Figure1.png")
 ```
@@
 We fitted a Structural Equation Model (SEM) suggesting that these manifest variables group to first-level illusion-specific latent factors, which then load on a general factor of illusion sensitivity (Factor \\textit{i})."}
-knitr::include_graphics("figures/Figure2.png")
+knitr::include_graphics("../figures/Figure2.png")
 ```
@@
 Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \\textit{i}) and personality traits.", out.width="100%"}
-knitr::include_graphics("figures/Figure3.png")
+knitr::include_graphics("../figures/Figure3.png")
 ```

### 2022-09-08 submission/manuscript.tex
ms

--- 2022-09-08 manuscript.tex
@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 5099}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 5114}
 \usepackage{lineno}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures/Figure1} \caption{The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+\includegraphics[width=1\linewidth]{../figures/Figure1} \caption{The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
 Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.}\label{fig:unnamed-chunk-2}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures/Figure2} \caption{Top: the effect of illusion strength and task difficulty on the error rate and reaction time (RT) for each individual illusion.
+\includegraphics[width=1\linewidth]{../figures/Figure2} \caption{Top: the effect of illusion strength and task difficulty on the error rate and reaction time (RT) for each individual illusion.
 The solid line represents the General Additive Model (GAM), and the dashed line corresponds to its approximation via linear models.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures/Figure3} \caption{The upper plots show the distribution of demographic and dispositional variables.
+\includegraphics[width=1\linewidth]{../figures/Figure3} \caption{The upper plots show the distribution of demographic and dispositional variables.
 The middle plots shows the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).

### 2022-09-08 submission/manuscript.Rmd
minor

@@
 Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
-The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008] and in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
+The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], and in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
 
@@
 To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
-This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (see **Figure 1**).
+This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (**Figure 1**).
 
@@
 Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
-Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we selected and fitted the best model (based on their indices of fit), and compared their output visually (see **Figure 2**).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we selected and fitted the best model (based on their indices of fit), and compared their output visually (**Figure 2**).
 
@@
 
-For error rates, most of the models closely matched their GAMs counterpart (see **Figure 2**), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
+For error rates, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
 
@@
 We used the *IPIP6* [24 items, @sibley2011] to measure 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the PID-5 [25 items, @hopwood2012] to measure "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
-The participants were the same as in study 2 (see **Figure 3**).
+The participants were the same as in study 2 (**Figure 3**).
 However, due to a technical issue, no personality data was recorded for the first eight participants.

### 2022-09-20 submission/manuscript.Rmd
typo

@@
 Finally, we found that the sensitivity to illusions was positively associated with "positive" personality traits, such as *Agreeableness* and *Honesty-Humility*, and negatively associated with maladaptive traits such as *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
-Beyond highlighting the relevance of illusions beyond the field of visual perception, these results point towards an association with high-level general-domain mechanisms.
+Beyond highlighting the relevance of illusions beyond the field of visual perception, these results point towards an association with high-level domain-general mechanisms.
 While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.

### 2022-09-20 study1/study1.Rmd
update study1

@@
 ---
-title: "Illusion Game Validation (Study 1)"
-output: html_document
+title: "Illusion Game Validation (Pilot Study)"
+author: "Dominique Makowski et al."
+output:
+  html_document:
+    toc: true
+    toc_depth: 2
+    theme: paper
+    code_folding: hide
 date: "`r Sys.Date()`"
@@
   chunk_output_type: console
+bibliography: "../submission/references.bib"
 ---
@@
 
-# Preprocessing
+# Introduction
+
+This pilot experiment aimed at gathering some preliminary data to assess if the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors), and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
+
+In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+
+# Methods
+
+## Procedure
+
+We generated 56 stimuli for each of the 10 illusion types.
+These stimuli resulted from the combination of 8 linearly-spread levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where negative values correspond to congruent illusion strengths).
+
+The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the blocks was also randomized.
+After the first series of 10 blocks, another series was administered (with new randomized orders of blocks and trials).
+In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
+The task was implemented using *jsPsych* [@de2015jspsych], and the instructions for each illusion type are available in the experiment code.
 
@@
 
-We removed `r length(outliers)` participants upon inspection of the average error rage (when close to 50%, suggesting random answers) and/or when the reaction time distribution was implausibly fast.
-
-For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).
-
 ### Error Rate
+
+We removed `r length(outliers)` participants (in red) upon inspection of the average error rage (when close to 50%, suggesting random answers) and/or when the reaction time distribution was implausibly fast.
 
@@
 knitr::kable(dfsub) |> 
-  kableExtra::row_spec(which(dfsub$Participant %in% outliers), background  = "#EF9A9A")
+  kableExtra::row_spec(which(dfsub$Participant %in% outliers), background  = "#EF9A9A") |> 
+  kableExtra::kable_styling(full_width = TRUE) |> 
+  kableExtra::scroll_box(width = "100%", height = "500px")
 ```
@@
 ### Error Rate per Illusion Block
+
+For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).
 
@@
 
-```{r message=FALSE, warning=FALSE}
+Fifty-two participants were recruited via *Prolific* ([www.prolificacademic.co.uk](www.prolificacademic.co.uk)), a crowd-sourcing platform providing high data quality [@peer2022].
+The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
+Participants were incentivised with a reward of about \textsterling 7.5 for completing the task, which took about 50 minutes to finish.
+Demographic variables (age, gender, and ethnicity) were self-reported on a voluntary basis.
+
+We removed 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
+For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that instructions got misunderstood; e.g., participants were selecting the shorter line instead of the longer one).
+Finally, we removed 692 (1.37%) trials based on an implausibly short or long response time (\< 150 ms or \> 3000 ms).
+
+The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: [19, 60]; Sex: 39.1% females, 56.5% males, and 4.4% other).
+
+```{r message=FALSE, warning=FALSE, results='hide'}
 dfsub <- df |>
@@
   ungroup()
-```
-
-The final sample included `r report::report_participants(dfsub, age="Age", sex="Sex")`.
+
+report::report_participants(dfsub, age="Age", sex="Sex")
+```
 
@@
 
+## Data Analysis
+
+The analysis focused on the probability of errors as the main outcome variable.
+For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some intuition on the underlying generative model.
+Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
+We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potential non-linear relationships (at the expense of model simplicity).
+
+The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle].
+
 # Results {.tabset}
+
+## Summary
+
+The statistical models suggested that the effect of task difficulty had a cubic relationship with error rate for the Delboeuf and Ebbinghaus illusions (both composed of circular shapes), square relationship for the Rod and Frame and Vertical-Horizontal illusions, cubic relationship for the Zöllner and Poggendorff illusions, exponential relationship for the White illusion, cubic relationship for the Müller-Lyer and Ponzo illusions (both based on line lengths), and linear relationship for the Contrast illusion.
+All models suggested a significant effect of illusion strength and task difficulty.
+See details and figures in the analysis script.
 
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r delboeuf_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models <- function(data) {
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r delboeuf_model, message=FALSE, warning=FALSE, cache=cache}
 cbrt <- function(x) sign(x) * abs(x)**(1/3)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r delboeuf_gam, message=FALSE, warning=FALSE, cache=cache}
 make_gam <- function(data) {
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r ebbinghaus_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r ebbinghausmodel, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r ebbinghaus_gam, message=FALSE, warning=FALSE, cache=cache}
 gam <- make_gam(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r rodframe_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r rodframe_model, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r rodframe_gam, message=FALSE, warning=FALSE, cache=cache}
 gam <- make_gam(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r verticalhorizontal_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r verticalhorizontal_model, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r verticalhorizontal_gam, message=FALSE, warning=FALSE, cache=cache}
 gam <- make_gam(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r zollner_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r zollner_model, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r white_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r white_model, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r white_gam, message=FALSE, warning=FALSE, cache=cache}
 gam <- make_gam(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r mullerlyer_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r mullerlyer_model, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r mullerlyer_gam, message=FALSE, warning=FALSE, cache=cache}
 gam <- make_gam(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r ponzo_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r ponzo_model, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r ponzo_gam, message=FALSE, warning=FALSE, cache=cache}
 gam <- make_gam(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r poggendorff_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r poggendorff_model, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r poggendorff_gam, message=FALSE, warning=FALSE, cache=cache}
 gam <- make_gam(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r contrast_bestmodels, message=FALSE, warning=FALSE, cache=cache}
 best_models(data)
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r contrast_model, message=FALSE, warning=FALSE, cache=cache}
 formula <- brms::bf(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r contrast_gam, message=FALSE, warning=FALSE, cache=cache}
 gam <- make_gam(data)
@@
 ```
+
+# Discussion
+
+This pilot study provided a clearer understanding of the magnitude of the parametric effects at stake and the type of interaction between them.
+Furthermore, it allowed us to better understand and test the stimuli generated by *Pyllusion*, as well as uncover incidental bugs and technical issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed in a new software release.
+Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
+
+In most illusions, the task difficulty exhibited monotonic power-law scaled effects, which is in line with the psychophysics literature on perceptual decisions [@ditzinger2010; @shekhar2021; @bogacz2006].
+One notable result was the illusion effect pattern for the Zöllner illusion, which suggested a non-linear relationship.
+By generating a wider range of illusion strength values, the [**next study**](https://github.com/RealityBending/IllusionGameValidation) will attempt at clarifying this point.
+
+# References

### 2022-09-20 study2/study2.Rmd
update study1

@@
   kableExtra::row_spec(which(dfsub$Participant %in% outliers) + 1, background = "#EF9A9A") |> 
-  kableExtra::row_spec(which(dfsub$Participant %in% partial_outliers) + 1, background = "#FFECB3")
+  kableExtra::row_spec(which(dfsub$Participant %in% partial_outliers) + 1, background = "#FFECB3") |> 
+  kableExtra::kable_styling(full_width = TRUE) |> 
+  kableExtra::scroll_box(width = "100%", height = "500px")
 ```

### 2022-09-20 study2/study2.Rmd
reorganize into 1 study

@@
 ---
-title: "Illusion Game Validation (Study 2)"
-output: html_document
+title: "Illusion Game Validation (part 1)"
+author: "Dominique Makowski et al."
+output:
+  html_document:
+    toc: true
+    toc_depth: 2
+    theme: paper
+    code_folding: hide
 date: "`r Sys.Date()`"
@@
 
-# Preprocessing
+# Introduction
+
+The aim of part 1 is two-fold.
+In the first part, we will carefully model the error rate and the reaction time of each illusion type in order to validate our novel paradigm and show that the effect of illusions can be manipulated continuously.
+In the second part, we will derive the participant-level scores from the models (i.e., the effect of illusion strength for each individual) and analyze their latent factors structure.
+
+# Methods
 
@@
 
-### Summary
+### Descriptive Table
 
@@
 ```
-
-See the next tab for the descriptive table.
-
-### Descriptive Table
 
@@
 ```
+
+# Discussion
+
+This study confirmed that it was possible to continuously manipulate the effect of illusion strength for 10 classic illusions.
+Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
+Future studies are needed to explore reaction times and identify the most appropriate models, and/or use models that integrate both errors and reaction time (e.g., drift diffusion models).
+
+The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
+For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with fewer errors than incongruent effects) were related to small and strong increases of errors, respectively.
+For the Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
+This was fixed in the latest release of *Pyllusion* (v1.2), which now generates outlines of the same size as the target circle.
+For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports [@kitaoka2000; @kitaoka2007], suggesting an acute angle contraction effect at very small - as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
+
+Finally, this study provided evidence for both the existence of illusion-specific factors, as well as for a common latent factor (labelled Factor *i*) that explained about half of the total variance.
+These participant-level scores were positively related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
+
+# References

### 2022-09-20 study3/study3.Rmd
reorganize into 1 study

@@
 ---
-title: "Illusion Game Validation (Study 3)"
-output: html_document
+title: "Illusion Game Validation (part 2)"
+author: "Dominique Makowski et al."
+output:
+  html_document:
+    toc: true
+    toc_depth: 2
+    theme: paper
+    code_folding: hide
 date: "`r Sys.Date()`"
@@
 
-The goal of study 3 was to...
+Part 2 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), contextual variables (pertaining to the experiment setting), such as screen size, demographic features (such as sex and age), and stable dispositional variables such as "general" personality traits.
+Indeed, despite the abundant literature on visual illusions, relatively few studies have investigated its ties with participants' characteristics.
+Research examining the influence of demographic variables such as gender and age have generally found inconsistent results [@cretenoud2020individual; @grzeczkowski2017; @lo2011; @papageorgiou2020].
+Regarding links with personality, most works focused on traits associated with psychopathology, such as impulsivity or sensation-seeking [@hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
 
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r sex_stats, message=FALSE, warning=FALSE, cache=cache}
 sig <- list()
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r education_stats, message=FALSE, warning=FALSE, cache=cache}
 preds <- data.frame()
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache}
+```{r age_stats, message=FALSE, warning=FALSE, cache=cache}
 preds <- data.frame()
@@
 p_age
+```
 
+```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
 ggsave("figures/figure_agesex.png",
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
+```{r ipip_stats, message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 cor <- make_correlation(
@@
 
-```{r message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
+```{r pid_stats, message=FALSE, warning=FALSE, cache=cache, eval=TRUE}
 cor <- make_correlation(
@@
 ```
+
+# Discussion
+
+Indeed, despite the abundant literature on visual illusions, relatively few studies have investigated its ties with participants' characteristics.
+Research examining the influence of demographic variables such as gender and age have generally found inconsistent results [@cretenoud2020individual; @grzeczkowski2017; @lo2011; @papageorgiou2020].
+Regarding links with personality, most works focused on traits associated with psychopathology, such as impulsivity or sensation-seeking [@hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
+
+<!-- Context -->
+
+We report significant links between inter-individual indices of illusion sensitivity and variables related to experimental context, demographic characteristics and personality.
+Firstly, screen size was found to have a significant negative relationship with the sensitivity to the two contrast-based illusions, namely the White and Contrast illusions.
+One possible explanation can be found in the mechanism by which visual systems filter through more low spatial frequencies when the size of the target object is small [@dixon2014].
+As this filtering process excludes illumination information from visual processing, smaller screen sizes could yield artifactual changes in brightness perception, which in turn could attenuate the illusory effect of luminance-related illusions.
+
+<!-- Demographic -->
+
+Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females for the Zöllner illusion.
+Although we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity [@lo2011; @miller2001; @papageorgiou2020].
+This inconsistency could be due to past studies using a measure of illusion sensitivity that conflates the effect of illusions *per se* with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
+On the contrary, the perceptual difficulty of the task and the illusion effect was independently modulated in our paradigm, and statistically dissociated.
+Our scores of illusion sensitivity might thus be less loaded with perceptual skills, thereby mitigating its effect.
+
+<!-- Normal personality + Antagonism -->
+
+Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*, and a negative link with *Antagonism*.
+Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing; @pessoa2008].
+While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
+Since narcissism and aggression tendencies are correlated with lower field dependence [i.e., a lesser reliance on external cues in ambiguous contexts, @witkin1976; @ohmann2016; @damour2021], opposite traits, such as *Honesty-Humility* and *Agreeableness*, could conversely be more biased by contextual cues and thus more sensitive to illusions.
+
+<!-- Pathological personality -->
+
+The positive relationship between illusion sensitivity and "positive" personality traits is mirrored by a negative relationship with several other pathological traits, including *Psychoticism*, *Disinhibition*, and *Negative Affect*.
+These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing [including illusion sensitivity, @konrath2009seeing].
+For instance, pathological egocentric beliefs [often observed alongside *Psychoticism*, @fox2006] have been related to reduced context integration [manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, @ohmann2016; @konrath2009seeing; @fox2006].
+As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits [@msetfi2009; @parkes1981field; @calamari2000field; @hoyle2006] .
+
+Furthermore, these results in favour of a link between illusion sensitivity and maladaptive personality traits in a non-clinical population could be put in relation with clinical findings, which could be seen as extreme cases where the relationship with illusion sensitivity is the most manifest.
+In line with our results (in particular on *Psychoticism* and *Disinhibition*), prior research has found greater illusion resistance in schizophrenia [@notredame2014; @pessoa2008; @grzeczkowski2018], and in particular, in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019; @lanyi2022can].
+
+# References

### 2022-09-20 submission/manuscript.Rmd
reorganize into 1 study

@@
 Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings.
-In this set of studies, we used the parametric framework for visual illusions implemented in the *Pyllusion* software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength.
-We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task, from which we extracted participant-level performance scores (n=250).
+In this study, we used the novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
+We tested the objective effect of the illusions on errors and reaction times, and extracted participant-level performance scores (n=250).
 Our results provide evidence in favour of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions.
 Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
-Data and code are available in open-access.
+The experiment, data and code are fully available for re-use at https://github.com/RealityBending/IllusionGameValidation.
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
@@
 
-In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously, and separately modeled statistically.
-Then, we will further utilize the paradigm to assess whether 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor.
-Finally, we will investigate how the the inter-individual sensitivity to illusions relates to dispositional variables, such as demographic characteristics and personality.
+The aim of the present preregistered study is three-fold.
+First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
+Next, we will investigate the latent factors structure of participant-level performance scores to these illusions and test the existence of a common factor of illusion sensitivity.
+Finally, we will explore how illusion sensitivity relates to demographic characteristics, contextual variables (pertaining to the experiment setting), and personality traits.
 
-In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as **Supplementary Materials** at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
 
-# Study 1
+# Methods
 
-## Aim
+## Stimuli
 
-Study 1 can be seen as a pilot experiment aiming to gather some preliminary data to assess if the stimuli generated by *Pyllusion* behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors), and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
+A pilot study (*n = 46*), which full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
+For each of the 10 illusion types, we generated a total of 134 stimuli.
+These stimuli resulted from the combination of 15 equally-spaced levels of illusion *strength* (7 negative, i.e., congruent effects; 7 positive, i.e., incongruent effects; and 0) overlapped with 16 non-linearly spaced task *difficulty* levels (i.e., with an exponential, square or cubic spacing depending on the pilot results).
+For instance, a linear space of [0.1, 0.4, 0.7, 1.0] can be transformed to an exponential space of [0.1, 0.34, 0.64, 1.0], where 0.1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets).
+For each illusion type, the stimuli were split into two series (56 and 72 stimuli per group) with alternating parameter values to maintain their homogeneity.
 
@@
 
-We generated 56 stimuli for each of the 10 illusion types.
-These stimuli resulted from the combination of 8 linearly-spread levels of task difficulty (e.g., [1, 2, 3, 4, 5, 6, 7], where 1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., [-3, -2, -1, 0, 1, 2, 3], where negative values correspond to congruent illusion strengths).
-
-The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the blocks was also randomized.
-After the first series of 10 blocks, another series was administered (with new randomized orders of blocks and trials).
-In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
+After a brief demographic survey, the first series of 10 illusion blocks was presented in a randomized order, with a randomized order of the stimuli within the blocks.
+Following the first series, two personality questionnaires were administered, the *IPIP6* [24 items, @sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the PID-5 [25 items, @hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
+In total, each participant underwent 1340 trials to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
 The task was implemented using *jsPsych* [@de2015jspsych], and the instructions for each illusion type are available in the experiment code.
@@
 
-Fifty-two participants were recruited via *Prolific* ([www.prolificacademic.co.uk](www.prolificacademic.co.uk)), a crowd-sourcing platform providing high data quality [@peer2022].
+Participants were recruited via *Prolific*, a crowd-sourcing platform recognized for providing high quality data [@peer2022].
 The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
@@
 
-We removed 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
-For the remaining participants, we discarded blocks where the error rate was higher than 50% (possibly indicating that instructions got misunderstood; e.g., participants were selecting the shorter line instead of the longer one).
-Finally, we removed 692 (1.37%) trials based on an implausibly short or long response time (\< 150 ms or \> 3000 ms).
+We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
+For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions got misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean).
+Additional, due to a technical issue, no personality data was recorded for the first eight participants.
 
-The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: [19, 60]; Sex: 39.1% females, 56.5% males, and 4.4% other).
-
-## Data Analysis
-
-The analysis of study 1 focused on the probability of errors as the main outcome variable.
-For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some intuition on the underlying generative model.
-Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up *vs.* down or left *vs.* right) as an additional predictor.
-We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potential non-linear relationships (at the expense of model simplicity).
-
-The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle].
-
-## Results
-
-The statistical models suggested that the effect of task difficulty had a cubic relationship with error rate for the Delboeuf and Ebbinghaus illusions (both composed of circular shapes), square relationship for the Rod and Frame and Vertical-Horizontal illusions, cubic relationship for the Zöllner and Poggendorff illusions, exponential relationship for the White illusion, cubic relationship for the Müller-Lyer and Ponzo illusions (both based on line lengths), and linear relationship for the Contrast illusion.
-All models suggested a significant effect of illusion strength and task difficulty.
-See details and figures in the analysis script.
-
-## Discussion
-
-This study provided a clearer understanding of the magnitude of the parametric effects at stake and the type of interaction between them.
-Furthermore, it allowed us to better understand and test the stimuli generated by *Pyllusion*, as well as uncover incidental bugs and technical issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed in a new software release.
-Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
-
-In most illusions, the task difficulty exhibited monotonic power-law scaled effects, which is in line with the psychophysics literature on perceptual decisions [@ditzinger2010; @shekhar2021; @bogacz2006].
-One notable result was the illusion effect pattern for the Zöllner illusion, which suggested a non-linear relationship.
-By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.
-
-# Study 2
-
-## Aim
-
-The aim of study 2 was two-fold.
-In the first part, we carefully modeled the error rate and the reaction time of each illusion type in order to validate our novel paradigm and show that the effect of illusions can be manipulated continuously.
-In the second part, we derived the participant-level scores from the models (i.e., the effect of illusion strength for each individual) and analyzed their latent factors structure.
-
-## Procedure
-
-The paradigm of study 2 was similar to that of study 1, with the following changes: the illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
-Moreover, taking into account the findings of study 1, we used non-linearly spaced difficulty levels, depending on the best underlying model (i.e., with an exponential, square or cubic spacing depending on the relationship).
-For instance, a linear space of [0.1, 0.4, 0.7, 1.0] can be transformed to an exponential space of [0.1, 0.34, 0.64, 1.0].
-
-Additionally, instead of repeating each stimulus two times, we generated illusions using more levels of difficulty and illusion strength.
-As such, for each illusion type, we generated a total of 134 stimuli that were split into two groups (67 stimuli per illusion block).
-Furthermore, instead of a simple break screen, we added two personality questionnaires between the two series of 10 illusion blocks (see study 3).
-
-## Participants
-
-Using the same recruitment procedure as in study 1, we recruited 256 participants, out of which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
-Please see study 3 for the full demographic breakdown.
-We discarded blocks with more than 50% of errors (2.16% of trials) and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean).
+The final sample included 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
 
@@
 The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT) within each illusion.
-In order to achieve this, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
+We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions.
 Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].
 
-Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
-Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we selected and fitted the best model (based on their indices of fit), and compared their output visually (**Figure 2**).
+Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear mixed models, which can be used to estimate the effects' participant-level variability (via random slopes).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we fitted the best model (based on their indices of fit), and compared their output visually (**Figure 2**).
 
-We then extracted the inter-individual variability in the effect of illusion strength and its interaction with task difficulty, and used it as participant-level scores.
-Finally, we explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
+We extracted the inter-individual variability in the effect of illusion strength and its interaction with task difficulty, and used it as participant-level scores.
+We explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and test the existence of a general of illusion sensitivity (Factor *i*).
 
-## Results
+Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) and personality.
 
-The best models were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; $cbrt(diff)*strength$ for Zöllner; $diff*sqrt(strength)$ and $log(diff)*strength$ respectively for errors and RT in White; $sqrt(diff)*sqrt(strength)$ and $sqrt(diff)*strength$ respectively for errors and RT in Müller-Lyer; $cbrt(diff)*strength$ for Ponzo; $cbrt(diff)*sqrt(strength)$ and $cbrt(diff)*strength$ respectively for errors and RT in Poggendorff; and $sqrt(diff)*sqrt(strength)$ for Contrast.
-For all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
-
-For error rates, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
+The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle].
+As the full results are available as supplementary materials, we will focus here on the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
 
@@
 
+# Results
+
+## Effects of Illusion Strength and Task Difficulty
+
+The best model specifications were $log(diff)*strength$ for Delboeuf; $sqrt(diff)*strength$ for Ebbinghaus; $log(diff)*log(strength)$ for Rod and Frame; $sqrt(diff)*sqrt(strength)$ for Vertical-Horizontal; $cbrt(diff)*strength$ for Zöllner; $diff*sqrt(strength)$ and $log(diff)*strength$ respectively for errors and RT in White; $sqrt(diff)*sqrt(strength)$ and $sqrt(diff)*strength$ respectively for errors and RT in Müller-Lyer; $cbrt(diff)*strength$ for Ponzo; $cbrt(diff)*sqrt(strength)$ and $cbrt(diff)*strength$ respectively for errors and RT in Poggendorff; and $sqrt(diff)*sqrt(strength)$ for Contrast.
+For all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
+
+For error rates, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
+
 For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants respond faster.
@@
 
-Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy (and are also more conservative estimates due to shrinkage) than basic empirical scores, such as the total number of errors, or the average RT.
+## Factor Structure
+
+Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy (and also provide more conservative estimates due to shrinkage) than basic empirical scores, such as the total number of errors, or the average RT.
 Thus, for each illusion and within each participant, we extracted the effect of illusion strength and its interaction with task difficulty when the illusion effect was incongruent.
@@
 
-## Discussion
+## Correlations with Inter-individual Characteristics
 
-This study confirmed that it was possible to continuously manipulate the effect of illusion strength for 10 classic illusions.
-Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
-Future studies are needed to explore reaction times and identify the most appropriate models, and/or use models that integrate both errors and reaction time (e.g., drift diffusion models).
-
-The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
-For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with fewer errors than incongruent effects) were related to small and strong increases of errors, respectively.
-For the Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
-This was fixed in the latest release of *Pyllusion* (v1.2), which now generates outlines of the same size as the target circle.
-For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports [@kitaoka2000; @kitaoka2007], suggesting an acute angle contraction effect at very small - as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
-
-Finally, this study provided evidence for both the existence of illusion-specific factors, as well as for a common latent factor (labelled Factor *i*) that explained about half of the total variance.
-These participant-level scores were positively related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
-
-# Study 3
-
-## Aim
-
-Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), contextual variables (pertaining to the experiment setting), such as screen size, demographic features (such as sex and age), and stable dispositional variables such as "general" personality traits.
-Indeed, despite the abundant literature on visual illusions, relatively few studies have investigated its ties with participants' characteristics.
-Research examining the influence of demographic variables such as gender and age have generally found inconsistent results [@cretenoud2020individual; @grzeczkowski2017; @lo2011; @papageorgiou2020].
-Regarding links with personality, most works focused on traits associated with psychopathology, such as impulsivity or sensation-seeking [@hlavata2018; @zhang2017; @lanyi2022can; @pessoa2008].
-
-## Procedure
-
-This study was based on the data collected in study 2.
-The variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
-We used the *IPIP6* [24 items, @sibley2011] to measure 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the PID-5 [25 items, @hopwood2012] to measure "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
-The participants were the same as in study 2 (**Figure 3**).
-However, due to a technical issue, no personality data was recorded for the first eight participants.
+The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions (**Figure 3**).
+Anecdotal ($BF_{10} = 2.05$) and moderate evidence ($BF_{10} = 4.11$) was found for a negative correlation between screen size and the sensitivity to the White and the Contrast illusion, respectively.
+To test whether this result could be an artifact related to the highly skewed screen size distribution (caused by very few participants with extreme screen sizes), we re-ran a robust correlation (with rank-transformed values), which provided even stronger evidence in favor of the effect existence ($BF_{10} = 28.19$, $BF_{10} = 4.31$ for White and Contrast, respectively).
 
@@
 ```
-
-## Data Analysis
-
-For each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) and personality.
-As the supplementary material contains the detailed results, we will only report the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
-
-## Results
-
-The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions.
-Anecdotal ($BF_{10} = 2.05$) and moderate evidence ($BF_{10} = 4.11$) was found for a negative correlation between screen size and the sensitivity to the White and the Contrast illusion, respectively.
-To test whether this result could be an artifact related to the highly skewed screen size distribution (caused by very few participants with extreme screen sizes), we re-ran a robust correlation (with rank-transformed values), which provided even stronger evidence in favor of the effect existence ($BF_{10} = 28.19$, $BF_{10} = 4.31$ for White and Contrast, respectively).
 
@@
 
-## Discussion
-
-<!-- Context -->
-
-We report significant links between inter-individual indices of illusion sensitivity and variables related to experimental context, demographic characteristics and personality.
-Firstly, screen size was found to have a significant negative relationship with the sensitivity to the two contrast-based illusions, namely the White and Contrast illusions.
-One possible explanation can be found in the mechanism by which visual systems filter through more low spatial frequencies when the size of the target object is small [@dixon2014].
-As this filtering process excludes illumination information from visual processing, smaller screen sizes could yield artifactual changes in brightness perception, which in turn could attenuate the illusory effect of luminance-related illusions.
-
-<!-- Demographic -->
-
-Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females for the Zöllner illusion.
-Although we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity [@lo2011; @miller2001; @papageorgiou2020].
-This inconsistency could be due to past studies using a measure of illusion sensitivity that conflates the effect of illusions *per se* with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
-On the contrary, the perceptual difficulty of the task and the illusion effect was independently modulated in our paradigm, and statistically dissociated.
-Our scores of illusion sensitivity might thus be less loaded with perceptual skills, thereby mitigating its effect.
-
-<!-- Normal personality + Antagonism -->
-
-Our findings also suggested a positive relationship between illusion sensitivity and two "normal" personality traits, namely *Honesty-Humility* and *Agreeableness*, and a negative link with *Antagonism*.
-Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism [@zhang2017; @konrath2009seeing; @pessoa2008].
-While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
-Since narcissism and aggression tendencies are correlated with lower field dependence [i.e., a lesser reliance on external cues in ambiguous contexts, @witkin1976; @ohmann2016; @damour2021], opposite traits, such as *Honesty-Humility* and *Agreeableness*, could conversely be more biased by contextual cues and thus more sensitive to illusions.
-
-<!-- Pathological personality -->
-
-The positive relationship between illusion sensitivity and "positive" personality traits is mirrored by a negative relationship with several other pathological traits, including *Psychoticism*, *Disinhibition*, and *Negative Affect*.
-These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing [including illusion sensitivity, @konrath2009seeing].
-For instance, pathological egocentric beliefs [often observed alongside *Psychoticism*, @fox2006] have been related to reduced context integration [manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, @ohmann2016; @konrath2009seeing; @fox2006].
-As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits [@msetfi2009; @parkes1981field; @calamari2000field; @hoyle2006] .
-
-Furthermore, these results in favour of a link between illusion sensitivity and maladaptive personality traits in a non-clinical population could be put in relation with clinical findings, which could be seen as extreme cases where the relationship with illusion sensitivity is the most manifest.
-In line with our results (in particular on *Psychoticism* and *Disinhibition*), prior research has found greater illusion resistance in schizophrenia [@notredame2014; @pessoa2008; @grzeczkowski2018], and in particular, in association with schizotypal traits, such as cognitive disorganization [@cretenoud2019; @lanyi2022can].
-
-# General Discussion
+# Discussion
 

### 2022-09-20 submission/manuscript.Rmd
manu

@@
 
-Notably, the presence of a common mechanism underlying the effect of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
+Notably, the presence of a common mechanism underlying the effects of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
 The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], and in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
@@
 One key challenge hindering the further development of illusion research is the relative difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
+To address this issue, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
 This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (**Figure 1**).
@@
 
-Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
+Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths, dependent on whether they had inwards or outwards pointing arrows at the ends).
 Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to match a reference stimulus relying only on their perception [@mylniec2016; @grzeczkowski2018].
@@
 
-A pilot study (*n = 46*), which full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
-For each of the 10 illusion types, we generated a total of 134 stimuli.
+A pilot study (*n = 46*), of which full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
+For each of the 10 illusion types, we generated a total of 134 stimuli (including 6 stimuli for practice trials).
 These stimuli resulted from the combination of 15 equally-spaced levels of illusion *strength* (7 negative, i.e., congruent effects; 7 positive, i.e., incongruent effects; and 0) overlapped with 16 non-linearly spaced task *difficulty* levels (i.e., with an exponential, square or cubic spacing depending on the pilot results).
 For instance, a linear space of [0.1, 0.4, 0.7, 1.0] can be transformed to an exponential space of [0.1, 0.34, 0.64, 1.0], where 0.1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets).
-For each illusion type, the stimuli were split into two series (56 and 72 stimuli per group) with alternating parameter values to maintain their homogeneity.
+For each illusion type, the stimuli were split into two series (56 and 72 stimuli respectively) with alternating parameter values to maintain their homogeneity.
 
@@
 
-After a brief demographic survey, the first series of 10 illusion blocks was presented in a randomized order, with a randomized order of the stimuli within the blocks.
-Following the first series, two personality questionnaires were administered, the *IPIP6* [24 items, @sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the PID-5 [25 items, @hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+After a brief demographic survey, the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order for trials within each block.
+Following this first series of blocks, two personality questionnaires were administered, the *IPIP6* [24 items, @sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the *PID-5* [25 items, @hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
 Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
-In total, each participant underwent 1340 trials to which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
-The task was implemented using *jsPsych* [@de2015jspsych], and the instructions for each illusion type are available in the experiment code.
+In total, each participant underwent 1340 trials of which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
+The task was implemented using *jsPsych* [@de2015jspsych], and the set of instructions for each illusion type is available in the experiment code.
 
@@
 
-We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and when the reaction time distribution was implausibly fast.
-For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions got misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean).
-Additional, due to a technical issue, no personality data was recorded for the first eight participants.
+We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when it was implausibly fast).
+For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean).
+Additionally, due to a technical issue, no personality data was recorded for the first eight participants.
 
@@
 
-We extracted the inter-individual variability in the effect of illusion strength and its interaction with task difficulty, and used it as participant-level scores.
-We explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and test the existence of a general of illusion sensitivity (Factor *i*).
+The inter-individual variability in the effect of illusion strength and its interaction with task difficulty, were extracted and used as participant-level scores.
+We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general of illusion sensitivity (Factor *i*).
 
-Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) and personality.
+Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) as well as personality traits.
 
@@
 
-For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants respond faster.
+For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants responded faster.
 While this is is not surprising (strong illusions are likely so effective in biasing perception that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
@@
 
-The parametric illusion generation framework developed in @makowski2021 proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
+The parametric illusion generation framework developed in @makowski2021 proposes to conceptualize illusions as being composed of targets and distractors that can be manipulated independently and continuously.
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
-This important methodological step opens the door for new illusions-based paradigms and tasks to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
+Specifically, increasing the illusion strength led to an increase in the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
+As such, this important methodological step not only helps us move away from traditional illusion research paradigms which are highly dependent on participants' subjective experiences of illusory effects; opening the door for new illusions-based paradigms and tasks, it also allows us to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
+For instance, using this novel framework accorded by *Pyllusion* to objectively manipulate the parameters of the Zöllner illusion, we were able to find an effect of illusion strength on errors that was distinct from the other 9 illusions examined.
+For a more detailed discussion on this specific illusion effect, please refer to the Supplementary Materials (Part 2 - Discussion).
 
@@
 
-However, although the illusions were relatively different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
-Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity, and assess the psychometric properties of similar paradigms, including stability (e.g., test-retest reliability) and validity.
+However, although the nature of the illusions were different in terms of the perceptual task underlying the illusory effect (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and general illusion sensitivity, and assess the psychometric properties of similar paradigms, including their stability (e.g., test-retest reliability) and validity.
 
@@
 
-Finally, we found that the sensitivity to illusions was positively associated with "positive" personality traits, such as *Agreeableness* and *Honesty-Humility*, and negatively associated with maladaptive traits such as *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
-Beyond highlighting the relevance of illusions beyond the field of visual perception, these results point towards an association with high-level domain-general mechanisms.
-While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
+Finally, we found the sensitivity to illusions to be positively associated with "positive" personality traits, such as *Agreeableness* and *Honesty-Humility*, and negatively associated with maladaptive traits such as *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
+Although empirical evidence proposing significant links between illusion sensitivity and maladaptive personality traits are scarce, these results are consistent with past findings relating pathological egocentric beliefs (often observed alongside *Psychoticism*, @fox2006) to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli[ [@ohmann2016; @konrath2009seeing; @fox2006].
+As such, the association between maladaptive traits and lower illusion sensitivity could be linked to a self-centered, decontextualized and disorganized information processing style.
+Conversely, the relationship between sensitivity to illusions and "positive" personality traits could be drawn from research relating cognitive styles (i.e., field dependence - a tendency to rely less on external cues in ambiguous contexts) to dispositional traits like hostility, aggression and narcissism [@zhang2017; @konrath2009seeing; @pessoa2008], which negatively correlate with "positive" traits like *Agreeableness* and *Honesty-Humility* (see Supplementary Materials, Discussion - Part 3 for an in-depth discussion on the relationship between personality traits, cognitive styles and illusion sensitivity).
+
+Importantly, such findings highlight the relevance of illusions beyond the field of visual perception, pointing towards an association with high-level domain-general mechanisms.
+In particular, the significant findings in favor of a relationship between maladaptive personality traits and illusion sensitivity could be viewed in relation to clinical findings, of which greater resistance to illusions have often been reported among patients with schizophrenia [@notredame2014; @pessoa2008; @grzeczkowski2018], especially in association with schizotypal traits such as cognitive disorganization [@cretenoud2019; @lanyi2022can].
+While the search for the exact mechanism(s) underlying these links is an important goal of future research, our observations unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
 

### 2022-09-22 submission/manuscript.Rmd
donzo

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "5114"
+wordcount         : "3621"
 bibliography      : references.bib
@@
 
-Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths, dependent on whether they had inwards or outwards pointing arrows at the ends).
+Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths if they end with inwards vs.
+outwards pointing arrows).
 Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to match a reference stimulus relying only on their perception [@mylniec2016; @grzeczkowski2018].
@@
 A pilot study (*n = 46*), of which full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
-For each of the 10 illusion types, we generated a total of 134 stimuli (including 6 stimuli for practice trials).
+Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
 These stimuli resulted from the combination of 15 equally-spaced levels of illusion *strength* (7 negative, i.e., congruent effects; 7 positive, i.e., incongruent effects; and 0) overlapped with 16 non-linearly spaced task *difficulty* levels (i.e., with an exponential, square or cubic spacing depending on the pilot results).
 For instance, a linear space of [0.1, 0.4, 0.7, 1.0] can be transformed to an exponential space of [0.1, 0.34, 0.64, 1.0], where 0.1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets).
-For each illusion type, the stimuli were split into two series (56 and 72 stimuli respectively) with alternating parameter values to maintain their homogeneity.
+For each illusion type, the stimuli were split into two series (56 and 72 stimuli per series) with alternating parameter values to maintain their homogeneity.
+Additionally, 6 stimuli per illusion type was generated for a practice series, with more extreme variations (i.e., containing very easy trials to help cement the task instructions).
 
@@
 
-After a brief demographic survey, the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order for trials within each block.
+After a brief demographic survey and a practice series of illusions, the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
 Following this first series of blocks, two personality questionnaires were administered, the *IPIP6* [24 items, @sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the *PID-5* [25 items, @hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
 Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
-In total, each participant underwent 1340 trials of which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
+In total, each participant underwent 1340 trials of which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key  (left/right,or up/down depending on the illusion type).
+For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
 The task was implemented using *jsPsych* [@de2015jspsych], and the set of instructions for each illusion type is available in the experiment code.
@@
 
-We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when it was implausibly fast).
+We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when implausibly fast).
 For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean).
@@
 
-The inter-individual variability in the effect of illusion strength and its interaction with task difficulty, were extracted and used as participant-level scores.
+The inter-individual variability in the effect of illusion strength and its interaction with task difficulty was extracted from the models and used as participant-level scores.
 We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general of illusion sensitivity (Factor *i*).
 
-Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) as well as personality traits.
+Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age), and personality traits.
 
@@
 For error rates, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
+A specific discussion regarding these 2 illusions is available in the Supplementary Materials (Part 1 - Discussion).
 
@@
 The last remaining trait, *Detachment*, did not share any significant relationship with illusion sensitivity.
+See Supplementary Materials (Part 2 - Discussion) for a detailed discussion regarding these associations.
 
@@
 
-The parametric illusion generation framework developed in @makowski2021 proposes to conceptualize illusions as being composed of targets and distractors that can be manipulated independently and continuously.
+The parametric illusion generation framework developed in @makowski2021 proposes to conceptualize illusions as composed of targets and distractors that can be manipulated independently and continuously.
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
 Specifically, increasing the illusion strength led to an increase in the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
-As such, this important methodological step not only helps us move away from traditional illusion research paradigms which are highly dependent on participants' subjective experiences of illusory effects; opening the door for new illusions-based paradigms and tasks, it also allows us to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
-For instance, using this novel framework accorded by *Pyllusion* to objectively manipulate the parameters of the Zöllner illusion, we were able to find an effect of illusion strength on errors that was distinct from the other 9 illusions examined.
-For a more detailed discussion on this specific illusion effect, please refer to the Supplementary Materials (Part 2 - Discussion).
+This important methodological step opens the door for new illusions-based paradigms and tasks to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes - such as accuracy or speed - instead of subjective meta-cognitive reports.
+This new and complementary approach will hopefully help address some of the longstanding literature gaps, as well as cement illusions as valuable stimuli for the study of cognition.
 
@@
 
-However, although the nature of the illusions were different in terms of the perceptual task underlying the illusory effect (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
-Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and general illusion sensitivity, and assess the psychometric properties of similar paradigms, including their stability (e.g., test-retest reliability) and validity.
+However, although the illusions did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+Future studies should investigate the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions), and assess the psychometric properties - such as stability (e.g., test-retest reliability) and validity - of similar illusion-based paradigms.
 
@@
 
-Finally, we found the sensitivity to illusions to be positively associated with "positive" personality traits, such as *Agreeableness* and *Honesty-Humility*, and negatively associated with maladaptive traits such as *Antagonism*, *Psychoticism*, *Disinhibition* and *Negative Affect*.
-Although empirical evidence proposing significant links between illusion sensitivity and maladaptive personality traits are scarce, these results are consistent with past findings relating pathological egocentric beliefs (often observed alongside *Psychoticism*, @fox2006) to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli[ [@ohmann2016; @konrath2009seeing; @fox2006].
+Finally, we found the sensitivity to illusions to be positively associated with "positive" personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
+Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs [often associated with psychoticism, @fox2006] to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli  [@ohmann2016; @konrath2009seeing; @fox2006].
 As such, the association between maladaptive traits and lower illusion sensitivity could be linked to a self-centered, decontextualized and disorganized information processing style.
-Conversely, the relationship between sensitivity to illusions and "positive" personality traits could be drawn from research relating cognitive styles (i.e., field dependence - a tendency to rely less on external cues in ambiguous contexts) to dispositional traits like hostility, aggression and narcissism [@zhang2017; @konrath2009seeing; @pessoa2008], which negatively correlate with "positive" traits like *Agreeableness* and *Honesty-Humility* (see Supplementary Materials, Discussion - Part 3 for an in-depth discussion on the relationship between personality traits, cognitive styles and illusion sensitivity).
+Conversely, the relationship between illusion sensitivity and adaptive personality traits is in line with the decreased field dependence (the tendency to rely on external cues in ambiguous contexts) associated with traits negatively correlated with agreeableness and honesty-humility, such as hostility, aggression and narcissism [@zhang2017; @konrath2009seeing; @pessoa2008].
 
-Importantly, such findings highlight the relevance of illusions beyond the field of visual perception, pointing towards an association with high-level domain-general mechanisms.
-In particular, the significant findings in favor of a relationship between maladaptive personality traits and illusion sensitivity could be viewed in relation to clinical findings, of which greater resistance to illusions have often been reported among patients with schizophrenia [@notredame2014; @pessoa2008; @grzeczkowski2018], especially in association with schizotypal traits such as cognitive disorganization [@cretenoud2019; @lanyi2022can].
-While the search for the exact mechanism(s) underlying these links is an important goal of future research, our observations unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
+Importantly, these findings highlight the relevance of illusions beyond the field of visual perception, pointing towards an association with high-level domain-general mechanisms.
+In particular, the evidence in favor of a relationship between maladaptive personality traits and illusion sensitivity is in line with clinical observations, in which a greater resistance to illusions have been reported among patients with schizophrenia [@notredame2014; @pessoa2008; @grzeczkowski2018], especially in association with schizotypal traits such as cognitive disorganization [@cretenoud2019; @lanyi2022can].
+While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
 
-To conclude, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+In conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
 The task, data and analysis script are available in open-access at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.

### 2022-09-22 submission/manuscript.tex
donzo

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 5114}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3621}
 \usepackage{lineno}
@@
 Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings.
-In this set of studies, we used the parametric framework for visual illusions implemented in the \emph{Pyllusion} software to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength.
-We tested the objective effect of the illusions on errors and reaction times in a perceptual discrimination task, from which we extracted participant-level performance scores (n=250).
+In this study, we used the novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
+We tested the objective effect of the illusions on errors and reaction times, and extracted participant-level performance scores (n=250).
 Our results provide evidence in favour of a general factor (labelled Factor \emph{i}) underlying the sensitivity to different illusions.
 Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
-Data and code are available in open-access.
+The experiment, data and code are fully available for re-use at \url{https://github.com/RealityBending/IllusionGameValidation}.
 }
@@
 
-Notably, the presence of a common mechanism underlying the effect of different illusions has been contested (Cretenoud et al., 2019; Cretenoud, Francis, et al., 2020; Hamburger, 2016); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018) are strongly debated.
-The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022; Pessoa et al., 2008) and in individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
+Notably, the presence of a common mechanism underlying the effects of different illusions has been contested (Cretenoud et al., 2019; Cretenoud et al., 2020; Hamburger, 2016); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018) are strongly debated.
+The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022; Pessoa et al., 2008), and in individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
 
 One key challenge hindering the further development of illusion research is the relative difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-To address this issue, we first developed a parametric framework to manipulate visual illusions, which we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021).
-This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (see \textbf{Figure 1}).
+To address this issue, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021).
+This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (\textbf{Figure 1}).
 
@@
 
-Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths when they end with inwards or outwards pointing arrows).
+Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths if they end with inwards vs.~outwards pointing arrows).
 Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to match a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
@@
 
-In the present set of preregistered studies, we will first test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously, and separately modeled statistically.
-Then, we will further utilize the paradigm to assess whether 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) share a common latent factor.
-Finally, we will investigate how the the inter-individual sensitivity to illusions relates to dispositional variables, such as demographic characteristics and personality.
+The aim of the present preregistered study is three-fold.
+First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
+Next, we will investigate the latent factors structure of participant-level performance scores to these illusions and test the existence of a common factor of illusion sensitivity.
+Finally, we will explore how illusion sensitivity relates to demographic characteristics, contextual variables (pertaining to the experiment setting), and personality traits.
 
-In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
+In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as \textbf{Supplementary Materials} at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
 
-\hypertarget{study-1}{%
-\section{Study 1}\label{study-1}}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
-\hypertarget{aim}{%
-\subsection{Aim}\label{aim}}
+\hypertarget{stimuli}{%
+\subsection{Stimuli}\label{stimuli}}
 
-Study 1 can be seen as a pilot experiment aiming to gather some preliminary data to assess if the stimuli generated by \emph{Pyllusion} behaves as expected for each of the 10 illusion types (i.e., whether an increase of task difficulty and illusion strength leads to an increase of errors), and develop an intuition about the magnitude of effects, to refine the stimuli parameters to a more sensible range (i.e., not overly easy and not impossibly hard) for the next study.
+A pilot study (\emph{n = 46}), of which full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
+Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
+These stimuli resulted from the combination of 15 equally-spaced levels of illusion \emph{strength} (7 negative, i.e., congruent effects; 7 positive, i.e., incongruent effects; and 0) overlapped with 16 non-linearly spaced task \emph{difficulty} levels (i.e., with an exponential, square or cubic spacing depending on the pilot results).
+For instance, a linear space of {[}0.1, 0.4, 0.7, 1.0{]} can be transformed to an exponential space of {[}0.1, 0.34, 0.64, 1.0{]}, where 0.1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets).
+For each illusion type, the stimuli were split into two series (56 and 72 stimuli per series) with alternating parameter values to maintain their homogeneity.
+Additionally, 6 stimuli per illusion type was generated for a practice series, with more extreme variations (i.e., containing very easy trials to help cement the task instructions).
 
@@
 
-We generated 56 stimuli for each of the 10 illusion types.
-These stimuli resulted from the combination of 8 linearly-spread levels of task difficulty (e.g., {[}1, 2, 3, 4, 5, 6, 7{]}, where 1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets) and 7 levels of illusion strength (3 values of strength on the congruent side, 3 on the incongruent side, and 0; e.g., {[}-3, -2, -1, 0, 1, 2, 3{]}, where negative values correspond to congruent illusion strengths).
-
-The 10 illusion blocks were randomly presented, and the order of the 56 stimuli within the blocks was also randomized.
-After the first series of 10 blocks, another series was administered (with new randomized orders of blocks and trials).
-In total, each participant saw 56 different trials per 10 illusion type, repeated 2 times (total = 1120 trials), to which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off).
-The task was implemented using \emph{jsPsych} (De Leeuw, 2015), and the instructions for each illusion type are available in the experiment code.
+After a brief demographic survey and a practice series of illusions, the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
+Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6} (24 items, Sibley et al., 2011) - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5} (25 items, Hopwood et al., 2012) - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
+In total, each participant underwent 1340 trials of which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key (left/right,or up/down depending on the illusion type).
+For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
+The task was implemented using \emph{jsPsych} (De Leeuw, 2015), and the set of instructions for each illusion type is available in the experiment code.
 
@@
 
-Fifty-two participants were recruited via \emph{Prolific} (\url{www.prolificacademic.co.uk}), a crowd-sourcing platform providing high data quality (Peer et al., 2022).
+Participants were recruited via \emph{Prolific}, a crowd-sourcing platform recognized for providing high quality data (Peer et al., 2022).
 The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
@@
 
-We removed 6 participants upon inspection of the average error rate (when close to 50\%, suggesting random answers), and when the reaction time distribution was implausibly fast.
-For the remaining participants, we discarded blocks where the error rate was higher than 50\% (possibly indicating that instructions got misunderstood; e.g., participants were selecting the shorter line instead of the longer one).
-Finally, we removed 692 (1.37\%) trials based on an implausibly short or long response time (\textless{} 150 ms or \textgreater{} 3000 ms).
+We excluded 6 participants upon inspection of the average error rate (when close to 50\%, suggesting random answers), and reaction time distribution (when implausibly fast).
+For the remaining participants, we discarded blocks with more than 50\% of errors (2.16\% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76\% trials with extreme response times (\textless{} 125 ms or \textgreater{} 4 SD above mean).
+Additionally, due to a technical issue, no personality data was recorded for the first eight participants.
 
-The final sample included 46 participants (Mean age = 26.7, SD = 7.7, range: {[}19, 60{]}; Sex: 39.1\% females, 56.5\% males, and 4.4\% other).
+The final sample included 250 participants (Mean age = 26.5, SD = 7.6, range: {[}18, 69{]}; Sex: 48\% females, 52\% males).
 
@@
 
-The analysis of study 1 focused on the probability of errors as the main outcome variable.
-For each illusion, we started by visualizing the average effect of task difficulty and illusion strength to gain some intuition on the underlying generative model.
-Next, we tested the performance of various logistic models differing in their specifications, such as: with or without a transformation of the task difficulty (log, square root or cubic root), with or without a 2nd order polynomial term for the illusion strength, and with or without the illusion side (up \emph{vs.} down or left \emph{vs.} right) as an additional predictor.
-We then fitted the best performing model under a Bayesian framework, and compared its visualization with that of a General Additive Model (GAM), which has an increased ability of mapping underlying potential non-linear relationships (at the expense of model simplicity).
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT) within each illusion.
+We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions.
+Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location \(\mu\) (mean), scale \(\sigma\) (spread) and tail-dominance \(\tau\) of the RT distribution (Balota \& Yap, 2011; Matzke \& Wagenmakers, 2009).
+
+Using GAMs as the ``ground-truth'' models, we attempted at approximating them using general linear mixed models, which can be used to estimate the effects' participant-level variability (via random slopes).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task \emph{difficulty} and illusion \emph{strength}), we fitted the best model (based on their indices of fit), and compared their output visually (\textbf{Figure 2}).
+
+The inter-individual variability in the effect of illusion strength and its interaction with task difficulty was extracted from the models and used as participant-level scores.
+We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general of illusion sensitivity (Factor \emph{i}).
+
+Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor \emph{i}), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age), and personality traits.
 
 The analysis was carried out using \emph{R 4.2} (R Core Team, 2022), \emph{brms} (Bürkner, 2017), the \emph{tidyverse} (Wickham et al., 2019), and the \emph{easystats} collection of packages (Lüdecke et al., 2021, 2019; Makowski et al., 2020; Makowski, Ben-Shachar, \& Lüdecke, 2019).
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-The statistical models suggested that the effect of task difficulty had a cubic relationship with error rate for the Delboeuf and Ebbinghaus illusions (both composed of circular shapes), square relationship for the Rod and Frame and Vertical-Horizontal illusions, cubic relationship for the Zöllner and Poggendorff illusions, exponential relationship for the White illusion, cubic relationship for the Müller-Lyer and Ponzo illusions (both based on line lengths), and linear relationship for the Contrast illusion.
-All models suggested a significant effect of illusion strength and task difficulty.
-See details and figures in the analysis script.
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
-This study provided a clearer understanding of the magnitude of the parametric effects at stake and the type of interaction between them.
-Furthermore, it allowed us to better understand and test the stimuli generated by \emph{Pyllusion}, as well as uncover incidental bugs and technical issues (for instance, the specification direction of the illusion strength was reversed for a few illusions), which were fixed in a new software release.
-Crucially, this study allowed us to refine the range of task difficulty and illusion strength values in order to maximize information gain.
-
-In most illusions, the task difficulty exhibited monotonic power-law scaled effects, which is in line with the psychophysics literature on perceptual decisions (Bogacz et al., 2006; Ditzinger, 2010; Shekhar \& Rahnev, 2021).
-One notable result was the illusion effect pattern for the Zöllner illusion, which suggested a non-linear relationship.
-By generating a wider range of illusion strength values, the next study will attempt at clarifying this point.
-
-\hypertarget{study-2}{%
-\section{Study 2}\label{study-2}}
-
-\hypertarget{aim-1}{%
-\subsection{Aim}\label{aim-1}}
-
-The aim of study 2 was two-fold.
-In the first part, we carefully modeled the error rate and the reaction time of each illusion type in order to validate our novel paradigm and show that the effect of illusions can be manipulated continuously.
-In the second part, we derived the participant-level scores from the models (i.e., the effect of illusion strength for each individual) and analyzed their latent factors structure.
-
-\hypertarget{procedure-1}{%
-\subsection{Procedure}\label{procedure-1}}
-
-The paradigm of study 2 was similar to that of study 1, with the following changes: the illusory stimuli were re-generated within a refined space of parameters based on the results of study 1.
-Moreover, taking into account the findings of study 1, we used non-linearly spaced difficulty levels, depending on the best underlying model (i.e., with an exponential, square or cubic spacing depending on the relationship).
-For instance, a linear space of {[}0.1, 0.4, 0.7, 1.0{]} can be transformed to an exponential space of {[}0.1, 0.34, 0.64, 1.0{]}.
-
-Additionally, instead of repeating each stimulus two times, we generated illusions using more levels of difficulty and illusion strength.
-As such, for each illusion type, we generated a total of 134 stimuli that were split into two groups (67 stimuli per illusion block).
-Furthermore, instead of a simple break screen, we added two personality questionnaires between the two series of 10 illusion blocks (see study 3).
-
-\hypertarget{participants-1}{%
-\subsection{Participants}\label{participants-1}}
-
-Using the same recruitment procedure as in study 1, we recruited 256 participants, out of which 6 were identified as outliers and excluded, leaving a final sample of 250 participants (Mean age = 26.5, SD = 7.6, range: {[}18, 69{]}; Sex: 48\% females, 52\% males).
-Please see study 3 for the full demographic breakdown.
-We discarded blocks with more than 50\% of errors (2.16\% of trials) and 0.76\% trials with extreme response times (\textless{} 125 ms or \textgreater{} 4 SD above mean).
-
-\hypertarget{data-analysis-1}{%
-\subsection{Data Analysis}\label{data-analysis-1}}
-
-The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT) within each illusion.
-In order to achieve this, we started by fitting General Additive Models (GAMs), which can accommodate possible non-linear effects and interactions.
-Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location \(\mu\) (mean), scale \(\sigma\) (spread) and tail-dominance \(\tau\) of the RT distribution (Balota \& Yap, 2011; Matzke \& Wagenmakers, 2009).
-
-Using GAMs as the ``ground-truth'' models, we attempted at approximating them using general linear models, which have the advantage of estimating the participant-level variability of the effects (via random slopes).
-Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task \emph{difficulty} and illusion \emph{strength}), we selected and fitted the best model (based on their indices of fit), and compared their output visually (see \textbf{Figure 2}).
-
-We then extracted the inter-individual variability in the effect of illusion strength and its interaction with task difficulty, and used it as participant-level scores.
-Finally, we explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM).
-
-\hypertarget{results-1}{%
-\subsection{Results}\label{results-1}}
-
-The best models were \(log(diff)*strength\) for Delboeuf; \(sqrt(diff)*strength\) for Ebbinghaus; \(log(diff)*log(strength)\) for Rod and Frame; \(sqrt(diff)*sqrt(strength)\) for Vertical-Horizontal; \(cbrt(diff)*strength\) for Zöllner; \(diff*sqrt(strength)\) and \(log(diff)*strength\) respectively for errors and RT in White; \(sqrt(diff)*sqrt(strength)\) and \(sqrt(diff)*strength\) respectively for errors and RT in Müller-Lyer; \(cbrt(diff)*strength\) for Ponzo; \(cbrt(diff)*sqrt(strength)\) and \(cbrt(diff)*strength\) respectively for errors and RT in Poggendorff; and \(sqrt(diff)*sqrt(strength)\) for Contrast.
-For all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
-
-For error rates, most of the models closely matched their GAMs counterpart (see \textbf{Figure 2}), with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
+As the full results are available as supplementary materials, we will focus here on the significant results (based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see Makowski, Ben-Shachar, Chen, et al., 2019).
 
@@
 
-For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants respond faster.
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{effects-of-illusion-strength-and-task-difficulty}{%
+\subsection{Effects of Illusion Strength and Task Difficulty}\label{effects-of-illusion-strength-and-task-difficulty}}
+
+The best model specifications were \(log(diff)*strength\) for Delboeuf; \(sqrt(diff)*strength\) for Ebbinghaus; \(log(diff)*log(strength)\) for Rod and Frame; \(sqrt(diff)*sqrt(strength)\) for Vertical-Horizontal; \(cbrt(diff)*strength\) for Zöllner; \(diff*sqrt(strength)\) and \(log(diff)*strength\) respectively for errors and RT in White; \(sqrt(diff)*sqrt(strength)\) and \(sqrt(diff)*strength\) respectively for errors and RT in Müller-Lyer; \(cbrt(diff)*strength\) for Ponzo; \(cbrt(diff)*sqrt(strength)\) and \(cbrt(diff)*strength\) respectively for errors and RT in Poggendorff; and \(sqrt(diff)*sqrt(strength)\) for Contrast.
+For all of these models, the effects of illusion strength, task difficulty and their interaction were significant.
+
+For error rates, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
+A specific discussion regarding these 2 illusions is available in the Supplementary Materials (Part 1 - Discussion).
+
+For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants responded faster.
 While this is is not surprising (strong illusions are likely so effective in biasing perception that it is ``easier'', i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
@@
 
-Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy (and are also more conservative estimates due to shrinkage) than basic empirical scores, such as the total number of errors, or the average RT.
+\hypertarget{factor-structure}{%
+\subsection{Factor Structure}\label{factor-structure}}
+
+Though imperfect, we believe that the random-slope models capture inter-individual differences with more accuracy (and also provide more conservative estimates due to shrinkage) than basic empirical scores, such as the total number of errors, or the average RT.
 Thus, for each illusion and within each participant, we extracted the effect of illusion strength and its interaction with task difficulty when the illusion effect was incongruent.
@@
 
-\hypertarget{discussion-1}{%
-\subsection{Discussion}\label{discussion-1}}
+\hypertarget{correlations-with-inter-individual-characteristics}{%
+\subsection{Correlations with Inter-individual Characteristics}\label{correlations-with-inter-individual-characteristics}}
 
-This study confirmed that it was possible to continuously manipulate the effect of illusion strength for 10 classic illusions.
-Increasing the illusion strength increased the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
-Future studies are needed to explore reaction times and identify the most appropriate models, and/or use models that integrate both errors and reaction time (e.g., drift diffusion models).
-
-The effect on errors was monotonic for most illusions, with the exception of Delboeuf and Zöllner.
-For both of them, mildly congruent illusion strengths (which theoretically were supposed to be associated with fewer errors than incongruent effects) were related to small and strong increases of errors, respectively.
-For the Delboeuf illusion, we believe that this was due to an artifact caused by the illusion generation algorithm: the outline of the target circles was always created as slightly bigger, which made the difference between them more obvious at an illusion strength of 0.
-This was fixed in the latest release of \emph{Pyllusion} (v1.2), which now generates outlines of the same size as the target circle.
-For the Zöllner illusion, the observed non-monotonic pattern is actually consistent with previous reports (Kitaoka, 2007; Kitaoka \& Ishihara, 2000), suggesting an acute angle contraction effect at very small - as well as at sufficiently large angles (below 10 degrees for the former and between 50 to 90 degrees for the latter) between the target horizontal line and the biasing horizontal bars when the illusion strength is weak.
-
-Finally, this study provided evidence for both the existence of illusion-specific factors, as well as for a common latent factor (labelled Factor \emph{i}) that explained about half of the total variance.
-These participant-level scores were positively related to the error rate and average reaction time, and can thus be interpreted as indices of illusion sensitivity.
-
-\hypertarget{study-3}{%
-\section{Study 3}\label{study-3}}
-
-\hypertarget{aim-2}{%
-\subsection{Aim}\label{aim-2}}
-
-Study 3 aimed at investigating the links between the inter-individual scores of illusion sensitivity (obtained in study 2), contextual variables (pertaining to the experiment setting), such as screen size, demographic features (such as sex and age), and stable dispositional variables such as ``general'' personality traits.
-Indeed, despite the abundant literature on visual illusions, relatively few studies have investigated its ties with participants' characteristics.
-Research examining the influence of demographic variables such as gender and age have generally found inconsistent results (Cretenoud, Grzeczkowski, et al., 2020; Grzeczkowski et al., 2017; Lo \& Dinov, 2011; Papageorgiou et al., 2020).
-Regarding links with personality, most works focused on traits associated with psychopathology, such as impulsivity or sensation-seeking (Hlavatá et al., 2018; Lányi et al., 2022; Pessoa et al., 2008; Zhang et al., 2017).
-
-\hypertarget{procedure-2}{%
-\subsection{Procedure}\label{procedure-2}}
-
-This study was based on the data collected in study 2.
-The variables of interest here were taken from the questionnaires that were inserted in between the two series of illusion blocks.
-We used the \emph{IPIP6} (24 items, Sibley et al., 2011) to measure 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the PID-5 (25 items, Hopwood et al., 2012) to measure ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
-The participants were the same as in study 2 (see \textbf{Figure 3}).
-However, due to a technical issue, no personality data was recorded for the first eight participants.
+The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions (\textbf{Figure 3}).
+Anecdotal (\(BF_{10} = 2.05\)) and moderate evidence (\(BF_{10} = 4.11\)) was found for a negative correlation between screen size and the sensitivity to the White and the Contrast illusion, respectively.
+To test whether this result could be an artifact related to the highly skewed screen size distribution (caused by very few participants with extreme screen sizes), we re-ran a robust correlation (with rank-transformed values), which provided even stronger evidence in favor of the effect existence (\(BF_{10} = 28.19\), \(BF_{10} = 4.31\) for White and Contrast, respectively).
 
@@
 \end{figure}
-
-\hypertarget{data-analysis-2}{%
-\subsection{Data Analysis}\label{data-analysis-2}}
-
-For each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor \emph{i}), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age) and personality.
-As the supplementary material contains the detailed results, we will only report the significant results (based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see Makowski, Ben-Shachar, Chen, et al., 2019).
-
-\hypertarget{results-2}{%
-\subsection{Results}\label{results-2}}
-
-The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions.
-Anecdotal (\(BF_{10} = 2.05\)) and moderate evidence (\(BF_{10} = 4.11\)) was found for a negative correlation between screen size and the sensitivity to the White and the Contrast illusion, respectively.
-To test whether this result could be an artifact related to the highly skewed screen size distribution (caused by very few participants with extreme screen sizes), we re-ran a robust correlation (with rank-transformed values), which provided even stronger evidence in favor of the effect existence (\(BF_{10} = 28.19\), \(BF_{10} = 4.31\) for White and Contrast, respectively).
 
@@
 The last remaining trait, \emph{Detachment}, did not share any significant relationship with illusion sensitivity.
+See Supplementary Materials (Part 2 - Discussion) for a detailed discussion regarding these associations.
 
-\hypertarget{discussion-2}{%
-\subsection{Discussion}\label{discussion-2}}
-
-We report significant links between inter-individual indices of illusion sensitivity and variables related to experimental context, demographic characteristics and personality.
-Firstly, screen size was found to have a significant negative relationship with the sensitivity to the two contrast-based illusions, namely the White and Contrast illusions.
-One possible explanation can be found in the mechanism by which visual systems filter through more low spatial frequencies when the size of the target object is small (Dixon et al., 2014).
-As this filtering process excludes illumination information from visual processing, smaller screen sizes could yield artifactual changes in brightness perception, which in turn could attenuate the illusory effect of luminance-related illusions.
-
-Our results suggested an inconsistent pattern of non-significant sex-differences, with the exception of greater sensitivity of males as compared to females for the Zöllner illusion.
-Although we do not consider this result as significant given its specificity, we note that the existing literature reports, if any differences, that females exhibited greater illusion sensitivity (Lo \& Dinov, 2011; Miller, 2001; Papageorgiou et al., 2020).
-This inconsistency could be due to past studies using a measure of illusion sensitivity that conflates the effect of illusions \emph{per se} with the perceptual abilities involved in the task, for which gender-related differences can be found (in fact, the authors mention sex-differences in visuospatial strategies as the potential mechanism underlying their findings).
-On the contrary, the perceptual difficulty of the task and the illusion effect was independently modulated in our paradigm, and statistically dissociated.
-Our scores of illusion sensitivity might thus be less loaded with perceptual skills, thereby mitigating its effect.
-
-Our findings also suggested a positive relationship between illusion sensitivity and two ``normal'' personality traits, namely \emph{Honesty-Humility} and \emph{Agreeableness}, and a negative link with \emph{Antagonism}.
-Although the past literature regarding the links between illusion sensitivity and personality traits remain scarce, convergent evidence can be found in studies reporting a negative relationship between illusion sensitivity and hostility, aggression and narcissism (Konrath et al., 2009; Pessoa et al., 2008; Zhang et al., 2017).
-While this result's interpretation is challenging, one possible explanation could be drawn from the literature on the cognitive style known as field dependence.
-Since narcissism and aggression tendencies are correlated with lower field dependence (D'Amour et al., 2021; Ohmann \& Burgmer, 2016; i.e., a lesser reliance on external cues in ambiguous contexts, Witkin \& Goodenough, 1976), opposite traits, such as \emph{Honesty-Humility} and \emph{Agreeableness}, could conversely be more biased by contextual cues and thus more sensitive to illusions.
-
-The positive relationship between illusion sensitivity and ``positive'' personality traits is mirrored by a negative relationship with several other pathological traits, including \emph{Psychoticism}, \emph{Disinhibition}, and \emph{Negative Affect}.
-These results are, in general, consistent with past findings and theories, suggesting a negative relationship between egocentric cognitive styles and context processing (including illusion sensitivity, Konrath et al., 2009).
-For instance, pathological egocentric beliefs (often observed alongside \emph{Psychoticism}, Fox, 2006) have been related to reduced context integration (Fox, 2006; Konrath et al., 2009; manifesting for instance in a tendency to separate objects from their surroundings when processing visual stimuli, Ohmann \& Burgmer, 2016).
-As such, it is possible to relate this higher resistance to illusions to a self-centered, decontextualized and disorganized information processing style, which can be found across the aforementioned maladaptive personality traits (Calamari et al., 2000; Hoyle, 2006; Msetfi et al., 2009; Parkes, 1981) .
-
-Furthermore, these results in favour of a link between illusion sensitivity and maladaptive personality traits in a non-clinical population could be put in relation with clinical findings, which could be seen as extreme cases where the relationship with illusion sensitivity is the most manifest.
-In line with our results (in particular on \emph{Psychoticism} and \emph{Disinhibition}), prior research has found greater illusion resistance in schizophrenia (Grzeczkowski et al., 2018; Notredame et al., 2014; Pessoa et al., 2008), and in particular, in association with schizotypal traits, such as cognitive disorganization (Cretenoud et al., 2019; Lányi et al., 2022).
-
-\hypertarget{general-discussion}{%
-\section{General Discussion}\label{general-discussion}}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
 The parametric illusion generation framework developed in Makowski et al.
-(2021) proposes to conceptualize illusions as made of targets and distractors, both of which can be manipulated independently and continuously.
+(2021) proposes to conceptualize illusions as composed of targets and distractors that can be manipulated independently and continuously.
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
-This important methodological step opens the door for new illusions-based paradigms and tasks to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes, such as accuracy or speed.
+Specifically, increasing the illusion strength led to an increase in the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
+This important methodological step opens the door for new illusions-based paradigms and tasks to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes - such as accuracy or speed - instead of subjective meta-cognitive reports.
+This new and complementary approach will hopefully help address some of the longstanding literature gaps, as well as cement illusions as valuable stimuli for the study of cognition.
 
@@
 This result comes in a field of mixed findings.
-In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals (Cretenoud et al., 2019; Cretenoud, Francis, et al., 2020; Grzeczkowski et al., 2017, 2018; Yang et al., 2012).
+In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals (Cretenoud et al., 2019; Cretenoud et al., 2020; Grzeczkowski et al., 2017, 2018; Yang et al., 2012).
 Instead, past findings suggest illusory effects are highly specific to the perceptual features of the illusions at stake (Cretenoud et al., 2019; Grzeczkowski et al., 2017).
@@
 
-However, although the illusions were relatively different in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
-Future studies should investigate the relationship between perceptual abilities (in a similar task, but without illusions) and illusion sensitivity, and assess the psychometric properties of similar paradigms, including stability (e.g., test-retest reliability) and validity.
+However, although the illusions did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+Future studies should investigate the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions), and assess the psychometric properties - such as stability (e.g., test-retest reliability) and validity - of similar illusion-based paradigms.
 
-Finally, we found that the sensitivity to illusions was positively associated with ``positive'' personality traits, such as \emph{Agreeableness} and \emph{Honesty-Humility}, and negatively associated with maladaptive traits such as \emph{Antagonism}, \emph{Psychoticism}, \emph{Disinhibition} and \emph{Negative Affect}.
-Beyond highlighting the relevance of illusions beyond the field of visual perception, these results point towards an association with high-level general-domain mechanisms.
+Finally, we found the sensitivity to illusions to be positively associated with ``positive'' personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
+Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs (often associated with psychoticism, Fox, 2006) to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli (Fox, 2006; Konrath et al., 2009; Ohmann \& Burgmer, 2016).
+As such, the association between maladaptive traits and lower illusion sensitivity could be linked to a self-centered, decontextualized and disorganized information processing style.
+Conversely, the relationship between illusion sensitivity and adaptive personality traits is in line with the decreased field dependence (the tendency to rely on external cues in ambiguous contexts) associated with traits negatively correlated with agreeableness and honesty-humility, such as hostility, aggression and narcissism (Konrath et al., 2009; Pessoa et al., 2008; Zhang et al., 2017).
+
+Importantly, these findings highlight the relevance of illusions beyond the field of visual perception, pointing towards an association with high-level domain-general mechanisms.
+In particular, the evidence in favor of a relationship between maladaptive personality traits and illusion sensitivity is in line with clinical observations, in which a greater resistance to illusions have been reported among patients with schizophrenia (Grzeczkowski et al., 2018; Notredame et al., 2014; Pessoa et al., 2008), especially in association with schizotypal traits such as cognitive disorganization (Cretenoud et al., 2019; Lányi et al., 2022).
 While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
 
-To conclude, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+In conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
 The task, data and analysis script are available in open-access at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bogacz2006}{}}%
-Bogacz, R., Brown, E., Moehlis, J., Holmes, P., \& Cohen, J.
-D.
-(2006).
-The physics of optimal decision making: A formal analysis of models of performance in two-alternative forced-choice tasks.
-\emph{Psychological Review}, \emph{113}(4), 700--765.
-\url{https://doi.org/10.1037/0033-295X.113.4.700}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
@@
 \emph{British Journal of Mathematical and Statistical Psychology}, \emph{73}(3), 420--451.
-
-\leavevmode\vadjust pre{\hypertarget{ref-calamari2000field}{}}%
-Calamari, E., Pini, M., \& Puleggio, A.
-(2000).
-Field dependence and verbalized strategies on the portable rod-and-frame test by depressed outpatients and normal controls.
-\emph{Perceptual and Motor Skills}, \emph{91}(3\_suppl), 1221--1229.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020individual}{}}%
-Cretenoud, A.
-F., Grzeczkowski, L., Bertamini, M., \& Herzog, M.
-H.
-(2020).
-Individual differences in the m{ü}ller-lyer and ponzo illusions are stable across different contexts.
-\emph{Journal of Vision}, \emph{20}(6), 4--4.
-
 \leavevmode\vadjust pre{\hypertarget{ref-cretenoud2019}{}}%
@@
 \url{https://doi.org/10.1167/19.14.12}
-
-\leavevmode\vadjust pre{\hypertarget{ref-damour2021}{}}%
-D'Amour, S., Harris, L.
-R., Berti, S., \& Keshavarz, B.
-(2021).
-The role of cognitive factors and personality traits in the perception of illusory self-motion (vection).
-\emph{Attention, Perception \& Psychophysics}, \emph{83}(4), 1804--1817.
-\url{https://doi.org/10.3758/s13414-020-02228-3}
 
@@
 De Leeuw, J. R. (2015). jsPsych: A JavaScript library for creating behavioral experiments in a web browser. \emph{Behavior Research Methods}, \emph{47}(1), 1--12.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ditzinger2010}{}}%
-Ditzinger, T. (2010). \emph{Optical illusions: Examples for nonlinear dynamics in perception} (Vol. 328, pp. 179--191).
-
-\leavevmode\vadjust pre{\hypertarget{ref-dixon2014}{}}%
-Dixon, E., Shapiro, A., \& Lu, Z.-L.
-(2014).
-Scale-invariance in brightness illusions implicates object-level visual processing.
-\emph{Scientific Reports}, \emph{4}(1), 3900.
-\url{https://doi.org/10.1038/srep03900}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hlavata2018}{}}%
-Hlavatá, P., Kašpárek, T., Linhartová, P., Ošlejšková, H., \& Bareš, M.
-(2018).
-Autism, impulsivity and inhibition a review of the literature.
-\emph{Basal Ganglia}, \emph{14}, 44--53.
-\url{https://doi.org/10.1016/j.baga.2018.10.002}
-
 \leavevmode\vadjust pre{\hypertarget{ref-hopwood2012}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hoyle2006}{}}%
-Hoyle, R.
-H.
-(2006).
-Personality and self-regulation: trait and information-processing perspectives.
-\emph{Journal of Personality}, \emph{74}(6), 1507--1525.
-\url{https://doi.org/10.1111/j.1467-6494.2006.00418.x}
-
 \leavevmode\vadjust pre{\hypertarget{ref-keane2014}{}}%
@@
 \url{https://doi.org/10.1016/j.neuropsychologia.2014.02.001}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kitaoka2007}{}}%
-Kitaoka, A.
-(2007).
-Tilt illusions after Oyama (1960): A review {\textsuperscript{1}}: Tilt illusions: A review.
-\emph{Japanese Psychological Research}, \emph{49}(1), 7--19.
-\url{https://doi.org/10.1111/j.1468-5884.2007.00328.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kitaoka2000}{}}%
-Kitaoka, A., \& Ishihara, M. (2000). Three elemental illusions determine the zöllner illusion. \emph{Perception \& Psychophysics}, \emph{62}, 569--575.
 
@@
 Lányi, O., Keri, S., Pálffy, Z., \& Polner, B. (2022). \emph{Can you believe your eyes? Positive schizotypy is associated with increased susceptibility to the m{ü}ller-lyer illusion}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lo2011}{}}%
-Lo, K. C., \& Dinov, D. I. (2011). \emph{Investigation of Optical Illusions on the Aspects of Gender and Age}. \emph{24}, 8.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-miller2001}{}}%
-Miller, R.
-J.
-(2001).
-Gender differences in illusion response: The inffluence of spatial strategy and sex ratio.
-\emph{Sex Roles}, \emph{44}(3/4), 209--225.
-\url{https://doi.org/10.1023/A:1010907204430}
-
-\leavevmode\vadjust pre{\hypertarget{ref-msetfi2009}{}}%
-Msetfi, R.
-M., Murphy, R.
-A., Kornbrot, D.
-E., \& Simpson, J.
-(2009).
-Short article: Impaired context maintenance in mild to moderately depressed students.
-\emph{Quarterly Journal of Experimental Psychology}, \emph{62}(4), 653--662.
-\url{https://doi.org/10.1080/17470210802486092}
-
 \leavevmode\vadjust pre{\hypertarget{ref-mylniec2016}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-papageorgiou2020}{}}%
-Papageorgiou, C., Stachtea, X., Papageorgiou, P., Alexandridis, A.
-T., Makris, G., Chrousos, G., \& Kosteletos, G.
-(2020).
-Gender-dependent variations in optical illusions: evidence from N400 waveforms.
-\emph{Physiological Measurement}, \emph{41}(9), 095006.
-\url{https://doi.org/10.1088/1361-6579/abb2eb}
-
 \leavevmode\vadjust pre{\hypertarget{ref-park2022}{}}%
@@
 \url{https://doi.org/10.1111/ejn.15739}
-
-\leavevmode\vadjust pre{\hypertarget{ref-parkes1981field}{}}%
-Parkes, K. R. (1981). Field dependence and the differentiation of affective states. \emph{The British Journal of Psychiatry}, \emph{139}(1), 52--58.
 
@@
 \url{https://doi.org/10.1080/03004430.2020.1802260}
-
-\leavevmode\vadjust pre{\hypertarget{ref-shekhar2021}{}}%
-Shekhar, M., \& Rahnev, D. (2021). The nature of metacognitive inefficiency in perceptual decision making. \emph{Psychological Review}, \emph{128}(1), 45--70. \url{https://doi.org/10.1037/rev0000249}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-witkin1976}{}}%
-Witkin, H.
-A., \& Goodenough, D.
-R.
-(1976).
-Field Dependence and Interpersonal Behavior.
-\emph{ETS Research Bulletin Series}, \emph{1976}(1), i--78.
-\url{https://doi.org/10.1002/j.2333-8504.1976.tb01098.x}
-
 \leavevmode\vadjust pre{\hypertarget{ref-yang2012}{}}%

### 2022-09-22 submission/manuscript.Rmd
fix sentence

@@
 Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
-The experiment, data and code are fully available for re-use at https://github.com/RealityBending/IllusionGameValidation.
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3621"
+wordcount         : "3611"
 bibliography      : references.bib
@@
 First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
-Next, we will investigate the latent factors structure of participant-level performance scores to these illusions and test the existence of a common factor of illusion sensitivity.
-Finally, we will explore how illusion sensitivity relates to demographic characteristics, contextual variables (pertaining to the experiment setting), and personality traits.
+Next, we will investigate the factor structure of illusion-specific performance scores and test the existence of a common latent factor of illusion sensitivity.
+Finally, we will explore how illusion sensitivity relates to demographic characteristics, contextual variables, and personality traits.
 
-In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as **Supplementary Materials** at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+Following open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as **Supplementary Materials** at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
 

### 2022-09-22 submission/manuscript.tex
fix sentence

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3621}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3611}
 \usepackage{lineno}
@@
 Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
-The experiment, data and code are fully available for re-use at \url{https://github.com/RealityBending/IllusionGameValidation}.
 }
@@
 First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
-Next, we will investigate the latent factors structure of participant-level performance scores to these illusions and test the existence of a common factor of illusion sensitivity.
-Finally, we will explore how illusion sensitivity relates to demographic characteristics, contextual variables (pertaining to the experiment setting), and personality traits.
+Next, we will investigate the factor structure of illusion-specific performance scores and test the existence of a common latent factor of illusion sensitivity.
+Finally, we will explore how illusion sensitivity relates to demographic characteristics, contextual variables, and personality traits.
 
-In line with open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as \textbf{Supplementary Materials} at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
+Following open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as \textbf{Supplementary Materials} at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
 

### 2022-09-22 submission/manuscript.Rmd
other minor

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3611"
+wordcount         : "3628"
 bibliography      : references.bib
@@
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
-Specifically, increasing the illusion strength led to an increase in the likelihood of errors, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
-This important methodological step opens the door for new illusions-based paradigms and tasks to study the effect of illusions under different conditions and to quantify illusion sensitivity using objective behavioral outcomes - such as accuracy or speed - instead of subjective meta-cognitive reports.
+Increasing the illusion strength led to an increase in error likelihood, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
+Using mixed models, we were able to statistically quantify the effect of illusions for each illusion and each participant separately.
+This important methodological step opens the door for new illusions-based paradigms and tasks to study the effect of illusions under different conditions and to measure illusion sensitivity using objective behavioral outcomes - such as accuracy or speed - instead of subjective meta-cognitive reports.
 This new and complementary approach will hopefully help address some of the longstanding literature gaps, as well as cement illusions as valuable stimuli for the study of cognition.

### 2022-09-22 submission/manuscript.tex
other minor

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3611}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3628}
 \usepackage{lineno}

### 2022-09-22 submission/manuscript.Rmd
reorganize intro

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3628"
+wordcount         : "3808"
 bibliography      : references.bib
@@
 
-Notably, the presence of a common mechanism underlying the effects of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] are strongly debated.
+Notably, the presence of a common mechanism underlying the effects of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] remain debated.
 The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], and in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
+
+<!-- Bayes -->
+
+One recent theory proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free], as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to provide a coherent perceptual experience.
+In this framework, certain traits or characteristics (e.g., psychoticism) are driven by alterations in the system's metacognitive components [@adams2013computational], resulting in underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions [@koethe2009binocular].
+
+<!-- Technical issues -->
+
+Despite strong theoretical foundations and hypotheses, the empirical evidence remains scarce, clouded by methodological hurdles.
+For instance, one key challenge is the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability.
+Moreover, the paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
+However, this reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity.
 
@@
 
-One key challenge hindering the further development of illusion research is the relative difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-To address this issue, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
+To address these issues, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software *Pyllusion* [@makowski2021].
 This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: *illusion strength* and *task difficulty* (**Figure 1**).
+Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths if they end with inwards vs.
+outwards pointing arrows).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to match a reference stimulus relying only on their perception [@mylniec2016; @grzeczkowski2018].
+Alternatively, *Pyllusion* allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
@@
 
-<!-- Target, strength, difficulty -->
-
-Indeed, many visual illusions can be seen as being composed of *targets* (e.g., same-length lines), of which perception is biased by the *context* (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths if they end with inwards vs.
-outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to match a reference stimulus relying only on their perception [@mylniec2016; @grzeczkowski2018].
-Alternatively, *Pyllusion* allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
-
 <!-- Paradigm -->
 
-This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can be either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier).
+This allows the creation of an experimental task in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
+Moreover, the illusion effect can specified as either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier).
 Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.

### 2022-09-22 submission/manuscript.tex
reorganize intro

@@
 
-Notably, the presence of a common mechanism underlying the effects of different illusions has been contested (Cretenoud et al., 2019; Cretenoud et al., 2020; Hamburger, 2016); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018) are strongly debated.
+Notably, the presence of a common mechanism underlying the effects of different illusions has been contested (Cretenoud et al., 2019; Cretenoud et al., 2020; Hamburger, 2016); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018) remain debated.
 The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022; Pessoa et al., 2008), and in individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
 
-One key challenge hindering the further development of illusion research is the relative difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-To address this issue, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021).
+One recent theory proposes to conceptualize illusions under the Bayesian brain hypothesis (Friston, 2010), as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to provide a coherent perceptual experience.
+In this framework, certain traits or characteristics (e.g., psychoticism) are driven by alterations in the system's metacognitive components (Adams et al., 2013), resulting in underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions (Koethe et al., 2009).
+
+Despite strong theoretical foundations and hypotheses, the empirical evidence remains scarce, clouded by methodological hurdles.
+For instance, one key challenge is the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability.
+Moreover, the paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to perceptually match a reference stimulus (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
+However, this reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity.
+
+To address these issues, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021).
 This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (\textbf{Figure 1}).
+Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths if they end with inwards vs.~outwards pointing arrows).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to match a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
+Alternatively, \emph{Pyllusion} allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
 
@@
 
-Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths if they end with inwards vs.~outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to match a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
-Alternatively, \emph{Pyllusion} allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
-
-This opens the door for an experimental task in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can be either ``incongruent'' (making the task more difficult by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
+This allows the creation of an experimental task in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
+Moreover, the illusion effect can specified as either ``incongruent'' (making the task more difficult by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
 Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the ``objective'' effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-adams2013computational}{}}%
+Adams, R. A., Stephan, K. E., Brown, H. R., Frith, C. D., \& Friston, K. J. (2013). The computational anatomy of psychosis. \emph{Frontiers in Psychiatry}, \emph{4}, 47.
+
 \leavevmode\vadjust pre{\hypertarget{ref-balota2011moving}{}}%
@@
 Fox, A. (2006). \emph{Adolescent self-development and psychopathology: Anorexia nervosa and psychosis} {[}PhD thesis{]}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-friston2010free}{}}%
+Friston, K. (2010). The free-energy principle: A unified brain theory? \emph{Nature Reviews Neuroscience}, \emph{11}(2), 127--138.
 
@@
 \url{https://doi.org/10.1016/j.neuropsychologia.2014.02.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-koethe2009binocular}{}}%
+Koethe, D., Kranaster, L., Hoyer, C., Gross, S., Neatby, M.
+A., Schultze-Lutter, F., Ruhrmann, S., Klosterkötter, J., Hellmich, M., \& Leweke, F.
+M.
+(2009).
+Binocular depth inversion as a paradigm of reduced visual information processing in prodromal state, antipsychotic-naive and treated schizophrenia.
+\emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{259}(4), 195--202.
 

### 2022-09-22 submission/manuscript.Rmd
update intro

@@
 
-Notably, the presence of a common mechanism underlying the effects of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions]; and the nature of the underlying processes - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] remain debated.
-The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism [@notredame2014; @park2022; @giaouri2011; @keane2014; @pessoa2008], and in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
+Notably, the presence of a common mechanism underlying the effects of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions].
+Indeed, while early research had suggested a common factor of illusion sensitivity indexed by overall vision proficiency [@halpern1999interindividual; @thurstone1944factorial], recent studies attempting to identify said factor have at most observed weak correlations between resistance to distinct illusions [@grzeczkowski2017; @grzeczkowski2018].
+To this end, the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] remain debated.
+To add, the existence of dispositional correlates of illusion sensitivity has also been another area of controversy, with some studies reporting higher illusion resistance in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
+In particular, an increasing amount of evidence in recent years purports lower illusion susceptibility in patients with schizophrenia and autism [@notredame2014; @park2022; @grzeczkowski2018; @gori2016; @razeghi2022].
 
@@
 
-One recent theory proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free], as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to provide a coherent perceptual experience.
+In an attempt to bridge the research gaps aforementioned, one recent theory proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free]; as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to provide a coherent perceptual experience.
 In this framework, certain traits or characteristics (e.g., psychoticism) are driven by alterations in the system's metacognitive components [@adams2013computational], resulting in underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions [@koethe2009binocular].
@@
 Despite strong theoretical foundations and hypotheses, the empirical evidence remains scarce, clouded by methodological hurdles.
-For instance, one key challenge is the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability.
-Moreover, the paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
-However, this reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity.
+For instance, one key challenge to empirical efforts involves the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability [@cretenoud2020illusions; @cretenoud2019z; @bressan2021most].
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
+However, this reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].
 
@@
 For each illusion type, the stimuli were split into two series (56 and 72 stimuli per series) with alternating parameter values to maintain their homogeneity.
-Additionally, 6 stimuli per illusion type was generated for a practice series, with more extreme variations (i.e., containing very easy trials to help cement the task instructions).
+Additionally, 6 stimuli per illusion type was generated for a practice series using parameters with more extreme variations (i.e., containing very easy trials to help cement the task instructions).
 
@@
 Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
-In total, each participant underwent 1340 trials of which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key  (left/right,or up/down depending on the illusion type).
+In total, each participant underwent 1340 trials of which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key  (left/right, or up/down depending on the illusion type).
 For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.

### 2022-09-23 submission/manuscript.Rmd
intro pass

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3808"
+wordcount         : "3876"
 bibliography      : references.bib
@@
 
-<!-- Intro  & Gaps-->
+<!-- Intro -->
 
@@
 
-Notably, the presence of a common mechanism underlying the effects of different illusions has been contested [@hamburger2016; @cretenoud2019; @cretenoud2020illusions].
-Indeed, while early research had suggested a common factor of illusion sensitivity indexed by overall vision proficiency [@halpern1999interindividual; @thurstone1944factorial], recent studies attempting to identify said factor have at most observed weak correlations between resistance to distinct illusions [@grzeczkowski2017; @grzeczkowski2018].
-To this end, the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences of prior beliefs [@teufel2018; @caporuscio2022] remain debated.
-To add, the existence of dispositional correlates of illusion sensitivity has also been another area of controversy, with some studies reporting higher illusion resistance in individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
-In particular, an increasing amount of evidence in recent years purports lower illusion susceptibility in patients with schizophrenia and autism [@notredame2014; @park2022; @grzeczkowski2018; @gori2016; @razeghi2022].
+<!-- Gaps-->
+
+One area of contention concerns the presence of a common mechanism underlying the effect of different illusions [@hamburger2016; @cretenoud2020illusions].
+While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency [@halpern1999interindividual; @thurstone1944factorial], recent empirical studies at most observed weak correlations between resistance to distinct illusions [@grzeczkowski2017; @grzeczkowski2018].
+The existence of dispositional correlates of illusion sensitivity has also been controversial, related to evidence suggesting a lower illusion sensitivity in in patients with schizophrenia and autism [@notredame2014; @park2022; @grzeczkowski2018; @gori2016; @razeghi2022], as well as individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
 
 <!-- Bayes -->
-
-In an attempt to bridge the research gaps aforementioned, one recent theory proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free]; as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to provide a coherent perceptual experience.
-In this framework, certain traits or characteristics (e.g., psychoticism) are driven by alterations in the system's metacognitive components [@adams2013computational], resulting in underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions [@koethe2009binocular].
+Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences [@teufel2018; @caporuscio2022] - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free], as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
+In this framework, certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components [@adams2013computational], resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions [@koethe2009binocular].
 
@@
 Despite strong theoretical foundations and hypotheses, the empirical evidence remains scarce, clouded by methodological hurdles.
-For instance, one key challenge to empirical efforts involves the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability [@cretenoud2020illusions; @cretenoud2019z; @bressan2021most].
+For instance, one key challenge can be found in the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability [@cretenoud2020illusions; @cretenoud2019; @bressan2021most].
 Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
-However, this reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].
+This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].
 
@@
 For each illusion type, the stimuli were split into two series (56 and 72 stimuli per series) with alternating parameter values to maintain their homogeneity.
-Additionally, 6 stimuli per illusion type was generated for a practice series using parameters with more extreme variations (i.e., containing very easy trials to help cement the task instructions).
+Additionally, 6 stimuli per illusion type were generated for a practice series using parameters with more extreme variations (i.e., containing very easy trials to help cement the task instructions).
 

### 2022-09-23 submission/manuscript.tex
intro pass

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3628}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3873}
 \usepackage{lineno}
@@
 
-Notably, the presence of a common mechanism underlying the effects of different illusions has been contested (Cretenoud et al., 2019; Cretenoud et al., 2020; Hamburger, 2016); and the nature of the underlying processes - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences of prior beliefs (Caporuscio et al., 2022; Teufel et al., 2018) remain debated.
-The existence of dispositional correlates of illusion sensitivity is another area of controversy, with some studies reporting higher illusion resistance in patients with schizophrenia and autism (Giaouri \& Alevriadou, 2011; Keane et al., 2014; Notredame et al., 2014; Park et al., 2022; Pessoa et al., 2008), and in individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
+One area of contention concerns the presence of a common mechanism underlying the effect of different illusions (Cretenoud et al., 2020; Hamburger, 2016).
+While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency (Halpern et al., 1999; Thurstone, 1944), recent empirical studies at most observed weak correlations between resistance to distinct illusions (Grzeczkowski et al., 2017, 2018).
+The existence of dispositional correlates of illusion sensitivity has also been controversial, related to evidence suggesting a lower illusion sensitivity in in patients with schizophrenia and autism (Gori et al., 2016; Grzeczkowski et al., 2018; Notredame et al., 2014; Park et al., 2022; Razeghi et al., 2022), as well as individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
 
-One recent theory proposes to conceptualize illusions under the Bayesian brain hypothesis (Friston, 2010), as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to provide a coherent perceptual experience.
-In this framework, certain traits or characteristics (e.g., psychoticism) are driven by alterations in the system's metacognitive components (Adams et al., 2013), resulting in underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions (Koethe et al., 2009).
+Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences (Caporuscio et al., 2022; Teufel et al., 2018) - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis (Friston, 2010), as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
+In this framework, certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components (Adams et al., 2013), resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions (Koethe et al., 2009).
 
 Despite strong theoretical foundations and hypotheses, the empirical evidence remains scarce, clouded by methodological hurdles.
-For instance, one key challenge is the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability.
-Moreover, the paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to perceptually match a reference stimulus (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
-However, this reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity.
+For instance, one key challenge can be found in the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
+Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability (Bressan \& Kramer, 2021; Cretenoud et al., 2019; Cretenoud et al., 2020).
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to perceptually match a reference stimulus (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
+This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity (Skottun \& Skoyles, 2014).
 
@@
 For each illusion type, the stimuli were split into two series (56 and 72 stimuli per series) with alternating parameter values to maintain their homogeneity.
-Additionally, 6 stimuli per illusion type was generated for a practice series, with more extreme variations (i.e., containing very easy trials to help cement the task instructions).
+Additionally, 6 stimuli per illusion type were generated for a practice series using parameters with more extreme variations (i.e., containing very easy trials to help cement the task instructions).
 
@@
 Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
-In total, each participant underwent 1340 trials of which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key (left/right,or up/down depending on the illusion type).
+In total, each participant underwent 1340 trials of which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key (left/right, or up/down depending on the illusion type).
 For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bressan2021most}{}}%
+Bressan, P., \& Kramer, P. (2021). Most findings obtained with untimed visual illusions are confounded. \emph{Psychological Science}, \emph{32}(8), 1238--1246.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-giaouri2011}{}}%
-Giaouri, S., \& Alevriadou, A.
-(2011).
-Are children with down syndrome susceptible to visual illusions?
-\emph{Procedia - Social and Behavioral Sciences}, \emph{15}, 1988--1992.
-\url{https://doi.org/10.1016/j.sbspro.2011.04.040}
-
 \leavevmode\vadjust pre{\hypertarget{ref-gomez-villa2022}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-halpern1999interindividual}{}}%
+Halpern, S. D., Andrews, T. J., \& Purves, D. (1999). Interindividual variation in human visual performance. \emph{Journal of Cognitive Neuroscience}, \emph{11}(5), 521--534.
+
 \leavevmode\vadjust pre{\hypertarget{ref-hamburger2016}{}}%
@@
 \url{https://doi.org/10.1037/a0026656}
-
-\leavevmode\vadjust pre{\hypertarget{ref-keane2014}{}}%
-Keane, B.
-P., Joseph, J., \& Silverstein, S.
-M.
-(2014).
-Late, not early, stages of Kanizsa shape perception are compromised in schizophrenia.
-\emph{Neuropsychologia}, \emph{56}, 302--311.
-\url{https://doi.org/10.1016/j.neuropsychologia.2014.02.001}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-skottun2014subjective}{}}%
+Skottun, B. C., \& Skoyles, J. R. (2014). Subjective criteria and illusions in visual testing: Some methodological limitations. \emph{Psychological Research}, \emph{78}(1), 136--140.
+
 \leavevmode\vadjust pre{\hypertarget{ref-teufel2018}{}}%
@@
 \url{https://doi.org/10.1073/pnas.1503916112}
+
+\leavevmode\vadjust pre{\hypertarget{ref-thurstone1944factorial}{}}%
+Thurstone, L. L. (1944). \emph{A factorial study of perception.}
 

### 2022-09-23 submission/manuscript.Rmd
update

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3876"
+wordcount         : "3873"
 bibliography      : references.bib
@@
 One area of contention concerns the presence of a common mechanism underlying the effect of different illusions [@hamburger2016; @cretenoud2020illusions].
-While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency [@halpern1999interindividual; @thurstone1944factorial], recent empirical studies at most observed weak correlations between resistance to distinct illusions [@grzeczkowski2017; @grzeczkowski2018].
-The existence of dispositional correlates of illusion sensitivity has also been controversial, related to evidence suggesting a lower illusion sensitivity in in patients with schizophrenia and autism [@notredame2014; @park2022; @grzeczkowski2018; @gori2016; @razeghi2022], as well as individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
+While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency [@halpern1999interindividual; @thurstone1944factorial], recent empirical studies observed at nost weak correlations between resistance to distinct illusions [@grzeczkowski2017; @grzeczkowski2018].
+The existence of dispositional correlates of illusion sensitivity has also been controversial, with evidence suggesting a lower illusion sensitivity in patients with schizophrenia and autism [@notredame2014; @park2022; @grzeczkowski2018; @gori2016; @razeghi2022], as well as individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
 
@@
 The inter-individual variability in the effect of illusion strength and its interaction with task difficulty was extracted from the models and used as participant-level scores.
-We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general of illusion sensitivity (Factor *i*).
+We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general factor of illusion sensitivity (Factor *i*).
 
@@
 For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants responded faster.
-While this is is not surprising (strong illusions are likely so effective in biasing perception that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
+While this is not surprising (strong illusions are likely so effective in biasing perception that it is "easier", i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
 As such, we decided not to use them for the individual scores analysis.
@@
 
-The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions (**Figure 3**).
+The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions.
 Anecdotal ($BF_{10} = 2.05$) and moderate evidence ($BF_{10} = 4.11$) was found for a negative correlation between screen size and the sensitivity to the White and the Contrast illusion, respectively.
 To test whether this result could be an artifact related to the highly skewed screen size distribution (caused by very few participants with extreme screen sizes), we re-ran a robust correlation (with rank-transformed values), which provided even stronger evidence in favor of the effect existence ($BF_{10} = 28.19$, $BF_{10} = 4.31$ for White and Contrast, respectively).
+
+The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females ($\Delta=-0.37$, 95% CI [-0.62, -0.13], $BF_{10} = 12.74$).
+We fitted Bayesian linear models with the education level entered as a monotonic predictor [appropriate for ordinal variables, @burkner2020modelling], which yielded no significant effects.
+For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
+These consistently suggested a significant positive linear relationship between age and Factor *i* ($pd=100\%$), as well as the sensitivity to Müller-Lyer ($pd=100\%$), Vertical-Horizontal ($pd=100\%$), Zöllner ($pd=100\%$) and Ebbinghaus ($pd=99\%$) illusions (**Figure 3**).
+
+Regarding "normal" personality traits, Bayesian correlations suggested substantial evidence in favor of a positive relationship between *Honesty-Humility* and Zöllner ($BF_{10} > 100$), Vertical-Horizontal ($BF_{10} = 9.78$) and the Factor *i* ($BF_{10} = 4.00$); as well as between *Agreeableness* and Vertical-Horizontal ($BF_{10} = 25.06$), Ponzo ($BF_{10} = 4.88$) and the Factor *i* ($BF_{10} = 19.65$).
 
@@
 ```
-
-The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females ($\Delta=-0.37$, 95% CI [-0.62, -0.13], $BF_{10} = 12.74$).
-We fitted Bayesian linear models with the education level entered as a monotonic predictor [appropriate for ordinal variables, @burkner2020modelling], which yielded no significant effects.
-For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
-These consistently suggested a significant positive linear relationship between age and Factor *i* ($pd=100\%$), as well as the sensitivity to Müller-Lyer ($pd=100\%$), Vertical-Horizontal ($pd=100\%$), Zöllner ($pd=100\%$) and Ebbinghaus ($pd=99\%$) illusions.
-
-Regarding "normal" personality traits, Bayesian correlations suggested substantial evidence in favor of a positive relationship between *Honesty-Humility* and Zöllner ($BF_{10} > 100$), Vertical-Horizontal ($BF_{10} = 9.78$) and the Factor *i* ($BF_{10} = 4.00$); as well as between *Agreeableness* and Vertical-Horizontal ($BF_{10} = 25.06$), Ponzo ($BF_{10} = 4.88$) and the Factor *i* ($BF_{10} = 19.65$).
 

### 2022-09-23 submission/manuscript.tex
update

@@
 One area of contention concerns the presence of a common mechanism underlying the effect of different illusions (Cretenoud et al., 2020; Hamburger, 2016).
-While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency (Halpern et al., 1999; Thurstone, 1944), recent empirical studies at most observed weak correlations between resistance to distinct illusions (Grzeczkowski et al., 2017, 2018).
-The existence of dispositional correlates of illusion sensitivity has also been controversial, related to evidence suggesting a lower illusion sensitivity in in patients with schizophrenia and autism (Gori et al., 2016; Grzeczkowski et al., 2018; Notredame et al., 2014; Park et al., 2022; Razeghi et al., 2022), as well as individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
+While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency (Halpern et al., 1999; Thurstone, 1944), recent empirical studies observed at nost weak correlations between resistance to distinct illusions (Grzeczkowski et al., 2017, 2018).
+The existence of dispositional correlates of illusion sensitivity has also been controversial, with evidence suggesting a lower illusion sensitivity in patients with schizophrenia and autism (Gori et al., 2016; Grzeczkowski et al., 2018; Notredame et al., 2014; Park et al., 2022; Razeghi et al., 2022), as well as individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
 
@@
 The inter-individual variability in the effect of illusion strength and its interaction with task difficulty was extracted from the models and used as participant-level scores.
-We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general of illusion sensitivity (Factor \emph{i}).
+We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general factor of illusion sensitivity (Factor \emph{i}).
 
@@
 For RTs, the GAMs suggested a consistent non-linear relationship between RT and illusion strength: as the illusion strength increases beyond a certain threshold, the participants responded faster.
-While this is is not surprising (strong illusions are likely so effective in biasing perception that it is ``easier'', i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
+While this is not surprising (strong illusions are likely so effective in biasing perception that it is ``easier'', i.e., faster, to make the wrong decision), the linear models were not designed to capture this - likely quadratic - pattern and hence are not good representatives of the underlying dynamics.
 As such, we decided not to use them for the individual scores analysis.
@@
 
-The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions (\textbf{Figure 3}).
+The Bayesian correlation analysis (with narrow priors centered around a null effect) between the illusion scores and contextual variables (screen size and refresh rate) provided weak evidence in favor of an absence of effect, with the exception of the two contrast-based illusions.
 Anecdotal (\(BF_{10} = 2.05\)) and moderate evidence (\(BF_{10} = 4.11\)) was found for a negative correlation between screen size and the sensitivity to the White and the Contrast illusion, respectively.
 To test whether this result could be an artifact related to the highly skewed screen size distribution (caused by very few participants with extreme screen sizes), we re-ran a robust correlation (with rank-transformed values), which provided even stronger evidence in favor of the effect existence (\(BF_{10} = 28.19\), \(BF_{10} = 4.31\) for White and Contrast, respectively).
+
+The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females (\(\Delta=-0.37\), 95\% CI {[}-0.62, -0.13{]}, \(BF_{10} = 12.74\)).
+We fitted Bayesian linear models with the education level entered as a monotonic predictor (appropriate for ordinal variables, Bürkner \& Charpentier, 2020), which yielded no significant effects.
+For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
+These consistently suggested a significant positive linear relationship between age and Factor \emph{i} (\(pd=100\%\)), as well as the sensitivity to Müller-Lyer (\(pd=100\%\)), Vertical-Horizontal (\(pd=100\%\)), Zöllner (\(pd=100\%\)) and Ebbinghaus (\(pd=99\%\)) illusions (\textbf{Figure 3}).
+
+Regarding ``normal'' personality traits, Bayesian correlations suggested substantial evidence in favor of a positive relationship between \emph{Honesty-Humility} and Zöllner (\(BF_{10} > 100\)), Vertical-Horizontal (\(BF_{10} = 9.78\)) and the Factor \emph{i} (\(BF_{10} = 4.00\)); as well as between \emph{Agreeableness} and Vertical-Horizontal (\(BF_{10} = 25.06\)), Ponzo (\(BF_{10} = 4.88\)) and the Factor \emph{i} (\(BF_{10} = 19.65\)).
 
@@
 \end{figure}
-
-The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females (\(\Delta=-0.37\), 95\% CI {[}-0.62, -0.13{]}, \(BF_{10} = 12.74\)).
-We fitted Bayesian linear models with the education level entered as a monotonic predictor (appropriate for ordinal variables, Bürkner \& Charpentier, 2020), which yielded no significant effects.
-For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
-These consistently suggested a significant positive linear relationship between age and Factor \emph{i} (\(pd=100\%\)), as well as the sensitivity to Müller-Lyer (\(pd=100\%\)), Vertical-Horizontal (\(pd=100\%\)), Zöllner (\(pd=100\%\)) and Ebbinghaus (\(pd=99\%\)) illusions.
-
-Regarding ``normal'' personality traits, Bayesian correlations suggested substantial evidence in favor of a positive relationship between \emph{Honesty-Humility} and Zöllner (\(BF_{10} > 100\)), Vertical-Horizontal (\(BF_{10} = 9.78\)) and the Factor \emph{i} (\(BF_{10} = 4.00\)); as well as between \emph{Agreeableness} and Vertical-Horizontal (\(BF_{10} = 25.06\)), Ponzo (\(BF_{10} = 4.88\)) and the Factor \emph{i} (\(BF_{10} = 19.65\)).
 

### 2022-09-23 submission/manuscript.Rmd
update

@@
 
-This allows the creation of an experimental task in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
+The systematic calibration of the stimuli with the use of pyllusion enables creation of experimental tasks in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
 Moreover, the illusion effect can specified as either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier).
@@
 The parametric illusion generation framework developed in @makowski2021 proposes to conceptualize illusions as composed of targets and distractors that can be manipulated independently and continuously.
+The goals of this research were ternary.
+First, to test and validate the novel paradigm, second to investigate the existence of a general factor underlying illusion sensitivity (Factor *I*), and finally to explore the relationship between illusion sensitivity and personality variables.
 In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
@@
 
-However, although the illusions did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+Although the illusions did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+
+Despite the robust results this research has provided, models using error-rates alone lead to over simplification and be a potential limitation of this research.
 Future studies should investigate the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions), and assess the psychometric properties - such as stability (e.g., test-retest reliability) and validity - of similar illusion-based paradigms.
@@
 
-Finally, we found the sensitivity to illusions to be positively associated with "positive" personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
+Finally, we found illusion sensitivity to be positively associated with "positive" personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
 Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs [often associated with psychoticism, @fox2006] to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli  [@ohmann2016; @konrath2009seeing; @fox2006].

### 2022-09-24 submission/manuscript.Rmd
another pass

@@
   Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings.
-In this study, we used the novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
+Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, leaving open the questions pertaining to a potential unique factor underlying the senstivity to different types of illusions.
+In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
 We tested the objective effect of the illusions on errors and reaction times, and extracted participant-level performance scores (n=250).
@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3873"
+wordcount         : "3977"
 bibliography      : references.bib
@@
 
-The systematic calibration of the stimuli with the use of pyllusion enables creation of experimental tasks in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
+This systematic calibration of the stimuli enables the creation of experimental tasks in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
 Moreover, the illusion effect can specified as either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier).
@@
 <!-- Parametric framework -->
-
-The parametric illusion generation framework developed in @makowski2021 proposes to conceptualize illusions as composed of targets and distractors that can be manipulated independently and continuously.
-The goals of this research were ternary.
-First, to test and validate the novel paradigm, second to investigate the existence of a general factor underlying illusion sensitivity (Factor *I*), and finally to explore the relationship between illusion sensitivity and personality variables.
-In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
+This study tested a novel illusion sensitivity task paradigm based on the parametric illusion generation framework [@makowski2021].
+Using the carefully generated stimuli in a perceptual decision task, we have shown that a gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
 Increasing the illusion strength led to an increase in error likelihood, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
@@
 
-Although the illusions did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
-
-Despite the robust results this research has provided, models using error-rates alone lead to over simplification and be a potential limitation of this research.
-Future studies should investigate the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions), and assess the psychometric properties - such as stability (e.g., test-retest reliability) and validity - of similar illusion-based paradigms.
-
 <!-- Personality -->
@@
 While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
+
+<!-- Limitations -->
+Future research is needed to address several limitations, such as investigating the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions).
+Although the illusions used in the present study did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+Moreover, using only the error rate models to extract individual-level scores might fail in capturing the whole range of behavioral dynamics.
+Future work should attempt at integrating the reaction times data (e.g., by jointly analyzing them using drift diffusion models), and assess the psychometric properties - such as stability (e.g., test-retest reliability) and validity - of similar illusion-based paradigms.
+Finally, while the personality measures used in this study highlight illusion sensitivity as an interesting measure rather than a mere perceptual artifact, further studies should test its relationship with more specific dispositional characteristics (e.g., autistic or schizotypal traits), cognitive styles and abilities, to help understanding the potential underlying mechanisms of these associations.
 

### 2022-09-24 submission/manuscript.tex
another pass

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3873}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3977}
 \usepackage{lineno}
@@
 Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings.
-In this study, we used the novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
+Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, leaving open the questions pertaining to a potential unique factor underlying the senstivity to different types of illusions.
+In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
 We tested the objective effect of the illusions on errors and reaction times, and extracted participant-level performance scores (n=250).
@@
 
-This allows the creation of an experimental task in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
+This systematic calibration of the stimuli enables the creation of experimental tasks in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
 Moreover, the illusion effect can specified as either ``incongruent'' (making the task more difficult by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
@@
 
-The parametric illusion generation framework developed in Makowski et al.
-(2021) proposes to conceptualize illusions as composed of targets and distractors that can be manipulated independently and continuously.
-In the present study, we have shown that such gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
+This study tested a novel illusion sensitivity task paradigm based on the parametric illusion generation framework (Makowski et al., 2021).
+Using the carefully generated stimuli in a perceptual decision task, we have shown that a gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
 Increasing the illusion strength led to an increase in error likelihood, as well as the average and spread of RTs (but only up to a point, after which participants become faster at responding with the wrong answer).
@@
 
-However, although the illusions did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
-Future studies should investigate the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions), and assess the psychometric properties - such as stability (e.g., test-retest reliability) and validity - of similar illusion-based paradigms.
-
-Finally, we found the sensitivity to illusions to be positively associated with ``positive'' personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
+Finally, we found illusion sensitivity to be positively associated with ``positive'' personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
 Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs (often associated with psychoticism, Fox, 2006) to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli (Fox, 2006; Konrath et al., 2009; Ohmann \& Burgmer, 2016).
@@
 While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
+
+Future research is needed to address several limitations, such as investigating the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions).
+Although the illusions used in the present study did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.
+Moreover, using only the error rate models to extract individual-level scores might fail in capturing the whole range of behavioral dynamics.
+Future work should attempt at integrating the reaction times data (e.g., by jointly analyzing them using drift diffusion models), and assess the psychometric properties - such as stability (e.g., test-retest reliability) and validity - of similar illusion-based paradigms.
+Finally, while the personality measures used in this study highlight illusion sensitivity as an interesting measure rather than a mere perceptual artifact, further studies should test its relationship with more specific dispositional characteristics (e.g., autistic or schizotypal traits), cognitive styles and abilities, to help understanding the potential underlying mechanisms of these associations.
 

### 2022-09-28 submission/manuscript.Rmd
minor

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3977"
+wordcount         : "3978"
 bibliography      : references.bib
@@
 They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience [@carbon2014].
-Despite the longstanding interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022; @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
+Despite the long-standing interest within the fields of visual perception [@day1972; @eagleman2001; @gomez-villa2022], consciousness science [@caporuscio2022; @lamme2020], and psychiatry [@notredame2014; @gori2016; @razeghi2022; @teufel2015], several important issues remain open.
 
@@
 <!-- Limitations -->
-Future research is needed to address several limitations, such as investigating the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions).
+Future research is needed to address several limitations.
+One key question concerns the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions).
 Although the illusions used in the present study did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.

### 2022-09-28 submission/manuscript.tex
minor

@@
 They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience (Carbon, 2014).
-Despite the longstanding interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science (Caporuscio et al., 2022; Lamme, 2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
+Despite the long-standing interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science (Caporuscio et al., 2022; Lamme, 2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
 
@@
 
-Future research is needed to address several limitations, such as investigating the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions).
+Future research is needed to address several limitations.
+One key question concerns the relationship of illusion sensitivity with perceptual abilities (e.g., using similar tasks, but without illusions).
 Although the illusions used in the present study did differ in terms of the perceptual task (contrast-based, size-estimation, angle-perception), the possibility of our general factor being driven by inter-individual perceptual skills variability (or other cognitive skills) cannot be discarded.

### 2022-09-28 study2/make_shortversion.R
minor

@@ -28,7 +28,7 @@ make_plot(gam_verticalhorizontal_err, gam_verticalhorizontal_rt,
           diff="Illusion_Difference=c(0.05, 0.225)",
           strength=c(-33, 33))
 
-# M??ller-Lyer
+# Muller-Lyer
 make_plot(gam_mullerlyer_err, gam_mullerlyer_rt,
           diff="Illusion_Difference=c(0.05, 0.35)",
           strength=c(-23, 23))

### 2022-10-08 submission/manuscript.Rmd
minor

@@
 ---
-title             : "**The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates**"
+title             : "**The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates**"
 shorttitle        : "Illusion Game Validation"
@@
   Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, leaving open the questions pertaining to a potential unique factor underlying the senstivity to different types of illusions.
+Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, leaving open the questions pertaining to a potential unique factor underlying the sensitivity to different types of illusions.
 In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.

### 2022-10-08 submission/manuscript.tex
minor

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3977}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3978}
 \usepackage{lineno}
@@
 \hypersetup{
-  pdftitle={The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates},
+  pdftitle={The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates},
   pdfauthor={Dominique Makowski1, An Shu Te1, Stephanie Kirk1, Ngoi Zi Liang1, \& S.H. Annabel Chen1, 2, 3, 4},
@@
 
-\title{\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence in Favour of a General Factor of Visual Illusion Sensitivity and Personality Correlates}}
+\title{\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates}}
 \author{Dominique Makowski\textsuperscript{1}, An Shu Te\textsuperscript{1}, Stephanie Kirk\textsuperscript{1}, Ngoi Zi Liang\textsuperscript{1}, \& S.H. Annabel Chen\textsuperscript{1, 2, 3, 4}}
@@
 Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, leaving open the questions pertaining to a potential unique factor underlying the senstivity to different types of illusions.
+Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, leaving open the questions pertaining to a potential unique factor underlying the sensitivity to different types of illusions.
 In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.

### 2022-10-08 study2/make_shortversion.R

@@ -43,3 +43,102 @@ summary(get_data(gam_poggendorff_err)$Illusion_Difference)
 make_plot(gam_poggendorff_err, gam_poggendorff_rt,
           diff="Illusion_Difference=c(0.002, 0.17)",
           strength=c(-28, 28))
+
+
+
+# ISI ---------------------------------------------------------------------
+library(tidyverse)
+library(ggdist)
+library(ggside)
+library(easystats)
+library(patchwork)
+library(brms)
+
+df <- read.csv("../data/study2_part1.csv") |>
+  rbind(read.csv("../data/study2_part2.csv")) |>
+  mutate(
+    Date = lubridate::dmy(Date),
+    Participant = fct_reorder(Participant, Date),
+    Screen_Refresh = as.character(Screen_Refresh),
+    Illusion_Side = as.factor(Illusion_Side),
+    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
+    Block = as.factor(Block)
+  )
+
+
+
+outliers <- c(
+  # Error rate of 48.8% Very short RT
+  # Prolific Status: REJECTED (06/08)
+  "S46",
+  # 2nd block of responses very fast
+  # Prolific Status: REJECTED (15/08)
+  "S221",
+  # Error rate of 44% and very short RTs
+  # Prolific Status: RETURN REQUESTED (22/08)
+  "S154",
+  # 2nd block bad, first block 1/3 bad
+  # Prolific Status: RETURN REQUESTED (26/08)
+  "S68",
+  # Prolific Status: RETURN REQUESTED (26/08)
+  "S238",
+  # Prolific status: accepted (not enough proof)
+  "S201"
+)
+
+partial_outliers <- c(
+  # 2nd block a bit bad
+  "S22",
+  # Entire 2nd block bad
+  "S235",
+  # Entire 2nd block bad
+  "S107",
+  # Half of 2nd block bad
+  "S204",
+  # 2nd block not good
+  "S140")
+
+
+df <- filter(df, !Participant %in% outliers)
+
+
+df <- df |>
+  group_by(Participant, Illusion_Type, Block) |>
+  mutate(ErrorRate_per_block = sum(Error) / n()) |>
+  ungroup() |>
+  filter(ErrorRate_per_block < 0.5) |>
+  select(-ErrorRate_per_block)
+
[…]

### 2022-10-17 submission/manuscript.Rmd
minor

@@
   Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (dom.makowski@gmail.com).
+
 abstract: |
@@
 
-<!-- Significance Statement (< 120 words) -->
-<!-- A novel paradigm to study the objective effect of visual illusions is presented, which yielded evidence in favor of a common factor to visual illusions (Factor *i*) and a relationship between a low illusion sensitivity and maladaptive personality traits, such as antagonism, psychoticism and disinhibition.
--->
+**Significance Statement.** A novel paradigm to study the objective effect of visual illusions yielded evidence in favor of a common factor to visual illusions (Factor *i*) and a relationship between illusion resistance and maladaptive personality traits, such as antagonism, psychoticism and disinhibition.
 
@@
 ```{r message=FALSE, warning=FALSE, fig.cap="The upper plots show the distribution of demographic and dispositional variables.
-The middle plots shows the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
+The middle plots show the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
 Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \\textit{i}) and personality traits.", out.width="100%"}
@@
 
+# Funding
+
+This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).
+
 # Acknowledgments

### 2022-10-17 submission/manuscript.tex
minor

@@
 \hypersetup{
-  pdftitle={The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates},
+  pdftitle={A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates},
   pdfauthor={Dominique Makowski1, An Shu Te1, Stephanie Kirk1, Ngoi Zi Liang1, \& S.H. Annabel Chen1, 2, 3, 4},
@@
 
-\title{\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates}}
+\title{\textbf{A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates}}
 \author{Dominique Makowski\textsuperscript{1}, An Shu Te\textsuperscript{1}, Stephanie Kirk\textsuperscript{1}, Ngoi Zi Liang\textsuperscript{1}, \& S.H. Annabel Chen\textsuperscript{1, 2, 3, 4}}
@@
 \maketitle
+
+\textbf{Significance Statement.} A novel paradigm to study the objective effect of visual illusions yielded evidence in favor of a common factor to visual illusions (Factor \emph{i}) and a relationship between illusion resistance and maladaptive personality traits, such as antagonism, psychoticism and disinhibition.
 
@@
 \includegraphics[width=1\linewidth]{../figures/Figure3} \caption{The upper plots show the distribution of demographic and dispositional variables.
-The middle plots shows the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
+The middle plots show the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
 Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \textit{i}) and personality traits.}\label{fig:unnamed-chunk-4}

### 2022-10-18 submission/manuscript.Rmd
include Paul's comments
@cratlagh thanks a lot :)

@@
   Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, leaving open the questions pertaining to a potential unique factor underlying the sensitivity to different types of illusions.
+Unfortunately, illusion research has been hampered by the difficulty to experimentally manipulate these stimuli, leaving unanswered the hypothesis of a potential unique factor underlying the sensitivity to different types of illusions.
 In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
@@
 One area of contention concerns the presence of a common mechanism underlying the effect of different illusions [@hamburger2016; @cretenoud2020illusions].
-While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency [@halpern1999interindividual; @thurstone1944factorial], recent empirical studies observed at nost weak correlations between resistance to distinct illusions [@grzeczkowski2017; @grzeczkowski2018].
+While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency [@halpern1999interindividual; @thurstone1944factorial], recent empirical studies observed at most weak correlations between inter-individual resistance to distinct illusions [@grzeczkowski2017; @grzeczkowski2018].
 The existence of dispositional correlates of illusion sensitivity has also been controversial, with evidence suggesting a lower illusion sensitivity in patients with schizophrenia and autism [@notredame2014; @park2022; @grzeczkowski2018; @gori2016; @razeghi2022], as well as individuals with stronger aggression and narcissism traits [@zhang2017; @konrath2009seeing].
@@
 <!-- Bayes -->
-Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences [@teufel2018; @caporuscio2022] - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free], as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
-In this framework, certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components [@adams2013computational], resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions [@koethe2009binocular].
+Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences [@teufel2018; @caporuscio2022] - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free].
+In this context, illusions are conceptualized as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
+The predictive coding account further provides an explanation regarding the observations from clinical populations.
+Certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components [@adams2013computational], resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions [@koethe2009binocular].
 
@@
 This systematic calibration of the stimuli enables the creation of experimental tasks in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can specified as either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier).
+Moreover, the illusion effect can be specified as either "incongruent" (making the task more difficult by biasing the perception in the opposite way) or "congruent" (making the task easier).
 Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the "objective" effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
@@
 
-A pilot study (*n = 46*), of which full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
+A pilot study (*n = 46*), of which a full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
 Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
@@
 The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
-Participants were incentivised with a reward of about \textsterling 7.5 for completing the task, which took about 50 minutes to finish.
+Participants were incentivised with a reward of about \textsterling 7.50 for completing the task, which took about 50 minutes to finish.
 Demographic variables (age, gender, and ethnicity) were self-reported on a voluntary basis.
@@
 
-The final sample included 250 participants (Mean age = 26.5, SD = 7.6, range: [18, 69]; Sex: 48% females, 52% males).
+The final sample included 250 participants (Mean age = 26.5, SD = 7.6, range: [18 - 69]; Sex: 48% females, 52% males).
 
@@
 
-The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and reaction time (RT) within each illusion.
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) within each illusion.
 We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions.
@@
 Future work should attempt at integrating the reaction times data (e.g., by jointly analyzing them using drift diffusion models), and assess the psychometric properties - such as stability (e.g., test-retest reliability) and validity - of similar illusion-based paradigms.
-Finally, while the personality measures used in this study highlight illusion sensitivity as an interesting measure rather than a mere perceptual artifact, further studies should test its relationship with more specific dispositional characteristics (e.g., autistic or schizotypal traits), cognitive styles and abilities, to help understanding the potential underlying mechanisms of these associations.
+Finally, while the personality measures used in this study highlight illusion sensitivity as an interesting measure rather than a mere perceptual artifact, further studies should test its relationship with more specific dispositional characteristics (e.g., autistic or schizotypal traits), cognitive styles and abilities, to help understand the potential underlying mechanisms of these associations.
 

### 2022-10-20 submission/manuscript.Rmd
add coren lit

@@
 Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability [@cretenoud2020illusions; @cretenoud2019; @bressan2021most].
-Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], having them estimate their perceptions of the targets' parametric properties (e.g., diameter of circular targets in the Ebbinghaus illusion) [@coren1976empirical] or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
 This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].

### 2022-10-22 study2/preprocessing.R
remove data with ID

@@ -95,14 +95,6 @@ preprocess_raw <- function(file) {
   # Correct duration
   df$Duration <- df$Duration - df$Break_Duration
 
-  # Manual fixes for reported errors (using demographic info from prolific)
-  df[df$Participant == "5d3c6e745602310001bca8aa_6bdtb", "Sex"] <- "Female"
-  df[df$Participant == "60f5fa488f7b1381d175ecb5_qwvfs", "Sex"] <- "Male"
-  df[df$Participant == "5d558c262d8a6200172a8b7b_deegh", "Sex"] <- "Male"
-  df[df$Participant == "614a4c9d8172751921452626_mc4l4", "Sex"] <- "Female"
-  df[df$Participant == "61323c38f727b9f84e45a9ef_ls925", "Sex"] <- "Female"
-  df[df$Participant == "61568a59110d8f1cc99558af_8tqre", "Sex"] <- "Female"
-
 
   # Format names
   df$Illusion_Type <- ifelse(df$Illusion_Type == "MullerLyer", "Müller-Lyer", df$Illusion_Type)

### 2022-10-22 submission/manuscript.Rmd
minor

--- 2022-10-18 submission/manuscript.Rmd
@@
 Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability [@cretenoud2020illusions; @cretenoud2019; @bressan2021most].
-Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], having them estimate the targets' physical properties [@coren1976empirical], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
 This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].

### 2022-10-26 submission/manuscript.Rmd
scirep

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3978"
+wordcount         : "4070"
 bibliography      : references.bib

### 2022-10-26 submission/manuscript.tex
scirep

--- 2022-10-08 submission/manuscript.tex
@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3978}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4070}
 \usepackage{lineno}
@@
 Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty of adapting these stimuli to experimental settings, leaving open the questions pertaining to a potential unique factor underlying the sensitivity to different types of illusions.
+Unfortunately, illusion research has been hampered by the difficulty to experimentally manipulate these stimuli, leaving unanswered the hypothesis of a potential unique factor underlying the sensitivity to different types of illusions.
 In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
@@
 
+\textbf{Significance Statement.} A novel paradigm to study the objective effect of visual illusions yielded evidence in favor of a common factor to visual illusions (Factor \emph{i}) and a relationship between illusion resistance and maladaptive personality traits, such as antagonism, psychoticism and disinhibition.
+
 \hypertarget{introduction}{%
@@
 One area of contention concerns the presence of a common mechanism underlying the effect of different illusions (Cretenoud et al., 2020; Hamburger, 2016).
-While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency (Halpern et al., 1999; Thurstone, 1944), recent empirical studies observed at nost weak correlations between resistance to distinct illusions (Grzeczkowski et al., 2017, 2018).
+While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency (Halpern et al., 1999; Thurstone, 1944), recent empirical studies observed at most weak correlations between inter-individual resistance to distinct illusions (Grzeczkowski et al., 2017, 2018).
 The existence of dispositional correlates of illusion sensitivity has also been controversial, with evidence suggesting a lower illusion sensitivity in patients with schizophrenia and autism (Gori et al., 2016; Grzeczkowski et al., 2018; Notredame et al., 2014; Park et al., 2022; Razeghi et al., 2022), as well as individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
 
-Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences (Caporuscio et al., 2022; Teufel et al., 2018) - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis (Friston, 2010), as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
-In this framework, certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components (Adams et al., 2013), resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions (Koethe et al., 2009).
+Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences (Caporuscio et al., 2022; Teufel et al., 2018) - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis (Friston, 2010).
+In this context, illusions are conceptualized as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
+The predictive coding account further provides an explanation regarding the observations from clinical populations.
+Certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components (Adams et al., 2013), resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions (Koethe et al., 2009).
 
@@
 Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability (Bressan \& Kramer, 2021; Cretenoud et al., 2019; Cretenoud et al., 2020).
-Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to perceptually match a reference stimulus (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), having them estimate the targets' physical properties (Coren et al., 1976), or having them adjust the targets to perceptually match a reference stimulus (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
 This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity (Skottun \& Skoyles, 2014).
@@
 This systematic calibration of the stimuli enables the creation of experimental tasks in which participants make perceptual judgments about the targets (e.g., which segment is the longest) under different conditions of objective difficulty and illusion strength.
-Moreover, the illusion effect can specified as either ``incongruent'' (making the task more difficult by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
+Moreover, the illusion effect can be specified as either ``incongruent'' (making the task more difficult by biasing the perception in the opposite way) or ``congruent'' (making the task easier).
 Although visual illusions are inherently tied to subjective perception, this framework allows a reversal of the traditional paradigm to potentially quantify the ``objective'' effect of illusions by measuring its behavioral effect (error rate and reaction times) on the performance in a perceptual task.
@@
 
-A pilot study (\emph{n = 46}), of which full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
+A pilot study (\emph{n = 46}), of which a full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
 Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
@@
 
-The final sample included 250 participants (Mean age = 26.5, SD = 7.6, range: {[}18, 69{]}; Sex: 48\% females, 52\% males).
+The final sample included 250 participants (Mean age = 26.5, SD = 7.6, range: {[}18 - 69{]}; Sex: 48\% females, 52\% males).
 
@@
 The task, data and analysis script are available in open-access at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
+
+\hypertarget{funding}{%
+\section{Funding}\label{funding}}
+
+This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).
 
@@
 Carbon, C.-C. (2014). Understanding human perception by human-made illusions. \emph{Frontiers in Human Neuroscience}, \emph{8}. \url{https://www.frontiersin.org/articles/10.3389/fnhum.2014.00566}
+
+\leavevmode\vadjust pre{\hypertarget{ref-coren1976empirical}{}}%
+Coren, S., Girgus, J. S., Erlichman, H., \& Hakstian, A. R. (1976). An empirical taxonomy of visual illusions. \emph{Perception \& Psychophysics}, \emph{20}(2), 129--137.
 

### 2022-10-31 submission/manuscript.Rmd
reformat to follow Nature reg

@@
 
-The aim of the present preregistered study is three-fold.
+The aim of the present study is three-fold.
 First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
@@
 
-Following open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as **Supplementary Materials** at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+This study has been preregistered to the Open Science Framework (OSF; [**https://osf.io/5d6xp**](https://osf.io/5d6xp{.uri})).
+
+<!-- Following open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as **Supplementary Materials** at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+-->
 
@@
 
+## Ethics Statement
+
+This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187) and all procedures performed were in accordance with the ethical standards of the institutional board and with the 1964 Helsinki Declaration.
+All participants provided their informed consent prior to participation and were incentivized after completing the study.
+
 ## Stimuli
 
-A pilot study (*n = 46*), of which a full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
+A pilot study (*n = 46*), of which a full description is available in the GitHub repository at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}, was first conducted to determine a sensitive range of stimuli parameters.
 Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
@@
 The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle].
-As the full results are available as supplementary materials, we will focus here on the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
+As the full results are available in the study documentation at the dedicated GitHub repository [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}, we will focus here on the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
 
@@
 For error rates, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
-A specific discussion regarding these 2 illusions is available in the Supplementary Materials (Part 1 - Discussion).
+A specific discussion regarding these 2 illusions is available in the study 1 documentation (Part 1 - Discussion) at the GitHub repository [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
 
@@
 The last remaining trait, *Detachment*, did not share any significant relationship with illusion sensitivity.
-See Supplementary Materials (Part 2 - Discussion) for a detailed discussion regarding these associations.
+See study 1 documentation (Part 2 - Discussion) in the GitHub repository [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri} for a detailed discussion regarding these associations.
 
@@
 
+# Data Availability
+
 In conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.

### 2022-11-01 submission/manuscript.Rmd
citations style

--- 2022-10-26 submission/manuscript.Rmd
@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "4070"
+wordcount         : "3826"
 bibliography      : references.bib
@@
 output            : papaja::apa6_pdf
-csl: utils/apa.csl
+csl: utils/scientific-reports.csl
 header-includes:
@@
 
-The aim of the present preregistered study is three-fold.
+The aim of the present preregistered (https://osf.io/5d6xp) study is three-fold.
 First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
@@
 
-Following open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as **Supplementary Materials** at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+<!-- Following open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+-->
 
@@
 
+## Ethics Statement
+
+This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187) and all procedures performed were in accordance with the ethical standards of the institutional board and with the 1964 Helsinki Declaration.
+All participants provided their informed consent prior to participation and were incentivized after completing the study.
+
 ## Stimuli
 
-A pilot study (*n = 46*), of which a full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
+A pilot study (*n = 46*), of which a full description is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}, was first conducted to determine a sensitive range of stimuli parameters.
 Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
@@
 The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle].
-As the full results are available as supplementary materials, we will focus here on the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
+As all the full results have been made available (see **Data Availability**), we will focus here on the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
 
@@
 For error rates, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
-A specific discussion regarding these 2 illusions is available in the Supplementary Materials (Part 1 - Discussion).
+A specific discussion regarding these 2 illusions is available in the study documentation (part 1) at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
 
@@
 The last remaining trait, *Detachment*, did not share any significant relationship with illusion sensitivity.
-See Supplementary Materials (Part 2 - Discussion) for a detailed discussion regarding these associations.
+<!-- See study 1 documentation (Part 2 - Discussion) in the GitHub repository [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri} for a detailed discussion regarding these associations.
+-->
 
@@
 
-In conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
+<!-- In conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
 The task, data and analysis script are available in open-access at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}.
+-->
+
+# Data Availability
+
+The datasets generated and/or analysed during the current study are available in the GitHub repository https://github.com/RealityBending/IllusionGameValidation
 

### 2022-11-01 submission/manuscript.tex
citations style

@@
 Visual illusions are fascinating stimuli capturing a key feature of our neurocognitive systems.
-They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience (Carbon, 2014).
-Despite the long-standing interest within the fields of visual perception (Day, 1972; Eagleman, 2001; Gomez-Villa et al., 2022), consciousness science (Caporuscio et al., 2022; Lamme, 2020), and psychiatry (Gori et al., 2016; Notredame et al., 2014; Razeghi et al., 2022; Teufel et al., 2015), several important issues remain open.
+They eloquently show that our brains did not evolve to be perfect perceptual devices providing veridical accounts of physical reality, but integrate prior knowledge and contextual information - blended together in our subjective conscious experience\textsuperscript{1}.
+Despite the long-standing interest within the fields of visual perception\textsuperscript{2--4}, consciousness science\textsuperscript{5,6}, and psychiatry\textsuperscript{7--10}, several important issues remain open.
 
-One area of contention concerns the presence of a common mechanism underlying the effect of different illusions (Cretenoud et al., 2020; Hamburger, 2016).
-While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency (Halpern et al., 1999; Thurstone, 1944), recent empirical studies observed at most weak correlations between inter-individual resistance to distinct illusions (Grzeczkowski et al., 2017, 2018).
-The existence of dispositional correlates of illusion sensitivity has also been controversial, with evidence suggesting a lower illusion sensitivity in patients with schizophrenia and autism (Gori et al., 2016; Grzeczkowski et al., 2018; Notredame et al., 2014; Park et al., 2022; Razeghi et al., 2022), as well as individuals with stronger aggression and narcissism traits (Konrath et al., 2009; Zhang et al., 2017).
+One area of contention concerns the presence of a common mechanism underlying the effect of different illusions\textsuperscript{11,12}.
+While early research has suggested a common factor of illusion sensitivity indexed by overall vision proficiency\textsuperscript{13,14}, recent empirical studies observed at most weak correlations between inter-individual resistance to distinct illusions\textsuperscript{15,16}.
+The existence of dispositional correlates of illusion sensitivity has also been controversial, with evidence suggesting a lower illusion sensitivity in patients with schizophrenia and autism\textsuperscript{7--9,16,17}, as well as individuals with stronger aggression and narcissism traits\textsuperscript{18,19}.
 
-Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system (Cretenoud et al., 2019; Gori et al., 2016) or to top-down influences (Caporuscio et al., 2022; Teufel et al., 2018) - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis (Friston, 2010).
+Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system\textsuperscript{8,20} or to top-down influences\textsuperscript{5,21} - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis\textsuperscript{22}.
 In this context, illusions are conceptualized as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
 The predictive coding account further provides an explanation regarding the observations from clinical populations.
-Certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components (Adams et al., 2013), resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions (Koethe et al., 2009).
+Certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components\textsuperscript{23}, resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions\textsuperscript{24}.
 
@@
 For instance, one key challenge can be found in the difficulty of adapting visual illusions to an experimental setting, which typically requires the controlled modulation of the specific variables of interest.
-Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability (Bressan \& Kramer, 2021; Cretenoud et al., 2019; Cretenoud et al., 2020).
-Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), having them estimate the targets' physical properties (Coren et al., 1976), or having them adjust the targets to perceptually match a reference stimulus (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
-This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity (Skottun \& Skoyles, 2014).
+Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability\textsuperscript{12,20,25}.
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different\textsuperscript{26}, having them estimate the targets' physical properties\textsuperscript{27}, or having them adjust the targets to perceptually match a reference stimulus\textsuperscript{16,28}.
+This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity\textsuperscript{29}.
 
-To address these issues, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software \emph{Pyllusion} (Makowski et al., 2021).
+To address these issues, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software \emph{Pyllusion}\textsuperscript{30}.
 This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (\textbf{Figure 1}).
 Indeed, many visual illusions can be seen as being composed of \emph{targets} (e.g., same-length lines), of which perception is biased by the \emph{context} (e.g., in the Müller-Lyer illusion, the same-length line segments appear to have different lengths if they end with inwards vs.~outwards pointing arrows).
-Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different (Lányi et al., 2022), or having them adjust the targets to match a reference stimulus relying only on their perception (Grzeczkowski et al., 2018; Mylniec \& Bednarek, 2016).
+Past illusion studies traditionally employed paradigms focusing on participants' subjective experience, by asking them the extent to which they perceive two identical targets as different\textsuperscript{26}, or having them adjust the targets to match a reference stimulus relying only on their perception\textsuperscript{16,28}.
 Alternatively, \emph{Pyllusion} allows the creation of illusions in which the targets are objectively different (e.g., one segment is truly more or less longer than the other), and in which the illusion varies in strength (the biasing angle of the arrows is more or less acute).
@@
 
-The aim of the present preregistered study is three-fold.
+The aim of the present preregistered (\url{https://osf.io/5d6xp}) study is three-fold.
 First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
@@
 
-Following open-science standards, all the material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, preregistration, etc.) is available as \textbf{Supplementary Materials} at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
-
 \hypertarget{methods}{%
 \section{Methods}\label{methods}}
+
+\hypertarget{ethics-statement}{%
+\subsection{Ethics Statement}\label{ethics-statement}}
+
+This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187) and all procedures performed were in accordance with the ethical standards of the institutional board and with the 1964 Helsinki Declaration.
+All participants provided their informed consent prior to participation and were incentivized after completing the study.
 
@@
 
-A pilot study (\emph{n = 46}), of which a full description is available in the Supplementary Materials, was first conducted to determine a sensitive range of stimuli parameters.
+A pilot study (\emph{n = 46}), of which a full description is available at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}, was first conducted to determine a sensitive range of stimuli parameters.
 Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
@@
 After a brief demographic survey and a practice series of illusions, the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
-Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6} (24 items, Sibley et al., 2011) - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5} (25 items, Hopwood et al., 2012) - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6}\textsuperscript{24 items, 31} - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5}\textsuperscript{25 items, 32} - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
 Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
@@
 For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
-The task was implemented using \emph{jsPsych} (De Leeuw, 2015), and the set of instructions for each illusion type is available in the experiment code.
+The task was implemented using \emph{jsPsych}\textsuperscript{33}, and the set of instructions for each illusion type is available in the experiment code.
 
@@
 
-Participants were recruited via \emph{Prolific}, a crowd-sourcing platform recognized for providing high quality data (Peer et al., 2022).
+Participants were recruited via \emph{Prolific}, a crowd-sourcing platform recognized for providing high quality data\textsuperscript{34}.
 The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
@@
 We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions.
-Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location \(\mu\) (mean), scale \(\sigma\) (spread) and tail-dominance \(\tau\) of the RT distribution (Balota \& Yap, 2011; Matzke \& Wagenmakers, 2009).
+Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location \(\mu\) (mean), scale \(\sigma\) (spread) and tail-dominance \(\tau\) of the RT distribution\textsuperscript{35,36}.
 
@@
 
-The analysis was carried out using \emph{R 4.2} (R Core Team, 2022), \emph{brms} (Bürkner, 2017), the \emph{tidyverse} (Wickham et al., 2019), and the \emph{easystats} collection of packages (Lüdecke et al., 2021, 2019; Makowski et al., 2020; Makowski, Ben-Shachar, \& Lüdecke, 2019).
-As the full results are available as supplementary materials, we will focus here on the significant results (based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see Makowski, Ben-Shachar, Chen, et al., 2019).
+The analysis was carried out using \emph{R 4.2}\textsuperscript{37}, \emph{brms}\textsuperscript{38}, the \emph{tidyverse}\textsuperscript{39}, and the \emph{easystats} collection of packages\textsuperscript{40--43}.
+As all the full results have been made available (see \textbf{Data Availability}), we will focus here on the significant results\textsuperscript{based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see 44}.
 
@@
 For error rates, most of the models closely matched their GAMs counterpart, with the exception of Delboeuf (for which the GAM suggested a non-monotonic effect of illusion strength with a local minimum at 0) and Zöllner (for which theoretically congruent illusion effects were related to increased error rate).
-A specific discussion regarding these 2 illusions is available in the Supplementary Materials (Part 1 - Discussion).
+A specific discussion regarding these 2 illusions is available in the study documentation (part 1) at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
 
@@
 These twenty participant-level scores were subjected to exploratory factor analysis (EFA).
-The Method Agreement Procedure (Lüdecke et al., 2020) suggested the presence of 7 latent factors.
+The Method Agreement Procedure\textsuperscript{45} suggested the presence of 7 latent factors.
 An oblique (\emph{oblimin} rotation) factor solution explaining 66.69\% of variance suggested separate dimensions for the effect of Zöllner, White, Poggendorff, Contrast, Ebbinghaus, Delboeuf, and a common factor for the parameters related to Müller-Lyer, Vertical-Horizontal, Ponzo and Rod and Frame.
@@
 The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females (\(\Delta=-0.37\), 95\% CI {[}-0.62, -0.13{]}, \(BF_{10} = 12.74\)).
-We fitted Bayesian linear models with the education level entered as a monotonic predictor (appropriate for ordinal variables, Bürkner \& Charpentier, 2020), which yielded no significant effects.
+We fitted Bayesian linear models with the education level entered as a monotonic predictor\textsuperscript{appropriate for ordinal variables, 46}, which yielded no significant effects.
 For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
@@
 The last remaining trait, \emph{Detachment}, did not share any significant relationship with illusion sensitivity.
-See Supplementary Materials (Part 2 - Discussion) for a detailed discussion regarding these associations.
 
@@
 
-This study tested a novel illusion sensitivity task paradigm based on the parametric illusion generation framework (Makowski et al., 2021).
+This study tested a novel illusion sensitivity task paradigm based on the parametric illusion generation framework\textsuperscript{30}.
 Using the carefully generated stimuli in a perceptual decision task, we have shown that a gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
@@
 This result comes in a field of mixed findings.
-In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals (Cretenoud et al., 2019; Cretenoud et al., 2020; Grzeczkowski et al., 2017, 2018; Yang et al., 2012).
-Instead, past findings suggest illusory effects are highly specific to the perceptual features of the illusions at stake (Cretenoud et al., 2019; Grzeczkowski et al., 2017).
+In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals\textsuperscript{12,15,16,20,47}.
+Instead, past findings suggest illusory effects are highly specific to the perceptual features of the illusions at stake\textsuperscript{15,20}.
 It should be noted, however, that most of these studies were low-powered and/or relied on conventional paradigms, such as the adjustment procedure to measure the participants' subjective perception.
@@
 Finally, we found illusion sensitivity to be positively associated with ``positive'' personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
-Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs (often associated with psychoticism, Fox, 2006) to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli (Fox, 2006; Konrath et al., 2009; Ohmann \& Burgmer, 2016).
+Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs\textsuperscript{often associated with psychoticism, 48} to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli\textsuperscript{19,48,49}.
 As such, the association between maladaptive traits and lower illusion sensitivity could be linked to a self-centered, decontextualized and disorganized information processing style.
-Conversely, the relationship between illusion sensitivity and adaptive personality traits is in line with the decreased field dependence (the tendency to rely on external cues in ambiguous contexts) associated with traits negatively correlated with agreeableness and honesty-humility, such as hostility, aggression and narcissism (Konrath et al., 2009; Pessoa et al., 2008; Zhang et al., 2017).
+Conversely, the relationship between illusion sensitivity and adaptive personality traits is in line with the decreased field dependence (the tendency to rely on external cues in ambiguous contexts) associated with traits negatively correlated with agreeableness and honesty-humility, such as hostility, aggression and narcissism\textsuperscript{18,19,50}.
 
 Importantly, these findings highlight the relevance of illusions beyond the field of visual perception, pointing towards an association with high-level domain-general mechanisms.
-In particular, the evidence in favor of a relationship between maladaptive personality traits and illusion sensitivity is in line with clinical observations, in which a greater resistance to illusions have been reported among patients with schizophrenia (Grzeczkowski et al., 2018; Notredame et al., 2014; Pessoa et al., 2008), especially in association with schizotypal traits such as cognitive disorganization (Cretenoud et al., 2019; Lányi et al., 2022).
+In particular, the evidence in favor of a relationship between maladaptive personality traits and illusion sensitivity is in line with clinical observations, in which a greater resistance to illusions have been reported among patients with schizophrenia\textsuperscript{7,16,50}, especially in association with schizotypal traits such as cognitive disorganization\textsuperscript{20,26}.
 While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
@@
 
-In conclusion, we strongly invite researchers to explore and re-analyze our dataset with other approaches and methods to push the understanding of visual illusions and illusion sensitivity further.
-The task, data and analysis script are available in open-access at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}.
+\hypertarget{data-availability}{%
+\section{Data Availability}\label{data-availability}}
+
+The datasets generated and/or analysed during the current study are available in the GitHub repository \url{https://github.com/RealityBending/IllusionGameValidation}
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
+\begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-carbon2014}{}}%
+\CSLLeftMargin{1. }%
+\CSLRightInline{Carbon, C.-C.
+\href{https://www.frontiersin.org/articles/10.3389/fnhum.2014.00566}{Understanding human perception by human-made illusions}.
+\emph{Frontiers in Human Neuroscience} \textbf{8}, (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-day1972}{}}%
+\CSLLeftMargin{2. }%
+\CSLRightInline{Day, R.
+H.
+\href{https://doi.org/10.1126/science.175.4028.1335}{Visual Spatial Illusions: A General Explanation: A wide range of visual illusions, including geometrical distortions, can be explained by a single principle.} \emph{Science} \textbf{175}, 1335--1340 (1972).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-eagleman2001}{}}%
+\CSLLeftMargin{3. }%
+\CSLRightInline{Eagleman, D. M. \href{https://doi.org/10.1038/35104092}{Visual illusions and neurobiology}. \emph{Nature Reviews Neuroscience} \textbf{2}, 920--926 (2001).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gomez-villa2022}{}}%
+\CSLLeftMargin{4. }%
+\CSLRightInline{Gomez-Villa, A., Martín, A., Vazquez-Corral, J., Bertalmío, M.
+\& Malo, J.
+\href{https://doi.org/10.1167/jov.22.8.2}{On the synthesis of visual illusions using deep generative models}.
+\emph{Journal of Vision} \textbf{22}, 2 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-caporuscio2022}{}}%
+\CSLLeftMargin{5. }%
+\CSLRightInline{Caporuscio, C., Fink, S.
+B., Sterzer, P.
+\& Martin, J.
+M.
+\href{https://doi.org/10.1016/j.concog.2022.103334}{When seeing is not believing: A mechanistic basis for predictive divergence}.
+\emph{Consciousness and Cognition} \textbf{102}, 103334 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lamme2020}{}}%
+\CSLLeftMargin{6. }%
+\CSLRightInline{Lamme, V. A. F. \href{https://www.frontiersin.org/articles/10.3389/fpsyg.2020.00083}{Visual functions generating conscious seeing}. \emph{Frontiers in Psychology} \textbf{11}, (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-notredame2014}{}}%
+\CSLLeftMargin{7. }%
+\CSLRightInline{Notredame, C.-E., Pins, D., Deneve, S.
+\& Jardri, R.
+\href{https://doi.org/10.3389/fnint.2014.00063}{What visual illusions teach us about schizophrenia}.
+\emph{Frontiers in Integrative Neuroscience} \textbf{8}, 63 (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gori2016}{}}%
+\CSLLeftMargin{8. }%
+\CSLRightInline{Gori, S., Molteni, M.
+\& Facoetti, A.
+\href{https://doi.org/10.3389/fnhum.2016.00175}{Visual illusions: An interesting tool to investigate developmental dyslexia and autism spectrum disorder}.
+\emph{Frontiers in Human Neuroscience} \textbf{10}, 175 (2016).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-razeghi2022}{}}%
+\CSLLeftMargin{9. }%
+\CSLRightInline{Razeghi, R., Arsham, S., Movahedi, A.
+\& Sammaknejad, N.
+\href{https://doi.org/10.1080/03004430.2020.1802260}{The effect of visual illusion on performance and quiet eye in autistic children}.
+\emph{Early Child Development and Care} \textbf{192}, 807--815 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-teufel2015}{}}%
+\CSLLeftMargin{10. }%
+\CSLRightInline{Teufel, C.
+\emph{et al.} \href{https://doi.org/10.1073/pnas.1503916112}{Shift toward prior knowledge confers a perceptual advantage in early psychosis and psychosis-prone healthy individuals}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{112}, 13401--13406 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hamburger2016}{}}%
+\CSLLeftMargin{11. }%
+\CSLRightInline{Hamburger, K.
+Visual Illusions Based on Processes: New Classification System Needed: \emph{Perception} (2016) doi:\href{https://doi.org/10.1177/0301006616629038}{10.1177/0301006616629038}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020illusions}{}}%
+\CSLLeftMargin{12. }%
+\CSLRightInline{Cretenoud, A. F., Francis, G. \& Herzog, M. H. When illusions merge. \emph{Journal of vision} \textbf{20}, 12--12 (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-halpern1999interindividual}{}}%
+\CSLLeftMargin{13. }%
+\CSLRightInline{Halpern, S. D., Andrews, T. J. \& Purves, D. Interindividual variation in human visual performance. \emph{Journal of cognitive neuroscience} \textbf{11}, 521--534 (1999).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-thurstone1944factorial}{}}%
+\CSLLeftMargin{14. }%
+\CSLRightInline{Thurstone, L. L. A factorial study of perception. (1944).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-grzeczkowski2017}{}}%
+\CSLLeftMargin{15. }%
+\CSLRightInline{Grzeczkowski, L., Clarke, A.
+M., Francis, G., Mast, F.
+W.
+\& Herzog, M.
+H.
+\href{https://doi.org/10.1016/j.visres.2016.10.006}{About individual differences in vision}.
+\emph{Vision Research} \textbf{141}, 282--292 (2017).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-grzeczkowski2018}{}}%
+\CSLLeftMargin{16. }%
+\CSLRightInline{Grzeczkowski, L.
+\emph{et al.} \href{https://doi.org/10.1016/j.psychres.2018.10.063}{Is the perception of illusions abnormal in schizophrenia?} \emph{Psychiatry Research} \textbf{270}, 929--939 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-park2022}{}}%
+\CSLLeftMargin{17. }%
+\CSLRightInline{Park, S., Zikopoulos, B.
+\& Yazdanbakhsh, A.
+\href{https://doi.org/10.1111/ejn.15739}{Visual illusion susceptibility in autism: A neural model}.
+\emph{European Journal of Neuroscience} \textbf{56}, (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-zhang2017}{}}%
+\CSLLeftMargin{18. }%
+\CSLRightInline{Zhang, Y.
+\emph{et al.} \href{https://doi.org/10.1515/tnsci-2017-0004}{Personality traits and perception of Müller-Lyer illusion in male Chinese military soldiers and university students}.
+\emph{Translational Neuroscience} \textbf{8}, 15--20 (2017).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-konrath2009seeing}{}}%
+\CSLLeftMargin{19. }%
+\CSLRightInline{Konrath, S., Bushman, B.
+J.
+\& Grove, T.
+Seeing my world in a million little pieces: Narcissism, self-construal, and cognitive--perceptual style.
+\emph{Journal of Personality} \textbf{77}, 1197--1228 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2019}{}}%
+\CSLLeftMargin{20. }%
+\CSLRightInline{Cretenoud, A.
+F.
+\emph{et al.} \href{https://doi.org/10.1167/19.14.12}{Factors underlying visual illusions are illusion-specific but not feature-specific}.
+\emph{Journal of Vision} \textbf{19}, 12 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-teufel2018}{}}%
+\CSLLeftMargin{21. }%
+\CSLRightInline{Teufel, C., Dakin, S.
+C.
+\& Fletcher, P.
+C.
+\href{https://doi.org/10.1038/s41598-018-28845-5}{Prior object-knowledge sharpens properties of early visual feature-detectors}.
+\emph{Scientific Reports} \textbf{8}, 10853 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-friston2010free}{}}%
+\CSLLeftMargin{22. }%
+\CSLRightInline{Friston, K. The free-energy principle: A unified brain theory? \emph{Nature reviews neuroscience} \textbf{11}, 127--138 (2010).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-adams2013computational}{}}%
-Adams, R. A., Stephan, K. E., Brown, H. R., Frith, C. D., \& Friston, K. J. (2013). The computational anatomy of psychosis. \emph{Frontiers in Psychiatry}, \emph{4}, 47.
+\CSLLeftMargin{23. }%
+\CSLRightInline{Adams, R. A., Stephan, K. E., Brown, H. R., Frith, C. D. \& Friston, K. J. The computational anatomy of psychosis. \emph{Frontiers in psychiatry} \textbf{4}, 47 (2013).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-balota2011moving}{}}%
-Balota, D.
-A., \& Yap, M.
-J.
-(2011).
-Moving beyond the mean in studies of mental chronometry: The power of response time distributional analyses.
-\emph{Current Directions in Psychological Science}, \emph{20}(3), 160--166.
+\leavevmode\vadjust pre{\hypertarget{ref-koethe2009binocular}{}}%
+\CSLLeftMargin{24. }%
+\CSLRightInline{Koethe, D.
+\emph{et al.} Binocular depth inversion as a paradigm of reduced visual information processing in prodromal state, antipsychotic-naive and treated schizophrenia.
+\emph{European Archives of Psychiatry and Clinical Neuroscience} \textbf{259}, 195--202 (2009).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bressan2021most}{}}%
-Bressan, P., \& Kramer, P. (2021). Most findings obtained with untimed visual illusions are confounded. \emph{Psychological Science}, \emph{32}(8), 1238--1246.
+\CSLLeftMargin{25. }%
+\CSLRightInline{Bressan, P. \& Kramer, P. Most findings obtained with untimed visual illusions are confounded. \emph{Psychological Science} \textbf{32}, 1238--1246 (2021).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
-Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel models using {Stan}. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
-
-\leavevmode\vadjust pre{\hypertarget{ref-burkner2020modelling}{}}%
-Bürkner, P.-C., \& Charpentier, E.
-(2020).
-Modelling monotonic effects of ordinal predictors in bayesian regression models.
-\emph{British Journal of Mathematical and Statistical Psychology}, \emph{73}(3), 420--451.
-
-\leavevmode\vadjust pre{\hypertarget{ref-caporuscio2022}{}}%
-Caporuscio, C., Fink, S.
-B., Sterzer, P., \& Martin, J.
-M.
-(2022).
-When seeing is not believing: A mechanistic basis for predictive divergence.
-\emph{Consciousness and Cognition}, \emph{102}, 103334.
-\url{https://doi.org/10.1016/j.concog.2022.103334}
-
-\leavevmode\vadjust pre{\hypertarget{ref-carbon2014}{}}%
-Carbon, C.-C. (2014). Understanding human perception by human-made illusions. \emph{Frontiers in Human Neuroscience}, \emph{8}. \url{https://www.frontiersin.org/articles/10.3389/fnhum.2014.00566}
+\leavevmode\vadjust pre{\hypertarget{ref-lanyi2022can}{}}%
+\CSLLeftMargin{26. }%
+\CSLRightInline{Lányi, O., Keri, S., Pálffy, Z. \& Polner, B. Can you believe your eyes? Positive schizotypy is associated with increased susceptibility to the m{ü}ller-lyer illusion. (2022).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-coren1976empirical}{}}%
-Coren, S., Girgus, J. S., Erlichman, H., \& Hakstian, A. R. (1976). An empirical taxonomy of visual illusions. \emph{Perception \& Psychophysics}, \emph{20}(2), 129--137.
+\CSLLeftMargin{27. }%
+\CSLRightInline{Coren, S., Girgus, J. S., Erlichman, H. \& Hakstian, A. R. An empirical taxonomy of visual illusions. \emph{Perception \& psychophysics} \textbf{20}, 129--137 (1976).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020illusions}{}}%
-Cretenoud, A. F., Francis, G., \& Herzog, M. H. (2020). When illusions merge. \emph{Journal of Vision}, \emph{20}(8), 12--12.
+\leavevmode\vadjust pre{\hypertarget{ref-mylniec2016}{}}%
+\CSLLeftMargin{28. }%
+\CSLRightInline{Mylniec, A.
+\& Bednarek, H.
+\href{https://doi.org/10.1515/ppb-2016-0012}{Field dependence, efficiency of information processing in working memory and susceptibility to orientation illusions among architects}.
+\emph{Polish Psychological Bulletin} \textbf{47}, 112--122 (2016).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2019}{}}%
-Cretenoud, A.
-F., Karimpur, H., Grzeczkowski, L., Francis, G., Hamburger, K., \& Herzog, M.
+\leavevmode\vadjust pre{\hypertarget{ref-skottun2014subjective}{}}%
+\CSLLeftMargin{29. }%
+\CSLRightInline{Skottun, B. C. \& Skoyles, J. R. Subjective criteria and illusions in visual testing: Some methodological limitations. \emph{Psychological research} \textbf{78}, 136--140 (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-makowski2021}{}}%
+\CSLLeftMargin{30. }%
+\CSLRightInline{Makowski, D., Lau, Z.
+J., Pham, T., Paul Boyce, W.
+\& Annabel Chen, S.
 H.
-(2019).
-Factors underlying visual illusions are illusion-specific but not feature-specific.
-\emph{Journal of Vision}, \emph{19}(14), 12.
-\url{https://doi.org/10.1167/19.14.12}
+\href{https://doi.org/10.1177/03010066211057347}{A Parametric Framework to Generate Visual Illusions Using Python}.
+\emph{Perception} \textbf{50}, 950--965 (2021).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-day1972}{}}%
-Day, R.
-H.
-(1972).
-Visual Spatial Illusions: A General Explanation: A wide range of visual illusions, including geometrical distortions, can be explained by a single principle.
-\emph{Science}, \emph{175}(4028), 1335--1340.
-\url{https://doi.org/10.1126/science.175.4028.1335}
-
-\leavevmode\vadjust pre{\hypertarget{ref-de2015jspsych}{}}%
-De Leeuw, J. R. (2015). jsPsych: A JavaScript library for creating behavioral experiments in a web browser. \emph{Behavior Research Methods}, \emph{47}(1), 1--12.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eagleman2001}{}}%
-Eagleman, D. M. (2001). Visual illusions and neurobiology. \emph{Nature Reviews Neuroscience}, \emph{2}(12), 920--926. \url{https://doi.org/10.1038/35104092}
-
-\leavevmode\vadjust pre{\hypertarget{ref-fox2006}{}}%
-Fox, A. (2006). \emph{Adolescent self-development and psychopathology: Anorexia nervosa and psychosis} {[}PhD thesis{]}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-friston2010free}{}}%
-Friston, K. (2010). The free-energy principle: A unified brain theory? \emph{Nature Reviews Neuroscience}, \emph{11}(2), 127--138.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gomez-villa2022}{}}%
-Gomez-Villa, A., Martín, A., Vazquez-Corral, J., Bertalmío, M., \& Malo, J.
-(2022).
-On the synthesis of visual illusions using deep generative models.
-\emph{Journal of Vision}, \emph{22}(8), 2.
-\url{https://doi.org/10.1167/jov.22.8.2}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gori2016}{}}%
-Gori, S., Molteni, M., \& Facoetti, A.
-(2016).
-Visual illusions: An interesting tool to investigate developmental dyslexia and autism spectrum disorder.
-\emph{Frontiers in Human Neuroscience}, \emph{10}, 175.
-\url{https://doi.org/10.3389/fnhum.2016.00175}
-
-\leavevmode\vadjust pre{\hypertarget{ref-grzeczkowski2017}{}}%
-Grzeczkowski, L., Clarke, A.
-M., Francis, G., Mast, F.
-W., \& Herzog, M.
-H.
-(2017).
-About individual differences in vision.
-\emph{Vision Research}, \emph{141}, 282--292.
-\url{https://doi.org/10.1016/j.visres.2016.10.006}
-
-\leavevmode\vadjust pre{\hypertarget{ref-grzeczkowski2018}{}}%
-Grzeczkowski, L., Roinishvili, M., Chkonia, E., Brand, A., Mast, F.
-W., Herzog, M.
-H., \& Shaqiri, A.
-(2018).
-Is the perception of illusions abnormal in schizophrenia?
-\emph{Psychiatry Research}, \emph{270}, 929--939.
-\url{https://doi.org/10.1016/j.psychres.2018.10.063}
-
-\leavevmode\vadjust pre{\hypertarget{ref-halpern1999interindividual}{}}%
-Halpern, S. D., Andrews, T. J., \& Purves, D. (1999). Interindividual variation in human visual performance. \emph{Journal of Cognitive Neuroscience}, \emph{11}(5), 521--534.
-
-\leavevmode\vadjust pre{\hypertarget{ref-hamburger2016}{}}%
-Hamburger, K. (2016). Visual Illusions Based on Processes: New Classification System Needed: \emph{Perception}. \url{https://doi.org/10.1177/0301006616629038}
+\leavevmode\vadjust pre{\hypertarget{ref-sibley2011}{}}%
+\CSLLeftMargin{31. }%
+\CSLRightInline{Sibley, C.
+\emph{et al.} The mini-IPIP6: Validation and extension of a short measure of the big-six factors of personality in new zealand.
+\emph{New Zealand Journal of Psychology} \textbf{40}, 142--159 (2011).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-hopwood2012}{}}%
-Hopwood, C.
+\CSLLeftMargin{32. }%
+\CSLRightInline{Hopwood, C.
 J., Thomas, K.
@@
 G.
-C., \& Krueger, R.
+C.
+\& Krueger, R.
 F.
-(2012).
-DSM-5 personality traits and DSM{\textendash}IV personality disorders.
-\emph{Journal of Abnormal Psychology}, \emph{121}(2), 424--432.
-\url{https://doi.org/10.1037/a0026656}
+\href{https://doi.org/10.1037/a0026656}{DSM-5 personality traits and DSM{\textendash}IV personality disorders.} \emph{Journal of Abnormal Psychology} \textbf{121}, 424--432 (2012).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-koethe2009binocular}{}}%
-Koethe, D., Kranaster, L., Hoyer, C., Gross, S., Neatby, M.
-A., Schultze-Lutter, F., Ruhrmann, S., Klosterkötter, J., Hellmich, M., \& Leweke, F.
-M.
-(2009).
-Binocular depth inversion as a paradigm of reduced visual information processing in prodromal state, antipsychotic-naive and treated schizophrenia.
-\emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{259}(4), 195--202.
+\leavevmode\vadjust pre{\hypertarget{ref-de2015jspsych}{}}%
+\CSLLeftMargin{33. }%
+\CSLRightInline{De Leeuw, J. R. jsPsych: A JavaScript library for creating behavioral experiments in a web browser. \emph{Behavior research methods} \textbf{47}, 1--12 (2015).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-konrath2009seeing}{}}%
-Konrath, S., Bushman, B.
-J., \& Grove, T.
-(2009).
-Seeing my world in a million little pieces: Narcissism, self-construal, and cognitive--perceptual style.
-\emph{Journal of Personality}, \emph{77}(4), 1197--1228.
+\leavevmode\vadjust pre{\hypertarget{ref-peer2022}{}}%
+\CSLLeftMargin{34. }%
+\CSLRightInline{Peer, E., Rothschild, D., Gordon, A., Evernden, Z.
+\& Damer, E.
+\href{https://doi.org/10.3758/s13428-021-01694-3}{Data quality of platforms and panels for online behavioral research}.
+\emph{Behavior Research Methods} \textbf{54}, 1643--1662 (2022).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-lamme2020}{}}%
-Lamme, V. A. F. (2020). Visual functions generating conscious seeing. \emph{Frontiers in Psychology}, \emph{11}. \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2020.00083}
+\leavevmode\vadjust pre{\hypertarget{ref-balota2011moving}{}}%
+\CSLLeftMargin{35. }%
+\CSLRightInline{Balota, D.
+A.
+\& Yap, M.
+J.
+Moving beyond the mean in studies of mental chronometry: The power of response time distributional analyses.
+\emph{Current Directions in Psychological Science} \textbf{20}, 160--166 (2011).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-lanyi2022can}{}}%
-Lányi, O., Keri, S., Pálffy, Z., \& Polner, B. (2022). \emph{Can you believe your eyes? Positive schizotypy is associated with increased susceptibility to the m{ü}ller-lyer illusion}.
+\leavevmode\vadjust pre{\hypertarget{ref-matzke2009psychological}{}}%
+\CSLLeftMargin{36. }%
+\CSLRightInline{Matzke, D.
+\& Wagenmakers, E.-J.
+Psychological interpretation of the ex-gaussian and shifted wald parameters: A diffusion model analysis.
+\emph{Psychonomic bulletin \& review} \textbf{16}, 798--817 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
+\CSLLeftMargin{37. }%
+\CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
+\CSLLeftMargin{38. }%
+\CSLRightInline{Bürkner, P.-C.
+\href{https://doi.org/10.18637/jss.v080.i01}{{brms}: An {R} package for {Bayesian} multilevel models using {Stan}}.
+\emph{Journal of Statistical Software} \textbf{80}, 1--28 (2017).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
+\CSLLeftMargin{39. }%
+\CSLRightInline{Wickham, H. \emph{et al.} \href{https://doi.org/10.21105/joss.01686}{Welcome to the tidyverse}. \emph{Journal of Open Source Software} \textbf{4}, 1686 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bayestestRArticle}{}}%
+\CSLLeftMargin{40. }%
+\CSLRightInline{Makowski, D., Ben-Shachar, M.
+\& Lüdecke, D.
+\href{https://doi.org/10.21105/joss.01541}{{bayestestR}: Describing effects and their uncertainty, existence and significance within the {Bayesian} framework}.
+\emph{JOSS} \textbf{4}, 1541 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-correlationArticle}{}}%
+\CSLLeftMargin{41. }%
+\CSLRightInline{Makowski, D., Ben-Shachar, M., Patil, I.
+\& Lüdecke, D.
+\href{https://doi.org/10.21105/joss.02306}{Methods and algorithms for correlation analysis in {R}}.
+\emph{JOSS} \textbf{5}, 2306 (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-performanceArticle}{}}%
+\CSLLeftMargin{42. }%
+\CSLRightInline{Lüdecke, D., Ben-Shachar, M., Patil, I., Waggoner, P.
+\& Makowski, D.
+\href{https://doi.org/10.21105/joss.03139}{{performance}: An {R} package for assessment, comparison and testing of statistical models}.
+\emph{JOSS} \textbf{6}, 3139 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-insightArticle}{}}%
+\CSLLeftMargin{43. }%
+\CSLRightInline{Lüdecke, D., Waggoner, P.
+\& Makowski, D.
+\href{https://doi.org/10.21105/joss.01412}{Insight: A unified interface to access information from model objects in {R}}.
+\emph{JOSS} \textbf{4}, 1412 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-makowski2019indices}{}}%
+\CSLLeftMargin{44. }%
+\CSLRightInline{Makowski, D., Ben-Shachar, M.
+S., Chen, S.
+A.
+\& Lüdecke, D.
+Indices of effect existence and significance in the bayesian framework.
+\emph{Frontiers in psychology} \textbf{10}, 2767 (2019).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-parametersArticle}{}}%
-Lüdecke, D., Ben-Shachar, M., Patil, I., \& Makowski, D.
-(2020).
-Extracting, computing and exploring the parameters of statistical models using {R}.
-\emph{Journal of Open Source Software}, \emph{5}(53), 2445.
-\url{https://doi.org/10.21105/joss.02445}
+\CSLLeftMargin{45. }%
+\CSLRightInline{Lüdecke, D., Ben-Shachar, M., Patil, I.
+\& Makowski, D.
+\href{https://doi.org/10.21105/joss.02445}{Extracting, computing and exploring the parameters of statistical models using {R}}.
+\emph{JOSS} \textbf{5}, 2445 (2020).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-performanceArticle}{}}%
-Lüdecke, D., Ben-Shachar, M., Patil, I., Waggoner, P., \& Makowski, D.
-(2021).
-{performance}: An {R} package for assessment, comparison and testing of statistical models.
-\emph{Journal of Open Source Software}, \emph{6}(60), 3139.
-\url{https://doi.org/10.21105/joss.03139}
+\leavevmode\vadjust pre{\hypertarget{ref-burkner2020modelling}{}}%
+\CSLLeftMargin{46. }%
+\CSLRightInline{Bürkner, P.-C.
+\& Charpentier, E.
+Modelling monotonic effects of ordinal predictors in bayesian regression models.
+\emph{British Journal of Mathematical and Statistical Psychology} \textbf{73}, 420--451 (2020).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-insightArticle}{}}%
-Lüdecke, D., Waggoner, P., \& Makowski, D.
-(2019).
-Insight: A unified interface to access information from model objects in {R}.
-\emph{Journal of Open Source Software}, \emph{4}(38), 1412.
-\url{https://doi.org/10.21105/joss.01412}
+\leavevmode\vadjust pre{\hypertarget{ref-yang2012}{}}%
+\CSLLeftMargin{47. }%
+\CSLRightInline{Yang, E.
+\emph{et al.} Visual Context Processing in Schizophrenia: \emph{Clinical Psychological Science} (2012) doi:\href{https://doi.org/10.1177/2167702612464618}{10.1177/2167702612464618}.}
 
-\leavevmode\vadjust pre{\hypertarget{ref-makowski2019indices}{}}%
-Makowski, D., Ben-Shachar, M. S., Chen, S. A., \& Lüdecke, D. (2019). Indices of effect existence and significance in the bayesian framework. \emph{Frontiers in Psychology}, \emph{10}, 2767.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bayestestRArticle}{}}%
-Makowski, D., Ben-Shachar, M., \& Lüdecke, D.
-(2019).
-{bayestestR}: Describing effects and their uncertainty, existence and significance within the {Bayesian} framework.
-\emph{Journal of Open Source Software}, \emph{4}(40), 1541.
-\url{https://doi.org/10.21105/joss.01541}
-
-\leavevmode\vadjust pre{\hypertarget{ref-correlationArticle}{}}%
-Makowski, D., Ben-Shachar, M., Patil, I., \& Lüdecke, D.
-(2020).
-Methods and algorithms for correlation analysis in {R}.
-\emph{Journal of Open Source Software}, \emph{5}(51), 2306.
-\url{https://doi.org/10.21105/joss.02306}
-
-\leavevmode\vadjust pre{\hypertarget{ref-makowski2021}{}}%
-Makowski, D., Lau, Z.
-J., Pham, T., Paul Boyce, W., \& Annabel Chen, S.
-H.
-(2021).
-A Parametric Framework to Generate Visual Illusions Using Python.
-\emph{Perception}, \emph{50}(11), 950--965.
-\url{https://doi.org/10.1177/03010066211057347}
-
-\leavevmode\vadjust pre{\hypertarget{ref-matzke2009psychological}{}}%
-Matzke, D., \& Wagenmakers, E.-J.
-(2009).
-Psychological interpretation of the ex-gaussian and shifted wald parameters: A diffusion model analysis.
-\emph{Psychonomic Bulletin \& Review}, \emph{16}(5), 798--817.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mylniec2016}{}}%
-Mylniec, A., \& Bednarek, H.
-(2016).
-Field dependence, efficiency of information processing in working memory and susceptibility to orientation illusions among architects.
-\emph{Polish Psychological Bulletin}, \emph{47}(1), 112--122.
-\url{https://doi.org/10.1515/ppb-2016-0012}
-
-\leavevmode\vadjust pre{\hypertarget{ref-notredame2014}{}}%
-Notredame, C.-E., Pins, D., Deneve, S., \& Jardri, R.
-(2014).
-What visual illusions teach us about schizophrenia.
-\emph{Frontiers in Integrative Neuroscience}, \emph{8}, 63.
-\url{https://doi.org/10.3389/fnint.2014.00063}
+\leavevmode\vadjust pre{\hypertarget{ref-fox2006}{}}%
+\CSLLeftMargin{48. }%
+\CSLRightInline{Fox, A. Adolescent self-development and psychopathology: Anorexia nervosa and psychosis. (2006).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-ohmann2016}{}}%
-Ohmann, K., \& Burgmer, P.
-(2016).
-Nothing compares to me: How narcissism shapes comparative thinking.
-\emph{Personality and Individual Differences}, \emph{98}, 162--170.
-\url{https://doi.org/10.1016/j.paid.2016.03.069}
-
-\leavevmode\vadjust pre{\hypertarget{ref-park2022}{}}%
-Park, S., Zikopoulos, B., \& Yazdanbakhsh, A.
-(2022).
-Visual illusion susceptibility in autism: A neural model.
-\emph{European Journal of Neuroscience}, \emph{56}.
-\url{https://doi.org/10.1111/ejn.15739}
-
-\leavevmode\vadjust pre{\hypertarget{ref-peer2022}{}}%
-Peer, E., Rothschild, D., Gordon, A., Evernden, Z., \& Damer, E.
-(2022).
-Data quality of platforms and panels for online behavioral research.
-\emph{Behavior Research Methods}, \emph{54}(4), 1643--1662.
-\url{https://doi.org/10.3758/s13428-021-01694-3}
+\CSLLeftMargin{49. }%
+\CSLRightInline{Ohmann, K.
+\& Burgmer, P.
+\href{https://doi.org/10.1016/j.paid.2016.03.069}{Nothing compares to me: How narcissism shapes comparative thinking}.
+\emph{Personality and Individual Differences} \textbf{98}, 162--170 (2016).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-pessoa2008}{}}%
-Pessoa, V.
+\CSLLeftMargin{50. }%
+\CSLRightInline{Pessoa, V.
 F., Monge-Fuentes, V., Simon, C.
-Y., Suganuma, E., \& Tavares, M.
+Y., Suganuma, E.
+\& Tavares, M.
 C.
 H.
-(2008).
-The müller-lyer illusion as a tool for schizophrenia screening.
-\emph{Reviews in the Neurosciences}, \emph{19}(2-3).
-\url{https://doi.org/10.1515/REVNEURO.2008.19.2-3.91}
-
-\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
-R Core Team. (2022). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-razeghi2022}{}}%
-Razeghi, R., Arsham, S., Movahedi, A., \& Sammaknejad, N.
-(2022).
-The effect of visual illusion on performance and quiet eye in autistic children.
-\emph{Early Child Development and Care}, \emph{192}(5), 807--815.
-\url{https://doi.org/10.1080/03004430.2020.1802260}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sibley2011}{}}%
-Sibley, C., Luyten, N., Wolfman, M., Mobberley, A., Wootton, L.
-W., Hammond, M., Sengupta, N., Perry, R., West-Newman, T., Wilson, M., McLellan, L., Hoverd, W.
-J., \& Robertson, A.
-(2011).
-The mini-IPIP6: Validation and extension of a short measure of the big-six factors of personality in new zealand.
-\emph{New Zealand Journal of Psychology}, \emph{40}, 142--159.
-
-\leavevmode\vadjust pre{\hypertarget{ref-skottun2014subjective}{}}%
-Skottun, B. C., \& Skoyles, J. R. (2014). Subjective criteria and illusions in visual testing: Some methodological limitations. \emph{Psychological Research}, \emph{78}(1), 136--140.
-
-\leavevmode\vadjust pre{\hypertarget{ref-teufel2018}{}}%
-Teufel, C., Dakin, S.
-C., \& Fletcher, P.
-C.
-(2018).
-Prior object-knowledge sharpens properties of early visual feature-detectors.
-\emph{Scientific Reports}, \emph{8}(1), 10853.
-\url{https://doi.org/10.1038/s41598-018-28845-5}
-
-\leavevmode\vadjust pre{\hypertarget{ref-teufel2015}{}}%
-Teufel, C., Subramaniam, N., Dobler, V., Perez, J., Finnemann, J., Mehta, P.
-R., Goodyer, I.
-M., \& Fletcher, P.
-C.
-(2015).
-Shift toward prior knowledge confers a perceptual advantage in early psychosis and psychosis-prone healthy individuals.
-\emph{Proceedings of the National Academy of Sciences}, \emph{112}(43), 13401--13406.
-\url{https://doi.org/10.1073/pnas.1503916112}
-
-\leavevmode\vadjust pre{\hypertarget{ref-thurstone1944factorial}{}}%
-Thurstone, L. L. (1944). \emph{A factorial study of perception.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T., Miller, E., Bache, S., Müller, K., Ooms, J., Robinson, D., Seidel, D., Spinu, V., \ldots{} Yutani, H.
-(2019).
-Welcome to the tidyverse.
-\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
-\url{https://doi.org/10.21105/joss.01686}
-
-\leavevmode\vadjust pre{\hypertarget{ref-yang2012}{}}%
-Yang, E., Tadin, D., Glasser, D.
-M., Hong, S.
-W., Blake, R., \& Park, S.
-(2012).
-Visual Context Processing in Schizophrenia: \emph{Clinical Psychological Science}.
-\url{https://doi.org/10.1177/2167702612464618}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zhang2017}{}}%
-Zhang, Y., Liu, J., Wang, Y., Huang, J., Wei, L., Zhang, B., Wang, W., \& Chen, W.
-(2017).
-Personality traits and perception of Müller-Lyer illusion in male Chinese military soldiers and university students.
-\emph{Translational Neuroscience}, \emph{8}(1), 15--20.
-\url{https://doi.org/10.1515/tnsci-2017-0004}
+\href{https://doi.org/10.1515/REVNEURO.2008.19.2-3.91}{The müller-lyer illusion as a tool for schizophrenia screening}.
+\emph{Reviews in the Neurosciences} \textbf{19}, (2008).}
 

### 2022-11-02 submission/manuscript.Rmd
edit abstract

@@
 abstract: |
-  Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty to experimentally manipulate these stimuli, leaving unanswered the hypothesis of a potential unique factor underlying the sensitivity to different types of illusions.
+  Visual illusions underscore how our perception of reality is constructed from an integration of sensory input with our prior beliefs.
+While the potential of there bieng a unique common factor that underlies our general sensitivity across illusions has been hypothesized and widely debated over, the field of illusion research has been hampered by the difficulty in experimentally manipulating these stimuli.
 In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.

### 2022-11-02 submission/manuscript.Rmd
tighten abstract

--- 2022-11-01 submission/manuscript.Rmd
@@
 abstract: |
-  Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty to experimentally manipulate these stimuli, leaving unanswered the hypothesis of a potential unique factor underlying the sensitivity to different types of illusions.
+  Visual illusions are a gateway to understand how we construct our experience of reality.
+Unfortunately, important questions remain open, such as the hypothesis of a common factor underlying the sensitivity to different types of illusions, as well as of personality correlates of illusion sensitivity.
 In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
-We tested the objective effect of the illusions on errors and reaction times, and extracted participant-level performance scores (n=250).
-Our results provide evidence in favour of a general factor (labelled Factor *i*) underlying the sensitivity to different illusions.
-Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
+We tested the objective effect of the illusions on errors and response times, and extracted participant-level performance scores (n=250) for each illusion.
+Our results provide evidence in favour of a general factor underlying the sensitivity to different illusions (labelled Factor *i*).
+Moreover, we report a positive link between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"

### 2022-12-06 submission/Initial/manuscript.tex
prepare revision file

--- 2022-11-01 submission/manuscript.tex
@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4070}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3826}
 \usepackage{lineno}
@@
 \abstract{%
-Visual illusions highlight how the brain uses contextual and prior information to inform our perception of reality.
-Unfortunately, illusion research has been hampered by the difficulty to experimentally manipulate these stimuli, leaving unanswered the hypothesis of a potential unique factor underlying the sensitivity to different types of illusions.
+Visual illusions are a gateway to understand how we construct our experience of reality.
+Unfortunately, important questions remain open, such as the hypothesis of a common factor underlying the sensitivity to different types of illusions, as well as of personality correlates of illusion sensitivity.
 In this study, we used a novel parametric framework for visual illusions to generate 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast) varying in strength, embedded in a perceptual discrimination task.
-We tested the objective effect of the illusions on errors and reaction times, and extracted participant-level performance scores (n=250).
-Our results provide evidence in favour of a general factor (labelled Factor \emph{i}) underlying the sensitivity to different illusions.
-Moreover, we report a positive relationship between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
+We tested the objective effect of the illusions on errors and response times, and extracted participant-level performance scores (n=250) for each illusion.
+Our results provide evidence in favour of a general factor underlying the sensitivity to different illusions (labelled Factor \emph{i}).
+Moreover, we report a positive link between illusion sensitivity and personality traits such as Agreeableness, Honesty-Humility, and negative relationships with Psychoticism, Antagonism, Disinhibition, and Negative Affect.
 }

### 2022-12-20 submission/R1/manuscript.Rmd
start reply

--- 2022-11-02 submission/manuscript.Rmd
@@
 
-The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) within each illusion.
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) **separately for each illusion**.
 We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions.

### 2022-12-26 submission/R1/manuscript.Rmd
updates to some comments

@@
 Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability [@cretenoud2020illusions; @cretenoud2019; @bressan2021most].
-Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], having them estimate the targets' physical properties [@coren1976empirical], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], having them estimate the targets' physical properties [@coren1976empirical], or most typically, through a method of adjustment, having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018; @cretenoud2020individual; @cretenoud2021visual].
 This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].
+While some recent studies have made efforts to implement paradigms that are more empirically rigorous [@cretenoud2021individual], varying the physical dimensions of the elements that make up the visual illusions, most of the manipulations made only alter the illusion magnitude without modulating the actual strength of the illusory effect.
 

### 2022-12-28 submission/R1/manuscript.Rmd
more replies

--- 2022-12-20 submission/R1/manuscript.Rmd
@@
 Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences [@teufel2018; @caporuscio2022] - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free].
-In this context, illusions are conceptualized as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
+In this context, illusions are conceptualized as anomalous beliefs that result from giving ample weight to prior knowledge to minimize prediction error in the face of ambiguous incoming sensory evidence.
 The predictive coding account further provides an explanation regarding the observations from clinical populations.
@@
 Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability [@cretenoud2020illusions; @cretenoud2019; @bressan2021most].
-Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], having them estimate the targets' physical properties [@coren1976empirical], or having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018].
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], having them estimate the targets' physical properties [@coren1976empirical], or through the method of adjustment, which involves having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018; @cretenoud2020individual; @cretenoud2021visual].
 This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].
+While some recent efforts have some made to implement more empirically rigorous paradigms [@cretenoud2021individual], most of the applied manipulations only focus on varying the physical dimensions of the illusion's target features without modulating its contextual elements, hence limiting the variability in the illusory effects of the stimuli presented.
+Furthermore, such prior studies have typically generated stimuli whose targets’ physical attributes vary over a relatively narrow range, thus further constraining the validity of their findings.
+As such, it is possible that the recent evidence reported against a common factor of illusions could be due to the low stimulus variance instead of a true reflection of a lack of common mechanism.
 
@@
 The aim of the present preregistered (https://osf.io/5d6xp) study is three-fold.
-First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
+First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Simultaneous Brighness Contrast).
 Next, we will investigate the factor structure of illusion-specific performance scores and test the existence of a common latent factor of illusion sensitivity.
@@
 
+We investigated the effect of 10 different classic illusions:
+
+<!-- Geometrical Illusions -->
+The Müller-Lyer illusion is represented by two equal-length line segments that end with arrow-like fins, either pointing inwards or outwards.
+The line with inward-pointing fins is typically perceived to be longer.
+
+The Ebbinghaus illusion is comprised of two circles of identical size placed near each other.
+While one circle is surrounded by smaller circles, the other is surrounded by larger circles.
+The latter circle usually appears smaller than the former.
+
+The Ponzo illusion is composed of two parallel horizontal lines embedded between a pair of converging lines.
+The upper line is perceived to be longer despite the two lines having the same objective length.
+
+The Poggendorff illusion is characterized by an oblique collinear line that appears as two misaligned line segments as a result of an intervening rectangular structure separating them.
+The upper line usually appears to be positioned higher than the point that would be collinear with the lower line.
+
+The Zollner illusion is represented by two parallel horizontal lines that are crossed with short, repeated lines.
+Notably, the direction in which the short repeated lines cross over the target parallel lines oppose one another, causing the target lines to appear not parallel.
+
+The Vertical-Horizontal illusion involves two equal-length line segments; one presented vertically and the other horizontally.
+The length of the vertical line is usually perceived as being longer, relative to that of the horizontal line.
+
+The Rod and Frame illusion consists of a vertical line segment ('rod') that is enclosed in a laterally tilted square ('frame').
+The orientation of the line segment typically looks congruent with the orientation of the frame.
+
+The Delboeuf illusion comprises two circles of identical size placed near each other; each surrounded by an outer ring.
+The central circle with the closer surrounding ring appears smaller than that with the distant surrounding ring in this visual illusion.
+
+<!-- Brightness illusions -->
+The Simultaneous Brightness Contrast illusion, henceforth referred to as the Contrast illusion, consists of two rectangles with identical shades of grey that are embedded in backgrounds of different color intensities.
+The rectangle situated on the dark background usually appears to be of a lighter shade than the rectangle on the light background.
+
+The White illusion is composed of a series of black and white horizontal bars that have been superimposed with rectangles that have an identical shade of grey.
+Rectangles that have been superimposed onto white bars appear to have a darker shade than those superimposed onto black bars.
+
 A pilot study (*n = 46*), of which a full description is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}, was first conducted to determine a sensitive range of stimuli parameters.
@@
 
-After a brief demographic survey and a practice series of illusions, the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
+Participants were first given a brief demographic survey, which collected information regarding their age, gender, country of birth, ethnicity and highest attained education level.
+This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
 Following this first series of blocks, two personality questionnaires were administered, the *IPIP6* [24 items, @sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the *PID-5* [25 items, @hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
@@
 For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
-The task was implemented using *jsPsych* [@de2015jspsych], and the set of instructions for each illusion type is available in the experiment code.
+All trials, which was estimated to take 50 minutes to complete, were required to be completed within a single-session.
+The task was implemented using *jsPsych* [@de2015jspsych] and was hosted on *Pavlovia* [(**https://pavlovia.org/**)](https://pavlovia.org/){.uri}.
+The set of instructions for each illusion type is available in the experiment code.
 
@@
 The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
-Participants were incentivised with a reward of about \textsterling 7.50 for completing the task, which took about 50 minutes to finish.
+Participants were incentivised with a reward of about \textsterling 7.50 for completing the task, which took approximately 50 minutes to finish.
 Demographic variables (age, gender, and ethnicity) were self-reported on a voluntary basis.
@@
 
-The inter-individual variability in the effect of illusion strength and its interaction with task difficulty was extracted from the models and used as participant-level scores.
+The inter-individual variability in the effect of illusion strength and its interaction with task difficulty (**diff**) was extracted from the models and used as participant-level scores.
 We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general factor of illusion sensitivity (Factor *i*).

### 2022-12-29 submission/R1/manuscript.Rmd
minor

@@
 While some recent efforts have some made to implement more empirically rigorous paradigms [@cretenoud2021individual], most of the applied manipulations only focus on varying the physical dimensions of the illusion's target features without modulating its contextual elements, hence limiting the variability in the illusory effects of the stimuli presented.
-Furthermore, such prior studies have typically generated stimuli whose targets’ physical attributes vary over a relatively narrow range, thus further constraining the validity of their findings.
+Furthermore, such prior studies have typically generated stimuli whose targets’ physical attributes vary over a relatively narrow range, thus further constraining the reliability of their findings.
 As such, it is possible that the recent evidence reported against a common factor of illusions could be due to the low stimulus variance instead of a true reflection of a lack of common mechanism.
@@
 For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
-All trials, which was estimated to take 50 minutes to complete, were required to be completed within a single-session.
+All trials were required to be completed within a single-session.
 The task was implemented using *jsPsych* [@de2015jspsych] and was hosted on *Pavlovia* [(**https://pavlovia.org/**)](https://pavlovia.org/){.uri}.

### 2022-12-30 study3/study3.Rmd
added labels to y-axis in fig 3

@@
   chunk_output_type: console
-bibliography: "../submission/references.bib"
+bibliography: "../submission/R1/references.bib"
 ---
@@
   scale_y_continuous(expand = c(0, 0)) +
-  labs(title = "Age", color = NULL) +
-  theme_modern() +
+  labs(title = expression("Age"[" (in years)"]), color = NULL, y = "Distribution") +
+  theme_modern(axis.title.space = 5) +
   theme(
@@
     plot.subtitle = element_text(face = "italic", hjust = 0.5),
-    axis.title.y = element_blank(),
+    axis.title.y = element_text(size = rel(0.8)),
     axis.text.y = element_blank(),
@@
   geom_bar(stat = "identity") +
-  labs(y = "Number", title = "Nationality") +
+  labs(y = "Number of participants", title = "Nationality") +
   scale_fill_material_d(guide = "none") +
@@
   scale_y_continuous(expand = c(0, 0)) +
-  theme_modern() +
+  theme_modern(axis.title.space = 5) +
   theme(
@@
     axis.title.x = element_blank(),
-    axis.title.y = element_blank(),
+    axis.title.y = element_text(size = rel(0.8)),
     axis.text.y = element_text(size = rel(0.8)),
@@
 
-# ggsave("figures/figure_dem1.png",
-#        patchwork::wrap_elements(p_age / p_nat),
-#        width=0.5*(fig.height),
-#        height=1*(fig.height), dpi=1000, bg="white")
+ggsave("figures/figure_dem1.png",
+       patchwork::wrap_elements(p_age / p_nat),
+       width=0.5*(fig.height),
+       height=1*(fig.height), dpi=1000, bg="white")
 
@@
 
-# ggsave("figures/figure_dem2.png",
-#        patchwork::wrap_elements(p_sex / p_edu / p_race), width=1*(fig.height), height=1*(fig.height), dpi=1000, bg="white")
+ggsave("figures/figure_dem2.png",
+       patchwork::wrap_elements(p_sex / p_edu / p_race), width=1*(fig.height), height=1*(fig.height), dpi=1000, bg="white")
 
@@
   scale_color_manual(values = c("Agreeableness" = "#FFC107", "Honesty-Humility" = "#00BCD4", "Extraversion" = "#9C27B0", "Conscientiousness" = "#3F51B5", "Openness" = "#4CAF50", "Neuroticism" = "#E91E63")) +
-  labs(x = "Score", title = "Normal Personality", color = NULL) +
-  theme_modern() +
+  labs(x = "Score", title = "Normal Personality", y = "Distribution", color = NULL) +
+  theme_modern(axis.title.space = 5) +
   theme(
     plot.title = element_text(size = rel(1), face = "bold", hjust = 0.5),
-    axis.title.y = element_blank(),
     axis.text.y = element_blank(),
+    axis.title.y = element_text(size = rel(0.8)),
     axis.text.x = element_text(size = rel(0.8)),
@@
   scale_color_manual(values = c("Antagonism" = "#FF9800", "Detachment" = "#03A9F4", "Disinhibition" = "#FF5722", "Negative Affect" = "#F44336", "Psychoticism" = "#673AB7")) +
-  labs(x = "Score", title = "Pathological Personality", color = NULL) +
-  theme_modern() +
+  labs(x = "Score", title = "Pathological Personality", y = "Distribution", color = NULL) +
+  theme_modern(axis.title.space = 5) +
   theme(
     plot.title = element_text(size = rel(1), face = "bold", hjust = 0.5),
-    axis.title.y = element_blank(),
     axis.text.y = element_blank(),
+    axis.title.y = element_text(size = rel(0.8)),
     axis.text.x = element_text(size = rel(0.8)),
@@
 
-# ggsave("figures/figure_dem3.png",
-#        patchwork::wrap_elements(p_ipip / p_pid), width=2/3*(fig.height), height=1*(fig.height), dpi=1000, bg="white")
+ggsave("figures/figure_dem3.png",
+       patchwork::wrap_elements(p_ipip / p_pid), width=2/3*(fig.height), height=1*(fig.height), dpi=1000, bg="white")
 
@@
 ```{r message=FALSE, warning=FALSE, cache=cache, eval=FALSE}
-p <- patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem.png"), interpolate = TRUE)) /
+p <- 
+  # patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_dem.png"), interpolate = TRUE)) /
   patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_agesex.png"), interpolate = TRUE)) /
   patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/figure_pers.png"), interpolate = TRUE)) +
-  plot_layout(heights = c(7200, 5000, 8000))
+  # plot_layout(heights = c(7200, 5000, 8000))
+  plot_layout(heights = c(5000, 8000))
 
-ggsave("../figures/figure3.png", p, height = 20200, width = 15166, units = "px", limitsize = FALSE)
+ggsave("../figures/figure4.png", p, height = 5000+8000, width = 15166, units = "px", limitsize = FALSE)
 ```

### 2022-12-30 study2/functions.R
add task in fig 3 and remove demographic from fig 4

@@ -482,7 +482,7 @@ plot_model_rt <- function(data, model, gam) {
 
   pred |>
     ggplot(aes_string(x = varstrength, y = "Predicted")) +
-    ggdist::stat_slab(data = data, aes_string(y = "RT", group = vardiff, fill = vardiff), alpha = 1/3) +
+    ggdist::stat_slab(data = data, aes_string(y = "RT", group = vardiff, fill = vardiff), alpha = 1 / 3) +
     geom_ribbon(data=gamdata, aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
     geom_vline(xintercept = 0, linetype = "dashed") +
     geom_line(data=gamdata, aes_string(color = vardiff, group = vardiff)) +
@@ -512,7 +512,7 @@ plot_model_rt <- function(data, model, gam) {
 
 
 
-plot_all <- function(data, p_err, p_rt) {
+plot_all <- function(data, p_err, p_rt, question="") {
   illname <- unique(data$Illusion_Type)
   # Get stimuli
   dat <- df |>
@@ -585,7 +585,9 @@ plot_all <- function(data, p_err, p_rt) {
   wrap_elements(((img_leftup | patchwork::plot_spacer() | img_rightup) / p / (img_leftdown | patchwork::plot_spacer() | img_rightdown) +
     patchwork::plot_layout(heights = c(0.5, 1.5, 0.5)) +
     patchwork::plot_annotation(title = paste(illname, "Illusion"),
-                               theme = theme(plot.title = element_text(size=rel(1.75), face="bold", hjust=0.5)))))
+                               subtitle = question,
+                               theme = theme(plot.title = element_text(size=rel(1.75), face="bold", hjust=0.5),
+                                             plot.subtitle = element_text(size=rel(1), face="italic", hjust=0.5, margin=margin(0,0,-30,0))))))
 }
 
 

### 2022-12-30 submission/R1/manuscript.Rmd
update

@@
 
-We investigated the effect of 10 different classic illusions:
+We investigated the effect of 10 different classic illusions [**inseert figure**]:
 
@@
 For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
-All trials were required to be completed within a single-session.
+All trials were required to be completed within a single-session (total experiment duration: ~55 minutes).
 The task was implemented using *jsPsych* [@de2015jspsych] and was hosted on *Pavlovia* [(**https://pavlovia.org/**)](https://pavlovia.org/){.uri}.
@@
 
-We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when implausibly fast).
-For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean).
+We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when implausibly fast relative to the typical RT distribution).
+For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean RT).
 Additionally, due to a technical issue, no personality data was recorded for the first eight participants.
@@
 Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age), and personality traits.
+It should be noted that the screen size is a proxy of the true screen size, measured using the number of pixels.
 

### 2022-12-30 study2/study2.Rmd

@@
   plot_model_err(filter(df, Illusion_Type == "Delboeuf"), model_delboeuf_err, gam_delboeuf_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Delboeuf", Error == 0), model_delboeuf_rt, gam_delboeuf_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Delboeuf", Error == 0), model_delboeuf_rt, gam_delboeuf_rt), 
+  question="Which red circle is bigger?")
 p_ebbinghaus <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Ebbinghaus"), model_ebbinghaus_err, gam_ebbinghaus_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Ebbinghaus", Error == 0), model_ebbinghaus_rt, gam_ebbinghaus_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Ebbinghaus", Error == 0), model_ebbinghaus_rt, gam_ebbinghaus_rt), 
+  question="Which red circle is bigger?")
 p_rodframe <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Rod-Frame"), model_rodframe_err, gam_rodframe_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Rod-Frame", Error == 0), model_rodframe_rt, gam_rodframe_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Rod-Frame", Error == 0), model_rodframe_rt, gam_rodframe_rt), 
+  question="Which side does the red line lean to?")
 p_verticalhorizontal <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Vertical-Horizontal"), model_verticalhorizontal_err, gam_verticalhorizontal_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0), model_verticalhorizontal_rt, gam_verticalhorizontal_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0), model_verticalhorizontal_rt, gam_verticalhorizontal_rt), 
+  question="Which red segment is longer?")
 p_zollner <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Zöllner"), model_zollner_err, gam_zollner_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Zöllner", Error == 0), model_zollner_rt, gam_zollner_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Zöllner", Error == 0), model_zollner_rt, gam_zollner_rt), 
+  question="Which side do the red lines close on?")
 p_white <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "White"), model_white_err, gam_white_err), 
-  plot_model_rt(filter(df, Illusion_Type == "White", Error == 0), model_white_rt, gam_white_rt) + coord_cartesian(ylim=c(0.125, 1.7), xlim = c(-17.5, 17.5)))
+  plot_model_rt(filter(df, Illusion_Type == "White", Error == 0), model_white_rt, gam_white_rt) + coord_cartesian(ylim=c(0.125, 1.7), xlim = c(-17.5, 17.5)), 
+  question="Which verticalcolumn\nis lighter?")
 p_mullerlyer <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Müller-Lyer"), model_mullerlyer_err, gam_mullerlyer_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Müller-Lyer", Error == 0), model_mullerlyer_rt, gam_mullerlyer_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Müller-Lyer", Error == 0), model_mullerlyer_rt, gam_mullerlyer_rt), 
+  question="Which red line is longer?")
 p_ponzo <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Ponzo"), model_ponzo_err, gam_ponzo_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Ponzo", Error == 0), model_ponzo_rt, gam_ponzo_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Ponzo", Error == 0), model_ponzo_rt, gam_ponzo_rt), 
+  question="Which red line\nis longer?")
 p_poggendorff <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Poggendorff"), model_poggendorff_err, gam_poggendorff_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Poggendorff", Error == 0), model_poggendorff_rt, gam_poggendorff_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Poggendorff", Error == 0), model_poggendorff_rt, gam_poggendorff_rt), 
+  question="Which side is the right\npart of the red segment\nmisplaced towards?")
 p_contrast <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Contrast"), model_contrast_err, gam_contrast_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Contrast", Error == 0), model_contrast_rt, gam_contrast_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Contrast", Error == 0), model_contrast_rt, gam_contrast_rt), 
+  question="Which small rectangle\nis lighter?")
 
@@
 
-ggsave("../figures/figure2.png", ((p_ill) / p_sem) + plot_layout(heights=c(2, 1.5)), height = 35, width = 30, dpi=300)
+ggsave("../figures/figure3.png", ((p_ill) / p_sem) + plot_layout(heights=c(2, 1.5)), height = 35, width = 30, dpi=300)
 ```

### 2022-12-30 submission/R1/manuscript.Rmd
minor edit

@@
 We fitted a Structural Equation Model (SEM) suggesting that these manifest variables group to first-level illusion-specific latent factors, which then load on a general factor of illusion sensitivity (Factor \\textit{i})."}
-knitr::include_graphics("../figures/Figure2.png")
+knitr::include_graphics("../figures/Figure3.png")
 ```
@@
 
-```{r message=FALSE, warning=FALSE, fig.cap="The upper plots show the distribution of demographic and dispositional variables.
-The middle plots show the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
-Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \\textit{i}) and personality traits.", out.width="100%"}
-knitr::include_graphics("../figures/Figure3.png")
+```{r message=FALSE, warning=FALSE, fig.cap="The upper plots show the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
+Bottom plots show the correlation between the general factor (Factor \\textit{i}) of illusion sensitivity (on the x-axes) and personality traits.", out.width="100%"}
+knitr::include_graphics("../figures/Figure4.png")
 ```

### 2022-12-31 study2/study2.Rmd
minor

--- 2022-09-20 study2/study2.Rmd
@@
   plot_model_err(filter(df, Illusion_Type == "Rod-Frame"), model_rodframe_err, gam_rodframe_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Rod-Frame", Error == 0), model_rodframe_rt, gam_rodframe_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Rod-Frame", Error == 0), model_rodframe_rt, gam_rodframe_rt), 
+  question="Which direction is the red line leaning towards?")
 p_verticalhorizontal <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Vertical-Horizontal"), model_verticalhorizontal_err, gam_verticalhorizontal_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0), model_verticalhorizontal_rt, gam_verticalhorizontal_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Vertical-Horizontal", Error == 0), model_verticalhorizontal_rt, gam_verticalhorizontal_rt), 
+  question="Which red line is longer?")
 p_zollner <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "Zöllner"), model_zollner_err, gam_zollner_err), 
-  plot_model_rt(filter(df, Illusion_Type == "Zöllner", Error == 0), model_zollner_rt, gam_zollner_rt))
+  plot_model_rt(filter(df, Illusion_Type == "Zöllner", Error == 0), model_zollner_rt, gam_zollner_rt), 
+  question="Which direction are the red\nlines converging towards?")
 p_white <- plot_all(
@@
   plot_model_err(filter(df, Illusion_Type == "White"), model_white_err, gam_white_err), 
-  plot_model_rt(filter(df, Illusion_Type == "White", Error == 0), model_white_rt, gam_white_rt) + coord_cartesian(ylim=c(0.125, 1.7), xlim = c(-17.5, 17.5)))
+  plot_model_rt(filter(df, Illusion_Type == "White", Error == 0), model_white_rt, gam_white_rt) + coord_cartesian(ylim=c(0.125, 1.7), xlim = c(-17.5, 17.5)), 
+  question="Which vertical rectangle\nis lighter?")
 p_mullerlyer <- plot_all(

### 2022-12-31 submission/R1/manuscript.Rmd
minor

@@
 
-We investigated the effect of 10 different classic illusions [**inseert figure**]:
+```{r message=FALSE, warning=FALSE, out.width="100%", fig.cap="Ten different illusions were used as stimuli in a perceptual task, where participants had to answer as fast as possibly, without making errors, according to specific instructions.
+For each illusion type, two parameters were experimentally manipulated, 1) the task difficulty (e.g., how large was the difference between the bigger and the smaller red circles in the Delboeuf illusion), and 2) the illusion strength (e.g., the size of the black circles in the Ebbinghaus illusion)."}
+knitr::include_graphics("../figures/Figure2.png")
+```
 
-<!-- Geometrical Illusions -->
-The Müller-Lyer illusion is represented by two equal-length line segments that end with arrow-like fins, either pointing inwards or outwards.
-The line with inward-pointing fins is typically perceived to be longer.
-
-The Ebbinghaus illusion is comprised of two circles of identical size placed near each other.
-While one circle is surrounded by smaller circles, the other is surrounded by larger circles.
-The latter circle usually appears smaller than the former.
-
-The Ponzo illusion is composed of two parallel horizontal lines embedded between a pair of converging lines.
-The upper line is perceived to be longer despite the two lines having the same objective length.
-
-The Poggendorff illusion is characterized by an oblique collinear line that appears as two misaligned line segments as a result of an intervening rectangular structure separating them.
-The upper line usually appears to be positioned higher than the point that would be collinear with the lower line.
-
-The Zollner illusion is represented by two parallel horizontal lines that are crossed with short, repeated lines.
-Notably, the direction in which the short repeated lines cross over the target parallel lines oppose one another, causing the target lines to appear not parallel.
-
-The Vertical-Horizontal illusion involves two equal-length line segments; one presented vertically and the other horizontally.
-The length of the vertical line is usually perceived as being longer, relative to that of the horizontal line.
-
-The Rod and Frame illusion consists of a vertical line segment ('rod') that is enclosed in a laterally tilted square ('frame').
-The orientation of the line segment typically looks congruent with the orientation of the frame.
-
-The Delboeuf illusion comprises two circles of identical size placed near each other; each surrounded by an outer ring.
-The central circle with the closer surrounding ring appears smaller than that with the distant surrounding ring in this visual illusion.
-
-<!-- Brightness illusions -->
-The Simultaneous Brightness Contrast illusion, henceforth referred to as the Contrast illusion, consists of two rectangles with identical shades of grey that are embedded in backgrounds of different color intensities.
-The rectangle situated on the dark background usually appears to be of a lighter shade than the rectangle on the light background.
-
-The White illusion is composed of a series of black and white horizontal bars that have been superimposed with rectangles that have an identical shade of grey.
-Rectangles that have been superimposed onto white bars appear to have a darker shade than those superimposed onto black bars.
-
+We investigated the effect of 10 different classic illusions (**Figure 2**).
 A pilot study (*n = 46*), of which a full description is available at [**https://github.com/RealityBending/IllusionGameValidation**](https://github.com/RealityBending/IllusionGameValidation){.uri}, was first conducted to determine a sensitive range of stimuli parameters.
@@
 Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear mixed models, which can be used to estimate the effects' participant-level variability (via random slopes).
-Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we fitted the best model (based on their indices of fit), and compared their output visually (**Figure 2**).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we fitted the best model (based on their indices of fit), and compared their output visually (**Figure 3**).
 
@@
 For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
-These consistently suggested a significant positive linear relationship between age and Factor *i* ($pd=100\%$), as well as the sensitivity to Müller-Lyer ($pd=100\%$), Vertical-Horizontal ($pd=100\%$), Zöllner ($pd=100\%$) and Ebbinghaus ($pd=99\%$) illusions (**Figure 3**).
+These consistently suggested a significant positive linear relationship between age and Factor *i* ($pd=100\%$), as well as the sensitivity to Müller-Lyer ($pd=100\%$), Vertical-Horizontal ($pd=100\%$), Zöllner ($pd=100\%$) and Ebbinghaus ($pd=99\%$) illusions (**Figure 4**).
 

### 2023-01-01 submission/R1/manuscript.Rmd
more

@@
 
-The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) **separately for each illusion**.
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) **separately for each illusion** under a Bayesian framework.
 We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions.
-Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].
+Errors were analyzed using logistic mixed models **(suited to estimate the error rate)**, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].
 
 Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear mixed models, which can be used to estimate the effects' participant-level variability (via random slopes).
-Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we fitted the best model (based on their indices of fit), and compared their output visually (**Figure 3**).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root; **which are types of relationship commonly found in perceptual tasks**) on the main predictors (task *difficulty* and illusion *strength*), we fitted the best model (**based on their BIC and R2**), and compared their output visually (**Figure 3**).
+**Note that the model comparison and the parameters used in the resulting models were not pre-registered.**
 
 The inter-individual variability in the effect of illusion strength and its interaction with task difficulty (**diff**) was extracted from the models and used as participant-level scores.
-We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general factor of illusion sensitivity (Factor *i*).
+We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA, **to gain insights into the structure**), and structural equation modelling (SEM, **to model and test different hierarchical models**), and tested the existence of a general factor of illusion sensitivity (Factor *i*).
 

### 2023-01-02 submission/R1/manuscript.Rmd
homogenize comments

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "3826"
+wordcount         : "4084"
 bibliography      : references.bib
@@
 Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences [@teufel2018; @caporuscio2022] - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free].
-In this context, illusions are conceptualized as anomalous beliefs that result from giving ample weight to prior knowledge to minimize prediction error in the face of ambiguous incoming sensory evidence.
-The predictive coding account further provides an explanation regarding the observations from clinical populations.
+In this context,illusions are conceptualized as **non-veridical perceptual experiences that result from giving ample weight to prior knowledge to minimize prediction error in the face of biasing sensory evidence.** The predictive coding account further provides an explanation regarding the observations from clinical populations.
 Certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components [@adams2013computational], resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions [@koethe2009binocular].
@@
 Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability [@cretenoud2020illusions; @cretenoud2019; @bressan2021most].
-Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], having them estimate the targets' physical properties [@coren1976empirical], or through the method of adjustment, which involves having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018; @cretenoud2020individual; @cretenoud2021visual].
-This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].
-While some recent efforts have some made to implement more empirically rigorous paradigms [@cretenoud2021individual], most of the applied manipulations only focus on varying the physical dimensions of the illusion's target features without modulating its contextual elements, hence limiting the variability in the illusory effects of the stimuli presented.
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different [@lanyi2022can], having them estimate the targets' physical properties [@coren1976empirical], **or through the method of adjustment, which involves having them adjust the targets to perceptually match a reference stimulus [@mylniec2016; @grzeczkowski2018; @cretenoud2020individual; @cretenoud2021visual].** This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity [@skottun2014subjective].
+**While some recent efforts have some made to implement more empirically rigorous paradigms [@cretenoud2021individual], most of the applied manipulations only focus on varying the physical dimensions of the illusion's target features without modulating its contextual elements, hence limiting the variability in the illusory effects of the stimuli presented.
 Furthermore, such prior studies have typically generated stimuli whose targets’ physical attributes vary over a relatively narrow range, thus further constraining the reliability of their findings.
-As such, it is possible that the recent evidence reported against a common factor of illusions could be due to the low stimulus variance instead of a true reflection of a lack of common mechanism.
+As such, it is possible that the recent evidence reported against a common factor of illusions could be due to the low stimulus variance instead of a true reflection of a lack of common mechanism.**
 
@@
 Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.", out.width="100%"}
-knitr::include_graphics("../figures/Figure1.png")
+knitr::include_graphics("../../figures/Figure1.png")
 ```
@@
 For each illusion type, two parameters were experimentally manipulated, 1) the task difficulty (e.g., how large was the difference between the bigger and the smaller red circles in the Delboeuf illusion), and 2) the illusion strength (e.g., the size of the black circles in the Ebbinghaus illusion)."}
-knitr::include_graphics("../figures/Figure2.png")
+knitr::include_graphics("../../figures/Figure2.png")
 ```
@@
 
-Participants were first given a brief demographic survey, which collected information regarding their age, gender, country of birth, ethnicity and highest attained education level.
+Participants were first given a brief demographic survey, which collected information regarding their **age, gender, country of birth, ethnicity and highest attained education level**.
 This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
@@
 For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
-All trials were required to be completed within a single-session (total experiment duration: ~55 minutes).
-The task was implemented using *jsPsych* [@de2015jspsych] and was hosted on *Pavlovia* [(**https://pavlovia.org/**)](https://pavlovia.org/){.uri}.
+**All trials were required to be completed within a single-session (total experiment duration: ~55 minutes)**.
+The task was implemented using *jsPsych* [@de2015jspsych] and was **hosted on *Pavlovia* ** [(**https://pavlovia.org/**)](https://pavlovia.org/){.uri}.
 The set of instructions for each illusion type is available in the experiment code.
@@
 
-We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when implausibly fast relative to the typical RT distribution).
-For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean RT).
+We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when implausibly fast relative to the average RT distribution).
+For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (< 125 ms or > 4 SD above mean RT).
 Additionally, due to a technical issue, no personality data was recorded for the first eight participants.
@@
 Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age), and personality traits.
-It should be noted that the screen size is a proxy of the true screen size, measured using the number of pixels.
+It should be noted that the measure of screen size used (measured using the number of pixels) is **only a proxy of the true physical screen size**.
 
@@
 We fitted a Structural Equation Model (SEM) suggesting that these manifest variables group to first-level illusion-specific latent factors, which then load on a general factor of illusion sensitivity (Factor \\textit{i})."}
-knitr::include_graphics("../figures/Figure3.png")
+knitr::include_graphics("../../figures/Figure3.png")
 ```
@@
 Bottom plots show the correlation between the general factor (Factor \\textit{i}) of illusion sensitivity (on the x-axes) and personality traits.", out.width="100%"}
-knitr::include_graphics("../figures/Figure4.png")
+knitr::include_graphics("../../figures/Figure4.png")
 ```

### 2023-01-02 submission/R1/manuscript.tex
homogenize comments

--- 2022-12-06 submission/Initial/manuscript.tex
@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 3826}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4084}
 \usepackage{lineno}
@@
 Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system\textsuperscript{8,20} or to top-down influences\textsuperscript{5,21} - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis\textsuperscript{22}.
-In this context, illusions are conceptualized as ambiguous percepts (noisy sensory evidence) giving ample weight to prior knowledge to minimize prediction error and provide a coherent perceptual experience.
-The predictive coding account further provides an explanation regarding the observations from clinical populations.
+In this context,illusions are conceptualized as \textbf{non-veridical perceptual experiences that result from giving ample weight to prior knowledge to minimize prediction error in the face of biasing sensory evidence.} The predictive coding account further provides an explanation regarding the observations from clinical populations.
 Certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components\textsuperscript{23}, resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions\textsuperscript{24}.
@@
 Instead, existing studies typically use only one or a small subset of illusion types, with few contrasting conditions, restricting the findings' generalizability\textsuperscript{12,20,25}.
-Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different\textsuperscript{26}, having them estimate the targets' physical properties\textsuperscript{27}, or having them adjust the targets to perceptually match a reference stimulus\textsuperscript{16,28}.
-This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity\textsuperscript{29}.
+Moreover, conventional paradigms often focus on the participants' subjective experience, by asking them the extent to which they perceive two identical targets as different\textsuperscript{26}, having them estimate the targets' physical properties\textsuperscript{27}, \textbf{or through the method of adjustment, which involves having them adjust the targets to perceptually match a reference stimulus\textsuperscript{16,28--30}.} This reliance on meta-cognitive judgements about one's subjective experience likely distorts the measurand, limiting the ability to reliably obtain more direct and objective measures of illusion sensitivity\textsuperscript{31}.
+\textbf{While some recent efforts have some made to implement more empirically rigorous paradigms\textsuperscript{32}, most of the applied manipulations only focus on varying the physical dimensions of the illusion's target features without modulating its contextual elements, hence limiting the variability in the illusory effects of the stimuli presented.
+Furthermore, such prior studies have typically generated stimuli whose targets' physical attributes vary over a relatively narrow range, thus further constraining the reliability of their findings.
+As such, it is possible that the recent evidence reported against a common factor of illusions could be due to the low stimulus variance instead of a true reflection of a lack of common mechanism.}
 
-To address these issues, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software \emph{Pyllusion}\textsuperscript{30}.
+To address these issues, we first developed a parametric framework to manipulate visual illusions that we implemented and made accessible in the open-source software \emph{Pyllusion}\textsuperscript{33}.
 This software allows us to generate different types of classic visual illusions with a continuous and independent modulation of two parameters: \emph{illusion strength} and \emph{task difficulty} (\textbf{Figure 1}).
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/Figure1} \caption{The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+\includegraphics[width=1\linewidth]{../../figures/Figure1} \caption{The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
 Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.}\label{fig:unnamed-chunk-2}
@@
 The aim of the present preregistered (\url{https://osf.io/5d6xp}) study is three-fold.
-First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Contrast).
+First, we will test this novel paradigm by investigating if the effect of illusion strength and task difficulty can be manipulated continuously for 10 different classic illusions (Delboeuf, Ebbinghaus, Rod and Frame, Vertical-Horizontal, Zöllner, White, Müller-Lyer, Ponzo, Poggendorff, Simultaneous Brighness Contrast).
 Next, we will investigate the factor structure of illusion-specific performance scores and test the existence of a common latent factor of illusion sensitivity.
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{../../figures/Figure2} \caption{Ten different illusions were used as stimuli in a perceptual task, where participants had to answer as fast as possibly, without making errors, according to specific instructions.
+For each illusion type, two parameters were experimentally manipulated, 1) the task difficulty (e.g., how large was the difference between the bigger and the smaller red circles in the Delboeuf illusion), and 2) the illusion strength (e.g., the size of the black circles in the Ebbinghaus illusion).}\label{fig:unnamed-chunk-3}
+\end{figure}
+
+We investigated the effect of 10 different classic illusions (\textbf{Figure 2}).
 A pilot study (\emph{n = 46}), of which a full description is available at \href{https://github.com/RealityBending/IllusionGameValidation}{\textbf{https://github.com/RealityBending/IllusionGameValidation}}, was first conducted to determine a sensitive range of stimuli parameters.
@@
 
-After a brief demographic survey and a practice series of illusions, the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
-Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6}\textsuperscript{24 items, 31} - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5}\textsuperscript{25 items, 32} - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+Participants were first given a brief demographic survey, which collected information regarding their \textbf{age, gender, country of birth, ethnicity and highest attained education level}.
+This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
+Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6}\textsuperscript{24 items, 34} - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5}\textsuperscript{25 items, 35} - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
 Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
@@
 For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest.
-The task was implemented using \emph{jsPsych}\textsuperscript{33}, and the set of instructions for each illusion type is available in the experiment code.
+\textbf{All trials were required to be completed within a single-session (total experiment duration: \textasciitilde55 minutes)}.
+The task was implemented using \emph{jsPsych}\textsuperscript{36} and was \textbf{hosted on \emph{Pavlovia} } \href{https://pavlovia.org/}{(\textbf{https://pavlovia.org/})}.
+The set of instructions for each illusion type is available in the experiment code.
 
@@
 
-Participants were recruited via \emph{Prolific}, a crowd-sourcing platform recognized for providing high quality data\textsuperscript{34}.
+Participants were recruited via \emph{Prolific}, a crowd-sourcing platform recognized for providing high quality data\textsuperscript{37}.
 The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
-Participants were incentivised with a reward of about \textsterling 7.50 for completing the task, which took about 50 minutes to finish.
+Participants were incentivised with a reward of about \textsterling 7.50 for completing the task, which took approximately 50 minutes to finish.
 Demographic variables (age, gender, and ethnicity) were self-reported on a voluntary basis.
 
-We excluded 6 participants upon inspection of the average error rate (when close to 50\%, suggesting random answers), and reaction time distribution (when implausibly fast).
-For the remaining participants, we discarded blocks with more than 50\% of errors (2.16\% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76\% trials with extreme response times (\textless{} 125 ms or \textgreater{} 4 SD above mean).
+We excluded 6 participants upon inspection of the average error rate (when close to 50\%, suggesting random answers), and reaction time distribution (when implausibly fast relative to the average RT distribution).
+For the remaining participants, we discarded blocks with more than 50\% of errors (2.16\% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76\% trials with extreme response times (\textless{} 125 ms or \textgreater{} 4 SD above mean RT).
 Additionally, due to a technical issue, no personality data was recorded for the first eight participants.
@@
 
-The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) within each illusion.
+The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) \textbf{separately for each illusion} under a Bayesian framework.
 We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions.
-Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location \(\mu\) (mean), scale \(\sigma\) (spread) and tail-dominance \(\tau\) of the RT distribution\textsuperscript{35,36}.
+Errors were analyzed using logistic mixed models \textbf{(suited to estimate the error rate)}, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location \(\mu\) (mean), scale \(\sigma\) (spread) and tail-dominance \(\tau\) of the RT distribution\textsuperscript{38,39}.
 
 Using GAMs as the ``ground-truth'' models, we attempted at approximating them using general linear mixed models, which can be used to estimate the effects' participant-level variability (via random slopes).
-Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task \emph{difficulty} and illusion \emph{strength}), we fitted the best model (based on their indices of fit), and compared their output visually (\textbf{Figure 2}).
+Following a comparison of models with a combination of transformations (raw, log, square root or cubic root; \textbf{which are types of relationship commonly found in perceptual tasks}) on the main predictors (task \emph{difficulty} and illusion \emph{strength}), we fitted the best model (\textbf{based on their BIC and R2}), and compared their output visually (\textbf{Figure 3}).
+\textbf{Note that the model comparison and the parameters used in the resulting models were not pre-registered.}
 
-The inter-individual variability in the effect of illusion strength and its interaction with task difficulty was extracted from the models and used as participant-level scores.
-We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general factor of illusion sensitivity (Factor \emph{i}).
+The inter-individual variability in the effect of illusion strength and its interaction with task difficulty (\textbf{diff}) was extracted from the models and used as participant-level scores.
+We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA, \textbf{to gain insights into the structure}), and structural equation modelling (SEM, \textbf{to model and test different hierarchical models}), and tested the existence of a general factor of illusion sensitivity (Factor \emph{i}).
 
 Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor \emph{i}), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age), and personality traits.
+It should be noted that the measure of screen size used (measured using the number of pixels) is \textbf{only a proxy of the true physical screen size}.
 
-The analysis was carried out using \emph{R 4.2}\textsuperscript{37}, \emph{brms}\textsuperscript{38}, the \emph{tidyverse}\textsuperscript{39}, and the \emph{easystats} collection of packages\textsuperscript{40--43}.
-As all the full results have been made available (see \textbf{Data Availability}), we will focus here on the significant results\textsuperscript{based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see 44}.
+The analysis was carried out using \emph{R 4.2}\textsuperscript{40}, \emph{brms}\textsuperscript{41}, the \emph{tidyverse}\textsuperscript{42}, and the \emph{easystats} collection of packages\textsuperscript{43--46}.
+As all the full results have been made available (see \textbf{Data Availability}), we will focus here on the significant results\textsuperscript{based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see 47}.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/Figure2} \caption{Top: the effect of illusion strength and task difficulty on the error rate and reaction time (RT) for each individual illusion.
+\includegraphics[width=1\linewidth]{../../figures/Figure3} \caption{Top: the effect of illusion strength and task difficulty on the error rate and reaction time (RT) for each individual illusion.
 The solid line represents the General Additive Model (GAM), and the dashed line corresponds to its approximation via linear models.
@@
 Bottom: We extracted the effect slope of the illusion strength and its interaction with task difficulty for each participant.
-We fitted a Structural Equation Model (SEM) suggesting that these manifest variables group to first-level illusion-specific latent factors, which then load on a general factor of illusion sensitivity (Factor \textit{i}).}\label{fig:unnamed-chunk-3}
+We fitted a Structural Equation Model (SEM) suggesting that these manifest variables group to first-level illusion-specific latent factors, which then load on a general factor of illusion sensitivity (Factor \textit{i}).}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 These twenty participant-level scores were subjected to exploratory factor analysis (EFA).
-The Method Agreement Procedure\textsuperscript{45} suggested the presence of 7 latent factors.
+The Method Agreement Procedure\textsuperscript{48} suggested the presence of 7 latent factors.
 An oblique (\emph{oblimin} rotation) factor solution explaining 66.69\% of variance suggested separate dimensions for the effect of Zöllner, White, Poggendorff, Contrast, Ebbinghaus, Delboeuf, and a common factor for the parameters related to Müller-Lyer, Vertical-Horizontal, Ponzo and Rod and Frame.
@@
 The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females (\(\Delta=-0.37\), 95\% CI {[}-0.62, -0.13{]}, \(BF_{10} = 12.74\)).
-We fitted Bayesian linear models with the education level entered as a monotonic predictor\textsuperscript{appropriate for ordinal variables, 46}, which yielded no significant effects.
+We fitted Bayesian linear models with the education level entered as a monotonic predictor\textsuperscript{appropriate for ordinal variables, 49}, which yielded no significant effects.
 For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.
-These consistently suggested a significant positive linear relationship between age and Factor \emph{i} (\(pd=100\%\)), as well as the sensitivity to Müller-Lyer (\(pd=100\%\)), Vertical-Horizontal (\(pd=100\%\)), Zöllner (\(pd=100\%\)) and Ebbinghaus (\(pd=99\%\)) illusions (\textbf{Figure 3}).
+These consistently suggested a significant positive linear relationship between age and Factor \emph{i} (\(pd=100\%\)), as well as the sensitivity to Müller-Lyer (\(pd=100\%\)), Vertical-Horizontal (\(pd=100\%\)), Zöllner (\(pd=100\%\)) and Ebbinghaus (\(pd=99\%\)) illusions (\textbf{Figure 4}).
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/Figure3} \caption{The upper plots show the distribution of demographic and dispositional variables.
-The middle plots show the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
-Bottom plots show the correlation between the general factor of illusion sensitivity (Factor \textit{i}) and personality traits.}\label{fig:unnamed-chunk-4}
+\includegraphics[width=1\linewidth]{../../figures/Figure4} \caption{The upper plots show the illusion sensitivity scores as a function of sex and age (solid lines indicate significant relationships).
+Bottom plots show the correlation between the general factor (Factor \textit{i}) of illusion sensitivity (on the x-axes) and personality traits.}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-This study tested a novel illusion sensitivity task paradigm based on the parametric illusion generation framework\textsuperscript{30}.
+This study tested a novel illusion sensitivity task paradigm based on the parametric illusion generation framework\textsuperscript{33}.
 Using the carefully generated stimuli in a perceptual decision task, we have shown that a gradual modulation of illusion strength is effectively possible across 10 different types of classic visual illusions.
@@
 This result comes in a field of mixed findings.
-In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals\textsuperscript{12,15,16,20,47}.
+In fact, contrary to early studies on visual illusions, more recent research have generally not found any significant evidence for a common stable factor across illusions within individuals\textsuperscript{12,15,16,20,50}.
 Instead, past findings suggest illusory effects are highly specific to the perceptual features of the illusions at stake\textsuperscript{15,20}.
@@
 Finally, we found illusion sensitivity to be positively associated with ``positive'' personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
-Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs\textsuperscript{often associated with psychoticism, 48} to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli\textsuperscript{19,48,49}.
+Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs\textsuperscript{often associated with psychoticism, 51} to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli\textsuperscript{19,51,52}.
 As such, the association between maladaptive traits and lower illusion sensitivity could be linked to a self-centered, decontextualized and disorganized information processing style.
-Conversely, the relationship between illusion sensitivity and adaptive personality traits is in line with the decreased field dependence (the tendency to rely on external cues in ambiguous contexts) associated with traits negatively correlated with agreeableness and honesty-humility, such as hostility, aggression and narcissism\textsuperscript{18,19,50}.
+Conversely, the relationship between illusion sensitivity and adaptive personality traits is in line with the decreased field dependence (the tendency to rely on external cues in ambiguous contexts) associated with traits negatively correlated with agreeableness and honesty-humility, such as hostility, aggression and narcissism\textsuperscript{18,19,53}.
 
 Importantly, these findings highlight the relevance of illusions beyond the field of visual perception, pointing towards an association with high-level domain-general mechanisms.
-In particular, the evidence in favor of a relationship between maladaptive personality traits and illusion sensitivity is in line with clinical observations, in which a greater resistance to illusions have been reported among patients with schizophrenia\textsuperscript{7,16,50}, especially in association with schizotypal traits such as cognitive disorganization\textsuperscript{20,26}.
+In particular, the evidence in favor of a relationship between maladaptive personality traits and illusion sensitivity is in line with clinical observations, in which a greater resistance to illusions have been reported among patients with schizophrenia\textsuperscript{7,16,53}, especially in association with schizotypal traits such as cognitive disorganization\textsuperscript{20,26}.
 While the search for the exact mechanism(s) underlying these links is an important goal of future research, our findings unlock the potential of illusion-based tasks as sensitive tools to capture specific inter-individual neuro-cognitive differences.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2020individual}{}}%
+\CSLLeftMargin{29. }%
+\CSLRightInline{Cretenoud, A.
+F., Grzeczkowski, L., Bertamini, M.
+\& Herzog, M.
+H.
+Individual differences in the m{ü}ller-lyer and ponzo illusions are stable across different contexts.
+\emph{Journal of Vision} \textbf{20}, 4--4 (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2021visual}{}}%
+\CSLLeftMargin{30. }%
+\CSLRightInline{Cretenoud, A. F. \emph{et al.} How do visual skills relate to action video game performance? \emph{Journal of vision} \textbf{21}, 10--10 (2021).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-skottun2014subjective}{}}%
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{Skottun, B. C. \& Skoyles, J. R. Subjective criteria and illusions in visual testing: Some methodological limitations. \emph{Psychological research} \textbf{78}, 136--140 (2014).}
 
+\leavevmode\vadjust pre{\hypertarget{ref-cretenoud2021individual}{}}%
+\CSLLeftMargin{32. }%
+\CSLRightInline{Cretenoud, A.
+F., Grzeczkowski, L., Kunchulia, M.
+\& Herzog, M.
+H.
+Individual differences in the perception of visual illusions are stable across eyes, time, and measurement methods.
+\emph{Journal of vision} \textbf{21}, 26--26 (2021).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-makowski2021}{}}%
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{Makowski, D., Lau, Z.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sibley2011}{}}%
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{Sibley, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hopwood2012}{}}%
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Hopwood, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-de2015jspsych}{}}%
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{De Leeuw, J. R. jsPsych: A JavaScript library for creating behavioral experiments in a web browser. \emph{Behavior research methods} \textbf{47}, 1--12 (2015).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-peer2022}{}}%
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Peer, E., Rothschild, D., Gordon, A., Evernden, Z.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-balota2011moving}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Balota, D.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-matzke2009psychological}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Matzke, D.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Bürkner, P.-C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Wickham, H. \emph{et al.} \href{https://doi.org/10.21105/joss.01686}{Welcome to the tidyverse}. \emph{Journal of Open Source Software} \textbf{4}, 1686 (2019).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bayestestRArticle}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Makowski, D., Ben-Shachar, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-correlationArticle}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Makowski, D., Ben-Shachar, M., Patil, I.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-performanceArticle}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Lüdecke, D., Ben-Shachar, M., Patil, I., Waggoner, P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-insightArticle}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Lüdecke, D., Waggoner, P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-makowski2019indices}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Makowski, D., Ben-Shachar, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-parametersArticle}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Lüdecke, D., Ben-Shachar, M., Patil, I.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-burkner2020modelling}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Bürkner, P.-C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-yang2012}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Yang, E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-fox2006}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Fox, A. Adolescent self-development and psychopathology: Anorexia nervosa and psychosis. (2006).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ohmann2016}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Ohmann, K.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pessoa2008}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Pessoa, V.

### 2023-01-02 submission/R1/manuscript.Rmd
accept all and ready to submit

@@
 keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "4084"
+wordcount         : "4085"
 bibliography      : references.bib
@@
 Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system [@cretenoud2019; @gori2016] or to top-down influences [@teufel2018; @caporuscio2022] - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis [@friston2010free].
-In this context,illusions are conceptualized as **non-veridical perceptual experiences that result from giving ample weight to prior knowledge to minimize prediction error in the face of biasing sensory evidence.** The predictive coding account further provides an explanation regarding the observations from clinical populations.
+In this context, illusions are conceptualized as **non-veridical perceptual experiences that result from giving ample weight to prior knowledge to minimize prediction error in the face of biasing sensory evidence.** The predictive coding account further provides an explanation regarding the observations from clinical populations.
 Certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components [@adams2013computational], resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions [@koethe2009binocular].

### 2023-01-02 submission/R1/manuscript.tex
accept all and ready to submit

@@
 Although the nature of the processes underlying illusion perception - whether related to low-level features of the visual processing system\textsuperscript{8,20} or to top-down influences\textsuperscript{5,21} - remains debated, a growing body of literature proposes to conceptualize illusions under the Bayesian brain hypothesis\textsuperscript{22}.
-In this context,illusions are conceptualized as \textbf{non-veridical perceptual experiences that result from giving ample weight to prior knowledge to minimize prediction error in the face of biasing sensory evidence.} The predictive coding account further provides an explanation regarding the observations from clinical populations.
+In this context, illusions are conceptualized as \textbf{non-veridical perceptual experiences that result from giving ample weight to prior knowledge to minimize prediction error in the face of biasing sensory evidence.} The predictive coding account further provides an explanation regarding the observations from clinical populations.
 Certain dispositional traits or characteristics (e.g., psychoticism) are seen as driven by alterations in the system's metacognitive components\textsuperscript{23}, resulting in an underweighting of priors during perceptual inferences, and manifesting as a decreased sensitivity to illusions\textsuperscript{24}.

### 2023-01-03 submission/R1/manuscript.Rmd
minor formatting

@@
 This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
-Following this first series of blocks, two personality questionnaires were administered, the *IPIP6* [24 items, @sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the *PID-5* [25 items, @hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+Following this first series of blocks, two personality questionnaires were administered, the *IPIP6* [@sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the *PID-5* [@hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
 Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
@@
 The analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle].
-As all the full results have been made available (see **Data Availability**), we will focus here on the significant results [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].
+As all the full results have been made available (see **Data Availability**), we will focus here on the significant results
+<!-- [based on the Bayes Factor *BF* or the Probability of Direction *pd*, see @makowski2019indices].  -->
+(based on the Bayes Factor *BF* or the Probability of Direction *pd* [@makowski2019indices]).
 
@@
 The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females ($\Delta=-0.37$, 95% CI [-0.62, -0.13], $BF_{10} = 12.74$).
-We fitted Bayesian linear models with the education level entered as a monotonic predictor [appropriate for ordinal variables, @burkner2020modelling], which yielded no significant effects.
+We fitted Bayesian linear models with the education level entered as a monotonic predictor
+(appropriate for ordinal variables [@burkner2020modelling]])
+<!-- [appropriate for ordinal variables, @burkner2020modelling], -->
+which yielded no significant effects.
 For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.

### 2023-01-03 submission/R1/manuscript.tex
minor formatting 2

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4084}
+\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4085}
 \usepackage{lineno}
@@
 This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block.
-Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6}\textsuperscript{24 items, 34} - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5}\textsuperscript{25 items, 35} - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
+Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6}\textsuperscript{34} - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5}\textsuperscript{35} - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism).
 Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials).
@@
 The analysis was carried out using \emph{R 4.2}\textsuperscript{40}, \emph{brms}\textsuperscript{41}, the \emph{tidyverse}\textsuperscript{42}, and the \emph{easystats} collection of packages\textsuperscript{43--46}.
-As all the full results have been made available (see \textbf{Data Availability}), we will focus here on the significant results\textsuperscript{based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see 47}.
+As all the full results have been made available (see \textbf{Data Availability}), we will focus here on the significant results
+(based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}\textsuperscript{47}).
 
@@
 The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females (\(\Delta=-0.37\), 95\% CI {[}-0.62, -0.13{]}, \(BF_{10} = 12.74\)).
-We fitted Bayesian linear models with the education level entered as a monotonic predictor\textsuperscript{appropriate for ordinal variables, 49}, which yielded no significant effects.
+We fitted Bayesian linear models with the education level entered as a monotonic predictor
+(appropriate for ordinal variables\textsuperscript{49}{]})
+which yielded no significant effects.
 For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model.

### 2023-04-10 submission/R1/manuscript.Rmd
final  comments

@@
 ---
-title             : "**The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates**"
+title             : "**A Novel Visual Illusion Paradigm Provides Evidence for a General Factor of Illusion Sensitivity and Personality Correlates**"
 shorttitle        : "Illusion Game Validation"
@@
     institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
-authornote: |
-  Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (dom.makowski@gmail.com).
+# authornote: |
+#   Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (dom.makowski@gmail.com).
 
@@
 Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
-These stimuli resulted from the combination of 15 equally-spaced levels of illusion *strength* (7 negative, i.e., congruent effects; 7 positive, i.e., incongruent effects; and 0) overlapped with 16 non-linearly spaced task *difficulty* levels (i.e., with an exponential, square or cubic spacing depending on the pilot results).
+These stimuli resulted from the combination of 15 equally-spaced levels of illusion *strength* (7 negative, i.e., congruent effects; 7 positive, i.e., incongruent effects) overlapped with 16 non-linearly spaced task *difficulty* levels (i.e., with an exponential, square or cubic spacing depending on the pilot results).
 For instance, a linear space of [0.1, 0.4, 0.7, 1.0] can be transformed to an exponential space of [0.1, 0.34, 0.64, 1.0], where 0.1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets).
@@
 Task difficulty (the objective difference between the targets of perceptual decision) levels are shown as colors, with lower values corresponding to harder trials.
-The results for each illusion are surrounded by 4 extreme examples of stimuli, corresponding to the hardest difficulty (on top) and the strongest illusion (on the right for incongruent illusions).
+The results for each illusion are surrounded by 4 extreme examples of stimuli, corresponding to lowest to the hardest difficulty (bottom to top) and the weakest to strongest illusion (from left to right for incongruent illusions).
 Bottom: We extracted the effect slope of the illusion strength and its interaction with task difficulty for each participant.
@@
 Finally, we found illusion sensitivity to be positively associated with "positive" personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
-Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs [often associated with psychoticism, @fox2006] to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli  [@ohmann2016; @konrath2009seeing; @fox2006].
+Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs (often associated with psychoticism [@fox2006]) to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli  [@ohmann2016; @konrath2009seeing; @fox2006].
 As such, the association between maladaptive traits and lower illusion sensitivity could be linked to a self-centered, decontextualized and disorganized information processing style.

### 2023-04-10 submission/R1/manuscript.tex
final  comments

@@
 \hypersetup{
-  pdftitle={The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates},
+  pdftitle={A Novel Visual Illusion Paradigm Provides Evidence for a General Factor of Illusion Sensitivity and Personality Correlates},
   pdfauthor={Dominique Makowski1, An Shu Te1, Stephanie Kirk1, Ngoi Zi Liang1, \& S.H. Annabel Chen1, 2, 3, 4},
@@
 
-\title{\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates}}
+\title{\textbf{A Novel Visual Illusion Paradigm Provides Evidence for a General Factor of Illusion Sensitivity and Personality Correlates}}
 \author{Dominique Makowski\textsuperscript{1}, An Shu Te\textsuperscript{1}, Stephanie Kirk\textsuperscript{1}, Ngoi Zi Liang\textsuperscript{1}, \& S.H. Annabel Chen\textsuperscript{1, 2, 3, 4}}
@@
 \authornote{
-
-Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (\href{mailto:dom.makowski@gmail.com}{\nolinkurl{dom.makowski@gmail.com}}).
 
@@
 Then, for each of the 10 illusion types, we generated a total of 134 stimuli.
-These stimuli resulted from the combination of 15 equally-spaced levels of illusion \emph{strength} (7 negative, i.e., congruent effects; 7 positive, i.e., incongruent effects; and 0) overlapped with 16 non-linearly spaced task \emph{difficulty} levels (i.e., with an exponential, square or cubic spacing depending on the pilot results).
+These stimuli resulted from the combination of 15 equally-spaced levels of illusion \emph{strength} (7 negative, i.e., congruent effects; 7 positive, i.e., incongruent effects) overlapped with 16 non-linearly spaced task \emph{difficulty} levels (i.e., with an exponential, square or cubic spacing depending on the pilot results).
 For instance, a linear space of {[}0.1, 0.4, 0.7, 1.0{]} can be transformed to an exponential space of {[}0.1, 0.34, 0.64, 1.0{]}, where 0.1 corresponds to the highest difficulty - i.e., the smallest objective difference between targets).
@@
 Finally, we found illusion sensitivity to be positively associated with ``positive'' personality traits, such as agreeableness and honesty-humility, and negatively associated with maladaptive traits such as antagonism, psychoticism, disinhibition, and negative affect.
-Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs\textsuperscript{often associated with psychoticism, 51} to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli\textsuperscript{19,51,52}.
+Although the existing evidence investigating links between illusion sensitivity and personality traits is scarce, these results are consistent with past findings relating pathological egocentric beliefs (often associated with psychoticism\textsuperscript{51}) to reduced context integration, manifesting in a tendency to separate objects from their surroundings when processing visual stimuli\textsuperscript{19,51,52}.
 As such, the association between maladaptive traits and lower illusion sensitivity could be linked to a self-centered, decontextualized and disorganized information processing style.

### 2023-04-22 submission/R1/manuscript.Rmd
preprint

@@
   - name          : "Dominique Makowski"
-    affiliation   : "1"
+    affiliation   : "1, 2"
     corresponding : yes    # Define only one corresponding author
@@
   - name          : "An Shu Te"
-    affiliation   : "1"
+    affiliation   : "2"
     orcid         : 0000-0001-8008-2824
@@
   - name          : "Stephanie Kirk"
-    affiliation   : "1"
+    affiliation   : "2"
     orcid         : 0000-0002-9312-5552
@@
   - name          : "Ngoi Zi Liang"
-    affiliation   : "1"
+    affiliation   : "2"
     role:
@@
   - name          : "S.H. Annabel Chen"
-    affiliation   : "1, 2, 3, 4"
+    affiliation   : "2, 3, 4, 5"
     orcid         : 0000-0002-1540-5516
@@
 affiliation:
-
   - id            : "1"
+    institution   : "School of Psychology, University of Sussex, UK"
+  - id            : "2"
     institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
-  - id            : "2"
+  - id            : "3"
     institution   : "LKC Medicine, Nanyang Technological University, Singapore"
-  - id            : "3"
+  - id            : "4"
     institution   : "National Institute of Education, Singapore"
-  - id            : "4"
+  - id            : "5"
     institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
